### 2019-01-19 writeup/paper-names.tex
paper draft

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

%\usepackage{pslatex}
%\usepackage{latexsym}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
%\usepackage[colorinlistoftodos]{todonotes}
\usepackage{rotating}
\usepackage{multirow}

\newcommand{\R}[0]{\mathbb{R}}
\newcommand{\Ff}[0]{\mathcal{F}}
\newcommand{\key}[1]{\textbf{#1}}

\newcommand{\soft}[1]{}
\newcommand{\nopreview}[1]{}
\newcommand\comment[1]{{\color{red}#1}}
\newcommand\mhahn[1]{{\color{red}(#1)}}
\newcommand{\rljf}[1]{{\color{blue}[rljf: #1]}}

\newcommand{\thetad}[0]{{\theta_d}}
\newcommand{\thetal}[0]{{\theta_{LM}}}
\newcommand{\thetap}[0]{{\theta_{P}}}

\title{Universals of word order result from optimization of grammars for efficient communication}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1,2]{Michael Hahn}
\author[a]{Daniel Jurafsky}
\author[b]{Richard Futrell}

\affil[a]{Stanford University}
\affil[b]{University of California, Irvine}

% Please give the surname of the lead author for the running footer
\leadauthor{Hahn}

% Please add here a significance statement to explain the relevance of your work
% limit 120 words 
\significancestatement{
What explains the universal properties of human languages? We present evidence that a major subset of these properties can be explained by viewing languages as codes for efficient communication among agents with highly generic cognitive constraints. In doing so, we provide the first full formalization and computational implementation of ideas which have been stated informally in the functional linguistics literature for decades. The success of this approach suggests a new way to conceptualize human language in quantitative and computational work, as an information-theoretic code dynamically shaped by communicative and cognitive pressures. Our results argue against the idea that the distinctive properties of human language result from essentially arbitrary genetic constraints.% (``universal grammar'').
}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{MH and RF designed research. MH implemented models and experiments. MH and RF wrote the paper. MH, DJ, and RF provided comments on the paper.}
\authordeclaration{The authors declare no conflict of interest.}
%\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: mhahn2@stanford.edu}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{language universals $|$ language processing $|$ computational linguistics}

\begin{abstract}
The universal properties of human languages have been the subject of intense study across disciplines. We report novel computational and corpus-based evidence for the hypothesis that a prominent subset of these universal properties---those related to word order---result from a process of optimization for efficient communication among humans. We develop a probabilistic, differentiable model of word order grammars: the means by which different languages convert underlying hierarchical structures into strings of words. We show how the parameters of a word order grammar can be optimized for robustness of information transfer, as quantified using mutual information with latent tree structures, and efficiency of processing, as quantified using incremental predictability. Applying these grammars to tree structures found in dependency corpora, we show that optimizing the grammar parameters for efficiency and robustness results in word order patterns that reproduce a large subset of the major word order correlations reported in the linguistic typological literature, and reproduce the predictions of previous heuristic theories such as dependency length minimization.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}
% TODO mention the preregistration somewhere
\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{F}or decades, researchers in fields ranging from philology to cognitive science to statistical physics have been involved in documenting and trying to explain the universal syntactic and statistical properties of human language \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical}. An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}. In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which certain syntactic universals can be explained through the statistical optimization of grammars for information-theoretic efficiency. % TODO add a cogsci person to citation chain at the beginning

We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}. These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages. For example, whether a language puts verbs before or after objects strongly determines whether it places adjectives before or after nouns, and whether it had prepositions or postpositions \cite{dryer1992greenbergian}.

A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.

Researchers working in the functional paradigm have proposed a range of criteria which languages should meet in order to enable efficient communication and processing. These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.

The contribution of this work is (i) to provide a principled formalization of the notion of communicative efficiency in word order, based on related work in information theory \cite{cover2006elements,tishby1999information}, bounded rationality \cite{genewein2015bounded}, and computational neuroscience \cite{friston2009predictive}; (ii) to introduce a computational framework in which we can produce simulated grammars directly optimized for our notion of efficiency, using state-of-the-art machine learning techniques \cite{hochreiter1997long,goldberg2017neural,dozat2017stanford}; and (iii) to demonstrate that in doing so, we reproduce the major word order patterns reported in \cite{greenberg1963universals} and related work \cite{dryer1992greenbergian}, as well as reproducing the predictions of previous heuristic theories of word order optimization \cite{hawkins1994performance,ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.

We formalize the notion of communicative efficiency using two terms: one having to do with minimizing the difficulty of linguistic communication, and one having to do with maximizing the information transferred in such communication. We quantify the difficulty of linguistic communication in terms of the word-by-word \key{predictability} of strings, which has been found to be a highly accurate and general predictor of human online comprehension difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and which can be justified in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}. We quantify the information transferred in language as \key{parseability}: the quantity of information contained in linguistic utterances about latent hierarchical tree structures. We also compare the results of optimizing for communicative efficiency to the results of optimizing for dependency length minimization, a previously proposed approximate metric of the amount of memory involved in language production and comprehension \cite{hudson1995measuring,gibson1998linguistic,gibson2000dependency,liu2017dependency,temperley2018minimizing}.

To evaluate whether communicative efficiency can explain word order universals, we apply the following method \cite{gildea2007optimizing}. First, we develop word order grammars: simple, interpretable models of word order, implemented as functions that take in an unordered syntactic dependency tree and output a distribution over possible word orders for that tree. Second, we take dependency treebanks of natural languages and compute word order grammars which are optimized for an objective functions representing communicative efficiency.
Third, we evaluate whether the optimized word order grammars reproduce the desired word order universals when applied to attested dependency trees from the treebanks across many languages. Our favorable results suggest that it may be possible to generally model human languages as efficient codes for communication among agents with generic human-like cognitive and communicative constraints. %The best way to describe a human language might be as a solution to an equation involving such constraints.

%We draw on data from the Universal Dependencies project \cite{nivre-universal-2017}, which has been developing treebanks in a unified format for several dozens of diverse languages.
%This universal annotation scheme is particularly useful for studying word order universals, as many of them can be operationalized  in terms of crosslinguistically valid syntactic notions. % such as \emph{verbs} and \emph{objects}.

\section{Background}

\subsection{Explananda: Word Order Correlations}

Working with a database of 30 languages, Greenberg \cite{greenberg1963universals} proposed 45 universals of language that he determined to be true of most or all languages.
31 of them were concerned with word order, the remainder with morphology.
Many of the word order universals state \emph{correlations} between patterns:
For instance, Universals 3 and 4 state that ``\emph{Languages with dominant VSO order are always prepositional.}''  (e.g. Arabic), and ``\emph{With overwhelmingly greater than chance frequency, languages with normal SOV order are postpositional.}'' (e.g. Japanese). 
These are illustrated in Figure~\ref{fig:arabic-japanese}:
Arabic has basic order verb-object (VO) and prepositions (\emph{li} `to'), while Japanese has order object-verb (OV) and postpositions (\emph{ni} `to'). Our goal is to explain Greenberg's reported word order correlations, as updated in \cite{dryer1992greenbergian}.

\begin{figure}
\centering
\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
	   katabt \& risāla \& li \& sadīq  \\
	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
	   wrote \& letter   \& to \& friend  \\
   \end{deptext}
	%   \deproot{3}{ROOT}
   \depedge{1}{2}{obj}
	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
   \depedge{1}{3}{}
   \depedge{3}{4}{}
   %\depedge[arc angle=50]{7}{6}{ATT}
\end{dependency}
\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
      tomodachi \& ni \& tegami-o \& kaita \\
 \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
	   friend \& to \& letter  \& wrote \\
   \end{deptext}
	%   \deproot{3}{ROOT}
   \depedge{4}{2}{}
	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
   \depedge{2}{1}{}
   \depedge{4}{3}{obj}
   %\depedge[arc angle=50]{7}{6}{ATT}
\end{dependency}

\begin{dependency}[theme=simple]
    \begin{deptext}
    I \& wrote \& a \& long \& letter \& to \& a \& friend \\
    \end{deptext}
    \depedge{2}{1}{nsubj}
    \depedge{2}{5}{obj}
    \depedge{5}{4}{ }
    \depedge{5}{3}{ }
    \depedge{2}{6}{}
    \depedge{6}{8}{ }
    \depedge{8}{7}{ }
\end{dependency}
	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese}
\end{figure}

% TODO what to do about these labels...? They're not UD...

\subsection{Dependency trees}

The sentences in Figure~\ref{fig:arabic-japanese} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words. For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} if the verb, rather than e.g. the subject or an indirect object. A dependency arc is drawn from a \key{head} (e.g. \emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g. \emph{tegami-o}). Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}. Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.

In order to interpret a sentence correctly, a listener must be able to reconstruct the information expressed in the dependency tree and its labels. This is because, according to the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.

\section{Optimizing word order grammars}

Here we describe our model of word order and how we derive optimal grammars within that model in terms of communicative efficiency.

Let $t$ be an unordered dependency tree, i.e. a structure as in Figure~\ref{fig:arabic-japanese} but without any specified ordering of the words. We think of an unordered dependency tree $t$ as representing the compositional meaning to be conveyed by a sentence. We define a \key{word order grammar} as a function $L_\theta(t)$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output a probability distribution over ordered sequences of words $\mathbf{w}$ linearizing the tree, or equivalently, a probability distribution over orders on the nodes of the tree. % do we need to define "linearize"?

%For example, TODO % We should have some pretty picture visualizing the process here

Next we define how a word order grammar linearizes trees in terms of its parameters $\theta$, and then we describe how $\theta$ can be set to optimize objective functions related to communicative efficiency.

\subsection{Specification of word order grammars}

All natural languages have some degree of word order regularity. Thus, it is not sufficient to optimize the word orders of individual sentences in the corpora---instead, we will optimize the word order rules of entire languages.

Our goal is to specify word order grammars that transduce unordered dependency trees into strings of words. We have three additional goals: first, we would like our grammars to be able to operate over dependency trees as specified in large human-annotated treebank corpora, such as those developed by the Universal Dependencies project \cite{nivre2017universal}. Second, we would like our grammars to be simple and easily interpretable, so we can determine whether they exhibit word order universals. Third, we would like our grammars to be differentiable, so that it is easy to optimize their parameters for efficiency.

Our model linearizes dependency trees such as the ones in Figure~\ref{fig:arabic-japanese}. We assume nodes are labeled with part-of-speech (POS) tags, and arcs are labeled with syntactic relations. The parameters of a word order grammar are as follows. 
For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p. 184) without replacement.

This simple parameterization lets us directly test whether a given grammar follows word order universals. For example, we can test whether a Greenbergian word order correlation holds in a grammar by checking whether $a_\tau>\frac{1}{2}$ for different syntactic relations $\tau$.

\subsection{Objective functions}
\label{ref:objective-funcs}

Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a word order grammar. Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
\begin{equation}
\label{eq:exp-over-utts}
J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim L_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
\end{equation}
where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$. In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.

\paragraph{Predictability}
Predictability represents the incremental difficulty of online language processing. We formalize predictability as the \key{surprisal} (negative log probability) of each word given its preceding context \cite{hale2001probabilistic,levy2008expectation,smith2013effect}. Optimization of predictability has been proposed as an explanation for certain word order patterns by \cite{ferrericancho2017optimality}.

We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
\begin{equation*}
R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
\end{equation*}
where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\L_\theta$ as accurately as possible (maximizing predictability):
\begin{equation*}
	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
\end{equation*}
We implement the language model $P_\phi$ using an LSTM neural network architecture \cite{hochreiter1997long}, currently the state of the art for this task \cite{goldberg2017neural}. Note that the language model parameters $\phi$ and the word order grammar parameters $\theta$ are recursively dependent on each other. See the SI for details on how $\theta$ and $\phi$ are jointly optimized, as well as on the neural network architecture.

\paragraph{Parseability}
Parseability is the extent to which the syntactic and semantic relationships among words can be recovered from the string of words. 
%Some notion of parseability has been proposed informally as a desirable property of languages in the linguistic literature going back at least to the 19th century \cite{humboldt,sapir,jesperson}, and parseability has been offered as an underlying justification for syntactic locality as a cost function for language processing \cite{hawkins-performance-1994}. 
%At a basic level, a language can encode information about syntactic relations into fixed word order patterns or into morphological marking on individual words. Much linguistic work has been devoted to investigating the trade-offs of various ways of indicating syntactic relations \cite{}. In this work, we present what is to our knowledge the first fully implemented computational formalization of this idea.
We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees. Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
\begin{align}
\nonumber
R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
\nonumber
&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
\end{align}
where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
\begin{equation*}
\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
\end{equation*}
We implement $P_\psi$ using a highly generic graph-based parsing architecture \cite{kiperwasser2016simple,dozat2017stanford} which recovers graph structures from word strings by solving a minimal spanning tree problem.
Due to data sparsity, we train parsers to predict dependency trees from POS tags alone, not from full wordforms, therefore we do not capture the extent to which morphological wordforms are informative about dependency tree structure.

See the SI for further details on the joint optimization of $\theta$ and $\psi$ and on the neural parser architecture.

\paragraph{Efficiency}
The two scoring functions---predictability and parseability---can be combined into an overall equation of the form:
\begin{equation}
\label{eq:combined-score}
R_{Efficiency} = R_{Pars} + \lambda R_{Pred},
\end{equation}
with an interpolation weight $\lambda \in [0,1)$. In all experiments in this paper we use $\lambda=.9$. This combined equation represents efficiency: bits of information transferred about dependency trees minus the cost of the transmission. In expectation over utterances (substituting Eq.~\ref{eq:combined-score} into Eq.~\ref{eq:exp-over-utts}), the overall objective function to be maximized by a word order grammar can be written in information-theoretic terms as:
\begin{equation}
    \label{eq:efficiency}
    J_T(\theta) = I[L_\theta(T);T] - \lambda H[L_\theta(T)],
\end{equation}
where $I[L_\theta(T);T]$ is the \key{mutual information} between strings and trees, and $H[L_\theta(T)]$ is the \key{entropy} of strings \cite{cover2006elements}. Maximizing Eq.~\ref{eq:efficiency} is equivalent to maximizing mutual information between strings and trees subject to a constraint on the maximum entropy of strings.

The two terms of Eq.~\ref{eq:efficiency} implement two opposing pressures. Parseability (the mutual information term) pushes to make strings more heterogeneous, so that they can indicate different tree structures unambiguously. Predictability (the entropy term) pushes to make strings more homogeneous, to reduce the complexity of using the language. The idea that natural language arises from a tension between these two forces goes back over a century \cite{gabelentz1901sprachwissenschaft}. Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking unordered dependency trees $T$ as the underlying meanings to be conveyed. Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI for the precise relationship).

\paragraph{Dependency length}

In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}. 
Dependency length has been proposed as a metric of processing difficulty on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, and minimization of dependency length is known to recover a large subset of the Greenbergian universals \cite{hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic}. 
We demonstrate that the minimization of dependency length can be formalized in our framework, reproducing the finding that optimizing grammars for dependency length recovers several word order universals \cite{gildea2010grammars}.

To quantify dependency length, we sum the lengths of syntactic dependencies in the sentence \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
\begin{equation}
    \label{eq:depl}
	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
\end{equation}
where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.

\subsection{Discussion and Related Work}

Our work builds on previous work on optimizing word order grammars for various objectives \cite{gildea2007optimizing,gildea2010grammars,gildea2015human}.
Previous work has used deterministic word order grammars optimized by simple hill-climbing. 
Our work departs from previous work in three ways. 
First, we use word order grammars that are probabilistic and differentiable, enabling greater expressivity and optimization by stochastic gradient descent.
Second, we use state-of-the-art neural language models for predictability, jointly optimizing parameters of word order grammars and language models.
Third, we implement parseability and efficiency as a novel objective functions, which would not be feasible with non-differentiable word order grammars.

In the space of word order grammars, more complex models than ours have been defined \cite{belz2011first,futrell2015experiments,wang2016galactic}. However, the simple parametric form of our model enables easy interpretation across languages.
Like these previous models, our model makes simplifying assumptions and will not fit the rules of natural languages perfectly.
For instance, none of the models accounts for word order phenomena that are conditioned on the larger context---e.g., differences in word order between embedded and main clauses. 
Also like previous models, our model only generates word orders that are \key{projective}, meaning that dependency lines do not cross. There are two reasons for constraint. First, there is currently no easily interpretable probabilistic model of non-projective word ordering to build on from the natural language processing literature (cf. \cite{bohnet2012generating}). Second, natural language dependency trees are overwhelmingly projective \cite{kuhlmann2013mildly}. Note that the preponderance of projectivity in natural language is itself a linguistic universal, equivalent to context-freeness in formal language theory \cite{marcus1965notion,kuhlmann2013mildly}, and explanations have been offered for it in terms of dependency length minimization  \cite{ferrericancho2006syntactic}.

We test efficiency metrics based both on information theory (Eq.~\ref{eq:efficiency}) and dependency length minimization (Eq.~\ref{eq:depl}). We do not see these as competing theories. Rather, we see the information-theoretic objective as a broad description of efficiency in general, and dependency length as a heuristic description of a particular component of complexity. There is reason to believe that dependency length minimization might emerge as a byproduct of optimization for parseability and predictability under resource constraints \cite{futrell2017memory}: in particular, short dependencies are optimal for predictability given lossily compressed representations of context \cite{futrell2017noisycontext,futrell2017generalizing}. In the results below, we demonstrate that optimizing for our Eq.~\ref{eq:efficiency} does indeed produce dependency length minimization as a byproduct.

\section{Results}

\subsection{Relative efficiency of languages}
\label{sec:relative-efficiency}

We first demonstrate that real languages are relatively efficient compared to random baselines. For each language, we generated ten baseline word order grammars for the language by choosing all word order grammar parameters randomly at uniform from $[0,1]$. Figure~\ref{fig:pareto-plane} shows the predictability and parseability for each real language relative to its baseline grammars. In order to control for limitations due to our word order grammar formalism, we represent real languages in the figure by maximum likelihood fits of word order grammars to the real language data.
%; in general these maximum-likelihood word order grammars end up with lower predictability and parseability than the real languages. % TODO verify this is true? TODO awkward sentence
For the calculation of predictability and parseability, we make all (baseline and real) word order grammars deterministic by always choosing the highest-probability linearization of each tree; by making the grammars deterministic in this way we eliminate an anticonservative bias toward low predictability in the baseline languages, which are highly nondeterministic. 
The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.

\begin{figure}
    \centering
    \includegraphics[scale=.4]{figures/pareto-plane-iso-best-balanced-legend.pdf}
    \caption{Predictability and parseability of 51 UD languages (red), indicated by ISO codes, compared to ten baseline word order grammars per language (green). Predictability and parseability scores are $z$-scored within language. Each point for a real language has a line pointing in the direction of the center of mass of its baselines. The green contour shows the density of baseline languages. Unlabeled dots represent the centroid for real languages (red), baseline languages (green), and languages optimized for predictability (yellow), parseability (pink), efficiency (blue), and dependency length (red). When a language is to the bottom--left of its baselines, this indicates that it is relatively optimal for efficiency.} % TODO colors, fontsize, etc could conceivably be made prettier
    \label{fig:pareto-plane}
\end{figure}

Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages. Languages appear to be attracted toward these points and away from the region of the baseline languages. We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
% TODO add something about dependency length-optimized languages?

These results are shown in more detail for two typologically distinct languages (English and Japanese) in Figure~\ref{fig:predict-pars}. Here we include also the predictability and parseability of the original language data and of the word order grammars optimized for dependency length. We see that real languages are even more predictable than any optimized or baseline language, while optimization for either parseability or dependency length suffices to raise parseability to the level of the real languages. This result is in line with the finding that languages with longer dependencies are harder to parse in general \cite{gulordava2016multilingual}.

\begin{figure*}
\includegraphics[width=0.98\textwidth]{figures/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right). We provide cross-entropy for predictability (lower is better) and unlabeled attachment scores---a standard metric from natural language processing---for parseability (higher is better).}\label{fig:predict-pars}
\end{figure*}

Finally we demonstrate the relationship between dependency length minimization and the maximization of efficiency. Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency. We see that optimizing for efficiency lowers dependency length relative to random baselines, in keeping with the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}. In 3/4 of the cases shown, optimizing explicitly for dependency length produces dependencies that overshoot the depenndency length of the real language; in Chinese and Hindi, the real language is best matched by efficiency optimization.

\begin{figure*} % need Efficiency and (maybe) MLE langs on here
\includegraphics[width=0.98\textwidth]{figures/depLength-facet.pdf}
\caption{Average dependency length as a function of sentence length in four languages. Across languages, real and optimized languages have shorter dependencies than random baseline orderings.}\label{fig:deplength}
\end{figure*}

\subsection{Greenbergian Word Order Correlations}
\label{sec:greenberg}
\begin{table*}
	\begin{center}
\begin{tabular}{|ll|l|l|l|ll|l|}
	\hline
	%&
	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       &     \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
	verb & object &      & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
adposition    &    NP    &    lifted\_case    &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
copula    &    NP    &    lifted\_cop    &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
auxiliary    &    VP    &    aux    &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
noun    &    genitive    &    nmod    &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
noun    &    relative clause    &    acl    &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
complementizer    &    S    &    lifted\_mark    &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
verb    &    PP    &    obl    &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
want    &    VP    &    xcomp    &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
verb    &    subject    &    nsubj    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
verb    &    manner adverb    &    advmod    &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenbergian Correlations. Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner' (the `Correlate' column), whose relative order correlates with that of verbs and objects.
	For clarity, we organize the correlations by the category of the `verb patterner'.
	For each correlation, we give our operationalization in terms of UD. A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation. 
	For each correlation we report how many of the languages in our sample satisfied it (`Real').
We then report, for each correlation and each objective function, how many of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
\end{table*}

%\begin{table*}
%\begin{tabular}{lllllllllllllllllll}
%	&	          & \multicolumn{2}{c}{Dep. Length}      & \multicolumn{2}{c}{Lang. Mod.}              &  \multicolumn{2}{c}{Parsing}  &\multicolumn{2}{c}{Operationalization} & & GU\\ \hline
%	Adp &	Adp+Noun &                  & \cmark       &             &            &         &             & \textsc{noun} $\xrightarrow{case}$ \textsc{adp} & - &  3, 4\\ \hline
%	Verb &	Verb+PP & \checkmark        & \cmark      & \checkmark  &   \cmark         &    &             & \textsc{verb} $\xrightarrow{obl}$ \textsc{noun}  & + & 7  \\
%	Verb &	Copula & \checkmark         &  \cmark     &  \checkmark &   \cmark    &  \checkmark  &     \cmark        & \textsc{noun} $\xrightarrow{cop}$  \textsc{aux} & - \\
%	Verb &	want & \checkmark           & \cmark      & \checkmark  & \cmark     &  \checkmark  &    \cmark         & \textsc{verb} $\xrightarrow{xcomp}$ \textsc{verb} & + & 15 \\
%	Verb &	subject &                   &       & \checkmark  &      \cmark      & \xmark   &    \xmark         & \textsc{verb} $\xrightarrow{nsubj}$ \textsc{noun} & + \\
%	Verb &	auxiliary & \cmark          &       & \cmark      & \cmark     &           &             & \textsc{verb}  $\xrightarrow{aux}$ \textsc{verb}/\textsc{aux}  & - & 16\\
%	verb & manner adverb &              &  -     &  \xmark     &   \xmark     &          &             & \textsc{verb} $\xrightarrow{advmod}$ ADV  & + & 7 \\ \hline
%	Comp &	comp; adv sub. &            &   \cmark    &             &   \cmark         &          &             & \textsc{verb} $\xrightarrow{mark}$ \textsc{sconj} & - & \\ \hline
%	Noun &	article N' &                &   \xmark/\xmark    &             &    -/\cmark        &          &   -/\cmark          & \textsc{noun} $\xrightarrow{det}$ \textsc{det}/\textsc{pron}  & -\\
%	Noun &	noun+genitive & \cmark      &   \cmark    &             &            &              &    \cmark  & \textsc{noun} $\xrightarrow{nmod}$ \textsc{noun}  & + & 2\\
%	Noun &	noun+relative clse & \cmark &  \cmark   &               &            &  \checkmark &      \cmark       & \textsc{noun} $\xrightarrow{acl}$ \textsc{verb} & + & 24 \\ \hline
%	Adj &	adj+std            &        &  \cmark   &               &            &             &             &  ADJ $\xrightarrow{nmod}$ \textsc{adp}  & + & 23\\
%\end{tabular}
%	\caption{Greenbergian Correlations, as provided by \cite{dryer-greenbergian-1992}. For each correlation, we give our operationalization in terms of UD and whether the direction of this dependency correlates positively ($+$) or negatively ($-$) with the verb-object-dependency. For each correlation and each objective function, we indicate whether the correlation is predicted correctly (\cmark), in the wrong direction (\xmark), or no prediction is made (blank). For each objective function, we show results for plain UD (left column) and UD with function word heads (right column). }\label{table:results-dryer}
%\end{table*}
%%, with corresponding numbers in the original Greenberg universals, where applicable.

%In the last sections, we confirmed the effectiveness of our optimization method and elucidated the relation between the three optimization functions on the level of .

We now examine to what extent we can recover Greenberg's word order correlations in optimized grammars. 
Dryer \cite{dryer1992greenbergian} presents a comprehensive updated version of the word order correlations, drawing on 625 languages, which we take as the basis of our evaluation.
In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
This dependency corresponds to dependencies of the form $\xrightarrow{obj}$ in UD.
Similarly, most of the other dependencies in Dryer's formulation can be formalized:
for instance, the dependency between nouns and relative clauses can be formalized as $\xrightarrow{acl}$.
Testing whether an ordering model satisfies the correlation between relative clauses and objects reduces to checking whether these two relation types have direction parameters $a_\tau$ that are either both $>0.5$ (both kinds of dependents precede their heads) or both $< 0.5$ (they both follow their heads).

Dryer \cite{dryer1992greenbergian} presents three correlations which do not correspond to dependencies annotated in UD: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles. Two pairs of Dryer's correlations, namely those for the dependencies between complementizers and adverbial subordinators and their complement clauses,and those for the dependencies between verbs and adpositional phrases, and adjectives and their standard of comparison, had to be collapsed into two correlations in UD. From Dryer's 15 correlations, we obtained 10 formalized correlations, roughly covering nine of Greenberg's original universals. %TODO update to 10? What happened to the 11th? Was that amod? [The 11th was adjective+standard of comparison, which is merged into obl] Maybe we could include amod as a "true rejection" of efficiency, but a false alarm for DepL

In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau$ show the same direction for the correlating dependency and for the verb-object dependency, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
%We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau > 0.5$ holds for the correlating dependency depending on whether $a_\tau > 0.5$ for the verb-object dependency, with per-language-family random intercept and slope.\footnote{We coded language families according to \url{universaldependencies.org}.}
We are interested in the direction and significance of this effect: 
If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.

We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages. To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks. The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) in all except four languages, for WALS did not have data.

\paragraph{Results}
Results are shown in Table~\ref{table:results-dryer}.
All correlations but two are confirmed in the models estimated from the real orderings. The exceptions are the subject--verb dependency and the verb--adverb dependency, which typically go in the opposite direction from the standard description.
We will discuss these exceptions further below.

In keeping with previous work, we see that optimizing for dependency length correctly accounts for nine word order correlations, missing only the verb--adverb dependency. 
Predictability and parseability predict five and seven correlations, respectively, making largely complementary predictions.
Efficiency significantly predicts all the word order correlations, each in the same direction as attested in the dependency corpora.

We now address the two word order correlations whose direction in the dependency corpora is opposite from what would be expected in the typological literature. The first is the correlation of the order of verb--subject and verb--object dependencies. Our sample of mainly European languages highly over-represents languages with the general order subject--verb--object (such as English), in which the order of the verb--subject and verb--object dependencies are anti-correlated. Surprisingly, given the sample of tree structures of these languages, it turns out that the optimal languages tend to have anti-correlated orders for subjects and objects order similar to the real languages.

The second anomalous dependency is the verb--manner adverb dependency. We believe the anti-correlation in the UD corpora arises because the \emph{advmod} dependency does not distinguish between manner adverbs---the subject of the typological judgment---and various other types of modifiers such as sentence-level adverbs. Nevertheless, the languages optimized for efficiency reproduce the anti-correlation of the orders of verb--object and verb--adverb at around the same rate as the real languages.

We further evaluate the word order predictions of efficiency, showing that efficiency is most successful in predicting correlations in the direction found in the UD corpora.
We constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency in languages optimized for efficiency, with random effects for language and language family, correlated across the ten dependencies.
We conducted the same analysis for predictability, parseability, and dependency length.
We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$. The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.
For dependency length, the posterior puts much of the probability mass on predicting only nine of the correlations; predictability and parseability predict significantly less correlations.

\begin{figure}
    \centering
    \includegraphics[scale=.2]{figures/posterior-satisfied-universals-efficiency.pdf}
     \includegraphics[scale=.2]{figures/posterior-satisfied-universals-langmod.pdf}
     \includegraphics[scale=.2]{figures/posterior-satisfied-universals-parseability.pdf}
    \includegraphics[scale=.2]{figures/posterior-satisfied-universals-depl.pdf}
	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks.}
    \label{fig:posterior-satisfied-count}
\end{figure}

\subsection{Discussion}

In Section~\ref{sec:relative-efficiency}, we found that word order grammars in the majority of UD languages have better efficiency than baseline word order grammars. Furthermore, in Section~\ref{sec:greenberg} we found that explicitly optimizing grammars for efficiency reproduces 10 major word order correlations reported in the typological literature, and with greater accuracy than the typological literature when it comes to the languages studied.

While the efficiency objective succeeds in predicting the presence of the most word order correlations, the dependency length objective seems to predict a greater prevalence for those correlations that it does predict. We believe this discrepancy is a result of noise in the optimization process. Optimizing for efficiency is computationally challenging, requiring stochastic gradient descent on word order grammar, language model, and parser parameters, while optimizing for dependency length is simpler and thus more optimal grammars can be reached easily. Inasmuch as dependency length minimization is computationally easier to implement than efficiency maximization, we believe it will remain useful as an easy-to-use heuristic for predicting and explaining word order universals.
% TODO we can also report (in the SI) prevalences among the best-performing of the optimized languages, i.e., those where optimization was most successful. I would expect that this should narrow the gap between DepL and Efficiency, which would confirm that the gap is really about the difficulty of optimization, not about the objective functions.

\section{Conclusion}

We found that a large subset of the Greenbergian word order correlations can be explained in terms of from optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods. We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals. Beyond our present results, we provide a complete formalization and computational framework in which theories of the functional optimization of languages can be tested. Other objective functions could be hypothesized and tested within our framework; furthermore, future advances in machine learning will enable the optimization of richer and richer models of grammar.

A major question for functional explanations for linguistic universals is: \emph{how} do languages come to be optimized? Do speakers actively seek out new communicative conventions that allow better efficiency? Or do languages change in response to biases that come into play during language acquisition \cite{fedzechkina2012language,culbertson2012learning}? Our work is neutral toward such questions. To the extent that language universals arise from biases in learning or in the representational capacity of the human brain, our results suggest that those biases tilt toward communicative efficiency. Our work does provide against the idea that word order universals are best explained in terms of learning biases that are irreducibly arbitrary and genetic in nature \cite{chomsky2010some}.

While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into dependency trees. The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication. Relatedly, there are many remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}. Our models cannot capture this trade-off for technical reasons: the parseability objective does not operate over wordforms, only POS tags, and so it does not take advantage of morphological cues to syntactic structure. Future work can investigate whether these and other remaining universals can be explained using more sophisticated models of how meaning representations are transduced into strings of words, based on larger databases with richer annotations.

\matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
We use all languages for which at least one treebank with a training partition was available, a total of 50 languages.
For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.
Punctuation was removed.
%We removed punctuation. %, which we operationalized by removing all dependents of \emph{punct} dependencies, and recursively removing their dependents.

Universal dependencies represents as dependents some words that are typically classified as heads in syntactic theory.
This particularly applies to the \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark} dependencies.
Following prior work studying dependency length minimization \cite{futrell2015largescale}, we modified each treebank by inverting these dependencies, promoting the dependent to the head position.
We report results on this modified version of UD.

The efficiency optimization results from Table~\ref{tab:results-dryer} were preregistered: \url{http://aspredicted.org/blind.php?x=8gp2bt}.

See the SI for details on the neural language models and parsers used, and on the optimization procedures.
}

\showmatmethods{} % Display the Materials and Methods section

\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, and audiences at CAMP 2018 for helpful discussion.}

\showacknow{} % Display the acknowledgments section

%\bibliographystyle{acl_natbib}
%\bibliography{extracted}
\bibliography{everything}

\end{document}

### 2019-01-19 writeup/pnas-word-order-si.tex
paper draft

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{hyperref}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{arydshln}

\newcommand{\key}[1]{\textbf{#1}}

\usepackage{mathtools}
\DeclarePairedDelimiter{\ceiling}{\lceil}{\rceil}

\title{Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''}
\author{
        Michael Hahn \\
                Department of Linguistics\\
       Stanford University
            \and
       Daniel Jurafsky\\
       Department of Linguistics\\
       Stanford University \\
       \and
       Richard Futrell\\
       Department of Language Science\\
       University of California, Irvine
}
\date{\today}

\begin{document}
\maketitle

\section{Results on all UD Relations}
Table~\ref{tab:all-predictions} shows the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.

\begin{table*} % TODO figure out away to make this table more aesthetic
	\begin{center}

\tiny{
\begin{tabular}{|l|l|l|ll|l|l|}
	\hline
	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
% From Dryer	
acl  &  80   &   \textbf{85}$^{***}$   &   43   &   \textbf{82}$^{***}$   &   \textbf{71}$^{***}$  & $>50\%$  \citep{dryer1992greenbergian} \\
aux  &  12   &   \textbf{26}$^{***}$   &   \textbf{19}$^{***}$   &   49   &   \textbf{35}$^{**}$  & $<50\%$ \citep{dryer1992greenbergian} \\
lifted\_case  &  86   &   \textbf{81}$^{***}$   &   \textbf{40}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{67}$^{***}$ & $> 50 \%$ \citep{dryer1992greenbergian}  \\
lifted\_cop  &  94   &   \textbf{80}$^{***}$   &   59   &   \textbf{72}$^{***}$   &   \textbf{63}$^{**}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
lifted\_mark  &  76   &   \textbf{85}$^{***}$   &   \textbf{57}$^{*}$   &   \textbf{76}$^{***}$   &   \textbf{72}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
nmod  &  80   &   \textbf{82}$^{***}$   &   49   &   \textbf{71}$^{***}$   &   \textbf{68}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\ 
obl  &  88   &   \textbf{78}$^{***}$   &   \textbf{71}$^{***}$   &   46   &   \textbf{68}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
xcomp  &  88   &   \textbf{90}$^{***}$   &   \textbf{76}$^{***}$   &   \textbf{89}$^{***}$   &   \textbf{85}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
\hdashline
nsubj  &  33   &   \textbf{29}$^{***}$   &   53   &   \textbf{7}$^{***}$   &   \textbf{23}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
advmod  &  35   &   51   &   \textbf{18}$^{***}$   &   47   &   \textbf{39}$^{**}$  & $>50\%$ for manner adverbs; $\approx 50\%$ for negation, TAM, intensifiers \citep{dryer1992greenbergian} \\

\hline
% Other universals, not in Dryer
advcl  &  86   &   \textbf{84}$^{***}$   &   51   &   \textbf{67}$^{***}$   &   \textbf{67}$^{*}$  & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
amod  &  51   &   \textbf{84}$^{***}$   &   42   &   \textbf{53}$^{*}$   &   51   & $\approx 50\%$ \citep{dryer1992greenbergian} \\
nummod  &  37   &   \textbf{74}$^{**}$   &   49   &   52   &   52 & $\approx 50\%$ \citep[][89A vs. 83A]{wals} \\

\hline
% Typological data not available
appos  &  77   &   \textbf{77}$^{***}$   &   49   &   \textbf{67}$^{***}$   &   \textbf{62}$^{**}$ &  Unknown \\%($>50\%$? correlates with \textit{nmod} according to Greenberg U 23)  \\ % \cite{cinque2009greenberg} claims it correlates with OV/VO. However, it's not clear to me how closely appos matches the subject of the typological claim
lifted\_cc  &  71   &   \textbf{83}$^{***}$   &   51   &   \textbf{81}$^{***}$   &   \textbf{71}$^{***}$  & Unknown \\
ccomp  &  86   &   \textbf{88}$^{***}$   &   \textbf{70}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{77}$^{**}$ & Unknown (cf. \cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
csubj  &  79   &   \textbf{78}$^{***}$   &   60   &   \textbf{66}$^{***}$   &   \textbf{68}$^{**}$  & Unknown (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
expl  &  12   &   36   &   62   &   40   &   53  & Unknown \\
iobj  &  74   &   49   &   \textbf{78}$^{***}$   &   \textbf{38}$^{*}$   &   62  & Unknown \\
vocative  &  46   &   \textbf{42}$^{*}$   &   59   &   45   &   \textbf{45}$^{*}$  & Unknown \\

\hline
% Uninterpretable
compound  &  66   &   \textbf{61}$^{*}$   &   48   &   \textbf{58}$^{*}$   &   52  & Uninterpretable \\
det  &  31   &   \textbf{71}$^{***}$   &   \textbf{56}$^{*}$   &   48   &   54  & Uninterpretable (conflates articles and others) \\%$.50\%$ for articles, $\approx 50\%$ for others, \citep{dryer1992greenbergian} \\ %TODO what else is conflated in det? why do we keep advmod but not det?
dislocated  &  47   &   61   &   \textbf{68}$^{**}$   &   55   &   58  & Uninterpretable \\
dep  &  56   &   \textbf{64}$^{*}$   &   52   &   58   &   \textbf{58}$^{*}$   & Uninterpretable \\

\hline
% UD artifacts
conj  &  72   &   \textbf{82}$^{***}$   &   \textbf{42}$^{**}$   &   \textbf{62}$^{***}$   &   \textbf{59}$^{*}$  & UD artifact  \\
discourse  &  24   &   \textbf{36}$^{**}$   &   \textbf{41}$^{*}$   &   49   &   46  & UD artifact   \\
fixed  &  78   &   \textbf{32}$^{***}$   &   45   &   47   &   44  & UD artifact   \\
flat  &  73   &   \textbf{66}$^{**}$   &   45   &   \textbf{59}$^{**}$   &   54  & UD artifact   \\
goeswith  &  61   &   57   &   44   &   57   &   52  & UD artifact   \\
list  &  94   &   59   &   50   &   54   &   52  & UD artifact  \\
orphan  &  82   &   \textbf{70}$^{**}$   &   50   &   \textbf{55}$^{*}$   &   59  & UD artifact   \\
parataxis  &  66   &   56   &   43   &   57   &   59  & UD artifact  \\
reparandum  &  18   &   53   &   57   &   43   &   48  & UD artifact  \\

 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
}

\end{center}
\caption{Predictions on all UD relations occurring in more than one language. Numbers are the fraction of languages where the direction of the given relation agrees in direction with the \emph{obj} relation. In the last column, we indicate what direction would be expected typologically. ``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful. ``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
\label{tab:all-predictions}
\end{table*}

\section{Results on Best Optimized Grammars}

In the main part, we reported the percentage of optimized grammars supporting correlations.
Here we report these numbers for the grammars that optimize each objective most successfully.
That is, for each language and objective, we chose among the eight optimized grammars the one that achieves the best value for the objective function on the validation partition of the treebanks -- that is, the one where the optimization procedure was most successful.

Numbers for these languages are shown in Table~\ref{table:results-dryer-best-only}.
%Confirming the results from the main paper, 
As expected, the correlations mostly have stronger prevalences than when considering all optimized languages.
For Dependency Length, the correlations that it predicts mostly have prevalences close to 100 \%, whereas both the UD languages and efficiency-optimized languages show somewhat lower percentages.
That is, dependency length mostly overshoots for the correlations that it predicts correctly.

\begin{table*}
	\begin{center}
\begin{tabular}{|ll|l|l|l|ll|l|}
	\hline
	%&
	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       &     \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
	verb & object &      & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
%adp. &
adposition    &    NP    &    lifted\_case    &    86    &    98    &    69    &    88    &    83   \\
copula    &    NP    &    lifted\_cop    &    94    &    98    &    58    &    85    &    74   \\
auxiliary    &    VP    &    aux    &    88    &    84    &    90    &    61    &    73   \\
noun    &    genitive    &    nmod    &    80    &    98    &    75    &    88    &    91   \\
noun    &    relative clause    &    acl    &    80    &    100    &    69    &    88    &    90   \\
complementizer    &    S    &    lifted\_mark    &    76    &    100    &    76    &    82    &    87   \\
verb    &    PP    &    obl    &    88    &    98    &    88    &    75    &    91   \\
want    &    VP    &    xcomp    &    88    &    100    &    92    &    94    &    96   \\
verb    &    subject    &    nsubj    &    33    &    25    &    53    &    10    &    21   \\
verb    &    manner adverb    &    advmod    &    35    &    61    &    27    &    55    &    36   \\
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
\end{tabular}
	\end{center}
	\caption{Percentages of grammars supporting correlations, among grammars that best optimize each objective on treebanks of a given language.}\label{table:results-dryer-best-only}
\end{table*}

\begin{figure}
    \centering

    \includegraphics[scale=.25]{figures/coverage-langmod-best.pdf}
    \includegraphics[scale=.25]{figures/coverage-parse-best.pdf}
    \includegraphics[scale=.25]{figures/coverage-two-lambda09-best.pdf}

    \includegraphics[scale=.25]{figures/coverage-ground.pdf}
    \includegraphics[scale=.25]{figures/coverage-depl.pdf}
	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
    \label{fig:posterior-prevalence}
\end{figure}

\section{Further Statistical Analyses}

In the main results table, we reported the fraction of optimized languages satisfying a correlation.
We show that essentially the same predicted fractions are obtained when controlling for the imbalance in distribution of language families.

To this end, we used the logistic mixed-effects regressions that were used to estimate significance to compute the expected prevalence of a correlation controlling for language families.
To obtain predictions controlling for languages and language families, we plot the logit-transformed posterior of $\alpha$ in Figure \ref{fig:posterior-prevalence}.

\begin{figure}
    \centering
    \includegraphics[scale=.7]{figures/posterior_joint_with_ud.pdf}
	\caption{Posteriors of the prevalences of the correlations among languages optimized for Predictability (red), parseability (green), and efficiency (blue), controlling for languages and language families, with prevalence in the UD languages.}
    \label{fig:posterior-prevalence}
\end{figure}

%\begin{figure}
%    \centering
%    \includegraphics[scale=.3]{figures/posterior-satisfied-universals-efficiency.pdf}
%     \includegraphics[scale=.3]{figures/posterior-satisfied-universals-langmod.pdf}
%     \includegraphics[scale=.3]{figures/posterior-satisfied-universals-parseability.pdf}
%    \includegraphics[scale=.3]{figures/posterior-satisfied-universals-depl.pdf}
%	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks.}
%    \label{fig:posterior-satisfied-count}
%\end{figure}

%We argued that efficiency predicts all 10 correlation universals in the direction manifested in the UD corpora, as for each universal, the posterior probability that less than $50 \%$ of the optimized grammars satisfy the correlation was less than $0.05$.
%To address potential multiple comparisons problems, we use a single logistic model to show that efficiency predicts all ten correlations with high posterior probability.

%For each objective function, we constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency, with random effects for language and language family, correlated across the ten dependencies.

%We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
%The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
%The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$. The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.

\section{Relation to the Information Bottleneck}

Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
\begin{equation}
    \label{eq:efficiency-again}
    J_T(L) = I[L(T); T] - \lambda H[L(T)].
\end{equation}
This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}. In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information. The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$. Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
\begin{equation}
    \label{eq:ib}
    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
\end{equation}
In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.

In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$. To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$. We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized. Then our objective is parallel to Eq.~\ref{eq:ib}:
\begin{equation}
    \label{eq:efficiency-kinda-ib}
    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
\end{equation}
Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$. Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
\begin{equation*}
    J_T(L) = I[U;T] - \lambda H[U].
\end{equation*}

Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.

\section{Possible values of $\lambda$}

In the efficiency objective~(\ref{eq:efficiency-again}), the value of $\lambda$ is constrained to be in $[0,1)$.
This means, surprisal must be weighted less strongly than parsability.

Greater values of $\lambda$ mathematically result in degenerate solutions.
To show this, note that the following inequality always holds:
\begin{equation}
I[L(T); T] \leq H[L(T)].
\end{equation}
Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
\begin{equation}
    J_T(L) = I[L(T); T] - \lambda H[L(T)] \leq 0.
\end{equation}
and it takes the maximally possible value zero if $L(T)$ maps all trees to a single utterance, in which case both $I[L(T); T]$ and $H[L(T)]$ are zero.
This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.

In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order     correlations and relative optimality of language.

\section{Joint optimization of grammar, parser, and language model parameters}

For each of our objective functions, we seek parameters $\theta$ of a word order grammar $L_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $L_\theta$. The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers. The optimization problems come out to:
\begin{align}
\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
\end{align}
Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.

The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$. A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.

The case of predictability, parseability, and efficiency are more challenging, as we need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
In each step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
\begin{equation}\label{eq:estimator}
{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
\end{equation}
%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
for the gradient
$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.

In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
%and $\ell_2$ regularization.

\section{Implementation Details}

\paragraph{Locality}
In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.

\paragraph{Language Model}
We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
Given the small size of the corpora, this limit is only attained only for few languages.
In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
The model predicts both the next word and its POS tags in each step.
Using POS tags is intended to prevent overfitting on small corpora.
These decisions were made before evaluating word order properties.

\paragraph{Parser}
We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}. This architecture is remarkably simple: the words of a sentence are encoded into context-sensitive embeddings using bidirectional LSTMs, then a classifier is trained to predict the head for each work. The classifier works by calculating a score for every pair of word embeddings $(w_i, w_j)$, indicating the likelihood that the $j$th word is the head of the $i$th word. This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.

To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags. Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged. Parsing from POS tags reduces early overfitting.
This decision was made before evaluating word order properties.

\paragraph{Hyperparameters}
For locality, predictability, and parseability, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function on the language. We then used this setting for creating optimized word order grammars. % TODO more detail and maybe an example here
% TODO all design changes were made before examining word order universal results

% TODO Discuss iconicity somewhere somehow

All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
We do not use pretrained embeddings \citep{peters2018deep}: these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.

%\begin{table}[]
%    \centering
%    \begin{tabular}{c|cccc}
%    Entropy penalty     &  0.0001, 0.001 \\
%    Learning Rate (Word order grammar)     & 5e-6, 1e-5, 2e-5, 5e-5 \\
%    Momentum (Word order grammar) & 0.8, 0,9 \\
%    Dropout (Language model) & 0.0, 0.3, 0.5 \\
%    
%    \end{tabular}
%    \caption{Hyperparameters}
%    \label{tab:my_label}
%\end{table}

\paragraph{Optimization details}

We employ two common variance reduction methods to improve the estimator~(\ref{eq:estimator}), while keeping it unbiased.
For dependency length and predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word (and its head, in the case of dependency length). We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:estimator}) with lower variance.
Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple}.
These two methods reduce the variance of the estimator and thereby increase the speed of optimization and reduce training time, without biasing the results.
For numerical stability, we represent $a_\tau \in [0,1]$ via its logit $\in \mathbb{R}$.
Furthermore, to encourage exploration of the parameter space, we add an entropy penalty \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$. The weight of the entropy penalty was chosen together with the other hyperparameters.
%, adding the negative entropy of the $Bernoulli(a_\tau)$ to the loss $R$. %, weighted with $\alpha$.

These techniques for improving (\ref{eq:estimator}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.

We update word order grammar parameters $\theta$ using SGD with momentum.
For the language model parameters $\phi$, we use plain SGD. 
For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
The learning rates and momentum weight of SGD, and the parameters of Adam were determined together with the other hyperparameters.

For each language and objective function, we apply early stopping using a Monte-Carlo estimate of the respective objective functions on the UD development set. We compute an estimate of the objective on the development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate. This technique, well known in the machine learning literature, helps prevent overfitting to the specific corpora used.

The choice of optimization methods and the stopping criterion were fixed before we investigated word order correlations.

Recall that, for each language, we created 8 optimized languages for each optimization criterion.
We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
This maximizes statistical power for detecting and quantifying correlations between the \textit{obj} relation and other relations.

 % TODO we could use some language at the top of this section on the general guiding principles for all these decisions. Were these implementation decisions made before or after looking at the word order universals results? If they were made before, we should highlight that.

\section{Robustness to different language models and parsers}

Here we take up the question of the extent to which our results are dependent on the particular parser and language model used in the optimization process. We want to know: when we optimize a word order grammar for parseability, have we produced a language which is highly parseable \emph{in general}, or one which is highly parseable \emph{for a specific parser}? We wish to argue that natural language syntax is optimized for parseability in general, meaning that syntactic trees are highly recoverable from word orders in principle. If it turns out that our optimized languages are only optimal for a certain parser from the NLP literature, then we run the risk of circularity: it may be that the reason this parser was successful in the NLP literature was because it implicitly encoded word order universals in its inductive biases, and thus it would be no surprise that languages which are optimized for parseability also show those universals.

In this connection, we note that the parser and language model architectures we use are highly generic, and do not encode any obvious bias toward natural-language-like word orders. The LSTM language model is a generic model of sequence data which is also been used to model financial time series \citep{sirignano2018universal} and purely theoretical chaotic dynamical systems \citep{ogunmolu2016nonlinear}; the neural graph-based parser is simply solving a minimal spanning tree problem \citep{mcdonald2005nonprojective}. Nevertheless, it may be the case that a bias toward word order universals is somehow encoded implicitly in the hyperparameters and architectures of these models.

Here we address this question by demonstrating that our languages optimized for prediction and parseability are also optimal under a range of different language models and parsers. These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on a particular prediction and parsing algorithms.

\subsection{CYK Parsers}

We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of Collins 1997 (Model 1).
In our PCFGs, each head independently generates a set of left and right dependents.
We formulate this as a PCFG where each rule has the form:
\begin{center}
	POS$_H$ $\rightarrow$ POS$_H$ POS$_D$
\end{center}
for head-initial structures, and
\begin{center}
	POS$_H$ $\rightarrow$ POS$_D$ POS$_H$
\end{center}
for head-final structures, where each symbol is a POS tag.
Thus, POS tags act both as terminals and as nonterminals.

We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CYK algorithm.

This parsing strategy is very different from the neural graph-based parser:
While the graph-based parser solves a minimum spanning tree problem, the CKY algorithm uses dynamic programming to compute the exact probabilities of trees given a sentence, as specified by the generative model encoded in the PCFG.
Second, while the graph-based neural parser uses machine learning to induce syntactic knowledge from data, the CKY parser performs exact probabilistic inference.
%In this sense, the CKY algorithm does not have any architectural biases in itself.
%On the other hand, the PCFG makes severely simplifying independence assumptions, compared to the universal approximation capabilities of neural network-based systems.

We used the CKY algorithm to compute the parsing loss $H[T|L_\theta(T)]$ on the validation partition of the English UD corpora, for random and optimized ordering grammars.
Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
% TODO get UAS for compatibility with other parts of SI
%We then compared parsability of random and optimized word order grammars on English data.

\begin{figure}
    \centering
    \includegraphics[scale=.7]{figures/cky-parse.pdf} 
	\caption{Parsability loss $H[T|L_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).}
    \label{fig:cky-parser}
\end{figure}

\subsection{Distorted graph-based parsers}

Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization. Therefore, we address here the idea that the graph-based parser might have a built-in bias toward parses involving short dependencies, which we call a \key{locality bias}.

A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words. Suppose that a parser $P$ has a bias toward positing a dependency between a word and the immediate previous word.

\paragraph{Even--odd order.} A sequence of $n$ words originally ordered as $w_1 w_2 w_3 w_4 \cdots w_n$ is reordered by separating the even and odd indices: $w_2 w_4 w_6 \cdots w_{n-1} w_1 w_3 w_5 \cdots w_n$ (assuming $n$ odd). Therefore all words that are adjacent in the original order will be separated by a distance of $\approx n/2$ in the distorted order, while all words of distance 2 in the original order will become adjacent.

\paragraph{Interleaving order.} In interleaving ordering, a sequence originally ordered as $w_1 w_2 w_3 \cdots w_n$ is split in half at the middle (index $m=\ceiling{n/2}$), and the two resulting sequences are interleaved, yielding $w_1 w_m w_2 w_{m+1} w_3 w_{m+3} \cdots w_n$. Thus all words that were originally adjacent will have distance 2 in the distorted order, with the intervening word coming from a very distant part of the sentence.

\paragraph{Inwards order.} A sequence originally ordered as $w_1 w_2 w_3 \cdots w_{n-1} w_n$ is ordered from the edges of the string inwards, as $w_1 w_n w_2 w_{n-1} \cdots w_{\ceiling{n/2}}$. This corresponds to folding the string in on itself once, or equivalently, splitting the sequence in half at the middle, then interleaving the two resulting sequences after reversing the second one. The result is that the most non-local possible dependencies in the original order become the most local dependencies in the distorted order.

\paragraph{Sorted order.} A sequence is reordered by sorting by POS tags, and randomizing the order within each block of identical POS tags.
To each word, we then add a symbol encoding the original position in the sequence.
For instance
\begin{center}
PRON VERB PRON
\end{center}
may be reordered as
\begin{center}
PRON 1 PRON 3 VERB 2
\end{center}
or
\begin{center}
PRON 3 PRON 1 VERB 2
\end{center}
The numbers are provided to the parser as atomic symbols from a vocabulary ranging from 1 to 200; numbers greater than 200 (which may occur in extremely long sentences) are replaced by an out-of-range token.

\paragraph{Experiments}
Using English and Japanese data, we trained parsers for the ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
Resulting parsing accuracy scores are shown in Figure~\ref{fig:distorted-parser}.
In all settings, the language optimized for efficiency achieved higher parsing accuracy than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
%Note that absolute values of parsing accuracy vary between the orderings, being higher for the actual orderings than for distorted orderings, in which trees become nonprojective.

\begin{figure}
    \centering
    English

    \includegraphics[scale=.5]{figures/adversarial-parse-loss-english.pdf}

    Japanese

    \includegraphics[scale=.5]{figures/adversarial-parse-loss-japanese.pdf}
	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').}
    \label{fig:distorted-parser}
\end{figure}

\subsection{n-gram language models}

We constructed a Bigram model with Kneser-Ney smoothing.
A bigram model predicts each word taking only the previous word into account.
This contrasts with LSTMs, which take the entire context into consideration.
Thus, bigram models and LSTMs stand on opposing ends of a spectrum of language models taking more and more aspects of the context into account.

We estimated language models on the training partitions, and used the validation partitions to estimate surprisal.
We conducted this for the ten random and the best optimized ordering grammars on English and Japanese data.
Results (Figure~\ref{fig:bigrams}) show that languages optimized for efficiency are optimal for a bigram language model.

\begin{figure}
    \centering
    \includegraphics[scale=.4]{figures/bigrams.pdf} 
	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).}
    \label{fig:bigrams}
\end{figure}

\subsection{Discussion}

It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.

\section{Effects of data sparsity}

Our efficiency objective comes down to

If the difference between random and optimized languages is due to data sparsity, we expect that it decreases as the amount of training data is increased.
If, on the other hand, there is an inherent difference in efficiency between random and optimized languages, we expect that the difference persists as training data is increased.

We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
We cosidered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.

The resulting curves are shown in Figure~\ref{fig:learning-czech}.
A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.

\begin{figure}
    \centering
    \includegraphics[scale=.4]{figures/learning-parser-czech.pdf} 
        \caption{Parsing accuracy (measured in UAS) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
    \label{fig:learning-czech}
\end{figure}

\section{}

\bibliographystyle{acl_natbib}
\bibliography{everything}

\end{document}

### 2019-01-19 writeup/paper-names.tex
rewired figure paths

@@
     \centering
-    \includegraphics[scale=.4]{figures/pareto-plane-iso-best-balanced-legend.pdf}
+    \includegraphics[scale=.4]{../results/plane/pareto-plane-iso-best-balanced-legend.pdf}
     \caption{Predictability and parseability of 51 UD languages (red), indicated by ISO codes, compared to ten baseline word order grammars per language (green).
@@
     \centering
-    \includegraphics[scale=.2]{figures/posterior-satisfied-universals-efficiency.pdf}
-     \includegraphics[scale=.2]{figures/posterior-satisfied-universals-langmod.pdf}
-     \includegraphics[scale=.2]{figures/posterior-satisfied-universals-parseability.pdf}
-    \includegraphics[scale=.2]{figures/posterior-satisfied-universals-depl.pdf}
+    \includegraphics[scale=.2]{../results/correlations/posterior-satisfied-universals-efficiency.pdf}
+     \includegraphics[scale=.2]{../results/correlations/posterior-satisfied-universals-langmod.pdf}
+     \includegraphics[scale=.2]{../results/correlations/posterior-satisfied-universals-parseability.pdf}
+    \includegraphics[scale=.2]{../results/correlations/posterior-satisfied-universals-depl.pdf}
 	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks.}

### 2019-01-19 writeup/pnas-word-order-si.tex
rewired figure paths

@@
 
-    \includegraphics[scale=.25]{figures/coverage-langmod-best.pdf}
-    \includegraphics[scale=.25]{figures/coverage-parse-best.pdf}
-    \includegraphics[scale=.25]{figures/coverage-two-lambda09-best.pdf}
+    \includegraphics[scale=.25]{../results/correlations/coverage-langmod-best.pdf}
+    \includegraphics[scale=.25]{../results/correlations/coverage-parse-best.pdf}
+    \includegraphics[scale=.25]{../results/correlations/coverage-two-lambda09-best.pdf}
 
-    \includegraphics[scale=.25]{figures/coverage-ground.pdf}
-    \includegraphics[scale=.25]{figures/coverage-depl.pdf}
+    \includegraphics[scale=.25]{../results/correlations/coverage-ground.pdf}
+    \includegraphics[scale=.25]{../results/correlations/coverage-depl.pdf}
 	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
@@
     \centering
-    \includegraphics[scale=.7]{figures/posterior_joint_with_ud.pdf}
+    \includegraphics[scale=.7]{../results/correlations/posterior_joint_with_ud.pdf}
 	\caption{Posteriors of the prevalences of the correlations among languages optimized for Predictability (red), parseability (green), and efficiency (blue), controlling for languages and language families, with prevalence in the UD languages.}
@@
     \centering
-    \includegraphics[scale=.7]{figures/cky-parse.pdf} 
+    \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
 	\caption{Parsability loss $H[T|L_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).}
@@
 
-    \includegraphics[scale=.5]{figures/adversarial-parse-loss-english.pdf}
+    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-english.pdf}
 
@@
 
-    \includegraphics[scale=.5]{figures/adversarial-parse-loss-japanese.pdf}
+    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
 	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').}
@@
     \centering
-    \includegraphics[scale=.4]{figures/bigrams.pdf} 
+    \includegraphics[scale=.4]{../results/bigrams/bigrams.pdf} 
 	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).}

### 2019-01-19 writeup/pnas-word-order-si.tex
rewiring correlations analysis

@@
 
-    \includegraphics[scale=.25]{../results/correlations/coverage-langmod-best.pdf}
-    \includegraphics[scale=.25]{../results/correlations/coverage-parse-best.pdf}
-    \includegraphics[scale=.25]{../results/correlations/coverage-two-lambda09-best.pdf}
+    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
+    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
+    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
 
-    \includegraphics[scale=.25]{../results/correlations/coverage-ground.pdf}
-    \includegraphics[scale=.25]{../results/correlations/coverage-depl.pdf}
+    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
+    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
 	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
@@
     \centering
-    \includegraphics[scale=.7]{../results/correlations/posterior_joint_with_ud.pdf}
+    \includegraphics[scale=.7]{../results/correlations/figures/posterior_joint_with_ud.pdf}
 	\caption{Posteriors of the prevalences of the correlations among languages optimized for Predictability (red), parseability (green), and efficiency (blue), controlling for languages and language families, with prevalence in the UD languages.}

### 2019-01-19 results/correlations/evaluation-coarse-depl-prevalence-multiple.R

@@ -1,5 +1,5 @@
 
-data = read.csv("CS_SCR/deps/manual_output_funchead_coarse_depl/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+data = read.csv("../../grammars/manual_output_funchead_coarse_depl/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
 
 #/u/scr/mhahn/deps/manual_output_funchead_coarse_depl/
 
@@ -12,7 +12,7 @@ dryer_greenberg_fine = data
 
 
 
-languages = read.csv("languages-iso_codes.tsv")
+languages = read.csv("../languages/languages-iso_codes.tsv")
 dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"), all.x=TRUE)
 
 
@@ -35,9 +35,7 @@ getCorrPair = function(dependency) {
 }
 
 corr_pair = getCorrPair("lifted_cop")
-#model2 = brm(correlator_s ~ obj_s + (1+obj_s|Family) + (1+obj_s|Language), family="bernoulli", data=corr_pair, iter=5000)
 
-# Simpler analysis for posterior prevalence of correlation
 model3 = brm(agree ~ (1|Family) + (1|Language), family="bernoulli", data=corr_pair)
    samples = posterior_samples(model3, "b_Intercept")[,]
    posteriorOpposite = ecdf(samples)(0.0)
@@ -51,11 +49,11 @@ dependencies = c("acl", "advmod", "aux", "lifted_case", "lifted_cop", "lifted_ma
 
 
 
-sink("results-prevalence-depl.tsv")
+sink("output/results-prevalence-depl.tsv")
 cat("")
 sink()
 
-cat(paste("dependency", "satisfiedFraction", "posteriorMean", "posteriorSD", "posteriorOpposite", sep="\t"), file="results-prevalence-depl.tsv", append=TRUE, sep="\n")
+cat(paste("dependency", "satisfiedFraction", "posteriorMean", "posteriorSD", "posteriorOpposite", sep="\t"), file="output/results-prevalence-depl.tsv", append=TRUE, sep="\n")
 
 
 for(dependency in dependencies) {
@@ -68,7 +66,7 @@ for(dependency in dependencies) {
    posteriorMean = mean(samples)
    posteriorSD = sd(samples)
[…]

### 2019-01-19 results/correlations/evaluation-coarse-depl.R

@@ -1,60 +1,0 @@
-
-data = read.csv("CS_SCR/deps/manual_output_funchead_coarse_depl/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
-
-#/u/scr/mhahn/deps/manual_output_funchead_coarse_depl/
-
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-
-dryer_greenberg_fine = data
-
-
-
-languages = read.csv("languages.tsv", sep="\t")
-dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"), all.x=TRUE)
-
-
-library("brms")
-
-#options(mc.cores = parallel::detectCores())
-rstan_options(auto_write = TRUE)
-
-dependency = "nmod"
-
-getCorrPair = function(dependency) {
-   corr_pair = dryer_greenberg_fine %>% filter((Dependency == dependency) | (Dependency == "obj"))
-   corr_pair = unique(corr_pair %>% select(Family, Language, FileName, Dependency, DH_Weight, ModelName)) %>% spread(Dependency, DH_Weight)
-   corr_pair$correlator = corr_pair[[dependency]]
-   corr_pair = corr_pair %>% mutate(correlator_s = pmax(0,sign(correlator)), obj_s=pmax(0,sign(obj)))
-   return(corr_pair)
-}
-
-corr_pair = getCorrPair("nmod")
-model2 = brm(correlator_s ~ obj_s + (1+obj_s|Family) + (1+obj_s|Language), family="bernoulli", data=corr_pair)
-
-dependencies = c("acl", "advcl", "advmod", "amod", "appos", "aux", "ccomp", "compound", "conj", "csubj", "dep", "det", "discourse", "dislocated", "expl", "fixed", "flat", "goeswith", "iobj", "lifted_case", "lifted_cc", "lifted_cop", "lifted_mark", "list", "nmod", "nsubj", "nummod", "obl", "orphan", "parataxis", "reparandum", "vocative", "xcomp")
-
-
-
-sink("results-depl.tsv")
-cat("")
-sink()
-
-cat(paste("dependency", "satisfiedFraction", "posteriorMean", "posteriorSD", "posteriorOpposite", sep="\t"), file="results-depl.tsv", append=TRUE, sep="\n")
-
-
-for(dependency in dependencies) {
-   corr_pair = getCorrPair(dependency)
[…]

### 2019-01-19 results/correlations/evaluation-coarse-ground.R

@@ -1,86 +1,0 @@
-
-data = read.csv("CS_SCR/deps/manual_output_funchead_ground_coarse_final/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
-
-
-
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-
-dryer_greenberg_fine = data %>% mutate(DH_Weight = DH_Mean_NoPunct)
-
-
-
-languages = read.csv("languages.tsv", sep="\t")
-dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"), all.x=TRUE)
-
-#dryer_greenberg_fine %>% group_by(Dependency) %>% summarise(DH_Weight = quantile(DH_Weight - 2*DH_Sigma_NoPunct,0.999)) %>% print(n=50
-#dryer_greenberg_fine %>% group_by(Dependency) %>% summarise(DH_Weight = quantile(DH_Weight + 2*DH_Sigma_NoPunct,0.001)) %>% print(n=50
-
-
-# Interest in general biases
-#dryer_greenberg_fine %>% group_by(Dependency) %>% summarise(DH_Weight = quantile(DH_Weight,0.1)) %>% print(n=50)
-#dryer_greenberg_fine %>% group_by(Dependency) %>% summarise(DH_Weight = quantile(DH_Weight,0.9)) %>% print(n=50)
-#+
-#advmod
-#clf
-#det
-#expl
-#nsubj
-#nummod
-#
-#- things that come after the head in basically all languages
-#appos
-#conj
-#fixed
-#flat
-#list
-# hist((dryer_greenberg_fine %>% filter(Dependency == "appos"))$DH_Weight) manual inspecton
-
-# dryer_greenberg_fine %>% group_by(Dependency) %>% summarise(Distance_Mean_NoPunct = quantile(Distance_Mean_NoPunct,0.9)) %>% print(n=50
-# dryer_greenberg_fine %>% group_by(Dependency) %>% summarise(Distance_Mean_NoPunct = quantile(Distance_Mean_NoPunct,0.1)) %>% print(n=50)
-
-library("rstan")
-library("brms")
-
-options(mc.cores = parallel::detectCores())
-rstan_options(auto_write = TRUE)
-
-dependency = "nmod"
-
-getCorrPair = function(dependency) {
-   corr_pair = dryer_greenberg_fine %>% filter((Dependency == dependency) | (Dependency == "obj"))
[…]

### 2019-01-19 results/correlations/evaluation-coarse-langmod-prevalence-multiple.R

@@ -1,10 +1,10 @@
-data = read.csv("CS_SCR/deps/manual_output_funchead_langmod_coarse_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+data = read.csv("../../grammars/manual_output_funchead_langmod_coarse_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
 library(forcats)
 library(dplyr)
 library(tidyr)
 library(ggplot2)
 data = data %>% mutate(Language = fct_recode(Language, "Ancient_Greek" = "Ancient", "Old_Church_Slavonic" = "Old"))
-languages = read.csv("languages-iso_codes.tsv")
+languages = read.csv("../languages/languages-iso_codes.tsv")
 data  = merge(data, languages, by=c("Language"), all.x=TRUE)
 library("brms")
 dependency = "nmod"

### 2019-01-19 results/correlations/evaluation-coarse-parser-prevalence-multiple.R

@@ -1,10 +1,10 @@
-data = read.csv("CS_SCR/deps/manual_output_funchead_two_coarse_parser_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+data = read.csv("../../grammars/manual_output_funchead_two_coarse_parser_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
 library(forcats)
 library(dplyr)
 library(tidyr)
 library(ggplot2)
 data = data %>% mutate(Language = fct_recode(Language, "Ancient_Greek" = "Ancient", "Old_Church_Slavonic" = "Old"))
-languages = read.csv("languages-iso_codes.tsv")
+languages = read.csv("../languages/languages-iso_codes.tsv")
 data  = merge(data, languages, by=c("Language"), all.x=TRUE)
 library("brms")
 dependency = "nmod"

### 2019-01-19 results/correlations/evaluation-coarse-parser.R

@@ -1,60 +1,0 @@
-
-data = read.csv("CS_SCR/deps/manual_output_funchead_two_coarse_parser/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
-
-
-
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-
-dryer_greenberg_fine = data
-
-
-
-languages = read.csv("languages.tsv", sep="\t")
-dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"), all.x=TRUE)
-
-
-library("brms")
-
-#options(mc.cores = parallel::detectCores())
-rstan_options(auto_write = TRUE)
-
-dependency = "nmod"
-
-getCorrPair = function(dependency) {
-   corr_pair = dryer_greenberg_fine %>% filter((CoarseDependency == dependency) | (CoarseDependency == "obj"))
-   corr_pair = unique(corr_pair %>% select(Family, Language, FileName, CoarseDependency, DH_Weight, ModelName)) %>% spread(CoarseDependency, DH_Weight)
-   corr_pair$correlator = corr_pair[[dependency]]
-   corr_pair = corr_pair %>% mutate(correlator_s = pmax(0,sign(correlator)), obj_s=pmax(0,sign(obj)))
-   return(corr_pair)
-}
-
-corr_pair = getCorrPair("nmod")
-model2 = brm(correlator_s ~ obj_s + (1+obj_s|Family) + (1+obj_s|Language), family="bernoulli", data=corr_pair)
-
-dependencies = c("acl", "advcl", "advmod", "amod", "appos", "aux", "ccomp", "compound", "conj", "csubj", "dep", "det", "discourse", "dislocated", "expl", "fixed", "flat", "goeswith", "iobj", "lifted_case", "lifted_cc", "lifted_cop", "lifted_mark", "list", "nmod", "nsubj", "nummod", "obl", "orphan", "parataxis", "reparandum", "vocative", "xcomp")
-
-
-
-sink("results-parser.tsv")
-cat("")
-sink()
-
-cat(paste("dependency", "satisfiedFraction", "posteriorMean", "posteriorSD", "posteriorOpposite", sep="\t"), file="results-parser.tsv", append=TRUE, sep="\n")
-
-
-for(dependency in dependencies) {
-   corr_pair = getCorrPair(dependency)
-   model2 = update(model2, newdata=corr_pair, iter=4000)
[…]

### 2019-01-19 results/correlations/evaluation-coarse-two.R

@@ -1,61 +1,0 @@
-
-# _final/
-data = read.csv("CS_SCR/deps/manual_output_funchead_two_coarse_final/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
-
-
-
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-
-dryer_greenberg_fine = data
-
-
-
-languages = read.csv("languages-iso_codes.tsv")
-dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"), all.x=TRUE)
-
-
-library("brms")
-
-#options(mc.cores = parallel::detectCores())
-rstan_options(auto_write = TRUE)
-
-dependency = "nmod"
-
-getCorrPair = function(dependency) {
-   corr_pair = dryer_greenberg_fine %>% filter((CoarseDependency == dependency) | (CoarseDependency == "obj"))
-   corr_pair = unique(corr_pair %>% select(Family, Language, FileName, CoarseDependency, DH_Weight )) %>% spread(CoarseDependency, DH_Weight)
-   corr_pair$correlator = corr_pair[[dependency]]
-   corr_pair = corr_pair %>% mutate(correlator_s = pmax(0,sign(correlator)), obj_s=pmax(0,sign(obj)))
-   return(corr_pair)
-}
-
-corr_pair = getCorrPair("nmod")
-model2 = brm(correlator_s ~ obj_s + (1+obj_s|Family) + (1+obj_s|Language), family="bernoulli", data=corr_pair)
-
-dependencies = c("acl", "advcl", "advmod", "amod", "appos", "aux", "ccomp", "compound", "conj", "csubj", "dep", "det", "discourse", "dislocated", "expl", "fixed", "flat", "goeswith", "iobj", "lifted_case", "lifted_cc", "lifted_cop", "lifted_mark", "list", "nmod", "nsubj", "nummod", "obl", "orphan", "parataxis", "reparandum", "vocative", "xcomp")
-
-
-
-sink("results-two.tsv")
-cat("")
-sink()
-
-cat(paste("dependency", "satisfiedFraction", "posteriorMean", "posteriorSD", "posteriorOpposite", sep="\t"), file="results-two.tsv", append=TRUE, sep="\n")
-
-
-for(dependency in dependencies) {
-   corr_pair = getCorrPair(dependency)
-   model2 = update(model2, newdata=corr_pair, iter=4000)
[…]

### 2019-01-19 results/correlations/tiles-depl-best.R

@@ -1,3 +1,4 @@
+# Creating the tiles plot
 
 data = read.csv("CS_SCR/deps/manual_output_funchead_coarse_depl/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
 

### 2019-01-19 writeup/pnas-word-order-si.tex
reproduced per-dependency results

@@
 
+A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
+Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
+We thus do not evaluate this objective function here.
+
 \section{Implementation Details}
@@
 For numerical stability, we represent $a_\tau \in [0,1]$ via its logit $\in \mathbb{R}$.
-Furthermore, to encourage exploration of the parameter space, we add an entropy penalty \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$.
-The weight of the entropy penalty was chosen together with the other hyperparameters.
+Furthermore, to encourage exploration of the parameter space, we add an entropy regularization term \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$.
+The weight of the entropy regularization was chosen together with the other hyperparameters.
 %, adding the negative entropy of the $Bernoulli(a_\tau)$ to the loss $R$. %, weighted with $\alpha$.
@@
 For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
-The learning rates and momentum weight of SGD, and the parameters of Adam were determined together with the other hyperparameters.
+The learning rates and other optimization hyperparameters were determined together with the other hyperparameters.
 
-For each language and objective function, we apply early stopping using a Monte-Carlo estimate of the respective objective functions on the UD development set.
-We compute an estimate of the objective on the development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
+For each language and objective function, we apply Early Stopping~\citep{prechelt1998early} as follows.
+We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
 This technique, well known in the machine learning literature, helps prevent overfitting to the specific corpora used.

### 2019-01-20 writeup/paper-names.tex
revising paper

--- 2019-01-19 writeup/paper-names.tex
@@
     \begin{deptext}
-    I \& wrote \& a \& long \& letter \& to \& a \& friend \\
+    I \& wrote \& a \& letter \& to \& a \& friend \\
     \end{deptext}
     \depedge{2}{1}{nsubj}
-    \depedge{2}{5}{obj}
-    \depedge{5}{4}{ }
-    \depedge{5}{3}{ }
-    \depedge{2}{6}{}
-    \depedge{6}{8}{ }
-    \depedge{8}{7}{ }
+    \depedge{2}{4}{obj}
+    \depedge{4}{3}{ }
+    \depedge{4}{3}{ }
+    \depedge{2}{5}{}
+    \depedge{5}{7}{ }
+    \depedge{7}{6}{ }
 \end{dependency}
@@
 % do we need to define "linearize"?
-
-%For example, TODO % We should have some pretty picture visualizing the process here
+This process is visualized in Figure~\ref{fig:process-visualization}.
 
 Next we define how a word order grammar linearizes trees in terms of its parameters $\theta$, and then we describe how $\theta$ can be set to optimize objective functions related to communicative efficiency.
+
+\begin{figure}
+\includegraphics[width=0.48\textwidth]{figures/process-visualization.png}
+\caption{Our method for constructing counterfactual versions of languages: We extract unordered dependency tree topologies from dependency corpora, and apply word order grammars to obtain counterfactually reordered corpora.}\label{fig:process-visualization}
+\end{figure}
 
@@
 
-See the SI for further details on the joint optimization of $\theta$ and $\psi$ and on the neural parser architecture.
+See the SI (appendix section 6-7) for further details on the joint optimization of $\theta$ and $\psi$ and on the neural parser architecture, and see SI (section 8) for robustness to different parsing models.
 
@@
 with an interpolation weight $\lambda \in [0,1)$.
-In all experiments in this paper we use $\lambda=.9$.
+In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification).
 This combined equation represents efficiency: bits of information transferred about dependency trees minus the cost of the transmission.
@@
 Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking unordered dependency trees $T$ as the underlying meanings to be conveyed.
-Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI for the precise relationship).
+Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
 
@@
     \centering
-    \includegraphics[scale=.4]{figures/pareto-plane-iso-best-balanced-legend.pdf}
+    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend.pdf}
     \caption{Predictability and parseability of 51 UD languages (red), indicated by ISO codes, compared to ten baseline word order grammars per language (green).
@@
 \begin{figure*}
-\includegraphics[width=0.98\textwidth]{figures/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
+\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
 \caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right).
@@
     \centering
-    \includegraphics[scale=.2]{figures/posterior-satisfied-universals-efficiency.pdf}
-     \includegraphics[scale=.2]{figures/posterior-satisfied-universals-langmod.pdf}
-     \includegraphics[scale=.2]{figures/posterior-satisfied-universals-parseability.pdf}
-    \includegraphics[scale=.2]{figures/posterior-satisfied-universals-depl.pdf}
+     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-efficiency.pdf}
+     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-langmod.pdf}
+     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}
+     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-depl.pdf}
 	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks.}
@@
 We believe this discrepancy is a result of noise in the optimization process.
-Optimizing for efficiency is computationally challenging, requiring stochastic gradient descent on word order grammar, language model, and parser parameters, while optimizing for dependency length is simpler and thus more optimal grammars can be reached easily.
+Optimizing for efficiency is computationally challenging, requiring stochastic gradient descent on word order grammar, language model, and parser parameters, while optimizing for dependency length is simpler and thus more optimal grammars can be reached easily (See SI section 2 for supporting data).
 Inasmuch as dependency length minimization is computationally easier to implement than efficiency maximization, we believe it will remain useful as an easy-to-use heuristic for predicting and explaining word order universals.
@@
 
-We found that a large subset of the Greenbergian word order correlations can be explained in terms of from optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
+We found that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
 We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals.

### 2019-01-20 results/correlations/visualize-joint-evaluation-depl.R

@@ -14,7 +14,9 @@ data2 = data %>% group_by(satisfiedCount) %>% summarise(posterior = NROW(satisfi
 
 #plot = ggplot(data = data, aes(x=satisfiedCount)) + geom_histogram() + theme_bw() + xlim(0,10.5) + ggtitle("Dependency Length") + xlab("Number of Predicted Correlations") + ylab("Number of Posterior Samples")
 
-plot = ggplot(data = data2, aes(x=satisfiedCount, y=posterior)) + geom_bar(stat="identity") + theme_bw() + xlim(0,10.5) + ggtitle("Dependency Length") + xlab("Number of Predicted Correlations") + ylab("Number of Posterior Samples") + geom_text(aes(label=posteriorProb), vjust=0) + ylim(0, 1.1*nrow(data))
+plot = ggplot(data = data2, aes(x=satisfiedCount, y=posterior)) + geom_bar(stat="identity") + theme_bw() + xlim(0,10.5) + ggtitle("Dependency Length") + xlab("Number of Predicted Correlations") + ylab("Number of Posterior Samples") + geom_text(aes(label=posteriorProb), vjust=0, size=4.3) + ylim(0, 1.1*nrow(data)) +
+    theme(text = element_text(size=20),
+        axis.text.x = element_text(angle=90, hjust=1)) 
 
 ggsave(plot=plot, filename="figures/posterior-satisfied-universals-depl.pdf")
 

### 2019-01-20 results/correlations/visualize-joint-evaluation-together.R

@@ -1,0 +1,36 @@
+
+library(dplyr)
+library(tidyr)
+
+
+u_depl = read.csv("/home/user/CS_SCR/posteriors/posterior-10-depl.csv") %>% mutate(Type = "Dependency Length")
+u_langmod = read.csv("/home/user/CS_SCR/posteriors/posterior-10-langmod.csv") %>% mutate(Type = "Predictability")
+u_parser = read.csv("/home/user/CS_SCR/posteriors/posterior-10-parseability.csv") %>% mutate(Type = "Parseability")
+u_efficiency = read.csv("/home/user/CS_SCR/posteriors/posterior-10-efficiency.csv") %>% mutate(Type = "Efficiency")
+
+
+u = rbind(u_depl, u_langmod, u_parser, u_efficiency)
+
+satisfied = 10 - ((u$b_acl_Intercept < 0) + (u$b_advmod_Intercept > 0)  + (u$b_aux_Intercept > 0 ) + (u$b_liftedcase_Intercept < 0 ) + (u$b_liftedcop_Intercept < 0 ) + (u$b_liftedmark_Intercept < 0 ) + (u$b_nmod_Intercept < 0 ) + (u$b_nsubj_Intercept > 0 ) + (u$b_obl_Intercept < 0 ) + (u$b_xcomp_Intercept < 0 ))
+
+library(ggplot2)
+
+data = data.frame(satisfiedCount=satisfied)
+
+samplesPerType = data %>% group_by(Type) %>% summarise(SamplesNum = nrow(satisfied))
+
+data = merge(data, samplesPerType, by=c("Type"))
+
+data2 = data %>% group_by(Type, satisfiedCount) %>% summarise(posterior = NROW(satisfiedCount)) %>% mutate(posteriorProb = posterior/SamplesNum)
+
+
+#plot = ggplot(data = data, aes(x=satisfiedCount)) + geom_histogram() + theme_bw() + xlim(0,10.5) + ggtitle("Dependency Length") + xlab("Number of Predicted Correlations") + ylab("Number of Posterior Samples")
+
+plot = ggplot(data = data2, aes(x=satisfiedCount, y=posterior)) + geom_bar(stat="identity") + theme_bw() + xlim(0,10.5) + ggtitle("Dependency Length") + xlab("Number of Predicted Correlations") + ylab("Number of Posterior Samples") + geom_text(aes(label=posteriorProb), vjust=0, size=4.3) + ylim(0, 1.1*nrow(data)) +
[…]

### 2019-01-21 writeup/paper-names.tex
Dependency length

@@
 \begin{figure*} % need Efficiency and (maybe) MLE langs on here
-\includegraphics[width=0.98\textwidth]{figures/depLength-facet.pdf}
+\includegraphics[width=0.98\textwidth]{../results/dependency-length/figures/depLength-facet.pdf}
 \caption{Average dependency length as a function of sentence length in four languages.
@@
     \centering
-     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-efficiency.pdf}
-     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-langmod.pdf}
-     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}
-     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-depl.pdf}
+%     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-efficiency.pdf}
+ %    \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-langmod.pdf}
+  %   \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}
+   %  \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-depl.pdf}
+     \includegraphics[scale=.45]{../results/correlations/figures/posterior-satisfied-universals-together.pdf}
+
 	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks.}

### 2019-01-21 results/correlations/visualize-joint-evaluation-together.R

@@ -22,9 +22,9 @@ u = NULL
 
 #plot = ggplot(data = data, aes(x=satisfiedCount)) + geom_histogram() + theme_bw() + xlim(0,10.5) + ggtitle("Dependency Length") + xlab("Number of Predicted Correlations") + ylab("Number of Posterior Samples")
 
-plot = ggplot(data = data2, aes(x=satisfied, y=posterior)) + geom_bar(stat="identity") + theme_bw() + xlim(0,10.5) + ggtitle("Dependency Length") + xlab("Number of Predicted Correlations") + ylab("Number of Posterior Samples") + geom_text(aes(label=posteriorProb), vjust=0, size=4.3) + ylim(0, 1.1*nrow(data)) +
+plot = ggplot(data = data2, aes(x=satisfied, y=posterior)) + geom_bar(stat="identity") + theme_bw() + xlim(0,10.5) + xlab("Number of Predicted Correlations") + ylab("Number of Posterior Samples") + geom_text(aes(label=posteriorProb), vjust=0, size=3.3) + ylim(0, 1.1*max(v$SamplesNum)) +
     theme(text = element_text(size=20),
-        axis.text.x = element_text(angle=90, hjust=1)) + facet_grid(~Type)
+        axis.text.x = element_text(angle=90, hjust=1)) + facet_wrap(~Type, ncol=2)
 
 ggsave(plot=plot, filename="figures/posterior-satisfied-universals-together.pdf")
 

### 2019-01-21 results/dependency-length/dependencyLength_FuncHead_Coarse.R

@@ -1,29 +1,16 @@
-#data = read.csv("CS_SCR/deps/dependency_length/total_summary_funchead_coarse.tsv", sep="\t")# %>% rename(Quality=AverageLength)
 library(tidyr)
 library(dplyr)
-
-#data = data %>% filter(grepl("FuncHead", ModelName)) %>% filter(grepl("Coarse", ModelName))
-##dataModels = data %>% filter(!(Model %in% c("RANDOM", "REAL")))
-#modelData = read.csv("CS_SCR/deps//models_log_summary.tsv", sep="\t")
-#data = merge(data, modelData %>% select(FileName, Command, Model) %>% rename(ModelOriginal=Model, Model=FileName), by=c("Model"), all.x=TRUE)
-#data = data[order(data$AverageLengthPerWord),]
-#data = data %>% mutate(LSTM = grepl("Running", ModelOriginal), DepL = grepl("DepL", ModelOriginal), Parsing = grepl("Reinf", ModelOriginal), Real = (Model == "REAL"), Random = (Model == "RANDOM") )
-#data = data %>% mutate(type = case_when(LSTM ~ "LSTM", DepL ~ "DepL", Parsing ~ "Parsing", Real ~ "Real", Random ~ "Random", TRUE ~ "OTHER"))
-#data = data %>% filter(Real | Random | Temperature == Inf)
-#bestLSTM = data %>% filter(LSTM | Parsing) %>% group_by(Language, type) %>% summarise(BestOriginalLoss = min(OriginalLoss))
-#data = merge(data, bestLSTM, by=c("Language", "type"), all=TRUE)
-#data = data %>% filter(Real | Random | DepL | OriginalLoss == BestOriginalLoss)
-#data = data[order(data$AverageLengthPerWord),]
-#perLanguage = data %>% group_by(Language, type) %>% summarise(AverageLengthPerWord = mean(AverageLengthPerWord))
-#perLanguageSpread = perLanguage %>% spread(type, AverageLengthPerWord)
 
 data = data.frame()
 
 for(LanguageName in c("English", "Chinese", "Hindi", "Japanese")) {
-   LanguageName = "English"
-   
+      data2 = read.csv(paste("/home/user/CS_SCR/deps/dependency_length/", LanguageName, "_computeDependencyLengthsByType_NEWPYTORCH_Deterministic_FuncHead_REAL_REAL.py_model_REAL_REAL.tsv", sep=""), sep="\t")
[…]

### 2019-01-23 writeup/paper-names.tex
unified color scheme (which may require reworking)

@@
 
+% TODO where do we say that we're basing our work on data from 51 languages?
+
 \subsection{Explananda: Word Order Correlations}
@@
 
-\begin{dependency}[theme=simple]
-    \begin{deptext}
-    I \& wrote \& a \& letter \& to \& a \& friend \\
-    \end{deptext}
-    \depedge{2}{1}{nsubj}
-    \depedge{2}{4}{obj}
-    \depedge{4}{3}{ }
-    \depedge{4}{3}{ }
-    \depedge{2}{5}{}
-    \depedge{5}{7}{ }
-    \depedge{7}{6}{ }
-\end{dependency}
+%\begin{dependency}[theme=simple]
+%    \begin{deptext}
+%    I \& wrote \& a \& letter \& to \& a \& friend \\
+%    \end{deptext}
+%    \depedge{2}{1}{nsubj}
+%    \depedge{2}{4}{obj}
+%    \depedge{4}{3}{ }
+%    \depedge{4}{3}{ }
+%    \depedge{2}{5}{}
+%    \depedge{5}{7}{ }
+%    \depedge{7}{6}{ }
+%\end{dependency}
 	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese}
@@
 	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right).
-We provide cross-entropy for predictability (lower is better) and unlabeled attachment scores---a standard metric from natural language processing---for parseability (higher is better).}\label{fig:predict-pars}
+We provide the entropy $H[L_\theta(T)]$ for predictability (lower is better) and the conditional entropy $H[T|L_\theta(T)]$ for parseability (lower is better).
+Both measures are normalized by sentence length.}\label{fig:predict-pars}
 \end{figure*}
@@
 We see that optimizing for efficiency lowers dependency length relative to random baselines, in keeping with the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}.
-In 3/4 of the cases shown, optimizing explicitly for dependency length produces dependencies that overshoot the depenndency length of the real language; in Chinese and Hindi, the real language is best matched by efficiency optimization.
+In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the depenndency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.
 
@@
 	\caption{Greenbergian Correlations.
-Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner' (the `Correlate' column), whose relative order correlates with that of verbs and objects.
-	For clarity, we organize the correlations by the category of the `verb patterner'.
+Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%	For clarity, we organize the correlations by the category of the `verb patterner'.
 	For each correlation, we give our operationalization in terms of UD.
-A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
-	For each correlation we report how many of the languages in our sample satisfied it (`Real').
-We then report, for each correlation and each objective function, how many of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
+%A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
+	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
+	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
 \end{table*}
@@
 To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
-The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) in all except four languages, for WALS did not have data.
+The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
+%in all except four languages, for WALS did not have data.
 
@@
 
-	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks.}
+	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks, computed from a mixed-effects logistic regression jointly modeling all ten dependencies.
+Efficiency predicts all ten correlations with high posterior probability.}
     \label{fig:posterior-satisfied-count}
@@
 
-\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, and audiences at CAMP 2018 for helpful discussion.}
+\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Chris Manning, and audiences at CAMP 2018 for helpful discussion.}
 

### 2019-01-23 results/dependency-length/dependencyLength_FuncHead_Coarse_visualize.R

@@ -1,15 +1,32 @@
 
-data= read.csv("/home/user/CS_SCR/dependencyLength-intermediate.csv")
 
 library(tidyr)
 library(dplyr)
 library(ggplot2)
 
+data= read.csv("/home/user/CS_SCR/dependencyLength-intermediate.csv") %>% filter(TypeN != "MLE")
+
+
+library(forcats)
+
+data = data %>% mutate(TypeN = fct_recode(TypeN, "Parseability" = "Optimized for Pars.", "Predictability" = "Optimized for Pred.", "Efficiency" = "Optimized for Eff.", "Dependency Length" = "Optimized for Dep.L.", "Real Languages" = "Actual Language", "Baseline Languages" = "Baseline"))
+
+
+
+data = data %>% mutate(TypeN = factor(TypeN, levels=c("Parseability", "Predictability", "Efficiency", "Dependency Length", "Real Languages", "Baseline Languages"), ordered=TRUE))
+forColoring  = factor(data$TypeN, levels=c("Real Languages", "Baseline Languages", "Parseability", "Predictability", "Efficiency", "Dependency Length"), ordered=TRUE)
+#Create a custom color scale
+library(RColorBrewer)
+myColors <- brewer.pal(6,"Set1")
+names(myColors) <- levels(forColoring)
+
+
+
 plot = ggplot(data %>% filter(TypeN != "MLE"), aes(x=SentenceLength,y=DepLength,color=TypeN)) +
   geom_smooth(method = 'auto', se=F) +
   xlab("Sentence Length") +
   ylab("Dependency Length") +
-       scale_color_brewer(palette="Set1") +
+       scale_colour_manual(name = "TypeN",values = myColors) +
   theme_classic() + ylim(0,220) + xlim(0,50) +
   facet_wrap(~ Language, ncol=4)  +
   theme(legend.title=element_blank())

### 2019-01-23 results/plot-by-lang/analyzeLangmod.R

@@ -41,6 +41,16 @@ data = data %>% mutate(TypeN = fct_recode(Type, "Parseability" = "manual_output_
 
 
 
+
+
+data = data %>% mutate(TypeN = factor(TypeN, levels=c("Parseability", "Predictability", "Efficiency", "Dependency Length", "Real Languages", "Baseline Languages"), ordered=TRUE))
+forColoring  = factor(data$TypeN, levels=c("Real Languages", "Baseline Languages", "Parseability", "Predictability", "Efficiency", "Dependency Length"), ordered=TRUE)
+#Create a custom color scale
+library(RColorBrewer)
+myColors <- brewer.pal(6,"Set1")
+names(myColors) <- levels(forColoring)
+
+
   dodge = position_dodge(width = .9)
   dataL = data %>% filter(Language %in% c("Japanese", "English"))
   dataBars = dataL %>% filter(TypeN != "Baseline Languages") %>% group_by(Language, TypeN) %>% summarise(y=1, Surp=mean(Surp))
@@ -49,11 +59,39 @@ data = data %>% mutate(TypeN = fct_recode(Type, "Parseability" = "manual_output_
   plot1 = ggplot(dataL %>% filter(TypeN == "Baseline Languages"), aes(x=Surp,fill=TypeN)) +
     geom_density(aes(x=Surp, y=..scaled.., fill=TypeN), adjust=3) +
     geom_bar(stat="identity", width=widthOfBars, data=dataBars, aes(x=Surp, y=y, fill=TypeN, group=TypeN), position = position_dodge()) +
+    scale_fill_manual(name = "TypeN",values = myColors) +
     theme_classic() +
     theme(axis.title.y=element_blank(),
           axis.text.y=element_blank(),
           axis.ticks.y=element_blank(),
-          axis.line.y=element_blank()) + facet_grid(~Language, scales="free") + scale_fill_brewer(palette="Set2")
+          axis.line.y=element_blank()) + facet_grid(~Language, scales="free") +
+    theme(legend.title=element_blank()) +
+    xlab("Predictability")
+ 
+
+
+
+
+#plot = ggplot(dataPlot)
+#plot = plot + theme_bw()
+#plot = plot + geom_segment(aes(x=Pars_z, y=Surp_z, xend=Pars_z_end, yend = Surp_z_end, color=TypeN, group=TypeN), size=0.6)
[…]

### 2019-01-25 second-draft/paper-names.tex
attempting revision

--- 2019-01-23 writeup/paper-names.tex
@@
 We develop a probabilistic, differentiable model of word order grammars: the means by which different languages convert underlying hierarchical structures into strings of words.
-We show how the parameters of a word order grammar can be optimized for robustness of information transfer, as quantified using mutual information with latent tree structures, and efficiency of processing, as quantified using incremental predictability.
-Applying these grammars to tree structures found in dependency corpora, we show that optimizing the grammar parameters for efficiency and robustness results in word order patterns that reproduce a large subset of the major word order correlations reported in the linguistic typological literature, and reproduce the predictions of previous heuristic theories such as dependency length minimization.
+We show how the parameters of a word order grammar can be optimized for efficiency of information transfer, quantified as a tradeoff between incremental predictability and mutual information with latent tree structures.
+Applying these grammars to tree structures found in dependency corpora from 51 languages, we show that optimizing the grammar parameters for efficiency results in word order patterns that reproduce a large subset of the major word order correlations reported in the linguistic typological literature, and reproduce the predictions of previous heuristic theories such as dependency length minimization.
 \end{abstract}
@@
 An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which certain syntactic universals can be explained through the statistical optimization of grammars for information-theoretic efficiency.
+In this work we examine this question from a computational perspective, demonstrating a fully formalized framework in which certain syntactic universals can be explained through the statistical optimization of grammars for information-theoretic efficiency.
 % TODO add a cogsci person to citation chain at the beginning
@@
 These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages.
-For example, whether a language puts verbs before or after objects strongly determines whether it places adjectives before or after nouns, and whether it had prepositions or postpositions \cite{dryer1992greenbergian}.
+For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
 
@@
 
-Researchers working in the functional paradigm have proposed a range of criteria which languages should meet in order to enable efficient communication and processing.
+Researchers working in the functional paradigm have proposed a range of criteria that languages should meet in order to enable efficient communication and processing.
 These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
@@
 We quantify the difficulty of linguistic communication in terms of the word-by-word \key{predictability} of strings, which has been found to be a highly accurate and general predictor of human online comprehension difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and which can be justified in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
-We quantify the information transferred in language as \key{parseability}: the quantity of information contained in linguistic utterances about latent hierarchical tree structures.
+We quantify the information transferred in language as \key{parseability}: the quantity of information linguistic utterances contain about latent hierarchical tree structures.
 We also compare the results of optimizing for communicative efficiency to the results of optimizing for dependency length minimization, a previously proposed approximate metric of the amount of memory involved in language production and comprehension \cite{hudson1995measuring,gibson1998linguistic,gibson2000dependency,liu2017dependency,temperley2018minimizing}.
@@
 \end{equation*}
-where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\L_\theta$ as accurately as possible (maximizing predictability):
+where the language model parameters $\phi$ are chosen to predict the language induced by tree distribution $T$ and word order grammar $\L_\theta$ as accurately as possible (maximizing predictability):
 \begin{equation*}
@@
 	%&
-	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       &     \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\
-	verb & object &      & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
+	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
 %adp. &
-adposition    &    NP    &    lifted\_case    &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
-copula    &    NP    &    lifted\_cop    &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
-auxiliary    &    VP    &    aux    &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
-noun    &    genitive    &    nmod    &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
-noun    &    relative clause    &    acl    &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
-complementizer    &    S    &    lifted\_mark    &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
-verb    &    PP    &    obl    &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
-want    &    VP    &    xcomp    &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
-verb    &    subject    &    nsubj    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
-verb    &    manner adverb    &    advmod    &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
+adposition    &    NP       &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
+copula    &    NP        &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
+auxiliary    &    VP       &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
+noun    &    genitive      &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
+noun    &    relative clause       &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
+complementizer    &    S        &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
+verb    &    PP     &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
+want    &    VP      &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
+verb    &    subject    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
+verb    &    manner adverb     &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
  \hline
@@
 In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
-This dependency corresponds to dependencies of the form $\xrightarrow{obj}$ in UD.
-Similarly, most of the other dependencies in Dryer's formulation can be formalized:
-for instance, the dependency between nouns and relative clauses can be formalized as $\xrightarrow{acl}$.
-Testing whether an ordering model satisfies the correlation between relative clauses and objects reduces to checking whether these two relation types have direction parameters $a_\tau$ that are either both $>0.5$ (both kinds of dependents precede their heads) or both $< 0.5$ (they both follow their heads).
+Most of them can be straightforwardly implemented in UD, allowing us to check which correlations a word order grammar satisfies.
+%This dependency corresponds to dependencies of the form $\xrightarrow{obj}$ in UD.
+%Similarly, most of the other dependencies in Dryer's formulation can be formalized:
+%for instance, the dependency between nouns and relative clauses can be formalized as $\xrightarrow{acl}$.
+%Testing whether an ordering model satisfies the correlation between relative clauses and objects reduces to checking whether these two relation types have direction parameters $a_\tau$ that are either both $>0.5$ (both kinds of dependents precede their heads) or both $< 0.5$ (they both follow their heads).
 
@@
 See the SI for details on the neural language models and parsers used, and on the optimization procedures.
+
+\subsection{Specification of word order grammars}
+
+All natural languages have some degree of word order regularity.
+Thus, it is not sufficient to optimize the word orders of individual sentences in the corpora---instead, we will optimize the word order rules of entire languages.
+
+Our goal is to specify word order grammars that transduce unordered dependency trees into strings of words.
+We have three additional goals: first, we would like our grammars to be able to operate over dependency trees as specified in large human-annotated treebank corpora, such as those developed by the Universal Dependencies project \cite{nivre2017universal}.
+Second, we would like our grammars to be simple and easily interpretable, so we can determine whether they exhibit word order universals.
+Third, we would like our grammars to be differentiable, so that it is easy to optimize their parameters for efficiency.
+
+Our model linearizes dependency trees such as the ones in Figure~\ref{fig:arabic-japanese}.
+We assume nodes are labeled with part-of-speech (POS) tags, and arcs are labeled with syntactic relations.
+The parameters of a word order grammar are as follows.
+For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
+Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
+Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
+184) without replacement.
+
+This simple parameterization lets us directly test whether a given grammar follows word order universals.
+For example, we can test whether a Greenbergian word order correlation holds in a grammar by checking whether $a_\tau>\frac{1}{2}$ for different syntactic relations $\tau$.
+
 }

### 2019-01-25 second-draft/pnas-word-order-si.tex
attempting revision

--- 2019-01-19 writeup/pnas-word-order-si.tex
@@
 \maketitle
+
+\section{UD Formalization}
+
+\begin{table*}
+	\begin{center}
+\begin{tabular}{|ll|l|l|l|ll|l|}
+	\hline
+	%&
+	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       &     \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\
+	verb & object &      & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%adp. &
+adposition    &    NP    &    lifted\_case    &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
+copula    &    NP    &    lifted\_cop    &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
+auxiliary    &    VP    &    aux    &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
+noun    &    genitive    &    nmod    &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
+noun    &    relative clause    &    acl    &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
+complementizer    &    S    &    lifted\_mark    &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
+verb    &    PP    &    obl    &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
+want    &    VP    &    xcomp    &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
+verb    &    subject    &    nsubj    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
+verb    &    manner adverb    &    advmod    &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+\end{tabular}
+	\end{center}
+	\caption{Greenbergian Correlations.
+Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%	For clarity, we organize the correlations by the category of the `verb patterner'.
+	For each correlation, we give our operationalization in terms of UD.
+%A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
+	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
+	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
+\end{table*}
 

### 2019-02-03 second-draft/paper-names.tex
revising

@@
 
-We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
-These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages.
-For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
+Natural languages vary a lot in the order in which they express information.
+Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
+Both sentences contain a verb meaning `wrote', a noun expressing  `letter', and a phrase translating to `to a friend'.
+In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
+While the two sentences containing words with the same meanings, the order of these words are entirely different in the two languages:
+%However, the languages differ in the order in which these words and phrases are arranged: 
+In Arabic, the verb stands at the beginning, followed by both the object and the adpositional phrase.
+In Japanese, the verb stands at the end; object and adpositional phrase precede it.
+What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
+It turns out that this is a deep and stable regularity:
+Languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world.
+However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
+In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization.
+%, over 90 \% of languages satisfy this generalization .
+
+This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
+These describe correlations between the relative positions of different types of expressions across languages.
+The example above documents that the position of the object (`letter') relative to the verb is \key{correlated} with the position of adpositional phrases (`to a friend') relative to the verb.
+Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
+The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.
+%For instance, the relative order of objects and nouns also strongly determines the positions of genitives, auxiliaries, and other
+
+These universals are supported by languages on all five continents, and they are valid across
+
+These cross-linguistic patterns call for an explanation, and the problem of explaining the word order correlations has inspired a long tradition of research.
 
@@
 These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
+
+%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
+These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages.
+For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
+
+\begin{figure}
+\centering
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   katabt \& risāla \& li \& sadīq  \\
+	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+	   wrote \& letter   \& to \& friend  \\
+   \end{deptext}
+\end{dependency}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+      tomodachi \& ni \& tegami-o \& kaita \\
+ \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+	   friend \& to \& letter  \& wrote \\
+   \end{deptext}
+\end{dependency}
+
+%\begin{dependency}[theme=simple]
+%    \begin{deptext}
+%    I \& wrote \& a \& letter \& to \& a \& friend \\
+%    \end{deptext}
+%    \depedge{2}{1}{nsubj}
+%    \depedge{2}{4}{obj}
+%    \depedge{4}{3}{ }
+%    \depedge{4}{3}{ }
+%    \depedge{2}{5}{}
+%    \depedge{5}{7}{ }
+%    \depedge{7}{6}{ }
+%\end{dependency}
+	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese-simple}
+\end{figure}
+
+\begin{table*}
+	\begin{center}
+\begin{tabular}{|ll|l|l|l|ll|l|llllllll}
+	\hline
+	%&
+	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
+	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%adp. &
+adposition    &    NP       & \\
+	to            & a friend \\
+copula    &    NP        &     \\
+	is        & a friend \\
+auxiliary    &    VP       &   \\
+	has          & written \\
+noun    &    genitive      &  \\
+TODO &  \\
+noun    &    relative clause       &  \\
+complementizer    &    S        & \\
+verb    &    PP     &    \\
+want    &    VP      &    \\
+verb    &    subject    &  \\
+verb    &    manner adverb     & \\
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+\end{tabular}
+	\end{center}
+	\caption{Greenbergian Correlations.
+Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+}\label{table:results-dryer}
+\end{table*}
 
@@
 
-\subsection{Explananda: Word Order Correlations}
-
-Working with a database of 30 languages, Greenberg \cite{greenberg1963universals} proposed 45 universals of language that he determined to be true of most or all languages.
-31 of them were concerned with word order, the remainder with morphology.
-Many of the word order universals state \emph{correlations} between patterns:
-For instance, Universals 3 and 4 state that ``\emph{Languages with dominant VSO order are always prepositional.}''  (e.g.
+%\subsection{Explananda: Word Order Correlations}
+%
+%Working with a database of 30 languages, Greenberg \cite{greenberg1963universals} proposed 45 universals of language that he determined to be true of most or all languages.
+%31 of them were concerned with word order, the remainder with morphology.
+%Many of the word order universals state \emph{correlations} between patterns:
+%For instance, Universals 3 and 4 state that ``\emph{Languages with dominant VSO order are always prepositional.}''  (e.g.
 Arabic), and ``\emph{With overwhelmingly greater than chance frequency, languages with normal SOV order are postpositional.}'' (e.g.
 Japanese).
-These are illustrated in Figure~\ref{fig:arabic-japanese}:
-Arabic has basic order verb-object (VO) and prepositions (\emph{li} `to'), while Japanese has order object-verb (OV) and postpositions (\emph{ni} `to').
+%These are illustrated in Figure~\ref{fig:arabic-japanese}:
+%Arabic has basic order verb-object (VO) and prepositions (\emph{li} `to'), while Japanese has order object-verb (OV) and postpositions (\emph{ni} `to').
 Our goal is to explain Greenberg's reported word order correlations, as updated in \cite{dryer1992greenbergian}.
@@
 
-\subsection{Specification of word order grammars}
+We think of it as the following problem:
+
+A speaker has a tree in mind, and they have to produce an utterance.
+
+%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a word order grammar.
+Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
+%\begin{equation}
+%\label{eq:exp-over-utts}
+%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim L_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
+%\end{equation}
+%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
+In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
+%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
+%
+%
 
@@
 Thus, it is not sufficient to optimize the word orders of individual sentences in the corpora---instead, we will optimize the word order rules of entire languages.
+We develop a method to create word order rules that optimize given processing measures on the resulting counterfactual corpus.
 
-Our goal is to specify word order grammars that transduce unordered dependency trees into strings of words.
-We have three additional goals: first, we would like our grammars to be able to operate over dependency trees as specified in large human-annotated treebank corpora, such as those developed by the Universal Dependencies project \cite{nivre2017universal}.
-Second, we would like our grammars to be simple and easily interpretable, so we can determine whether they exhibit word order universals.
-Third, we would like our grammars to be differentiable, so that it is easy to optimize their parameters for efficiency.
+\subsection{DLM Predicts Universals}
 
-Our model linearizes dependency trees such as the ones in Figure~\ref{fig:arabic-japanese}.
-We assume nodes are labeled with part-of-speech (POS) tags, and arcs are labeled with syntactic relations.
-The parameters of a word order grammar are as follows.
-For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
-Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
-Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
-184) without replacement.
+We first apply our method to show that DLM predicts the correlations.
 
-This simple parameterization lets us directly test whether a given grammar follows word order universals.
-For example, we can test whether a Greenbergian word order correlation holds in a grammar by checking whether $a_\tau>\frac{1}{2}$ for different syntactic relations $\tau$.
+%Dependency length has been proposed as a metric of processing difficulty on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, and minimization of dependency length is known to recover a large subset of the Greenbergian universals \cite{hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic}.
+%We demonstrate that the minimization of dependency length can be formalized in our framework, reproducing the finding that optimizing grammars for dependency length recovers several word order universals \cite{gildea2010grammars}.
+%
+%To quantify dependency length, we sum the lengths of syntactic dependencies in the sentence \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
+%\begin{equation}
+%    \label{eq:depl}
+%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+%\end{equation}
+%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
+Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
 
-\subsection{Objective functions}
-\label{ref:objective-funcs}
+\subsection{Efficiency Predicts Universals}
 
-Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a word order grammar.
-Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
-\begin{equation}
-\label{eq:exp-over-utts}
-J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim L_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
-\end{equation}
-where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
-In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
-Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
+DLM is stipulative and at most a heuristic approximation to psycholinguistic findings
+
+We now introduce a general information-theoretic notion of language efficiency.
+
+use the general speaker-listener picture
 
@@
 \end{equation*}
-where the language model parameters $\phi$ are chosen to predict the language induced by tree distribution $T$ and word order grammar $\L_\theta$ as accurately as possible (maximizing predictability):
-\begin{equation*}
-	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-\end{equation*}
+%where the language model parameters $\phi$ are chosen to predict the language induced by tree distribution $T$ and word order grammar $\L_\theta$ as accurately as possible (maximizing predictability):
+%\begin{equation*}
+%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation*}
 We implement the language model $P_\phi$ using an LSTM neural network architecture \cite{hochreiter1997long}, currently the state of the art for this task \cite{goldberg2017neural}.
-Note that the language model parameters $\phi$ and the word order grammar parameters $\theta$ are recursively dependent on each other.
+%Note that the language model parameters $\phi$ and the word order grammar parameters $\theta$ are recursively dependent on each other.
 See the SI for details on how $\theta$ and $\phi$ are jointly optimized, as well as on the neural network architecture.
@@
 \end{align}
-where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
-\begin{equation*}
-\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-\end{equation*}
+%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
+%\begin{equation*}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation*}
 We implement $P_\psi$ using a highly generic graph-based parsing architecture \cite{kiperwasser2016simple,dozat2017stanford} which recovers graph structures from word strings by solving a minimal spanning tree problem.
-Due to data sparsity, we train parsers to predict dependency trees from POS tags alone, not from full wordforms, therefore we do not capture the extent to which morphological wordforms are informative about dependency tree structure.
-
+%Due to data sparsity, we train parsers to predict dependency trees from POS tags alone, not from full wordforms, therefore we do not capture the extent to which morphological wordforms are informative about dependency tree structure.
 See the SI (appendix section 6-7) for further details on the joint optimization of $\theta$ and $\psi$ and on the neural parser architecture, and see SI (section 8) for robustness to different parsing models.
@@
 \label{eq:combined-score}
-R_{Efficiency} = R_{Pars} + \lambda R_{Pred},
+	R_{\textit{Eff}} = R_{Pars} + \lambda R_{Pred},
 \end{equation}
@@
 
-\paragraph{Dependency length}
-
-In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-Dependency length has been proposed as a metric of processing difficulty on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, and minimization of dependency length is known to recover a large subset of the Greenbergian universals \cite{hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic}.
-We demonstrate that the minimization of dependency length can be formalized in our framework, reproducing the finding that optimizing grammars for dependency length recovers several word order universals \cite{gildea2010grammars}.
-
-To quantify dependency length, we sum the lengths of syntactic dependencies in the sentence \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
-\begin{equation}
-    \label{eq:depl}
-	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
-\end{equation}
-where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
-Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
-
-\subsection{Discussion and Related Work}
-
-Our work builds on previous work on optimizing word order grammars for various objectives \cite{gildea2007optimizing,gildea2010grammars,gildea2015human}.
-Previous work has used deterministic word order grammars optimized by simple hill-climbing. 
-Our work departs from previous work in three ways. 
-First, we use word order grammars that are probabilistic and differentiable, enabling greater expressivity and optimization by stochastic gradient descent.
-Second, we use state-of-the-art neural language models for predictability, jointly optimizing parameters of word order grammars and language models.
-Third, we implement parseability and efficiency as a novel objective functions, which would not be feasible with non-differentiable word order grammars.
-
-In the space of word order grammars, more complex models than ours have been defined \cite{belz2011first,futrell2015experiments,wang2016galactic}.
-However, the simple parametric form of our model enables easy interpretation across languages.
-Like these previous models, our model makes simplifying assumptions and will not fit the rules of natural languages perfectly.
-For instance, none of the models accounts for word order phenomena that are conditioned on the larger context---e.g., differences in word order between embedded and main clauses. 
-Also like previous models, our model only generates word orders that are \key{projective}, meaning that dependency lines do not cross.
-There are two reasons for constraint.
-First, there is currently no easily interpretable probabilistic model of non-projective word ordering to build on from the natural language processing literature (cf.
-\cite{bohnet2012generating}).
-Second, natural language dependency trees are overwhelmingly projective \cite{kuhlmann2013mildly}.
-Note that the preponderance of projectivity in natural language is itself a linguistic universal, equivalent to context-freeness in formal language theory \cite{marcus1965notion,kuhlmann2013mildly}, and explanations have been offered for it in terms of dependency length minimization  \cite{ferrericancho2006syntactic}.
-
-We test efficiency metrics based both on information theory (Eq.~\ref{eq:efficiency}) and dependency length minimization (Eq.~\ref{eq:depl}).
-We do not see these as competing theories.
-Rather, we see the information-theoretic objective as a broad description of efficiency in general, and dependency length as a heuristic description of a particular component of complexity.
-There is reason to believe that dependency length minimization might emerge as a byproduct of optimization for parseability and predictability under resource constraints \cite{futrell2017memory}: in particular, short dependencies are optimal for predictability given lossily compressed representations of context \cite{futrell2017noisycontext,futrell2017generalizing}.
-In the results below, we demonstrate that optimizing for our Eq.~\ref{eq:efficiency} does indeed produce dependency length minimization as a byproduct.
-
+%In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
 \section{Results}
@@
 
-While the efficiency objective succeeds in predicting the presence of the most word order correlations, the dependency length objective seems to predict a greater prevalence for those correlations that it does predict.
+%While the efficiency objective succeeds in predicting the presence of the most word order correlations, the dependency length objective seems to predict a greater prevalence for those correlations that it does predict.
 We believe this discrepancy is a result of noise in the optimization process.
@@
 
+\subsection{Objective functions}
+\label{ref:objective-funcs}
+
+Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a word order grammar.
+Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
+\begin{equation}
+\label{eq:exp-over-utts}
+J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim L_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
+\end{equation}
+where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
+In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
+Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
+
+\paragraph{Predictability}
+Predictability represents the incremental difficulty of online language processing.
+We formalize predictability as the \key{surprisal} (negative log probability) of each word given its preceding context \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+Optimization of predictability has been proposed as an explanation for certain word order patterns by \cite{ferrericancho2017optimality}.
+
+We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
+\begin{equation*}
+R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
+\end{equation*}
+where the language model parameters $\phi$ are chosen to predict the language induced by tree distribution $T$ and word order grammar $\L_\theta$ as accurately as possible (maximizing predictability):
+\begin{equation*}
+	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+\end{equation*}
+We implement the language model $P_\phi$ using an LSTM neural network architecture \cite{hochreiter1997long}, currently the state of the art for this task \cite{goldberg2017neural}.
+Note that the language model parameters $\phi$ and the word order grammar parameters $\theta$ are recursively dependent on each other.
+See the SI for details on how $\theta$ and $\phi$ are jointly optimized, as well as on the neural network architecture.
+
+\paragraph{Parseability}
+Parseability is the extent to which the syntactic and semantic relationships among words can be recovered from the string of words. 
+%Some notion of parseability has been proposed informally as a desirable property of languages in the linguistic literature going back at least to the 19th century \cite{humboldt,sapir,jesperson}, and parseability has been offered as an underlying justification for syntactic locality as a cost function for language processing \cite{hawkins-performance-1994}.
+%At a basic level, a language can encode information about syntactic relations into fixed word order patterns or into morphological marking on individual words.
+Much linguistic work has been devoted to investigating the trade-offs of various ways of indicating syntactic relations \cite{}.
+In this work, we present what is to our knowledge the first fully implemented computational formalization of this idea.
+We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
+Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
+\begin{align}
+\nonumber
+R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
+\nonumber
+&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
+\end{align}
+where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
+\begin{equation*}
+\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+\end{equation*}
+We implement $P_\psi$ using a highly generic graph-based parsing architecture \cite{kiperwasser2016simple,dozat2017stanford} which recovers graph structures from word strings by solving a minimal spanning tree problem.
+Due to data sparsity, we train parsers to predict dependency trees from POS tags alone, not from full wordforms, therefore we do not capture the extent to which morphological wordforms are informative about dependency tree structure.
+
+See the SI (appendix section 6-7) for further details on the joint optimization of $\theta$ and $\psi$ and on the neural parser architecture, and see SI (section 8) for robustness to different parsing models.
+
+\paragraph{Efficiency}
+The two scoring functions---predictability and parseability---can be combined into an overall equation of the form:
+\begin{equation}
+\label{eq:combined-score}
+R_{Efficiency} = R_{Pars} + \lambda R_{Pred},
+\end{equation}
+with an interpolation weight $\lambda \in [0,1)$.
+In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification).
+This combined equation represents efficiency: bits of information transferred about dependency trees minus the cost of the transmission.
+In expectation over utterances (substituting Eq.~\ref{eq:combined-score} into Eq.~\ref{eq:exp-over-utts}), the overall objective function to be maximized by a word order grammar can be written in information-theoretic terms as:
+\begin{equation}
+    \label{eq:efficiency}
+    J_T(\theta) = I[L_\theta(T);T] - \lambda H[L_\theta(T)],
+\end{equation}
+where $I[L_\theta(T);T]$ is the \key{mutual information} between strings and trees, and $H[L_\theta(T)]$ is the \key{entropy} of strings \cite{cover2006elements}.
+Maximizing Eq.~\ref{eq:efficiency} is equivalent to maximizing mutual information between strings and trees subject to a constraint on the maximum entropy of strings.
+
+The two terms of Eq.~\ref{eq:efficiency} implement two opposing pressures.
+Parseability (the mutual information term) pushes to make strings more heterogeneous, so that they can indicate different tree structures unambiguously.
+Predictability (the entropy term) pushes to make strings more homogeneous, to reduce the complexity of using the language.
+The idea that natural language arises from a tension between these two forces goes back over a century \cite{gabelentz1901sprachwissenschaft}.
+Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking unordered dependency trees $T$ as the underlying meanings to be conveyed.
+Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
+
+\paragraph{Dependency length}
+
+In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+Dependency length has been proposed as a metric of processing difficulty on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, and minimization of dependency length is known to recover a large subset of the Greenbergian universals \cite{hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic}.
+We demonstrate that the minimization of dependency length can be formalized in our framework, reproducing the finding that optimizing grammars for dependency length recovers several word order universals \cite{gildea2010grammars}.
+
+To quantify dependency length, we sum the lengths of syntactic dependencies in the sentence \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
+\begin{equation}
+    \label{eq:depl}
+	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+\end{equation}
+where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
+Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
+
+\subsection{Discussion and Related Work}
+
+Our work builds on previous work on optimizing word order grammars for various objectives \cite{gildea2007optimizing,gildea2010grammars,gildea2015human}.
+Previous work has used deterministic word order grammars optimized by simple hill-climbing. 
+Our work departs from previous work in three ways. 
+First, we use word order grammars that are probabilistic and differentiable, enabling greater expressivity and optimization by stochastic gradient descent.
+Second, we use state-of-the-art neural language models for predictability, jointly optimizing parameters of word order grammars and language models.
+Third, we implement parseability and efficiency as a novel objective functions, which would not be feasible with non-differentiable word order grammars.
+
+In the space of word order grammars, more complex models than ours have been defined \cite{belz2011first,futrell2015experiments,wang2016galactic}.
+However, the simple parametric form of our model enables easy interpretation across languages.
+Like these previous models, our model makes simplifying assumptions and will not fit the rules of natural languages perfectly.
+For instance, none of the models accounts for word order phenomena that are conditioned on the larger context---e.g., differences in word order between embedded and main clauses. 
+Also like previous models, our model only generates word orders that are \key{projective}, meaning that dependency lines do not cross.
+There are two reasons for constraint.
+First, there is currently no easily interpretable probabilistic model of non-projective word ordering to build on from the natural language processing literature (cf.
+\cite{bohnet2012generating}).
+Second, natural language dependency trees are overwhelmingly projective \cite{kuhlmann2013mildly}.
+Note that the preponderance of projectivity in natural language is itself a linguistic universal, equivalent to context-freeness in formal language theory \cite{marcus1965notion,kuhlmann2013mildly}, and explanations have been offered for it in terms of dependency length minimization  \cite{ferrericancho2006syntactic}.
+
+We test efficiency metrics based both on information theory (Eq.~\ref{eq:efficiency}) and dependency length minimization (Eq.~\ref{eq:depl}).
+We do not see these as competing theories.
+Rather, we see the information-theoretic objective as a broad description of efficiency in general, and dependency length as a heuristic description of a particular component of complexity.
+There is reason to believe that dependency length minimization might emerge as a byproduct of optimization for parseability and predictability under resource constraints \cite{futrell2017memory}: in particular, short dependencies are optimal for predictability given lossily compressed representations of context \cite{futrell2017noisycontext,futrell2017generalizing}.
+In the results below, we demonstrate that optimizing for our Eq.~\ref{eq:efficiency} does indeed produce dependency length minimization as a byproduct.
+
 }

### 2019-02-05 second-draft/paper-names.tex
continue revising

@@
 In Arabic, the verb stands at the beginning, followed by both the object and the adpositional phrase.
-In Japanese, the verb stands at the end; object and adpositional phrase precede it.
-What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
-It turns out that this is a deep and stable regularity:
-Languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world.
-However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
+`To' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
+In Japanese, the verb stands at the end; object and adpositional phrase precede it; `to' is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
+%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
+%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
+%It turns out that these are deep and stable regularities:
+It turns out that this variation reflects a deep and stable regularity:
+While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
+this is strongly correlated with the occurrence of pre- or postpositions:
+Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
+%However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
+This correlation has an extremely strong empirical basis:
 In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization.
 %, over 90 \% of languages satisfy this generalization .
-
-This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
-These describe correlations between the relative positions of different types of expressions across languages.
-The example above documents that the position of the object (`letter') relative to the verb is \key{correlated} with the position of adpositional phrases (`to a friend') relative to the verb.
-Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
-The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.
-%For instance, the relative order of objects and nouns also strongly determines the positions of genitives, auxiliaries, and other
-
-These universals are supported by languages on all five continents, and they are valid across
-
-These cross-linguistic patterns call for an explanation, and the problem of explaining the word order correlations has inspired a long tradition of research.
-
-A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
-Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.
-
-Researchers working in the functional paradigm have proposed a range of criteria that languages should meet in order to enable efficient communication and processing.
-These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
-
-%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
-These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages.
-For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
 
@@
 
+This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
+These describe correlations between the relative positions of different types of expressions across languages.
+The example above documents that the position of the object (`letter') relative to the verb is \key{correlated} with the position of the adposition (`to').
+%adpositional phrases (`to a friend') relative to the verb.
+Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
+The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.
+%For instance, the relative order of objects and nouns also strongly determines the positions of genitives, auxiliaries, and other
+
+%These universals are supported by languages on all five continents, cutting across language families.
+
+%These cross-linguistic patterns call for an explanation, and 
+Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
+%and the problem of explaining the word order correlations has inspired a long tradition of research.
+A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
+Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.
+
+Researchers working in the functional paradigm have proposed a range of criteria that languages should meet in order to enable efficient communication and processing.
+These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
+
+We formalize efficiency of language as a tradeoff between minimizing the
+
+We go back to early functional work.
+\cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
+According to Zipf, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
+Conversely, to minimize listener effort, the language should provide different utterances for different meanings, so that the listeenr can unambiguously identify the meaning from the utterance.
+Minimizing speaker and listener represent two opposing forces:
+A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
+A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
+The idea that language results from the tension between these two pressures has a long and fruitful history.
+\cite{horn-toward-1984}, RSA, Color naming
+
+This idea has been shown to account for phenomena such as pragmatic inference (CITE) and color naming (CITE).
+
+TODO somewhere start calling this Efficiency
+
+TODO explain why surprisal. some production/speaker work. (plan reuse)
+
+%In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
+%Even earlier than Zipf, \cite{gabelentz1901sprachwissenschaft} ...
+%This theory blabla Horn RSA
+%Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
+
+In this work, we show computationally that efficiency optimization accounts for the Greenbergian word order correlations.
+That is, we show that word orders in natural language
+
+%\section{DLM}
+
+%
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=0.45\textwidth]{figures/letter-options.png}
+%	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure.
+In this example, Grammar 1 achieves the lowest overall dependency length.
+Indeed, this is the }
+%	\label{fig:pareto-plane}
+%\end{figure}
+%
+%
+
 \begin{table*}
@@
 %adp. &
-adposition    &    NP       & \\
-	to            & a friend \\
+	adposition    &    NP       & 	\emph{to}            & \emph{a friend} \\
+
 copula    &    NP        &     \\
-	is        & a friend \\
+	\emph{	is}        & \emph{a friend} \\
 auxiliary    &    VP       &   \\
@@
 noun    &    genitive      &  \\
-TODO &  \\
+friend &  of John \\
 noun    &    relative clause       &  \\
+	books & that you read \\
 complementizer    &    S        & \\
+	that & Mary \\
 verb    &    PP     &    \\
+	went & to school \\
 want    &    VP      &    \\
+	wants   & to to leave \\
 verb    &    subject    &  \\
+	(there) entered & a tall man \\
 verb    &    manner adverb     & \\
+	ran & quickly\\
  \hline
@@
 Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+For each correlation, we provide an example.
+Given the statistical nature of the correlations, not every language satisfies every one of them.
+Not all correlations are satisfied by every natu
+}\label{table:corr-dryer}
+\end{table*}
+
+\section{Formalizing Efficiency in Word Order}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/communication.png}
+	\caption{Our model of communication: A speaker (left) expresses a dependency structure into a string of words forming a sentence.
+A listener probabilistically recovers a dependency structure.
+In this example, the grammar of English allows the listener to recover the correct dependency structure with very high confidence.}
+	\label{fig:comm}
+\end{figure*}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figures/grammars.png}
+	\caption{Given a set of dependency structures, grammars define a set of sentences. For simplicity, we use English words.  }
+	\label{fig:pareto-plane}
+\end{figure}
+
+We model the process where a speaker communicates an utterance to a listener (Figure~\ref{fig:comm}).
+
+We model the meaning that the speaker communicates via the predicate-argument structure.
+
+We represent these in the format of Dependency Grammar:
+This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
+
+By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.
+
+We can view this as a formalization of the Language of Thought (CITE) TODO explain why
+
+When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
+
+We are interested in the following hypothesis:
+If speakers choose to order their sentences in such a way that efficiency is optimized, will this lead to the observed language universals?
+
+Do languages optimize their word order for efficiency?
+Does this optimization explain the observed language universals?
+
+All natural languages have some degree of word order regularities that are specified in their grammar.
+It is thus not realistic to assume that speakers optimize each utterance individually.
+Instead, we hypothesize that the word order rules of languages as a whole are optimized for efficiency, and that this optimization produces the observed universals.
+
+We formalize a language as a distribution over dependency structures $T$ together with an ordering grammar $\theta$ that specifies how dependency structures are ordered into strings of words.
+Applying the ordering grammar to the dependency structures, a language defines a probability distribution over sentences $L$.
+Efficiency is defined as a tradeoff between
+
+The \key{predictability} of the language is the negative entropy of the sentences:
+\begin{equation}
+	R_{Pred} := H[L] = \sum_{l} p(l) \log p(l)
+\end{equation}
+where the sum runs over all possible sentences $l$ that belong to the language.
+This quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
+%how many distinct sentences there are.
+%It is larger
+
+The opposing pressure is \key{parsability}, which is the mutual information beteen sentences and dependency structures:
+\begin{equation}
+	R_{Pars} := I[L,T] = \sum_{t,l} p(t,l) \log \frac{p(t|l)}{p(t)}
+\end{equation}
+where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
+This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
+It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
+
+The \key{efficiency} of a language is a weighted combination
+\begin{equation}
+	R_{Eff} := R_{Pars} + \lambda R_{Pred}
+\end{equation}
+with an interpolation weight $\lambda \in [0,1)$. In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification).
+
+%This combined equation represents efficiency: bits of information transferred about dependency trees minus the cost of the transmission.
+In expectation over utterances (substituting Eq.~\ref{eq:combined-score} into Eq.~\ref{eq:exp-over-utts}), the overall objective function to be maximized by a word order grammar can be written in information-theoretic terms as:
+
+%The two terms of Eq.~\ref{eq:efficiency} implement two opposing pressures.
+Parseability (the mutual information term) pushes to make strings more heterogeneous, so that they can indicate different tree structures unambiguously.
+Predictability (the entropy term) pushes to make strings more homogeneous, to reduce the complexity of using the language.
+%The idea that natural language arises from a tension between these two forces goes back over a century \cite{gabelentz1901sprachwissenschaft}.
+Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking unordered dependency trees $T$ as the underlying meanings to be conveyed.
+Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
+
+We create
+
+Fortunately, we have exacly the kind of data we need
+
+For the dependency structures, we take real tree structures from corpora of 51 languages.
+These corpora represent a typologically and genetically diverse group of languages.
+
+For each of these corpora, we construct grammars that optimize efficiency of the linearized sentences.
+(The original word orders of the languages do not enter.)
+
+See SI for our method for creating optimized grammars.
+
+%\subsection{Dependency Length Minimization}
+
+\section{Results}
+
+\subsection{Relative efficiency of languages}
+\label{sec:relative-efficiency}
+
+We first demonstrate that real languages are relatively efficient compared to random baselines.
+For each language, we generated ten baseline word order grammars for the language by choosing all word order grammar parameters randomly at uniform from $[0,1]$.
+Figure~\ref{fig:pareto-plane} shows the predictability and parseability for each real language relative to its baseline grammars.
+In order to control for limitations due to our word order grammar formalism, we represent real languages in the figure by maximum likelihood fits of word order grammars to the real language data.
+%; in general these maximum-likelihood word order grammars end up with lower predictability and parseability than the real languages. % TODO verify this is true? TODO awkward sentence
+For the calculation of predictability and parseability, we make all (baseline and real) word order grammars deterministic by always choosing the highest-probability linearization of each tree; by making the grammars deterministic in this way we eliminate an anticonservative bias toward low predictability in the baseline languages, which are highly nondeterministic.
+The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend.pdf}
+    \caption{Predictability and parseability of 51 UD languages (red), indicated by ISO codes, compared to ten baseline word order grammars per language (green).
+Predictability and parseability scores are $z$-scored within language.
+Each point for a real language has a line pointing in the direction of the center of mass of its baselines.
+The green contour shows the density of baseline languages.
+Unlabeled dots represent the centroid for real languages (red), baseline languages (green), and languages optimized for predictability (yellow), parseability (pink), efficiency (blue), and dependency length (red).
+When a language is to the bottom--left of its baselines, this indicates that it is relatively optimal for efficiency.} % TODO colors, fontsize, etc could conceivably be made prettier
+    \label{fig:pareto-plane}
+\end{figure}
+
+Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
+Languages appear to be attracted toward these points and away from the region of the baseline languages.
+We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
+% TODO add something about dependency length-optimized languages?
+
+%These results are shown in more detail for two typologically distinct languages (English and Japanese) in Figure~\ref{fig:predict-pars}.
+Here we include also the predictability and parseability of the original language data and of the word order grammars optimized for dependency length.
+We see that real languages are even more predictable than any optimized or baseline language, while optimization for either parseability or dependency length suffices to raise parseability to the level of the real languages.
+This result is in line with the finding that languages with longer dependencies are harder to parse in general \cite{gulordava2016multilingual}.
+
+%\begin{figure*}
+%\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
+%	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right).
+We provide the entropy $H[L_\theta(T)]$ for predictability (lower is better) and the conditional entropy $H[T|L_\theta(T)]$ for parseability (lower is better).
+Both measures are normalized by sentence length.}\label{fig:predict-pars}
+%\end{figure*}
+%
+
+\subsection{Efficiency Explains DLM}
+Finally we demonstrate the relationship between dependency length minimization and the maximization of efficiency.
+Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency.
+We see that optimizing for efficiency lowers dependency length relative to random baselines, in keeping with the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}.
+In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the depenndency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.
+
+\begin{figure*} % need Efficiency and (maybe) MLE langs on here
+\includegraphics[width=0.98\textwidth]{../results/dependency-length/figures/depLength-facet.pdf}
+\caption{Average dependency length as a function of sentence length in four languages.
+Across languages, real and optimized languages have shorter dependencies than random baseline orderings.}\label{fig:deplength}
+\end{figure*}
+
+\subsection{Greenbergian Word Order Correlations}
+\label{sec:greenberg}
+\begin{table*}
+	\begin{center}
+\begin{tabular}{|ll|l|l|l|ll|l|}
+	\hline
+	%&
+	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
+	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%adp. &
+adposition    &    NP       &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
+copula    &    NP        &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
+auxiliary    &    VP       &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
+noun    &    genitive      &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
+noun    &    relative clause       &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
+complementizer    &    S        &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
+verb    &    PP     &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
+want    &    VP      &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
+verb    &    subject    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
+verb    &    manner adverb     &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+\end{tabular}
+	\end{center}
+	\caption{Greenbergian Correlations.
+Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%	For clarity, we organize the correlations by the category of the `verb patterner'.
+	For each correlation, we give our operationalization in terms of UD.
+%A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
+	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
+	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
+\end{table*}
+
+%\begin{table*}
+%\begin{tabular}{lllllllllllllllllll}
+%	&	          & \multicolumn{2}{c}{Dep. Length}      & \multicolumn{2}{c}{Lang. Mod.}              &  \multicolumn{2}{c}{Parsing}  &\multicolumn{2}{c}{Operationalization} & & GU\\ \hline
+%	Adp &	Adp+Noun &                  & \cmark       &             &            &         &             & \textsc{noun} $\xrightarrow{case}$ \textsc{adp} & - &  3, 4\\ \hline
+%	Verb &	Verb+PP & \checkmark        & \cmark      & \checkmark  &   \cmark         &    &             & \textsc{verb} $\xrightarrow{obl}$ \textsc{noun}  & + & 7  \\
+%	Verb &	Copula & \checkmark         &  \cmark     &  \checkmark &   \cmark    &  \checkmark  &     \cmark        & \textsc{noun} $\xrightarrow{cop}$  \textsc{aux} & - \\
+%	Verb &	want & \checkmark           & \cmark      & \checkmark  & \cmark     &  \checkmark  &    \cmark         & \textsc{verb} $\xrightarrow{xcomp}$ \textsc{verb} & + & 15 \\
+%	Verb &	subject &                   &       & \checkmark  &      \cmark      & \xmark   &    \xmark         & \textsc{verb} $\xrightarrow{nsubj}$ \textsc{noun} & + \\
+%	Verb &	auxiliary & \cmark          &       & \cmark      & \cmark     &           &             & \textsc{verb}  $\xrightarrow{aux}$ \textsc{verb}/\textsc{aux}  & - & 16\\
+%	verb & manner adverb &              &  -     &  \xmark     &   \xmark     &          &             & \textsc{verb} $\xrightarrow{advmod}$ ADV  & + & 7 \\ \hline
+%	Comp &	comp; adv sub. &            &   \cmark    &             &   \cmark         &          &             & \textsc{verb} $\xrightarrow{mark}$ \textsc{sconj} & - & \\ \hline
+%	Noun &	article N' &                &   \xmark/\xmark    &             &    -/\cmark        &          &   -/\cmark          & \textsc{noun} $\xrightarrow{det}$ \textsc{det}/\textsc{pron}  & -\\
+%	Noun &	noun+genitive & \cmark      &   \cmark    &             &            &              &    \cmark  & \textsc{noun} $\xrightarrow{nmod}$ \textsc{noun}  & + & 2\\
+%	Noun &	noun+relative clse & \cmark &  \cmark   &               &            &  \checkmark &      \cmark       & \textsc{noun} $\xrightarrow{acl}$ \textsc{verb} & + & 24 \\ \hline
+%	Adj &	adj+std            &        &  \cmark   &               &            &             &             &  ADJ $\xrightarrow{nmod}$ \textsc{adp}  & + & 23\\
+%\end{tabular}
+%	\caption{Greenbergian Correlations, as provided by \cite{dryer-greenbergian-1992}.
+For each correlation, we give our operationalization in terms of UD and whether the direction of this dependency correlates positively ($+$) or negatively ($-$) with the verb-object-dependency.
+For each correlation and each objective function, we indicate whether the correlation is predicted correctly (\cmark), in the wrong direction (\xmark), or no prediction is made (blank).
+For each objective function, we show results for plain UD (left column) and UD with function word heads (right column).
 }\label{table:results-dryer}
-\end{table*}
+%\end{table*}
+%%, with corresponding numbers in the original Greenberg universals, where applicable.
+
+%In the last sections, we confirmed the effectiveness of our optimization method and elucidated the relation between the three optimization functions on the level of .
+
+We now examine to what extent we can recover Greenberg's word order correlations in optimized grammars. 
+Dryer \cite{dryer1992greenbergian} presents a comprehensive updated version of the word order correlations, drawing on 625 languages, which we take as the basis of our evaluation.
+In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
+Most of them can be straightforwardly implemented in UD, allowing us to check which correlations a word order grammar satisfies.
+%This dependency corresponds to dependencies of the form $\xrightarrow{obj}$ in UD.
+%Similarly, most of the other dependencies in Dryer's formulation can be formalized:
+%for instance, the dependency between nouns and relative clauses can be formalized as $\xrightarrow{acl}$.
+%Testing whether an ordering model satisfies the correlation between relative clauses and objects reduces to checking whether these two relation types have direction parameters $a_\tau$ that are either both $>0.5$ (both kinds of dependents precede their heads) or both $< 0.5$ (they both follow their heads).
+
+Dryer \cite{dryer1992greenbergian} presents three correlations which do not correspond to dependencies annotated in UD: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
+Two pairs of Dryer's correlations, namely those for the dependencies between complementizers and adverbial subordinators and their complement clauses,and those for the dependencies between verbs and adpositional phrases, and adjectives and their standard of comparison, had to be collapsed into two correlations in UD.
+From Dryer's 15 correlations, we obtained 10 formalized correlations, roughly covering nine of Greenberg's original universals.
+%TODO update to 10?
+What happened to the 11th?
+Was that amod?
+[The 11th was adjective+standard of comparison, which is merged into obl] Maybe we could include amod as a "true rejection" of efficiency, but a false alarm for DepL
+
+In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
+We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau$ show the same direction for the correlating dependency and for the verb-object dependency, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
+%We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau > 0.5$ holds for the correlating dependency depending on whether $a_\tau > 0.5$ for the verb-object dependency, with per-language-family random intercept and slope.\footnote{We coded language families according to \url{universaldependencies.org}.}
+We are interested in the direction and significance of this effect: 
+If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
+
+We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
+To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
+The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
+%in all except four languages, for WALS did not have data.
+
+\paragraph{Results}
+Results are shown in Table~\ref{table:results-dryer}.
+All correlations but two are confirmed in the models estimated from the real orderings.
+The exceptions are the subject--verb dependency and the verb--adverb dependency, which typically go in the opposite direction from the standard description.
+We will discuss these exceptions further below.
+
+In keeping with previous work, we see that optimizing for dependency length correctly accounts for nine word order correlations, missing only the verb--adverb dependency. 
+Predictability and parseability predict five and seven correlations, respectively, making largely complementary predictions.
+Efficiency significantly predicts all the word order correlations, each in the same direction as attested in the dependency corpora.
+
+We now address the two word order correlations whose direction in the dependency corpora is opposite from what would be expected in the typological literature.
+The first is the correlation of the order of verb--subject and verb--object dependencies.
+Our sample of mainly European languages highly over-represents languages with the general order subject--verb--object (such as English), in which the order of the verb--subject and verb--object dependencies are anti-correlated.
+Surprisingly, given the sample of tree structures of these languages, it turns out that the optimal languages tend to have anti-correlated orders for subjects and objects order similar to the real languages.
+
+The second anomalous dependency is the verb--manner adverb dependency.
+We believe the anti-correlation in the UD corpora arises because the \emph{advmod} dependency does not distinguish between manner adverbs---the subject of the typological judgment---and various other types of modifiers such as sentence-level adverbs.
+Nevertheless, the languages optimized for efficiency reproduce the anti-correlation of the orders of verb--object and verb--adverb at around the same rate as the real languages.
+
+We further evaluate the word order predictions of efficiency, showing that efficiency is most successful in predicting correlations in the direction found in the UD corpora.
+We constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency in languages optimized for efficiency, with random effects for language and language family, correlated across the ten dependencies.
+We conducted the same analysis for predictability, parseability, and dependency length.
+We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
+The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
+The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$.
+The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.
+For dependency length, the posterior puts much of the probability mass on predicting only nine of the correlations; predictability and parseability predict significantly less correlations.
+
+\begin{figure}
+    \centering
+%     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-efficiency.pdf}
+ %    \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-langmod.pdf}
+  %   \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}
+   %  \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-depl.pdf}
+     \includegraphics[scale=.45]{../results/correlations/figures/posterior-satisfied-universals-together.pdf}
+
+	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks, computed from a mixed-effects logistic regression jointly modeling all ten dependencies.
+Efficiency predicts all ten correlations with high posterior probability.}
+    \label{fig:posterior-satisfied-count}
+\end{figure}
+
+\section{DLM}
+
+We compare Efficiency to the predictions of Dependency Length Minimization.
+This theory states that natural languages order information in such a way that the distances between syntactically linked words are minimized.
+For instance, in Figure~\ref{fig:arabic-japanese-simple}, there would be syntactic links between the adposition (to) and the noun (friend), and between the verb and both the object (letter) and the adposition (to).
+There is strong evidence that natural language minimizes the length of these dependencies~\cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
+Theoretical work in the functional linguistics literature has proposed that this principle explains the correlations.
+However, the principle itself is stipulative and not a first-principles explanation.
+\cite{hawkins1994performance} theoetically argues that it increases  parsability.
+
+However, these arguments have been made on a theoretical basis.
+We will show that (1) Dependency Length Minimization indeeds predicts most of the correlations, and (2) Dependency Length Minimization is explained by efficiency optimization.
+
+These computational results confirm theoretical ideas  that have been stated informally by authors in the functional linguistics literature at least since the 1980s.
+
+%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
+These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages.
+For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
+
+\showmatmethods{} % Display the Materials and Methods section
+
+\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Chris Manning, and audiences at CAMP 2018 for helpful discussion.}
+
+\showacknow{} % Display the acknowledgments section
+
+%\bibliographystyle{acl_natbib}
+%\bibliography{extracted}
+\bibliography{everything}
+
+\end{document}
+
+%the meaning
+
+\begin{figure}
+\centering
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   katabt \& risāla \& li \& sadīq  \\
+	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+	   wrote \& letter   \& to \& friend  \\
+   \end{deptext}
+\end{dependency}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+      tomodachi \& ni \& tegami-o \& kaita \\
+ \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+	   friend \& to \& letter  \& wrote \\
+   \end{deptext}
+\end{dependency}
+
+%\begin{dependency}[theme=simple]
+%    \begin{deptext}
+%    I \& wrote \& a \& letter \& to \& a \& friend \\
+%    \end{deptext}
+%    \depedge{2}{1}{nsubj}
+%    \depedge{2}{4}{obj}
+%    \depedge{4}{3}{ }
+%    \depedge{4}{3}{ }
+%    \depedge{2}{5}{}
+%    \depedge{5}{7}{ }
+%    \depedge{7}{6}{ }
+%\end{dependency}
+	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese-simple}
+\end{figure}
+
+\section{Minimizing Dependency Length Predicts Universals}
+
+..
+
+\begin{figure}
+\centering
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   katabt \& risāla \& li \& sadīq  \\
+	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+	   wrote \& letter   \& to \& friend  \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{1}{2}{obj}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{1}{3}{}
+   \depedge{3}{4}{}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+      tomodachi \& ni \& tegami-o \& kaita \\
+ \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+	   friend \& to \& letter  \& wrote \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{4}{2}{}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{2}{1}{}
+   \depedge{4}{3}{obj}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+\end{figure}
+
+Dependency Length:
+
+The sentences in Figure~\ref{fig:arabic-japanese} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words.
+For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} if the verb, rather than e.g.
+the subject or an indirect object.
+A dependency arc is drawn from a \key{head} (e.g.
+\emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g.
+\emph{tegami-o}).
+Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}.
+Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.
+
+In order to interpret a sentence correctly, a listener must be able to reconstruct the information expressed in the dependency tree and its labels.
+This is because, according to the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.
 
@@
 \subsection{Dependency trees}
-
-The sentences in Figure~\ref{fig:arabic-japanese} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words.
-For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} if the verb, rather than e.g.
-the subject or an indirect object.
-A dependency arc is drawn from a \key{head} (e.g.
-\emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g.
-\emph{tegami-o}).
-Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}.
-Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.
-
-In order to interpret a sentence correctly, a listener must be able to reconstruct the information expressed in the dependency tree and its labels.
-This is because, according to the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.
 

### 2019-02-06 second-draft/paper-names.tex
continuing revision

@@
 %and the problem of explaining the word order correlations has inspired a long tradition of research.
+
+%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
+
 A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
@@
 
-We go back to early functional work.
 \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
@@
 
-In this work, we show computationally that efficiency optimization accounts for the Greenbergian word order correlations.
-That is, we show that word orders in natural language
+In this work, we show computationally, using corpus data from 51 languages, that efficiency optimization accounts for the Greenbergian word order correlations.
+That is, we show that the word orders in natural languages have evolved to optimize efficiency, and that this optimization accounts for the prevalence of the word order correlations.
+%word orders in natural language
 
@@
 
-\begin{figure*}
+\begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{figures/communication.png}
+    \includegraphics[width=0.45\textwidth]{figures/communication.png}
 	\caption{Our model of communication: A speaker (left) expresses a dependency structure into a string of words forming a sentence.
@@
 	\label{fig:comm}
-\end{figure*}
+\end{figure}
 
@@
     \includegraphics[width=0.45\textwidth]{figures/grammars.png}
-	\caption{Given a set of dependency structures, grammars define a set of sentences. For simplicity, we use English words.  }
-	\label{fig:pareto-plane}
+	\caption{Given a set of dependency structures, grammars define a set of sentences.
+For simplicity, we use English words.
+Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb.
+}
+	\label{fig:grammars}
 \end{figure}
@@
 We model the process where a speaker communicates an utterance to a listener (Figure~\ref{fig:comm}).
-
-We model the meaning that the speaker communicates via the predicate-argument structure.
-
+TODO Shannon
+We model the meaning that the speaker communicates as the basic predicate-argument structure.
 We represent these in the format of Dependency Grammar:
@@
 By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.
+Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
 
@@
 When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
-
-We are interested in the following hypothesis:
-If speakers choose to order their sentences in such a way that efficiency is optimized, will this lead to the observed language universals?
+The listener receives a string of words, and attempts to reconstruct the underlying syntactic relations.
+%We formalize the listener's understanding as probabilistic inference, that is, the listener 
+If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).
+
+Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
+While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
+This is illustrated in Figure~\ref{fig:grammars}: Grammars specify how dependency structures are encoded into strings of words.
+For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
+Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
+Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
+In a speech community, speakers and listeners have shared knowledge of these rules, which allows the listener to decode the dependency tree.
+In the terminology of information theory, these rules specify a \emph{channel} through which the speaker transmits the dependency structures.
+
+Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance (TODO can we find an example for this?)
+
+The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances provide about their underlying tree structures:
+\begin{equation}
+	R_{Pars} := I[L,T] = \sum_{t,l} p(t,l) \log \frac{p(t|l)}{p(t)}
+\end{equation}
+where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
+This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
+It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
+This quantity formalizes Zipf's principle of listener effort and his Force of Diversification:
+It can be maximized when trees can be decoded fully unambiguously from utterances.
+This formalization is a standard formalization of Zipf's principle of listener economy (CITE RSA, Regier, ...).
+
+The counteracting Force of Unification is formalized by the cost of encoding the dependency structure and producing the utterance.
+TODO motivate surprisal here.
+
+%Note that, in order for this process to be successful, speakers and listeners need to have shared knowledge of the rules by which dependency structures are encoded into sentences.
+%We hypothesize that the rules found in actual natural languages are optimized for the efficiency of this process, and that this optimization accounts for the observed word order universals.
+
+%If speakers choose to order their sentences in such a way that efficiency is optimized, will this lead to the observed language universals?
 

### 2019-02-07 second-draft/paper-names.tex
continuing revision

@@
 
-We model the process where a speaker communicates an utterance to a listener (Figure~\ref{fig:comm}).
-TODO Shannon
-We model the meaning that the speaker communicates as the basic predicate-argument structure.
-We represent these in the format of Dependency Grammar:
+We model the process where a speaker communicates an utterance to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
+In Shannon's model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
+Applying this model to word order, we take the message to consist of a set of syntactic and semantic relations.
+Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
 This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
-
-By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.
+In Figure~\ref{fig:comm}, the message consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
+
+When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
+The listener receives this string of words.
+By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
+The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
 Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
-
-We can view this as a formalization of the Language of Thought (CITE) TODO explain why
-
-When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
-The listener receives a string of words, and attempts to reconstruct the underlying syntactic relations.
+Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
+If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).
+
+%We can view this as a formalization of the Language of Thought (CITE) TODO explain why
+
 %We formalize the listener's understanding as probabilistic inference, that is, the listener
-If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).
-
-Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
-While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
+
+All natural languages have some degree of word order regularities that are specified in their grammar.
+For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
+That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
+Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
+%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
+%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
 This is illustrated in Figure~\ref{fig:grammars}: Grammars specify how dependency structures are encoded into strings of words.
@@
 Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
-In a speech community, speakers and listeners have shared knowledge of these rules, which allows the listener to decode the dependency tree.
-In the terminology of information theory, these rules specify a \emph{channel} through which the speaker transmits the dependency structures.
-
+%In a speech community, speakers and listeners have shared knowledge of these rules, which allows the listener to decode the dependency tree.
+%In the terminology of information theory, these rules specify a \emph{channel} through which the speaker transmits the dependency structures.
+
+In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes.
+In our model, these will depend on the grammar:
 Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance (TODO can we find an example for this?)
@@
 
-All natural languages have some degree of word order regularities that are specified in their grammar.
-It is thus not realistic to assume that speakers optimize each utterance individually.
-Instead, we hypothesize that the word order rules of languages as a whole are optimized for efficiency, and that this optimization produces the observed universals.
-
-We formalize a language as a distribution over dependency structures $T$ together with an ordering grammar $\theta$ that specifies how dependency structures are ordered into strings of words.
-Applying the ordering grammar to the dependency structures, a language defines a probability distribution over sentences $L$.
-Efficiency is defined as a tradeoff between
+%We formalize a language as a distribution over dependency structures $T$ together with an ordering grammar $\theta$ that specifies how dependency structures are ordered into strings of words.
+%Applying the ordering grammar to the dependency structures, a language defines a probability distribution over sentences $L$.
+%Efficiency is defined as a tradeoff between
 
@@
 
-We create
-
-Fortunately, we have exacly the kind of data we need
-
-For the dependency structures, we take real tree structures from corpora of 51 languages.
+\subsection{Optimizing Grammars for Efficiency}
+
+We now ask:
+(1) Do the grammars of natural languages evolve towards optimizing efficiency of communication?
+(2) Does this process of optimization account for Greenberg's order correlation universals?
+
+%For answering these questions,
+
+Answering these questions requires a sample of dependency structures as actually used by speakers of difefrent languages.
+Such samples have recently become available with the Universal Dependencies project, which has collected and created dependency annotations for several dozens of different languages.
+We use data from 51 languages.
 These corpora represent a typologically and genetically diverse group of languages.
 
-For each of these corpora, we construct grammars that optimize efficiency of the linearized sentences.
-(The original word orders of the languages do not enter.)
-
+To answer whether languages evolve to have efficient word order, we compare the efficiency of the actual grammars of these 51 languages to randomly constructed baseline grammars.
+To show that this process of efficiency optimization accounts for Greenberg's correlation universals, we computationally construct grammars that optimize efficiency.
+We then show that these grammars mostly exhibit the Greenbergian order correlation universals.
+We furthermore show that this process of optimization also \emph{explains} DLM, providing a first-principles explanation of this heuristic generalization.
+
+For every one of the 51 corpora, we computationally construct eight optimal grammars.
+Note that the original orders of the actual languages do not enter this objective.
 See SI for our method for creating optimized grammars.
+
+%
+%All natural languages have some degree of word order regularities that are specified in their grammar.
+%It is thus not realistic to assume that speakers optimize each utterance individually.
+%Instead, we hypothesize that the word order rules of languages as a whole are optimized for efficiency, and that this optimization produces the observed universals.
+%
 
@@
 
-\section{DLM}
+\section{DLM TODO put these at the appropriate place}
 

### 2019-02-09 second-draft/2-paper-names.tex
revising paper

--- 2019-02-07 second-draft/paper-names.tex
@@
 An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
+
 In this work we examine this question from a computational perspective, demonstrating a fully formalized framework in which certain syntactic universals can be explained through the statistical optimization of grammars for information-theoretic efficiency.
@@
 
+We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
+These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages.
+For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
+
+A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
+Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.
+
+Researchers working in the functional paradigm have proposed a range of criteria which languages should meet in order to enable efficient communication and processing.
+These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
+
+The contribution of this work is (i) to provide a principled formalization of the notion of communicative efficiency in word order, based on related work in information theory \cite{cover2006elements,tishby1999information}, bounded rationality \cite{genewein2015bounded}, and computational neuroscience \cite{friston2009predictive}; (ii) to introduce a computational framework in which we can produce simulated grammars directly optimized for our notion of efficiency, using state-of-the-art machine learning techniques \cite{hochreiter1997long,goldberg2017neural,dozat2017stanford}; and (iii) to demonstrate that in doing so, we reproduce the major word order patterns reported in \cite{greenberg1963universals} and related work \cite{dryer1992greenbergian}, as well as reproducing the predictions of previous heuristic theories of word order optimization \cite{hawkins1994performance,ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
+
+We formalize the notion of communicative efficiency using two terms: one having to do with minimizing the difficulty of linguistic communication, and one having to do with maximizing the information transferred in such communication.
+We quantify the difficulty of linguistic communication in terms of the word-by-word \key{predictability} of strings, which has been found to be a highly accurate and general predictor of human online comprehension difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and which can be justified in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+We quantify the information transferred in language as \key{parseability}: the quantity of information contained in linguistic utterances about latent hierarchical tree structures.
+We also compare the results of optimizing for communicative efficiency to the results of optimizing for dependency length minimization, a previously proposed approximate metric of the amount of memory involved in language production and comprehension \cite{hudson1995measuring,gibson1998linguistic,gibson2000dependency,liu2017dependency,temperley2018minimizing}.
+
+%To evaluate whether communicative efficiency can explain word order universals, we apply the following method \cite{gildea2007optimizing}.
+First, we develop word order grammars: simple, interpretable models of word order, implemented as functions that take in an unordered syntactic dependency tree and output a distribution over possible word orders for that tree.
+Second, we take dependency treebanks of natural languages and compute word order grammars which are optimized for an objective functions representing communicative efficiency.
+%Third, we evaluate whether the optimized word order grammars reproduce the desired word order universals when applied to attested dependency trees from the treebanks across many languages.
+Our favorable results suggest that it may be possible to generally model human languages as efficient codes for communication among agents with generic human-like cognitive and communicative constraints.
+%The best way to describe a human language might be as a solution to an equation involving such constraints.
+
+\section{Greenberg's Word Order Correlations}
 Natural languages vary a lot in the order in which they express information.
@@
 
+TODO maybe being more general in the beginning, as in the first version. make sure the reader doesn't think 'it's just headedness, why bother'
+
 \begin{figure}
@@
 %These cross-linguistic patterns call for an explanation, and 
-Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
 %and the problem of explaining the word order correlations has inspired a long tradition of research.
-
-%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-
+Supported by languages on all continents, these correlations are among the language universals with the strongest empirical support.
+Importantly, their validity is also independent from specific assumptions about theories of grammar.
+
+Consequently, explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
+%
+%%supported by languages on all continents, and their validity is independent from 
+%
+%TODO say more about why these are extremely important, and that it is a puzzle why they are true
+%
+%%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
+%
 A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
@@
 
-We formalize efficiency of language as a tradeoff between minimizing the
-
-\cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
-According to Zipf, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
+Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
+According to this theory, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
 Conversely, to minimize listener effort, the language should provide different utterances for different meanings, so that the listeenr can unambiguously identify the meaning from the utterance.
-Minimizing speaker and listener represent two opposing forces:
+Minimizing speaker and listener effort represent two opposing forces:
 A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
 A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
+%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
+Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
 The idea that language results from the tension between these two pressures has a long and fruitful history.
@@
 
-This idea has been shown to account for phenomena such as pragmatic inference (CITE) and color naming (CITE).
-
-TODO somewhere start calling this Efficiency
-
-TODO explain why surprisal. some production/speaker work. (plan reuse)
+%In the domain of word order,
+
+%TODO introduce the terms \key{cost} and \key{informativity}
+%
+This idea has been shown to account for phenomena such as pragmatic inference \cite{horn-toward-1984, frank2012predicting} and color naming~\cite{zaslavsky2018efficient}.
+
+%TODO somewhere start calling this Efficiency
+%TODO explain why surprisal. some production/speaker work. (plan reuse)
 
@@
 
-In this work, we show computationally, using corpus data from 51 languages, that efficiency optimization accounts for the Greenbergian word order correlations.
+In this work, we show that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
+
+%In this work, we show computationally, using corpus data from 51 languages, that efficiency optimization accounts for the Greenbergian word order correlations.
 That is, we show that the word orders in natural languages have evolved to optimize efficiency, and that this optimization accounts for the prevalence of the word order correlations.
@@
 %adp. &
-	adposition    &    NP       & 	\emph{to}            & \emph{a friend} \\
-
+	adposition    &    NP       \\
+	\emph{to}            & \emph{a friend} \\
 copula    &    NP        &     \\
-	\emph{	is}        & \emph{a friend} \\
+	\emph{is}        & \emph{a friend} \\
 auxiliary    &    VP       &   \\
-	has          & written \\
+	\emph{has}          & \emph{written} \\
 noun    &    genitive      &  \\
-friend &  of John \\
+	\emph{friend} &  \emph{of John} \\
 noun    &    relative clause       &  \\
-	books & that you read \\
+	\emph{books} & \emph{that you read} \\
 complementizer    &    S        & \\
-	that & Mary \\
+	\emph{that} & \emph{Mary} \\
 verb    &    PP     &    \\
-	went & to school \\
+	\emph{went} & \emph{to school} \\
 want    &    VP      &    \\
-	wants   & to to leave \\
+	\emph{wants}   &  \emph{to leave} \\
 verb    &    subject    &  \\
-	(there) entered & a tall man \\
+	\emph{(there) entered} & \emph{a person} \\
 verb    &    manner adverb     & \\
-	ran & quickly\\
+	\emph{ran} & \emph{quickly} \\
  \hline
@@
 
-We model the process where a speaker communicates an utterance to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
+We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
 In Shannon's model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
-Applying this model to word order, we take the message to consist of a set of syntactic and semantic relations.
+Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
 Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
@@
 Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance (TODO can we find an example for this?)
-
-The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances provide about their underlying tree structures:
+This reflects the \key{parseability} of the language.
+
+The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
 \begin{equation}
-	R_{Pars} := I[L,T] = \sum_{t,l} p(t,l) \log \frac{p(t|l)}{p(t)}
+	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
 \end{equation}
@@
 It can be maximized when trees can be decoded fully unambiguously from utterances.
-This formalization is a standard formalization of Zipf's principle of listener economy (CITE RSA, Regier, ...).
-
-The counteracting Force of Unification is formalized by the cost of encoding the dependency structure and producing the utterance.
-TODO motivate surprisal here.
+Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's principle of listener economy~\cite{frank2012predicting,regier2015word}.
+
+%The counteracting Force of Unification is formalized by the cost of encoding the dependency structure and producing the utterance.
+%TODO motivate surprisal here.
 
@@
 
-Do languages optimize their word order for efficiency?
-Does this optimization explain the observed language universals?
+%Do languages optimize their word order for efficiency?
+%Does this optimization explain the observed language universals?
 
@@
 
-The \key{predictability} of the language is the negative entropy of the sentences:
+% based on related work in information theory \cite{cover2006elements,tishby1999information}, bounded rationality \cite{genewein2015bounded}, and computational neuroscience \cite{friston2009predictive}; (ii)
+%
+%We formalize the notion of communicative efficiency using two terms: one having to do with minimizing the difficulty of linguistic communication, and one having to do with maximizing the information transferred in such communication.
+We quantify the difficulty of linguistic communication in terms of the word-by-word \key{predictability} of strings, which has been found to be a highly accurate and general predictor of human online comprehension difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect},
+% in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+We quantify the information transferred in language as \key{parseability}: the quantity of information linguistic utterances contain about latent hierarchical tree structures.
+We also compare the results of optimizing for communicative efficiency to the results of optimizing for dependency length minimization, a previously proposed approximate metric of the amount of memory involved in language production and comprehension \cite{hudson1995measuring,gibson1998linguistic,gibson2000dependency,liu2017dependency,temperley2018minimizing}.
+
+We formalize the cost of an utterance as its word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
+Surprisal has been found to be is a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
+%The \key{predictability} of the language is the negative entropy of the sentences:
 \begin{equation}
-	R_{Pred} := H[L] = \sum_{l} p(l) \log p(l)
+	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
 \end{equation}
 where the sum runs over all possible sentences $l$ that belong to the language.
-This quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
+In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
 %how many distinct sentences there are.
+
 %It is larger
 
-The opposing pressure is \key{parsability}, which is the mutual information beteen sentences and dependency structures:
-\begin{equation}
-	R_{Pars} := I[L,T] = \sum_{t,l} p(t,l) \log \frac{p(t|l)}{p(t)}
-\end{equation}
-where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
-This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
-It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
+%
+%The opposing pressure is \key{parsability}, which is the mutual information beteen sentences and dependency structures:
+%\begin{equation}
+%	R_{Pars} := I[L,T] = \sum_{t,l} p(t,l) \log \frac{p(t|l)}{p(t)}
+%\end{equation}
+%where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
+%This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
+%It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
 
 The \key{efficiency} of a language is a weighted combination
-\begin{equation}
+\begin{equation}\label{eq:efficiency}
 	R_{Eff} := R_{Pars} + \lambda R_{Pred}
@@
 %The idea that natural language arises from a tension between these two forces goes back over a century \cite{gabelentz1901sprachwissenschaft}.
-Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking unordered dependency trees $T$ as the underlying meanings to be conveyed.
+Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking dependency structures as the underlying meanings to be conveyed.
 Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
 
-\subsection{Optimizing Grammars for Efficiency}
-
-We now ask:
-(1) Do the grammars of natural languages evolve towards optimizing efficiency of communication?
-(2) Does this process of optimization account for Greenberg's order correlation universals?
+%\subsection{Optimizing Grammars for Efficiency}
+
+Our goals are to show that
+(1) the grammars of natural languages evolve towards optimizing efficiency of communication, and
+(2) this process of optimization accounts for Greenberg's order correlation universals.
 
@@
 
-We first demonstrate that real languages are relatively efficient compared to random baselines.
-For each language, we generated ten baseline word order grammars for the language by choosing all word order grammar parameters randomly at uniform from $[0,1]$.
+We first demonstrate that real languages are relatively efficient compared to random baselines. For each language, we generated ten baseline word order grammars for the language.
+These grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}).
+%by choosing all word order grammar parameters randomly at uniform from $[0,1]$.
 Figure~\ref{fig:pareto-plane} shows the predictability and parseability for each real language relative to its baseline grammars.
@@
 \subsection{Efficiency Explains DLM}
-Finally we demonstrate the relationship between dependency length minimization and the maximization of efficiency.
+We demonstrate the relationship between dependency length minimization and the maximization of efficiency.
 Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency.
 We see that optimizing for efficiency lowers dependency length relative to random baselines, in keeping with the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}.
-In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the depenndency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.
+In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the dependency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.
 
@@
 
-We now examine to what extent we can recover Greenberg's word order correlations in optimized grammars. 
-Dryer \cite{dryer1992greenbergian} presents a comprehensive updated version of the word order correlations, drawing on 625 languages, which we take as the basis of our evaluation.
+We now show that efficiency optimization predicts Greenberg's word order correlations. % in optimized grammars. 
+We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
 In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
-Most of them can be straightforwardly implemented in UD, allowing us to check which correlations a word order grammar satisfies.
+Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
 %This dependency corresponds to dependencies of the form $\xrightarrow{obj}$ in UD.
@@
 
-Dryer \cite{dryer1992greenbergian} presents three correlations which do not correspond to dependencies annotated in UD: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
-Two pairs of Dryer's correlations, namely those for the dependencies between complementizers and adverbial subordinators and their complement clauses,and those for the dependencies between verbs and adpositional phrases, and adjectives and their standard of comparison, had to be collapsed into two correlations in UD.
+Dryer \cite{dryer1992greenbergian} presents three correlations that are only relevant to part of the world's languages and are not annotated reliably in UD.
+%: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
+Further collapsing correlations that cannot be formalized individually in UD, we otained 10 formalized correlations from Dryer's 15 correlations.
+%Two pairs of Dryer's correlations, namely those for the dependencies between complementizers and adverbial subordinators and their complement clauses,and those for the dependencies between verbs and adpositional phrases, and adjectives and their standard of comparison, had to be collapsed into two correlations in UD.
 From Dryer's 15 correlations, we obtained 10 formalized correlations, roughly covering nine of Greenberg's original universals.
@@
 
-\begin{figure}
-    \centering
-%     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-efficiency.pdf}
- %    \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-langmod.pdf}
-  %   \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}
-   %  \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-depl.pdf}
-     \includegraphics[scale=.45]{../results/correlations/figures/posterior-satisfied-universals-together.pdf}
-
-	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks, computed from a mixed-effects logistic regression jointly modeling all ten dependencies.
-Efficiency predicts all ten correlations with high posterior probability.}
-    \label{fig:posterior-satisfied-count}
-\end{figure}
+\subsection{Discussion}
+
+In Section~\ref{sec:relative-efficiency}, we found that word order grammars in the majority of UD languages have better efficiency than baseline word order grammars.
+Furthermore, in Section~\ref{sec:greenberg} we found that explicitly optimizing grammars for efficiency reproduces 10 major word order correlations reported in the typological literature, and with greater accuracy than the typological literature when it comes to the languages studied.
+
+%While the efficiency objective succeeds in predicting the presence of the most word order correlations, the dependency length objective seems to predict a greater prevalence for those correlations that it does predict.
+We believe this discrepancy is a result of noise in the optimization process.
+Optimizing for efficiency is computationally challenging, requiring stochastic gradient descent on word order grammar, language model, and parser parameters, while optimizing for dependency length is simpler and thus more optimal grammars can be reached easily (See SI section 2 for supporting data).
+Inasmuch as dependency length minimization is computationally easier to implement than efficiency maximization, we believe it will remain useful as an easy-to-use heuristic for predicting and explaining word order universals.
+% TODO we can also report (in the SI) prevalences among the best-performing of the optimized languages, i.e., those where optimization was most successful.
+I would expect that this should narrow the gap between DepL and Efficiency, which would confirm that the gap is really about the difficulty of optimization, not about the objective functions.
+
+\section{Conclusion}
+
+We found that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
+We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals.
+Beyond our present results, we provide a complete formalization and computational framework in which theories of the functional optimization of languages can be tested.
+Other objective functions could be hypothesized and tested within our framework; furthermore, future advances in machine learning will enable the optimization of richer and richer models of grammar.
+
+A major question for functional explanations for linguistic universals is: \emph{how} do languages come to be optimized?
+Do speakers actively seek out new communicative conventions that allow better efficiency?
+Or do languages change in response to biases that come into play during language acquisition \cite{fedzechkina2012language,culbertson2012learning}?
+Our work is neutral toward such questions.
+To the extent that language universals arise from biases in learning or in the representational capacity of the human brain, our results suggest that those biases tilt toward communicative efficiency.
+Our work does provide against the idea that word order universals are best explained in terms of learning biases that are irreducibly arbitrary and genetic in nature \cite{chomsky2010some}.
+
+While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into dependency trees.
+The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication.
+Relatedly, there are many remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
+Our models cannot capture this trade-off for technical reasons: the parseability objective does not operate over wordforms, only POS tags, and so it does not take advantage of morphological cues to syntactic structure.
+Future work can investigate whether these and other remaining universals can be explained using more sophisticated models of how meaning representations are transduced into strings of words, based on larger databases with richer annotations.
 
@@
 For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
+
+\matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
+We use all languages for which at least one treebank with a training partition was available, a total of 50 languages.
+For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.
+Punctuation was removed.
+%We removed punctuation. %, which we operationalized by removing all dependents of \emph{punct} dependencies, and recursively removing their dependents.
+
+Universal dependencies represents as dependents some words that are typically classified as heads in syntactic theory.
+This particularly applies to the \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark} dependencies.
+Following prior work studying dependency length minimization \cite{futrell2015largescale}, we modified each treebank by inverting these dependencies, promoting the dependent to the head position.
+We report results on this modified version of UD.
+
+The efficiency optimization results from Table~\ref{tab:results-dryer} were preregistered: \url{http://aspredicted.org/blind.php?x=8gp2bt}.
+
+See the SI for details on the neural language models and parsers used, and on the optimization procedures.}
 
@@
 
-See the SI for details on the neural language models and parsers used, and on the optimization procedures.
+See the SI for details on the neural language models and parsers used, and on the optimization procedures.}
 

### 2019-02-09 second-draft/pnas-word-order-si.tex
revising paper

@@
 \end{figure}
+
+\begin{figure}
+    \centering
+%     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-efficiency.pdf}
+ %    \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-langmod.pdf}
+  %   \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}
+   %  \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-depl.pdf}
+     \includegraphics[scale=.45]{../results/correlations/figures/posterior-satisfied-universals-together.pdf}
+
+	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks, computed from a mixed-effects logistic regression jointly modeling all ten dependencies.
+Efficiency predicts all ten correlations with high posterior probability.}
+    \label{fig:posterior-satisfied-count}
+\end{figure}
+
+We further evaluate the word order predictions of efficiency, showing that efficiency is most successful in predicting correlations in the direction found in the UD corpora.
+We constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency in languages optimized for efficiency, with random effects for language and language family, correlated across the ten dependencies.
+We conducted the same analysis for predictability, parseability, and dependency length.
+We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
+The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
+The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$.
+The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.
+For dependency length, the posterior puts much of the probability mass on predicting only nine of the correlations; predictability and parseability predict significantly less correlations.
 

### 2019-02-09 results/correlations/evaluation-depl_graphic.R

@@ -1,0 +1,102 @@
+
+data = read.csv("../../grammars/manual_output_funchead_coarse_depl/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+
+
+library(forcats)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+
+data = data %>% mutate(Language = fct_recode(Language, "Ancient_Greek" = "Ancient", "Old_Church_Slavonic" = "Old"))
+
+
+
+
+dryer_greenberg_fine = data
+
+
+
+languages = read.csv("../languages/languages-iso_codes.tsv")
+dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"), all.x=TRUE)
+
+
+
+#options(mc.cores = parallel::detectCores())
+#rstan_options(auto_write = TRUE)
+
+
+dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS = rbind(dataS, dataS2) %>% filter(Type == "manual_output_funchead_two_coarse_lambda09_best_balanced")
+
+dataP = read.csv("../../grammars/plane/plane-parse.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP2 = read.csv("../../grammars/plane/plane-parse-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP = rbind(dataP, dataP2) %>% filter(Type == "manual_output_funchead_two_coarse_lambda09_best_balanced")
+
+
+dataE = merge(dataS, dataP, by=c("Type", "Language", "Model")) %>% mutate(Eff = Pars+0.9*Surp) %>% group_by(Type, Language, Model) %>% summarise(Eff = mean(Eff))
+
+dataE = dataE[order(dataE$Language, dataE$Model),]
+
+eff = dataE$Eff
+models = c()
+for(i in (1:202)) {
+   if(eff[2*i] < eff[2*i+1]) {
+	   models = c(models, dataE$Model[2*i])
+   } else {
+	   models = c(models, dataE$Model[2*i+1])
+   }
+}
+
+
+dryer_greenberg_fine = dryer_greenberg_fine%>% filter(FileName %in% models)
+
+
+
+
[…]

### 2019-02-09 results/correlations/evaluation-efficiency09_test.R

@@ -1,0 +1,73 @@
+
+# _final/
+data = read.csv("../../grammars/manual_output_funchead_two_coarse_lambda09_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+
+library(forcats)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+
+data = data %>% mutate(Language = fct_recode(Language, "Ancient_Greek" = "Ancient", "Old_Church_Slavonic" = "Old"))
+
+
+
+
+dryer_greenberg_fine = data
+
+
+
+languages = read.csv("../languages/languages-iso_codes.tsv")
+dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"), all.x=TRUE)
+
+
+library("brms")
+
+#options(mc.cores = parallel::detectCores())
+#rstan_options(auto_write = TRUE)
+
+dependency = "nmod"
+
+getCorrPair = function(dependency) {
+   corr_pair = dryer_greenberg_fine %>% filter((CoarseDependency == dependency) | (CoarseDependency == "obj"))
+   corr_pair = unique(corr_pair %>% select(Family, Language, FileName, CoarseDependency, DH_Weight )) %>% spread(CoarseDependency, DH_Weight)
+   corr_pair$correlator = corr_pair[[dependency]]
+   corr_pair = corr_pair %>% mutate(correlator_s = pmax(0,sign(correlator)), obj_s=pmax(0,sign(obj)))
+#   corr_pair = corr_pair %>% mutate(obj_s = obj_s - mean(obj_s))
+   corr_pair = corr_pair %>% mutate(correlator_s = ifelse(correlator == 0, NA, correlator_s)) # special case of actual zero, indicating fully random ordering: has to be excluded from analysis. Such cases occur when a dependency does not occur in the training partition, but does occur in the validation partition.
+   corr_pair$agree = (corr_pair$correlator_s == corr_pair$obj_s)
+   return(corr_pair)
+}
+
+mean(getCorrPair("lifted_cop")$agree)
+
+
+
+
+dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
[…]

### 2019-02-09 results/correlations/evaluation-langmod_test.R

@@ -1,0 +1,68 @@
+
+# _final/
+data = read.csv("../../grammars/manual_output_funchead_langmod_coarse_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+
+library(forcats)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+
+data = data %>% mutate(Language = fct_recode(Language, "Ancient_Greek" = "Ancient", "Old_Church_Slavonic" = "Old"))
+
+
+
+
+dryer_greenberg_fine = data
+
+
+
+languages = read.csv("../languages/languages-iso_codes.tsv")
+dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"), all.x=TRUE)
+
+
+library("brms")
+
+#options(mc.cores = parallel::detectCores())
+#rstan_options(auto_write = TRUE)
+
+dependency = "nmod"
+
+getCorrPair = function(dependency) {
+   corr_pair = dryer_greenberg_fine %>% filter((CoarseDependency == dependency) | (CoarseDependency == "obj"))
+   corr_pair = unique(corr_pair %>% select(Family, Language, FileName, CoarseDependency, DH_Weight )) %>% spread(CoarseDependency, DH_Weight)
+   corr_pair$correlator = corr_pair[[dependency]]
+   corr_pair = corr_pair %>% mutate(correlator_s = pmax(0,sign(correlator)), obj_s=pmax(0,sign(obj)))
+#   corr_pair = corr_pair %>% mutate(obj_s = obj_s - mean(obj_s))
+   corr_pair = corr_pair %>% mutate(correlator_s = ifelse(correlator == 0, NA, correlator_s)) # special case of actual zero, indicating fully random ordering: has to be excluded from analysis. Such cases occur when a dependency does not occur in the training partition, but does occur in the validation partition.
+   corr_pair$agree = (corr_pair$correlator_s == corr_pair$obj_s)
+   return(corr_pair)
+}
+
+mean(getCorrPair("lifted_cop")$agree)
+
+
+
+
+dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
[…]

### 2019-02-20 second-draft/3-paper-names.tex
new draft

--- 2019-02-09 second-draft/2-paper-names.tex
@@
 An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-
 In this work we examine this question from a computational perspective, demonstrating a fully formalized framework in which certain syntactic universals can be explained through the statistical optimization of grammars for information-theoretic efficiency.
@@
 
-We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
-These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages.
-For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
-
-A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
-Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.
-
-Researchers working in the functional paradigm have proposed a range of criteria which languages should meet in order to enable efficient communication and processing.
-These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
-
-The contribution of this work is (i) to provide a principled formalization of the notion of communicative efficiency in word order, based on related work in information theory \cite{cover2006elements,tishby1999information}, bounded rationality \cite{genewein2015bounded}, and computational neuroscience \cite{friston2009predictive}; (ii) to introduce a computational framework in which we can produce simulated grammars directly optimized for our notion of efficiency, using state-of-the-art machine learning techniques \cite{hochreiter1997long,goldberg2017neural,dozat2017stanford}; and (iii) to demonstrate that in doing so, we reproduce the major word order patterns reported in \cite{greenberg1963universals} and related work \cite{dryer1992greenbergian}, as well as reproducing the predictions of previous heuristic theories of word order optimization \cite{hawkins1994performance,ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
-
-We formalize the notion of communicative efficiency using two terms: one having to do with minimizing the difficulty of linguistic communication, and one having to do with maximizing the information transferred in such communication.
-We quantify the difficulty of linguistic communication in terms of the word-by-word \key{predictability} of strings, which has been found to be a highly accurate and general predictor of human online comprehension difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and which can be justified in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
-We quantify the information transferred in language as \key{parseability}: the quantity of information contained in linguistic utterances about latent hierarchical tree structures.
-We also compare the results of optimizing for communicative efficiency to the results of optimizing for dependency length minimization, a previously proposed approximate metric of the amount of memory involved in language production and comprehension \cite{hudson1995measuring,gibson1998linguistic,gibson2000dependency,liu2017dependency,temperley2018minimizing}.
-
-%To evaluate whether communicative efficiency can explain word order universals, we apply the following method \cite{gildea2007optimizing}.
-First, we develop word order grammars: simple, interpretable models of word order, implemented as functions that take in an unordered syntactic dependency tree and output a distribution over possible word orders for that tree.
-Second, we take dependency treebanks of natural languages and compute word order grammars which are optimized for an objective functions representing communicative efficiency.
-%Third, we evaluate whether the optimized word order grammars reproduce the desired word order universals when applied to attested dependency trees from the treebanks across many languages.
-Our favorable results suggest that it may be possible to generally model human languages as efficient codes for communication among agents with generic human-like cognitive and communicative constraints.
-%The best way to describe a human language might be as a solution to an equation involving such constraints.
-
-\section{Greenberg's Word Order Correlations}
+\textbf{TODO maybe should be more general in the beginning, as in the first version. Also make sure the reader doesn't think 'it's just headedness, why bother'.}
+
 Natural languages vary a lot in the order in which they express information.
@@
 
-TODO maybe being more general in the beginning, as in the first version. make sure the reader doesn't think 'it's just headedness, why bother'
-
 \begin{figure}
@@
 
+A commonly proposed functional explanation of Greenberg's correlations is in terms of Dependency Length Minimization \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
+This has been confirmed on the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, and has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.
+
+%In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+%Dependency length has been proposed as a metric of processing difficulty on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, and minimization of dependency length is known to recover a large subset of the Greenbergian universals \cite{hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic}.
+%We demonstrate that the minimization of dependency length can be formalized in our framework, reproducing the finding that optimizing grammars for dependency length recovers several word order universals \cite{gildea2010grammars}.
+
 Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
@@
 The idea that language results from the tension between these two pressures has a long and fruitful history.
-\cite{horn-toward-1984}, RSA, Color naming
-
+%\cite{horn-toward-1984}, RSA, Color naming
+%
 %In the domain of word order, 
-
+%
 %TODO introduce the terms \key{cost} and \key{informativity}
@@
 Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb.
-}
+\textbf{TODO This figure is just a draft sketch so far.}}
 	\label{fig:grammars}

### 2019-02-24 second-draft/4-paper-names.tex
visualization

--- 2019-02-20 second-draft/3-paper-names.tex
@@
 % TODO add a cogsci person to citation chain at the beginning
-
-\textbf{TODO maybe should be more general in the beginning, as in the first version. Also make sure the reader doesn't think 'it's just headedness, why bother'.}
 
@@
 
-Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
-According to this theory, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
-Conversely, to minimize listener effort, the language should provide different utterances for different meanings, so that the listeenr can unambiguously identify the meaning from the utterance.
-Minimizing speaker and listener effort represent two opposing forces:
-A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
-A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
-%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
-Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
-The idea that language results from the tension between these two pressures has a long and fruitful history.
-%\cite{horn-toward-1984}, RSA, Color naming
-%
-%In the domain of word order, 
-%
-%TODO introduce the terms \key{cost} and \key{informativity}
-%
-This idea has been shown to account for phenomena such as pragmatic inference \cite{horn-toward-1984, frank2012predicting} and color naming~\cite{zaslavsky2018efficient}.
-
-%TODO somewhere start calling this Efficiency
-%TODO explain why surprisal. some production/speaker work. (plan reuse)
-
-%In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
-%Even earlier than Zipf, \cite{gabelentz1901sprachwissenschaft} ...
-%This theory blabla Horn RSA
-%Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
-
 In this work, we show that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
@@
 
-%\section{DLM}
-
-%
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.45\textwidth]{figures/letter-options.png}
-%	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure.
+\section{DLM}
+
+A commonly proposed functional explanation of Greenberg's correlations is in terms of Dependency Length Minimization \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
+This has been confirmed on the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, and has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.7\textwidth]{figures/four-grammars-dlm.png}
+	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure.
 In this example, Grammar 1 achieves the lowest overall dependency length.
 Indeed, this is the }
-%	\label{fig:pareto-plane}
-%\end{figure}
-%
-%
+	\label{fig:dlm-grammars}
+\end{figure*}
+
+\begin{figure}
+\centering
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   katabt \& risāla \& li \& sadīq  \\
+	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+	   wrote \& letter   \& to \& friend  \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{1}{2}{obj}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{1}{3}{}
+   \depedge{3}{4}{}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+      tomodachi \& ni \& tegami-o \& kaita \\
+ \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+	   friend \& to \& letter  \& wrote \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{4}{2}{}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{2}{1}{}
+   \depedge{4}{3}{obj}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+\end{figure}
+
+The sentences in Figure~\ref{fig:arabic-japanese} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words.
+For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} if the verb, rather than e.g.
+the subject or an indirect object.
+A dependency arc is drawn from a \key{head} (e.g.
+\emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g.
+\emph{tegami-o}).
+Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}.
+Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.
+
+In order to interpret a sentence correctly, a listener must be able to reconstruct the information expressed in the dependency tree and its labels.
+This is because, according to the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.
 
@@
 
+Dependency Length Minimization is heuristic and stipulative.
+
+Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
+According to this theory, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
+Conversely, to minimize listener effort, the language should provide different utterances for different meanings, so that the listeenr can unambiguously identify the meaning from the utterance.
+Minimizing speaker and listener effort represent two opposing forces:
+A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
+A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
+%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
+Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
+The idea that language results from the tension between these two pressures has a long and fruitful history.
+%\cite{horn-toward-1984}, RSA, Color naming
+%
+%In the domain of word order, 
+%
+%TODO introduce the terms \key{cost} and \key{informativity}
+%
+This idea has been shown to account for phenomena such as pragmatic inference \cite{horn-toward-1984, frank2012predicting} and color naming~\cite{zaslavsky2018efficient}.
+
+%TODO somewhere start calling this Efficiency
+%TODO explain why surprisal. some production/speaker work. (plan reuse)
+
+%In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
+%Even earlier than Zipf, \cite{gabelentz1901sprachwissenschaft} ...
+%This theory blabla Horn RSA
+%Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
+
 We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
@@
 In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the dependency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.
+
+\begin{figure*} % need Efficiency and (maybe) MLE langs on here
+\includegraphics[width=0.98\textwidth]{../results/dependency-length/figures/depLength-facet.pdf}
+\caption{Average dependency length as a function of sentence length in four languages.
+Across languages, real and optimized languages have shorter dependencies than random baseline orderings.}\label{fig:deplength}
+\end{figure*}
+
+\subsection{Greenbergian Word Order Correlations}
+\label{sec:greenberg}
+\begin{table*}
+	\begin{center}
+\begin{tabular}{|ll|l|l|l|ll|l|}
+	\hline
+	%&
+	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
+	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%adp. &
+adposition    &    NP       
+	&   \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf}   
+	&    \textbf{81}$^{***}$    
+	&   \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf}   
+	&    \textbf{76}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}  \\
+copula    &    NP       
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf}
+	&    \textbf{81}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf}
+	&    \textbf{79}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}  \\
+auxiliary    &    VP       
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf} 
+	&    \textbf{74}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf} 
+	&    55    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  \\
+noun    &    genitive      
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf} 
+	&    \textbf{82}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf} 
+	&    \textbf{74}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  \\
+noun    &    relative clause       
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf} 
+	&    \textbf{85}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf} 
+	&    \textbf{77}$^{**}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  \\
+complementizer    &    S        
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf} 
+	&    \textbf{85}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf} 
+	&    \textbf{80}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  \\
+verb    &    PP     
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf} 
+	&    \textbf{78}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf} 
+	&    59    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  \\
+want    &    VP      
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf} 
+	&    \textbf{90}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf} 
+	&    \textbf{92}$^{***}$    
+	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf} 
+  \\
+ \hline
+    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+\end{tabular}
+	\end{center}
+	\caption{Greenbergian Correlations.
+Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%	For clarity, we organize the correlations by the category of the `verb patterner'.
+	For each correlation, we give our operationalization in terms of UD.
+%A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
+	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
+	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
+\end{table*}
+
+We now show that efficiency optimization predicts Greenberg's word order correlations. % in optimized grammars. 
+We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
+In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
+Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
+
+Dryer \cite{dryer1992greenbergian} presents three correlations that are only relevant to part of the world's languages and are not annotated reliably in UD.
+One further dependency is not annotated reliably in UD. %: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
+Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations from Dryer's 15 correlations.
+
+In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
+We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau$ show the same direction for the correlating dependency and for the verb-object dependency, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
+%We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau > 0.5$ holds for the correlating dependency depending on whether $a_\tau > 0.5$ for the verb-object dependency, with per-language-family random intercept and slope.\footnote{We coded language families according to \url{universaldependencies.org}.}
+We are interested in the direction and significance of this effect: 
+If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
+
+We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
+To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
+The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
+%in all except four languages, for WALS did not have data.
+
+\paragraph{Results}
+Results are shown in Table~\ref{table:results-dryer}.
+All eight correlations are confirmed in the models estimated from the real orderings. 
+We will discuss these exceptions further below.
+
+Confirming previous theoretical arguments, we see that optimizing for dependency length correctly accounts for all word order correlations. 
+Predictability and parseability predict four and six correlations, respectively, making complementary predictions.
+Efficiency correctly predicts all the word order correlations.
+
+\subsection{Discussion}
+
+In Section~\ref{sec:relative-efficiency}, we found that word order grammars in the majority of UD languages have better efficiency than baseline word order grammars.
+Furthermore, in Section~\ref{sec:greenberg} we found that explicitly optimizing grammars for efficiency reproduces the Greenbergian correlations to the extent that they can be tested on the basis of available data.
+
+\section{Conclusion}
+
+We found that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
+We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals.
+Beyond our present results, we provide a complete formalization and computational framework in which theories of the functional optimization of languages can be tested.
+Other objective functions could be hypothesized and tested within our framework; furthermore, future advances in machine learning will enable the optimization of richer and richer models of grammar.
+
+A major question for functional explanations for linguistic universals is: \emph{how} do languages come to be optimized?
+Do speakers actively seek out new communicative conventions that allow better efficiency?
+Or do languages change in response to biases that come into play during language acquisition \cite{fedzechkina2012language,culbertson2012learning}?
+Our work is neutral toward such questions.
+To the extent that language universals arise from biases in learning or in the representational capacity of the human brain, our results suggest that those biases tilt toward communicative efficiency.
+Our work does provide against the idea that word order universals are best explained in terms of learning biases that are irreducibly arbitrary and genetic in nature \cite{chomsky2010some}.
+
+While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into dependency trees.
+The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication.
+Relatedly, there are many remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
+Our models cannot capture this trade-off for technical reasons: the parseability objective does not operate over wordforms, only POS tags, and so it does not take advantage of morphological cues to syntactic structure.
+Future work can investigate whether these and other remaining universals can be explained using more sophisticated models of how meaning representations are transduced into strings of words, based on larger databases with richer annotations.
+
+\section{DLM TODO put these at the appropriate place}
+
+We compare Efficiency to the predictions of Dependency Length Minimization.
+This theory states that natural languages order information in such a way that the distances between syntactically linked words are minimized.
+For instance, in Figure~\ref{fig:arabic-japanese-simple}, there would be syntactic links between the adposition (to) and the noun (friend), and between the verb and both the object (letter) and the adposition (to).
+There is strong evidence that natural language minimizes the length of these dependencies~\cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
+Theoretical work in the functional linguistics literature has proposed that this principle explains the correlations.
+However, the principle itself is stipulative and not a first-principles explanation.
+\cite{hawkins1994performance} theoetically argues that it increases  parsability.
+
+However, these arguments have been made on a theoretical basis.
+We will show that (1) Dependency Length Minimization indeeds predicts most of the correlations, and (2) Dependency Length Minimization is explained by efficiency optimization.
+
+These computational results confirm theoretical ideas  that have been stated informally by authors in the functional linguistics literature at least since the 1980s.
+
+%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
+These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages.
+For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
+
+\matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
+We use all languages for which at least one treebank with a training partition was available, a total of 50 languages.
+For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.
+Punctuation was removed.
+%We removed punctuation. %, which we operationalized by removing all dependents of \emph{punct} dependencies, and recursively removing their dependents.
+
+Universal dependencies represents as dependents some words that are typically classified as heads in syntactic theory.
+This particularly applies to the \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark} dependencies.
+Following prior work studying dependency length minimization \cite{futrell2015largescale}, we modified each treebank by inverting these dependencies, promoting the dependent to the head position.
+We report results on this modified version of UD.
+
+The efficiency optimization results from Table~\ref{tab:results-dryer} were preregistered: \url{http://aspredicted.org/blind.php?x=8gp2bt}.
+
+See the SI for details on the neural language models and parsers used, and on the optimization procedures.}
+
+\showmatmethods{} % Display the Materials and Methods section
+
+\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Chris Manning, and audiences at CAMP 2018 for helpful discussion.}
+
+\showacknow{} % Display the acknowledgments section
+
+%\bibliographystyle{acl_natbib}
+%\bibliography{extracted}
+\bibliography{everything}
+
+\end{document}
+
+%the meaning
+
+\begin{figure}
+\centering
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   katabt \& risāla \& li \& sadīq  \\
+	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+	   wrote \& letter   \& to \& friend  \\
+   \end{deptext}
+\end{dependency}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+      tomodachi \& ni \& tegami-o \& kaita \\
+ \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+	   friend \& to \& letter  \& wrote \\
+   \end{deptext}
+\end{dependency}
+
+%\begin{dependency}[theme=simple]
+%    \begin{deptext}
+%    I \& wrote \& a \& letter \& to \& a \& friend \\
+%    \end{deptext}
+%    \depedge{2}{1}{nsubj}
+%    \depedge{2}{4}{obj}
+%    \depedge{4}{3}{ }
+%    \depedge{4}{3}{ }
+%    \depedge{2}{5}{}
+%    \depedge{5}{7}{ }
+%    \depedge{7}{6}{ }
+%\end{dependency}
+	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese-simple}
+\end{figure}
+
+\section{Minimizing Dependency Length Predicts Universals}
+
+..
+
+The contribution of this work is (i) to provide a principled formalization of the notion of communicative efficiency in word order, based on related work in information theory \cite{cover2006elements,tishby1999information}, bounded rationality \cite{genewein2015bounded}, and computational neuroscience \cite{friston2009predictive}; (ii) to introduce a computational framework in which we can produce simulated grammars directly optimized for our notion of efficiency, using state-of-the-art machine learning techniques \cite{hochreiter1997long,goldberg2017neural,dozat2017stanford}; and (iii) to demonstrate that in doing so, we reproduce the major word order patterns reported in \cite{greenberg1963universals} and related work \cite{dryer1992greenbergian}, as well as reproducing the predictions of previous heuristic theories of word order optimization \cite{hawkins1994performance,ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
+
+We formalize the notion of communicative efficiency using two terms: one having to do with minimizing the difficulty of linguistic communication, and one having to do with maximizing the information transferred in such communication.
+We quantify the difficulty of linguistic communication in terms of the word-by-word \key{predictability} of strings, which has been found to be a highly accurate and general predictor of human online comprehension difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and which can be justified in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+We quantify the information transferred in language as \key{parseability}: the quantity of information linguistic utterances contain about latent hierarchical tree structures.
+We also compare the results of optimizing for communicative efficiency to the results of optimizing for dependency length minimization, a previously proposed approximate metric of the amount of memory involved in language production and comprehension \cite{hudson1995measuring,gibson1998linguistic,gibson2000dependency,liu2017dependency,temperley2018minimizing}.
+
+To evaluate whether communicative efficiency can explain word order universals, we apply the following method \cite{gildea2007optimizing}.
+First, we develop word order grammars: simple, interpretable models of word order, implemented as functions that take in an unordered syntactic dependency tree and output a distribution over possible word orders for that tree.
+Second, we take dependency treebanks of natural languages and compute word order grammars which are optimized for an objective functions representing communicative efficiency.
+Third, we evaluate whether the optimized word order grammars reproduce the desired word order universals when applied to attested dependency trees from the treebanks across many languages.
+Our favorable results suggest that it may be possible to generally model human languages as efficient codes for communication among agents with generic human-like cognitive and communicative constraints.
+%The best way to describe a human language might be as a solution to an equation involving such constraints.
+
+%We draw on data from the Universal Dependencies project \cite{nivre-universal-2017}, which has been developing treebanks in a unified format for several dozens of diverse languages.
+%This universal annotation scheme is particularly useful for studying word order universals, as many of them can be operationalized  in terms of crosslinguistically valid syntactic notions.
+% such as \emph{verbs} and \emph{objects}.
+
+\section{Background}
+
+% TODO where do we say that we're basing our work on data from 51 languages?
+
+%\subsection{Explananda: Word Order Correlations}
+%
+%Working with a database of 30 languages, Greenberg \cite{greenberg1963universals} proposed 45 universals of language that he determined to be true of most or all languages.
+%31 of them were concerned with word order, the remainder with morphology.
+%Many of the word order universals state \emph{correlations} between patterns:
+%For instance, Universals 3 and 4 state that ``\emph{Languages with dominant VSO order are always prepositional.}''  (e.g.
+Arabic), and ``\emph{With overwhelmingly greater than chance frequency, languages with normal SOV order are postpositional.}'' (e.g.
+Japanese).
+%These are illustrated in Figure~\ref{fig:arabic-japanese}:
+%Arabic has basic order verb-object (VO) and prepositions (\emph{li} `to'), while Japanese has order object-verb (OV) and postpositions (\emph{ni} `to').
+Our goal is to explain Greenberg's reported word order correlations, as updated in \cite{dryer1992greenbergian}.
+
+\begin{figure}
+\centering
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   katabt \& risāla \& li \& sadīq  \\
+	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+	   wrote \& letter   \& to \& friend  \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{1}{2}{obj}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{1}{3}{}
+   \depedge{3}{4}{}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+      tomodachi \& ni \& tegami-o \& kaita \\
+ \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+	   friend \& to \& letter  \& wrote \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{4}{2}{}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{2}{1}{}
+   \depedge{4}{3}{obj}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+
+%\begin{dependency}[theme=simple]
+%    \begin{deptext}
+%    I \& wrote \& a \& letter \& to \& a \& friend \\
+%    \end{deptext}
+%    \depedge{2}{1}{nsubj}
+%    \depedge{2}{4}{obj}
+%    \depedge{4}{3}{ }
+%    \depedge{4}{3}{ }
+%    \depedge{2}{5}{}
+%    \depedge{5}{7}{ }
+%    \depedge{7}{6}{ }
+%\end{dependency}
+	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese}
+\end{figure}
+
+% TODO what to do about these labels...? They're not UD...
+
+\subsection{Dependency trees}
+
+\section{Optimizing word order grammars}
+
+Here we describe our model of word order and how we derive optimal grammars within that model in terms of communicative efficiency.
+
+Let $t$ be an unordered dependency tree, i.e.
+a structure as in Figure~\ref{fig:arabic-japanese} but without any specified ordering of the words.
+We think of an unordered dependency tree $t$ as representing the compositional meaning to be conveyed by a sentence.
+We define a \key{word order grammar} as a function $L_\theta(t)$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output a probability distribution over ordered sequences of words $\mathbf{w}$ linearizing the tree, or equivalently, a probability distribution over orders on the nodes of the tree.
+% do we need to define "linearize"?
+This process is visualized in Figure~\ref{fig:process-visualization}.
+
+Next we define how a word order grammar linearizes trees in terms of its parameters $\theta$, and then we describe how $\theta$ can be set to optimize objective functions related to communicative efficiency.
+
+\begin{figure}
+\includegraphics[width=0.48\textwidth]{figures/process-visualization.png}
+\caption{Our method for constructing counterfactual versions of languages: We extract unordered dependency tree topologies from dependency corpora, and apply word order grammars to obtain counterfactually reordered corpora.}\label{fig:process-visualization}
+\end{figure}
+
+We think of it as the following problem:
+
+A speaker has a tree in mind, and they have to produce an utterance.
+
+%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a word order grammar.
+Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
+%\begin{equation}
+%\label{eq:exp-over-utts}
+%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim L_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
+%\end{equation}
+%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
+In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
+%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
+%
+%
+
+All natural languages have some degree of word order regularity.
+Thus, it is not sufficient to optimize the word orders of individual sentences in the corpora---instead, we will optimize the word order rules of entire languages.
+We develop a method to create word order rules that optimize given processing measures on the resulting counterfactual corpus.
+
+\subsection{DLM Predicts Universals}
+
+We first apply our method to show that DLM predicts the correlations.
+
+%Dependency length has been proposed as a metric of processing difficulty on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, and minimization of dependency length is known to recover a large subset of the Greenbergian universals \cite{hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic}.
+%We demonstrate that the minimization of dependency length can be formalized in our framework, reproducing the finding that optimizing grammars for dependency length recovers several word order universals \cite{gildea2010grammars}.
+%
+%To quantify dependency length, we sum the lengths of syntactic dependencies in the sentence \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
+%\begin{equation}
+%    \label{eq:depl}
+%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+%\end{equation}
+%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
+Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
+
+\subsection{Efficiency Predicts Universals}
+
+DLM is stipulative and at most a heuristic approximation to psycholinguistic findings
+
+We now introduce a general information-theoretic notion of language efficiency.
+
+use the general speaker-listener picture
+
+\paragraph{Predictability}
+Predictability represents the incremental difficulty of online language processing.
+We formalize predictability as the \key{surprisal} (negative log probability) of each word given its preceding context \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+Optimization of predictability has been proposed as an explanation for certain word order patterns by \cite{ferrericancho2017optimality}.
+
+We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
+\begin{equation*}
+R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
+\end{equation*}
+%where the language model parameters $\phi$ are chosen to predict the language induced by tree distribution $T$ and word order grammar $\L_\theta$ as accurately as possible (maximizing predictability):
+%\begin{equation*}
+%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation*}
+We implement the language model $P_\phi$ using an LSTM neural network architecture \cite{hochreiter1997long}, currently the state of the art for this task \cite{goldberg2017neural}.
+%Note that the language model parameters $\phi$ and the word order grammar parameters $\theta$ are recursively dependent on each other.
+See the SI for details on how $\theta$ and $\phi$ are jointly optimized, as well as on the neural network architecture.
+
+\paragraph{Parseability}
+Parseability is the extent to which the syntactic and semantic relationships among words can be recovered from the string of words. 
+%Some notion of parseability has been proposed informally as a desirable property of languages in the linguistic literature going back at least to the 19th century \cite{humboldt,sapir,jesperson}, and parseability has been offered as an underlying justification for syntactic locality as a cost function for language processing \cite{hawkins-performance-1994}.
+%At a basic level, a language can encode information about syntactic relations into fixed word order patterns or into morphological marking on individual words.
+Much linguistic work has been devoted to investigating the trade-offs of various ways of indicating syntactic relations \cite{}.
+In this work, we present what is to our knowledge the first fully implemented computational formalization of this idea.
+We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
+Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
+\begin{align}
+\nonumber
+R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
+\nonumber
+&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
+\end{align}
+%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
+%\begin{equation*}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation*}
+We implement $P_\psi$ using a highly generic graph-based parsing architecture \cite{kiperwasser2016simple,dozat2017stanford} which recovers graph structures from word strings by solving a minimal spanning tree problem.
+%Due to data sparsity, we train parsers to predict dependency trees from POS tags alone, not from full wordforms, therefore we do not capture the extent to which morphological wordforms are informative about dependency tree structure.
+See the SI (appendix section 6-7) for further details on the joint optimization of $\theta$ and $\psi$ and on the neural parser architecture, and see SI (section 8) for robustness to different parsing models.
+
+\paragraph{Efficiency}
+The two scoring functions---predictability and parseability---can be combined into an overall equation of the form:
+\begin{equation}
+\label{eq:combined-score}
+	R_{\textit{Eff}} = R_{Pars} + \lambda R_{Pred},
+\end{equation}
+with an interpolation weight $\lambda \in [0,1)$.
+In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification).
+This combined equation represents efficiency: bits of information transferred about dependency trees minus the cost of the transmission.
+In expectation over utterances (substituting Eq.~\ref{eq:combined-score} into Eq.~\ref{eq:exp-over-utts}), the overall objective function to be maximized by a word order grammar can be written in information-theoretic terms as:
+\begin{equation}
+    \label{eq:efficiency}
+    J_T(\theta) = I[L_\theta(T);T] - \lambda H[L_\theta(T)],
+\end{equation}
+where $I[L_\theta(T);T]$ is the \key{mutual information} between strings and trees, and $H[L_\theta(T)]$ is the \key{entropy} of strings \cite{cover2006elements}.
+Maximizing Eq.~\ref{eq:efficiency} is equivalent to maximizing mutual information between strings and trees subject to a constraint on the maximum entropy of strings.
+
+The two terms of Eq.~\ref{eq:efficiency} implement two opposing pressures.
+Parseability (the mutual information term) pushes to make strings more heterogeneous, so that they can indicate different tree structures unambiguously.
+Predictability (the entropy term) pushes to make strings more homogeneous, to reduce the complexity of using the language.
+The idea that natural language arises from a tension between these two forces goes back over a century \cite{gabelentz1901sprachwissenschaft}.
+Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking unordered dependency trees $T$ as the underlying meanings to be conveyed.
+Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
+
+%In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+\section{Results}
+
+\subsection{Relative efficiency of languages}
+\label{sec:relative-efficiency}
+
+We first demonstrate that real languages are relatively efficient compared to random baselines.
+For each language, we generated ten baseline word order grammars for the language by choosing all word order grammar parameters randomly at uniform from $[0,1]$.
+Figure~\ref{fig:pareto-plane} shows the predictability and parseability for each real language relative to its baseline grammars.
+In order to control for limitations due to our word order grammar formalism, we represent real languages in the figure by maximum likelihood fits of word order grammars to the real language data.
+%; in general these maximum-likelihood word order grammars end up with lower predictability and parseability than the real languages. % TODO verify this is true? TODO awkward sentence
+For the calculation of predictability and parseability, we make all (baseline and real) word order grammars deterministic by always choosing the highest-probability linearization of each tree; by making the grammars deterministic in this way we eliminate an anticonservative bias toward low predictability in the baseline languages, which are highly nondeterministic.
+The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend.pdf}
+    \caption{Predictability and parseability of 51 UD languages (red), indicated by ISO codes, compared to ten baseline word order grammars per language (green).
+Predictability and parseability scores are $z$-scored within language.
+Each point for a real language has a line pointing in the direction of the center of mass of its baselines.
+The green contour shows the density of baseline languages.
+Unlabeled dots represent the centroid for real languages (red), baseline languages (green), and languages optimized for predictability (yellow), parseability (pink), efficiency (blue), and dependency length (red).
+When a language is to the bottom--left of its baselines, this indicates that it is relatively optimal for efficiency.} % TODO colors, fontsize, etc could conceivably be made prettier
+    \label{fig:pareto-plane}
+\end{figure}
+
+Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
+Languages appear to be attracted toward these points and away from the region of the baseline languages.
+We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
+% TODO add something about dependency length-optimized languages?
+
+These results are shown in more detail for two typologically distinct languages (English and Japanese) in Figure~\ref{fig:predict-pars}.
+Here we include also the predictability and parseability of the original language data and of the word order grammars optimized for dependency length.
+We see that real languages are even more predictable than any optimized or baseline language, while optimization for either parseability or dependency length suffices to raise parseability to the level of the real languages.
+This result is in line with the finding that languages with longer dependencies are harder to parse in general \cite{gulordava2016multilingual}.
+
+\begin{figure*}
+\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
+	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right).
+We provide the entropy $H[L_\theta(T)]$ for predictability (lower is better) and the conditional entropy $H[T|L_\theta(T)]$ for parseability (lower is better).
+Both measures are normalized by sentence length.}\label{fig:predict-pars}
+\end{figure*}
+
+Finally we demonstrate the relationship between dependency length minimization and the maximization of efficiency.
+Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency.
+We see that optimizing for efficiency lowers dependency length relative to random baselines, in keeping with the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}.
+In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the depenndency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.
 
@@
 
-We now show that efficiency optimization predicts Greenberg's word order correlations. % in optimized grammars. 
-We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
-In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
-Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
-%This dependency corresponds to dependencies of the form $\xrightarrow{obj}$ in UD.
-%Similarly, most of the other dependencies in Dryer's formulation can be formalized:
-%for instance, the dependency between nouns and relative clauses can be formalized as $\xrightarrow{acl}$.
-%Testing whether an ordering model satisfies the correlation between relative clauses and objects reduces to checking whether these two relation types have direction parameters $a_\tau$ that are either both $>0.5$ (both kinds of dependents precede their heads) or both $< 0.5$ (they both follow their heads).
-
-Dryer \cite{dryer1992greenbergian} presents three correlations that are only relevant to part of the world's languages and are not annotated reliably in UD.
-%: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
-Further collapsing correlations that cannot be formalized individually in UD, we otained 10 formalized correlations from Dryer's 15 correlations.
-%Two pairs of Dryer's correlations, namely those for the dependencies between complementizers and adverbial subordinators and their complement clauses,and those for the dependencies between verbs and adpositional phrases, and adjectives and their standard of comparison, had to be collapsed into two correlations in UD.
-From Dryer's 15 correlations, we obtained 10 formalized correlations, roughly covering nine of Greenberg's original universals.
-%TODO update to 10?
-What happened to the 11th?
-Was that amod?
-[The 11th was adjective+standard of comparison, which is merged into obl] Maybe we could include amod as a "true rejection" of efficiency, but a false alarm for DepL
-
-In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
-We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau$ show the same direction for the correlating dependency and for the verb-object dependency, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
-%We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau > 0.5$ holds for the correlating dependency depending on whether $a_\tau > 0.5$ for the verb-object dependency, with per-language-family random intercept and slope.\footnote{We coded language families according to \url{universaldependencies.org}.}
-We are interested in the direction and significance of this effect: 
-If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
-
-We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
-To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
-The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
-%in all except four languages, for WALS did not have data.
-
-\paragraph{Results}
-Results are shown in Table~\ref{table:results-dryer}.
-All correlations but two are confirmed in the models estimated from the real orderings.
-The exceptions are the subject--verb dependency and the verb--adverb dependency, which typically go in the opposite direction from the standard description.
-We will discuss these exceptions further below.
-
-In keeping with previous work, we see that optimizing for dependency length correctly accounts for nine word order correlations, missing only the verb--adverb dependency. 
-Predictability and parseability predict five and seven correlations, respectively, making largely complementary predictions.
-Efficiency significantly predicts all the word order correlations, each in the same direction as attested in the dependency corpora.
-
-We now address the two word order correlations whose direction in the dependency corpora is opposite from what would be expected in the typological literature.
-The first is the correlation of the order of verb--subject and verb--object dependencies.
-Our sample of mainly European languages highly over-represents languages with the general order subject--verb--object (such as English), in which the order of the verb--subject and verb--object dependencies are anti-correlated.
-Surprisingly, given the sample of tree structures of these languages, it turns out that the optimal languages tend to have anti-correlated orders for subjects and objects order similar to the real languages.
-
-The second anomalous dependency is the verb--manner adverb dependency.
-We believe the anti-correlation in the UD corpora arises because the \emph{advmod} dependency does not distinguish between manner adverbs---the subject of the typological judgment---and various other types of modifiers such as sentence-level adverbs.
-Nevertheless, the languages optimized for efficiency reproduce the anti-correlation of the orders of verb--object and verb--adverb at around the same rate as the real languages.
-
-We further evaluate the word order predictions of efficiency, showing that efficiency is most successful in predicting correlations in the direction found in the UD corpora.
-We constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency in languages optimized for efficiency, with random effects for language and language family, correlated across the ten dependencies.
-We conducted the same analysis for predictability, parseability, and dependency length.
-We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
-The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
-The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$.
-The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.
-For dependency length, the posterior puts much of the probability mass on predicting only nine of the correlations; predictability and parseability predict significantly less correlations.
-
-\subsection{Discussion}
-
-In Section~\ref{sec:relative-efficiency}, we found that word order grammars in the majority of UD languages have better efficiency than baseline word order grammars.
-Furthermore, in Section~\ref{sec:greenberg} we found that explicitly optimizing grammars for efficiency reproduces 10 major word order correlations reported in the typological literature, and with greater accuracy than the typological literature when it comes to the languages studied.
-
-%While the efficiency objective succeeds in predicting the presence of the most word order correlations, the dependency length objective seems to predict a greater prevalence for those correlations that it does predict.
-We believe this discrepancy is a result of noise in the optimization process.
-Optimizing for efficiency is computationally challenging, requiring stochastic gradient descent on word order grammar, language model, and parser parameters, while optimizing for dependency length is simpler and thus more optimal grammars can be reached easily (See SI section 2 for supporting data).
-Inasmuch as dependency length minimization is computationally easier to implement than efficiency maximization, we believe it will remain useful as an easy-to-use heuristic for predicting and explaining word order universals.
-% TODO we can also report (in the SI) prevalences among the best-performing of the optimized languages, i.e., those where optimization was most successful.
-I would expect that this should narrow the gap between DepL and Efficiency, which would confirm that the gap is really about the difficulty of optimization, not about the objective functions.
-
-\section{Conclusion}
-
-We found that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
-We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals.
-Beyond our present results, we provide a complete formalization and computational framework in which theories of the functional optimization of languages can be tested.
-Other objective functions could be hypothesized and tested within our framework; furthermore, future advances in machine learning will enable the optimization of richer and richer models of grammar.
-
-A major question for functional explanations for linguistic universals is: \emph{how} do languages come to be optimized?
-Do speakers actively seek out new communicative conventions that allow better efficiency?
-Or do languages change in response to biases that come into play during language acquisition \cite{fedzechkina2012language,culbertson2012learning}?
-Our work is neutral toward such questions.
-To the extent that language universals arise from biases in learning or in the representational capacity of the human brain, our results suggest that those biases tilt toward communicative efficiency.
-Our work does provide against the idea that word order universals are best explained in terms of learning biases that are irreducibly arbitrary and genetic in nature \cite{chomsky2010some}.
-
-While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into dependency trees.
-The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication.
-Relatedly, there are many remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
-Our models cannot capture this trade-off for technical reasons: the parseability objective does not operate over wordforms, only POS tags, and so it does not take advantage of morphological cues to syntactic structure.
-Future work can investigate whether these and other remaining universals can be explained using more sophisticated models of how meaning representations are transduced into strings of words, based on larger databases with richer annotations.
-
-\section{DLM TODO put these at the appropriate place}
-
-We compare Efficiency to the predictions of Dependency Length Minimization.
-This theory states that natural languages order information in such a way that the distances between syntactically linked words are minimized.
-For instance, in Figure~\ref{fig:arabic-japanese-simple}, there would be syntactic links between the adposition (to) and the noun (friend), and between the verb and both the object (letter) and the adposition (to).
-There is strong evidence that natural language minimizes the length of these dependencies~\cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
-Theoretical work in the functional linguistics literature has proposed that this principle explains the correlations.
-However, the principle itself is stipulative and not a first-principles explanation.
-\cite{hawkins1994performance} theoetically argues that it increases  parsability.
-
-However, these arguments have been made on a theoretical basis.
-We will show that (1) Dependency Length Minimization indeeds predicts most of the correlations, and (2) Dependency Length Minimization is explained by efficiency optimization.
-
-These computational results confirm theoretical ideas  that have been stated informally by authors in the functional linguistics literature at least since the 1980s.
-
-%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
-These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages.
-For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
-
-\matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
-We use all languages for which at least one treebank with a training partition was available, a total of 50 languages.
-For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.
-Punctuation was removed.
-%We removed punctuation. %, which we operationalized by removing all dependents of \emph{punct} dependencies, and recursively removing their dependents.
-
-Universal dependencies represents as dependents some words that are typically classified as heads in syntactic theory.
-This particularly applies to the \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark} dependencies.
-Following prior work studying dependency length minimization \cite{futrell2015largescale}, we modified each treebank by inverting these dependencies, promoting the dependent to the head position.
-We report results on this modified version of UD.
-
-The efficiency optimization results from Table~\ref{tab:results-dryer} were preregistered: \url{http://aspredicted.org/blind.php?x=8gp2bt}.
-
-See the SI for details on the neural language models and parsers used, and on the optimization procedures.}
-
-\showmatmethods{} % Display the Materials and Methods section
-
-\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Chris Manning, and audiences at CAMP 2018 for helpful discussion.}
-
-\showacknow{} % Display the acknowledgments section
-
-%\bibliographystyle{acl_natbib}
-%\bibliography{extracted}
-\bibliography{everything}
-
-\end{document}
-
-%the meaning
-
-\begin{figure}
-\centering
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
-	   katabt \& risāla \& li \& sadīq  \\
-	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
-	   wrote \& letter   \& to \& friend  \\
-   \end{deptext}
-\end{dependency}
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-      tomodachi \& ni \& tegami-o \& kaita \\
- \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
-	   friend \& to \& letter  \& wrote \\
-   \end{deptext}
-\end{dependency}
-
-%\begin{dependency}[theme=simple]
-%    \begin{deptext}
-%    I \& wrote \& a \& letter \& to \& a \& friend \\
-%    \end{deptext}
-%    \depedge{2}{1}{nsubj}
-%    \depedge{2}{4}{obj}
-%    \depedge{4}{3}{ }
-%    \depedge{4}{3}{ }
-%    \depedge{2}{5}{}
-%    \depedge{5}{7}{ }
-%    \depedge{7}{6}{ }
-%\end{dependency}
-	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese-simple}
-\end{figure}
-
-\section{Minimizing Dependency Length Predicts Universals}
-
-..
-
-\begin{figure}
-\centering
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
-	   katabt \& risāla \& li \& sadīq  \\
-	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
-	   wrote \& letter   \& to \& friend  \\
-   \end{deptext}
-	%   \deproot{3}{ROOT}
-   \depedge{1}{2}{obj}
-	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
-   \depedge{1}{3}{}
-   \depedge{3}{4}{}
-   %\depedge[arc angle=50]{7}{6}{ATT}
-\end{dependency}
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-      tomodachi \& ni \& tegami-o \& kaita \\
- \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
-	   friend \& to \& letter  \& wrote \\
-   \end{deptext}
-	%   \deproot{3}{ROOT}
-   \depedge{4}{2}{}
-	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
-   \depedge{2}{1}{}
-   \depedge{4}{3}{obj}
-   %\depedge[arc angle=50]{7}{6}{ATT}
-\end{dependency}
-\end{figure}
-
-Dependency Length:
-
-The sentences in Figure~\ref{fig:arabic-japanese} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words.
-For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} if the verb, rather than e.g.
-the subject or an indirect object.
-A dependency arc is drawn from a \key{head} (e.g.
-\emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g.
-\emph{tegami-o}).
-Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}.
-Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.
-
-In order to interpret a sentence correctly, a listener must be able to reconstruct the information expressed in the dependency tree and its labels.
-This is because, according to the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.
-
-The contribution of this work is (i) to provide a principled formalization of the notion of communicative efficiency in word order, based on related work in information theory \cite{cover2006elements,tishby1999information}, bounded rationality \cite{genewein2015bounded}, and computational neuroscience \cite{friston2009predictive}; (ii) to introduce a computational framework in which we can produce simulated grammars directly optimized for our notion of efficiency, using state-of-the-art machine learning techniques \cite{hochreiter1997long,goldberg2017neural,dozat2017stanford}; and (iii) to demonstrate that in doing so, we reproduce the major word order patterns reported in \cite{greenberg1963universals} and related work \cite{dryer1992greenbergian}, as well as reproducing the predictions of previous heuristic theories of word order optimization \cite{hawkins1994performance,ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
-
-We formalize the notion of communicative efficiency using two terms: one having to do with minimizing the difficulty of linguistic communication, and one having to do with maximizing the information transferred in such communication.
-We quantify the difficulty of linguistic communication in terms of the word-by-word \key{predictability} of strings, which has been found to be a highly accurate and general predictor of human online comprehension difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and which can be justified in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
-We quantify the information transferred in language as \key{parseability}: the quantity of information linguistic utterances contain about latent hierarchical tree structures.
-We also compare the results of optimizing for communicative efficiency to the results of optimizing for dependency length minimization, a previously proposed approximate metric of the amount of memory involved in language production and comprehension \cite{hudson1995measuring,gibson1998linguistic,gibson2000dependency,liu2017dependency,temperley2018minimizing}.
-
-To evaluate whether communicative efficiency can explain word order universals, we apply the following method \cite{gildea2007optimizing}.
-First, we develop word order grammars: simple, interpretable models of word order, implemented as functions that take in an unordered syntactic dependency tree and output a distribution over possible word orders for that tree.
-Second, we take dependency treebanks of natural languages and compute word order grammars which are optimized for an objective functions representing communicative efficiency.
-Third, we evaluate whether the optimized word order grammars reproduce the desired word order universals when applied to attested dependency trees from the treebanks across many languages.
-Our favorable results suggest that it may be possible to generally model human languages as efficient codes for communication among agents with generic human-like cognitive and communicative constraints.
-%The best way to describe a human language might be as a solution to an equation involving such constraints.
-
-%We draw on data from the Universal Dependencies project \cite{nivre-universal-2017}, which has been developing treebanks in a unified format for several dozens of diverse languages.
-%This universal annotation scheme is particularly useful for studying word order universals, as many of them can be operationalized  in terms of crosslinguistically valid syntactic notions.
-% such as \emph{verbs} and \emph{objects}.
-
-\section{Background}
-
-% TODO where do we say that we're basing our work on data from 51 languages?
-
-%\subsection{Explananda: Word Order Correlations}
-%
-%Working with a database of 30 languages, Greenberg \cite{greenberg1963universals} proposed 45 universals of language that he determined to be true of most or all languages.
-%31 of them were concerned with word order, the remainder with morphology.
-%Many of the word order universals state \emph{correlations} between patterns:
-%For instance, Universals 3 and 4 state that ``\emph{Languages with dominant VSO order are always prepositional.}''  (e.g.
-Arabic), and ``\emph{With overwhelmingly greater than chance frequency, languages with normal SOV order are postpositional.}'' (e.g.
-Japanese).
-%These are illustrated in Figure~\ref{fig:arabic-japanese}:
-%Arabic has basic order verb-object (VO) and prepositions (\emph{li} `to'), while Japanese has order object-verb (OV) and postpositions (\emph{ni} `to').
-Our goal is to explain Greenberg's reported word order correlations, as updated in \cite{dryer1992greenbergian}.
-
-\begin{figure}
-\centering
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
-	   katabt \& risāla \& li \& sadīq  \\
-	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
-	   wrote \& letter   \& to \& friend  \\
-   \end{deptext}
-	%   \deproot{3}{ROOT}
-   \depedge{1}{2}{obj}
-	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
-   \depedge{1}{3}{}
-   \depedge{3}{4}{}
-   %\depedge[arc angle=50]{7}{6}{ATT}
-\end{dependency}
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-      tomodachi \& ni \& tegami-o \& kaita \\
- \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
-	   friend \& to \& letter  \& wrote \\
-   \end{deptext}
-	%   \deproot{3}{ROOT}
-   \depedge{4}{2}{}
-	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
-   \depedge{2}{1}{}
-   \depedge{4}{3}{obj}
-   %\depedge[arc angle=50]{7}{6}{ATT}
-\end{dependency}
-
-%\begin{dependency}[theme=simple]
-%    \begin{deptext}
-%    I \& wrote \& a \& letter \& to \& a \& friend \\
-%    \end{deptext}
-%    \depedge{2}{1}{nsubj}
-%    \depedge{2}{4}{obj}
-%    \depedge{4}{3}{ }
-%    \depedge{4}{3}{ }
-%    \depedge{2}{5}{}
-%    \depedge{5}{7}{ }
-%    \depedge{7}{6}{ }
-%\end{dependency}
-	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese}
-\end{figure}
-
-% TODO what to do about these labels...? They're not UD...
-
-\subsection{Dependency trees}
-
-\section{Optimizing word order grammars}
-
-Here we describe our model of word order and how we derive optimal grammars within that model in terms of communicative efficiency.
-
-Let $t$ be an unordered dependency tree, i.e.
-a structure as in Figure~\ref{fig:arabic-japanese} but without any specified ordering of the words.
-We think of an unordered dependency tree $t$ as representing the compositional meaning to be conveyed by a sentence.
-We define a \key{word order grammar} as a function $L_\theta(t)$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output a probability distribution over ordered sequences of words $\mathbf{w}$ linearizing the tree, or equivalently, a probability distribution over orders on the nodes of the tree.
-% do we need to define "linearize"?
-This process is visualized in Figure~\ref{fig:process-visualization}.
-
-Next we define how a word order grammar linearizes trees in terms of its parameters $\theta$, and then we describe how $\theta$ can be set to optimize objective functions related to communicative efficiency.
-
-\begin{figure}
-\includegraphics[width=0.48\textwidth]{figures/process-visualization.png}
-\caption{Our method for constructing counterfactual versions of languages: We extract unordered dependency tree topologies from dependency corpora, and apply word order grammars to obtain counterfactually reordered corpora.}\label{fig:process-visualization}
-\end{figure}
-
-We think of it as the following problem:
-
-A speaker has a tree in mind, and they have to produce an utterance.
-
-%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a word order grammar.
-Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
-%\begin{equation}
-%\label{eq:exp-over-utts}
-%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim L_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
-%\end{equation}
-%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
-In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
-%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
-%
-%
-
-All natural languages have some degree of word order regularity.
-Thus, it is not sufficient to optimize the word orders of individual sentences in the corpora---instead, we will optimize the word order rules of entire languages.
-We develop a method to create word order rules that optimize given processing measures on the resulting counterfactual corpus.
-
-\subsection{DLM Predicts Universals}
-
-We first apply our method to show that DLM predicts the correlations.
-
-%Dependency length has been proposed as a metric of processing difficulty on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, and minimization of dependency length is known to recover a large subset of the Greenbergian universals \cite{hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic}.
-%We demonstrate that the minimization of dependency length can be formalized in our framework, reproducing the finding that optimizing grammars for dependency length recovers several word order universals \cite{gildea2010grammars}.
-%
-%To quantify dependency length, we sum the lengths of syntactic dependencies in the sentence \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
-%\begin{equation}
-%    \label{eq:depl}
-%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
-%\end{equation}
-%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
-Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
-
-\subsection{Efficiency Predicts Universals}
-
-DLM is stipulative and at most a heuristic approximation to psycholinguistic findings
-
-We now introduce a general information-theoretic notion of language efficiency.
-
-use the general speaker-listener picture
-
-\paragraph{Predictability}
-Predictability represents the incremental difficulty of online language processing.
-We formalize predictability as the \key{surprisal} (negative log probability) of each word given its preceding context \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
-Optimization of predictability has been proposed as an explanation for certain word order patterns by \cite{ferrericancho2017optimality}.
-
-We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
-\begin{equation*}
-R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
-\end{equation*}
-%where the language model parameters $\phi$ are chosen to predict the language induced by tree distribution $T$ and word order grammar $\L_\theta$ as accurately as possible (maximizing predictability):
-%\begin{equation*}
-%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-%\end{equation*}
-We implement the language model $P_\phi$ using an LSTM neural network architecture \cite{hochreiter1997long}, currently the state of the art for this task \cite{goldberg2017neural}.
-%Note that the language model parameters $\phi$ and the word order grammar parameters $\theta$ are recursively dependent on each other.
-See the SI for details on how $\theta$ and $\phi$ are jointly optimized, as well as on the neural network architecture.
-
-\paragraph{Parseability}
-Parseability is the extent to which the syntactic and semantic relationships among words can be recovered from the string of words. 
-%Some notion of parseability has been proposed informally as a desirable property of languages in the linguistic literature going back at least to the 19th century \cite{humboldt,sapir,jesperson}, and parseability has been offered as an underlying justification for syntactic locality as a cost function for language processing \cite{hawkins-performance-1994}.
-%At a basic level, a language can encode information about syntactic relations into fixed word order patterns or into morphological marking on individual words.
-Much linguistic work has been devoted to investigating the trade-offs of various ways of indicating syntactic relations \cite{}.
-In this work, we present what is to our knowledge the first fully implemented computational formalization of this idea.
-We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
-Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
-\begin{align}
-\nonumber
-R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
-\nonumber
-&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
-\end{align}
-%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
-%\begin{equation*}
-%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-%\end{equation*}
-We implement $P_\psi$ using a highly generic graph-based parsing architecture \cite{kiperwasser2016simple,dozat2017stanford} which recovers graph structures from word strings by solving a minimal spanning tree problem.
-%Due to data sparsity, we train parsers to predict dependency trees from POS tags alone, not from full wordforms, therefore we do not capture the extent to which morphological wordforms are informative about dependency tree structure.
-See the SI (appendix section 6-7) for further details on the joint optimization of $\theta$ and $\psi$ and on the neural parser architecture, and see SI (section 8) for robustness to different parsing models.
-
-\paragraph{Efficiency}
-The two scoring functions---predictability and parseability---can be combined into an overall equation of the form:
-\begin{equation}
-\label{eq:combined-score}
-	R_{\textit{Eff}} = R_{Pars} + \lambda R_{Pred},
-\end{equation}
-with an interpolation weight $\lambda \in [0,1)$.
-In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification).
-This combined equation represents efficiency: bits of information transferred about dependency trees minus the cost of the transmission.
-In expectation over utterances (substituting Eq.~\ref{eq:combined-score} into Eq.~\ref{eq:exp-over-utts}), the overall objective function to be maximized by a word order grammar can be written in information-theoretic terms as:
-\begin{equation}
-    \label{eq:efficiency}
-    J_T(\theta) = I[L_\theta(T);T] - \lambda H[L_\theta(T)],
-\end{equation}
-where $I[L_\theta(T);T]$ is the \key{mutual information} between strings and trees, and $H[L_\theta(T)]$ is the \key{entropy} of strings \cite{cover2006elements}.
-Maximizing Eq.~\ref{eq:efficiency} is equivalent to maximizing mutual information between strings and trees subject to a constraint on the maximum entropy of strings.
-
-The two terms of Eq.~\ref{eq:efficiency} implement two opposing pressures.
-Parseability (the mutual information term) pushes to make strings more heterogeneous, so that they can indicate different tree structures unambiguously.
-Predictability (the entropy term) pushes to make strings more homogeneous, to reduce the complexity of using the language.
-The idea that natural language arises from a tension between these two forces goes back over a century \cite{gabelentz1901sprachwissenschaft}.
-Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking unordered dependency trees $T$ as the underlying meanings to be conveyed.
-Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
-
-%In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-\section{Results}
-
-\subsection{Relative efficiency of languages}
-\label{sec:relative-efficiency}
-
-We first demonstrate that real languages are relatively efficient compared to random baselines.
-For each language, we generated ten baseline word order grammars for the language by choosing all word order grammar parameters randomly at uniform from $[0,1]$.
-Figure~\ref{fig:pareto-plane} shows the predictability and parseability for each real language relative to its baseline grammars.
-In order to control for limitations due to our word order grammar formalism, we represent real languages in the figure by maximum likelihood fits of word order grammars to the real language data.
-%; in general these maximum-likelihood word order grammars end up with lower predictability and parseability than the real languages. % TODO verify this is true? TODO awkward sentence
-For the calculation of predictability and parseability, we make all (baseline and real) word order grammars deterministic by always choosing the highest-probability linearization of each tree; by making the grammars deterministic in this way we eliminate an anticonservative bias toward low predictability in the baseline languages, which are highly nondeterministic.
-The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend.pdf}
-    \caption{Predictability and parseability of 51 UD languages (red), indicated by ISO codes, compared to ten baseline word order grammars per language (green).
-Predictability and parseability scores are $z$-scored within language.
-Each point for a real language has a line pointing in the direction of the center of mass of its baselines.
-The green contour shows the density of baseline languages.
-Unlabeled dots represent the centroid for real languages (red), baseline languages (green), and languages optimized for predictability (yellow), parseability (pink), efficiency (blue), and dependency length (red).
-When a language is to the bottom--left of its baselines, this indicates that it is relatively optimal for efficiency.} % TODO colors, fontsize, etc could conceivably be made prettier
-    \label{fig:pareto-plane}
-\end{figure}
-
-Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
-Languages appear to be attracted toward these points and away from the region of the baseline languages.
-We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
-% TODO add something about dependency length-optimized languages?
-
-These results are shown in more detail for two typologically distinct languages (English and Japanese) in Figure~\ref{fig:predict-pars}.
-Here we include also the predictability and parseability of the original language data and of the word order grammars optimized for dependency length.
-We see that real languages are even more predictable than any optimized or baseline language, while optimization for either parseability or dependency length suffices to raise parseability to the level of the real languages.
-This result is in line with the finding that languages with longer dependencies are harder to parse in general \cite{gulordava2016multilingual}.
-
-\begin{figure*}
-\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
-	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right).
-We provide the entropy $H[L_\theta(T)]$ for predictability (lower is better) and the conditional entropy $H[T|L_\theta(T)]$ for parseability (lower is better).
-Both measures are normalized by sentence length.}\label{fig:predict-pars}
-\end{figure*}
-
-Finally we demonstrate the relationship between dependency length minimization and the maximization of efficiency.
-Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency.
-We see that optimizing for efficiency lowers dependency length relative to random baselines, in keeping with the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}.
-In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the depenndency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.
-
-\begin{figure*} % need Efficiency and (maybe) MLE langs on here
-\includegraphics[width=0.98\textwidth]{../results/dependency-length/figures/depLength-facet.pdf}
-\caption{Average dependency length as a function of sentence length in four languages.
-Across languages, real and optimized languages have shorter dependencies than random baseline orderings.}\label{fig:deplength}
-\end{figure*}
-
-\subsection{Greenbergian Word Order Correlations}
-\label{sec:greenberg}
-\begin{table*}
-	\begin{center}
-\begin{tabular}{|ll|l|l|l|ll|l|}
-	\hline
-	%&
-	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
-	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%adp. &
-adposition    &    NP       &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
-copula    &    NP        &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
-auxiliary    &    VP       &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
-noun    &    genitive      &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
-noun    &    relative clause       &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
-complementizer    &    S        &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
-verb    &    PP     &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
-want    &    VP      &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
-verb    &    subject    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
-verb    &    manner adverb     &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
-\end{tabular}
-	\end{center}
-	\caption{Greenbergian Correlations.
-Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%	For clarity, we organize the correlations by the category of the `verb patterner'.
-	For each correlation, we give our operationalization in terms of UD.
-%A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
-	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
-	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
-\end{table*}
-
-%\begin{table*}
-%\begin{tabular}{lllllllllllllllllll}
-%	&	          & \multicolumn{2}{c}{Dep. Length}      & \multicolumn{2}{c}{Lang. Mod.}              &  \multicolumn{2}{c}{Parsing}  &\multicolumn{2}{c}{Operationalization} & & GU\\ \hline
-%	Adp &	Adp+Noun &                  & \cmark       &             &            &         &             & \textsc{noun} $\xrightarrow{case}$ \textsc{adp} & - &  3, 4\\ \hline
-%	Verb &	Verb+PP & \checkmark        & \cmark      & \checkmark  &   \cmark         &    &             & \textsc{verb} $\xrightarrow{obl}$ \textsc{noun}  & + & 7  \\
-%	Verb &	Copula & \checkmark         &  \cmark     &  \checkmark &   \cmark    &  \checkmark  &     \cmark        & \textsc{noun} $\xrightarrow{cop}$  \textsc{aux} & - \\
-%	Verb &	want & \checkmark           & \cmark      & \checkmark  & \cmark     &  \checkmark  &    \cmark         & \textsc{verb} $\xrightarrow{xcomp}$ \textsc{verb} & + & 15 \\
-%	Verb &	subject &                   &       & \checkmark  &      \cmark      & \xmark   &    \xmark         & \textsc{verb} $\xrightarrow{nsubj}$ \textsc{noun} & + \\
-%	Verb &	auxiliary & \cmark          &       & \cmark      & \cmark     &           &             & \textsc{verb}  $\xrightarrow{aux}$ \textsc{verb}/\textsc{aux}  & - & 16\\
-%	verb & manner adverb &              &  -     &  \xmark     &   \xmark     &          &             & \textsc{verb} $\xrightarrow{advmod}$ ADV  & + & 7 \\ \hline
-%	Comp &	comp; adv sub. &            &   \cmark    &             &   \cmark         &          &             & \textsc{verb} $\xrightarrow{mark}$ \textsc{sconj} & - & \\ \hline
-%	Noun &	article N' &                &   \xmark/\xmark    &             &    -/\cmark        &          &   -/\cmark          & \textsc{noun} $\xrightarrow{det}$ \textsc{det}/\textsc{pron}  & -\\
-%	Noun &	noun+genitive & \cmark      &   \cmark    &             &            &              &    \cmark  & \textsc{noun} $\xrightarrow{nmod}$ \textsc{noun}  & + & 2\\
-%	Noun &	noun+relative clse & \cmark &  \cmark   &               &            &  \checkmark &      \cmark       & \textsc{noun} $\xrightarrow{acl}$ \textsc{verb} & + & 24 \\ \hline
-%	Adj &	adj+std            &        &  \cmark   &               &            &             &             &  ADJ $\xrightarrow{nmod}$ \textsc{adp}  & + & 23\\
-%\end{tabular}
-%	\caption{Greenbergian Correlations, as provided by \cite{dryer-greenbergian-1992}.
-For each correlation, we give our operationalization in terms of UD and whether the direction of this dependency correlates positively ($+$) or negatively ($-$) with the verb-object-dependency.
-For each correlation and each objective function, we indicate whether the correlation is predicted correctly (\cmark), in the wrong direction (\xmark), or no prediction is made (blank).
-For each objective function, we show results for plain UD (left column) and UD with function word heads (right column).
-}\label{table:results-dryer}
-%\end{table*}
-%%, with corresponding numbers in the original Greenberg universals, where applicable.
-
-%In the last sections, we confirmed the effectiveness of our optimization method and elucidated the relation between the three optimization functions on the level of .
-
 We now examine to what extent we can recover Greenberg's word order correlations in optimized grammars. 

### 2019-02-24 results/correlations/evaluation-efficiency09_graphic.R

@@ -96,6 +96,18 @@ corr_pairs_sum = corr_pairs %>% group_by(Dependency, obj_s, correlator_s) %>% su
 corr_pairs_sum = corr_pairs_sum %>% mutate(agree_color = ifelse(agree, "red", "blue"))
 plot = ggplot(corr_pairs_sum, aes(x=obj_s, y=correlator_s, group=Dependency)) + geom_point(aes(size=count, color=agree_color)) + facet_wrap(~Dependency) + theme_bw() + xlim(-0.5, 1.5) + ylim(-0.5, 1.5)  + theme(legend.position="none")
 
+for(dependency in dependencies) {
+   plot = ggplot(corr_pairs_sum %>% filter(Dependency == dependency), aes(x=obj_s, y=correlator_s)) + geom_point(aes(size=count, color=agree_color))  + theme_bw() + xlim(-0.5, 1.5) + ylim(-0.5, 1.5)  + theme(legend.position="none") + xlab(NULL) + ylab(NULL)+
+  theme(axis.title=element_blank(),
+        axis.text=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.ticks=element_blank()) +  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
+  ggsave(plot, file=paste("figures/correlations/correlation-efficiency-", dependency, ".pdf", sep=""), width=1, height=1)
+
+
+}
+
+
 
 
 

### 2019-02-24 results/correlations/evaluation-ground_graphic.R

@@ -61,6 +61,16 @@ corr_pairs_sum = corr_pairs_sum %>% mutate(agree_color = ifelse(agree, "red", "b
 plot = ggplot(corr_pairs_sum, aes(x=obj_s, y=correlator_s, group=Dependency)) + geom_point(aes(size=count, color=agree_color)) + facet_wrap(~Dependency, ncol=1) + theme_bw() + xlim(-0.5, 1.5) + ylim(-0.5, 1.5)  + theme(legend.position="none")
 
 
+for(dependency in dependencies) {
+   plot = ggplot(corr_pairs_sum %>% filter(Dependency == dependency), aes(x=obj_s, y=correlator_s)) + geom_point(aes(size=count, color=agree_color))  + theme_bw() + xlim(-0.5, 1.5) + ylim(-0.5, 1.5)  + theme(legend.position="none") + xlab(NULL) + ylab(NULL)+
+  theme(axis.title=element_blank(),
+        axis.text=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.ticks=element_blank()) +  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
+  ggsave(plot, file=paste("figures/correlations/correlation-ground-", dependency, ".pdf", sep=""), width=1, height=1)
+
+
+}
 
 
 

### 2019-02-24 results/correlations/evaluation-langmod_graphic.R

@@ -1,14 +1,18 @@
 
-data = read.csv("../../grammars/manual_output_funchead_coarse_depl/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
-
+# _final/
+data = read.csv("../../grammars/manual_output_funchead_langmod_coarse_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
 
 library(forcats)
 library(dplyr)
 library(tidyr)
 library(ggplot2)
 
+
+best = read.csv("../strongest_models/best-langmod-best-balanced.csv")
+
 data = data %>% mutate(Language = fct_recode(Language, "Ancient_Greek" = "Ancient", "Old_Church_Slavonic" = "Old"))
 
+data = merge(data %>% mutate(FileName = as.character(FileName)), best %>% rename(FileName = Model), by=c("Language", "FileName"))
 
 
 
@@ -24,41 +28,9 @@ dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"),
 #options(mc.cores = parallel::detectCores())
 #rstan_options(auto_write = TRUE)
 
-
-dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataS = rbind(dataS, dataS2) %>% filter(Type == "manual_output_funchead_two_coarse_lambda09_best_balanced")
-
-dataP = read.csv("../../grammars/plane/plane-parse.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataP2 = read.csv("../../grammars/plane/plane-parse-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataP = rbind(dataP, dataP2) %>% filter(Type == "manual_output_funchead_two_coarse_lambda09_best_balanced")
-
-
-dataE = merge(dataS, dataP, by=c("Type", "Language", "Model")) %>% mutate(Eff = Pars+0.9*Surp) %>% group_by(Type, Language, Model) %>% summarise(Eff = mean(Eff))
-
-dataE = dataE[order(dataE$Language, dataE$Model),]
-
-eff = dataE$Eff
[…]

### 2019-02-25 second-draft/4-paper-names.tex
revising

@@
 
-In this work, we show that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
+In this work, we show that the word orders of languages have evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
 
@@
 
-\section{DLM}
-
-A commonly proposed functional explanation of Greenberg's correlations is in terms of Dependency Length Minimization \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+\section{Dependency Length Minimization}
+
+Some prior work has has suggested a functional explanation of Greenberg's correlations in terms of Dependency Length Minimization \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
 This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
-This has been confirmed on the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, and has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=0.7\textwidth]{figures/four-grammars-dlm.png}
-	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure.
-In this example, Grammar 1 achieves the lowest overall dependency length.
-Indeed, this is the }
-	\label{fig:dlm-grammars}
-\end{figure*}
+On the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, it has been confirmed that languages reduce dependency length.
+This has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.
+
+(CITE) have made theoretical arguments that dependency length minimization should produce word order correlations.
+Here, we present the first large-scale test of this idea.
 
@@
 \centering
+
 \begin{dependency}[theme = simple]
@@
 %	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
-	   katabt \& risāla \& li \& sadīq  \\
-	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
-	   wrote \& letter   \& to \& friend  \\
+	   I \&	   visited \& friends  \\
    \end{deptext}
 	%   \deproot{3}{ROOT}
-   \depedge{1}{2}{obj}
-	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
-   \depedge{1}{3}{}
+   \depedge{2}{1}{nsubj}
+   \depedge{2}{3}{obj}
+\end{dependency}
+
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   I \&	   went   \& to \& school  \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{2}{1}{nsubj}
+   \depedge{2}{3}{}
    \depedge{3}{4}{}
@@
 \end{dependency}
+
 \begin{dependency}[theme = simple]
    \begin{deptext}[column sep=1em]
-      tomodachi \& ni \& tegami-o \& kaita \\
- \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
-	   friend \& to \& letter  \& wrote \\
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   I \&	   wrote \& letters   \& to \& friends  \\
    \end{deptext}
 	%   \deproot{3}{ROOT}
-   \depedge{4}{2}{}
+   \depedge{2}{1}{nsubj}
+   \depedge{2}{3}{obj}
 	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
-   \depedge{2}{1}{}
-   \depedge{4}{3}{obj}
+   \depedge{2}{4}{}
+   \depedge{4}{5}{}
    %\depedge[arc angle=50]{7}{6}{ATT}
 \end{dependency}
+
+	\caption{Dependency trees}\label{fig:dep-trees}
 \end{figure}
 
-The sentences in Figure~\ref{fig:arabic-japanese} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words.
-For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} if the verb, rather than e.g.
+The sentences in Figure~\ref{fig:dep-trees} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words.
+For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} of the verb, rather than e.g.
 the subject or an indirect object.
@@
 
-In order to interpret a sentence correctly, a listener must be able to reconstruct the information expressed in the dependency tree and its labels.
+The dependency length of a sentence is the sum of the lengths of its syntactic dependencies \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
+\begin{equation}
+    \label{eq:depl}
+\operatorname{DepL}(t; {\bf w}) = \sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+\end{equation}
+where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost.
+
+All natural languages have some degree of word order regularities that are specified in their grammar.
+For instance, English places subjects before and objects after the verb (Figure~\ref{fig:dep-trees}).
+Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
+This is illustrated in Figure~\ref{fig:dlm-grammars}: Grammars specify how dependency structures are encoded into strings of words.
+For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
+Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
+Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
+
+We create optimized grammars that minimize the average dependency length.
+
+Answering these questions requires a sample of dependency structures as actually used by speakers of difefrent languages.
+Such samples have recently become available with the Universal Dependencies project, which has collected and created dependency annotations for several dozens of different languages.
+We use data from 51 languages.
+These corpora represent a typologically and genetically diverse group of languages.
+
+%See SI for our method for creating optimized grammars.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.7\textwidth]{figures/four-grammars-dlm.png}
+	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure.
+In this example, Grammar 1 achieves the lowest overall dependency length.
+Indeed, this is the Given a set of dependency structures, grammars define a set of sentences.
+For simplicity, we use English words.
+Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb.
+\textbf{TODO This figure is just a draft sketch so far.}}
+	\label{fig:dlm-grammars}
+\end{figure*}
+
+%In order to interpret a sentence correctly, a listener must be able to reconstruct the information expressed in the dependency tree and its labels.
 This is because, according to the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.
@@
 
+\paragraph{Results}
+Now present the correlations and the results.
+We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
+In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
+Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
+
+Dryer \cite{dryer1992greenbergian} presents three correlations that are only relevant to part of the world's languages and are not annotated reliably in UD.
+One further dependency is not annotated reliably in UD. %: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
+Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations from Dryer's 15 correlations.
+
+In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
+We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau$ show the same direction for the correlating dependency and for the verb-object dependency, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
+%We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau > 0.5$ holds for the correlating dependency depending on whether $a_\tau > 0.5$ for the verb-object dependency, with per-language-family random intercept and slope.\footnote{We coded language families according to \url{universaldependencies.org}.}
+We are interested in the direction and significance of this effect: 
+If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
+
+We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
+To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
+The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
+%in all except four languages, for WALS did not have data.
+
+\paragraph{Results}
+Results are shown in Table~\ref{table:results-dryer}.
+All eight correlations are confirmed in the models estimated from the real orderings. 
+We will discuss these exceptions further below.
+
+Confirming previous theoretical arguments, we see that optimizing for dependency length correctly accounts for all word order correlations.
+
 \section{Formalizing Efficiency in Word Order}
+
+We have verified that dependency length minimization produces Greenberg's correlation universals.
+However, dependency length is both heuristic and stipulative.
+The actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
+It is also stipulative: There is no a priori reason why dependency length should be the crucial determining factor determining word order.
+There are other models of processing cost that make different predictions.
+
+We show that the word order predictions of DLM arise from deeper principles of efficient communication.
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figures/grammars.png}
-	\caption{Given a set of dependency structures, grammars define a set of sentences.
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=0.45\textwidth]{figures/grammars.png}
+%	\caption{Given a set of dependency structures, grammars define a set of sentences.
 For simplicity, we use English words.
@@
 \textbf{TODO This figure is just a draft sketch so far.}}
-	\label{fig:grammars}
-\end{figure}
-
-Dependency Length Minimization is heuristic and stipulative.
+%	\label{fig:grammars}
+%\end{figure}
+%
+%
+%
 
@@
 Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
-Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
-This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
+These are encoded in the dependency structure.
+%Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
+%This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
 In Figure~\ref{fig:comm}, the message consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
@@
 
-All natural languages have some degree of word order regularities that are specified in their grammar.
+Speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences
 For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
-That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
-Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
-%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
-%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
-This is illustrated in Figure~\ref{fig:grammars}: Grammars specify how dependency structures are encoded into strings of words.
-For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
-Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
-Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
-%In a speech community, speakers and listeners have shared knowledge of these rules, which allows the listener to decode the dependency tree.
-%In the terminology of information theory, these rules specify a \emph{channel} through which the speaker transmits the dependency structures.
 
@@
 
-%For answering these questions,
-
-Answering these questions requires a sample of dependency structures as actually used by speakers of difefrent languages.
-Such samples have recently become available with the Universal Dependencies project, which has collected and created dependency annotations for several dozens of different languages.
-We use data from 51 languages.
-These corpora represent a typologically and genetically diverse group of languages.
+We use the same 51 corpora.
 
@@
 We now show that efficiency optimization predicts Greenberg's word order correlations. % in optimized grammars. 
-We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
-In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
-Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
-
-Dryer \cite{dryer1992greenbergian} presents three correlations that are only relevant to part of the world's languages and are not annotated reliably in UD.
-One further dependency is not annotated reliably in UD. %: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
-Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations from Dryer's 15 correlations.
-
-In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
-We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau$ show the same direction for the correlating dependency and for the verb-object dependency, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
-%We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau > 0.5$ holds for the correlating dependency depending on whether $a_\tau > 0.5$ for the verb-object dependency, with per-language-family random intercept and slope.\footnote{We coded language families according to \url{universaldependencies.org}.}
-We are interested in the direction and significance of this effect: 
-If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
-
-We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
-To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
-The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
-%in all except four languages, for WALS did not have data.
-
-\paragraph{Results}
-Results are shown in Table~\ref{table:results-dryer}.
-All eight correlations are confirmed in the models estimated from the real orderings. 
-We will discuss these exceptions further below.
-
-Confirming previous theoretical arguments, we see that optimizing for dependency length correctly accounts for all word order correlations. 
 Predictability and parseability predict four and six correlations, respectively, making complementary predictions.
@@
 \end{document}
+
+\begin{figure}
+\centering
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   katabt \& risāla \& li \& sadīq  \\
+	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+	   wrote \& letter   \& to \& friend  \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{1}{2}{obj}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{1}{3}{}
+   \depedge{3}{4}{}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+      tomodachi \& ni \& tegami-o \& kaita \\
+ \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+	   friend \& to \& letter  \& wrote \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{4}{2}{}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{2}{1}{}
+   \depedge{4}{3}{obj}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+\end{figure}
+
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   katabt \& risāla \& li \& sadīq  \\
+	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+	   wrote \& letter   \& to \& friend  \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{1}{2}{obj}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{1}{3}{}
+   \depedge{3}{4}{}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+      tomodachi \& ni \& tegami-o \& kaita \\
+ \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+	   friend \& to \& letter  \& wrote \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{4}{2}{}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{2}{1}{}
+   \depedge{4}{3}{obj}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}

### 2019-02-27 second-draft/4-paper-names.tex
revising

@@
 
-\section{Dependency Length Minimization}
-
-Some prior work has has suggested a functional explanation of Greenberg's correlations in terms of Dependency Length Minimization \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+\section{Dependency Length Minimization Predicts Greenberg's Correlations}
+
+Theoretical studies of word order typology have proposed Dependency Length Minimization as a unifying principle behind Greenberg's correlations \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
 This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
-On the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, it has been confirmed that languages reduce dependency length.
+On the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, it has been confirmed that languages reduce dependency length, though a direct link to Greenger's correlations has not been established beyond theoretical arguments.
 This has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.
 
-(CITE) have made theoretical arguments that dependency length minimization should produce word order correlations.
-Here, we present the first large-scale test of this idea.
+%(CITE) have made theoretical arguments that dependency length minimization should produce word order correlations.
+%Here, we present the first large-scale test of this idea.
+
+%It has been shown that languages reduce dependency length, compared to most randomly chosen orderings.
+%However, the link to correlation universals is not trivial:
 
@@
 
+Large-scale corpus evidence across languages shows that natural languages order words in such a way as to reduce this quantity, compared to most random orderings \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+However, the link to Greenberg's correlations has not been established this directly.
+While theoretical arguments suggest that DLM predicts Greenberg's correlations, this idea has not been tested on the basis of actual large-scale corpus data.
+
+We test this idea by constructing languages that have minimal dependency length:
+If Greenberg's correlations emerge in these languages, then this provides evidence that DLM gives rise to these universals.
 All natural languages have some degree of word order regularities that are specified in their grammar.
@@
 We create optimized grammars that minimize the average dependency length.
-
+Prior theoretical work on dependency length minimization has argued that certain general patterns minimize dependency length \cite{temperley2008dependency}, however, this depends on the topology of the dependency trees.
 Answering these questions requires a sample of dependency structures as actually used by speakers of difefrent languages.
@@
 These corpora represent a typologically and genetically diverse group of languages.
-
+Previously, \cite{temperley2008dependency, CITE} have created grammars optimized for DLM; ours is the first study applying this to study correlation universals on a large-scale basis.
+
+% (Frazier, 1985; Rijkhoff, 1990; Hawkins, 1994)
+%Frazier, L.
+(1985).
+Syntactic complexity.
+In D.
+R.
+Dowty, L.
+Karttunen & A.
+M.
+Zwicky(Eds),Natural Language Parsing: Psychological, Computational, and TheoreticalPerspectives(pp.
+129–189).
+Cambridge, UK: Cambridge University Press.
 %See SI for our method for creating optimized grammars.
@@
 	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
-	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+	verb & object     & & & &  & \\ 
+	\emph{visited} & \emph{friends} \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+
 %adp. &
-	adposition    &    NP       \\
-	\emph{to}            & \emph{a friend} \\
-copula    &    NP        &     \\
-	\emph{is}        & \emph{a friend} \\
+	adposition    &    NP       
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} \textbf{86}  }
+	&   \multirow{2}{*}{  \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf}   \textbf{47 \%}$^{}$   }
+	&   \multirow{2}{*}{  \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf} \textbf{68 \%}$^{***}$    } \\
+	\emph{to}            & \emph{a friend} \\ \hline
+copula    &    NP         
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf}}
+	&   \multirow{2}{*}{  \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf}   }
+	&   \multirow{2}{*}{  \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}}  \\
+	\emph{is}        & \emph{a friend} \\ \hline
 auxiliary    &    VP       &   \\
@@
 complementizer    &    S        & \\
-	\emph{that} & \emph{Mary} \\
+	\emph{that} & \emph{Mary has arrived} \\
 verb    &    PP     &    \\
@@
 \paragraph{Results}
-Now present the correlations and the results.
+We show that optimizing grammars for DLM produces Greenberg's correlations.
 We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
+Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
+
+Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages, or are not annotated reliably in UD.
+%One further dependency is not annotated reliably in UD. %: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
+Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations (See SI for information on the other correlations).
+% from Dryer's 15 correlations.
+These are described in Table~\ref{table:corr-dryer}.
 In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
-Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
-
-Dryer \cite{dryer1992greenbergian} presents three correlations that are only relevant to part of the world's languages and are not annotated reliably in UD.
-One further dependency is not annotated reliably in UD. %: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
-Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations from Dryer's 15 correlations.
-
-In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
-We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau$ show the same direction for the correlating dependency and for the verb-object dependency, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
+
+For each optimized grammar, we can check which correlations it satisfies.
+%In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
+We conducted, for each correlation, a mixed-effects logistic regression model predicting whether a grammar satisfies the correlation, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
 %We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau > 0.5$ holds for the correlating dependency depending on whether $a_\tau > 0.5$ for the verb-object dependency, with per-language-family random intercept and slope.\footnote{We coded language families according to \url{universaldependencies.org}.}
-We are interested in the direction and significance of this effect: 
-If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
+%We are interested in the direction and significance of this effect: 
+%If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
 
@@
 
-\paragraph{Results}
-Results are shown in Table~\ref{table:results-dryer}.
-All eight correlations are confirmed in the models estimated from the real orderings. 
-We will discuss these exceptions further below.
-
-Confirming previous theoretical arguments, we see that optimizing for dependency length correctly accounts for all word order correlations.
+%\paragraph{Results}
+%Results are shown in Table~\ref{table:results-dryer}.
+%All eight correlations are confirmed in the models estimated from the real orderings. 
+%We will discuss these exceptions further below.
+
+Optimizing for dependency length correctly accounts for all word order correlations.
 
@@
 We have verified that dependency length minimization produces Greenberg's correlation universals.
-However, dependency length is both heuristic and stipulative.
+However, dependency length is only a heuristic and stipulative measure of complexity: % heuristic and stipulative.
 The actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
@@
 There are other models of processing cost that make different predictions.
-
-We show that the word order predictions of DLM arise from deeper principles of efficient communication.
+DLM is also insufficient as an underlying principle of word order, as absolutely minimizing dependency length results in languages that have no consistent grammatical orders, making comprehension difficult.
+Indeed, theoretical work has justified DLM as describing an aspect of a broader notion of language efficiency.
+\cite{hawkins1994performance} theoetically argues that DLM enables more efficient parsing.
+\cite{futrell2017noisycontext} argues that DLM arises from the optimization of predictability.
 
@@
 %
+
+We show that the word order predictions of DLM arise from deeper principles of efficient communication.
 
@@
 
-\section{DLM TODO put these at the appropriate place}
-
-We compare Efficiency to the predictions of Dependency Length Minimization.
-This theory states that natural languages order information in such a way that the distances between syntactically linked words are minimized.
-For instance, in Figure~\ref{fig:arabic-japanese-simple}, there would be syntactic links between the adposition (to) and the noun (friend), and between the verb and both the object (letter) and the adposition (to).
-There is strong evidence that natural language minimizes the length of these dependencies~\cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
-Theoretical work in the functional linguistics literature has proposed that this principle explains the correlations.
-However, the principle itself is stipulative and not a first-principles explanation.
-\cite{hawkins1994performance} theoetically argues that it increases  parsability.
-
-However, these arguments have been made on a theoretical basis.
-We will show that (1) Dependency Length Minimization indeeds predicts most of the correlations, and (2) Dependency Length Minimization is explained by efficiency optimization.
-
-These computational results confirm theoretical ideas  that have been stated informally by authors in the functional linguistics literature at least since the 1980s.
-
-%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
+%
+%
+%\section{DLM TODO put these at the appropriate place}
+%
+%We compare Efficiency to the predictions of Dependency Length Minimization.
+%This theory states that natural languages order information in such a way that the distances between syntactically linked words are minimized.
+%For instance, in Figure~\ref{fig:arabic-japanese-simple}, there would be syntactic links between the adposition (to) and the noun (friend), and between the verb and both the object (letter) and the adposition (to).
+%There is strong evidence that natural language minimizes the length of these dependencies~\cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
+%Theoretical work in the functional linguistics literature has proposed that this principle explains the correlations.
+%However, the principle itself is stipulative and not a first-principles explanation.
+%\cite{hawkins1994performance} theoetically argues that it increases  parsability.
+%
+%However, these arguments have been made on a theoretical basis.
+%We will show that (1) Dependency Length Minimization indeeds predicts most of the correlations, and (2) Dependency Length Minimization is explained by efficiency optimization.
+%
+%These computational results confirm theoretical ideas  that have been stated informally by authors in the functional linguistics literature at least since the 1980s. 
+%
+%%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
 These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages.
 For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
+%
+%
 

### 2019-03-04 second-draft/3-paper-names.tex
revising

@@
 % Please include corresponding author, author contribution and author declaration information
-\authorcontributions{MH and RF designed research. MH implemented models and experiments. MH and RF wrote the paper. MH, DJ, and RF provided comments on the paper.}
+%\authorcontributions{MH and RF designed research. MH implemented models and experiments. MH and RF wrote the paper. MH, DJ, and RF provided comments on the paper.}
 \authordeclaration{The authors declare no conflict of interest.}
@@
 We show how the parameters of a word order grammar can be optimized for efficiency of information transfer, quantified as a tradeoff between incremental predictability and mutual information with latent tree structures.
-Applying these grammars to tree structures found in dependency corpora from 51 languages, we show that optimizing the grammar parameters for efficiency results in word order patterns that reproduce a large subset of the major word order correlations reported in the linguistic typological literature, and reproduce the predictions of previous heuristic theories such as dependency length minimization.
+Applying these grammars to tree structures found in dependency corpora from 51 languages, we show that optimizing the grammar parameters for efficiency results in word order patterns that predict a large subset of the major word order correlations reported in the linguistic typological literature, and subsumes the predictions of previous heuristic theories such as dependency length minimization.
 \end{abstract}
@@
 
-\textbf{TODO maybe should be more general in the beginning, as in the first version. Also make sure the reader doesn't think 'it's just headedness, why bother'.}
+%\textbf{TODO maybe should be more general in the beginning, as in the first version. Also make sure the reader doesn't think 'it's just headedness, why bother'.}
 
@@
 this is strongly correlated with the occurrence of pre- or postpositions:
+This is illustrated in Figure~\ref{fig:corr-table}.
 Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
 %However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
-This correlation has an extremely strong empirical basis:
+As Figure~\ref{fig:corr-table} shows, this correlation has an extremely strong empirical basis:
 In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization.
 %, over 90 \% of languages satisfy this generalization .
+
+\begin{figure}
+\centering
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   katabt \& risāla \& li \& sadīq  \\
+%	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+	   wrote \& letter   \& to \& friend  \\
+   \end{deptext}
+\end{dependency}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+      tomodachi \& ni \& tegami-o \& kaita \\
+ %\textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+	   friend \& to \& letter  \& wrote \\
+   \end{deptext}
+\end{dependency}
+
+    \centering
+    \includegraphics[width=0.4\textwidth]{figures/corr-table.png}
+
+%\begin{dependency}[theme=simple]
+%    \begin{deptext}
+%    I \& wrote \& a \& letter \& to \& a \& friend \\
+%    \end{deptext}
+%    \depedge{2}{1}{nsubj}
+%    \depedge{2}{4}{obj}
+%    \depedge{4}{3}{ }
+%    \depedge{4}{3}{ }
+%    \depedge{2}{5}{}
+%    \depedge{5}{7}{ }
+%    \depedge{7}{6}{ }
+%\end{dependency}
+	\caption{ The idea of a word order correlation: Languages can order the object before or after the verb (rows), and have prepositions or postpositions (columns).
+For each combination, we indicate how many languages satisfy it, as documented in the World Atlas of Language Structure~\cite{wals}.
+Combinations on the diagonal are vastly more common than off-diagonal ones.
+A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese-simple}	\label{fig:corr-table}
+\end{figure}
+
+This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
+These describe correlations between the relative positions of different types of expressions across languages.
+The example above documents that the position of the object (`letter') relative to the verb is \key{correlated} with the position of the adposition (`to').
+%adpositional phrases (`to a friend') relative to the verb.
+Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
+The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.
+%For instance, the relative order of objects and nouns also strongly determines the positions of genitives, auxiliaries, and other
+
+%These universals are supported by languages on all five continents, cutting across language families.
+
+%These cross-linguistic patterns call for an explanation, and 
+%and the problem of explaining the word order correlations has inspired a long tradition of research.
+Supported by languages on all continents, these correlations are among the language universals with the strongest empirical support.
+Importantly, their validity is also independent from specific assumptions about theories of grammar.
+
+Consequently, explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
+%
+%%supported by languages on all continents, and their validity is independent from 
+%
+%TODO say more about why these are extremely important, and that it is a puzzle why they are true
+%
+%%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
+%
+A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
+Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.
+
+Researchers working in the functional paradigm have proposed a range of criteria that languages should meet in order to enable efficient communication and processing.
+These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
+
+Theoretical studies of word order typology have proposed Dependency Length Minimization as a unifying principle behind Greenberg's correlations \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
+%Our first contribution is to show that this principle indeed correctly predicts Greenberg's correlations.
+%We do this by creating counterfactual grammars that minimize dependency length of 
+This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in language processing \cite{gibson1998linguistic,gibson2000dependency}.
+It has been argued theoretically that this principle should account for Greenberg's correlations, though this link is nontrivial and depends on the precise distributions of tree topologies in syntactic structures.
+
+%We have verified that dependency length minimization produces Greenberg's correlation universals.
+%However, dependency length is only a heuristic and stipulative measure of complexity: % heuristic and stipulative.
+%The actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
+%It is also stipulative: There is no a priori reason why dependency length should be the crucial determining factor determining word order.
+%While DLM has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, there are other models of processing cost that make different predictions.
+%In particular, surprisal is known to be a strong predictor of human processing effort, but it makes predictions different from and sometimes opposite to those made by dependency length.
+%DLM is also insufficient as an underlying principle of word order, as absolutely minimizing dependency length results in languages that have no consistent grammatical orders, making comprehension difficult.
+%Indeed, theoretical work has justified DLM as describing an aspect of a broader notion of language efficiency.
+%\cite{hawkins1994performance} theoetically argues that DLM enables more efficient parsing.
+%\cite{futrell2017noisycontext} argues that DLM arises from the optimization of predictability.
+
+%A commonly proposed functional explanation of Greenberg's correlations is in terms of Dependency Length Minimization \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+%This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
+%This has been confirmed on the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, and has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.
+
+%In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+%Dependency length has been proposed as a metric of processing difficulty on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, and minimization of dependency length is known to recover a large subset of the Greenbergian universals \cite{hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic}.
+%We demonstrate that the minimization of dependency length can be formalized in our framework, reproducing the finding that optimizing grammars for dependency length recovers several word order universals \cite{gildea2010grammars}.
+
+Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
+According to this theory, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
+Conversely, to minimize listener effort, the language should provide different utterances for different meanings, so that the listeenr can unambiguously identify the meaning from the utterance.
+Minimizing speaker and listener effort represent two opposing forces:
+A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
+A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
+%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
+Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
+The idea that language results from the tension between these two pressures has a long and fruitful history.
+%\cite{horn-toward-1984}, RSA, Color naming
+%
+%In the domain of word order, 
+%
+%TODO introduce the terms \key{cost} and \key{informativity}
+%
+This idea has been shown to provide a formalization and derivation of pragmatic inference \cite{horn-toward-1984, frank2012predicting}.
+It also has been shown to account for the distribution of word meanings across languages, such as color naming~\cite{zaslavsky2018efficient}.
+
+%TODO somewhere start calling this Efficiency
+%TODO explain why surprisal. some production/speaker work. (plan reuse)
+
+%In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
+%Even earlier than Zipf, \cite{gabelentz1901sprachwissenschaft} ...
+%This theory blabla Horn RSA
+%Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
+
+In this work, we show that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations, and for Dependency Length Minimization.
+
+%In this work, we show computationally, using corpus data from 51 languages, that efficiency optimization accounts for the Greenbergian word order correlations.
+That is, we show that the word orders in natural languages have evolved to optimize efficiency, and that this optimization accounts for the prevalence of the word order correlations.
+%word orders in natural language
+
+%\section{DLM}
+
+%
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=0.45\textwidth]{figures/letter-options.png}
+%	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure.
+In this example, Grammar 1 achieves the lowest overall dependency length.
+Indeed, this is the }
+%	\label{fig:pareto-plane}
+%\end{figure}
+%
+%
+
+\begin{table*}
+	\begin{center}
+\begin{tabular}{|ll|l|l|l|ll|l|llllllll}
+	\hline
+	%&
+	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
+	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%adp. &
+	adposition    &    NP       
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} \textbf{86}  }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}   \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf}   \textbf{47 \%}$^{}$   }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}  \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf} \textbf{68 \%}$^{***}$    } \\
+	\emph{to}            & \emph{a friend} \\ \hline
+copula    &    NP         
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} \textbf{94} }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}   \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf}  \textbf{53}     }
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}  \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf} \textbf{61}$^{***}$  }  \\
+	\emph{is}        & \emph{a friend} \\ \hline
+auxiliary    &    VP       
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}   \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}   }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}   \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}}  \\
+	\emph{has}          & \emph{written} \\ \hline
+noun    &    genitive      
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}}  \\
+	\emph{friend} &  \emph{of John} \\ \hline
+noun    &    relative clause      
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}\textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf} \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}}  \\
+	\emph{books} & \emph{that you read} \\ \hline
+complementizer    &    S        
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf} \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}   }
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf}\textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}}  \\
+	\emph{that} & \emph{Mary has arrived} \\ \hline
+verb    &    PP         
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf} \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}   }
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf}\textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}}  \\
+	\emph{went} & \emph{to school} \\ \hline
+want    &    VP        
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}   }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}}  \\
+	\emph{wants}   &  \emph{to leave} \\ \hline
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+\end{tabular}
+	\end{center}
+	\caption{Greenbergian Correlations.
+Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+For each correlation, we provide an example.
+Given the statistical nature of the correlations, not every language satisfies every one of them.
+Not all correlations are satisfied by every natu
+}\label{table:corr-dryer}
+\end{table*}
+
+\section{Formalizing Efficiency in Word Order}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figures/communication.png}
+	\caption{Our model of communication: A speaker (left) expresses a dependency structure into a string of words forming a sentence.
+A listener probabilistically recovers a dependency structure.
+In this example, the grammar of English allows the listener to recover the correct dependency structure with very high confidence.}
+	\label{fig:comm}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figures/grammars.png}
+	\caption{Given a set of dependency structures, grammars define a set of sentences.
+For simplicity, we use English words.
+Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb.
+\textbf{TODO This figure is just a draft sketch so far.}}
+	\label{fig:grammars}
+\end{figure}
+
+We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
+In Shannon's model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
+Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
+Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
+This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
+In Figure~\ref{fig:comm}, the message consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
+
+When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
+The listener receives this string of words.
+By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
+The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
+Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
+Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
+If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).
+
+%We can view this as a formalization of the Language of Thought (CITE) TODO explain why
+
+%We formalize the listener's understanding as probabilistic inference, that is, the listener
+
+All natural languages have some degree of word order regularities that are specified in their grammar.
+For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
+That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
+Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
+%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
+%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
+This is illustrated in Figure~\ref{fig:grammars}: Grammars specify how dependency structures are encoded into strings of words.
+For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
+Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
+Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
+%In a speech community, speakers and listeners have shared knowledge of these rules, which allows the listener to decode the dependency tree.
+%In the terminology of information theory, these rules specify a \emph{channel} through which the speaker transmits the dependency structures.
+We formalize word order grammars as specifying how different types of dependents (subjects, objects, etc.) are ordered, employing a cross-linguistically meaningful inventory of relations (CITE UD).
+
+In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes.
+In our model, these will depend on the grammar:
+Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance (TODO can we find an example for this?)
+This reflects the \key{parseability} of the language.
+
+The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
+\begin{equation}
+	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+\end{equation}
+where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
+This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
+It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
+This quantity formalizes Zipf's principle of listener effort and his Force of Diversification:
+It can be maximized when trees can be decoded fully unambiguously from utterances.
+Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's principle of listener economy~\cite{frank2012predicting,regier2015word}.
+
+%The counteracting Force of Unification is formalized by the cost of encoding the dependency structure and producing the utterance.
+%TODO motivate surprisal here.
+
+%Note that, in order for this process to be successful, speakers and listeners need to have shared knowledge of the rules by which dependency structures are encoded into sentences.
+%We hypothesize that the rules found in actual natural languages are optimized for the efficiency of this process, and that this optimization accounts for the observed word order universals.
+
+%If speakers choose to order their sentences in such a way that efficiency is optimized, will this lead to the observed language universals?
+
+%Do languages optimize their word order for efficiency?
+%Does this optimization explain the observed language universals?
+
+%We formalize a language as a distribution over dependency structures $T$ together with an ordering grammar $\theta$ that specifies how dependency structures are ordered into strings of words.
+%Applying the ordering grammar to the dependency structures, a language defines a probability distribution over sentences $L$.
+%Efficiency is defined as a tradeoff between
+
+% based on related work in information theory \cite{cover2006elements,tishby1999information}, bounded rationality \cite{genewein2015bounded}, and computational neuroscience \cite{friston2009predictive}; (ii)
+%
+%We formalize the notion of communicative efficiency using two terms: one having to do with minimizing the difficulty of linguistic communication, and one having to do with maximizing the information transferred in such communication.
+We quantify the difficulty of linguistic communication in terms of the word-by-word \key{predictability} of strings, which has been found to be a highly accurate and general predictor of human online comprehension difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect},
+% in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+We quantify the information transferred in language as \key{parseability}: the quantity of information linguistic utterances contain about latent hierarchical tree structures.
+We also compare the results of optimizing for communicative efficiency to the results of optimizing for dependency length minimization, a previously proposed approximate metric of the amount of memory involved in language production and comprehension \cite{hudson1995measuring,gibson1998linguistic,gibson2000dependency,liu2017dependency,temperley2018minimizing}.
+
+We formalize the cost of an utterance as its word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
+Surprisal has been found to be is a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
+%The \key{predictability} of the language is the negative entropy of the sentences:
+\begin{equation}
+	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+\end{equation}
+where the sum runs over all possible sentences $l$ that belong to the language.
+In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
+%how many distinct sentences there are.
+
+%It is larger
+
+%
+%The opposing pressure is \key{parsability}, which is the mutual information beteen sentences and dependency structures:
+%\begin{equation}
+%	R_{Pars} := I[L,T] = \sum_{t,l} p(t,l) \log \frac{p(t|l)}{p(t)}
+%\end{equation}
+%where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
+%This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
+%It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
+
+The \key{efficiency} of a language is a weighted combination
+\begin{equation}\label{eq:efficiency}
+	R_{Eff} := R_{Pars} + \lambda R_{Pred}
+\end{equation}
+with an interpolation weight $\lambda \in [0,1)$. In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification).
+
+%This combined equation represents efficiency: bits of information transferred about dependency trees minus the cost of the transmission.
+In expectation over utterances (substituting Eq.~\ref{eq:combined-score} into Eq.~\ref{eq:exp-over-utts}), the overall objective function to be maximized by a word order grammar can be written in information-theoretic terms as:
+
+%The two terms of Eq.~\ref{eq:efficiency} implement two opposing pressures.
+Parseability (the mutual information term) pushes to make strings more heterogeneous, so that they can indicate different tree structures unambiguously.
+Predictability (the entropy term) pushes to make strings more homogeneous, to reduce the complexity of using the language.
+%The idea that natural language arises from a tension between these two forces goes back over a century \cite{gabelentz1901sprachwissenschaft}.
+Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking dependency structures as the underlying meanings to be conveyed.
+Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
+
+%\subsection{Optimizing Grammars for Efficiency}
+
+Our goals are to show that
+(1) the grammars of natural languages evolve towards optimizing efficiency of communication, and
+(2) this process of optimization accounts for Greenberg's order correlation universals.
+
+%For answering these questions,
+
+Answering these questions requires a sample of dependency structures as actually used by speakers of difefrent languages.
+Such samples have recently become available with the Universal Dependencies project, which has collected and created dependency annotations for several dozens of different languages.
+We use data from 51 languages.
+These corpora represent a typologically and genetically diverse group of languages.
+
+To answer whether languages evolve to have efficient word order, we compare the efficiency of the actual grammars of these 51 languages to randomly constructed baseline grammars.
+To show that this process of efficiency optimization accounts for Greenberg's correlation universals, we computationally construct grammars that optimize efficiency.
+We then show that these grammars mostly exhibit the Greenbergian order correlation universals.
+We furthermore show that this process of optimization also \emph{explains} DLM, providing a first-principles explanation of this heuristic generalization.
+
+For every one of the 51 corpora, we computationally construct eight optimal grammars.
+Note that the original orders of the actual languages do not enter this objective.
+See SI for our method for creating optimized grammars.
+
+%
+%All natural languages have some degree of word order regularities that are specified in their grammar.
+%It is thus not realistic to assume that speakers optimize each utterance individually.
+%Instead, we hypothesize that the word order rules of languages as a whole are optimized for efficiency, and that this optimization produces the observed universals.
+%
+
+Discuss DLM now again
+
+%\subsection{Dependency Length Minimization}
+
+\section{Results}
+
+\subsection{Relative efficiency of languages}
+\label{sec:relative-efficiency}
+
+We first demonstrate that real languages are relatively efficient compared to random baselines. For each language, we generated ten baseline word order grammars for the language.
+These grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}).
+%by choosing all word order grammar parameters randomly at uniform from $[0,1]$.
+Figure~\ref{fig:pareto-plane} shows the predictability and parseability for each real language relative to its baseline grammars.
+In order to control for limitations due to our word order grammar formalism, we represent real languages in the figure by maximum likelihood fits of word order grammars to the real language data.
+%; in general these maximum-likelihood word order grammars end up with lower predictability and parseability than the real languages. % TODO verify this is true? TODO awkward sentence
+For the calculation of predictability and parseability, we make all (baseline and real) word order grammars deterministic by always choosing the highest-probability linearization of each tree; by making the grammars deterministic in this way we eliminate an anticonservative bias toward low predictability in the baseline languages, which are highly nondeterministic.
+The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend.pdf}
+    \caption{Predictability and parseability of 51 UD languages (red), indicated by ISO codes, compared to ten baseline word order grammars per language (green).
+Predictability and parseability scores are $z$-scored within language.
+Each point for a real language has a line pointing in the direction of the center of mass of its baselines.
+The green contour shows the density of baseline languages.
+Unlabeled dots represent the centroid for real languages (red), baseline languages (green), and languages optimized for predictability (yellow), parseability (pink), efficiency (blue), and dependency length (red).
+When a language is to the bottom--left of its baselines, this indicates that it is relatively optimal for efficiency.} % TODO colors, fontsize, etc could conceivably be made prettier
+    \label{fig:pareto-plane}
+\end{figure}
+
+Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
+Languages appear to be attracted toward these points and away from the region of the baseline languages.
+We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
+% TODO add something about dependency length-optimized languages?
+
+%These results are shown in more detail for two typologically distinct languages (English and Japanese) in Figure~\ref{fig:predict-pars}.
+Here we include also the predictability and parseability of the original language data and of the word order grammars optimized for dependency length.
+We see that real languages are even more predictable than any optimized or baseline language, while optimization for either parseability or dependency length suffices to raise parseability to the level of the real languages.
+This result is in line with the finding that languages with longer dependencies are harder to parse in general \cite{gulordava2016multilingual}.
+
+%\begin{figure*}
+%\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
+%	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right).
+We provide the entropy $H[L_\theta(T)]$ for predictability (lower is better) and the conditional entropy $H[T|L_\theta(T)]$ for parseability (lower is better).
+Both measures are normalized by sentence length.}\label{fig:predict-pars}
+%\end{figure*}
+%
+
+\subsection{Efficiency Explains DLM}
+We demonstrate the relationship between dependency length minimization and the maximization of efficiency.
+Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency.
+We see that optimizing for efficiency lowers dependency length relative to random baselines, in keeping with the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}.
+In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the dependency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.
+
+\begin{figure*} % need Efficiency and (maybe) MLE langs on here
+\includegraphics[width=0.98\textwidth]{../results/dependency-length/figures/depLength-facet.pdf}
+\caption{Average dependency length as a function of sentence length in four languages.
+Across languages, real and optimized languages have shorter dependencies than random baseline orderings.}\label{fig:deplength}
+\end{figure*}
+
+\subsection{Greenbergian Word Order Correlations}
+\label{sec:greenberg}
+
+%\begin{table*}
+%	\begin{center}
+%\begin{tabular}{|ll|l|l|l|ll|l|}
+%	\hline
+%	%&
+%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
+%	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%%adp. &
+%adposition    &    NP       &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
+%copula    &    NP        &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
+%auxiliary    &    VP       &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
+%noun    &    genitive      &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
+%noun    &    relative clause       &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
+%complementizer    &    S        &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
+%verb    &    PP     &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
+%want    &    VP      &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
+%verb    &    subject    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
+%verb    &    manner adverb     &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+%\end{tabular}
+%	\end{center}
+%	\caption{Greenbergian Correlations.
+Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%%	For clarity, we organize the correlations by the category of the `verb patterner'.
+%	For each correlation, we give our operationalization in terms of UD.
+%A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
+%	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
+%	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
+%\end{table*}
+%
+
+We now show that efficiency optimization predicts Greenberg's word order correlations. % in optimized grammars. 
+We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
+In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
+Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
+%This dependency corresponds to dependencies of the form $\xrightarrow{obj}$ in UD.
+%Similarly, most of the other dependencies in Dryer's formulation can be formalized:
+%for instance, the dependency between nouns and relative clauses can be formalized as $\xrightarrow{acl}$.
+%Testing whether an ordering model satisfies the correlation between relative clauses and objects reduces to checking whether these two relation types have direction parameters $a_\tau$ that are either both $>0.5$ (both kinds of dependents precede their heads) or both $< 0.5$ (they both follow their heads).
+
+Dryer \cite{dryer1992greenbergian} presents three correlations that are only relevant to part of the world's languages and are not annotated reliably in UD.
+%: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
+Further collapsing correlations that cannot be formalized individually in UD, we otained 10 formalized correlations from Dryer's 15 correlations.
+%Two pairs of Dryer's correlations, namely those for the dependencies between complementizers and adverbial subordinators and their complement clauses,and those for the dependencies between verbs and adpositional phrases, and adjectives and their standard of comparison, had to be collapsed into two correlations in UD.
+From Dryer's 15 correlations, we obtained 10 formalized correlations, roughly covering nine of Greenberg's original universals.
+%TODO update to 10?
+What happened to the 11th?
+Was that amod?
+[The 11th was adjective+standard of comparison, which is merged into obl] Maybe we could include amod as a "true rejection" of efficiency, but a false alarm for DepL
+
+In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
+We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau$ show the same direction for the correlating dependency and for the verb-object dependency, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
+%We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau > 0.5$ holds for the correlating dependency depending on whether $a_\tau > 0.5$ for the verb-object dependency, with per-language-family random intercept and slope.\footnote{We coded language families according to \url{universaldependencies.org}.}
+We are interested in the direction and significance of this effect: 
+If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
+
+We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
+To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
+The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
+%in all except four languages, for WALS did not have data.
+
+\paragraph{Results}
+Results are shown in Table~\ref{table:results-dryer}.
+All correlations but two are confirmed in the models estimated from the real orderings.
+The exceptions are the subject--verb dependency and the verb--adverb dependency, which typically go in the opposite direction from the standard description.
+We will discuss these exceptions further below.
+
+In keeping with previous work, we see that optimizing for dependency length correctly accounts for nine word order correlations, missing only the verb--adverb dependency. 
+Predictability and parseability predict five and seven correlations, respectively, making largely complementary predictions.
+Efficiency significantly predicts all the word order correlations, each in the same direction as attested in the dependency corpora.
+
+We now address the two word order correlations whose direction in the dependency corpora is opposite from what would be expected in the typological literature.
+The first is the correlation of the order of verb--subject and verb--object dependencies.
+Our sample of mainly European languages highly over-represents languages with the general order subject--verb--object (such as English), in which the order of the verb--subject and verb--object dependencies are anti-correlated.
+Surprisingly, given the sample of tree structures of these languages, it turns out that the optimal languages tend to have anti-correlated orders for subjects and objects order similar to the real languages.
+
+The second anomalous dependency is the verb--manner adverb dependency.
+We believe the anti-correlation in the UD corpora arises because the \emph{advmod} dependency does not distinguish between manner adverbs---the subject of the typological judgment---and various other types of modifiers such as sentence-level adverbs.
+Nevertheless, the languages optimized for efficiency reproduce the anti-correlation of the orders of verb--object and verb--adverb at around the same rate as the real languages.
+
+We further evaluate the word order predictions of efficiency, showing that efficiency is most successful in predicting correlations in the direction found in the UD corpora.
+We constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency in languages optimized for efficiency, with random effects for language and language family, correlated across the ten dependencies.
+We conducted the same analysis for predictability, parseability, and dependency length.
+We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
+The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
+The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$.
+The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.
+For dependency length, the posterior puts much of the probability mass on predicting only nine of the correlations; predictability and parseability predict significantly less correlations.
+
+\subsection{Discussion}
+
+In Section~\ref{sec:relative-efficiency}, we found that word order grammars in the majority of UD languages have better efficiency than baseline word order grammars.
+Furthermore, in Section~\ref{sec:greenberg} we found that explicitly optimizing grammars for efficiency reproduces 10 major word order correlations reported in the typological literature, and with greater accuracy than the typological literature when it comes to the languages studied.
+
+%While the efficiency objective succeeds in predicting the presence of the most word order correlations, the dependency length objective seems to predict a greater prevalence for those correlations that it does predict.
+We believe this discrepancy is a result of noise in the optimization process.
+Optimizing for efficiency is computationally challenging, requiring stochastic gradient descent on word order grammar, language model, and parser parameters, while optimizing for dependency length is simpler and thus more optimal grammars can be reached easily (See SI section 2 for supporting data).
+Inasmuch as dependency length minimization is computationally easier to implement than efficiency maximization, we believe it will remain useful as an easy-to-use heuristic for predicting and explaining word order universals.
+% TODO we can also report (in the SI) prevalences among the best-performing of the optimized languages, i.e., those where optimization was most successful.
+I would expect that this should narrow the gap between DepL and Efficiency, which would confirm that the gap is really about the difficulty of optimization, not about the objective functions.
+
+\section{Conclusion}
+
+We found that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
+We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals.
+Beyond our present results, we provide a complete formalization and computational framework in which theories of the functional optimization of languages can be tested.
+Other objective functions could be hypothesized and tested within our framework; furthermore, future advances in machine learning will enable the optimization of richer and richer models of grammar.
+
+A major question for functional explanations for linguistic universals is: \emph{how} do languages come to be optimized?
+Do speakers actively seek out new communicative conventions that allow better efficiency?
+Or do languages change in response to biases that come into play during language acquisition \cite{fedzechkina2012language,culbertson2012learning}?
+Our work is neutral toward such questions.
+To the extent that language universals arise from biases in learning or in the representational capacity of the human brain, our results suggest that those biases tilt toward communicative efficiency.
+Our work does provide against the idea that word order universals are best explained in terms of learning biases that are irreducibly arbitrary and genetic in nature \cite{chomsky2010some}.
+
+While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into dependency trees.
+The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication.
+Relatedly, there are many remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
+Our models cannot capture this trade-off for technical reasons: the parseability objective does not operate over wordforms, only POS tags, and so it does not take advantage of morphological cues to syntactic structure.
+Future work can investigate whether these and other remaining universals can be explained using more sophisticated models of how meaning representations are transduced into strings of words, based on larger databases with richer annotations.
+
+\section{DLM TODO put these at the appropriate place}
+
+We compare Efficiency to the predictions of Dependency Length Minimization.
+This theory states that natural languages order information in such a way that the distances between syntactically linked words are minimized.
+For instance, in Figure~\ref{fig:arabic-japanese-simple}, there would be syntactic links between the adposition (to) and the noun (friend), and between the verb and both the object (letter) and the adposition (to).
+There is strong evidence that natural language minimizes the length of these dependencies~\cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
+Theoretical work in the functional linguistics literature has proposed that this principle explains the correlations.
+However, the principle itself is stipulative and not a first-principles explanation.
+\cite{hawkins1994performance} theoetically argues that it increases  parsability.
+
+However, these arguments have been made on a theoretical basis.
+We will show that (1) Dependency Length Minimization indeeds predicts most of the correlations, and (2) Dependency Length Minimization is explained by efficiency optimization.
+
+These computational results confirm theoretical ideas  that have been stated informally by authors in the functional linguistics literature at least since the 1980s.
+
+%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
+These universals are called \key{implicational} universals, and take the form of correlations between certain word order patterns and others across languages.
+For example, whether a language puts verbs before or after objects strongly determines whether it places genitives before or after nouns, and whether it has prepositions or postpositions \cite{dryer1992greenbergian}.
+
+\matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
+We use all languages for which at least one treebank with a training partition was available, a total of 50 languages.
+For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.
+Punctuation was removed.
+%We removed punctuation. %, which we operationalized by removing all dependents of \emph{punct} dependencies, and recursively removing their dependents.
+
+Universal dependencies represents as dependents some words that are typically classified as heads in syntactic theory.
+This particularly applies to the \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark} dependencies.
+Following prior work studying dependency length minimization \cite{futrell2015largescale}, we modified each treebank by inverting these dependencies, promoting the dependent to the head position.
+We report results on this modified version of UD.
+
+The efficiency optimization results from Table~\ref{tab:results-dryer} were preregistered: \url{http://aspredicted.org/blind.php?x=8gp2bt}.
+
+See the SI for details on the neural language models and parsers used, and on the optimization procedures.}
+
+\showmatmethods{} % Display the Materials and Methods section
+
+\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Chris Manning, and audiences at CAMP 2018 for helpful discussion.}
+
+\showacknow{} % Display the acknowledgments section
+
+%\bibliographystyle{acl_natbib}
+%\bibliography{extracted}
+\bibliography{everything}
+
+\end{document}
+
+%the meaning
 
@@
 
-This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
-These describe correlations between the relative positions of different types of expressions across languages.
-The example above documents that the position of the object (`letter') relative to the verb is \key{correlated} with the position of the adposition (`to').
-%adpositional phrases (`to a friend') relative to the verb.
-Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
-The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.
-%For instance, the relative order of objects and nouns also strongly determines the positions of genitives, auxiliaries, and other
-
-%These universals are supported by languages on all five continents, cutting across language families.
-
-%These cross-linguistic patterns call for an explanation, and 
-%and the problem of explaining the word order correlations has inspired a long tradition of research.
-Supported by languages on all continents, these correlations are among the language universals with the strongest empirical support.
-Importantly, their validity is also independent from specific assumptions about theories of grammar.
-
-Consequently, explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
-%
-%%supported by languages on all continents, and their validity is independent from 
-%
-%TODO say more about why these are extremely important, and that it is a puzzle why they are true
-%
-%%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-%
-A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
-Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.
-
-Researchers working in the functional paradigm have proposed a range of criteria that languages should meet in order to enable efficient communication and processing.
-These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
-
-A commonly proposed functional explanation of Greenberg's correlations is in terms of Dependency Length Minimization \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
-This has been confirmed on the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, and has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.
-
-%In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+\section{Minimizing Dependency Length Predicts Universals}
+
+..
+
+\begin{figure}
+\centering
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   katabt \& risāla \& li \& sadīq  \\
+	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+	   wrote \& letter   \& to \& friend  \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{1}{2}{obj}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{1}{3}{}
+   \depedge{3}{4}{}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+      tomodachi \& ni \& tegami-o \& kaita \\
+ \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+	   friend \& to \& letter  \& wrote \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{4}{2}{}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{2}{1}{}
+   \depedge{4}{3}{obj}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+\end{figure}
+
+Dependency Length:
+
+The sentences in Figure~\ref{fig:arabic-japanese} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words.
+For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} if the verb, rather than e.g.
+the subject or an indirect object.
+A dependency arc is drawn from a \key{head} (e.g.
+\emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g.
+\emph{tegami-o}).
+Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}.
+Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.
+
+In order to interpret a sentence correctly, a listener must be able to reconstruct the information expressed in the dependency tree and its labels.
+This is because, according to the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.
+
+The contribution of this work is (i) to provide a principled formalization of the notion of communicative efficiency in word order, based on related work in information theory \cite{cover2006elements,tishby1999information}, bounded rationality \cite{genewein2015bounded}, and computational neuroscience \cite{friston2009predictive}; (ii) to introduce a computational framework in which we can produce simulated grammars directly optimized for our notion of efficiency, using state-of-the-art machine learning techniques \cite{hochreiter1997long,goldberg2017neural,dozat2017stanford}; and (iii) to demonstrate that in doing so, we reproduce the major word order patterns reported in \cite{greenberg1963universals} and related work \cite{dryer1992greenbergian}, as well as reproducing the predictions of previous heuristic theories of word order optimization \cite{hawkins1994performance,ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
+
+We formalize the notion of communicative efficiency using two terms: one having to do with minimizing the difficulty of linguistic communication, and one having to do with maximizing the information transferred in such communication.
+We quantify the difficulty of linguistic communication in terms of the word-by-word \key{predictability} of strings, which has been found to be a highly accurate and general predictor of human online comprehension difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and which can be justified in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+We quantify the information transferred in language as \key{parseability}: the quantity of information linguistic utterances contain about latent hierarchical tree structures.
+We also compare the results of optimizing for communicative efficiency to the results of optimizing for dependency length minimization, a previously proposed approximate metric of the amount of memory involved in language production and comprehension \cite{hudson1995measuring,gibson1998linguistic,gibson2000dependency,liu2017dependency,temperley2018minimizing}.
+
+To evaluate whether communicative efficiency can explain word order universals, we apply the following method \cite{gildea2007optimizing}.
+First, we develop word order grammars: simple, interpretable models of word order, implemented as functions that take in an unordered syntactic dependency tree and output a distribution over possible word orders for that tree.
+Second, we take dependency treebanks of natural languages and compute word order grammars which are optimized for an objective functions representing communicative efficiency.
+Third, we evaluate whether the optimized word order grammars reproduce the desired word order universals when applied to attested dependency trees from the treebanks across many languages.
+Our favorable results suggest that it may be possible to generally model human languages as efficient codes for communication among agents with generic human-like cognitive and communicative constraints.
+%The best way to describe a human language might be as a solution to an equation involving such constraints.
+
+%We draw on data from the Universal Dependencies project \cite{nivre-universal-2017}, which has been developing treebanks in a unified format for several dozens of diverse languages.
+%This universal annotation scheme is particularly useful for studying word order universals, as many of them can be operationalized  in terms of crosslinguistically valid syntactic notions.
+% such as \emph{verbs} and \emph{objects}.
+
+\section{Background}
+
+% TODO where do we say that we're basing our work on data from 51 languages?
+
+%\subsection{Explananda: Word Order Correlations}
+%
+%Working with a database of 30 languages, Greenberg \cite{greenberg1963universals} proposed 45 universals of language that he determined to be true of most or all languages.
+%31 of them were concerned with word order, the remainder with morphology.
+%Many of the word order universals state \emph{correlations} between patterns:
+%For instance, Universals 3 and 4 state that ``\emph{Languages with dominant VSO order are always prepositional.}''  (e.g.
+Arabic), and ``\emph{With overwhelmingly greater than chance frequency, languages with normal SOV order are postpositional.}'' (e.g.
+Japanese).
+%These are illustrated in Figure~\ref{fig:arabic-japanese}:
+%Arabic has basic order verb-object (VO) and prepositions (\emph{li} `to'), while Japanese has order object-verb (OV) and postpositions (\emph{ni} `to').
+Our goal is to explain Greenberg's reported word order correlations, as updated in \cite{dryer1992greenbergian}.
+
+\begin{figure}
+\centering
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+	   katabt \& risāla \& li \& sadīq  \\
+	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+	   wrote \& letter   \& to \& friend  \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{1}{2}{obj}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{1}{3}{}
+   \depedge{3}{4}{}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+      tomodachi \& ni \& tegami-o \& kaita \\
+ \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+	   friend \& to \& letter  \& wrote \\
+   \end{deptext}
+	%   \deproot{3}{ROOT}
+   \depedge{4}{2}{}
+	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
+   \depedge{2}{1}{}
+   \depedge{4}{3}{obj}
+   %\depedge[arc angle=50]{7}{6}{ATT}
+\end{dependency}
+
+%\begin{dependency}[theme=simple]
+%    \begin{deptext}
+%    I \& wrote \& a \& letter \& to \& a \& friend \\
+%    \end{deptext}
+%    \depedge{2}{1}{nsubj}
+%    \depedge{2}{4}{obj}
+%    \depedge{4}{3}{ }
+%    \depedge{4}{3}{ }
+%    \depedge{2}{5}{}
+%    \depedge{5}{7}{ }
+%    \depedge{7}{6}{ }
+%\end{dependency}
+	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese}
+\end{figure}
+
+% TODO what to do about these labels...? They're not UD...
+
+\subsection{Dependency trees}
+
+\section{Optimizing word order grammars}
+
+Here we describe our model of word order and how we derive optimal grammars within that model in terms of communicative efficiency.
+
+Let $t$ be an unordered dependency tree, i.e.
+a structure as in Figure~\ref{fig:arabic-japanese} but without any specified ordering of the words.
+We think of an unordered dependency tree $t$ as representing the compositional meaning to be conveyed by a sentence.
+We define a \key{word order grammar} as a function $L_\theta(t)$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output a probability distribution over ordered sequences of words $\mathbf{w}$ linearizing the tree, or equivalently, a probability distribution over orders on the nodes of the tree.
+% do we need to define "linearize"?
+This process is visualized in Figure~\ref{fig:process-visualization}.
+
+Next we define how a word order grammar linearizes trees in terms of its parameters $\theta$, and then we describe how $\theta$ can be set to optimize objective functions related to communicative efficiency.
+
+\begin{figure}
+\includegraphics[width=0.48\textwidth]{figures/process-visualization.png}
+\caption{Our method for constructing counterfactual versions of languages: We extract unordered dependency tree topologies from dependency corpora, and apply word order grammars to obtain counterfactually reordered corpora.}\label{fig:process-visualization}
+\end{figure}
+
+We think of it as the following problem:
+
+A speaker has a tree in mind, and they have to produce an utterance.
+
+%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a word order grammar.
+Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
+%\begin{equation}
+%\label{eq:exp-over-utts}
+%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim L_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
+%\end{equation}
+%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
+In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
+%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
+%
+%
+
+All natural languages have some degree of word order regularity.
+Thus, it is not sufficient to optimize the word orders of individual sentences in the corpora---instead, we will optimize the word order rules of entire languages.
+We develop a method to create word order rules that optimize given processing measures on the resulting counterfactual corpus.
+
+\subsection{DLM Predicts Universals}
+
+We first apply our method to show that DLM predicts the correlations.
+
 %Dependency length has been proposed as a metric of processing difficulty on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, and minimization of dependency length is known to recover a large subset of the Greenbergian universals \cite{hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic}.
 %We demonstrate that the minimization of dependency length can be formalized in our framework, reproducing the finding that optimizing grammars for dependency length recovers several word order universals \cite{gildea2010grammars}.
-
-Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
-According to this theory, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
-Conversely, to minimize listener effort, the language should provide different utterances for different meanings, so that the listeenr can unambiguously identify the meaning from the utterance.
-Minimizing speaker and listener effort represent two opposing forces:
-A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
-A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
-%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
-Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
-The idea that language results from the tension between these two pressures has a long and fruitful history.
-%\cite{horn-toward-1984}, RSA, Color naming
-%
-%In the domain of word order, 
-%
-%TODO introduce the terms \key{cost} and \key{informativity}
-%
-This idea has been shown to account for phenomena such as pragmatic inference \cite{horn-toward-1984, frank2012predicting} and color naming~\cite{zaslavsky2018efficient}.
-
-%TODO somewhere start calling this Efficiency
-%TODO explain why surprisal. some production/speaker work. (plan reuse)
-
-%In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
-%Even earlier than Zipf, \cite{gabelentz1901sprachwissenschaft} ...
-%This theory blabla Horn RSA
-%Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
-
-In this work, we show that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
-
-%In this work, we show computationally, using corpus data from 51 languages, that efficiency optimization accounts for the Greenbergian word order correlations.
-That is, we show that the word orders in natural languages have evolved to optimize efficiency, and that this optimization accounts for the prevalence of the word order correlations.
-%word orders in natural language
-
-%\section{DLM}
-
-%
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.45\textwidth]{figures/letter-options.png}
-%	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure.
-In this example, Grammar 1 achieves the lowest overall dependency length.
-Indeed, this is the }
-%	\label{fig:pareto-plane}
-%\end{figure}
-%
-%
-
-\begin{table*}
-	\begin{center}
-\begin{tabular}{|ll|l|l|l|ll|l|llllllll}
-	\hline
-	%&
-	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
-	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%adp. &
-	adposition    &    NP       \\
-	\emph{to}            & \emph{a friend} \\
-copula    &    NP        &     \\
-	\emph{is}        & \emph{a friend} \\
-auxiliary    &    VP       &   \\
-	\emph{has}          & \emph{written} \\
-noun    &    genitive      &  \\
-	\emph{friend} &  \emph{of John} \\
-noun    &    relative clause       &  \\
-	\emph{books} & \emph{that you read} \\
-complementizer    &    S        & \\
-	\emph{that} & \emph{Mary} \\
-verb    &    PP     &    \\
-	\emph{went} & \emph{to school} \\
-want    &    VP      &    \\
-	\emph{wants}   &  \emph{to leave} \\
-verb    &    subject    &  \\
-	\emph{(there) entered} & \emph{a person} \\
-verb    &    manner adverb     & \\
-	\emph{ran} & \emph{quickly} \\
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
-\end{tabular}
-	\end{center}
-	\caption{Greenbergian Correlations.
-Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-For each correlation, we provide an example.
-Given the statistical nature of the correlations, not every language satisfies every one of them.
-Not all correlations are satisfied by every natu
-}\label{table:corr-dryer}
-\end{table*}
-
-\section{Formalizing Efficiency in Word Order}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figures/communication.png}
-	\caption{Our model of communication: A speaker (left) expresses a dependency structure into a string of words forming a sentence.
-A listener probabilistically recovers a dependency structure.
-In this example, the grammar of English allows the listener to recover the correct dependency structure with very high confidence.}
-	\label{fig:comm}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figures/grammars.png}
-	\caption{Given a set of dependency structures, grammars define a set of sentences.
-For simplicity, we use English words.
-Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb.
-\textbf{TODO This figure is just a draft sketch so far.}}
-	\label{fig:grammars}
-\end{figure}
-
-We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
-In Shannon's model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
-Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
-Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
-This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
-In Figure~\ref{fig:comm}, the message consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
-
-When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
-The listener receives this string of words.
-By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
-The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
-Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
-Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
-If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).
-
-%We can view this as a formalization of the Language of Thought (CITE) TODO explain why
-
-%We formalize the listener's understanding as probabilistic inference, that is, the listener
-
-All natural languages have some degree of word order regularities that are specified in their grammar.
-For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
-That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
-Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
-%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
-%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
-This is illustrated in Figure~\ref{fig:grammars}: Grammars specify how dependency structures are encoded into strings of words.
-For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
-Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
-Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
-%In a speech community, speakers and listeners have shared knowledge of these rules, which allows the listener to decode the dependency tree.
-%In the terminology of information theory, these rules specify a \emph{channel} through which the speaker transmits the dependency structures.
-
-In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes.
-In our model, these will depend on the grammar:
-Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance (TODO can we find an example for this?)
-This reflects the \key{parseability} of the language.
-
-The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
+%
+%To quantify dependency length, we sum the lengths of syntactic dependencies in the sentence \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
+%\begin{equation}
+%    \label{eq:depl}
+%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+%\end{equation}
+%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
+Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
+
+\subsection{Efficiency Predicts Universals}
+
+DLM is stipulative and at most a heuristic approximation to psycholinguistic findings
+
+We now introduce a general information-theoretic notion of language efficiency.
+
+use the general speaker-listener picture
+
+\paragraph{Predictability}
+Predictability represents the incremental difficulty of online language processing.
+We formalize predictability as the \key{surprisal} (negative log probability) of each word given its preceding context \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+Optimization of predictability has been proposed as an explanation for certain word order patterns by \cite{ferrericancho2017optimality}.
+
+We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
+\begin{equation*}
+R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
+\end{equation*}
+%where the language model parameters $\phi$ are chosen to predict the language induced by tree distribution $T$ and word order grammar $\L_\theta$ as accurately as possible (maximizing predictability):
+%\begin{equation*}
+%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation*}
+We implement the language model $P_\phi$ using an LSTM neural network architecture \cite{hochreiter1997long}, currently the state of the art for this task \cite{goldberg2017neural}.
+%Note that the language model parameters $\phi$ and the word order grammar parameters $\theta$ are recursively dependent on each other.
+See the SI for details on how $\theta$ and $\phi$ are jointly optimized, as well as on the neural network architecture.
+
+\paragraph{Parseability}
+Parseability is the extent to which the syntactic and semantic relationships among words can be recovered from the string of words. 
+%Some notion of parseability has been proposed informally as a desirable property of languages in the linguistic literature going back at least to the 19th century \cite{humboldt,sapir,jesperson}, and parseability has been offered as an underlying justification for syntactic locality as a cost function for language processing \cite{hawkins-performance-1994}.
+%At a basic level, a language can encode information about syntactic relations into fixed word order patterns or into morphological marking on individual words.
+Much linguistic work has been devoted to investigating the trade-offs of various ways of indicating syntactic relations \cite{}.
+In this work, we present what is to our knowledge the first fully implemented computational formalization of this idea.
+We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
+Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
+\begin{align}
+\nonumber
+R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
+\nonumber
+&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
+\end{align}
+%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
+%\begin{equation*}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation*}
+We implement $P_\psi$ using a highly generic graph-based parsing architecture \cite{kiperwasser2016simple,dozat2017stanford} which recovers graph structures from word strings by solving a minimal spanning tree problem.
+%Due to data sparsity, we train parsers to predict dependency trees from POS tags alone, not from full wordforms, therefore we do not capture the extent to which morphological wordforms are informative about dependency tree structure.
+See the SI (appendix section 6-7) for further details on the joint optimization of $\theta$ and $\psi$ and on the neural parser architecture, and see SI (section 8) for robustness to different parsing models.
+
+\paragraph{Efficiency}
+The two scoring functions---predictability and parseability---can be combined into an overall equation of the form:
 \begin{equation}
-	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+\label{eq:combined-score}
+	R_{\textit{Eff}} = R_{Pars} + \lambda R_{Pred},
 \end{equation}
-where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
-This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
-It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
-This quantity formalizes Zipf's principle of listener effort and his Force of Diversification:
-It can be maximized when trees can be decoded fully unambiguously from utterances.
-Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's principle of listener economy~\cite{frank2012predicting,regier2015word}.
-
-%The counteracting Force of Unification is formalized by the cost of encoding the dependency structure and producing the utterance.
-%TODO motivate surprisal here.
-
-%Note that, in order for this process to be successful, speakers and listeners need to have shared knowledge of the rules by which dependency structures are encoded into sentences.
-%We hypothesize that the rules found in actual natural languages are optimized for the efficiency of this process, and that this optimization accounts for the observed word order universals.
-
-%If speakers choose to order their sentences in such a way that efficiency is optimized, will this lead to the observed language universals?
-
-%Do languages optimize their word order for efficiency?
-%Does this optimization explain the observed language universals?
-
-%We formalize a language as a distribution over dependency structures $T$ together with an ordering grammar $\theta$ that specifies how dependency structures are ordered into strings of words.
-%Applying the ordering grammar to the dependency structures, a language defines a probability distribution over sentences $L$.
-%Efficiency is defined as a tradeoff between
-
-% based on related work in information theory \cite{cover2006elements,tishby1999information}, bounded rationality \cite{genewein2015bounded}, and computational neuroscience \cite{friston2009predictive}; (ii)
-%
-%We formalize the notion of communicative efficiency using two terms: one having to do with minimizing the difficulty of linguistic communication, and one having to do with maximizing the information transferred in such communication.
-We quantify the difficulty of linguistic communication in terms of the word-by-word \key{predictability} of strings, which has been found to be a highly accurate and general predictor of human online comprehension difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect},
-% in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
-We quantify the information transferred in language as \key{parseability}: the quantity of information linguistic utterances contain about latent hierarchical tree structures.
-We also compare the results of optimizing for communicative efficiency to the results of optimizing for dependency length minimization, a previously proposed approximate metric of the amount of memory involved in language production and comprehension \cite{hudson1995measuring,gibson1998linguistic,gibson2000dependency,liu2017dependency,temperley2018minimizing}.
-
-We formalize the cost of an utterance as its word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
-Surprisal has been found to be is a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
-In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
-%The \key{predictability} of the language is the negative entropy of the sentences:
+with an interpolation weight $\lambda \in [0,1)$.
+In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification).
+This combined equation represents efficiency: bits of information transferred about dependency trees minus the cost of the transmission.
+In expectation over utterances (substituting Eq.~\ref{eq:combined-score} into Eq.~\ref{eq:exp-over-utts}), the overall objective function to be maximized by a word order grammar can be written in information-theoretic terms as:
 \begin{equation}
-	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+    \label{eq:efficiency}
+    J_T(\theta) = I[L_\theta(T);T] - \lambda H[L_\theta(T)],
 \end{equation}
-where the sum runs over all possible sentences $l$ that belong to the language.
-In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
-%how many distinct sentences there are.
-
-%It is larger
-
-%
-%The opposing pressure is \key{parsability}, which is the mutual information beteen sentences and dependency structures:
-%\begin{equation}
-%	R_{Pars} := I[L,T] = \sum_{t,l} p(t,l) \log \frac{p(t|l)}{p(t)}
-%\end{equation}
-%where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
-%This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
-%It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
-
-The \key{efficiency} of a language is a weighted combination
-\begin{equation}\label{eq:efficiency}
-	R_{Eff} := R_{Pars} + \lambda R_{Pred}
-\end{equation}
-with an interpolation weight $\lambda \in [0,1)$. In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification).
-
-%This combined equation represents efficiency: bits of information transferred about dependency trees minus the cost of the transmission.
-In expectation over utterances (substituting Eq.~\ref{eq:combined-score} into Eq.~\ref{eq:exp-over-utts}), the overall objective function to be maximized by a word order grammar can be written in information-theoretic terms as:
-
-%The two terms of Eq.~\ref{eq:efficiency} implement two opposing pressures.
+where $I[L_\theta(T);T]$ is the \key{mutual information} between strings and trees, and $H[L_\theta(T)]$ is the \key{entropy} of strings \cite{cover2006elements}.
+Maximizing Eq.~\ref{eq:efficiency} is equivalent to maximizing mutual information between strings and trees subject to a constraint on the maximum entropy of strings.
+
+The two terms of Eq.~\ref{eq:efficiency} implement two opposing pressures.
 Parseability (the mutual information term) pushes to make strings more heterogeneous, so that they can indicate different tree structures unambiguously.
 Predictability (the entropy term) pushes to make strings more homogeneous, to reduce the complexity of using the language.
-%The idea that natural language arises from a tension between these two forces goes back over a century \cite{gabelentz1901sprachwissenschaft}.
-Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking dependency structures as the underlying meanings to be conveyed.
+The idea that natural language arises from a tension between these two forces goes back over a century \cite{gabelentz1901sprachwissenschaft}.
+Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking unordered dependency trees $T$ as the underlying meanings to be conveyed.
 Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
 
-%\subsection{Optimizing Grammars for Efficiency}
-
-Our goals are to show that
-(1) the grammars of natural languages evolve towards optimizing efficiency of communication, and
-(2) this process of optimization accounts for Greenberg's order correlation universals.
-
-%For answering these questions,
-
-Answering these questions requires a sample of dependency structures as actually used by speakers of difefrent languages.
-Such samples have recently become available with the Universal Dependencies project, which has collected and created dependency annotations for several dozens of different languages.
-We use data from 51 languages.
-These corpora represent a typologically and genetically diverse group of languages.
-
-To answer whether languages evolve to have efficient word order, we compare the efficiency of the actual grammars of these 51 languages to randomly constructed baseline grammars.
-To show that this process of efficiency optimization accounts for Greenberg's correlation universals, we computationally construct grammars that optimize efficiency.
-We then show that these grammars mostly exhibit the Greenbergian order correlation universals.
-We furthermore show that this process of optimization also \emph{explains} DLM, providing a first-principles explanation of this heuristic generalization.
-
-For every one of the 51 corpora, we computationally construct eight optimal grammars.
-Note that the original orders of the actual languages do not enter this objective.
-See SI for our method for creating optimized grammars.
-
-%
-%All natural languages have some degree of word order regularities that are specified in their grammar.
-%It is thus not realistic to assume that speakers optimize each utterance individually.
-%Instead, we hypothesize that the word order rules of languages as a whole are optimized for efficiency, and that this optimization produces the observed universals.
-%
-
-%\subsection{Dependency Length Minimization}
-
+%In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
 \section{Results}
@@
 
-We first demonstrate that real languages are relatively efficient compared to random baselines. For each language, we generated ten baseline word order grammars for the language.
-These grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}).
-%by choosing all word order grammar parameters randomly at uniform from $[0,1]$.
+We first demonstrate that real languages are relatively efficient compared to random baselines.
+For each language, we generated ten baseline word order grammars for the language by choosing all word order grammar parameters randomly at uniform from $[0,1]$.
 Figure~\ref{fig:pareto-plane} shows the predictability and parseability for each real language relative to its baseline grammars.
@@
 
-%These results are shown in more detail for two typologically distinct languages (English and Japanese) in Figure~\ref{fig:predict-pars}.
+These results are shown in more detail for two typologically distinct languages (English and Japanese) in Figure~\ref{fig:predict-pars}.
 Here we include also the predictability and parseability of the original language data and of the word order grammars optimized for dependency length.
@@
 
-%\begin{figure*}
-%\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
-%	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right).
+\begin{figure*}
+\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
+	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right).
 We provide the entropy $H[L_\theta(T)]$ for predictability (lower is better) and the conditional entropy $H[T|L_\theta(T)]$ for parseability (lower is better).
 Both measures are normalized by sentence length.}\label{fig:predict-pars}
-%\end{figure*}
-%
-
-\subsection{Efficiency Explains DLM}
-We demonstrate the relationship between dependency length minimization and the maximization of efficiency.
+\end{figure*}
+
+Finally we demonstrate the relationship between dependency length minimization and the maximization of efficiency.
 Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency.
 We see that optimizing for efficiency lowers dependency length relative to random baselines, in keeping with the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}.
-In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the dependency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.
+In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the depenndency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.
 

### 2019-03-04 second-draft/4-paper-names.tex
revising

@@
 this is strongly correlated with the occurrence of pre- or postpositions:
+This is illustrated in Figure~\ref{fig:corr-table}.
 Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
 %However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
-This correlation has an extremely strong empirical basis:
+As Figure~\ref{fig:corr-table} shows, this correlation has an extremely strong empirical basis:
 In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization.
@@
 	   katabt \& risāla \& li \& sadīq  \\
-	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+%	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
 	   wrote \& letter   \& to \& friend  \\
@@
       tomodachi \& ni \& tegami-o \& kaita \\
- \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+ %\textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
 	   friend \& to \& letter  \& wrote \\
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.4\textwidth]{figures/corr-table.png}
+	\caption{The idea of a word order correlation: Languages can order the object before or after the verb (rows), and have prepositions or postpositions (columns).
+For each combination, we indicate how many languages satisfy it, as documented in the World Atlas of Language Structure~\cite{wals}.
+Combinations on the diagonal are vastly more common than off-diagonal ones.}
+	\label{fig:corr-table}
+\end{figure}
+
 This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
@@
 
-\section{Dependency Length Minimization Predicts Greenberg's Correlations}
-
 Theoretical studies of word order typology have proposed Dependency Length Minimization as a unifying principle behind Greenberg's correlations \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
 This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
-On the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, it has been confirmed that languages reduce dependency length, though a direct link to Greenger's correlations has not been established beyond theoretical arguments.
-This has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.
+Our first contribution is to show that this principle indeed correctly predicts Greenberg's correlations.
+We do this by creating counterfactual grammars that minimize dependency length of
+
+\section{Dependency Length Minimization Predicts Greenberg's Correlations}
+
+Theoretical studies of word order typology have proposed Dependency Length Minimization as a unifying principle behind Greenberg's correlations \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
 
@@
 
-The sentences in Figure~\ref{fig:dep-trees} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words.
-For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} of the verb, rather than e.g.
-the subject or an indirect object.
+The sentences in Figure~\ref{fig:dep-trees} are annotated with \key{dependency trees}:
+these are directed trees describing the grammatical relations among words.
+For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} of the verb, rather than e.g. the subject or an indirect object.
 A dependency arc is drawn from a \key{head} (e.g.
-\emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g.
-\emph{tegami-o}).
+\emph{visited} in Figure~\ref{fig:dep-trees}) to a \key{dependent} (e.g.
+\emph{friends}).
 Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}.
@@
 
-The dependency length of a sentence is the sum of the lengths of its syntactic dependencies \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
-\begin{equation}
-    \label{eq:depl}
-\operatorname{DepL}(t; {\bf w}) = \sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
-\end{equation}
-where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost.
-
-Large-scale corpus evidence across languages shows that natural languages order words in such a way as to reduce this quantity, compared to most random orderings \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-However, the link to Greenberg's correlations has not been established this directly.
+The \key{dependency length} of a sentence ${\bf w}$ with dependency structure $t$ is the sum of the lengths of its dependency arcs, measured as the number of words spanned \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
+%\begin{equation}
+%    \label{eq:depl}
+%\operatorname{DepL}(t; {\bf w}) = \sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+%\end{equation}
+%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$. 
+%On the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, it has been confirmed that languages reduce dependency length, though a direct link to Greenger's correlations has not been established beyond theoretical arguments.
+Large-scale corpus evidence across languages shows that natural languages order words in such a way as to reduce dependency length, compared to most random orderings \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+These data provide strong evidence that minimization of dependency length is itself a linguistic universal~\cite{temperley2018minimizing}
+
 While theoretical arguments suggest that DLM predicts Greenberg's correlations, this idea has not been tested on the basis of actual large-scale corpus data.
+Prior theoretical work on dependency length minimization has argued that Greenberg's correlations minimize dependency length \cite{temperley2008dependency}; however, the link is not trivial:
+Whether or not Greenberg's correlations theoretically minimize dependency length will depend on the distribution of tree topologies in actual language usage.
+%this depends on the topology of the dependency trees.
 
 We test this idea by constructing languages that have minimal dependency length:
+That is, we ask: Under the assumption that language has evolved to minimize dependency length, will Greenberg's correlations emerge?
+We computationally construct hypothetical variants of real languages that are optimized to minimize dependency length.
 If Greenberg's correlations emerge in these languages, then this provides evidence that DLM gives rise to these universals.
+
 All natural languages have some degree of word order regularities that are specified in their grammar.
@@
 Answering these questions requires a sample of dependency structures as actually used by speakers of difefrent languages.
-Such samples have recently become available with the Universal Dependencies project, which has collected and created dependency annotations for several dozens of different languages.
+Such samples have recently become available with the Universal Dependencies (UD) project, which has collected and created dependency annotations for several dozens of different languages.
 We use data from 51 languages.
@@
 
+%adposition    &    NP       &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
+%copula    &    NP        &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
+%auxiliary    &    VP       &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
+%noun    &    genitive      &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
+%noun    &    relative clause       &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
+%complementizer    &    S        &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
+%verb    &    PP     &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
+%want    &    VP      &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
+%verb    &    subject    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
+%verb    &    manner adverb     &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
+%
+
 \begin{table*}
@@
 	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} \textbf{86}  }
-	&   \multirow{2}{*}{  \textbf{81}$^{***}$    }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}   \textbf{81}$^{***}$    }
 	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf}   \textbf{47 \%}$^{}$   }
-	&   \multirow{2}{*}{  \textbf{76}$^{***}$ }   
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}  \textbf{76}$^{***}$ }   
 	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf} \textbf{68 \%}$^{***}$    } \\
@@
 copula    &    NP         
-	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf}}
-	&   \multirow{2}{*}{  \textbf{81}$^{***}$    }
-	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf}   }
-	&   \multirow{2}{*}{  \textbf{76}$^{***}$ }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}}  \\
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} \textbf{94} }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}   \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf}  \textbf{53}     }
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}  \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf} \textbf{61}$^{***}$  }  \\
 	\emph{is}        & \emph{a friend} \\ \hline
-auxiliary    &    VP       &   \\
-	\emph{has}          & \emph{written} \\
-noun    &    genitive      &  \\
-	\emph{friend} &  \emph{of John} \\
-noun    &    relative clause       &  \\
-	\emph{books} & \emph{that you read} \\
-complementizer    &    S        & \\
-	\emph{that} & \emph{Mary has arrived} \\
-verb    &    PP     &    \\
-	\emph{went} & \emph{to school} \\
-want    &    VP      &    \\
-	\emph{wants}   &  \emph{to leave} \\
-verb    &    subject    &  \\
-	\emph{(there) entered} & \emph{a person} \\
-verb    &    manner adverb     & \\
-	\emph{ran} & \emph{quickly} \\
+auxiliary    &    VP       
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}   \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}   }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}   \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}}  \\
+	\emph{has}          & \emph{written} \\ \hline
+noun    &    genitive      
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}}  \\
+	\emph{friend} &  \emph{of John} \\ \hline
+noun    &    relative clause      
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}\textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf} \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}}  \\
+	\emph{books} & \emph{that you read} \\ \hline
+complementizer    &    S        
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf} \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}   }
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf}\textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}}  \\
+	\emph{that} & \emph{Mary has arrived} \\ \hline
+verb    &    PP         
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf} \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}   }
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf}\textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}}  \\
+	\emph{went} & \emph{to school} \\ \hline
+want    &    VP        
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} \textbf{81}$^{***}$    }
+	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}   }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} \textbf{76}$^{***}$ }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}}  \\
+	\emph{wants}   &  \emph{to leave} \\ \hline
  \hline
@@
 We show that optimizing grammars for DLM produces Greenberg's correlations.
-We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
-Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
-
-Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages, or are not annotated reliably in UD.
+%Dryer \cite{dryer1992greenbergian} provides an authoritative
+We base our evaluation on the authoritative compilation of Greenberg's correlations by Dryer \cite{dryer1992greenbergian}, which documents 15 correlations, drawing on 625 languages.
+%, which we take as the basis of our evaluation.
+%Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
+%Considering those correlations that are annotated reliably in UD, we obtain eight correlations
+We exclude three correlations that are only relevant to some of the world's languages, and collapse correlations that are not distinguished in UD.
+%Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages, or are not annotated reliably in UD.
 %One further dependency is not annotated reliably in UD. %: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
-Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations (See SI for information on the other correlations).
+%Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations (See SI for information on the other correlations).
 % from Dryer's 15 correlations.
-These are described in Table~\ref{table:corr-dryer}.
-In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
-
-For each optimized grammar, we can check which correlations it satisfies.
-%In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
-We conducted, for each correlation, a mixed-effects logistic regression model predicting whether a grammar satisfies the correlation, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
-%We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau > 0.5$ holds for the correlating dependency depending on whether $a_\tau > 0.5$ for the verb-object dependency, with per-language-family random intercept and slope.\footnote{We coded language families according to \url{universaldependencies.org}.}
-%We are interested in the direction and significance of this effect: 
-%If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
+We describe the resulting eight correlations in Table~\ref{table:corr-dryer}.
+%In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
+
+For each optimized grammar, we can check which of these correlations it satisfies.
+We conducted a multivariate mixed-effects logistic regression model predicting, for each grammar, which of the eight correlations it satisfies.
+We entered the language and language family of the underlying corpus as random effects, accounting for language-specific properties of tree topologies.\footnote{We coded language families according to \url{universaldependencies.org}.}
 
 We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
-To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
+To evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
 The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
 %in all except four languages, for WALS did not have data.
+
+Optimizing for dependency length correctly accounts for all word order correlations.
 
@@
 
-Optimizing for dependency length correctly accounts for all word order correlations.
-
 \section{Formalizing Efficiency in Word Order}
@@
 It is also stipulative: There is no a priori reason why dependency length should be the crucial determining factor determining word order.
-There are other models of processing cost that make different predictions.
+While DLM has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, there are other models of processing cost that make different predictions.
+In particular, surprisal is known to be a strong predictor of human processing effort, but it makes predictions different from and sometimes opposite to those made by dependency length.
 DLM is also insufficient as an underlying principle of word order, as absolutely minimizing dependency length results in languages that have no consistent grammatical orders, making comprehension difficult.
@@
 
-Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
+Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort, the \emph{Force of Unification}, and listener effort, the \emph{Force of Diversification}.
 According to this theory, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
@@
 Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
+
 The idea that language results from the tension between these two pressures has a long and fruitful history.
@@
 %
-This idea has been shown to account for phenomena such as pragmatic inference \cite{horn-toward-1984, frank2012predicting} and color naming~\cite{zaslavsky2018efficient}.
+This idea has been shown to provide a formalization and derivation of pragmatic inference \cite{horn-toward-1984, frank2012predicting}.
+It also has been shown to account for the distribution of word meanings across languages, such as color naming~\cite{zaslavsky2018efficient}.
 
@@
 Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
-These are encoded in the dependency structure.
-%Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
+
+Dependency structures provide a natural formalization of such structures (CITE).
+%These are encoded in the dependency structure.
 %This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
@@
 
-Speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences
+Speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
 For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
 
-In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes.
+In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes.
+%Note that these terms correspond to Zipf's Forces of Unification Diversification
+In keeping with Zipf's Forces of Unification and Diversification, transmission cost is lowered by channels that support fewer distinct codes, whereas recoverability is increased when the channel supports of larger range of distinct codess for different messages.
 In our model, these will depend on the grammar:
-Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance (TODO can we find an example for this?)
+Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance. % (TODO can we find an example for this?)
 This reflects the \key{parseability} of the language.
@@
 It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
-This quantity formalizes Zipf's principle of listener effort and his Force of Diversification:
+This quantity formalizes Zipf's Force of Diversification:
 It can be maximized when trees can be decoded fully unambiguously from utterances.
-Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's principle of listener economy~\cite{frank2012predicting,regier2015word}.
+Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's Force of Diversification~\cite{frank2012predicting,regier2015word}.
 
@@
 Our goals are to show that
-(1) the grammars of natural languages evolve towards optimizing efficiency of communication, and
+(1) the grammars of natural languages have evolved towards optimizing efficiency of communication, and
 (2) this process of optimization accounts for Greenberg's order correlation universals.
 
-We use the same 51 corpora.
+%We use the same 51 corpora.
 
@@
 For every one of the 51 corpora, we computationally construct eight optimal grammars.
-Note that the original orders of the actual languages do not enter this objective.
+Again, the original orders of the actual languages do not enter this objective.
 See SI for our method for creating optimized grammars.
@@
 
-\section{Results}
-
-\subsection{Relative efficiency of languages}
-\label{sec:relative-efficiency}
-
-We first demonstrate that real languages are relatively efficient compared to random baselines. For each language, we generated ten baseline word order grammars for the language.
-These grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}).
+\subsection{Results}
+
+%\section{Results}
+
+%\subsection{Relative efficiency of languages}
+%\label{sec:relative-efficiency}
+
+We first demonstrate that real languages are relatively efficient compared to random baselines. For each of the 51 corpora, we generated ten baseline word order grammars.
+These grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:dlm-grammars}).
 %by choosing all word order grammar parameters randomly at uniform from $[0,1]$.
@@
 %; in general these maximum-likelihood word order grammars end up with lower predictability and parseability than the real languages. % TODO verify this is true? TODO awkward sentence
-For the calculation of predictability and parseability, we make all (baseline and real) word order grammars deterministic by always choosing the highest-probability linearization of each tree; by making the grammars deterministic in this way we eliminate an anticonservative bias toward low predictability in the baseline languages, which are highly nondeterministic.
+%For the calculation of predictability and parseability, we make all (baseline and real) word order grammars deterministic by always choosing the highest-probability linearization of each tree; by making the grammars deterministic in this way we eliminate an anticonservative bias toward low predictability in the baseline languages, which are highly nondeterministic.
 The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
@@
 
-Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
-Languages appear to be attracted toward these points and away from the region of the baseline languages.
-We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
+Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages. Languages appear to be attracted toward these points and away from the region of the baseline languages.
+%We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
 % TODO add something about dependency length-optimized languages?
@@
 
-\subsection{Efficiency Explains DLM}
-We demonstrate the relationship between dependency length minimization and the maximization of efficiency.
+%\subsection{Efficiency Explains DLM}
+We next demonstrate the relationship between dependency length minimization and the maximization of efficiency.
 Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency.
@@
 
-\subsection{Greenbergian Word Order Correlations}
+%\subsection{Efficiency Explains Greenbergian Word Order Correlations}
 \label{sec:greenberg}
-\begin{table*}
-	\begin{center}
-\begin{tabular}{|ll|l|l|l|ll|l|}
-	\hline
-	%&
-	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
-	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%adp. &
-adposition    &    NP       
-	&   \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf}   
-	&    \textbf{81}$^{***}$    
-	&   \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf}   
-	&    \textbf{76}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}  \\
-copula    &    NP       
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf}
-	&    \textbf{81}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf}
-	&    \textbf{79}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}  \\
-auxiliary    &    VP       
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf} 
-	&    \textbf{74}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf} 
-	&    55    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  \\
-noun    &    genitive      
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf} 
-	&    \textbf{82}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf} 
-	&    \textbf{74}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  \\
-noun    &    relative clause       
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf} 
-	&    \textbf{85}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf} 
-	&    \textbf{77}$^{**}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  \\
-complementizer    &    S        
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf} 
-	&    \textbf{85}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf} 
-	&    \textbf{80}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  \\
-verb    &    PP     
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf} 
-	&    \textbf{78}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf} 
-	&    59    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  \\
-want    &    VP      
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf} 
-	&    \textbf{90}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf} 
-	&    \textbf{92}$^{***}$    
-	&    \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf} 
-  \\
- \hline
-    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
-\end{tabular}
-	\end{center}
-	\caption{Greenbergian Correlations.
-Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%	For clarity, we organize the correlations by the category of the `verb patterner'.
-	For each correlation, we give our operationalization in terms of UD.
-%A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
-	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
-	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
-\end{table*}
-
-We now show that efficiency optimization predicts Greenberg's word order correlations. % in optimized grammars. 
+
+We finally show that efficiency optimization predicts Greenberg's word order correlations. % in optimized grammars. 
 Predictability and parseability predict four and six correlations, respectively, making complementary predictions.
@@
 
-\subsection{Discussion}
-
-In Section~\ref{sec:relative-efficiency}, we found that word order grammars in the majority of UD languages have better efficiency than baseline word order grammars.
+%\subsection{Discussion}
+
+%In Section~\ref{sec:relative-efficiency}, we found that word order grammars in the majority of UD languages have better efficiency than baseline word order grammars.
 Furthermore, in Section~\ref{sec:greenberg} we found that explicitly optimizing grammars for efficiency reproduces the Greenbergian correlations to the extent that they can be tested on the basis of available data.

### 2019-03-04 results/correlations/evaluation-depl_graphic.R

@@ -9,6 +9,9 @@ library(ggplot2)
 
 data = data %>% mutate(Language = fct_recode(Language, "Ancient_Greek" = "Ancient", "Old_Church_Slavonic" = "Old"))
 
+#best = read.csv("../strongest_models/best-depl.csv")
+
+#data = merge(data %>% mutate(FileName = as.character(FileName)), best %>% rename(FileName = Model), by=c("Language", "FileName"))
 
 
 
@@ -20,39 +23,39 @@ languages = read.csv("../languages/languages-iso_codes.tsv")
 dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"), all.x=TRUE)
 
 
-
-#options(mc.cores = parallel::detectCores())
-#rstan_options(auto_write = TRUE)
-
-
-dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataS = rbind(dataS, dataS2) %>% filter(Type == "manual_output_funchead_two_coarse_lambda09_best_balanced")
-
-dataP = read.csv("../../grammars/plane/plane-parse.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataP2 = read.csv("../../grammars/plane/plane-parse-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataP = rbind(dataP, dataP2) %>% filter(Type == "manual_output_funchead_two_coarse_lambda09_best_balanced")
-
-
-dataE = merge(dataS, dataP, by=c("Type", "Language", "Model")) %>% mutate(Eff = Pars+0.9*Surp) %>% group_by(Type, Language, Model) %>% summarise(Eff = mean(Eff))
-
-dataE = dataE[order(dataE$Language, dataE$Model),]
-
-eff = dataE$Eff
-models = c()
-for(i in (1:202)) {
-   if(eff[2*i] < eff[2*i+1]) {
-	   models = c(models, dataE$Model[2*i])
-   } else {
-	   models = c(models, dataE$Model[2*i+1])
-   }
-}
-
-
-dryer_greenberg_fine = dryer_greenberg_fine%>% filter(FileName %in% models)
-
-
-
+#
[…]

### 2019-03-04 results/correlations/evaluation-parser_graphic.R

@@ -1,9 +1,6 @@
 
 # _final/
-data = read.csv("../../grammars/manual_output_funchead_two_coarse_lambda09_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
-
-
-best = read.csv("../strongest_models/best-two-lambda09-best-balanced.csv")
+data = read.csv("../../grammars/manual_output_funchead_two_coarse_parser_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
 
 library(forcats)
 library(dplyr)
@@ -12,13 +9,18 @@ library(ggplot2)
 
 data = data %>% mutate(Language = fct_recode(Language, "Ancient_Greek" = "Ancient", "Old_Church_Slavonic" = "Old"))
 
+best = read.csv("../strongest_models/best-parse-best-balanced.csv")
+
+data = merge(data %>% mutate(FileName = as.character(FileName)), best %>% rename(FileName = Model), by=c("Language", "FileName"))
+
+
+
 
 
 
 dryer_greenberg_fine = data
 
 
-data = merge(data %>% mutate(FileName = as.character(FileName)), best %>% rename(FileName = Model), by=c("Language", "FileName"))
 
 languages = read.csv("../languages/languages-iso_codes.tsv")
 dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"), all.x=TRUE)
@@ -27,10 +29,6 @@ dryer_greenberg_fine  = merge(dryer_greenberg_fine, languages, by=c("Language"),
 
 #options(mc.cores = parallel::detectCores())
 #rstan_options(auto_write = TRUE)
-
-
-
-dependency = "nmod"
 
 getCorrPair = function(dependency) {
    corr_pair = dryer_greenberg_fine %>% filter((CoarseDependency == dependency) | (CoarseDependency == "obj"))
@@ -44,6 +42,7 @@ getCorrPair = function(dependency) {
        corr_pair = corr_pair %>% mutate(correlator_s=1-correlator_s)
    }
 
+   corr_pair$agree = (corr_pair$correlator_s == corr_pair$obj_s)
    return(corr_pair)
 }
 
@@ -78,7 +77,7 @@ for(dependency in dependencies) {
         axis.text=element_blank(),
         axis.ticks.x=element_blank(),
[…]

### 2019-03-04 results/correlations/joint-evaluation-efficiency09-large-full.R

@@ -1,0 +1,45 @@
+data = read.csv("../../grammars/manual_output_funchead_two_coarse_lambda09_best_large/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+
+library(forcats)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+data = data %>% mutate(Language = fct_recode(Language, "Ancient_Greek" = "Ancient", "Old_Church_Slavonic" = "Old"))
+
+
+languages = read.csv("../languages/languages-iso_codes.tsv")
+data  = merge(data, languages, by=c("Language"), all.x=TRUE)
+library("brms")
+dependency = "nmod"
+dependencies = c("acl", "advmod", "aux", "lifted_case", "lifted_cop", "lifted_mark", "nmod", "nsubj", "obl", "xcomp")
+objs = data %>% filter(CoarseDependency == "obj") %>% mutate(obj = pmax(0, sign(DH_Weight))) %>% select(obj, Language, FileName)
+data = data %>% filter((CoarseDependency %in% dependencies))
+data = merge(data, objs, by=c("Language", "FileName"))
+data = data %>% mutate(dir = pmax(0, sign(DH_Weight)))
+data = data %>% mutate(dir = ifelse(DH_Weight == 0, NA, dir))
+data$agree = (data$dir == data$obj)
+data = unique(data %>% select(Family, Language, FileName, CoarseDependency, agree)) %>% spread(CoarseDependency, agree)
+
+model3 = brm(cbind(acl, advmod, aux, lifted_case, lifted_cop, lifted_mark, nmod, nsubj, obl, xcomp) ~ (1|p|Family) + (1|q|Language), family="bernoulli", data=data, iter=5000)
+
+
+
+u = posterior_samples(model3) #cbind(posterior_samples(model3, "acl_Intercept"), posterior_samples(model3, "advmod_Intercept"))
[…]

### 2019-03-13 second-draft/3-paper-names.tex
visualization

--- 2019-02-20 second-draft/3-paper-names.tex
@@
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,1,2]{Michael Hahn}
-\author[a]{Daniel Jurafsky}
-\author[b]{Richard Futrell}
-
-\affil[a]{Stanford University}
-\affil[b]{University of California, Irvine}
+%\author[a,1,2]{Michael Hahn}
+%\author[a]{Daniel Jurafsky}
+%\author[b]{Richard Futrell}
+
+%\affil[a]{Stanford University}
+%\affil[b]{University of California, Irvine}
 
@@
 \begin{abstract}
+	TODO rewrite
 The universal properties of human languages have been the subject of intense study across disciplines.
 	We report novel computational and corpus-based evidence for the hypothesis that a prominent subset of these universal properties---those related to word order---result from a process of optimization for efficient communication among humans.
-We develop a probabilistic, differentiable model of word order grammars: the means by which different languages convert underlying hierarchical structures into strings of words.
-We show how the parameters of a word order grammar can be optimized for efficiency of information transfer, quantified as a tradeoff between incremental predictability and mutual information with latent tree structures.
-Applying these grammars to tree structures found in dependency corpora from 51 languages, we show that optimizing the grammar parameters for efficiency results in word order patterns that reproduce a large subset of the major word order correlations reported in the linguistic typological literature, and reproduce the predictions of previous heuristic theories such as dependency length minimization.
+	TODO explain
+	%We develop a probabilistic, differentiable model of word order grammars: the means by which different languages convert underlying hierarchical structures into strings of words.
+	%We show how the parameters of a word order grammar can be optimized for efficiency of information transfer, quantified as a tradeoff between incremental predictability and mutual information with latent tree structures.
+	Applying these grammars to tree structures found in dependency corpora from 51 languages, we show that optimizing the grammar parameters for efficiency results in word order patterns that predict a large subset of the major word order correlations reported in the linguistic typological literature, and subsumes the predictions of previous heuristic theories such as dependency length minimization.
 \end{abstract}
@@
 
-\dropcap{F}or decades, researchers in fields ranging from philology to cognitive science to statistical physics have been involved in documenting and trying to explain the universal syntactic and statistical properties of human language \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical}.
+\dropcap{F}or TODO rewrite first sentence.
+decades, researchers in fields ranging from philology to cognitive science to statistical physics have been involved in documenting and trying to explain the universal syntactic and statistical properties of human language \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical}.
 An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
@@
 While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
-this is strongly correlated with the occurrence of pre- or postpositions:
+this is strongly correlated with the occurrence of pre- or postpositions (Figure~\ref{fig:corr-table}, top):
 Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
 %However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
-This correlation has an extremely strong empirical basis:
-In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization.
+%As Figure~\ref{fig:corr-table} shows, this correlation has an extremely strong empirical basis:
+%In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization.
 %, over 90 \% of languages satisfy this generalization .
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.4\textwidth]{figures/corr-full.png}
+%
+%\centering
+%\begin{dependency}[theme = simple]
+%   \begin{deptext}[column sep=1em]
+%%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
+%	   katabt \& risāla \& li \& sadīq  \\
+%%	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
+%	   wrote \& letter   \& to \& friend  \\
+%   \end{deptext}
+%\end{dependency}
+%\begin{dependency}[theme = simple]
+%   \begin{deptext}[column sep=1em]
+%      tomodachi \& ni \& tegami-o \& kaita \\
+% %\textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
+%	   friend \& to \& letter  \& wrote \\
+%   \end{deptext}
+%\end{dependency}
+%
+%    \centering
+%    \includegraphics[width=0.4\textwidth]{figures/corr-table.png}
+%
+%%\begin{dependency}[theme=simple]
+%%    \begin{deptext}
+%%    I \& wrote \& a \& letter \& to \& a \& friend \\
+%%    \end{deptext}
+%%    \depedge{2}{1}{nsubj}
+%%    \depedge{2}{4}{obj}
+%%    \depedge{4}{3}{ }
+%%    \depedge{4}{3}{ }
+%%    \depedge{2}{5}{}
+%%    \depedge{5}{7}{ }
+%%    \depedge{7}{6}{ }
+%%\end{dependency}
+	\caption{ The concept of a word order correlation: Languages can order the object after (Arabic) or before (Japanese) the verb, and have prepositions (Arabic) or postpositions (Japanese).
+For each combination, we indicate how many languages satisfy it, as documented in the World Atlas of Language Structure~\cite{wals}.
+Combinations on the diagonal are vastly more common than off-diagonal ones.
+% A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.
+	}\label{fig:arabic-japanese-simple}	\label{fig:corr-table}
+\end{figure}
+
+This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
+These describe correlations between the relative positions of different types of expressions across languages.
+The example above documents that the position of the object (`letter') relative to the verb is \key{correlated} with the position of the adposition (`to').
+%adpositional phrases (`to a friend') relative to the verb.
+Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
+The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.
+
+Supported by languages on all continents, these correlations are among the language universals with the strongest empirical support.
+Importantly, their validity is also independent from specific assumptions about theories of grammar.
+
+Consequently, explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
+An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
+
+A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
+Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.
+
+Researchers working in the functional paradigm have proposed a range of criteria that languages should meet in order to enable efficient communication and processing.
+These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
+
+Prior studies of word order typology have proposed Dependency Length Minimization (DLM) as general principle of word order \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+This is the idea that languages order words so as to minimize the linear distance between words linked in grammatical relationships.
+This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in language processing \cite{gibson1998linguistic,gibson2000dependency}.
+Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+It has also been argued theoretically that this principle should account for Greenberg's correlations, though this link is nontrivial and depends on the precise distributions of tree topologies in syntactic structures.
+As a principle of word order, DLM is also stipulative: There are other processing models, such as surprisal, that are strongly predictive of human processing cost, and make predictions distinct from, and partly conflicting with, those of DLM~\cite{hale2001probabilistic,levy2008expectation}.
+
+Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between the Forces of Unification and Diversification.
+The Force of Unification is a pressure to reduce the complexity of the language to make production and processing as easy as possible.
+Conversely, the Force of Diversification favors languages that provide different utterances for different meanings, so that the listener can unambiguously identify the meaning from the utterance.
+%Minimizing speaker and listener effort represent two opposing forces:
+%A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
+%A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
+%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
+These two forces act in opposing directions:
+Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
+The idea that language results from the tension between these two pressures has a long and fruitful history.
+%\cite{horn-toward-1984}, RSA, Color naming
+%
+%In the domain of word order, 
+%
+%TODO introduce the terms \key{cost} and \key{informativity}
+%
+This idea has been shown to provide a formalization and derivation of pragmatic inference \cite{horn-toward-1984, frank2012predicting}.
+It also has been shown to account for the distribution of word meanings across languages, such as color naming~\cite{zaslavsky2018efficient}.
+
+%
+
+%
+%
+
+%TODO somewhere start calling this Efficiency
+%TODO explain why surprisal. some production/speaker work. (plan reuse)
+
+%In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
+%Even earlier than Zipf, \cite{gabelentz1901sprachwissenschaft} ...
+%This theory blabla Horn RSA
+%Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
+
+In this work, we first provide a formalization of efficiency in the domain of word order, drawing on information-theoretic notions.
+We then show that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts both for the Greenbergian word order correlations and for Dependency Length Minimization.
+
+%In this work, we show computationally, using corpus data from 51 languages, that efficiency optimization accounts for the Greenbergian word order correlations.
+%That is, we show that the word orders in natural languages have evolved to optimize efficiency, and that this optimization accounts for the prevalence of the word order correlations.
+%word orders in natural language
+
+%OR say DLM and efficiency, as in the talks
+
+%\section{DLM}
+
+%
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=0.45\textwidth]{figures/letter-options.png}
+%	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure.
+In this example, Grammar 1 achieves the lowest overall dependency length.
+Indeed, this is the }
+%	\label{fig:pareto-plane}
+%\end{figure}
+%
+%
+
+%
+%
+%\begin{table}
+%	\begin{center}
+%%		\setlength\tabcolsep{1.5pt}
+%\begin{tabular}{|ll|c|cccc|llllllll}
+%	\hline
+%	%&
+%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Eff}  \\ 
+%	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%%adp. &
+%	adposition    &    NP       
+%	& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}       }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}   }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}     }  \\
+%	\emph{to}            & \emph{a friend} &&&&&\\ \hline
+%copula    &    NP         
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}       }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}   }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}   }  \\
+%	\emph{is}        & \emph{a friend}  &&&&&\\ \hline
+%auxiliary    &    VP       
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}       }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  }  \\
+%	\emph{has}          & \emph{written}  &&&&&\\ \hline
+%noun    &    genitive      
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  }  \\
+%	\emph{friend} &  \emph{of John}  &&&&&\\ \hline
+%noun    &    relative clause      
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  }  \\
+%	\emph{books} & \emph{that you read}  &&&&&\\ \hline
+%complementizer    &    S        
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf} }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  }  \\
+%	\emph{that} & \emph{she has arrived}  &&&&&\\ \hline
+%verb    &    PP         
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  }  \\
+%	\emph{went} & \emph{to school}  &&&&&\\ \hline
+%want    &    VP        
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}  }  \\
+%	\emph{wants}   &  \emph{to leave}  &&&&&\\ \hline
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+%\end{tabular}
+%	\end{center}
+%	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%For each correlation, we provide an English example.
+%%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%%Not all correlations are satisfied by every natu
+%}\label{table:corr-dryer}
+%\end{table}
+%
+%
+
+%
+%
+%\begin{table*}
+%	\begin{center}
+%\begin{tabular}{|ll|l|l|l|ll|l|llllllll}
+%	\hline
+%	%&
+%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
+%	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%%adp. &
+%	adposition    &    NP       
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}   $^{***}$    }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}  $^{***}$ }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf} $^{***}$    } \\
+%	\emph{to}            & \emph{a friend} \\ \hline
+%copula    &    NP         
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}   $^{***}$    }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}  $^{***}$ }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf} $^{**}$  }  \\
+%	\emph{is}        & \emph{a friend} \\ \hline
+%auxiliary    &    VP       
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}   $^{***}$    }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}  $^{***}$   }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}$^{**}$  }  \\
+%	\emph{has}          & \emph{written} \\ \hline
+%noun    &    genitive      
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} $^{***}$    }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} $^{***}$ }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}$^{***}$  }  \\
+%	\emph{friend} &  \emph{of John} \\ \hline
+%noun    &    relative clause      
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf} $^{***}$    }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf} $^{**}$ }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}$^{***}$  }  \\
+%	\emph{books} & \emph{that you read} \\ \hline
+%complementizer    &    S        
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf} $^{***}$    }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}  $^{**}$   }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf}$^{***}$ }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}$^{**}$  }  \\
+%	\emph{that} & \emph{Mary has arrived} \\ \hline
+%verb    &    PP         
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf} $^{***}$    }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}  $^{***}$   }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}$^{**}$  }  \\
+%	\emph{went} & \emph{to school} \\ \hline
+%want    &    VP        
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} $^{***}$    }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf} $^{**}$    }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} $^{***}$ }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}$^{***}$  }  \\
+%	\emph{wants}   &  \emph{to leave} \\ \hline
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+%\end{tabular}
+%	\end{center}
+%	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%For each correlation, we provide an English example.
+%%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%%Not all correlations are satisfied by every natu
+%}\label{table:corr-dryer}
+%\end{table*}
+%
+
+\section{Formalizing Efficiency in Word Order}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figures/communication.png}
+	\caption{Our model of communication: A speaker (left) expresses a dependency structure into a string of words forming a sentence.
+A listener probabilistically recovers a dependency structure.
+In this example, the grammar of English allows the listener to recover the correct dependency structure with very high confidence.}
+	\label{fig:comm}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.45\textwidth]{figures/grammars-colors-trees.png}
+	\caption{Given a set of dependency structures, grammars define a set of sentences.
+Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb.
+\textbf{TODO This figure is just a draft sketch so far.}}
+	\label{fig:grammars}
+\end{figure}
+
+%
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=0.45\textwidth]{../results/correlations/figures/coverage-ground_eff-circles.pdf}
+%	\caption{For each language, we show which of the eight correlations (columns) are satisfied in the actual language (right), and in grammars optimized for efficiency (left).}
+%	\label{fig:per-lang}
+%\end{figure}
+%
+
+We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
+In this model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
+Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
+Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
+This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
+In Figure~\ref{fig:comm}, the message consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
+
+When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
+The listener receives this string of words.
+By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
+The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
+Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
+Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
+If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).
+
+All natural languages have some degree of word order regularities that are specified in their grammar.
+For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
+That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
+Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
+%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
+%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
+This is illustrated in Figure~\ref{fig:grammars}: Grammars specify how dependency structures are encoded into strings of words.
+For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
+Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
+Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
+We formalize word order grammars as specifying how different types of dependents (subjects, objects, etc.) are ordered, employing a cross-linguistically meaningful inventory of relations~\cite{nivre2015universal}.
+
+In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes.
+In our model, these will depend on the grammar:
+Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
+This reflects the \key{parseability} of the language.
+
+The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
+\begin{equation}
+	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+\end{equation}
+where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
+This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
+It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
+This quantity formalizes Zipf's principle of listener effort and his Force of Diversification:
+It can be maximized when trees can be decoded fully unambiguously from utterances.
+Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's principle of listener economy~\cite{frank2012predicting,regier2015word}.
+
+We formalize the cost of an utterance as its word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
+Surprisal has been found to be is a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
+%The \key{predictability} of the language is the negative entropy of the sentences:
+\begin{equation}
+	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+\end{equation}
+where the sum runs over all possible sentences $l$ that belong to the language.
+In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
+%how many distinct sentences there are.
+
+The \key{efficiency} of a language is a weighted combination
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
+\end{equation}
+with an interpolation weight $\lambda \in [0,1)$. In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
+
+Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking dependency structures as the underlying meanings to be conveyed.
+Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
+
+Our goals are to show that
+(1) the grammars of natural languages evolve towards optimizing efficiency of communication, and
+(2) this process of optimization accounts both for DLM and for Greenberg's order correlation universals.
+
+Answering these questions requires a sample of dependency structures as actually used by speakers of different languages.
+Such samples have recently become available with the Universal Dependencies project~\cite{nivre2017universal}, which has collected and created dependency annotations for several dozens of different languages.
+51 languages had sufficient data for our purposes. %There was sufficient data from 51 languages for our We use data from 51 languages.
+These corpora represent a typologically and genetically diverse group of languages.
+
+To answer whether languages evolve to have efficient word order, we compare the efficiency of the actual grammars of these 51 languages to randomly constructed baseline grammars.
+To show that this process of efficiency optimization accounts for both DLM and Greenberg's correlation universals, we computationally construct grammars that optimize efficiency, and evaluate these for dependency length and for Greenberg's universals.
+If hypothetical grammars that are optimal for efficiency exhibit these typological patterns, this shows that these language universals can be explained by the process of efficiency optimization.
+%For every one of the 51 corpora, we computationally construct eight optimal grammars.
+%Note that the original orders of the actual languages do not enter this objective.
+See SI for our method for creating optimized grammars.
+
+\section{Results}
+
+\subsection{Relative efficiency of languages}
+\label{sec:relative-efficiency}
+
+We first demonstrate that real languages are relatively efficient compared to random baselines. For each language, we generated ten baseline word order grammars for the language.
+These grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}).
+The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend.pdf}
+    \caption{Predictability and parseability of 51 languages (red), indicated by ISO codes, compared to baseline word order grammars per language (blue distribution).
+Predictability and parseability scores are $z$-scored within language.
+Each point for a real language has a line pointing in the direction of the center of mass of its baselines.
+Unlabeled dots represent the centroid for real languages (red), baseline languages (blue), and languages optimized for predictability (purple), parseability (green), efficiency (orange), and dependency length (yellow).
+When a language is to the bottom--left of its baselines, this indicates that it is relatively optimal for efficiency.} % TODO colors, fontsize, etc could conceivably be made prettier
+    \label{fig:pareto-plane}
+\end{figure}
+
+Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
+Languages appear to be attracted toward these points and away from the region of the baseline languages.
+We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
+
+%\begin{figure*}
+%\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
+%	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right).
+We provide the entropy $H[L_\theta(T)]$ for predictability (lower is better) and the conditional entropy $H[T|L_\theta(T)]$ for parseability (lower is better).
+Both measures are normalized by sentence length.}\label{fig:predict-pars}
+%\end{figure*}
+%
+
+\subsection{Efficiency Explains DLM}
+We demonstrate the relationship between dependency length minimization and the maximization of efficiency.
+Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency.
+We see that optimizing for efficiency lowers dependency length relative to random baselines, confirming the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}.
+In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the dependency length of the real language; on average, the real language is best matched by efficiency optimization.
+
+\begin{figure} % need Efficiency and (maybe) MLE langs on here
+\includegraphics[width=0.4\textwidth]{../results/dependency-length/figures/depl-violin.pdf}
+	\caption{Average dependency length for real languages (blue), random baselines (purple), and for languages optimized for efficiency (green) and explicitly for dependency length (red).
+Dependency length was computed from sentences of lengths 8-12 and averaged over words.}\label{fig:deplength}
+\end{figure}
+
+\subsection{Greenbergian Word Order Correlations}
+\label{sec:greenberg}
+
+\begin{figure} % need Efficiency and (maybe) MLE langs on here
+	\includegraphics[width=0.4\textwidth]{../results/correlations/figures/coverage-ground_eff-circles.pdf}
+	\caption{per-languag predictions}
+\end{figure}
+
+\begin{table}
+	\begin{center}
+%		\setlength\tabcolsep{1.5pt}
+\begin{tabular}{|ll|c|cc|}
+	\hline
+	%&
+	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{DepL}   &  \multirow{2}{*}{Eff}  \\ 
+	verb & object     & & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%adp. &
+	adposition    &    NP       
+	& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_case.pdf} } %\textbf{86}  }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_case.pdf}       }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_case.pdf}   }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_case.pdf}     }  \\
+	\emph{to}            & \emph{a friend} &&&\\ \hline
+copula    &    NP         
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_cop.pdf} }% \textbf{94} }
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_cop.pdf} } % \textbf{53}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_cop.pdf}   }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_cop.pdf}   }  \\
+	\emph{is}        & \emph{a friend}  &&&\\ \hline
+auxiliary    &    VP       
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-aux.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-aux.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-aux.pdf}  }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-aux.pdf}  }  \\
+	\emph{has}          & \emph{written}  &&&\\ \hline
+noun    &    genitive      
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-nmod.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-nmod.pdf}   }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-nmod.pdf}  }  \\
+	\emph{friend} &  \emph{of John}  &&&\\ \hline
+noun    &    relative clause      
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-acl.pdf}}
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-acl.pdf}   }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-acl.pdf}  }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-acl.pdf}  }  \\
+	\emph{books} & \emph{that you read}  &&&\\ \hline
+complementizer    &    S        
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_mark.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_mark.pdf}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_mark.pdf} }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_mark.pdf}  }  \\
+	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+verb    &    PP         
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-obl.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-obl.pdf}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-obl.pdf} }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-obl.pdf}  }  \\
+	\emph{went} & \emph{to school}  &&&\\ \hline
+want    &    VP        
+	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-xcomp.pdf}}
+	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-xcomp.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-xcomp.pdf}  }  \\
+	\emph{wants}   &  \emph{to leave}  &&&\\ \hline
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+\end{tabular}
+	\end{center}
+	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+For each correlation, we provide an English example, and show the distribution of real languages, of languages optimized for DLM, and of languages optimized for efficiency.
+The process of optimization fir efficiency and DLM explains all eight universals.
+%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%Not all correlations are satisfied by every natu
+}\label{table:corr-dryer}
+\end{table}
+
+\begin{table}
+	\begin{center}
+%		\setlength\tabcolsep{1.5pt}
+\begin{tabular}{|ll|c|cc|}
+	\hline
+	%&
+	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  &  \multirow{2}{*}{Eff}  \\ 
+	verb & object     & & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%adp. &
+	adposition    &    NP       
+	& \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}       }
+%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
+%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}   }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}     }  \\
+	\emph{to}            & \emph{a friend} &&&\\ \hline
+copula    &    NP         
+	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}       }
+%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}   }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}   }  \\
+	\emph{is}        & \emph{a friend}  &&&\\ \hline
+auxiliary    &    VP       
+	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}       }
+%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  }  \\
+	\emph{has}          & \emph{written}  &&&\\ \hline
+noun    &    genitive      
+	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  }  \\
+	\emph{friend} &  \emph{of John}  &&&\\ \hline
+noun    &    relative clause      
+	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
+%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf}  }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  }  \\
+	\emph{books} & \emph{that you read}  &&&\\ \hline
+complementizer    &    S        
+	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf} }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  }  \\
+	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+verb    &    PP         
+	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
+	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  }  \\
+	\emph{went} & \emph{to school}  &&&\\ \hline
+want    &    VP        
+	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
+%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
+	&   \parbox[c]{0.05\textwidth}{\multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}  }}  \\
+	\emph{wants}   &  \emph{to leave}  &&&\\ \hline
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+\end{tabular}
+	\end{center}
+	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+For each correlation, we provide an English example, and show the distribution of real languages, of languages optimized for DLM, and of languages optimized for efficiency.
+The process of optimization fir efficiency and DLM explains all eight universals.
+%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%Not all correlations are satisfied by every natu
+}\label{table:corr-dryer}
+\end{table}
+
+%
+%
+%\begin{table}
+%	\begin{center}
+%%		\setlength\tabcolsep{1.5pt}
+%\begin{tabular}{|ll|c|cc|cc|}
+%	\hline
+%	%&
+%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  &  \multirow{2}{*}{Parser}  &  \multirow{2}{*}{Eff}  \\ 
+%	verb & object     & & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%%adp. &
+%	adposition    &    NP       
+%	& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_case.pdf} } %\textbf{86}  }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_case.pdf}       }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
+%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_case.pdf}   }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_case.pdf}     }  \\
+%	\emph{to}            & \emph{a friend} &&&\\ \hline
+%copula    &    NP         
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_cop.pdf} }% \textbf{94} }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_cop.pdf} } % \textbf{53}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_cop.pdf}   }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_cop.pdf}   }  \\
+%	\emph{is}        & \emph{a friend}  &&&\\ \hline
+%auxiliary    &    VP       
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-aux.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-aux.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-aux.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-aux.pdf}  }  \\
+%	\emph{has}          & \emph{written}  &&&\\ \hline
+%noun    &    genitive      
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-nmod.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-nmod.pdf}   }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-nmod.pdf}  }  \\
+%	\emph{friend} &  \emph{of John}  &&&\\ \hline
+%noun    &    relative clause      
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-acl.pdf}}
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-acl.pdf}   }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-acl.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-acl.pdf}  }  \\
+%	\emph{books} & \emph{that you read}  &&&\\ \hline
+%complementizer    &    S        
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_mark.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_mark.pdf}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_mark.pdf} }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_mark.pdf}  }  \\
+%	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+%verb    &    PP         
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-obl.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-obl.pdf}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-obl.pdf} }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-obl.pdf}  }  \\
+%	\emph{went} & \emph{to school}  &&&\\ \hline
+%want    &    VP        
+%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-xcomp.pdf}}
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-xcomp.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-xcomp.pdf}  }  \\
+%	\emph{wants}   &  \emph{to leave}  &&&\\ \hline
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+%\end{tabular}
+%	\end{center}
+%	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%For each correlation, we provide an English example, and show the distribution of real languages, of languages optimized for DLM, and of languages optimized for efficiency.
+The process of optimization fir efficiency and DLM explains all eight universals.
+%%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%%Not all correlations are satisfied by every natu
+%}\label{table:corr-dryer}
+%\end{table}
+%
+%
+
+%\begin{table*}
+%	\begin{center}
+%\begin{tabular}{|ll|l|l|l|ll|l|}
+%	\hline
+%	%&
+%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
+%	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%%adp. &
+%adposition    &    NP       &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
+%copula    &    NP        &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
+%auxiliary    &    VP       &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
+%noun    &    genitive      &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
+%noun    &    relative clause       &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
+%complementizer    &    S        &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
+%verb    &    PP     &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
+%want    &    VP      &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
+%verb    &    subject    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
+%verb    &    manner adverb     &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+%\end{tabular}
+%	\end{center}
+%	\caption{Greenbergian Correlations.
+Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%%	For clarity, we organize the correlations by the category of the `verb patterner'.
+%	For each correlation, we give our operationalization in terms of UD.
+%A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
+%	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
+%	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
+%\end{table*}
+%
+
+We now show that efficiency optimization predicts Greenberg's word order correlation universals.
+We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages.
+Most correlations can be straightforwardly implemented in Universal Dependencies (UD), allowing us to check which correlations a word order grammar satisfies.
+Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages or are not annotated reliably in UD.
+Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations.
+
+These are shown in Table~\ref{table:corr-dryer}.
+%Each correlation is visualized 
+%
+%
+We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
+To evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
+%The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
+%in all except four languages, for WALS did not have data.
+
+Table~\ref{table:corr-dryer} shows that grammars optimized for efficiency show correlations 
+In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
+We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
+We are interested in the direction and significance of this effect: 
+If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
+
+\paragraph{Results}
+Results are shown in Table~\ref{table:corr-dryer}.
+All eight correlations are confirmed in the models estimated from the real orderings.
+Efficiency significantly predicts all the word order correlations, each in the same direction as attested in the dependency corpora.
+As efficiency predicts DLM, we predict that DLM also accounts for all dependencies.
+
+We confirmed this finding by estimating the posterior distribution of the number of correlations that the optimization process correctly predicts.
+The estimated posterior probability that efficiency optimization predicts less than all eight correlations is $0.0242$.
+The probability that it predicts less than seven of the correlations is $<10^{-4}$.
+Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.
+
+\section{Conclusion}
+
+We found that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
+We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals.
+Beyond our present results, we provide a complete formalization and computational framework in which theories of the functional optimization of languages can be tested.
+Other objective functions could be hypothesized and tested within our framework; furthermore, future advances in machine learning will enable the optimization of richer and richer models of grammar.
+
+A major question for functional explanations for linguistic universals is: \emph{how} do languages come to be optimized?
+Do speakers actively seek out new communicative conventions that allow better efficiency?
+Or do languages change in response to biases that come into play during language acquisition \cite{fedzechkina2012language,culbertson2012learning}?
+Our work is neutral toward such questions.
+To the extent that language universals arise from biases in learning or in the representational capacity of the human brain, our results suggest that those biases tilt toward communicative efficiency.
+Our work does provide against the idea that word order universals are best explained in terms of learning biases that are irreducibly arbitrary and genetic in nature \cite{chomsky2010some}.
+
+While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into dependency trees.
+The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication.
+Relatedly, there are many remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
+Our models cannot capture this trade-off for technical reasons: the parseability objective does not operate over wordforms, only POS tags, and so it does not take advantage of morphological cues to syntactic structure.
+Future work can investigate whether these and other remaining universals can be explained using more sophisticated models of how meaning representations are transduced into strings of words, based on larger databases with richer annotations.
+
+\matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
+We use all languages for which at least one treebank with a training partition was available, a total of 50 languages.
+For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.
+Punctuation was removed.
+%We removed punctuation. %, which we operationalized by removing all dependents of \emph{punct} dependencies, and recursively removing their dependents.
+
+Universal dependencies represents as dependents some words that are typically classified as heads in syntactic theory.
+This particularly applies to the \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark} dependencies.
+Following prior work studying dependency length minimization \cite{futrell2015largescale}, we modified each treebank by inverting these dependencies, promoting the dependent to the head position.
+We report results on this modified version of UD.
+
+The efficiency optimization results from Table~\ref{tab:results-dryer} were preregistered: \url{http://aspredicted.org/blind.php?x=8gp2bt}.
+
+See the SI for details on the neural language models and parsers used, and on the optimization procedures.}
+
+\showmatmethods{} % Display the Materials and Methods section
+
+\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Noah Goodman, Chris Manning, and audiences at CAMP 2018 for helpful discussion.}
+
+\showacknow{} % Display the acknowledgments section
+
+%\bibliographystyle{acl_natbib}
+%\bibliography{extracted}
+\bibliography{everything}
+
+\end{document}
+
+%the meaning
 

### 2019-03-13 results/correlations/tiles-real_eff-byObj-restricted_full.R

@@ -25,15 +25,15 @@ D$Direction = ifelse(D$DirObj == 1, "OV", "VO")
 ordersByLanguage = unique(data.frame(Language = D$Language, Direction = as.character(D$Direction)))
 
 
-data = read.csv("../../grammars/manual_output_funchead_two_coarse_lambda09_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
-#best = read.csv("../strongest_models/best-two-lambda09-best-balanced.csv")
+data = read.csv("../../grammars/manual_output_funchead_two_coarse_lambda09_best_large/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+best = read.csv("../../grammars/manual_output_funchead_two_coarse_lambda09_best_large/successful-seeds.tsv")
 library(dplyr)
 library(tidyr)
 library(ggplot2)
 library(forcats)
 data = data %>% mutate(Language = fct_recode(Language, "Old_Church_Slavonic" = "Old", "Ancient_Greek"= "Ancient"))
-#best = best %>% mutate(Language = fct_recode(Language, "Old_Church_Slavonic" = "Old", "Ancient_Greek"= "Ancient"))
-#data = merge(data %>% mutate(FileName = as.character(FileName)), best %>% rename(FileName = Model), by=c("Language", "FileName"))
+best = best %>% mutate(Language = fct_recode(Language, "Old_Church_Slavonic" = "Old", "Ancient_Greek"= "Ancient"))
+data = merge(data %>% mutate(FileName = as.character(FileName)), best %>% rename(FileName = Model), by=c("Language", "FileName"))
 data = data
 languages = read.csv("../languages/languages-iso_codes.tsv", sep=",")
 data  = merge(data, languages, by=c("Language"), all.x=TRUE)
@@ -80,7 +80,7 @@ plot = ggplot(D, aes(x = CoarseDependency, y = Language_Ordered)) +
   theme(axis.title=element_blank()) + 
   theme(legend.position="none") + facet_wrap(~Type)
 
-ggsave(file="figures/coverage-ground_eff-byObj-circles.pdf", plot=plot)
+ggsave(file="figures/coverage-ground_eff_large_full-byObj-circles.pdf", plot=plot)
 
 
 #E = D %>% group_by(Language) %>% mutate(AgreeSum = sum(Agree))

### 2019-03-13 results/dependency-length/depLengthPerObjective.R

@@ -37,7 +37,119 @@ deplReal = depl %>% filter(Type == "REAL_REAL")
 depl = merge(depl, bestModels, by=c("Language", "Model", "Type"))
 depl = rbind(depl, deplRand, deplReal)
 
-depl2 = depl %>% group_by(Language, Type) %>% summarise(AverageLengthPerWord = mean(AverageLengthPerWord)) %>% spread(Type, AverageLengthPerWord)
+depl2 = depl %>% group_by(Language, Type) %>% summarise(AverageLengthPerWord = mean(AverageLengthPerWord)) #%>% spread(Type, AverageLengthPerWord)
+
+depl4 = depl2 %>% spread(Type, AverageLengthPerWord)
+
+
+meanPerLang = depl2 %>% group_by(Language) %>% summarise(MeanLength = mean(AverageLengthPerWord))
+depl2 = merge(depl2, meanPerLang, by=c("Language"))
+depl2 = depl2 %>% filter(!(Type %in% c("manual_output_funchead_langmod_coarse_best_balanced", "manual_output_funchead_two_coarse_parser_best_balanced")))
+
+plot = ggplot(data=depl2, aes(x=AverageLengthPerWord-MeanLength, group=Type, fill=Type)) + geom_density(alpha=0.3)
+
+depl3 = depl %>% filter(!(Type %in% c("manual_output_funchead_langmod_coarse_best_balanced", "manual_output_funchead_two_coarse_parser_best_balanced")))
+depl3$Type = factor(depl3$Type, levels=c("manual_output_funchead_coarse_depl", "manual_output_funchead_two_coarse_lambda09_best_balanced", "REAL_REAL",  "manual_output_funchead_RANDOM"  ))
+library(forcats)
+depl3$TypeN = fct_recode(depl3$Type, DLM="manual_output_funchead_coarse_depl", Eff="manual_output_funchead_two_coarse_lambda09_best_balanced", Real="REAL_REAL", Baseline="manual_output_funchead_RANDOM")
+
+
+library(ggrepel)
[…]

### 2019-03-15 second-draft/3-paper-names.tex
revising

@@
 
-\dropcap{F}or TODO rewrite first sentence.
-decades, researchers in fields ranging from philology to cognitive science to statistical physics have been involved in documenting and trying to explain the universal syntactic and statistical properties of human language \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical}.
-An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-In this work we examine this question from a computational perspective, demonstrating a fully formalized framework in which certain syntactic universals can be explained through the statistical optimization of grammars for information-theoretic efficiency.
-% TODO add a cogsci person to citation chain at the beginning
+\dropcap{F} \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
 
@@
 
-Natural languages vary a lot in the order in which they express information.
+Natural languages vary in the order in which they express information.
 Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
@@
 In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
-While the two sentences containing words with the same meanings, the order of these words are entirely different in the two languages:
+Yet, the order of these words are entirely different in the two languages:
 %However, the languages differ in the order in which these words and phrases are arranged: 
-In Arabic, the verb stands at the beginning, followed by both the object and the adpositional phrase.
+In Arabic, the verb stands at the beginning, followed by the object and the adpositional phrase.
 `To' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
@@
 %It turns out that these are deep and stable regularities:
-It turns out that this variation reflects a deep and stable regularity:
+This variation reflects a deep and stable regularity:
 While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
@@
 
-Consequently, explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
+Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
 An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
+(TODO cite linguists)
 
@@
 These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
-
-Prior studies of word order typology have proposed Dependency Length Minimization (DLM) as general principle of word order \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-This is the idea that languages order words so as to minimize the linear distance between words linked in grammatical relationships.
-This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in language processing \cite{gibson1998linguistic,gibson2000dependency}.
-Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-It has also been argued theoretically that this principle should account for Greenberg's correlations, though this link is nontrivial and depends on the precise distributions of tree topologies in syntactic structures.
-As a principle of word order, DLM is also stipulative: There are other processing models, such as surprisal, that are strongly predictive of human processing cost, and make predictions distinct from, and partly conflicting with, those of DLM~\cite{hale2001probabilistic,levy2008expectation}.
 
@@
 When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
+Should the speaker do it as
+Mary saw John (like in English)
+or as 
+Mary John saw (like in Japanese)
+
+Depending on the language, the listener decodes this.
+
 The listener receives this string of words.
@@
 In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes.
+
+Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking dependency structures as the underlying meanings to be conveyed.
+Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
+
 In our model, these will depend on the grammar:
@@
 We formalize the cost of an utterance as its word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
-Surprisal has been found to be is a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+Surprisal has been found to be a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
 In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
@@
 
-Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking dependency structures as the underlying meanings to be conveyed.
-Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
-
-Our goals are to show that
+\section{Results}
+
+We examined whether
 (1) the grammars of natural languages evolve towards optimizing efficiency of communication, and
-(2) this process of optimization accounts both for DLM and for Greenberg's order correlation universals.
+(2) this process of optimization accounts both for Greenberg's order correlation universals.
 
@@
 
-\section{Results}
-
 \subsection{Relative efficiency of languages}
@@
 The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
+X languages improve over the baselines in parsability, Y improve in predictability.
+Z improve in overall efficiency.
+Pareto frontier
+
+Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
+Languages appear to be attracted toward these points and away from the region of the baseline languages.
+We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
 
@@
 
-Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
-Languages appear to be attracted toward these points and away from the region of the baseline languages.
-We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
-
-%\begin{figure*}
-%\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
-%	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right).
-We provide the entropy $H[L_\theta(T)]$ for predictability (lower is better) and the conditional entropy $H[T|L_\theta(T)]$ for parseability (lower is better).
-Both measures are normalized by sentence length.}\label{fig:predict-pars}
-%\end{figure*}
-%
-
-\subsection{Efficiency Explains DLM}
-We demonstrate the relationship between dependency length minimization and the maximization of efficiency.
-Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency.
-We see that optimizing for efficiency lowers dependency length relative to random baselines, confirming the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}.
-In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the dependency length of the real language; on average, the real language is best matched by efficiency optimization.
-
-\begin{figure} % need Efficiency and (maybe) MLE langs on here
-\includegraphics[width=0.4\textwidth]{../results/dependency-length/figures/depl-violin.pdf}
-	\caption{Average dependency length for real languages (blue), random baselines (purple), and for languages optimized for efficiency (green) and explicitly for dependency length (red).
-Dependency length was computed from sentences of lengths 8-12 and averaged over words.}\label{fig:deplength}
-\end{figure}
-
 \subsection{Greenbergian Word Order Correlations}
@@
 	%&
-	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{DepL}   &  \multirow{2}{*}{Eff}  \\ 
+	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Eff}  \\ 
 	verb & object     & & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
@@
 	& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_case.pdf} } %\textbf{86}  }
-	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_case.pdf}       }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_case.pdf}       }
 %	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
@@
 	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_cop.pdf} }% \textbf{94} }
-	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
 %	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_cop.pdf} } % \textbf{53}     }
@@
 	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-aux.pdf}}
-	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
 %	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-aux.pdf}     }
@@
 	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-nmod.pdf}}
-	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
 %	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-nmod.pdf}   }
@@
 	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-acl.pdf}}
-	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
+%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
 %	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-acl.pdf}   }
@@
 	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_mark.pdf}}
-	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
 %	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_mark.pdf}     }
@@
 	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-obl.pdf}}
-	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
 %	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-obl.pdf}     }
@@
 	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-xcomp.pdf}}
-	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
 %	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-xcomp.pdf}     }
@@
 	\end{center}
-	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
+Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
 For each correlation, we provide an English example, and show the distribution of real languages, of languages optimized for DLM, and of languages optimized for efficiency.
@@
 
-\begin{table}
-	\begin{center}
-%		\setlength\tabcolsep{1.5pt}
-\begin{tabular}{|ll|c|cc|}
-	\hline
-	%&
-	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  &  \multirow{2}{*}{Eff}  \\ 
-	verb & object     & & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%adp. &
-	adposition    &    NP       
-	& \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}       }
-%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
-%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}   }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}     }  \\
-	\emph{to}            & \emph{a friend} &&&\\ \hline
-copula    &    NP         
-	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}       }
-%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}   }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}   }  \\
-	\emph{is}        & \emph{a friend}  &&&\\ \hline
-auxiliary    &    VP       
-	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}       }
-%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}     }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  }  \\
-	\emph{has}          & \emph{written}  &&&\\ \hline
-noun    &    genitive      
-	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
-%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  }  \\
-	\emph{friend} &  \emph{of John}  &&&\\ \hline
-noun    &    relative clause      
-	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
-%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}     }
-%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf}  }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  }  \\
-	\emph{books} & \emph{that you read}  &&&\\ \hline
-complementizer    &    S        
-	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf}     }
-%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}     }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf} }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  }  \\
-	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-verb    &    PP         
-	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf}     }
-%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}     }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  }  \\
-	\emph{went} & \emph{to school}  &&&\\ \hline
-want    &    VP        
-	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
-%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}     }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
-	&   \parbox[c]{0.05\textwidth}{\multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}  }}  \\
-	\emph{wants}   &  \emph{to leave}  &&&\\ \hline
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-\end{tabular}
-	\end{center}
-	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-For each correlation, we provide an English example, and show the distribution of real languages, of languages optimized for DLM, and of languages optimized for efficiency.
+We now show that efficiency optimization predicts Greenberg's word order correlation universals.
+We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages.
+Most correlations can be straightforwardly implemented in Universal Dependencies (UD), allowing us to check which correlations a word order grammar satisfies.
+Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages or are not annotated reliably in UD.
+Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations.
+
+These are shown in Table~\ref{table:corr-dryer}.
+We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
+To evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
+The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
+%in all except four languages, for WALS did not have data.
+
+Table~\ref{table:corr-dryer} shows that grammars optimized for efficiency accurately predict correlations to hold at prevalences similar to actual languages.
+
+%In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
+We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
+We are interested in the direction and significance of this effect: 
+If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
+
+We estimated the posterior distribution of the number of correlations that the optimization process correctly predicts.
+The estimated posterior probability that efficiency optimization predicts less than all eight correlations is $0.0242$.
+The probability that it predicts less than seven of the correlations is $<10^{-4}$.
+%Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.
+
+%
+%\begin{table}
+%	\begin{center}
+%%		\setlength\tabcolsep{1.5pt}
+%\begin{tabular}{|ll|c|cc|}
+%	\hline
+%	%&
+%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  &  \multirow{2}{*}{Eff}  \\ 
+%	verb & object     & & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%%adp. &
+%	adposition    &    NP       
+%	& \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
+%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}       }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}   }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}     }  \\
+%	\emph{to}            & \emph{a friend} &&&\\ \hline
+%copula    &    NP         
+%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
+%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}       }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
+%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}   }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}   }  \\
+%	\emph{is}        & \emph{a friend}  &&&\\ \hline
+%auxiliary    &    VP       
+%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
+%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}       }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}     }
+%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  }  \\
+%	\emph{has}          & \emph{written}  &&&\\ \hline
+%noun    &    genitive      
+%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
+%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
+%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  }  \\
+%	\emph{friend} &  \emph{of John}  &&&\\ \hline
+%noun    &    relative clause      
+%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
+%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}     }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
+%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf}  }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  }  \\
+%	\emph{books} & \emph{that you read}  &&&\\ \hline
+%complementizer    &    S        
+%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
+%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf}     }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}     }
+%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf} }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  }  \\
+%	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+%verb    &    PP         
+%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
+%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf}     }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}     }
+%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
+%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  }  \\
+%	\emph{went} & \emph{to school}  &&&\\ \hline
+%want    &    VP        
+%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
+%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
+%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}     }
+%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
+%	&   \parbox[c]{0.05\textwidth}{\multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}  }}  \\
+%	\emph{wants}   &  \emph{to leave}  &&&\\ \hline
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+%\end{tabular}
+%	\end{center}
+%	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%For each correlation, we provide an English example, and show the distribution of real languages, of languages optimized for DLM, and of languages optimized for efficiency.
 The process of optimization fir efficiency and DLM explains all eight universals.
-%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%Not all correlations are satisfied by every natu
-}\label{table:corr-dryer}
-\end{table}
+%%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%%Not all correlations are satisfied by every natu
+%}\label{table:corr-dryer}
+%\end{table}
+%
+%
 
@@
 
-We now show that efficiency optimization predicts Greenberg's word order correlation universals.
-We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages.
-Most correlations can be straightforwardly implemented in Universal Dependencies (UD), allowing us to check which correlations a word order grammar satisfies.
-Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages or are not annotated reliably in UD.
-Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations.
-
-These are shown in Table~\ref{table:corr-dryer}.
-%Each correlation is visualized 
-%
-%
-We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
-To evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
-%The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
-%in all except four languages, for WALS did not have data.
-
-Table~\ref{table:corr-dryer} shows that grammars optimized for efficiency show correlations 
-In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
-We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
-We are interested in the direction and significance of this effect: 
-If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
-
-\paragraph{Results}
-Results are shown in Table~\ref{table:corr-dryer}.
-All eight correlations are confirmed in the models estimated from the real orderings.
-Efficiency significantly predicts all the word order correlations, each in the same direction as attested in the dependency corpora.
-As efficiency predicts DLM, we predict that DLM also accounts for all dependencies.
-
-We confirmed this finding by estimating the posterior distribution of the number of correlations that the optimization process correctly predicts.
-The estimated posterior probability that efficiency optimization predicts less than all eight correlations is $0.0242$.
-The probability that it predicts less than seven of the correlations is $<10^{-4}$.
-Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.
+\section{Alternative Theories}
+
+Prior studies have proposed Dependency Length Minimization (DLM) as general principle of word order \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+This is the idea that languages order words so as to minimize the linear distance between words linked in grammatical relationships.
+This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in language processing \cite{gibson1998linguistic,gibson2000dependency}.
+Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+It has also been argued theoretically that this principle should account for Greenberg's correlations, though this link is nontrivial and depends on the precise distributions of tree topologies in syntactic structures.
+As a principle of word order, DLM is also stipulative: There are other processing models, such as surprisal, that are strongly predictive of human processing cost, and make predictions distinct from, and partly conflicting with, those of DLM~\cite{hale2001probabilistic,levy2008expectation}.
+\cite{futrell2017memory} suggests that efficiency predicts DLM.
+
+We show that Efficiency optimization predicts DLM, providing a first-principles functional explanation.
+Figure~\ref{fig:deplength} shows average dependency length for real languages, random baselines, and languages optimized for dependency length and efficiency.
+Optimizing for efficiency lowers dependency length relative to random baselines.
+
+Our results suggest that dependency length minimization is a by-product of efficiency maximization.
+In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the dependency length of the real language; on average, the real language is best matched by efficiency optimization.
+
+\begin{figure} % need Efficiency and (maybe) MLE langs on here
+\includegraphics[width=0.4\textwidth]{../results/dependency-length/figures/depl-violin.pdf}
+	\caption{Average dependency length for real languages (blue), random baselines (purple), and for languages optimized for efficiency (green) and explicitly for dependency length (red).
+Dependency length was computed from sentences of lengths 8-12 and averaged over words.}\label{fig:deplength}
+\end{figure}
 

### 2019-03-17 second-draft/3-paper-names.tex
revising

@@
 
-In this work, we first provide a formalization of efficiency in the domain of word order, drawing on information-theoretic notions.
-We then show that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts both for the Greenbergian word order correlations and for Dependency Length Minimization.
+The contribution of our work is to show  that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
+%In this work, we first provide a formalization of efficiency in the domain of word order. %, drawing on information-theoretic notions.
 
@@
 
+%\subsection{Counterfactual Grammars}
 We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
@@
 When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
-Should the speaker do it as
-Mary saw John (like in English)
-or as 
-Mary John saw (like in Japanese)
-
-Depending on the language, the listener decodes this.
-
-The listener receives this string of words.
-By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
-The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
-Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
+Imagine the speaker wants to communicate that Mary saw John.
+Should the speaker say (1) as in English, or (2) as in Japanese?
+
+(1)
+\begin{dependency}[theme=simple]
+    \begin{deptext}
+    Mary \& saw \& John \\
+    \end{deptext}
+    \depedge{2}{1}{nsubj}
+    \depedge{2}{3}{obj}
+\end{dependency}
+(2)
+\begin{dependency}[theme=simple]
+    \begin{deptext}
+    Mary \& John  \& saw \\
+    \end{deptext}
+    \depedge{3}{1}{nsubj}
+    \depedge{3}{2}{obj}
+\end{dependency}
+
+The listener receives a sentence -- a linear string of words -- and decodes the underlying message.
 Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
 If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).
+
+%The listener receives this string of words.
+%By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
+%The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
+%Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
 
@@
 
+%\subsection{Efficiency in Word Order}
+Zipf
+
 In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes.
@@
 Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
+
+Efficiency is formalized as a weighted combination of terms representing the amount of information conveyed and the cost of communication.
+%The \key{efficiency} of a language is a weighted combination
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
+\end{equation}
+with an interpolation weight $\lambda \in [0,1)$. %In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
 
@@
 This reflects the \key{parseability} of the language.
-
+Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's principle of listener economy~\cite{frank2012predicting,regier2015word}.
 The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
@@
 It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
-This quantity formalizes Zipf's principle of listener effort and his Force of Diversification:
+This quantity formalizes Zipf's Force of Diversification:
 It can be maximized when trees can be decoded fully unambiguously from utterances.
-Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's principle of listener economy~\cite{frank2012predicting,regier2015word}.
 
@@
 
-The \key{efficiency} of a language is a weighted combination
-\begin{equation}\label{eq:efficiency}
-	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
-\end{equation}
-with an interpolation weight $\lambda \in [0,1)$. In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
-
 \section{Results}
@@
 These grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}).
-The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
-X languages improve over the baselines in parsability, Y improve in predictability.
-Z improve in overall efficiency.
-Pareto frontier
-
-Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
-Languages appear to be attracted toward these points and away from the region of the baseline languages.
+In Figure~\ref{fig:pareto-plane}, we plot predictability and parsability of the grammars of 51 languages, together with the distribution of random baseline grammars, and the Pareto frontier defined by computationally optimized grammars.
+The majority of real languages are above and to the right of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
+X languages improve over the baselines in parsability ($p < ...$ each), Y improve in predictability ($p < ...$ each).
+Z improve in overall efficiency ($p < ...$ each).
+Languages are attracted towards the Pareto frontier and away from the region of the baseline languages.
+A few languages end up \emph{more} efficient than the computationally optimized languages.
+
+%Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
+%Languages appear to be attracted toward these points and away from the region of the baseline languages.
 We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
@@
     \centering
-    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend.pdf}
-    \caption{Predictability and parseability of 51 languages (red), indicated by ISO codes, compared to baseline word order grammars per language (blue distribution).
-Predictability and parseability scores are $z$-scored within language.
-Each point for a real language has a line pointing in the direction of the center of mass of its baselines.
-Unlabeled dots represent the centroid for real languages (red), baseline languages (blue), and languages optimized for predictability (purple), parseability (green), efficiency (orange), and dependency length (yellow).
-When a language is to the bottom--left of its baselines, this indicates that it is relatively optimal for efficiency.} % TODO colors, fontsize, etc could conceivably be made prettier
+    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend-viz.pdf}
+    \caption{Predictability and parseability of 51 languages (red), indicated by ISO codes, compared to baseline word order grammars (blue distribution).
+Predictability and parseability scores are $z$-scored within language.} % TODO colors, fontsize, etc could conceivably be made prettier
     \label{fig:pareto-plane}
@@
     \centering
-    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend.pdf}
+    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend-viz.pdf}
     \caption{Predictability and parseability of 51 UD languages (red), indicated by ISO codes, compared to ten baseline word order grammars per language (green).

### 2019-03-17 results/plane/plane-2D-best-visualization-large.R

@@ -303,7 +303,7 @@ names(myColors) <- levels(forColoring)
 
 plot = ggplot(dataPlot)  
 plot = plot + theme_bw() 
-plot = plot + geom_segment(aes(x=-Pars_z, y=-Surp_z, xend=-Pars_z_end, yend = -Surp_z_end, color=TypeN, group=TypeN), size=0.6)
+#plot = plot + geom_segment(aes(x=-Pars_z, y=-Surp_z, xend=-Pars_z_end, yend = -Surp_z_end, color=TypeN, group=TypeN), size=0.6)
 plot = plot + geom_density_2d(data=dataPlot %>% filter(grepl("Baseline", TypeN)), aes(x=-Pars, y=-Surp, color=TypeN, group=TypeN), size=0.3)
 plot = plot + geom_path(data=dataPlot %>% filter(TypeN %in% c("Parseability", "Efficiency", "Predictability")), aes(x=-ParsMean, y=-SurpMean, group=1), color="gray", size=5)
 #plot = plot + geom_point(data=dataPlot, aes(x=-ParsMean, y=-SurpMean, color=TypeN, group=TypeN), size=6)

### 2019-03-22 second-draft/3-paper-names.tex
revising

@@
 \end{figure}
+
+TODO maybe Arabic becomes English?
 
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{figures/communication.png}
-	\caption{Our model of communication: A speaker (left) expresses a dependency structure into a string of words forming a sentence.
-A listener probabilistically recovers a dependency structure.
-In this example, the grammar of English allows the listener to recover the correct dependency structure with very high confidence.}
+    \includegraphics[width=0.45\textwidth]{figures/communication-two-langs.png}
+	\caption{Grammars define communication channels: A speaker (left) expresses a dependency structure into a string of words forming a sentence.
+The order of words is determined by the grammar of the language.
+A listener recovers a dependency structure.}
 	\label{fig:comm}
@@
 Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
-Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
+Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar (CITE):
 This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
-In Figure~\ref{fig:comm}, the message consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
-
+
+Consider Figure~\ref{fig:comm}.
+Imagine the speaker wants to communicate that Mary saw John.
+The message (left) consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
 When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
-Imagine the speaker wants to communicate that Mary saw John.
-Should the speaker say (1) as in English, or (2) as in Japanese?
+In natural language, this order is largely determined by the \key{grammar} of the language:
+The speaker will order it as (1) in English, and as (2) in Japanese.
 
@@
 
-The listener receives a sentence -- a linear string of words -- and decodes the underlying message.
-Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
-If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).
+The listener receives a sentence -- a linear string of words -- and decodes the underlying message in accordance with the grammar of the language (right).
+%Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
+%If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).
 
@@
 
-All natural languages have some degree of word order regularities that are specified in their grammar.
-For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
-That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
-Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
+The grammars of natural languages provide consistent ordering rules for all syntactic relations.
+%All natural languages have some degree of word order regularities that are specified in their grammar.
+%For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
+%That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
+%Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
 %Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
 %While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
-This is illustrated in Figure~\ref{fig:grammars}: Grammars specify how dependency structures are encoded into strings of words.
+This is illustrated in Figure~\ref{fig:grammars}:
+We show how four different grammars order objects, adpositional phrases, and adpositions.
 For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
@@
 Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
-We formalize word order grammars as specifying how different types of dependents (subjects, objects, etc.) are ordered, employing a cross-linguistically meaningful inventory of relations~\cite{nivre2015universal}.
 
 %\subsection{Efficiency in Word Order}
-Zipf
-
-In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes.
+%Zipf
+
+%We formalize Zipf's Forces of Unification and Diversification in
+
+In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes (REF).
 
@@
 (2) this process of optimization accounts both for Greenberg's order correlation universals.
+
+We formalize word order grammars as specifying how different types of dependents (subjects, objects, etc.) are ordered, employing a cross-linguistically meaningful inventory of relations~\cite{nivre2015universal}.
 

### 2019-03-22 results/plane/plane-2D-best-visualization.R

@@ -1,0 +1,321 @@
+
+
+
+
+library(lme4)
+
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+
+
+
+
+depl = read.csv("../../grammars/dependency_length/total_summary_funchead_coarse.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+library(tidyr)
+library(dplyr)
+depl = depl %>% filter(grepl("FuncHead", ModelName)) %>% filter(grepl("Coarse", ModelName))
+
+#deplBest = depl %>% group_by(Language) %>% filter(Type == "manual_output_funchead_coarse_depl")
+#deplBest = merge(deplBest, deplBest %>% summarise(AverageLengthPerWord = min(AverageLengthPerWord)), by=c("Language", "AverageLengthPerWord"))
+#
+#deplBest = deplBest$Model
+#
+
+
+
+dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS = rbind(dataS, dataS2)
+
+dataP = read.csv("../../grammars/plane/plane-parse.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP2 = read.csv("../../grammars/plane/plane-parse-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP = rbind(dataP, dataP2)
+
+
+summary(lmer(Surp ~ Type + (1|Language), data=dataS %>% filter(Type %in% c("manual_output_funchead_langmod_coarse_final", "manual_output_funchead_langmod_coarse_best_balanced"))))
+summary(lmer(Surp ~ Type + (1|Language), data=dataS %>% filter(Type %in% c("manual_output_funchead_two_coarse_lambda09_best_balanced", "manual_output_funchead_langmod_coarse_best_balanced"))))
+
+
+summary(lmer(Pars ~ Type + (1|Language), data=dataP %>% filter(grepl("balanc", Type))))
+summary(lmer(Pars ~ Type + (1|Model) + (1|Language), data=dataP %>% filter(grepl("balanc", Type))))
+summary(lmer(UAS ~ Type + (1|Language), data=dataP %>% filter(grepl("balanc", Type))))
+
[…]

### 2019-03-23 second-draft/3-paper-names.tex
revising

@@
 
+% Given a set of dependency structures, grammars define a set of sentences. Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., 
 \begin{figure}
@@
     \includegraphics[width=0.45\textwidth]{figures/grammars-colors-trees.png}
-	\caption{Given a set of dependency structures, grammars define a set of sentences.
-Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb.
+	\caption{Grammars define consistent ordering rules for dependency structures.
+Here, Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb.
 \textbf{TODO This figure is just a draft sketch so far.}}
@@
 \end{figure}
+
+TODO at some point need to foreshadow more of the Greenberg correlations
 
@@
 When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
-In natural language, this order is largely determined by the \key{grammar} of the language:
+In natural language, this order is largely determined by the \key{grammar} of the language (CITE):
 The speaker will order it as (1) in English, and as (2) in Japanese.
@@
 (2) this process of optimization accounts both for Greenberg's order correlation universals.
-
-We formalize word order grammars as specifying how different types of dependents (subjects, objects, etc.) are ordered, employing a cross-linguistically meaningful inventory of relations~\cite{nivre2015universal}.
-
-Answering these questions requires a sample of dependency structures as actually used by speakers of different languages.
+%We formalize word order grammars as specifying how different types of dependents (subjects, objects, etc.) are ordered, employing a cross-linguistically meaningful inventory of relations~\cite{nivre2015universal}.
+Answering these questions requires a sample of dependency structures as actually used by speakers across different languages.
 Such samples have recently become available with the Universal Dependencies project~\cite{nivre2017universal}, which has collected and created dependency annotations for several dozens of different languages.
@@
 
-To answer whether languages evolve to have efficient word order, we compare the efficiency of the actual grammars of these 51 languages to randomly constructed baseline grammars.
-To show that this process of efficiency optimization accounts for both DLM and Greenberg's correlation universals, we computationally construct grammars that optimize efficiency, and evaluate these for dependency length and for Greenberg's universals.
-If hypothetical grammars that are optimal for efficiency exhibit these typological patterns, this shows that these language universals can be explained by the process of efficiency optimization.
+\subsection{Relative efficiency of languages}
+\label{sec:relative-efficiency}
+
+We first demonstrate that real languages are relatively efficient compared to random baselines, providing evidence that grammars evolve towards efficiency.
+We compare the efficiency of the actual grammars of the 51 languages from the Universal Dependencies datasets to randomly constructed baseline grammars.
+For every one of the 51 languages, we construct ten counterfactual versions by randomly creating ten grammars.
+These baseline grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}, see \textit{Materials and Methods} for details).
+We estimate predictability and parsability using start-of-the-art machine learning methods (see \textit{Materials and Methods}).
+Furthermore, for every one of the 51 languages, we computationally construct grammars that are \emph{optimized} for parsability, predictability, and efficiency (see \textit{Materials and Methods}).
+
 %For every one of the 51 corpora, we computationally construct eight optimal grammars.
 %Note that the original orders of the actual languages do not enter this objective.
-See SI for our method for creating optimized grammars.
-
-\subsection{Relative efficiency of languages}
-\label{sec:relative-efficiency}
-
-We first demonstrate that real languages are relatively efficient compared to random baselines. For each language, we generated ten baseline word order grammars for the language.
-These grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}).
+%See SI for our method for creating optimized grammars.
+% For each language, we generated ten baseline word order grammars for the language.
 In Figure~\ref{fig:pareto-plane}, we plot predictability and parsability of the grammars of 51 languages, together with the distribution of random baseline grammars, and the Pareto frontier defined by computationally optimized grammars.
@@
 Languages are attracted towards the Pareto frontier and away from the region of the baseline languages.
-A few languages end up \emph{more} efficient than the computationally optimized languages.
+%A few languages end up \emph{more} efficient than the computationally optimized languages.
 
@@
     \caption{Predictability and parseability of 51 languages (red), indicated by ISO codes, compared to baseline word order grammars (blue distribution).
-Predictability and parseability scores are $z$-scored within language.} % TODO colors, fontsize, etc could conceivably be made prettier
+Predictability and parseability scores are $z$-scored within language.
+The gray curve indicates the Pareto frontier of computationally optimized grammars, averaged over the 51 languages.} % TODO colors, fontsize, etc could conceivably be made prettier
     \label{fig:pareto-plane}
@@
 	\includegraphics[width=0.4\textwidth]{../results/correlations/figures/coverage-ground_eff-circles.pdf}
-	\caption{per-languag predictions}
+	\caption{per-languag predictions}\label{fig:per-lang}
 \end{figure}
@@
 
-We now show that efficiency optimization predicts Greenberg's word order correlation universals.
-We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages.
-Most correlations can be straightforwardly implemented in Universal Dependencies (UD), allowing us to check which correlations a word order grammar satisfies.
-Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages or are not annotated reliably in UD.
-Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations.
+We now show that this process of efficiency optimization accounts for Greenberg's correlation universals.
+% we examine the computatinally optimized grammars.
+%computationally construct grammars that optimize efficiency, and evaluate these for dependency length and for Greenberg's universals.
+For every one of the 51 languages, we create hypothetical grammars that are computationally optimized for efficiency (see \textit{Materials and Methods}).
+If hypothetical grammars that are optimal for efficiency exhibit these typological patterns, this shows that efficiency optimization predicts Greenberg's word order correlation universals.
+
+The Greenbergian correlations have been documented in great detail by Dryer \cite{dryer1992greenbergian}, drawing on 625 languages.
+For most of these correlations, it is straightforward to test which of our grammars satisfy them (see \textit{Materials and Methods}).
+Collapsing some correlations that cannot be tested individually, and excluding those for which Universal Dependencies does not provide annotation, we obtained eight correlations  (see \textit{Materials and Methods}).
+%We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages.
+%Most correlations can be straightforwardly implemented in Universal Dependencies (UD), allowing us to check which correlations a word order grammar satisfies.
+%Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages or are not annotated reliably in UD.
+%Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations.
 
 These are shown in Table~\ref{table:corr-dryer}.
-We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
-To evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
-The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
+We compare the prevalence of the word order correlations in optimized languages (right) to their prevalence in the real languages (left).
+%To evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
+%The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
 %in all except four languages, for WALS did not have data.
-
 Table~\ref{table:corr-dryer} shows that grammars optimized for efficiency accurately predict correlations to hold at prevalences similar to actual languages.
+
+In Figure~\ref{fig:per-lang}, we show results on the levels of individual languages.
+For every one of the 51 languages, we show which correlations are satisfied by the actual grammar (left) and by the grammar optimized for efficiency (right).
 
@@
 We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
-We are interested in the direction and significance of this effect: 
-If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
-
+%We are interested in the direction and significance of this effect: 
+%If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
 We estimated the posterior distribution of the number of correlations that the optimization process correctly predicts.

### 2019-03-23 second-draft/3-paper-names.tex
revising

@@
 
-\dropcap{F} \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
+%Ray Jackendoff. 2002. Foundations of Language:
+%Brain, Meaning, Grammar, Evolution. Oxford
+%University Press, Oxford, UK
+
+%William Croft and Alan Cruse. 2004. Cognitive
+%Linguistics. Cambridge University Press, Cam-
+%bridge, UK.
+% --> emphasize language use
+
+%Adele Goldberg. 2005. Constructions at work:
+%The nature of generalization in language. Ox-
+%ford University Press, Oxford, UK.
+% --> emphasize usage-based grammar
+% we are all born with the same basic conceptual apparatus, with the
+%same basic communicative demands, and we all live in the physical world with
+%forces of gravity, bodies, and night and day (cf. also LakoV 1987; Webelhuth and
+%Ackerman 1998). Certain semantic/pragmatic functions are so relevant and
+%useful that they can be found in language after language (Croft 2001).
+%Cross-linguistic generalizations that relate form and function are explained
+%by appeal to general cognitive constraints together with the functions of the
+%constructions involved. Constructionists turn to grammar-external explan-
+%ations such as universal functional pressures, iconic principles, and processing
+%and learning constraints to explain such empirically observable cross-linguis-
+%tic generalizations.
+%For example, the fact
+%that that most languages seem to have noun and verb categories may be
+%explained by the existence of corresponding basic semantic categories (Baker
+%2004). Hauser, Chomsky, and Fitch go so far as to suggest that the only innate
+%ability speciWc to language that may be required are the recursive interfaces to
+%phonology and semantics, and they allow that even these may turn out not to
+%be speciWc to language (Hauser, Chomsky, and Fitch 2002).
+
+% Haspelmath 1999 talks about types of explanations
+
+%Joan Bresnan, Ash Asudeh, Ida Toivonen, and
+%Stephen Wechsler. 2016. Lexical-Functional
+%Syntax, 2nd ed. Blackwell, Malden, MA.
+
+%Andrew Radford. 2006. Minimalist syntax re-
+%visited. http://www.public.asu.edu/
+%~gelderen/Radford2009.pdf.
+
+%Givon
+
+% TODO optimality theory and syntax citation?
+%Blutner R. (2000) Some aspects of optimality in natural language interpretation. Journal of Semantics 17: 189–216
+
+% ,lin2017critical
+\dropcap{U}nderstanding and explaining the universal properties of human languages is one of the central goals of Linguistics \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,chomsky1993lectures,...}.
+% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
+Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{berwick1984grammatical,chomsky2005three, hawkins2007processing} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005}.
+%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
+In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which certain syntactic universals can be explained through the computational optimization of grammars for communicative efficiency.
+% information-theoretic
+
+%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
+(TODO cite linguists)
+
+% TODO add a cogsci person to citation chain at the beginning
+
+%\dropcap{F}or decades, researchers in fields ranging from philology to cognitive science to statistical physics have been involved in documenting and trying to explain the universal syntactic and statistical properties of human language \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical}.
+An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
+
+%TODO at some point address head direction idea
+%- Vennemann 1972, 1974ab, 1976
+
+%-Hawkins 1980 limitations of head direction, also Hawkins 1982
+
+%Berwick and Chomsky ‘gene rative process  is optimal  ’, ‘efficient computa  tion ’(71), ‘this newly  emerged  computationalsystem forthought  ...ispe rfe ct, in sofar asSMT is corr ect ’(80).
+‘op -tim al’or ‘effici ent ’or ‘per fect ’.
+
+%Chomsky Three Factors, 2003:
+% --> emphasizes 'computational efficiency'
 
@@
 
+%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}. 
 Natural languages vary in the order in which they express information.
@@
 
+TODO can we find citations from all camps?
+
 Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
-An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-(TODO cite linguists)
-
 A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
@@
 Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
-Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar (CITE):
+Following a long tradition in formal and computational liguistics, we formalize these in the format of dependency trees~\cite{hays1964dependency,melcuk1988dependency,corbett1993heads,tesniere2015elements}:
 This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
+Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.
+
+%The sentences in Figure~\ref{fig:arabic-japanese} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words.
+For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} if the verb, rather than e.g.
+the subject or an indirect object.
+A dependency arc is drawn from a \key{head} (e.g.
+\emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g.
+\emph{tegami-o}).
+Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}.
+Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.
 

### 2019-03-24 second-draft/3-paper-names.tex
revising

@@
 % ,lin2017critical
-\dropcap{U}nderstanding and explaining the universal properties of human languages is one of the central goals of Linguistics \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,chomsky1993lectures,...}.
+\dropcap{U}nderstanding and explaining the universal properties of human languages is one of the central goals of Linguistics \cite{zipf1949human,greenberg1963universals,chomsky1993lectures,...}.
+% behaghel1909beziehungen,
 % \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
+An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
 Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{berwick1984grammatical,chomsky2005three, hawkins2007processing} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005}.
 %Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
-In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which certain syntactic universals can be explained through the computational optimization of grammars for communicative efficiency.
+In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which certain syntactic universals can be explained through the computational optimization of grammars for efficiency.
 % information-theoretic
-
-%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-(TODO cite linguists)
 
@@
 
-The contribution of our work is to show  that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
+Drawing on data from 51 languages, we show  that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
 %In this work, we first provide a formalization of efficiency in the domain of word order. %, drawing on information-theoretic notions.
@@
 
-In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes (REF).
-
-Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking dependency structures as the underlying meanings to be conveyed.
-Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
-
-Efficiency is formalized as a weighted combination of terms representing the amount of information conveyed and the cost of communication.
+In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
+This idea formalizes Zipf's forces of Unification and Diversification, and overall efficiency can be formalized as a weighted combination of terms representing the amount of information conveyed and the cost of communication:
 %The \key{efficiency} of a language is a weighted combination
 \begin{equation}\label{eq:efficiency}
-	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
+	R_{\textit{Eff}} := R_{\textit{Informativity}} - \lambda R_\textit{Cost}
 \end{equation}
 with an interpolation weight $\lambda \in [0,1)$. %In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
-
-In our model, these will depend on the grammar:
+Such notions of efficiency have been proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems.
+%, taking dependency structures as the underlying meanings to be conveyed.
+Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
+%Word meaning \cite{regier2015word}
+
+In our model, both aspects of efficiency will depend on the grammar:
 Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
 This reflects the \key{parseability} of the language.
-Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's principle of listener economy~\cite{frank2012predicting,regier2015word}.
+Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's Force of Diversification~\cite{frank2012predicting,regier2015word}. % also goodman2013knowledge,
 The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
@@
 It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
-This quantity formalizes Zipf's Force of Diversification:
-It can be maximized when trees can be decoded fully unambiguously from utterances.
+In line with Zipf's Force of Diversification, this quantity is maximized when trees can be decoded fully unambiguously from utterances.
 
@@
 \begin{figure} % need Efficiency and (maybe) MLE langs on here
-	\includegraphics[width=0.4\textwidth]{../results/correlations/figures/coverage-ground_eff-circles.pdf}
-	\caption{per-languag predictions}\label{fig:per-lang}
+	\includegraphics[width=0.3\textwidth]{../results/correlations/figures/coverage-ground_eff-circles.pdf}
+	\caption{Across 51 languages, efficiency optimization predicts Greenberg's correlations.
+For every one of the 51 languages, we show which universals are satisfied in the actual grammar (left), and in the grammar optimized for efficiency (right).
+Efficiency optimization replicates the universals at rates similar to the real grammars.}\label{fig:per-lang}
 \end{figure}
@@
 The probability that it predicts less than seven of the correlations is $<10^{-4}$.
-%Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.
+
+We conducted the same analysis for grammars optimized for predictability and parseability separately.
+Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.
 

### 2019-03-24 second-draft/3-paper-names.tex
revising

@@
 
-We first demonstrate that real languages are relatively efficient compared to random baselines, providing evidence that grammars evolve towards efficiency.
+We first demonstrate that the grammars of real languages are relatively efficient compared to random baselines, providing evidence that grammars evolve towards efficiency.
 We compare the efficiency of the actual grammars of the 51 languages from the Universal Dependencies datasets to randomly constructed baseline grammars.
@@
 In Figure~\ref{fig:pareto-plane}, we plot predictability and parsability of the grammars of 51 languages, together with the distribution of random baseline grammars, and the Pareto frontier defined by computationally optimized grammars.
+Languages are attracted towards the Pareto frontier and away from the region of the baseline languages.
 The majority of real languages are above and to the right of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
-X languages improve over the baselines in parsability ($p < ...$ each), Y improve in predictability ($p < ...$ each).
-Z improve in overall efficiency ($p < ...$ each).
-Languages are attracted towards the Pareto frontier and away from the region of the baseline languages.
-%A few languages end up \emph{more} efficient than the computationally optimized languages.
+By a one-sided \textit{t}-test, 84~\% of languages improve over the baselines in parsability ($p < 0.05$), 78~\% improve in predictability ($p < 0.05$).
+98~\% of languages improve on either predictability or parsability ($p<0.05$, with Bonferroni correction).
+At $\lambda = 0.9$, 86~\% of languages improve in overall efficiency ($p<0.05$).
+TODO fix bug with czech data
 

### 2019-03-24 results/plane/plane-2D-best-visualization-large-tests.R

@@ -1,0 +1,115 @@
+library(lme4)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+depl = read.csv("../../grammars/dependency_length/total_summary_funchead_coarse.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+library(tidyr)
+library(dplyr)
+depl = depl %>% filter(grepl("FuncHead", ModelName)) %>% filter(grepl("Coarse", ModelName))
+dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS3 = read.csv("../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS = rbind(dataS, dataS2, dataS3)
+dataP = read.csv("../../grammars/plane/plane-parse.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP2 = read.csv("../../grammars/plane/plane-parse-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP3 = read.csv("../../grammars/plane/plane-parse-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP = rbind(dataP, dataP2, dataP3)
+dataS = dataS %>% group_by(Language, Type, Model) %>% summarise(Surp = mean(Surp, na.rm=TRUE))
+dataP = dataP %>% group_by(Language, Type, Model) %>% summarise(UAS = mean(UAS, na.rm=TRUE), Pars = mean(Pars, na.rm=TRUE))
+dataS = as.data.frame(dataS)
+dataP = as.data.frame(dataP)
+library(lme4)
+summary(lmer(Surp ~ Type + (1|Language), data=dataS %>% filter(grepl("langm", Type))))
+summary(lmer(Surp ~ Type + (1|Language), data=dataS %>% filter(grepl("RL", Type) | grepl("ground", Type))))
+dataS = dataS %>% mutate(Type = as.character(Type))
[…]

### 2019-03-24 second-draft/3-paper-names.tex
revising

@@
 % ,lin2017critical
-\dropcap{U}nderstanding and explaining the universal properties of human languages is one of the central goals of Linguistics \cite{zipf1949human,greenberg1963universals,chomsky1993lectures,...}.
+\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics \cite{zipf1949human,greenberg1963universals,chomsky1993lectures,...}.
 % behaghel1909beziehungen,
@@
 
+Computational studies
+
+word order difficult, requires large data sets
+
 Drawing on data from 51 languages, we show  that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
 %In this work, we first provide a formalization of efficiency in the domain of word order. %, drawing on information-theoretic notions.
+We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.
 
@@
 
+Second, this requires a computational method for estimating predictability and parseability.
+We draw on recent machine learning methods that form the foundation of state-of-the-art natural language processing.
+We estimate predictability using LSTM recurrent neural networks, general sequence models that are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
+We estimate parseability using a generic neural network architecture that casts recovery of dependency structures as a minimum spanning tree problem \citep{dozat2017stanford, zhang2017dependency}.
+See SI for robustness of our results to these choices.
+
 \subsection{Relative efficiency of languages}
@@
 \begin{figure} % need Efficiency and (maybe) MLE langs on here
-	\includegraphics[width=0.3\textwidth]{../results/correlations/figures/coverage-ground_eff-circles.pdf}
-	\caption{Across 51 languages, efficiency optimization predicts Greenberg's correlations.
+%	\includegraphics[width=0.3\textwidth]{../results/correlations/figures/coverage-ground_eff-circles.pdf}
+	\begin{center}
+	\includegraphics[width=0.3\textwidth]{../results/correlations/figures/pred-eff-families.pdf}
+	\end{center}
+	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
+Dark blue: Verb patterner \emph{precedes} object patterner.
+Light blue: Verb patterner \emph{follows} object patterner.}\label{fig:per-lang}
+	% efficiency optimization predicts Greenberg's correlations.
 For every one of the 51 languages, we show which universals are satisfied in the actual grammar (left), and in the grammar optimized for efficiency (right).
-Efficiency optimization replicates the universals at rates similar to the real grammars.}\label{fig:per-lang}
+Efficiency optimization replicates the universals at rates similar to the real grammars.
 \end{figure}
@@
 	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Eff}  \\ 
-	verb & object     & & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+	verb & object     & & &   \\ 
+	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
 %adp. &
@@
 In Figure~\ref{fig:per-lang}, we show results on the levels of individual languages.
-For every one of the 51 languages, we show which correlations are satisfied by the actual grammar (left) and by the grammar optimized for efficiency (right).
+For every one of the 51 languages, we show orders predicted by efficiency, for the eight correlating pairs.
+For each language, we construct a grammar optimized for efficiency with the direction of the verb-object relation fixed to correspond to the actual language.
 
@@
 
-\section{Alternative Theories}
-
-Prior studies have proposed Dependency Length Minimization (DLM) as general principle of word order \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-This is the idea that languages order words so as to minimize the linear distance between words linked in grammatical relationships.
-This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in language processing \cite{gibson1998linguistic,gibson2000dependency}.
-Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-It has also been argued theoretically that this principle should account for Greenberg's correlations, though this link is nontrivial and depends on the precise distributions of tree topologies in syntactic structures.
-As a principle of word order, DLM is also stipulative: There are other processing models, such as surprisal, that are strongly predictive of human processing cost, and make predictions distinct from, and partly conflicting with, those of DLM~\cite{hale2001probabilistic,levy2008expectation}.
-\cite{futrell2017memory} suggests that efficiency predicts DLM.
-
-We show that Efficiency optimization predicts DLM, providing a first-principles functional explanation.
-Figure~\ref{fig:deplength} shows average dependency length for real languages, random baselines, and languages optimized for dependency length and efficiency.
-Optimizing for efficiency lowers dependency length relative to random baselines.
-
-Our results suggest that dependency length minimization is a by-product of efficiency maximization.
+\section{Discussion}
+
+We have directly implemented parseability and predictability using general-purpose statistical models of language and dependency structures.
+Prior functional studies have considered Dependency Length Minimization (DLM) as an approximate formalization of efficiency optimization~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+%Functional studies of word order have proposed Dependency Length Minimization (DLM) as general principle of word order \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+This is the idea that languages order words so as to minimize the overall length of dependencies.
+This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
+TODO also frazier, hawkins.
+Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is itself a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+%It has also been argued theoretically that this principle should account for Greenberg's correlations, though this link is nontrivial and depends on the precise distributions of tree topologies in syntactic structures.
+%As a principle of word order, DLM is stipulative: There are other processing models, such as surprisal, that are strongly predictive of human processing cost, and make predictions distinct from, and partly conflicting with, those of DLM~\cite{hale2001probabilistic,levy2008expectation}.
+%\cite{futrell2017memory} suggests that efficiency predicts DLM.
+Indeed, our experiments show that efficiency optimization predicts DLM, and efficiency optimization appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
+Our work thus provides the first direct confirmation of the idea of DLM as an approximation to efficiency optimization -- DLM emerges as a by-product of efficiency maximization.
+
+%We show that Efficiency optimization predicts DLM, providing a first-principles functional explanation.
+%Figure~\ref{fig:deplength} shows average dependency length for real languages, random baselines, and languages optimized for dependency length and efficiency.
+%Optimizing for efficiency lowers dependency length relative to random baselines.
+
+%Our results suggest that dependency length minimization is a by-product of efficiency maximization.
 In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the dependency length of the real language; on average, the real language is best matched by efficiency optimization.
 
-\begin{figure} % need Efficiency and (maybe) MLE langs on here
-\includegraphics[width=0.4\textwidth]{../results/dependency-length/figures/depl-violin.pdf}
-	\caption{Average dependency length for real languages (blue), random baselines (purple), and for languages optimized for efficiency (green) and explicitly for dependency length (red).
+%\begin{figure} % need Efficiency and (maybe) MLE langs on here
+%\includegraphics[width=0.4\textwidth]{../results/dependency-length/figures/depl-violin.pdf}
+%	\caption{Average dependency length for real languages (blue), random baselines (purple), and for languages optimized for efficiency (green) and explicitly for dependency length (red).
 Dependency length was computed from sentences of lengths 8-12 and averaged over words.}\label{fig:deplength}
-\end{figure}
+%\end{figure}
 
@@
 The efficiency optimization results from Table~\ref{tab:results-dryer} were preregistered: \url{http://aspredicted.org/blind.php?x=8gp2bt}.
+
+Code and results are available at \url{https://github.com/m-hahn/grammar-optim}.
 

### 2019-03-24 results/correlations/plot_families_num.R

@@ -1,0 +1,52 @@
+# tiles-real_eff_large-byObj-restricted-viz-pred-large.R
+
+
+
+D$LanguageNumeric = as.numeric(D$Language_Ordered)
+
+plot_orders_real = ggplot(D %>% filter(Type == "Real Languages"), aes(x = 1, y = LanguageNumeric, group=CoarseDependency)) + 
+  geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
+#  scale_color_gradient() + #values=c("blue", "green")) +
+  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
+                     axis.text.y=element_blank(),axis.ticks=element_blank(),
+                     plot.title=element_text(size=11)) +
+  theme(axis.title=element_blank()) + 
+  theme(legend.position="none") + labs(x=NULL)
+plot_orders_eff = ggplot(D %>% filter(Type == "Efficiency"), aes(x = 1, y = LanguageNumeric, group=CoarseDependency)) + 
+  geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
+#  scale_color_gradient() + #values=c("blue", "green")) +
+  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
+                     axis.text.y=element_blank(),axis.ticks=element_blank(),
+                     plot.title=element_text(size=11)) +
+  theme(axis.title=element_blank()) + 
+  theme(legend.position="none") + labs(x=NULL)
+DLang = unique(D %>% select(Language_Ordered, iso_Ordered, LanguageNumeric))
+DFam = D %>% group_by(Family) %>% summarise(Start = min(LanguageNumeric), End = max(LanguageNumeric), Mean = mean(LanguageNumeric))
+DFam = DFam %>% mutate(Family = ifelse(Family == "Malayo-Sumbawan", "Mal.-Sum.", as.character(Family)))
+DFam = DFam %>% mutate(Family = ifelse(Family == "Sino-Tibetan", "Sin.-Tib.", as.character(Family)))
+plot_langs = ggplot(DLang) 
+plot_langs = plot_langs +  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
[…]

### 2019-03-25 second-draft/3-paper-names.tex
revising

@@
 
-These are shown in Table~\ref{table:corr-dryer}.
+We compare grammars optimized for efficiency with the actual grammars of the 51 languages, as extracted from the orderings attested in the original corpora (see \textit{Materials and Methods}).
+
+The eight correlations are shown in Table~\ref{table:corr-dryer}.
 We compare the prevalence of the word order correlations in optimized languages (right) to their prevalence in the real languages (left).
@@
 For each language, we construct a grammar optimized for efficiency with the direction of the verb-object relation fixed to correspond to the actual language.
+Predictions of efficiency optimization broadly agree with the actual word order patterns.
+%Note that perfect agreement is not to be expected, as grammars will be impacted by factors such as historical development.
+%For instance, Dutch (nld) and Afrikaans (afr) have a prevalence of object-verb order, and
 

### 2019-03-25 results/correlations/plot_families_num2.R

@@ -1,0 +1,64 @@
+# tiles-real_eff_large-byObj-restricted-viz-pred-large.R
+
+
+
+D$LanguageNumeric = as.numeric(D$Language_Ordered)
+
+D$FamilyPrint = as.character(D$Family)
+D = D %>% mutate(FamilyPrint = ifelse(FamilyPrint == "Malayo-Sumbawan", "Mal.-Sum.", as.character(FamilyPrint)))
+D = D %>% mutate(FamilyPrint = ifelse(FamilyPrint == "Sino-Tibetan", "Sin.-Tib.", as.character(FamilyPrint)))
+D = D %>% mutate(FamilyPrint = ifelse(FamilyPrint == "Viet-Muong", "Viet-M.", as.character(FamilyPrint)))
+
+
+DFam = D %>% group_by(FamilyPrint) %>% summarise(Start = min(LanguageNumeric), End = max(LanguageNumeric), Mean = mean(LanguageNumeric))
+
+DFam$yOffset = 0.2*(1:(nrow(DFam))) 
+D$yOffset=NULL
+D = merge(D, DFam %>% select(FamilyPrint, yOffset), by=c("FamilyPrint"))
+
+
+DLang = unique(D %>% select(Language_Ordered, iso_Ordered, LanguageNumeric, yOffset))
+
+
+plot_orders_real = ggplot(D %>% filter(Type == "Real Languages"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
+  geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
+#  scale_color_gradient() + #values=c("blue", "green")) +
+  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
+                     axis.text.y=element_blank(),axis.ticks=element_blank(),
+                     plot.title=element_text(size=11)) +
+  theme(axis.title=element_blank()) + 
+  theme(legend.position="none") + labs(x=NULL)
+plot_orders_eff = ggplot(D %>% filter(Type == "Efficiency"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
+  geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
+#  scale_color_gradient() + #values=c("blue", "green")) +
+  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
[…]

### 2019-03-25 results/correlations/tiles-real_eff_large-byObj-restricted-viz-pred-large.R

@@ -78,8 +78,6 @@ D$Language = ifelse(D$Language == "Old_Church_Slavonic", "OCS", as.character(D$L
 D$Language = ifelse(D$Language == "Ancient_Greek", "A Greek", as.character(D$Language))
 
 
-D_Ground_Mean$Language = ifelse(D_Ground_Mean$Language == "Old_Church_Slavonic", "OCS", as.character(D_Ground_Mean$Language))
-D_Ground_Mean$Language = ifelse(D_Ground_Mean$Language == "Ancient_Greek", "A Greek", as.character(D_Ground_Mean$Language))
 
 
 
@@ -90,6 +88,7 @@ D_Ground$Language = ifelse(D_Ground$Language == "Ancient_Greek", "A Greek", as.c
 D$DirB = as.factor(as.character(round(pmax(0, pmin(1, D$Dir)))))
 D$DirB=D$Dir
 D_Ground_Mean = D_Ground %>% group_by(Language, iso_code, Family) %>% summarise(Dir = sum(DirObj))
+
 D$Language_Ordered = factor(D$Language, levels=unique(D_Ground_Mean[order(D_Ground_Mean$Family),]$Language), ordered=TRUE)
 D$iso_Ordered = factor(D$iso_code, levels=unique(D_Ground_Mean[order(D_Ground_Mean$Family),]$iso_code), ordered=TRUE)
 

### 2019-03-26 second-draft/3-paper-names.tex
revising

@@
 \begin{abstract}
-	TODO rewrite
+%	TODO rewrite
 The universal properties of human languages have been the subject of intense study across disciplines.
 	We report novel computational and corpus-based evidence for the hypothesis that a prominent subset of these universal properties---those related to word order---result from a process of optimization for efficient communication among humans.
-	TODO explain
+	We computationally formalize functional theories of language efficiencies using information theory and state-of-the-art machine learning techniques.
+Applying these methods to data from 51 languages, we show that evolution towards efficiency results in word order patterns that predict a large subset of the major word order correlations reported in the linguistic typological literature.
+%	TODO explain
 	%We develop a probabilistic, differentiable model of word order grammars: the means by which different languages convert underlying hierarchical structures into strings of words.
 	%We show how the parameters of a word order grammar can be optimized for efficiency of information transfer, quantified as a tradeoff between incremental predictability and mutual information with latent tree structures.
-	Applying these grammars to tree structures found in dependency corpora from 51 languages, we show that optimizing the grammar parameters for efficiency results in word order patterns that predict a large subset of the major word order correlations reported in the linguistic typological literature, and subsumes the predictions of previous heuristic theories such as dependency length minimization.
+%	Applying these grammars to tree structures found in dependency corpora from 51 languages, we show that optimizing the grammar parameters for efficiency results in word order patterns that predict a large subset of the major word order correlations reported in the linguistic typological literature, and subsumes the predictions of previous heuristic theories such as dependency length minimization.
 \end{abstract}
@@
 An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{berwick1984grammatical,chomsky2005three, hawkins2007processing} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005}.
+Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three, berwick1984grammatical,hawkins2007processing} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005}.
 %Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
@@
 
-TODO maybe Arabic becomes English?
-
 This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
@@
 %adpositional phrases (`to a friend') relative to the verb.
+Greenberg also found that the order between verb and object is correlated with other aspects of a language's word order, such the order of verb and adpositional phrase (`wrote -- to friend' in Arabic vs `friend to -- wrote' in Japanese), and that of noun and genitive (`book -- of friend' in Arabic, `friend of -- book' in Japanese).
 Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
@@
 
-TODO can we find citations from all camps?
-
-Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
+%TODO can we find citations from all camps?
+
+Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study~\cite{lehmann1973structural, vennemann1974theoretical,jackendoff1977x,frazier1985syntactic,chomsky1988language, dryer1992greenbergian, hawkins1994performance}.
 A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
@@
 
-We examined whether
-(1) the grammars of natural languages evolve towards optimizing efficiency of communication, and
-(2) this process of optimization accounts both for Greenberg's order correlation universals.
+We examined 
+(1) whether the grammars of natural languages evolve towards optimizing efficiency of communication, and
+(2) whether this process of optimization accounts  for Greenberg's order correlation universals.
 %We formalize word order grammars as specifying how different types of dependents (subjects, objects, etc.) are ordered, employing a cross-linguistically meaningful inventory of relations~\cite{nivre2015universal}.
+
 Answering these questions requires a sample of dependency structures as actually used by speakers across different languages.
@@
 
-Second, this requires a computational method for estimating predictability and parseability.
+Second, addressing these questions requires a computational method for estimating predictability and parseability.
 We draw on recent machine learning methods that form the foundation of state-of-the-art natural language processing.
@@
 We estimate parseability using a generic neural network architecture that casts recovery of dependency structures as a minimum spanning tree problem \citep{dozat2017stanford, zhang2017dependency}.
-See SI for robustness of our results to these choices.
+See SI for details and for robustness of our results to these choices.
 
@@
 The majority of real languages are above and to the right of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
-By a one-sided \textit{t}-test, 84~\% of languages improve over the baselines in parsability ($p < 0.05$), 78~\% improve in predictability ($p < 0.05$).
-98~\% of languages improve on either predictability or parsability ($p<0.05$, with Bonferroni correction).
-At $\lambda = 0.9$, 86~\% of languages improve in overall efficiency ($p<0.05$).
-TODO fix bug with czech data
+100~\% of languages improve on either predictability or parsability ($p<0.05$ each, by one-sided $t$-test, with Bonferroni correction).
+%By a one-sided \textit{t}-test, 
+84~\% of languages improve over the baselines in parsability ($p < 0.05$ each), 78~\% improve in predictability ($p < 0.05$ each).
+%At $\lambda = 0.9$, 86~\% of languages improve in overall efficiency ($p<0.05$ each).
+%TODO fix bug with czech data
 
@@
 	\begin{center}
-	\includegraphics[width=0.3\textwidth]{../results/correlations/figures/pred-eff-families.pdf}
+	\includegraphics[width=0.32\textwidth]{../results/correlations/figures/pred-eff-families-2.pdf}
 	\end{center}
@@
 %		\setlength\tabcolsep{1.5pt}
-\begin{tabular}{|ll|c|cc|}
+\begin{tabular}{c|ll|c|cc|ccc}
 	\hline
 	%&
-	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Eff}  \\ 
-	verb & object     & & &   \\ 
-	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%adp. &
-	adposition    &    NP       
+	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Optimized} & \\ 
+	&	verb & object     & & &   \\ 
+	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
 		& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_case.pdf} } %\textbf{86}  }
@@
 	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_case.pdf}   }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_case.pdf}     }  \\
-	\emph{to}            & \emph{a friend} &&&\\ \hline
-copula    &    NP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_case.pdf}     } & \\
+	&		\emph{to}            & \emph{a friend} &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
 		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_cop.pdf} }% \textbf{94} }
@@
 	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_cop.pdf}   }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_cop.pdf}   }  \\
-	\emph{is}        & \emph{a friend}  &&&\\ \hline
-auxiliary    &    VP       
+		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_cop.pdf}   } & \\
+	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
 		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-aux.pdf}}
@@
 	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-aux.pdf}  }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-aux.pdf}  }  \\
-	\emph{has}          & \emph{written}  &&&\\ \hline
-noun    &    genitive      
+		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-aux.pdf}  } & \\
+	&	\emph{has}          & \emph{written}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
 		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-nmod.pdf}}
@@
 	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-nmod.pdf}  }  \\
-	\emph{friend} &  \emph{of John}  &&&\\ \hline
-noun    &    relative clause      
+		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-nmod.pdf}  } & \\
+	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
 		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-acl.pdf}}
@@
 	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-acl.pdf}  }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-acl.pdf}  }  \\
-	\emph{books} & \emph{that you read}  &&&\\ \hline
-complementizer    &    S        
+		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-acl.pdf}  } & \\
+	&	\emph{books} & \emph{that you read}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
 		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_mark.pdf}}
@@
 	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_mark.pdf} }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_mark.pdf}  }  \\
-	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-verb    &    PP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_mark.pdf}  }  &\\
+	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
 		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-obl.pdf}}
@@
 	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-obl.pdf} }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-obl.pdf}  }  \\
-	\emph{went} & \emph{to school}  &&&\\ \hline
-want    &    VP        
+		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-obl.pdf}  } &  \\
+	&	\emph{went} & \emph{to school}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
 		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-xcomp.pdf}}
@@
 	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
-	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-xcomp.pdf}  }  \\
-	\emph{wants}   &  \emph{to leave}  &&&\\ \hline
+		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-xcomp.pdf}  } & \\
+	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
  \hline
@@
 %	AP &
-    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
 \end{tabular}
@@
 For every one of the 51 languages, we create hypothetical grammars that are computationally optimized for efficiency (see \textit{Materials and Methods}).
-If hypothetical grammars that are optimal for efficiency exhibit these typological patterns, this shows that efficiency optimization predicts Greenberg's word order correlation universals.
+In all results reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
+(See SI appendix section 5 for mathematical justification and for robustness to other choices).
+We test whether the process of efficiency optimization can explain these typological patterns. %, this shows that efficiency optimization predicts Greenberg's word order correlation universals.
 
@@
 Prior functional studies have considered Dependency Length Minimization (DLM) as an approximate formalization of efficiency optimization~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-%Functional studies of word order have proposed Dependency Length Minimization (DLM) as general principle of word order \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
 This is the idea that languages order words so as to minimize the overall length of dependencies.
 This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
-TODO also frazier, hawkins.
 Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is itself a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+Indeed, directly modeling efficiency optimization predicts DLM, and appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
+%Our work thus provides the first direct confirmation of the idea of DLM as an approximation to efficiency optimization -- DLM emerges as a by-product of efficiency maximization.
+
+%We show that Efficiency optimization predicts DLM, providing a first-principles functional explanation.
+%%Functional studies of word order have proposed Dependency Length Minimization (DLM) as general principle of word order \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+%Figure~\ref{fig:deplength} shows average dependency length for real languages, random baselines, and languages optimized for dependency length and efficiency.
+%Optimizing for efficiency lowers dependency length relative to random baselines.
+
 %It has also been argued theoretically that this principle should account for Greenberg's correlations, though this link is nontrivial and depends on the precise distributions of tree topologies in syntactic structures.
@@
 %\cite{futrell2017memory} suggests that efficiency predicts DLM.
-Indeed, our experiments show that efficiency optimization predicts DLM, and efficiency optimization appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
-Our work thus provides the first direct confirmation of the idea of DLM as an approximation to efficiency optimization -- DLM emerges as a by-product of efficiency maximization.
-
-%We show that Efficiency optimization predicts DLM, providing a first-principles functional explanation.
-%Figure~\ref{fig:deplength} shows average dependency length for real languages, random baselines, and languages optimized for dependency length and efficiency.
-%Optimizing for efficiency lowers dependency length relative to random baselines.
 

### 2019-03-26 results/correlations/plot_families_num2.R

@@ -23,21 +23,30 @@ DLang = unique(D %>% select(Language_Ordered, iso_Ordered, LanguageNumeric, yOff
 plot_orders_real = ggplot(D %>% filter(Type == "Real Languages"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
   geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
 #  scale_color_gradient() + #values=c("blue", "green")) +
-  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
+  theme_classic() +
+  #theme_bw() + 
+  theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
                      axis.text.y=element_blank(),axis.ticks=element_blank(),
                      plot.title=element_text(size=11)) +
   theme(axis.title=element_blank()) + 
-  theme(legend.position="none") + labs(x=NULL)
+  theme(legend.position="none") + labs(x=NULL) +
+  scale_x_continuous(breaks = NULL) +
+  scale_y_continuous(breaks = NULL)
+
 plot_orders_eff = ggplot(D %>% filter(Type == "Efficiency"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
   geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
 #  scale_color_gradient() + #values=c("blue", "green")) +
-  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
+  theme_classic() +
+ theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
                      axis.text.y=element_blank(),axis.ticks=element_blank(),
                      plot.title=element_text(size=11)) +
   theme(axis.title=element_blank()) + 
-  theme(legend.position="none") + labs(x=NULL)
+  theme(legend.position="none") + labs(x=NULL) +
+  scale_x_continuous(breaks = NULL) +
+  scale_y_continuous(breaks = NULL)
+
 plot_langs = ggplot(DLang) 
-plot_langs = plot_langs +  theme_bw() 
[…]

### 2019-03-26 results/plane/plane-2D-best-visualization-large.R

@@ -286,5 +286,25 @@ ggsave(plot, file="pareto-plane-iso-best-balanced-legend-viz-2.pdf")
 
 
 
+plot = ggplot(dataPlot)  
+plot = plot + theme_bw() 
+#plot = plot + geom_segment(aes(x=-Pars_z, y=-Surp_z, xend=-Pars_z_end, yend = -Surp_z_end, color=TypeN, group=TypeN), size=0.6)
+plot = plot + geom_density_2d(data=dataPlot %>% filter(grepl("Baseline", TypeN)), aes(x=-Pars, y=-Surp, color=TypeN, group=TypeN), size=0.3)
+plot = plot + geom_path(data=dataPlot %>% filter(TypeN %in% c("Parseability", "Efficiency", "Predictability")), aes(x=-ParsMean, y=-SurpMean, group=1), color="gray", size=5)
+#plot = plot + geom_point(data=dataPlot, aes(x=-ParsMean, y=-SurpMean, color=TypeN, group=TypeN), size=6)
+plot = plot + scale_x_continuous(name="Parsability") + scale_y_continuous(name="Predictability")
+plot = plot + geom_text(data=dataPlot %>% filter(grepl("Real", TypeN)), aes(x=-Pars, y=-Surp, color=TypeN, group=TypeN, label=iso_code),hjust=0.8, vjust=0.8)
+plot = plot + theme(legend.title = element_blank())  
+plot = plot + guides(color=guide_legend(nrow=2,ncol=4,byrow=TRUE)) 
+plot = plot + theme(legend.title = element_blank(), legend.position="bottom")
+plot = plot + scale_colour_manual(name = "TypeN",values = myColors)
+plot = plot + theme(axis.title.x = element_text(size=17))
+plot = plot + theme(axis.title.y = element_text(size=17))
+plot = plot + theme(legend.text = element_text(size=12))
+#plot = plot + annotate("text", x=-1.5, y=-1.3, label = "Baselines", hjust=0, size=8, color="blue")
+
+ggsave(plot, file="pareto-plane-iso-best-balanced-legend-viz-3.pdf")
+
+
 
 

### 2019-03-26 second-draft/3-paper-names.tex
revising

@@
     \centering
-    \includegraphics[width=0.45\textwidth]{figures/grammars-colors-trees.png}
+    \includegraphics[width=0.45\textwidth]{figures/grammars-langs.png}
 	\caption{Grammars define consistent ordering rules for dependency structures.
 Here, Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb.
-\textbf{TODO This figure is just a draft sketch so far.}}
+Grammars 1 and 3 conform to the Greenberg correlations and are common around the world; Grammars 2 and 4 are rare or impossible.}
 	\label{fig:grammars}
@@
 	%&
-	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Optimized} & \\ 
+	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
 	&	verb & object     & & &   \\ 
@@
 Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-For each correlation, we provide an English example, and show the distribution of real languages, of languages optimized for DLM, and of languages optimized for efficiency.
-The process of optimization fir efficiency and DLM explains all eight universals.
+For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
 %Given the statistical nature of the correlations, not every language satisfies every one of them.
@@
 We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
-%We are interested in the direction and significance of this effect: 
-%If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
 We estimated the posterior distribution of the number of correlations that the optimization process correctly predicts.
-The estimated posterior probability that efficiency optimization predicts less than all eight correlations is $0.0242$.
+We conducted the same analysis for grammars optimized for predictability and parseability separately~(Figure~\ref{fig:posterior}).
+%Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.
+Efficiency optimization predicts all eight correlations to be universal with posterior probability $0.9911$.
+%The estimated posterior probability that efficiency optimization predicts less than all eight correlations is $0.0242$.
 The probability that it predicts less than seven of the correlations is $<10^{-4}$.
-
+Predictability and parseability individually also predict six to eight correlations.
+
+\begin{figure}
+	\begin{center}
+	\includegraphics[width=0.48\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
+	\end{center}
+	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
+\end{figure}
+
+TODO better than this:
+
+We further compared to the predictions made individually by optimizing for predictability and parseability.
+We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
+We estimated the posterior distribution of the number of correlations that the optimization process correctly predicts.
 We conducted the same analysis for grammars optimized for predictability and parseability separately.
-Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.
+We used all eight runs of optimization to maximize statistical precision.
+For every one of the correlations, we report the posterior estimated.
+(These numbers control for random variation between language families, and are thus not directly comparable to data from the 51 languages).
 

### 2019-03-26 results/correlations/plot_families_num2.R

@@ -112,3 +112,9 @@ plot
 
 ggsave(plot=plot, "figures/pred-eff-families-2.pdf", width=4, height=8)
 
+D2 = (D %>% select(Family, Language, CoarseDependency, DirB, Type) %>% spread(Type, DirB) %>% rename(Real = 'Real Languages') %>% rename(Predicted = Efficiency))
+
+D2$Agree = (D2$Real == D2$Predicted)
+summary(glmer(Agree ~ (1|CoarseDependency) + (1|Family), data=D2, family="binomial"))
+mean(D2$Agree)
+

### 2019-03-26 results/correlations/visualize-joint-evaluation-together-large-three.R

@@ -1,0 +1,44 @@
+
+library(dplyr)
+library(tidyr)
+
+
+u_langmod = read.csv("/home/user/CS_SCR/posteriors/posterior-10-langmod.csv") %>% mutate(Type = "Predictability")
+u_parser = read.csv("/home/user/CS_SCR/posteriors/posterior-10-parseability.csv") %>% mutate(Type = "Parseability")
+u_efficiency = read.csv("/home/user/CS_SCR/posteriors/posterior-10-efficiency-large.csv") %>% mutate(Type = "Efficiency")
+
+
+u = rbind(u_langmod, u_parser, u_efficiency)
+
+#u$satisfied = 10 - ((u$b_acl_Intercept < 0) + (u$b_advmod_Intercept > 0)  + (u$b_aux_Intercept > 0 ) + (u$b_liftedcase_Intercept < 0 ) + (u$b_liftedcop_Intercept < 0 ) + (u$b_liftedmark_Intercept < 0 ) + (u$b_nmod_Intercept < 0 ) + (u$b_nsubj_Intercept > 0 ) + (u$b_obl_Intercept < 0 ) + (u$b_xcomp_Intercept < 0 ))
+
+u$satisfied = 8 - ((u$b_acl_Intercept < 0) + (u$b_aux_Intercept > 0 ) + (u$b_liftedcase_Intercept < 0 ) + (u$b_liftedcop_Intercept < 0 ) + (u$b_liftedmark_Intercept < 0 ) + (u$b_nmod_Intercept < 0 ) + (u$b_obl_Intercept < 0 ) + (u$b_xcomp_Intercept < 0 ))
+
+
+library(ggplot2)
+
+v = u %>% group_by(Type) %>% summarise(SamplesNum = NROW(satisfied))
+u = merge(u, v, by=c("Type"))
+data2 = u %>% group_by(Type, satisfied) %>% summarise(SamplesNum = mean(SamplesNum), posterior = NROW(satisfied)) %>% mutate(posteriorProb = posterior/SamplesNum)
+u = NULL
+
+#plot = ggplot(data = data, aes(x=satisfiedCount)) + geom_histogram() + theme_bw() + xlim(0,10.5) + ggtitle("Dependency Length") + xlab("Number of Predicted Correlations") + ylab("Number of Posterior Samples")
+
+
+data2$Type <- factor(data2$Type, levels = c("Predictability", "Parseability", "Efficiency"))
+
[…]

### 2019-03-27 second-draft/3-paper-names.tex
revising

@@
 
+%
+%
+
 \begin{table}
@@
 	\end{center}
-\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
+\caption{OPTION 1: Efficiency optimization accurately predicts the Greenbergian correlations.
 Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
@@
 \end{table}
+
+\begin{table}
+	\begin{center}
+%		\setlength\tabcolsep{1.5pt}
+\begin{tabular}{c|ll|c|cc|ccc}
+	\hline
+	%&
+	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
+	&	verb & object     & & &   \\ 
+	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
+				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_case.pdf}       }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_case.pdf}   }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
+	&		\emph{to}            & \emph{a friend} &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_cop.pdf} } % \textbf{53}     }
+	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_cop.pdf}   }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
+	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-aux.pdf}     }
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-aux.pdf}  }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
+	&	\emph{has}          & \emph{written}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-nmod.pdf}   }
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
+	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
+	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-acl.pdf}   }
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-acl.pdf}  }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
+	&	\emph{books} & \emph{that you read}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_mark.pdf}     }
+	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_mark.pdf} }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
+	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-obl.pdf}     }
+	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-obl.pdf} }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
+	&	\emph{went} & \emph{to school}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-xcomp.pdf}     }
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
+	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+\end{tabular}
+	\end{center}
+\caption{OPTION 2: Efficiency optimization accurately predicts the Greenbergian correlations.
+Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
+%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%Not all correlations are satisfied by every natu
+}\label{table:corr-dryer}
+\end{table}
+
+\begin{table*}
+	\begin{center}
+%		\setlength\tabcolsep{1.5pt}
+\begin{tabular}{c|ll|c|c|cc|ccccccccccc}
+	\hline
+	%&
+	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Efficiency} &  \multirow{2}{*}{Predictability} &  \multirow{2}{*}{Parseability} & \\ 
+	&	verb & object     & & &   \\ 
+	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
+				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_case.pdf}     }  & \\
+	&		\emph{to}            & \emph{a friend} &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     }
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_cop.pdf}     } 
+& \\
+	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_aux.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_aux.pdf}     } & \\
+	&	\emph{has}          & \emph{written}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_nmod.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_nmod.pdf}     } & \\
+	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_cop.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_acl.pdf}     } & \\
+	&	\emph{books} & \emph{that you read}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_mark.pdf}     } & \\
+	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_obl.pdf}     } & \\
+	&	\emph{went} & \emph{to school}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_xcomp.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_xcomp.pdf}     } & \\
+	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+\end{tabular}
+	\end{center}
+\caption{OPTION 3: Efficiency optimization accurately predicts the Greenbergian correlations.
+Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
+%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%Not all correlations are satisfied by every natu
+}\label{table:corr-dryer}
+\end{table*}
 

### 2019-03-27 results/correlations/visualize-posterior-2-large.R

@@ -76,7 +76,6 @@ E = data %>% group_by(Language, Family, Dependency) %>% summarise(Agree = mean(A
 
 dependencies = c("acl", "advcl", "advmod", "amod", "appos", "aux", "ccomp", "compound", "conj", "csubj", "dep", "det", "discourse", "dislocated", "expl", "fixed", "flat", "goeswith", "iobj", "lifted_case", "lifted_cc", "lifted_cop", "lifted_mark", "list", "nmod", "nsubj", "nummod", "obl", "orphan", "parataxis", "reparandum", "vocative", "xcomp")
 
-dependencies =  c("acl", "advmod", "aux", "lifted_case", "lifted_cop", "lifted_mark", "nmod", "nsubj", "obl", "xcomp")
 
 
 #balanced = read.csv("results-ground-agree.tsv", sep="\t")
@@ -94,7 +93,7 @@ library(tidyr)
 library(ggplot2)
 
 
-
+cat("\nReading posterior samples\n")
 parse = read.csv("/home/user/CS_SCR/posteriors/posterior-10-parseability.csv")
 two = read.csv("/home/user/CS_SCR/posteriors/posterior-10-efficiency-large.csv")
 langmod = read.csv("/home/user/CS_SCR/posteriors/posterior-10-langmod.csv")
@@ -128,6 +127,46 @@ data2 = data2 %>% mutate(Prevalence = case_when(Dependency == "aux" ~ 1-Prevalen
 
 #  plot = ggplot(data=data2) + geom_density(aes(x=Prevalence, y=..scaled.., fill=Type, group=Type), alpha=.5) + xlim(0,1) + geom_bar(data=D, stat="identity", width = 0.01, aes(x=Agree, y=1, fill=Type, group=Type)) + facet_wrap( ~ Dependency, ncol=1) + geom_bar(data=E, stat="identity", width = 0.01, aes(x=Agree, y=1, fill=Type, group=Type))
 #  ggsave(plot, file=paste("posterior/posterior_joint_with_ud_balanced" , ".pdf", sep=""))
+
+
+#for(type in c("
+  type = "Efficiency"
+  dependency = "acl"
+
+  dependencies =  c("acl", "aux", "lifted_case", "lifted_cop", "lifted_mark", "nmod", "obl", "xcomp")
+for(type in c("Efficiency", "Predictability", "Parseability")) {
+	if(type == "Efficiency") {
+		color = "blue"
+	} else if(type == "Predictability") {
+		color = "red"
+	} else if (type == "Parseability") {
[…]

### 2019-03-31 second-draft/3-paper-names.tex
revising

@@
 An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three, berwick1984grammatical,hawkins2007processing} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005}.
+Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005}.
 %Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
@@
 
-Computational studies
-
-word order difficult, requires large data sets
-
-Drawing on data from 51 languages, we show  that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
+In this work, we apply this approach to the problem of explaining Greenberg's word order correlation universals.
+Drawing on large sets of text data from 51 languages, we show  that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
+We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.
+
 %In this work, we first provide a formalization of efficiency in the domain of word order. %, drawing on information-theoretic notions.
-We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.
 
@@
 
-TODO at some point need to foreshadow more of the Greenberg correlations
+%TODO at some point need to foreshadow more of the Greenberg correlations
 
@@
 When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
-In natural language, this order is largely determined by the \key{grammar} of the language (CITE):
-The speaker will order it as (1) in English, and as (2) in Japanese.
-
-(1)
-\begin{dependency}[theme=simple]
-    \begin{deptext}
-    Mary \& saw \& John \\
-    \end{deptext}
-    \depedge{2}{1}{nsubj}
-    \depedge{2}{3}{obj}
-\end{dependency}
-(2)
-\begin{dependency}[theme=simple]
-    \begin{deptext}
-    Mary \& John  \& saw \\
-    \end{deptext}
-    \depedge{3}{1}{nsubj}
-    \depedge{3}{2}{obj}
-\end{dependency}
-
-The listener receives a sentence -- a linear string of words -- and decodes the underlying message in accordance with the grammar of the language (right).
+In natural language, this order is largely determined by the \key{grammar} of the language (CITE).
+The speaker will order it as \emph{Mary saw John} (Figure~\ref{fig:comm}, top) in English, and as \emph{Mary John saw} (Figure~\ref{fig:comm}, bottom) in Japanese.
+%
+%(1)
+%\begin{dependency}[theme=simple]
+%    \begin{deptext}
+%    Mary \& saw \& John \\
+%    \end{deptext}
+%    \depedge{2}{1}{nsubj}
+%    \depedge{2}{3}{obj}
+%\end{dependency}
+%(2)
+%\begin{dependency}[theme=simple]
+%    \begin{deptext}
+%    Mary \& John  \& saw \\
+%    \end{deptext}
+%    \depedge{3}{1}{nsubj}
+%    \depedge{3}{2}{obj}
+%\end{dependency}
+%
+The listener receives a sentence -- a linear string of words -- and decodes the underlying message in accordance with the grammar of the language (Figure~\ref{fig:comm}, right).
 %Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
@@
 
-The grammars of natural languages provide consistent ordering rules for all syntactic relations.
 %All natural languages have some degree of word order regularities that are specified in their grammar.
@@
 %While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
+The grammars of natural languages provide consistent ordering rules for all syntactic relations.
 This is illustrated in Figure~\ref{fig:grammars}:
@@
 
+%In information-theoretic formalizations of Zipf's ideas, the efficiency of a 
 In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
@@
 %how many distinct sentences there are.
+
+%CITE other syntactic work? LFG
+%TODO the following sentences should go somewhere
+%By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
+%The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
+%Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
 
@@
 
+TODO the following paragraphs depend on what visualization we choose
+
 We compare grammars optimized for efficiency with the actual grammars of the 51 languages, as extracted from the orderings attested in the original corpora (see \textit{Materials and Methods}).
@@
 %For instance, Dutch (nld) and Afrikaans (afr) have a prevalence of object-verb order, and
+
+\begin{figure}
+	\begin{center}
+	\includegraphics[width=0.48\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
+	\end{center}
+	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
+\end{figure}
 
@@
 Predictability and parseability individually also predict six to eight correlations.
-
-\begin{figure}
-	\begin{center}
-	\includegraphics[width=0.48\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
-	\end{center}
-	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
-\end{figure}
 

### 2019-04-03 second-draft/3-paper-names.tex
revising

@@
 % ,lin2017critical
-\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics \cite{zipf1949human,greenberg1963universals,chomsky1993lectures,...}.
+\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics \cite{greenberg1963universals,chomsky1993lectures,...}.
 % behaghel1909beziehungen,
@@
 An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005}.
+Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural}.
+TODO improve this
 %Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
@@
 % information-theoretic
+
+foreshadow functional explanations?
+
+- these linguistic patterns reflect general principles of communication
+
+- big idea has been around for a long time
+
+- here for the first time direct computational test
+
+- we can predict broad patterns, and make language-specific predictions based on these general principles
 
@@
 Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
-The idea that language results from the tension between these two pressures has a long and fruitful history.
-%\cite{horn-toward-1984}, RSA, Color naming
+The idea that language results from the tension between these two pressures has a long and fruitful history in theoretical studies of language~\cite{gabelentz1901sprachwissenschaft,horn1984toward}.
+%\cite{horn1984toward}, RSA, Color naming
 %
@@
 %
-This idea has been shown to provide a formalization and derivation of pragmatic inference \cite{horn-toward-1984, frank2012predicting}.
-It also has been shown to account for the distribution of word meanings across languages, such as color naming~\cite{zaslavsky2018efficient}.
-
-%
-
-%
-%
-
+
+More recent work has used an information-theoretic formalization to computationally test these ideas in various domains of language.
+This idea has been shown to provide a formalization and derivation of pragmatic inference \cite{frank2012predicting}.
+It also has been shown to account for the distribution of word meanings across languages, such as color naming~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship},
+and for basic statistical properties of languages~\cite{ferreri2003least}.
+Common to these accounts is that the Force of Diversification is formalized as maximizing the amount of information that linguistic forms provide about meanings, whereas the Force of Unification is formalized as reducing the diversity of forms.
+
+While it has been suggested that these ideas should also apply to grammar~\cite{hawkins2004efficiency}, testing them in syntax and word order is more difficult, as this requires both large amounts of data representative of language use across languages and cultures, and computational methods for estimating the efficiency of entire languages.
 %TODO somewhere start calling this Efficiency
 %TODO explain why surprisal. some production/speaker work. (plan reuse)
-
 %In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
@@
 %Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
-
 In this work, we apply this approach to the problem of explaining Greenberg's word order correlation universals.
 Drawing on large sets of text data from 51 languages, we show  that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
+Furthermore, efficiency optimization makes fine-grained per-language predictions about individual syntactic patterns.
 We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.
@@
 In this model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
-Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
+Applying this model to word order, we take the message to consist of people, objects, events, and a set of syntactic and semantic relations that hold between them~\cite{fodor1975language,jackendoff1983semantics,pinker1990natural}.
 Following a long tradition in formal and computational liguistics, we formalize these in the format of dependency trees~\cite{hays1964dependency,melcuk1988dependency,corbett1993heads,tesniere2015elements}:
@@
 
-\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Noah Goodman, Chris Manning, and audiences at CAMP 2018 for helpful discussion.}
+\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Noah Goodman, Chris Manning, and audiences at CAMP 2018 and CUNY 2019 for helpful discussion.}
 

### 2019-04-05 second-draft/3-paper-names.tex
revising

--- 2019-03-31 second-draft/3-paper-names.tex
@@
 % ,lin2017critical
-\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics \cite{zipf1949human,greenberg1963universals,chomsky1993lectures,...}.
+\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
 % behaghel1909beziehungen,
@@
 An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005}.
+Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural}.
 %Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
-In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which certain syntactic universals can be explained through the computational optimization of grammars for efficiency.
+In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which a large class of language universals can be explained through the computational optimization of grammars for efficiency.
 % information-theoretic
+
+%foreshadow functional explanations?
+%
+%- these linguistic patterns reflect general principles of communication
+%
+%- big idea has been around for a long time
+%
+%- here for the first time direct computational test
+%
+%- we can predict broad patterns, and make language-specific predictions based on these general principles 
+%
 
@@
 Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
-Both sentences contain a verb meaning `wrote', a noun expressing  `letter', and a phrase translating to `to a friend'.
-In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
+Both sentences contain a verb meaning `wrote', a noun expressing  the object `letter', and an adpositional phrase translating to `to a friend'.
+%In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
 Yet, the order of these words are entirely different in the two languages:
 %However, the languages differ in the order in which these words and phrases are arranged: 
-In Arabic, the verb stands at the beginning, followed by the object and the adpositional phrase.
-`To' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
-In Japanese, the verb stands at the end; object and adpositional phrase precede it; `to' is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
+The verb stands at the beginning in Arabic, and at the end in Japanese.
+Arabic expresses `to' by a \emph{preposition} (\emph{preceding} the noun `friend'); Japanese uses a \emph{postposition} (\emph{following} it). % and a postposition in Japanese, following `friend'.
+%In Arabic, the verb stands at the beginning, followed by the objecat and the adpositional phrase.
+%In Arabic, `to' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
+%In Japanese, it is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
 %What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
@@
 %It turns out that these are deep and stable regularities:
+
 This variation reflects a deep and stable regularity:
@@
 
-Researchers working in the functional paradigm have proposed a range of criteria that languages should meet in order to enable efficient communication and processing.
+%Researchers working in the functional paradigm have proposed a range of criteria that languages should meet in order to enable efficient communication and processing.
 These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
 
-Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between the Forces of Unification and Diversification.
+Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between the Forces of \emph{Unification} and \emph{Diversification}.
 The Force of Unification is a pressure to reduce the complexity of the language to make production and processing as easy as possible.
@@
 These two forces act in opposing directions:
-Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
-The idea that language results from the tension between these two pressures has a long and fruitful history.
-%\cite{horn-toward-1984}, RSA, Color naming
-%
-%In the domain of word order, 
-%
-%TODO introduce the terms \key{cost} and \key{informativity}
-%
-This idea has been shown to provide a formalization and derivation of pragmatic inference \cite{horn-toward-1984, frank2012predicting}.
-It also has been shown to account for the distribution of word meanings across languages, such as color naming~\cite{zaslavsky2018efficient}.
-
-%
-
-%
-%
-
+Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required in order to provide higher informativity.
+The idea that many properties of language arise from the tension between these two pressures has a long and fruitful history in theoretical studies of language~\cite{gabelentz1901sprachwissenschaft,horn1984toward, schwartz1997dispersion}.
+
+Recent work has used an information-theoretic formalization to computationally test these ideas in various domains of language, showing that they predict
+both basic statistical properties of languages~\cite{ferreri2003least},
+and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.
+Common to these accounts is the idea of maximizing the amount of information that linguistic forms provide about meanings, while constraining the diversity of forms.
+
+While it has been suggested that these ideas should also apply to grammar~\cite{hawkins2004efficiency}, testing these accounts is more difficult in the domain of grammar, as this requires both large amounts of data representative of language use across languages and cultures, and computational methods for estimating the efficiency of entire languages.
 %TODO somewhere start calling this Efficiency
 %TODO explain why surprisal. some production/speaker work. (plan reuse)
-
 %In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
@@
 %Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
-
-In this work, we apply this approach to the problem of explaining Greenberg's word order correlation universals.
-Drawing on large sets of text data from 51 languages, we show  that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
-We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.
+In this work, we address these challenges by combining large-scale text data from 51 languages with state-of-the-art machine learning techiques to estimate the communicative efficiency of grammar.
+We apply this approach to the problem of explaining Greenberg's word order correlation universals.
+We first show that the word order of languages has evolved to optimize efficiency.
+We then show that efficiency optimization accounts for the Greenbergian word order correlations, and makes fine-grained per-language predictions about individual syntactic patterns.
+%We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.
 
@@
 In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
-This idea formalizes Zipf's forces of Unification and Diversification, and overall efficiency can be formalized as a weighted combination of terms representing the amount of information conveyed and the cost of communication:
+%This idea formalizes Zipf's forces of Unification and Diversification
+Overall efficiency can be formalized as a weighted combination of terms representing the amount of information conveyed and the cost of communication:
 %The \key{efficiency} of a language is a weighted combination
@@
 with an interpolation weight $\lambda \in [0,1)$. %In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
-Such notions of efficiency have been proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems.
+This objective function has been proposed in prior computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship}.
+
+%Such notions of efficiency have been proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems.
 %, taking dependency structures as the underlying meanings to be conveyed.
-Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
+%Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
 %Word meaning \cite{regier2015word}
@@
 %how many distinct sentences there are.
+
+With these choices, our objective function is equivalent to that proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems.
+%, taking dependency structures as the underlying meanings to be conveyed.
+These works mainly differ in the cost function applied: can use the number of different utterances (CITE), the complexity of meanings \cite{kemp2012kinship}, and the entropy of possible utterances \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory}.
 
@@
 (See SI appendix section 5 for mathematical justification and for robustness to other choices).
-We test whether the process of efficiency optimization can explain these typological patterns. %, this shows that efficiency optimization predicts Greenberg's word order correlation universals.
+We test whether the process of efficiency optimization can explain these typological patterns, and make correct predictions about the orderings used in individual languages.
+%, this shows that efficiency optimization predicts Greenberg's word order correlation universals.
 
@@
 
-We compare grammars optimized for efficiency with the actual grammars of the 51 languages, as extracted from the orderings attested in the original corpora (see \textit{Materials and Methods}).
+We compare grammars optimized for efficiency with the actual grammars of the 51 languages, as extracted from the orderings attested in the original corpora and validated against expert judgments (see \textit{Materials and Methods}).
 
 The eight correlations are shown in Table~\ref{table:corr-dryer}.
+
 We compare the prevalence of the word order correlations in optimized languages (right) to their prevalence in the real languages (left).
-%To evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
-%The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
-%in all except four languages, for WALS did not have data.
-Table~\ref{table:corr-dryer} shows that grammars optimized for efficiency accurately predict correlations to hold at prevalences similar to actual languages.
-
-In Figure~\ref{fig:per-lang}, we show results on the levels of individual languages.
+
+For the real languages, we indicate how many of the 51 languages satisfy a correlation.
+For the optimized languages, we indicate the posterior distribution of the proportion of satisfying languages, obtained from a mixed-effects logistic regression with language families as random effects (see \textit{Materials and Methods}).
+Table~\ref{table:corr-dryer} shows that grammars optimized for efficiency accurately predict all eight correlations to hold at prevalences similar to actual languages.
+
+We further evaluated predictions of efficiency optimization on the level of individual laguages (Figure~\ref{fig:per-lang}).
+%In Figure~\ref{fig:per-lang}, we show results on the levels of individual languages.
 For every one of the 51 languages, we show orders predicted by efficiency, for the eight correlating pairs.
@@
 TODO better than this:
+
+Finally, we compare to the predictions made by optimizing only for predictability, and only for parseability.
 
@@
 
-\section{Discussion}
-
+\paragraph{Discussion}
 We have directly implemented parseability and predictability using general-purpose statistical models of language and dependency structures.
@@
 This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
-Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is itself a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+%Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is itself a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
 Indeed, directly modeling efficiency optimization predicts DLM, and appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
@@
 The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication.
-Relatedly, there are many remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
+Relatedly, there are  remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
 Our models cannot capture this trade-off for technical reasons: the parseability objective does not operate over wordforms, only POS tags, and so it does not take advantage of morphological cues to syntactic structure.

### 2019-04-06 second-draft/3-paper-names.tex
revising

@@
 An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural}.
+Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic}.
 %Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
@@
 %We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}. 
-Natural languages vary in the order in which they express information.
+Human languages vary in the order in which they express information.
 Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
@@
 %adpositional phrases (`to a friend') relative to the verb.
-Greenberg also found that the order between verb and object is correlated with other aspects of a language's word order, such the order of verb and adpositional phrase (`wrote -- to friend' in Arabic vs `friend to -- wrote' in Japanese), and that of noun and genitive (`book -- of friend' in Arabic, `friend of -- book' in Japanese).
-Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
-The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.
+Greenberg also found that the order between verb and object is correlated with other aspects of a language's word order (Table~\ref{table:corr-dryer}), such the order of verb and adpositional phrase (`wrote -- to friend' in Arabic vs `friend to -- wrote' in Japanese), and that of noun and genitive (`book -- of friend' in Arabic, `friend of -- book' in Japanese).
+%Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
+%The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.
 
@@
 These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
+
+% alternatively: expressivity vs simplicity
 
@@
 Recent work has used an information-theoretic formalization to computationally test these ideas in various domains of language, showing that they predict
-both basic statistical properties of languages~\cite{ferreri2003least},
+both basic statistical properties of languages~\cite{ferreri2003least} and language development~\cite{smith2013linguistic},
 and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.
@@
 
+\begin{table}
+	\begin{center}
+%		\setlength\tabcolsep{1.5pt}
+\begin{tabular}{|c|ll|ll|}
+	\hline
+	%&
+%	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
+%	&	Verb & Object     & Object & Verb    \\ \hline\hline
+	& kataba & rasaa'il 	&tegami-o & kaita \\
+	&	\emph{wrote} & \emph{letters} & 	\emph{letter} & \emph{wrote} \\ \hline \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    sadiiq       &	tomodachi & ni \\
+	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&kaana    &    sadiiq         &	tomodachi & datta \\
+	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak & udesho \\
+	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& sadiiq  &    John    & 	John no & tomodachi \\
+	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uha       & 	anata-ga yonda & hon \\
+	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tusil        & 	toochaku suru & koto \\
+	&	\emph{that} & \emph{she arrives}  &	\emph{has.arrived} & \emph{that} \\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'ila lmadrasa     & 	gakkoo ni & ittekimashita \\
+	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'uriid    &    'an 'ughaadir         & 	yom & itai \\
+	& \emph{wants}   &  \emph{to leave}  &	\emph{to leave} & \emph{want} \\ \hline \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+\end{tabular}
+	\end{center}
+	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples.
+Based on Dryer \cite{dryer1992greenbergian}, see Materials and Methods.}\label{table:corr-dryer}
+\end{table}
+
 \section{Formalizing Efficiency in Word Order}
@@
 The order of words is determined by the grammar of the language.
-A listener recovers a dependency structure.}
+A listener recovers a dependency structure.
+Communication is successful if the listener recovers the correct depenency structure.}
 	\label{fig:comm}
@@
 When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
-In natural language, this order is largely determined by the \key{grammar} of the language (CITE).
+In human language, this order is largely determined by the \key{grammar} of the language (CITE).
 The speaker will order it as \emph{Mary saw John} (Figure~\ref{fig:comm}, top) in English, and as \emph{Mary John saw} (Figure~\ref{fig:comm}, bottom) in Japanese.
@@
 %While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
-The grammars of natural languages provide consistent ordering rules for all syntactic relations.
+The grammars of human languages provide consistent ordering rules for all syntactic relations.
 This is illustrated in Figure~\ref{fig:grammars}:
@@
 
+%Prior computational tests of efficiency explanations for various aspects of language have
+
 %In information-theoretic formalizations of Zipf's ideas, the efficiency of a
-In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
+
+%In Information Theory, the usefulness of a communication channel 
+The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
 %This idea formalizes Zipf's forces of Unification and Diversification
-Overall efficiency can be formalized as a weighted combination of terms representing the amount of information conveyed and the cost of communication:
+Accordingly, computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge} have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication:
 %The \key{efficiency} of a language is a weighted combination
@@
 with an interpolation weight $\lambda \in [0,1)$. %In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
-This objective function has been proposed in prior computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship}.
 
@@
 
-In our model, both aspects of efficiency will depend on the grammar:
-Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
+%In our model, both aspects of efficiency will depend on the grammar:
+Grammars differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
 This reflects the \key{parseability} of the language.
-Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's Force of Diversification~\cite{frank2012predicting,regier2015word}. % also goodman2013knowledge,
+%Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's Force of Diversification~\cite{frank2012predicting,regier2015word}. % also goodman2013knowledge,
 The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
@@
 This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
-It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
-In line with Zipf's Force of Diversification, this quantity is maximized when trees can be decoded fully unambiguously from utterances.
-
-We formalize the cost of an utterance as its word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
+%It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
+%In line with Zipf's Force of Diversification, it is maximized when trees can be decoded fully unambiguously from utterances.
+
+%This idea formalizes Zipf's forces of Unification and Diversification
+%Accordingly, computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge} have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication:
+
+We formalize the cost of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
+This is the average of the word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
 Surprisal has been found to be a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
@@
 
-With these choices, our objective function is equivalent to that proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems.
+%With these choices, our objective function is equivalent to that proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems.
 %, taking dependency structures as the underlying meanings to be conveyed.
-These works mainly differ in the cost function applied: can use the number of different utterances (CITE), the complexity of meanings \cite{kemp2012kinship}, and the entropy of possible utterances \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory}.
+%These works mainly differ in the cost function applied: can use the number of different utterances (CITE), the complexity of meanings \cite{kemp2012kinship}, and the entropy of possible utterances \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory}.
 
@@
 We examined 
-(1) whether the grammars of natural languages evolve towards optimizing efficiency of communication, and
+(1) whether the grammars of human languages evolve towards optimizing efficiency of communication, and
 (2) whether this process of optimization accounts  for Greenberg's order correlation universals.
@@
 
-%
-%
-
 \begin{table}
@@
 %		\setlength\tabcolsep{1.5pt}
-\begin{tabular}{c|ll|c|cc|ccc}
-	\hline
-	%&
-	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
-	&	verb & object     & & &   \\ 
-	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
-		& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_case.pdf} } %\textbf{86}  }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_case.pdf}       }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_case.pdf}   }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_case.pdf}     } & \\
-	&		\emph{to}            & \emph{a friend} &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
-		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_cop.pdf} }% \textbf{94} }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_cop.pdf} } % \textbf{53}     }
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_cop.pdf}   }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_cop.pdf}   } & \\
-	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
-		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-aux.pdf}}
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-aux.pdf}     }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-aux.pdf}  }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-aux.pdf}  } & \\
-	&	\emph{has}          & \emph{written}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
-		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-nmod.pdf}}
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-nmod.pdf}   }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-nmod.pdf}  } & \\
-	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
-		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-acl.pdf}}
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-acl.pdf}   }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-acl.pdf}  }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-acl.pdf}  } & \\
-	&	\emph{books} & \emph{that you read}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
-		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_mark.pdf}}
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_mark.pdf}     }
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_mark.pdf} }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_mark.pdf}  }  &\\
-	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
-		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-obl.pdf}}
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-obl.pdf}     }
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-obl.pdf} }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-obl.pdf}  } &  \\
-	&	\emph{went} & \emph{to school}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
-		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-xcomp.pdf}}
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-xcomp.pdf}     }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-xcomp.pdf}  } & \\
-	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-\end{tabular}
-	\end{center}
-\caption{OPTION 1: Efficiency optimization accurately predicts the Greenbergian correlations.
-Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
-Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
-%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%Not all correlations are satisfied by every natu
-}\label{table:corr-dryer}
-\end{table}
-
-\begin{table}
-	\begin{center}
-%		\setlength\tabcolsep{1.5pt}
-\begin{tabular}{c|ll|c|cc|ccc}
+\begin{tabular}{|c|ll|c|cc|ccc}
 	\hline
@@
 				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_case.pdf}       }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_case.pdf}   }   
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
@@
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_cop.pdf} } % \textbf{53}     }
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_cop.pdf}   }   
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
@@
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-aux.pdf}     }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-aux.pdf}  }   
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
@@
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-nmod.pdf}   }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
@@
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-acl.pdf}   }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-acl.pdf}  }   
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
@@
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_mark.pdf}     }
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_mark.pdf} }   
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
@@
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-obl.pdf}     }
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-obl.pdf} }   
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
@@
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-xcomp.pdf}     }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
@@
 	\end{center}
-\caption{OPTION 2: Efficiency optimization accurately predicts the Greenbergian correlations.
-Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
+	%Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
 Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
@@
 %Not all correlations are satisfied by every natu
-}\label{table:corr-dryer}
+}\label{table:corr-resu}
 \end{table}
-
-\begin{table*}
-	\begin{center}
-%		\setlength\tabcolsep{1.5pt}
-\begin{tabular}{c|ll|c|c|cc|ccccccccccc}
-	\hline
-	%&
-	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Efficiency} &  \multirow{2}{*}{Predictability} &  \multirow{2}{*}{Parseability} & \\ 
-	&	verb & object     & & &   \\ 
-	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
-				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_case.pdf}     }  & \\
-	&		\emph{to}            & \emph{a friend} &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     }
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_cop.pdf}     } 
-& \\
-	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } 
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_aux.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_aux.pdf}     } & \\
-	&	\emph{has}          & \emph{written}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } 
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_nmod.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_nmod.pdf}     } & \\
-	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_cop.pdf}     }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_acl.pdf}     } & \\
-	&	\emph{books} & \emph{that you read}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_mark.pdf}     } & \\
-	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_obl.pdf}     } & \\
-	&	\emph{went} & \emph{to school}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_xcomp.pdf}     }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_xcomp.pdf}     } & \\
-	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-\end{tabular}
-	\end{center}
-\caption{OPTION 3: Efficiency optimization accurately predicts the Greenbergian correlations.
-Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
-Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
-%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%Not all correlations are satisfied by every natu
-}\label{table:corr-dryer}
-\end{table*}
 
@@
 
-The Greenbergian correlations have been documented in great detail by Dryer \cite{dryer1992greenbergian}, drawing on 625 languages.
-For most of these correlations, it is straightforward to test which of our grammars satisfy them (see \textit{Materials and Methods}).
-Collapsing some correlations that cannot be tested individually, and excluding those for which Universal Dependencies does not provide annotation, we obtained eight correlations  (see \textit{Materials and Methods}).
-%We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages.
-%Most correlations can be straightforwardly implemented in Universal Dependencies (UD), allowing us to check which correlations a word order grammar satisfies.
-%Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages or are not annotated reliably in UD.
-%Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations.
-
-TODO the following paragraphs depend on what visualization we choose
+%The Greenbergian correlations have been documented in great detail by Dryer \cite{dryer1992greenbergian}, drawing on 625 languages.
+%For most of these correlations, it is straightforward to test which of our grammars satisfy them (see \textit{Materials and Methods}).
+%Collapsing some correlations that cannot be tested individually, and excluding those for which Universal Dependencies does not provide annotation, we obtained eight correlations  (see \textit{Materials and Methods}).
+%%We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages.
+%%Most correlations can be straightforwardly implemented in Universal Dependencies (UD), allowing us to check which correlations a word order grammar satisfies.
+%%Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages or are not annotated reliably in UD.
+%%Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations.
 
@@
 
-The eight correlations are shown in Table~\ref{table:corr-dryer}.
-
-We compare the prevalence of the word order correlations in optimized languages (right) to their prevalence in the real languages (left).
-
+We first evaluated which of the eight correlations emerge from efficiency optimization across language families.
+We controlled for variation across different optima by creating eight optimized grammars for each language.
+We controlled for variation between languages and language families by constructing a single Bayesian mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family as random effects.
+In this analysis, efficiency optimization predicts all eight correlations to hold across languages with posterior probability $0.9911$.
+
+In Table~\ref{table:corr-resu}, we compare the prevalence of the eight correlations in real and optimized languages.
 For the real languages, we indicate how many of the 51 languages satisfy a correlation.
-For the optimized languages, we indicate the posterior distribution of the proportion of satisfying languages, obtained from a mixed-effects logistic regression with language families as random effects (see \textit{Materials and Methods}).
-Table~\ref{table:corr-dryer} shows that grammars optimized for efficiency accurately predict all eight correlations to hold at prevalences similar to actual languages.
-
-We further evaluated predictions of efficiency optimization on the level of individual laguages (Figure~\ref{fig:per-lang}).
-%In Figure~\ref{fig:per-lang}, we show results on the levels of individual languages.
+For the optimized languages, we indicate the posterior distribution of the proportion of satisfying languages, obtained from the mixed-effects analysis (see \textit{Materials and Methods}).
+Grammars optimized for efficiency accurately predict all eight correlations to hold at prevalences significantly greater than 50\%, similar to actual human languages.
+Optimizing for only predicability or only parseability does not predict all of the correlations (see SI).
+
+We confirmed these results by evaluating predictions of efficiency optimization on the level of individual laguages (Figure~\ref{fig:per-lang}).
 For every one of the 51 languages, we show orders predicted by efficiency, for the eight correlating pairs.
-For each language, we construct a grammar optimized for efficiency with the direction of the verb-object relation fixed to correspond to the actual language.
-Predictions of efficiency optimization broadly agree with the actual word order patterns.
-%Note that perfect agreement is not to be expected, as grammars will be impacted by factors such as historical development.
-%For instance, Dutch (nld) and Afrikaans (afr) have a prevalence of object-verb order, and
-
-\begin{figure}
-	\begin{center}
-	\includegraphics[width=0.48\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
-	\end{center}
-	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
-\end{figure}
-
-%In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
-We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
-We estimated the posterior distribution of the number of correlations that the optimization process correctly predicts.
-We conducted the same analysis for grammars optimized for predictability and parseability separately~(Figure~\ref{fig:posterior}).
-%Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.
-Efficiency optimization predicts all eight correlations to be universal with posterior probability $0.9911$.
+For each language, we consider the optimized grammar that had the best efficiency value, among those with the direction of the verb-object relation fixed to correspond to the actual language.
+Predictions of efficiency optimization agree with the actual word order patterns in XXX~\% of the cases.
+Note that perfect agreement is not to be expected, as grammars will be impacted by factors other than efficiency optimization, such as their historical development.
+For instance, Dutch (nld) and Afrikaans (afr) have a prevalence of object-verb order in corpus data, and efficiency optimization accordingly predicts them to have corresponding orders (light blue).
+However, Dutch and Afrikaans have orders similar to the other Germanic languages (dark blue), which they are historically closely related to.
+
+We have directly implemented parseability and predictability using general-purpose statistical models of language and dependency structures.
+Prior work has considered Dependency Length Minimization (DLM) as an approximate formalization of efficiency optimization~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+This is the idea that languages order words so as to minimize the overall length of dependencies.
+This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
+%Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is itself a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+Indeed, directly modeling efficiency optimization predicts DLM, and appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
+%Our work thus provides the first direct confirmation of the idea of DLM as an approximation to efficiency optimization -- DLM emerges as a by-product of efficiency maximization.
+
+%The eight correlations are shown in Table~\ref{table:corr-dryer}.
+
+%%We compare the prevalence of the word order correlations in optimized languages (right) to their prevalence in the real languages (left).
+%
+%
+%
+%
+%%In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
+%We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
+%We estimated the posterior distribution of the number of correlations that the optimization process correctly predicts.
+%We conducted the same analysis for grammars optimized for predictability and parseability separately~(Figure~\ref{fig:posterior}).
+%%Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.
+%Efficiency optimization predicts all eight correlations to be universal with posterior probability $0.9911$.
 %The estimated posterior probability that efficiency optimization predicts less than all eight correlations is $0.0242$.
 The probability that it predicts less than seven of the correlations is $<10^{-4}$.
-Predictability and parseability individually also predict six to eight correlations.
-
-TODO better than this:
-
-Finally, we compare to the predictions made by optimizing only for predictability, and only for parseability.
-
-We further compared to the predictions made individually by optimizing for predictability and parseability.
-We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
-We estimated the posterior distribution of the number of correlations that the optimization process correctly predicts.
-We conducted the same analysis for grammars optimized for predictability and parseability separately.
-We used all eight runs of optimization to maximize statistical precision.
-For every one of the correlations, we report the posterior estimated.
-(These numbers control for random variation between language families, and are thus not directly comparable to data from the 51 languages).
+%Predictability and parseability individually also predict six to eight correlations.
+%
+%
+%%
+
+%TODO better than this:
+%
+%Finally, we compare to the predictions made by optimizing only for predictability, and only for parseability.
+%
+%We further compared to the predictions made individually by optimizing for predictability and parseability.
+%We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
+%We estimated the posterior distribution of the number of correlations that the optimization process correctly predicts.
+%We conducted the same analysis for grammars optimized for predictability and parseability separately.
+%We used all eight runs of optimization to maximize statistical precision.
+%For every one of the correlations, we report the posterior estimated.
+%(These numbers control for random variation between language families, and are thus not directly comparable to data from the 51 languages).
+%
+%
+%
+
+%%
+%%
+%
+%\begin{table}
+%	\begin{center}
+%%		\setlength\tabcolsep{1.5pt}
+%\begin{tabular}{c|ll|c|cc|ccc}
+%	\hline
+%	%&
+%	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
+%	&	verb & object     & & &   \\ 
+%	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
+%		& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_case.pdf} } %\textbf{86}  }
+%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_case.pdf}       }
+%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
+%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_case.pdf}   }   
+%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_case.pdf}     } & \\
+%	&		\emph{to}            & \emph{a friend} &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
+%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_cop.pdf} }% \textbf{94} }
+%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
+%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_cop.pdf} } % \textbf{53}     }
+%	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_cop.pdf}   }   
+%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_cop.pdf}   } & \\
+%	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
+%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-aux.pdf}}
+%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
+%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-aux.pdf}     }
+%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-aux.pdf}  }   
+%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-aux.pdf}  } & \\
+%	&	\emph{has}          & \emph{written}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
+%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-nmod.pdf}}
+%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-nmod.pdf}   }
+%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
+%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-nmod.pdf}  } & \\
+%	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
+%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-acl.pdf}}
+%	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
+%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-acl.pdf}   }
+%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-acl.pdf}  }   
+%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-acl.pdf}  } & \\
+%	&	\emph{books} & \emph{that you read}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
+%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_mark.pdf}}
+%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
+%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_mark.pdf}     }
+%	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_mark.pdf} }   
+%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_mark.pdf}  }  &\\
+%	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+%	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
+%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-obl.pdf}}
+%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
+%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-obl.pdf}     }
+%	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-obl.pdf} }   
+%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-obl.pdf}  } &  \\
+%	&	\emph{went} & \emph{to school}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
+%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-xcomp.pdf}}
+%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-xcomp.pdf}     }
+%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
+%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-xcomp.pdf}  } & \\
+%	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+%\end{tabular}
+%	\end{center}
+%\caption{OPTION 1: Efficiency optimization accurately predicts the Greenbergian correlations.
+Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+%Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
+%%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%%Not all correlations are satisfied by every natu
+%}\label{table:corr-dryer}
+%\end{table}
+%
 
@@
 
-\paragraph{Discussion}
-We have directly implemented parseability and predictability using general-purpose statistical models of language and dependency structures.
-Prior functional studies have considered Dependency Length Minimization (DLM) as an approximate formalization of efficiency optimization~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-This is the idea that languages order words so as to minimize the overall length of dependencies.
-This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
-%Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is itself a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-Indeed, directly modeling efficiency optimization predicts DLM, and appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
-%Our work thus provides the first direct confirmation of the idea of DLM as an approximation to efficiency optimization -- DLM emerges as a by-product of efficiency maximization.
-
 %We show that Efficiency optimization predicts DLM, providing a first-principles functional explanation.
@@
 \end{document}
+
+\begin{table}
+	\begin{center}
+%		\setlength\tabcolsep{1.5pt}
+\begin{tabular}{|c|ll|c|cc|ccc}
+	\hline
+	%&
+	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
+	&	verb & object     & & &   \\ 
+	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
+				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
+	&		\emph{to}            & \emph{a friend} &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
+	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
+	&	\emph{has}          & \emph{written}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
+	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
+	&	\emph{books} & \emph{that you read}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
+	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
+	&	\emph{went} & \emph{to school}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
+	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+\end{tabular}
+	\end{center}
+\caption{OPTION 2: Efficiency optimization accurately predicts the Greenbergian correlations.
+Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
+%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%Not all correlations are satisfied by every natu
+}\label{table:corr-dryer}
+\end{table}
+
+\begin{table}
+	\begin{center}
+%		\setlength\tabcolsep{1.5pt}
+\begin{tabular}{c|ll|c|cc|ccc}
+	\hline
+	%&
+	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
+	&	verb & object     & & &   \\ 
+	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
+				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
+		&    & \\
+	&		\emph{to}            & \emph{a friend} &&&\\ 
+	&&& {  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_cop.pdf} } % \textbf{53}     }
+	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_cop.pdf}   }   
+		& & \\
+	&	\emph{is}        & \emph{a friend}  &&&\\ 
+	&&&   {  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-aux.pdf}     }
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-aux.pdf}  }   
+		&  & \\
+	&	\emph{has}          & \emph{written}  &&&\\ 
+	&&&  {  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-nmod.pdf}   }
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
+	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
+	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-acl.pdf}   }
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-acl.pdf}  }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
+	&	\emph{books} & \emph{that you read}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_mark.pdf}     }
+	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_mark.pdf} }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
+	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-obl.pdf}     }
+	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-obl.pdf} }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
+	&	\emph{went} & \emph{to school}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
+	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-xcomp.pdf}     }
+	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
+	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+\end{tabular}
+	\end{center}
+\caption{OPTION 2: Efficiency optimization accurately predicts the Greenbergian correlations.
+Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
+%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%Not all correlations are satisfied by every natu
+}\label{table:corr-dryer}
+\end{table}
 

### 2019-04-06 second-draft/pnas-word-order-si.tex
revising

@@
 \label{tab:all-predictions}
+\end{table*}
+
+\section{Posterior}
+
+\begin{figure}
+	\begin{center}
+	\includegraphics[width=0.48\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
+	\end{center}
+	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
+\end{figure}
+
+\begin{table*}
+	\begin{center}
+%		\setlength\tabcolsep{1.5pt}
+\begin{tabular}{c|ll|c|c|cc|ccccccccccc}
+	\hline
+	%&
+	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Efficiency} &  \multirow{2}{*}{Predictability} &  \multirow{2}{*}{Parseability} & \\ 
+	&	verb & object     & & &   \\ 
+	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
+				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_case.pdf}     }  & \\
+	&		\emph{to}            & \emph{a friend} &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     }
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_cop.pdf}     } 
+& \\
+	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_aux.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_aux.pdf}     } & \\
+	&	\emph{has}          & \emph{written}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_nmod.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_nmod.pdf}     } & \\
+	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_cop.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_acl.pdf}     } & \\
+	&	\emph{books} & \emph{that you read}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_mark.pdf}     } & \\
+	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_obl.pdf}     } & \\
+	&	\emph{went} & \emph{to school}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_xcomp.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_xcomp.pdf}     } & \\
+	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+\end{tabular}
+	\end{center}
+\caption{OPTION 3: Efficiency optimization accurately predicts the Greenbergian correlations.
+Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
+%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%Not all correlations are satisfied by every natu
+}\label{table:corr-dryer}
 \end{table*}

### 2019-04-06 results/plane/plane-2D-best-visualization-large-2.R
revising

@@ -132,7 +132,7 @@ data = merge(data, depl %>% select(Language, Model,AverageLengthPerWord) %>% mut
 data = data %>% mutate(Two = 0.9*Surp+Pars)
 
 data2 = rbind(data)
-data2 = data2 %>% group_by(Language) %>% mutate(MeanSurp = mean(Surp, na.rm=TRUE), SDSurp = sd(Surp, na.rm=TRUE)) %>% mutate(MeanPars = mean(Pars, na.rm=TRUE), SDPars = sd(Pars, na.rm=TRUE))
+data2 = data2 %>% group_by(Language, Type) %>% summarise(Surp=mean(Surp, na.rm=TRUE), Pars=mean(Pars, na.rm=TRUE)) %>% group_by(Language) %>% mutate(MeanSurp = mean(Surp, na.rm=TRUE), SDSurp = sd(Surp, na.rm=TRUE)) %>% mutate(MeanPars = mean(Pars, na.rm=TRUE), SDPars = sd(Pars, na.rm=TRUE))
 plot = ggplot(data2, aes(x=(Pars-MeanPars)/SDPars, y=(Surp-MeanSurp)/SDSurp, color=Type, group=Type)) +geom_point()
 #plot = ggplot(data, aes(x=(Pars), y=(Surp), color=Type, group=Type)) +geom_point()
 
@@ -185,10 +185,13 @@ data = rbind(data, dataDepL)
 data = rbind(data, dataGround)
 #data = rbind(data, dataReal)
 
-dataMean = data %>% group_by(Language) %>% summarise(MeanSurp = mean(Surp, na.rm=TRUE), SDSurp = sd(Surp, na.rm=TRUE)+0.0001)
+
+#data2 = data2 %>% %>% group_by(Language) %>% mutate(MeanSurp = mean(Surp, na.rm=TRUE), SDSurp = sd(Surp, na.rm=TRUE)) %>% mutate(MeanPars = mean(Pars, na.rm=TRUE), SDPars = sd(Pars, na.rm=TRUE))
+
+dataMean = data %>% group_by(Language, Type) %>% summarise(Surp=mean(Surp, na.rm=TRUE)) %>% group_by(Language) %>% summarise(MeanSurp = mean(Surp, na.rm=TRUE), SDSurp = sd(Surp, na.rm=TRUE)+0.0001)
 data = merge(data, dataMean, by=c("Language"))
 dataDepL = merge(dataDepL, dataMean, by=c("Language"))
[…]

### 2019-04-07 second-draft/3-paper-names.tex
revising

@@
 % ,lin2017critical
-\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
+\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics. % \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
 % behaghel1909beziehungen,
 % \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
-An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
+%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
 Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic}.
@@
 Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study~\cite{lehmann1973structural, vennemann1974theoretical,jackendoff1977x,frazier1985syntactic,chomsky1988language, dryer1992greenbergian, hawkins1994performance}.
-A prominent line of research has argued that universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
+
+A prominent line of research has argued that language universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
 Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.
@@
 Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required in order to provide higher informativity.
-The idea that many properties of language arise from the tension between these two pressures has a long and fruitful history in theoretical studies of language~\cite{gabelentz1901sprachwissenschaft,horn1984toward, schwartz1997dispersion}.
+The idea that many properties of language arise from the tension between these two pressures has a long and fruitful history in theoretical studies of language~\cite{gabelentz1901sprachwissenschaft,horn1984toward, lindblom1990explaining, schwartz1997dispersion, hawkins2004efficiency}.
 
@@
 We then show that efficiency optimization accounts for the Greenbergian word order correlations, and makes fine-grained per-language predictions about individual syntactic patterns.
+Our results show how general properties of efficient communication gives rise to the universal word order properties of human language.
 %We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.
@@
 
+%
+%\begin{table}
+%	\begin{center}
+%%		\setlength\tabcolsep{1.5pt}
+%\begin{tabular}{|c|ll|c|cc|ccc}
+%	\hline
+%	%&
+%	&	\multicolumn{2}{|c|}{Correlates with...}   &         \\ 
+%	&	verb & object     & & &   \\ 
+%	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	Adposition    &    Noun Phrase
+%\\
+%	&		\emph{to}            & \emph{a friend} &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
+%\\
+%	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
+% \\
+%	&	\emph{has}          & \emph{written}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
+%\\
+%	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
+%\\
+%	&	\emph{books} & \emph{that you read}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
+%\\
+%	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+%	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
+%\\
+%	&	\emph{went} & \emph{to school}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
+%\\
+%	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+%\end{tabular}
+%	\end{center}
+%\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
+%	%Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+%%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+%Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
+%%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%%Not all correlations are satisfied by every natu
+%}\label{table:corr-resu}
+%\end{table}
+%
+%
+%
+%
+
 \begin{table}
@@
 	%&
-%	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
-%	&	Verb & Object     & Object & Verb    \\ \hline\hline
+	&	\multicolumn{2}{c|}{\textbf{Arabic} (English, ...)}   &        \multicolumn{2}{c|}{\textbf{Japanese} (Hindi, ...)} \\ \hline\hline
+	& \multicolumn{2}{c}{Correlates with...} & \multicolumn{2}{|c|}{Correlates with...}  \\
+	&	Verb & Object     & Object & Verb    \\ \hline\hline
 	& kataba & rasaa'il 	&tegami-o & kaita \\
-	&	\emph{wrote} & \emph{letters} & 	\emph{letter} & \emph{wrote} \\ \hline \hline
+	&	\emph{wrote} & \emph{letters} & 	\emph{letter} & \emph{wrote} \\ \hline 
 	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    sadiiq       &	tomodachi & ni \\
@@
 	\end{center}
-	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples.
-Based on Dryer \cite{dryer1992greenbergian}, see Materials and Methods.}\label{table:corr-dryer}
+	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples: Across the world, the orders of different constituents are strikingly correlated with that of verb and object.
+See Materials and Methods.}\label{table:corr-dryer}
 \end{table}
@@
 %This idea formalizes Zipf's forces of Unification and Diversification
-Accordingly, computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge} have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication:
-%The \key{efficiency} of a language is a weighted combination
-\begin{equation}\label{eq:efficiency}
-	R_{\textit{Eff}} := R_{\textit{Informativity}} - \lambda R_\textit{Cost}
-\end{equation}
-with an interpolation weight $\lambda \in [0,1)$. %In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
-
-%Such notions of efficiency have been proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems.
-%, taking dependency structures as the underlying meanings to be conveyed.
-%Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
-%Word meaning \cite{regier2015word}
-
-%In our model, both aspects of efficiency will depend on the grammar:
-Grammars differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
-This reflects the \key{parseability} of the language.
-%Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's Force of Diversification~\cite{frank2012predicting,regier2015word}. % also goodman2013knowledge,
-The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
+Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
+
+As we take the underlying messages to be dependency structures, the first term is the degree to which listeners can reconstruct dependency structures from an utterance, i.e., the \key{parseability} if the language.
+This is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
 \begin{equation}
@@
 where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
-This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
-%It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
-%In line with Zipf's Force of Diversification, it is maximized when trees can be decoded fully unambiguously from utterances.
-
-%This idea formalizes Zipf's forces of Unification and Diversification
-%Accordingly, computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge} have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication:
+This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
 
 We formalize the cost of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
-This is the average of the word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
+In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
 Surprisal has been found to be a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
@@
 
+Then, the efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
+%The \key{efficiency} of a language is a weighted combination
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+\end{equation}
+with an interpolation weight $\lambda \in [0,1)$ (See SI appendix section 5 for mathematical justification).
+%In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
+
+%Such notions of efficiency have been proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems.
+%, taking dependency structures as the underlying meanings to be conveyed.
+%Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
+%Word meaning \cite{regier2015word}
+
+%In our model, both aspects of efficiency will depend on the grammar:
+%Grammars differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
+%This reflects the \key{parseability} of the language.
+%%Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's Force of Diversification~\cite{frank2012predicting,regier2015word}.
+% also goodman2013knowledge,
+%The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
+%\begin{equation}
+%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+%\end{equation}
+%where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
+%This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
+%It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
+%In line with Zipf's Force of Diversification, it is maximized when trees can be decoded fully unambiguously from utterances.
+
+%This idea formalizes Zipf's forces of Unification and Diversification
+%Accordingly, computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge} have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication:
+
 %With these choices, our objective function is equivalent to that proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems.
@@
 We compare the efficiency of the actual grammars of the 51 languages from the Universal Dependencies datasets to randomly constructed baseline grammars.
-For every one of the 51 languages, we construct ten counterfactual versions by randomly creating ten grammars.
+For every one of the 51 languages, we construct 20 counterfactual versions by randomly creating 20 grammars.
 These baseline grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}, see \textit{Materials and Methods} for details).
@@
 Furthermore, for every one of the 51 languages, we computationally construct grammars that are \emph{optimized} for parsability, predictability, and efficiency (see \textit{Materials and Methods}).
+For each grammar, we report predictability and parsability as estimated on the data resulting from ordering the dependency structures from the corpus according to the grammar.
 
@@
 %By a one-sided \textit{t}-test, 
-84~\% of languages improve over the baselines in parsability ($p < 0.05$ each), 78~\% improve in predictability ($p < 0.05$ each).
+92~\% of languages improve over the baselines in parsability ($p < 0.05$ each), 84~\% improve in predictability ($p < 0.05$ each).
 %At $\lambda = 0.9$, 86~\% of languages improve in overall efficiency ($p<0.05$ each).
@@
     \centering
-    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend-viz.pdf}
+    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend-viz-10.pdf}
     \caption{Predictability and parseability of 51 languages (red), indicated by ISO codes, compared to baseline word order grammars (blue distribution).
@@
 	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
-Dark blue: Verb patterner \emph{precedes} object patterner.
-Light blue: Verb patterner \emph{follows} object patterner.}\label{fig:per-lang}
+Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
+Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).}\label{fig:per-lang}
 	% efficiency optimization predicts Greenberg's correlations.
@@
 \end{figure}
+
+%
+%
+%
+%
+%\begin{table}
+%	\begin{center}
+%%		\setlength\tabcolsep{1.5pt}
+%\begin{tabular}{|c|ll|c|cc|ccc}
+%	\hline
+%	%&
+%	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
+%	&	verb & object     & & &   \\ 
+%	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
+%				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
+%	&		\emph{to}            & \emph{a friend} &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
+%	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
+%	&	\emph{has}          & \emph{written}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
+%	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
+%	&	\emph{books} & \emph{that you read}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
+%	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+%	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
+%	&	\emph{went} & \emph{to school}  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
+%	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+%\end{tabular}
+%	\end{center}
+%\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
+%	%Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+%%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+%Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
+%%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%%Not all correlations are satisfied by every natu
+%}\label{table:corr-resu}
+%\end{table}
+%
+%
+%
+%
+
+%
+%\begin{table}
+%	\begin{center}
+%%		\setlength\tabcolsep{1.5pt}
+%\begin{tabular}{|c|ll|c|cc|ccc}
+%	\hline
+%	%&
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	
+%				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
+%	&& &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
+%  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
+%  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
+%  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
+%  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
+%  &&&\\ \hline
+%	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
+%  &&&\\ \hline
+%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	  
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
+%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
+%  &&&\\ \hline
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+%\end{tabular}
+%	\end{center}
+%\caption{OPTION 1: Efficiency optimization accurately predicts the Greenbergian correlations.
+%	%Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+%For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+%%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+%Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
+%%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%%Not all correlations are satisfied by every natu
+%}\label{table:corr-resu}
+%\end{table}
 
@@
 For each language, we consider the optimized grammar that had the best efficiency value, among those with the direction of the verb-object relation fixed to correspond to the actual language.
-Predictions of efficiency optimization agree with the actual word order patterns in XXX~\% of the cases.
+Predictions of efficiency optimization agree with the actual word order patterns in 82~\% of the cases.
 Note that perfect agreement is not to be expected, as grammars will be impacted by factors other than efficiency optimization, such as their historical development.
@@
 %
+
+\section{Conclusion}
+
+We found that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
+We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals.
+Beyond our present results, we provide a complete formalization and computational framework in which theories of the functional optimization of languages can be tested.
+Other objective functions could be hypothesized and tested within our framework; furthermore, future advances in machine learning will enable the optimization of richer and richer models of grammar.
+
+A major question for functional explanations for linguistic universals is: \emph{how} do languages come to be optimized?
+Do speakers actively seek out new communicative conventions that allow better efficiency?
+Or do languages change in response to biases that come into play during language acquisition \cite{fedzechkina2012language,culbertson2012learning}?
+Our work is neutral toward such questions.
+To the extent that language universals arise from biases in learning or in the representational capacity of the human brain, our results suggest that those biases tilt toward communicative efficiency.
+Our work does provide evidence against the idea that word order universals are best explained in terms of learning biases that are irreducibly arbitrary and genetic in nature \cite{chomsky2010some}.
+
+While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into dependency trees.
+The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication.
+Relatedly, there are  remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
+Our models cannot capture this trade-off for technical reasons: the parseability objective does not operate over wordforms, only POS tags, and so it does not take advantage of morphological cues to syntactic structure.
+Future work can investigate whether these and other remaining universals can be explained using more sophisticated models of how meaning representations are transduced into strings of words, based on larger databases with richer annotations.
+
+\matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
+We use all languages for which at least one treebank with a training partition was available, a total of 50 languages.
+For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.
+Punctuation was removed.
+%We removed punctuation. %, which we operationalized by removing all dependents of \emph{punct} dependencies, and recursively removing their dependents.
+
+Universal dependencies represents as dependents some words that are typically classified as heads in syntactic theory.
+This particularly applies to the \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark} dependencies.
+Following prior work studying dependency length minimization \cite{futrell2015largescale}, we modified each treebank by inverting these dependencies, promoting the dependent to the head position.
+We report results on this modified version of UD.
+
+The efficiency optimization results from Table~\ref{tab:results-dryer} were preregistered: \url{http://aspredicted.org/blind.php?x=8gp2bt}.
+
+Code and results are available at \url{https://github.com/m-hahn/grammar-optim}.
+
+See the SI for details on the neural language models and parsers used, and on the optimization procedures.}
+
+\showmatmethods{} % Display the Materials and Methods section
+
+%\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Noah Goodman, Chris Manning, and audiences at CAMP 2018 and CUNY 2019 for helpful discussion.}
+
+\showacknow{} % Display the acknowledgments section
+
+%\bibliographystyle{acl_natbib}
+%\bibliography{extracted}
+\bibliography{everything}
+
+\end{document}
 
@@
 
-\section{Conclusion}
-
-We found that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
-We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals.
-Beyond our present results, we provide a complete formalization and computational framework in which theories of the functional optimization of languages can be tested.
-Other objective functions could be hypothesized and tested within our framework; furthermore, future advances in machine learning will enable the optimization of richer and richer models of grammar.
-
-A major question for functional explanations for linguistic universals is: \emph{how} do languages come to be optimized?
-Do speakers actively seek out new communicative conventions that allow better efficiency?
-Or do languages change in response to biases that come into play during language acquisition \cite{fedzechkina2012language,culbertson2012learning}?
-Our work is neutral toward such questions.
-To the extent that language universals arise from biases in learning or in the representational capacity of the human brain, our results suggest that those biases tilt toward communicative efficiency.
-Our work does provide against the idea that word order universals are best explained in terms of learning biases that are irreducibly arbitrary and genetic in nature \cite{chomsky2010some}.
-
-While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into dependency trees.
-The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication.
-Relatedly, there are  remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
-Our models cannot capture this trade-off for technical reasons: the parseability objective does not operate over wordforms, only POS tags, and so it does not take advantage of morphological cues to syntactic structure.
-Future work can investigate whether these and other remaining universals can be explained using more sophisticated models of how meaning representations are transduced into strings of words, based on larger databases with richer annotations.
-
-\matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
-We use all languages for which at least one treebank with a training partition was available, a total of 50 languages.
-For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.
-Punctuation was removed.
-%We removed punctuation. %, which we operationalized by removing all dependents of \emph{punct} dependencies, and recursively removing their dependents.
-
-Universal dependencies represents as dependents some words that are typically classified as heads in syntactic theory.
-This particularly applies to the \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark} dependencies.
-Following prior work studying dependency length minimization \cite{futrell2015largescale}, we modified each treebank by inverting these dependencies, promoting the dependent to the head position.
-We report results on this modified version of UD.
-
-The efficiency optimization results from Table~\ref{tab:results-dryer} were preregistered: \url{http://aspredicted.org/blind.php?x=8gp2bt}.
-
-Code and results are available at \url{https://github.com/m-hahn/grammar-optim}.
-
-See the SI for details on the neural language models and parsers used, and on the optimization procedures.}
-
-\showmatmethods{} % Display the Materials and Methods section
-
-\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Noah Goodman, Chris Manning, and audiences at CAMP 2018 and CUNY 2019 for helpful discussion.}
-
-\showacknow{} % Display the acknowledgments section
-
-%\bibliographystyle{acl_natbib}
-%\bibliography{extracted}
-\bibliography{everything}
-
-\end{document}
-
 \begin{table}

### 2019-04-07 results/correlations/visualize-posterior-2-large-2in1.R

@@ -1,0 +1,174 @@
+
+dependencies = c("acl", "advcl", "advmod", "amod", "appos", "aux", "ccomp", "compound", "conj", "csubj", "dep", "det", "discourse", "dislocated", "expl", "fixed", "flat", "goeswith", "iobj", "lifted_case", "lifted_cc", "lifted_cop", "lifted_mark", "list", "nmod", "nsubj", "nummod", "obl", "orphan", "parataxis", "reparandum", "vocative", "xcomp")
+
+dependencies =  c("acl", "advmod", "aux", "lifted_case", "lifted_cop", "lifted_mark", "nmod", "nsubj", "obl", "xcomp")
+
+data = read.csv("../../grammars/manual_output_funchead_ground_coarse_final/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+
+
+
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+
+data = data %>% mutate(DH_Weight = DH_Mean_NoPunct)
+
+
+library(forcats)
+data = data %>% mutate(Language = fct_recode(Language, "Old_Church_Slavonic" = "Old", "Ancient_Greek" = "Ancient"))
+
+languages = read.csv("../languages/languages-iso_codes.tsv", sep=",")
+data  = merge(data, languages, by=c("Language"), all.x=TRUE)
+
+
+
+
+dependency = "nmod"
+
+getCorrPair = function(dependency) {
+   corr_pair = data %>% filter((Dependency == dependency) | (Dependency == "obj"))
+   corr_pair = unique(corr_pair %>% select(Family, Language, FileName, Dependency, DH_Weight, ModelName)) %>% spread(Dependency, DH_Weight)
+   corr_pair$correlator = corr_pair[[dependency]]
+   corr_pair = corr_pair %>% mutate(correlator_s = pmax(0,sign(correlator)), obj_s=pmax(0,sign(obj)))
+   return(corr_pair)
+}
+
+corr_pair = getCorrPair("nmod")
+
+dependencies = c("acl", "advcl", "advmod", "amod", "appos", "aux", "ccomp", "compound", "conj", "csubj", "dep", "det", "discourse", "dislocated", "expl", "fixed", "flat", "goeswith", "iobj", "lifted_case", "lifted_cc", "lifted_cop", "lifted_mark", "list", "nmod", "nsubj", "nummod", "obl", "orphan", "parataxis", "reparandum", "vocative", "xcomp")
+
[…]

### 2019-04-13 second-draft/3-paper-names-backup.tex
cleaning up

--- 2019-04-07 second-draft/3-paper-names.tex
@@
 % information-theoretic
+Specifically, we show that efficiency optimization explains  linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.
 
@@
 and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.
-Common to these accounts is the idea of maximizing the amount of information that linguistic forms provide about meanings, while constraining the diversity of forms.
 
@@
 Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
+Common to these accounts is the idea of maximizing the amount of information that linguistic forms provide about meanings, while constraining the diversity of forms.
 
@@
 In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
-Surprisal has been found to be a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
 In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
@@
 \end{equation}
-where the sum runs over all possible sentences $l$ that belong to the language.
+where the sum runs over all possible sentences $u$ that belong to the language.
 In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
@@
 \matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
-We use all languages for which at least one treebank with a training partition was available, a total of 50 languages.
+We use all languages for which at least one treebank with a training partition was available, a total of 51 languages.
 For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.

### 2019-04-13 results/correlations/visualize-posterior-2-large-2in1.R

@@ -94,13 +94,13 @@ library(ggplot2)
 
 
 cat("\nReading posterior samples\n")
-parse = read.csv("/home/user/CS_SCR/posteriors/posterior-10-parseability.csv")
+#parse = read.csv("/home/user/CS_SCR/posteriors/posterior-10-parseability.csv")
 two = read.csv("/home/user/CS_SCR/posteriors/posterior-10-efficiency-large.csv")
-langmod = read.csv("/home/user/CS_SCR/posteriors/posterior-10-langmod.csv")
+#langmod = read.csv("/home/user/CS_SCR/posteriors/posterior-10-langmod.csv")
 
 
 
-data = rbind(langmod %>% mutate(Type="Predictability"), parse %>% mutate(Type="Parseability"), two %>% mutate(Type="Efficiency"))
+data = rbind(two %>% mutate(Type="Efficiency"))
 
 data2 = data %>% gather(Dependency, Intercept, b_acl_Intercept,b_advmod_Intercept                                   ,b_aux_Intercept                                       ,  b_liftedcase_Intercept,b_liftedcop_Intercept  ,                                 b_liftedmark_Intercept  ,                                b_nmod_Intercept         ,                               b_nsubj_Intercept         ,                              b_obl_Intercept            ,                             b_xcomp_Intercept           )
 
@@ -112,7 +112,6 @@ data2 = data2 %>% mutate(Dependency = case_when(Dependency == "liftedcase" ~ "li
 data2 = data2 %>% mutate(Prevalence = case_when(Dependency == "aux" ~ 1-Prevalence, TRUE ~ Prevalence))
 
   plot = ggplot(data=data2) + geom_density(aes(x=Prevalence, y=..scaled.., fill=Type, group=Type), alpha=.5) + xlim(0,1) +  facet_wrap( ~ Dependency, ncol=1) 
-  ggsave(plot, file=paste("figures/posterior_joint_large" , ".pdf", sep=""))
 
 
 
@@ -123,7 +122,6 @@ data2 = data2 %>% mutate(Prevalence = case_when(Dependency == "aux" ~ 1-Prevalen
 
 
[…]

### 2019-04-18 second-draft/3-paper-names.tex
figure

@@
 %An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
+% TODO mention universals here
 Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic}.
@@
 % information-theoretic
+Specifically, we show that efficiency optimization explains  linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}. 
+% TODO change wording so it's clear that this is a very major thing
+
+% Things TODO
+% - make clear that efficiency is something both functional and generative linguists have been appealing to
+% - at some point address head direction idea
+%TODO at some point address head direction idea
+%    - Vennemann 1972, 1974ab, 1976
+%    -Hawkins 1980 limitations of head direction, also Hawkins 1982
 
@@
 An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-
-%TODO at some point address head direction idea
-%- Vennemann 1972, 1974ab, 1976
-
-%-Hawkins 1980 limitations of head direction, also Hawkins 1982
 
@@
 Arabic expresses `to' by a \emph{preposition} (\emph{preceding} the noun `friend'); Japanese uses a \emph{postposition} (\emph{following} it). % and a postposition in Japanese, following `friend'.
-%In Arabic, the verb stands at the beginning, followed by the objecat and the adpositional phrase.
-%In Arabic, `to' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
-%In Japanese, it is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
-%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
-%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
-%It turns out that these are deep and stable regularities:
 
@@
 Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
-%However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
-%As Figure~\ref{fig:corr-table} shows, this correlation has an extremely strong empirical basis:
-%In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization.
-%, over 90 \% of languages satisfy this generalization .
 
@@
     \includegraphics[width=0.4\textwidth]{figures/corr-full.png}
-%
-%\centering
-%\begin{dependency}[theme = simple]
-%   \begin{deptext}[column sep=1em]
-%%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
-%	   katabt \& risāla \& li \& sadīq  \\
-%%	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
-%	   wrote \& letter   \& to \& friend  \\
-%   \end{deptext}
-%\end{dependency}
-%\begin{dependency}[theme = simple]
-%   \begin{deptext}[column sep=1em]
-%      tomodachi \& ni \& tegami-o \& kaita \\
-% %\textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
-%	   friend \& to \& letter  \& wrote \\
-%   \end{deptext}
-%\end{dependency}
-%
-%    \centering
-%    \includegraphics[width=0.4\textwidth]{figures/corr-table.png}
-%
-%%\begin{dependency}[theme=simple]
-%%    \begin{deptext}
-%%    I \& wrote \& a \& letter \& to \& a \& friend \\
-%%    \end{deptext}
-%%    \depedge{2}{1}{nsubj}
-%%    \depedge{2}{4}{obj}
-%%    \depedge{4}{3}{ }
-%%    \depedge{4}{3}{ }
-%%    \depedge{2}{5}{}
-%%    \depedge{5}{7}{ }
-%%    \depedge{7}{6}{ }
-%%\end{dependency}
 \caption{ The concept of a word order correlation: Languages can order the object after (Arabic) or before (Japanese) the verb, and have prepositions (Arabic) or postpositions (Japanese).
@@
 Greenberg also found that the order between verb and object is correlated with other aspects of a language's word order (Table~\ref{table:corr-dryer}), such the order of verb and adpositional phrase (`wrote -- to friend' in Arabic vs `friend to -- wrote' in Japanese), and that of noun and genitive (`book -- of friend' in Arabic, `friend of -- book' in Japanese).
-%Greenberg originally worked with 30 languages; the correlation universals have since been confirmed on the basis of much larger samples of languages.
-%The authoritative study by Dryer \cite{dryer1992greenbergian} draws on 625 languages and documents fifteen such correlations.
 
@@
 
-%TODO can we find citations from all camps?
-
 Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study~\cite{lehmann1973structural, vennemann1974theoretical,jackendoff1977x,frazier1985syntactic,chomsky1988language, dryer1992greenbergian, hawkins1994performance}.
@@
 Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.
-
-%Researchers working in the functional paradigm have proposed a range of criteria that languages should meet in order to enable efficient communication and processing.
-These criteria are based on theories of online processing difficulty (see \cite{jaeger2011language} for a review) and on information theoretic notions of efficiency and robustness in communication \cite{ferrericancho2001two,piantadosi2011word,gibson2013noisy}, and they have been formalized to varying degrees.
-
-% alternatively: expressivity vs simplicity
 
@@
 Conversely, the Force of Diversification favors languages that provide different utterances for different meanings, so that the listener can unambiguously identify the meaning from the utterance.
-%Minimizing speaker and listener effort represent two opposing forces:
-%A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
-%A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
-%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
 These two forces act in opposing directions:
@@
 and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.
-Common to these accounts is the idea of maximizing the amount of information that linguistic forms provide about meanings, while constraining the diversity of forms.
 
 While it has been suggested that these ideas should also apply to grammar~\cite{hawkins2004efficiency}, testing these accounts is more difficult in the domain of grammar, as this requires both large amounts of data representative of language use across languages and cultures, and computational methods for estimating the efficiency of entire languages.
-%TODO somewhere start calling this Efficiency
-%TODO explain why surprisal. some production/speaker work. (plan reuse)
-%In restricted settings such as modeling reference games and the distribution of meanings in certain classes of words, these ideas have received computational and information-theoretic formalizations, showing that they successfully account for phenomena such as pragmatic inference (CITE frank and goodman) and color naming (CITE).
-%Even earlier than Zipf, \cite{gabelentz1901sprachwissenschaft} ...
-%This theory blabla Horn RSA
-%Prior computational implementations of this idea have mostly been in the restricted small-scale settings, such as simple reference games.
 In this work, we address these challenges by combining large-scale text data from 51 languages with state-of-the-art machine learning techiques to estimate the communicative efficiency of grammar.
@@
 Our results show how general properties of efficient communication gives rise to the universal word order properties of human language.
-%We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.
-
-%In this work, we first provide a formalization of efficiency in the domain of word order. %, drawing on information-theoretic notions.
-
-%In this work, we show computationally, using corpus data from 51 languages, that efficiency optimization accounts for the Greenbergian word order correlations.
-%That is, we show that the word orders in natural languages have evolved to optimize efficiency, and that this optimization accounts for the prevalence of the word order correlations.
-%word orders in natural language
-
-%OR say DLM and efficiency, as in the talks
-
-%\section{DLM}
-
-%
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.45\textwidth]{figures/letter-options.png}
-%	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure.
-In this example, Grammar 1 achieves the lowest overall dependency length.
-Indeed, this is the }
-%	\label{fig:pareto-plane}
-%\end{figure}
-%
-%
-
-%
-%
-%\begin{table}
-%	\begin{center}
-%%		\setlength\tabcolsep{1.5pt}
-%\begin{tabular}{|ll|c|cccc|llllllll}
-%	\hline
-%	%&
-%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Eff}  \\ 
-%	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%%adp. &
-%	adposition    &    NP       
-%	& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}       }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}   }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}     }  \\
-%	\emph{to}            & \emph{a friend} &&&&&\\ \hline
-%copula    &    NP         
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}       }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}   }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}   }  \\
-%	\emph{is}        & \emph{a friend}  &&&&&\\ \hline
-%auxiliary    &    VP       
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}       }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}     }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  }  \\
-%	\emph{has}          & \emph{written}  &&&&&\\ \hline
-%noun    &    genitive      
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  }  \\
-%	\emph{friend} &  \emph{of John}  &&&&&\\ \hline
-%noun    &    relative clause      
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}     }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  }  \\
-%	\emph{books} & \emph{that you read}  &&&&&\\ \hline
-%complementizer    &    S        
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf}     }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}     }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf} }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  }  \\
-%	\emph{that} & \emph{she has arrived}  &&&&&\\ \hline
-%verb    &    PP         
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf}     }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}     }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  }  \\
-%	\emph{went} & \emph{to school}  &&&&&\\ \hline
-%want    &    VP        
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}     }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}  }  \\
-%	\emph{wants}   &  \emph{to leave}  &&&&&\\ \hline
-% \hline
-%%comp. &
-%%\multirow{3}{*}{NP}&
-%%	AP &
-%    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-%\end{tabular}
-%	\end{center}
-%	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%For each correlation, we provide an English example.
-%%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%%Not all correlations are satisfied by every natu
-%}\label{table:corr-dryer}
-%\end{table}
-%
-%
-
-%
-%
-%\begin{table*}
-%	\begin{center}
-%\begin{tabular}{|ll|l|l|l|ll|l|llllllll}
-%	\hline
-%	%&
-%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
-%	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%%adp. &
-%	adposition    &    NP       
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}   $^{***}$    }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}  $^{***}$ }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf} $^{***}$    } \\
-%	\emph{to}            & \emph{a friend} \\ \hline
-%copula    &    NP         
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}   $^{***}$    }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}  $^{***}$ }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf} $^{**}$  }  \\
-%	\emph{is}        & \emph{a friend} \\ \hline
-%auxiliary    &    VP       
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}   $^{***}$    }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}  $^{***}$   }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}$^{**}$  }  \\
-%	\emph{has}          & \emph{written} \\ \hline
-%noun    &    genitive      
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} $^{***}$    }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} $^{***}$ }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}$^{***}$  }  \\
-%	\emph{friend} &  \emph{of John} \\ \hline
-%noun    &    relative clause      
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf} $^{***}$    }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf} $^{**}$ }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}$^{***}$  }  \\
-%	\emph{books} & \emph{that you read} \\ \hline
-%complementizer    &    S        
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf} $^{***}$    }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}  $^{**}$   }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf}$^{***}$ }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}$^{**}$  }  \\
-%	\emph{that} & \emph{Mary has arrived} \\ \hline
-%verb    &    PP         
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf} $^{***}$    }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}  $^{***}$   }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}$^{**}$  }  \\
-%	\emph{went} & \emph{to school} \\ \hline
-%want    &    VP        
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} $^{***}$    }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf} $^{**}$    }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} $^{***}$ }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}$^{***}$  }  \\
-%	\emph{wants}   &  \emph{to leave} \\ \hline
-% \hline
-%%comp. &
-%%\multirow{3}{*}{NP}&
-%%	AP &
-%    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
-%\end{tabular}
-%	\end{center}
-%	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%For each correlation, we provide an English example.
-%%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%%Not all correlations are satisfied by every natu
-%}\label{table:corr-dryer}
-%\end{table*}
-%
-
-%
-%\begin{table}
-%	\begin{center}
-%%		\setlength\tabcolsep{1.5pt}
-%\begin{tabular}{|c|ll|c|cc|ccc}
-%	\hline
-%	%&
-%	&	\multicolumn{2}{|c|}{Correlates with...}   &         \\ 
-%	&	verb & object     & & &   \\ 
-%	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	Adposition    &    Noun Phrase
-%\\
-%	&		\emph{to}            & \emph{a friend} &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
-%\\
-%	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
-% \\
-%	&	\emph{has}          & \emph{written}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
-%\\
-%	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
-%\\
-%	&	\emph{books} & \emph{that you read}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
-%\\
-%	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-%	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
-%\\
-%	&	\emph{went} & \emph{to school}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
-%\\
-%	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
-% \hline
-%%comp. &
-%%\multirow{3}{*}{NP}&
-%%	AP &
-%%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-%\end{tabular}
-%	\end{center}
-%\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
-%	%Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-%%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
-%Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
-%%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%%Not all correlations are satisfied by every natu
-%}\label{table:corr-resu}
-%\end{table}
-%
-%
-%
-%
 
@@
 	& \emph{wants}   &  \emph{to leave}  &	\emph{to leave} & \emph{want} \\ \hline \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
 \end{tabular}
@@
 
-%TODO at some point need to foreshadow more of the Greenberg correlations
-
-%
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.45\textwidth]{../results/correlations/figures/coverage-ground_eff-circles.pdf}
-%	\caption{For each language, we show which of the eight correlations (columns) are satisfied in the actual language (right), and in grammars optimized for efficiency (left).}
-%	\label{fig:per-lang}
-%\end{figure}
-%
-
-%\subsection{Counterfactual Grammars}
 We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
@@
 
-%The sentences in Figure~\ref{fig:arabic-japanese} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words.
-For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} if the verb, rather than e.g.
-the subject or an indirect object.
-A dependency arc is drawn from a \key{head} (e.g.
-\emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g.
-\emph{tegami-o}).
-Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}.
-Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.
+%TODO also need to say these are syntax \& principle of compositionality
 
@@
 The speaker will order it as \emph{Mary saw John} (Figure~\ref{fig:comm}, top) in English, and as \emph{Mary John saw} (Figure~\ref{fig:comm}, bottom) in Japanese.
-%
-%(1)
-%\begin{dependency}[theme=simple]
-%    \begin{deptext}
-%    Mary \& saw \& John \\
-%    \end{deptext}
-%    \depedge{2}{1}{nsubj}
-%    \depedge{2}{3}{obj}
-%\end{dependency}
-%(2)
-%\begin{dependency}[theme=simple]
-%    \begin{deptext}
-%    Mary \& John  \& saw \\
-%    \end{deptext}
-%    \depedge{3}{1}{nsubj}
-%    \depedge{3}{2}{obj}
-%\end{dependency}
-%
 The listener receives a sentence -- a linear string of words -- and decodes the underlying message in accordance with the grammar of the language (Figure~\ref{fig:comm}, right).
-%Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
-%If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).
-
-%The listener receives this string of words.
-%By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
-%The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
-%Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
-
-%All natural languages have some degree of word order regularities that are specified in their grammar.
-%For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
-%That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
-%Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
-%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
-%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
+
 The grammars of human languages provide consistent ordering rules for all syntactic relations.
@@
 
-%\subsection{Efficiency in Word Order}
-%Zipf
-
-%We formalize Zipf's Forces of Unification and Diversification in
-
-%Prior computational tests of efficiency explanations for various aspects of language have
-
-%In information-theoretic formalizations of Zipf's ideas, the efficiency of a
-
-%In Information Theory, the usefulness of a communication channel 
 The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
-%This idea formalizes Zipf's forces of Unification and Diversification
 Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
+Common to these accounts is the idea of maximizing the amount of information that linguistic forms provide about meanings, while constraining the diversity of forms.
 
@@
 
-%Such notions of efficiency have been proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems.
-%, taking dependency structures as the underlying meanings to be conveyed.
-%Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
-%Word meaning \cite{regier2015word}
-
-%In our model, both aspects of efficiency will depend on the grammar:
-%Grammars differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
-%This reflects the \key{parseability} of the language.
-%%Maximizing the information provided by utterances about meanings is a standard formalization of Zipf's Force of Diversification~\cite{frank2012predicting,regier2015word}.
-% also goodman2013knowledge,
-%The degree to which listeners can reconstruct dependency structures from an utterance is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
-%\begin{equation}
-%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
-%\end{equation}
-%where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
-%This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences.
-%It is large when they can mostly be recovered, and small if sentences have high amounts of ambiguity about the dependency structure.
-%In line with Zipf's Force of Diversification, it is maximized when trees can be decoded fully unambiguously from utterances.
-
-%This idea formalizes Zipf's forces of Unification and Diversification
-%Accordingly, computational tests of functional explanations~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge} have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication:
-
-%With these choices, our objective function is equivalent to that proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems.
-%, taking dependency structures as the underlying meanings to be conveyed.
-%These works mainly differ in the cost function applied: can use the number of different utterances (CITE), the complexity of meanings \cite{kemp2012kinship}, and the entropy of possible utterances \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory}.
-
-%CITE other syntactic work? LFG
-%TODO the following sentences should go somewhere
-%By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
-%The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
-%Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
-
 \section{Results}
@@
 (2) whether this process of optimization accounts  for Greenberg's order correlation universals.
-%We formalize word order grammars as specifying how different types of dependents (subjects, objects, etc.) are ordered, employing a cross-linguistically meaningful inventory of relations~\cite{nivre2015universal}.
 
@@
 
-%Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
-%Languages appear to be attracted toward these points and away from the region of the baseline languages.
-We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
-
 \begin{figure}
@@
 Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).}\label{fig:per-lang}
-	% efficiency optimization predicts Greenberg's correlations.
-For every one of the 51 languages, we show which universals are satisfied in the actual grammar (left), and in the grammar optimized for efficiency (right).
-Efficiency optimization replicates the universals at rates similar to the real grammars.
 \end{figure}
-
-%
-%
-%
-%
-%\begin{table}
-%	\begin{center}
-%%		\setlength\tabcolsep{1.5pt}
-%\begin{tabular}{|c|ll|c|cc|ccc}
-%	\hline
-%	%&
-%	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
-%	&	verb & object     & & &   \\ 
-%	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
-%				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
-%	&		\emph{to}            & \emph{a friend} &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
-%	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
-%	&	\emph{has}          & \emph{written}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
-%	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
-%	&	\emph{books} & \emph{that you read}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
-%	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-%	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
-%	&	\emph{went} & \emph{to school}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
-%	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
-% \hline
-%%comp. &
-%%\multirow{3}{*}{NP}&
-%%	AP &
-%%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-%\end{tabular}
-%	\end{center}
-%\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
-%	%Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-%%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
-%Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
-%%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%%Not all correlations are satisfied by every natu
-%}\label{table:corr-resu}
-%\end{table}
-%
-%
-%
-%
-
-%
-%\begin{table}
-%	\begin{center}
-%%		\setlength\tabcolsep{1.5pt}
-%\begin{tabular}{|c|ll|c|cc|ccc}
-%	\hline
-%	%&
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	
-%				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
-%	&& &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
-%  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
-%  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
-%  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
-%  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
-%  &&&\\ \hline
-%	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
-%  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	  
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
-%		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
-%  &&&\\ \hline
-% \hline
-%%comp. &
-%%\multirow{3}{*}{NP}&
-%%	AP &
-%%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-%\end{tabular}
-%	\end{center}
-%\caption{OPTION 1: Efficiency optimization accurately predicts the Greenbergian correlations.
-%	%Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-%%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
-%Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
-%%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%%Not all correlations are satisfied by every natu
-%}\label{table:corr-resu}
-%\end{table}
 
@@
 \caption{Efficiency optimization accurately predicts the Greenbergian correlations.
-	%Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
 For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
 Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
-%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%Not all correlations are satisfied by every natu
 }\label{table:corr-resu}
@@
 
-%The Greenbergian correlations have been documented in great detail by Dryer \cite{dryer1992greenbergian}, drawing on 625 languages.
-%For most of these correlations, it is straightforward to test which of our grammars satisfy them (see \textit{Materials and Methods}).
-%Collapsing some correlations that cannot be tested individually, and excluding those for which Universal Dependencies does not provide annotation, we obtained eight correlations  (see \textit{Materials and Methods}).
-%%We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages.
-%%Most correlations can be straightforwardly implemented in Universal Dependencies (UD), allowing us to check which correlations a word order grammar satisfies.
-%%Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages or are not annotated reliably in UD.
-%%Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations.
-
 We compare grammars optimized for efficiency with the actual grammars of the 51 languages, as extracted from the orderings attested in the original corpora and validated against expert judgments (see \textit{Materials and Methods}).
@@
 This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
-%Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is itself a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
 Indeed, directly modeling efficiency optimization predicts DLM, and appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
-%Our work thus provides the first direct confirmation of the idea of DLM as an approximation to efficiency optimization -- DLM emerges as a by-product of efficiency maximization.
-
-%The eight correlations are shown in Table~\ref{table:corr-dryer}.
-
-%%We compare the prevalence of the word order correlations in optimized languages (right) to their prevalence in the real languages (left).
-%
-%
-%
-%
-%%In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
-%We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
-%We estimated the posterior distribution of the number of correlations that the optimization process correctly predicts.
-%We conducted the same analysis for grammars optimized for predictability and parseability separately~(Figure~\ref{fig:posterior}).
-%%Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.
-%Efficiency optimization predicts all eight correlations to be universal with posterior probability $0.9911$.
-%The estimated posterior probability that efficiency optimization predicts less than all eight correlations is $0.0242$.
-The probability that it predicts less than seven of the correlations is $<10^{-4}$.
-%Predictability and parseability individually also predict six to eight correlations.
-%
-%
-%%
-
-%TODO better than this:
-%
-%Finally, we compare to the predictions made by optimizing only for predictability, and only for parseability.
-%
-%We further compared to the predictions made individually by optimizing for predictability and parseability.
-%We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
-%We estimated the posterior distribution of the number of correlations that the optimization process correctly predicts.
-%We conducted the same analysis for grammars optimized for predictability and parseability separately.
-%We used all eight runs of optimization to maximize statistical precision.
-%For every one of the correlations, we report the posterior estimated.
-%(These numbers control for random variation between language families, and are thus not directly comparable to data from the 51 languages).
-%
-%
-%
 

### 2019-05-02 results/grammars/comparison-table.tex
WALS

Afrikaans  & DH  & ?  & HD  & ?  & HD  & ?  & --  & ?  & HD  & ?  & HD  & ?  & \\ 
Anc.Grk.  & DH  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & \\ 
Arabic  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & ?  & HD  & HD  & HD  & HD  & \\ 
Basque  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & \\ 
Belarusian  & HD  & *  & HD  & ?  & HD  & ?  & HD  & HD  & HD  & HD  & HD  & *  & \\ 
Bulgarian  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & *  & HD  & HD  & \\ 
Catalan  & HD  & HD  & HD  & HD  & HD  & ?  & HD  & HD  & HD  & HD  & HD  & ?  & \\ 
Chinese  & HD  & HD  & HD  & *  & DH  & ?  & DH  & DH  & DH  & DH  & DH  & DH  & \\ 
Coptic  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & \\ 
Croatian  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & ?  & HD  & *  & HD  & ?  & \\ 
Czech  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & *  & HD  & ?  & \\ 
Danish  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & \textit{HD} & DH  & HD  & HD  & \\ 
Dutch  & DH  & *  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & DH  & *  & \\ 
English  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & *  & HD  & HD  & \\ 
Estonian  & HD  & HD  & DH  & DH  & HD  & HD  & \textit{DH} & HD  & DH  & DH  & HD  & HD  & \\ 
Finnish  & HD  & HD  & DH  & DH  & HD  & HD  & \textit{DH} & HD  & DH  & DH  & HD  & HD  & \\ 
French  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & \\ 
Galician  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & \\ 
German  & HD  & *  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & DH  & *  & \\ 
Gothic  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & \\ 
Greek  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & ?  & \\ 
Hebrew  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & ?  & \\ 
Hindi  & DH  & DH  & DH  & DH  & HD  & HD  & DH  & *  & DH  & DH  & DH  & ?  & \\ 
Hungarian  & \textit{DH} & HD  & DH  & DH  & HD  & HD  & HD  & *  & DH  & DH  & DH  & ?  & \\ 
Indonesian  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & \\ 
Irish  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & \\ 
Italian  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & ?  & \\ 
Japanese  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & DH  & \\ 
Korean  & DH  & DH  & DH  & DH  & \textit{HD} & DH  & DH  & DH  & DH  & DH  & DH  & ?  & \\ 
Latin  & DH  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & DH  & ?  & \\ 
Latvian  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & DH  & DH  & DH  & ?  & \\ 
Lithuanian  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & DH  & DH  & DH  & ?  & \\ 
Marathi  & DH  & DH  & DH  & DH  & HD  & *  & DH  & DH  & DH  & DH  & DH  & ?  & \\ 
Norwegian  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & *  & HD  & ?  & \\ 
O.C.Slav.  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & \\ 
Persian  & DH  & DH  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & DH  & ?  & \\ 
Polish  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & ?  & \\ 
Portuguese  & HD  & HD  & HD  & HD  & HD  & ?  & HD  & HD  & HD  & HD  & HD  & ?  & \\ 
Romanian  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & ?  & \\ 
Russian  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & ?  & \\ 
Serbian  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & \\ 
Slovak  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & HD  & ?  & \\ 
Slovenian  & HD  & HD  & HD  & HD  & HD  & ?  & HD  & ?  & HD  & *  & HD  & ?  & \\ 
Spanish  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & \\ 
Swedish  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & \textit{HD} & DH  & HD  & HD  & \\ 
Tamil  & DH  & DH  & DH  & DH  & DH  & *  & \textit{HD} & DH  & DH  & DH  & DH  & DH  & \\ 
Telugu  & DH  & DH  & DH  & DH  & DH  & *  & DH  & DH  & DH  & DH  & DH  & ?  & \\ 
Turkish  & DH  & DH  & DH  & DH  & DH  & *  & DH  & DH  & DH  & DH  & DH  & DH  & \\ 
Ukrainian  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & HD  & ?  & HD  & ?  & \\ 
Urdu  & DH  & DH  & DH  & DH  & HD  & HD  & HD  & *  & DH  & DH  & DH  & ?  & \\ 
Vietnamese  & HD  & HD  & HD  & HD  & \textit{DH} & HD  & --  & HD  & HD  & HD  & HD  & HD  & \\

### 2019-05-02 second-draft/pnas-word-order-si.tex
WALS

@@
 \DeclareMathOperator{\E}{\mathop{\mathbb{E}}}
+
+\usepackage{longtable}
+
+\usepackage{natbib}
 
@@
 
-\section{UD Formalization}
+\section{Formalization of Correlations}
 
-\begin{table*}
+Here we describe how we selected the correlations in Table 1 of the main paper, and how we formalized these using syntactic relations defined by Universal Dependencies.
+
+We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding recent objections by~\cite{dunn2011evolved}, see follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
+Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
+\cite{dryer1992greenbergian} formulated these universals as correlations between the order of objects and verbs and the orders of other syntactic relations.
+We test our ordering grammars for these correlations by testing whether the coefficients for these syntactic relations have the same sign as the coefficient of the verb-object relation.
+Testing correlations is therefore constrained by the degree to which these relations are annotated in UD.
+The verb-object relation corresponds to the  \emph{obj} relation defined by UD.
+While most of the other relations also correspond to UD relations, some are not annotated reliably.
+We were able formalize eleven out of Dryer's seventeen correlations in UD. Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations.
+
+We excluded four correlations that are not annotated reliably in UD, and are only relevant to some of the world's languages: Question particles, plural words (i.e., independent plural markers), and articles.
+All three types of elements occur at most in parts of the 51 UD languages, and none of them is annotated reliably in those languages where they occur.
+Among these three types of elements, the one most prominent in our sample of 51 languages is articles (occurring in many European languages); UD subsumes them under the \emph{det} relation, which is also used for other frequent elements, such as demonstratives and quantifiers.
+
+We also excluded the verb-manner adverb correlation.
+UD does not distinguish manner adverbs from other elements labeled as adverbs, such as sentence-level adverbs and negation markers, whose ordering is very different from manner adverbs.
+All types of adverbs are unified under the \emph{advmod} relation.
+In the real orderings in our sample of 51 UD languages, the dominant ordering of \emph{advmod} almost always matches that of subjects -- that is, \emph{advmod} dependents are ordered after the verb only in VSO languages.
+This ordering behavior is very different from that documented for manner adverbs by Dryer.
+
+We further excluded the verb-subject correlation, which is not satisfied by much more than half of the world's languages (51 \% among those with annotation in \cite{wals-81}, with clear violation in 35.4 \%).
+It is satisfied only in 33\% of our sample of 51 UD languages, as quantified using the grammars we extracted.
+Dryer \cite{dryer1992greenbergian} counts this as a correlation since he describes the distribution of subject order as an interaction between a weak correlation with object order, and a very strong dominance principle favoring SV/SO orderings.
+We focus on the modeling of correlations, and such dominance principles are beyond the scope of our experiments.
+We therefore excluded this correlation.
+%We note that efficiency makes predictions compatible with the pattern observed in our sample of 51 languages, potentially in line with the noisy-channel explanation of SVO described by~\cite{gibson2013noisy} (see Table~\ref{tab:all-predictions}).
+
+TODO collapsing
+
+The resulting operationalization is shown in Table~\ref{table:greenberg-dryer}.
+
+\begin{table*}[ht]
 	\begin{center}
-\begin{tabular}{|ll|l|l|l|ll|l|}
+\begin{tabular}{|ll|l|l}
 	\hline
-	%&
-	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       &     \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\
-	verb & object &      & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       \\ 
+	verb & object &    \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
 %adp. &
-adposition    &    NP    &    lifted\_case    &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
-copula    &    NP    &    lifted\_cop    &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
-auxiliary    &    VP    &    aux    &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
-noun    &    genitive    &    nmod    &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
-noun    &    relative clause    &    acl    &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
-complementizer    &    S    &    lifted\_mark    &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
-verb    &    PP    &    obl    &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
-want    &    VP    &    xcomp    &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
-verb    &    subject    &    nsubj    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
-verb    &    manner adverb    &    advmod    &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
+adposition    &    NP    &    lifted\_case   & G3, G4   \\ \hline
+copula  verb  &    predicate    &    lifted\_cop   & --    \\\hline
+`want'    &    VP    &    xcomp   & G15   \\\hline
+tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{aux}   & G16, G13  \\
+negative auxiliary    &    VP    &     \\ \hline
+complementizer    &    S    &   \multirow{2}{*}{lifted\_mark}  & --    \\
+adverbial subordinator & S & \\ \hline
+noun    &    genitive    &    nmod & G2, G23   \\ \hline
+noun    &    relative clause    &    acl   &  G24    \\ \hline
+adjective & standard of comparison & \multirow{2}{*}{obl}\\
+verb    &    PP    &    & G22   \\
  \hline
@@
 %	AP &
-    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+    \multicolumn{3}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
 \end{tabular}
 	\end{center}
-	\caption{Greenbergian Correlations.
-Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%	For clarity, we organize the correlations by the category of the `verb patterner'.
-	For each correlation, we give our operationalization in terms of UD.
-%A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
-	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
-	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
+	\caption{Greenbergian Correlations according to Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies.
+We also provide the corresponding numbers in Greenberg's original study.}\label{table:greenberg-dryer}
 \end{table*}
+
+%croft2011greenbergian
+%dryer2011evidence
+
+\section{Supporting Analyses for Study 1}
+
+\paragraph{Other tests, maybe som viz}
+
+\paragraph{Pareto Curve}
+
+It has to be convex, so don't need to do more.
 
@@
 
+We also report results for grammars optimized for predictability and parseability individually.
+
+Further, we report results for grammars optimized for dependency length minimization (DLM).
+DLM can be seen as a heuristic approximation to efficiency.
+We also found that DLM is predicted by efficiency optimization.
+Indeed, we show that optimizing for DLM makes predictions similar to efficiency optimization.
+
 \begin{table*} % TODO figure out away to make this table more aesthetic
@@
 
-\tiny{
+\small{
 \begin{tabular}{|l|l|l|ll|l|l|}
@@
 nsubj  &  33   &   \textbf{29}$^{***}$   &   53   &   \textbf{7}$^{***}$   &   \textbf{23}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
-advmod  &  35   &   51   &   \textbf{18}$^{***}$   &   47   &   \textbf{39}$^{**}$  & $>50\%$ for manner adverbs; $\approx 50\%$ for negation, TAM, intensifiers \citep{dryer1992greenbergian} \\
 
@@
 advcl  &  86   &   \textbf{84}$^{***}$   &   51   &   \textbf{67}$^{***}$   &   \textbf{67}$^{*}$  & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
+ccomp  &  86   &   \textbf{88}$^{***}$   &   \textbf{70}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{77}$^{**}$ & $>50\%$ (cf.
+\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
+csubj  &  79   &   \textbf{78}$^{***}$   &   60   &   \textbf{66}$^{***}$   &   \textbf{68}$^{**}$  & $>50\%$ (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
 amod  &  51   &   \textbf{84}$^{***}$   &   42   &   \textbf{53}$^{*}$   &   51   & $\approx 50\%$ \citep{dryer1992greenbergian} \\
-nummod  &  37   &   \textbf{74}$^{**}$   &   49   &   52   &   52 & $\approx 50\%$ \citep[][89A vs. 83A]{wals} \\
+nummod  &  37   &   \textbf{74}$^{**}$   &   49   &   52   &   52 & $\approx 50\%$ \citep[][89A, 83A]{wals} \\ %
 
@@
 lifted\_cc  &  71   &   \textbf{83}$^{***}$   &   51   &   \textbf{81}$^{***}$   &   \textbf{71}$^{***}$  & Unknown \\
-ccomp  &  86   &   \textbf{88}$^{***}$   &   \textbf{70}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{77}$^{**}$ & Unknown (cf.
-\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
-csubj  &  79   &   \textbf{78}$^{***}$   &   60   &   \textbf{66}$^{***}$   &   \textbf{68}$^{**}$  & Unknown (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
 expl  &  12   &   36   &   62   &   40   &   53  & Unknown \\
@@
 compound  &  66   &   \textbf{61}$^{*}$   &   48   &   \textbf{58}$^{*}$   &   52  & Uninterpretable \\
-det  &  31   &   \textbf{71}$^{***}$   &   \textbf{56}$^{*}$   &   48   &   54  & Uninterpretable (conflates articles and others) \\%$.50\%$ for articles, $\approx 50\%$ for others, \citep{dryer1992greenbergian} \\ %TODO what else is conflated in det?
-why do we keep advmod but not det?
+det  &  31   &   \textbf{71}$^{***}$   &   \textbf{56}$^{*}$   &   48   &   54  & Uninterpretable \\%$.50\%$ for articles, $\approx 50\%$ for others, \citep{dryer1992greenbergian} \\
 dislocated  &  47   &   61   &   \textbf{68}$^{**}$   &   55   &   58  & Uninterpretable \\
 dep  &  56   &   \textbf{64}$^{*}$   &   52   &   58   &   \textbf{58}$^{*}$   & Uninterpretable \\
+advmod  &  35   &   51   &   \textbf{18}$^{***}$   &   47   &   \textbf{39}$^{**}$  & Uninterpretable \\
 
@@
 In the last column, we indicate what direction would be expected typologically.
+The first section correspond to universals from Dryer.
 ``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
@@
 
-\section{Results on Best Optimized Grammars}
-
-In the main part, we reported the percentage of optimized grammars supporting correlations.
-Here we report these numbers for the grammars that optimize each objective most successfully.
-That is, for each language and objective, we chose among the eight optimized grammars the one that achieves the best value for the objective function on the validation partition of the treebanks -- that is, the one where the optimization procedure was most successful.
-
-Numbers for these languages are shown in Table~\ref{table:results-dryer-best-only}.
-%Confirming the results from the main paper, 
-As expected, the correlations mostly have stronger prevalences than when considering all optimized languages.
-For Dependency Length, the correlations that it predicts mostly have prevalences close to 100 \%, whereas both the UD languages and efficiency-optimized languages show somewhat lower percentages.
-That is, dependency length mostly overshoots for the correlations that it predicts correctly.
-
-\begin{table*}
-	\begin{center}
-\begin{tabular}{|ll|l|l|l|ll|l|}
-	\hline
-	%&
-	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       &     \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\
-	verb & object &      & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%adp. &
-adposition    &    NP    &    lifted\_case    &    86    &    98    &    69    &    88    &    83   \\
-copula    &    NP    &    lifted\_cop    &    94    &    98    &    58    &    85    &    74   \\
-auxiliary    &    VP    &    aux    &    88    &    84    &    90    &    61    &    73   \\
-noun    &    genitive    &    nmod    &    80    &    98    &    75    &    88    &    91   \\
-noun    &    relative clause    &    acl    &    80    &    100    &    69    &    88    &    90   \\
-complementizer    &    S    &    lifted\_mark    &    76    &    100    &    76    &    82    &    87   \\
-verb    &    PP    &    obl    &    88    &    98    &    88    &    75    &    91   \\
-want    &    VP    &    xcomp    &    88    &    100    &    92    &    94    &    96   \\
-verb    &    subject    &    nsubj    &    33    &    25    &    53    &    10    &    21   \\
-verb    &    manner adverb    &    advmod    &    35    &    61    &    27    &    55    &    36   \\
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-\end{tabular}
-	\end{center}
-	\caption{Percentages of grammars supporting correlations, among grammars that best optimize each objective on treebanks of a given language.}\label{table:results-dryer-best-only}
-\end{table*}
-
 \begin{figure}
@@
 \end{figure}
-
-\section{Further Statistical Analyses}
-
-In the main results table, we reported the fraction of optimized languages satisfying a correlation.
-We show that essentially the same predicted fractions are obtained when controlling for the imbalance in distribution of language families.
-
-To this end, we used the logistic mixed-effects regressions that were used to estimate significance to compute the expected prevalence of a correlation controlling for language families.
-To obtain predictions controlling for languages and language families, we plot the logit-transformed posterior of $\alpha$ in Figure \ref{fig:posterior-prevalence}.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.7]{../results/correlations/figures/posterior_joint_with_ud.pdf}
-	\caption{Posteriors of the prevalences of the correlations among languages optimized for Predictability (red), parseability (green), and efficiency (blue), controlling for languages and language families, with prevalence in the UD languages.}
-    \label{fig:posterior-prevalence}
-\end{figure}
-
-\begin{figure}
-    \centering
-%     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-efficiency.pdf}
- %    \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-langmod.pdf}
-  %   \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}
-   %  \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-depl.pdf}
-     \includegraphics[scale=.45]{../results/correlations/figures/posterior-satisfied-universals-together.pdf}
-
-	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks, computed from a mixed-effects logistic regression jointly modeling all ten dependencies.
-Efficiency predicts all ten correlations with high posterior probability.}
-    \label{fig:posterior-satisfied-count}
-\end{figure}
-
-We further evaluate the word order predictions of efficiency, showing that efficiency is most successful in predicting correlations in the direction found in the UD corpora.
-We constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency in languages optimized for efficiency, with random effects for language and language family, correlated across the ten dependencies.
-We conducted the same analysis for predictability, parseability, and dependency length.
-We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
-The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
-The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$.
-The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.
-For dependency length, the posterior puts much of the probability mass on predicting only nine of the correlations; predictability and parseability predict significantly less correlations.
-
-%\begin{figure}
-%    \centering
-%    \includegraphics[scale=.3]{figures/posterior-satisfied-universals-efficiency.pdf}
-%     \includegraphics[scale=.3]{figures/posterior-satisfied-universals-langmod.pdf}
-%     \includegraphics[scale=.3]{figures/posterior-satisfied-universals-parseability.pdf}
-%    \includegraphics[scale=.3]{figures/posterior-satisfied-universals-depl.pdf}
-%	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks.}
-%    \label{fig:posterior-satisfied-count}
-%\end{figure}
-
-%We argued that efficiency predicts all 10 correlation universals in the direction manifested in the UD corpora, as for each universal, the posterior probability that less than $50 \%$ of the optimized grammars satisfy the correlation was less than $0.05$.
-%To address potential multiple comparisons problems, we use a single logistic model to show that efficiency predicts all ten correlations with high posterior probability.
-
-%For each objective function, we constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency, with random effects for language and language family, correlated across the ten dependencies.
-
-%We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
-%The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
-%The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$.
-The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.
 
@@
 
+The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
+We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (TODO check).
+
 \section{Joint optimization of grammar, parser, and language model parameters}
+
+\subsection{Differentiable Ordering Grammars}
+
+In order to carry out optimization of grammars, we extend the space of grammars to include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
+This makes efficiency a differentiable function of the grammar parameters, and enables optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.
+
+The parameters of such a differentiable word order grammar are as follows. 
+For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
+Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
+Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
+184) without replacement.
+
+If $a_\tau \in \{0, 1\}$, and the distances between values of $b_\tau$ (for different $\tau$) become very large, such a differentiable grammar becomes deterministic, assigning almost full probability to exactly one ordering for each syntactic structure.
+In this case, the grammar can be converted into an equivalent grammar of the form described in Materials and Methods, by extracting a single parameter in $[-1, 1]$ for each relation $\tau$.
+
+\subsection{Extracting Grammars from Datasets}
+We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
+To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
+We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.
+
+\subsection{Optimizing Grammars for Efficiency}\label{sec:optim-eff}
+
+\textbf{rewrite this}
 
@@
 
-\section{Implementation Details}
+%We extract grammars for the actual languages using a Bayesian data analysis method.
+%We specify a simple Bayesian prior over differentiable ordering grammars:
+%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
+%for each $\tau$.
+%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
+%We extract
 
-\paragraph{Locality}
-In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
+\section{Neural Network Architectures}
 
-\paragraph{Language Model}
+In this section, we describe the details of the neural network architectures.
+All choices described in this section were fixed before evaluating word order properties and the efficiency of real language.
+
+%\paragraph{Locality}
+%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
+
+\paragraph{Estimating Predictability}
 We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
@@
 
-\paragraph{Parser}
+\paragraph{Estimating Parseability}
 We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}.
@@
 To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags.
-Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged.
-Parsing from POS tags reduces early overfitting.
+Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged; parsing from POS tags prevents early overfitting.
 This decision was made before evaluating word order properties.
@@
 \paragraph{Hyperparameters}
-For locality, predictability, and parseability, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
-Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function on the language.
+
+\textbf{TODO study 1 only has 1 set of parameters}
+
+Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
+For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
+Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (predictability or parseability) on the language.
 We then used this setting for creating optimized word order grammars.
-% TODO more detail and maybe an example here
-% TODO all design changes were made before examining word order universal results
 
-% TODO Discuss iconicity somewhere somehow
+\textbf{TODO random seeds for efficiency}
 
 All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
-We do not use pretrained embeddings \citep{peters2018deep}: these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
+We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
 
@@
 We employ two common variance reduction methods to improve the estimator~(\ref{eq:estimator}), while keeping it unbiased.
-For dependency length and predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word (and its head, in the case of dependency length).
+%For dependency length and 
+For predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word (and its head, in the case of dependency length).
 We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:estimator}) with lower variance.
@@
 
-We update word order grammar parameters $\theta$ using SGD with momentum.
+We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
 For the language model parameters $\phi$, we use plain SGD. 
@@
 We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
-This maximizes statistical power for detecting and quantifying correlations between the \textit{obj} relation and other relations.
+This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.
 
@@
 Here we take up the question of the extent to which our results are dependent on the particular parser and language model used in the optimization process.
-We want to know: when we optimize a word order grammar for parseability, have we produced a language which is highly parseable \emph{in general}, or one which is highly parseable \emph{for a specific parser}?
-We wish to argue that natural language syntax is optimized for parseability in general, meaning that syntactic trees are highly recoverable from word orders in principle.
+We want to know: when we optimize a word order grammar for efficiency, have we produced a language which is highly efficient \emph{in general}, or one which is highly efficient \emph{for a specific parser}?
+We wish to argue that natural language syntax is optimized for efficiency in general, meaning that syntactic trees are highly recoverable from word orders in principle.
 If it turns out that our optimized languages are only optimal for a certain parser from the NLP literature, then we run the risk of circularity: it may be that the reason this parser was successful in the NLP literature was because it implicitly encoded word order universals in its inductive biases, and thus it would be no surprise that languages which are optimized for parseability also show those universals.
@@
 
-\subsection{CYK Parsers}
+\subsection{CKY Parsers}
 
@@
 We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
-For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CYK algorithm.
+For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm.
 
@@
 
-Our efficiency objective comes down to
-
-If the difference between random and optimized languages is due to data sparsity, we expect that it decreases as the amount of training data is increased.
-If, on the other hand, there is an inherent difference in efficiency between random and optimized languages, we expect that the difference persists as training data is increased.
+Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
+If the difference between random and real grammars is due to data sparsity, we expect that it decreases as the amount of training data is increased.
+If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference persists as training data is increased.
 
@@
 Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.
+
+\textbf{do same for language model, and for the other random grammars}
 
@@
 
-\section{}
+\section{Study 2, Correlation between Universals and Efficiency}
 
-\bibliographystyle{acl_natbib}
+\begin{figure}
+    \centering
+
+    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
+    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
+    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
+
+   	\caption{.}
+    \label{fig:todo}
+\end{figure}
+
+\section{Comparing Extracted Grammars to Linguistic Judgments}
+
+% user@user-X510UAR:~/grammar-optim/results/grammars$ git add comparison-table.tex
+
+Here, we validate our grammars by comparing the dominant orders of six syntactic relations that are also annotated in WALS.
+
+WALS has data for 74\% of the entries, and lists a dominant order for 91\% of these.
+96~\% of entries where WALS records a dominant word order agree with the WALS judgment.
+
+$*$ indicates that WALS does not list a dominant order, ? indicates that WALS has no data, and $--$ indicates that the relation does not occur in the corpus.
+
+\small{
+\begin{longtable}{l||ll|ll|ll|ll|ll|ll|llllll}
+%	   for(dep in c("obj", "lifted_mark", "acl", "nmod", "obl")) {
+		   Language 
+		   &	\multicolumn{2}{|c|}{Objects} 
+		   &	\multicolumn{2}{|c|}{Adpositions} 
+		   &	\multicolumn{2}{|c|}{Compl.} 
+		   &	\multicolumn{2}{|c|}{Rel.Cl.} 
+		   &	\multicolumn{2}{|c|}{Genitive} 
+		   &	\multicolumn{2}{|c|}{PP}  \\ \hline\hline
+	\input{../results/grammars/comparison-table.tex}
+\end{longtable}
+}
+
+\bibliographystyle{unsrtnat}
 \bibliography{everything}

### 2019-05-02 results/correlations/correlationsByGrammar.R

@@ -1,3 +1,6 @@
+# Study 1: Correlation between efficiency and the number of satisfied correlations, for real grammars.
+
+
 
 library(tidyr)
 library(dplyr)
@@ -127,6 +130,32 @@ cor.test(u$Agree, 0.9*u$Surp_z+u$Pars_z)
 
 # Robustness check:
 cor.test(u$Agree, 0.9*u$Surp_z+u$Pars_z, method="spearman")
+cor.test(u$Agree, 0.9*u$Surp_z+u$Pars_z, method="kendal")
+
+
+plot = ggplot(u, aes(x=Agree, y=-(0.9*u$Surp_z+u$Pars_z)))
+plot = plot + geom_point()
+plot = plot + geom_smooth(method="lm")
+plot = plot + xlab("Satisfied Correlations")
+plot = plot + ylab("Efficiency")
+plot = plot + theme_bw()
+ggsave(plot, file="correlations-by-grammar/ground-corrs-efficiency.pdf")
+
+plot = ggplot(u, aes(x=Agree, y=-u$Surp_z))
+plot = plot + geom_point()
+plot = plot + geom_smooth(method="lm")
+plot = plot + xlab("Satisfied Correlations")
+plot = plot + ylab("Predictability")
+plot = plot + theme_bw()
+ggsave(plot, file="correlations-by-grammar/ground-corrs-predictability.pdf")
+
+plot = ggplot(u, aes(x=Agree, y=-u$Pars_z))
+plot = plot + geom_point()
+plot = plot + geom_smooth(method="lm")
+plot = plot + xlab("Satisfied Correlations")
+plot = plot + ylab("Parseability")
+plot = plot + theme_bw()
+ggsave(plot, file="correlations-by-grammar/ground-corrs-parseability.pdf")
 
 
 

### 2019-05-02 results/correlations/plot_families_num2-dlm.R

@@ -1,0 +1,120 @@
+# tiles-real_eff_large-byObj-restricted-viz-pred-large-dlm.R
+
+
+
+D$LanguageNumeric = as.numeric(D$Language_Ordered)
+
+D$FamilyPrint = as.character(D$Family)
+D = D %>% mutate(FamilyPrint = ifelse(FamilyPrint == "Malayo-Sumbawan", "Mal.-Sum.", as.character(FamilyPrint)))
+D = D %>% mutate(FamilyPrint = ifelse(FamilyPrint == "Sino-Tibetan", "Sin.-Tib.", as.character(FamilyPrint)))
+D = D %>% mutate(FamilyPrint = ifelse(FamilyPrint == "Viet-Muong", "Viet-M.", as.character(FamilyPrint)))
+
+
+DFam = D %>% group_by(FamilyPrint) %>% summarise(Start = min(LanguageNumeric), End = max(LanguageNumeric), Mean = mean(LanguageNumeric))
+
+DFam$yOffset = 0.2*(1:(nrow(DFam))) 
+D$yOffset=NULL
+D = merge(D, DFam %>% select(FamilyPrint, yOffset), by=c("FamilyPrint"))
+
+
+DLang = unique(D %>% select(Language_Ordered, iso_Ordered, LanguageNumeric, yOffset))
+
+
+plot_orders_real = ggplot(D %>% filter(Type == "Real Languages"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
+  geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
+#  scale_color_gradient() + #values=c("blue", "green")) +
+  theme_classic() +
+  #theme_bw() + 
+  theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
+                     axis.text.y=element_blank(),axis.ticks=element_blank(),
+                     plot.title=element_text(size=11)) +
+  theme(axis.title=element_blank()) + 
+  theme(legend.position="none") + labs(x=NULL) +
+  scale_x_continuous(breaks = NULL) +
+  scale_y_continuous(breaks = NULL)
+
+plot_orders_eff = ggplot(D %>% filter(Type == "Efficiency"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
+  geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
[…]

### 2019-05-02 results/grammars/compareToWALS.R

@@ -1,0 +1,160 @@
+
+library(tidyr)
+library(dplyr)
+library(lme4)
+library(ggplot2)
+library(forcats)
+
+# Note the complication that Serbian-Croatian is one language in WALS, and two languages in UD. In contrast, Hindi and Urdu are separate in both.
+
+wals_path = "~/mhahn_files/stnfclasses/deps/writeup/language.csv"
+wals = read.csv(wals_path)
+
+
+wals = wals %>% mutate(obj_w_ = as.character(X83A.Order.of.Object.and.Verb))
+wals = wals %>% mutate(lifted_mark_w_ = as.character(wals$X94))
+wals = wals %>% mutate(acl_w_ = as.character(wals$X90A))
+wals = wals %>% mutate(nmod_w_ = as.character(wals$X86))
+wals = wals %>% mutate(lifted_case_w_ = as.character(wals$X85))
+wals = wals %>% mutate(obl_w_ = as.character(wals$X84A))
+
+
+wals = wals %>% mutate(obj_w = as.character(as.factor(case_when(obj_w_ ==  "2 VO" ~ "HD", obj_w_ == "1 OV" ~ "DH"))))
+wals = wals %>% mutate(lifted_mark_w = as.character(case_when(lifted_mark_w_ == "1 Initial subordinator word" ~ "HD", lifted_mark_w_ == "2 Final subordinator word" ~ "DH")))
+wals = wals %>% mutate(acl_w = as.character(case_when(acl_w_ == "1 Noun-Relative clause" ~ "HD", acl_w_ == "2 Relative clause-Noun" ~ "DH")))
+wals = wals %>% mutate(nmod_w = as.character(as.factor(case_when(nmod_w_ == "2 Noun-Genitive" ~ "HD", nmod_w_ == "1 Genitive-Noun" ~ "DH"))))
+wals = wals %>% mutate(lifted_case_w = as.character(as.factor(case_when(lifted_case_w_ %in% c("1 Postpositions" )  ~ "DH", lifted_case_w_ %in% c("2 Prepositions") ~ "HD"))))
+wals = wals %>% mutate(obl_w = as.character(as.factor(case_when(obl_w_ %in% c("1 VOX","5 OVX" )  ~ "HD", obl_w_ %in% c("2 XVO", "4 OXV", "3 XOV") ~ "DH"))))
+
+
+languages_wals_mapping = read.csv("languages-wals-mapping.csv")
+wals$Language = NULL
+wals$iso_code = NULL
[…]

### 2019-05-04 second-draft/pnas-word-order-si.tex
si

@@
 While most of the other relations also correspond to UD relations, some are not annotated reliably.
-We were able formalize eleven out of Dryer's seventeen correlations in UD. Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations.
+We were able formalize eleven out of Dryer's seventeen correlations in UD.
+Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations:
+First, tense/aspect and negative auxiliaries are together represented by the \emph{aux} relation in UD.
+Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{lifted\_mark} relation.
+Third, we both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.
 
-We excluded four correlations that are not annotated reliably in UD, and are only relevant to some of the world's languages: Question particles, plural words (i.e., independent plural markers), and articles.
+We excluded three correlations that are not annotated reliably in UD, and are only relevant to some of the world's languages: Question particles, plural words (i.e., independent plural markers), and articles.
 All three types of elements occur at most in parts of the 51 UD languages, and none of them is annotated reliably in those languages where they occur.
@@
 It is satisfied only in 33\% of our sample of 51 UD languages, as quantified using the grammars we extracted.
-Dryer \cite{dryer1992greenbergian} counts this as a correlation since he describes the distribution of subject order as an interaction between a weak correlation with object order, and a very strong dominance principle favoring SV/SO orderings.
-We focus on the modeling of correlations, and such dominance principles are beyond the scope of our experiments.
-We therefore excluded this correlation.
+Dryer \cite{dryer1992greenbergian} counts this as a correlation since he describes the distribution of subject order as an interaction between a weak correlation with object order, and a very strong dominance principle favoring SV orderings.
+We focus on the modeling of correlations, and leave dominance principles to future research.
+We therefore excluded this correlation here.
 %We note that efficiency makes predictions compatible with the pattern observed in our sample of 51 languages, potentially in line with the noisy-channel explanation of SVO described by~\cite{gibson2013noisy} (see Table~\ref{tab:all-predictions}).
-
-TODO collapsing
 
@@
 	\begin{center}
-\begin{tabular}{|ll|l|l}
+\begin{tabular}{|ll|l|l|}
 	\hline
-	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       \\ 
+	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}   & \multirow{2}{*}{\cite{greenberg1963universals}}    \\ 
 	verb & object &    \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
 %adp. &
-adposition    &    NP    &    lifted\_case   & G3, G4   \\ \hline
-copula  verb  &    predicate    &    lifted\_cop   & --    \\\hline
-`want'    &    VP    &    xcomp   & G15   \\\hline
-tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{aux}   & G16, G13  \\
-negative auxiliary    &    VP    &     \\ \hline
-complementizer    &    S    &   \multirow{2}{*}{lifted\_mark}  & --    \\
-adverbial subordinator & S & \\ \hline
-noun    &    genitive    &    nmod & G2, G23   \\ \hline
-noun    &    relative clause    &    acl   &  G24    \\ \hline
-adjective & standard of comparison & \multirow{2}{*}{obl}\\
-verb    &    PP    &    & G22   \\
+adposition    &    NP    &    case   & 3, 4   \\ \hline
+copula  verb  &    predicate    &    cop   & --    \\\hline
+`want'    &    VP    &    xcomp   & 15   \\\hline
+tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{aux}   & 16, 13  \\
+	negative auxiliary    &    VP    &    & -- \\ \hline
+complementizer    &    S    &   \multirow{2}{*}{mark}  & --    \\
+	adverbial subordinator & S &  & -- \\ \hline
+noun    &    genitive    &    nmod & 2, 23   \\ \hline
+noun    &    relative clause    &    acl   &  24    \\ \hline
+	adjective & standard of comparison & \multirow{2}{*}{obl} & --\\
+verb    &    PP    &    & 22   \\
  \hline
@@
 
-   	\caption{.}
+	\caption{Correlation between the number of satisfied correlations ($x$-axis) and efficiency, parseability, and predictability ($y$-axis), for the 51 real languages.}
     \label{fig:todo}
@@
 Here, we validate our grammars by comparing the dominant orders of six syntactic relations that are also annotated in WALS.
+Among the eight relations that we were able to test, five are annotated in WALS.
+Here, we compare our grammars with WALS on these five relations, and the verb-object relation.
 
+WALS has some data for all but seven languages.\footnote{Note that Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
+In the table, we compare those with the grammar we extracted for Croatian, noting that it agrees with the Serbian grammar.}
 WALS has data for 74\% of the entries, and lists a dominant order for 91\% of these.
-96~\% of entries where WALS records a dominant word order agree with the WALS judgment.
+The grammars we extracted from the corpora agree with WALS in 96~\% out of those cases where WALS records a dominant word order.
 
-$*$ indicates that WALS does not list a dominant order, ? indicates that WALS has no data, and $--$ indicates that the relation does not occur in the corpus.
+For each of the six syntactic relation, the first column provides the ordered coded in the extracted grammar; the second column provides the order coded in WALS (DH for dependent-head, HD for head-dependent order).
+`?' indicates that WALS has no data.
+$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
+Finally, `--' indicates that the relation does not occur in the corpus.
 

### 2019-05-04 second-draft/pnas-word-order-si.tex
si

@@
 	\hline
-	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}   & \multirow{2}{*}{\cite{greenberg1963universals}}    \\ 
+	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{UD Relation}   & \multirow{2}{*}{\cite{greenberg1963universals}}    \\ 
 	verb & object &    \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
@@
 noun    &    relative clause    &    acl   &  24    \\ \hline
-	adjective & standard of comparison & \multirow{2}{*}{obl} & --\\
+	adjective & std. of comp. & \multirow{2}{*}{obl} & --\\
 verb    &    PP    &    & 22   \\
@@
 	\caption{Greenbergian Correlations according to Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies.
-We also provide the corresponding numbers in Greenberg's original study.}\label{table:greenberg-dryer}
+	For reference, we also provide the numbers of the closest corresponding universals stated in Greenberg's original study.
+	}\label{table:greenberg-dryer}
 \end{table*}
@@
 
-\section{Posterior}
+\section{Comparing Efficiency to its Components}
 
-\begin{figure}
+\begin{figure}[ht]
 	\begin{center}
-	\includegraphics[width=0.48\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
+	\includegraphics[width=0.98\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
 	\end{center}
@@
 
-\section{Relation to the Information Bottleneck}
-
-Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
-\begin{equation}
-    \label{eq:efficiency-again}
-    J_T(L) = I[L(T); T] - \lambda H[L(T)].
-\end{equation}
-This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}.
-In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information.
-The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$.
-Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
-\begin{equation}
-    \label{eq:ib}
-    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
-\end{equation}
-In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
-
-In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$.
-To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$.
-We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized.
-Then our objective is parallel to Eq.~\ref{eq:ib}:
-\begin{equation}
-    \label{eq:efficiency-kinda-ib}
-    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
-\end{equation}
-Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$.
-Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
-\begin{equation*}
-    J_T(L) = I[U;T] - \lambda H[U].
-\end{equation*}
-
-Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
-
 \section{Possible values of $\lambda$}
@@
 In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
-We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations and relative optimality of language.
 
 The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
+We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations.
 We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (TODO check).
@@
 
-\begin{figure}
+\begin{figure}[ht]
     \centering
@@
 
+\section{Relation to the Information Bottleneck}
+
+Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
+\begin{equation}
+    \label{eq:efficiency-again}
+    J_T(L) = I[L(T); T] - \lambda H[L(T)].
+\end{equation}
+This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}.
+In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information.
+The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$.
+Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
+\begin{equation}
+    \label{eq:ib}
+    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
+\end{equation}
+In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
+
+In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$.
+To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$.
+We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized.
+Then our objective is parallel to Eq.~\ref{eq:ib}:
+\begin{equation}
+    \label{eq:efficiency-kinda-ib}
+    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
+\end{equation}
+Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$.
+Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
+\begin{equation*}
+    J_T(L) = I[U;T] - \lambda H[U].
+\end{equation*}
+
+Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
+
 \bibliographystyle{unsrtnat}

### 2019-05-05 second-draft/pnas-word-order-si.tex
si

@@
 \maketitle
+
+TODO
+
+- about the regression
+
+-- prior, how many samples, how many chains, convergence criteria
+
+-- robustness to priors
 
@@
 
+maybe per-language curves?
+
 \section{Results on all UD Relations}
-Table~\ref{tab:all-predictions} shows the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
-
+In this section, we provide the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
 We also report results for grammars optimized for predictability and parseability individually.
@@
 
+We considered all UD syntactic relations occurring in at least two of the 51 languages.
+In Table~\ref{tab:all-predictions-1}, we present the data for the eight correlations discussed in the main paper, and for those other relations for which the typological literature provides data.
+Additionally, in Table~\ref{tab:all-predictions-1} we present data for the other UD relations, for which either no typological data is available, or which are not linguistically meaningful.
+
+%posterior_perRelation_Real_conj.pdf         posterior_perRelation_Real_goeswith.pdf     posterior_perRelation_Real_nummod.pdf       
+%user@user-G751JY:~/grammar-optim/results/correlations$ evince figures/posteriors/posterior_perRelation_Real_csubj.pdf 
+%user@user-G751JY:~/grammar-optim/results/correlations$
+
 \begin{table*} % TODO figure out away to make this table more aesthetic
 	\begin{center}	
-
 \small{
-\begin{tabular}{|l|l|l|ll|l|l|}
+\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
 	\hline
-	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+&	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
 % From Dryer	
-acl  &  80   &   \textbf{85}$^{***}$   &   43   &   \textbf{82}$^{***}$   &   \textbf{71}$^{***}$  & $>50\%$  \citep{dryer1992greenbergian} \\
-aux  &  12   &   \textbf{26}$^{***}$   &   \textbf{19}$^{***}$   &   49   &   \textbf{35}$^{**}$  & $<50\%$ \citep{dryer1992greenbergian} \\
-lifted\_case  &  86   &   \textbf{81}$^{***}$   &   \textbf{40}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{67}$^{***}$ & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-lifted\_cop  &  94   &   \textbf{80}$^{***}$   &   59   &   \textbf{72}$^{***}$   &   \textbf{63}$^{**}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
-lifted\_mark  &  76   &   \textbf{85}$^{***}$   &   \textbf{57}$^{*}$   &   \textbf{76}$^{***}$   &   \textbf{72}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
-nmod  &  80   &   \textbf{82}$^{***}$   &   49   &   \textbf{71}$^{***}$   &   \textbf{68}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\ 
-obl  &  88   &   \textbf{78}$^{***}$   &   \textbf{71}$^{***}$   &   46   &   \textbf{68}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
-xcomp  &  88   &   \textbf{90}$^{***}$   &   \textbf{76}$^{***}$   &   \textbf{89}$^{***}$   &   \textbf{85}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
-\hdashline
-nsubj  &  33   &   \textbf{29}$^{***}$   &   53   &   \textbf{7}$^{***}$   &   \textbf{23}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+%\hdashline
 
@@
 % Other universals, not in Dryer
-advcl  &  86   &   \textbf{84}$^{***}$   &   51   &   \textbf{67}$^{***}$   &   \textbf{67}$^{*}$  & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
-ccomp  &  86   &   \textbf{88}$^{***}$   &   \textbf{70}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{77}$^{**}$ & $>50\%$ (cf.
+&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
+&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf.
 \cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
-csubj  &  79   &   \textbf{78}$^{***}$   &   60   &   \textbf{66}$^{***}$   &   \textbf{68}$^{**}$  & $>50\%$ (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
-amod  &  51   &   \textbf{84}$^{***}$   &   42   &   \textbf{53}$^{*}$   &   51   & $\approx 50\%$ \citep{dryer1992greenbergian} \\
-nummod  &  37   &   \textbf{74}$^{**}$   &   49   &   52   &   52 & $\approx 50\%$ \citep[][89A, 83A]{wals} \\ %
+&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf.
+\cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
+&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section (REF) \\
+&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
+&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+\end{tabular}
+}
 
+\end{center}
+\caption{Predictions on UD relations with predictions from the typological literature.
+The first section contains the eight correlations discussed in the main paper (See Section REF); the second section provides other relations for which predictions are available.
+In the last column, we indicate what direction would be expected typologically.}
+\label{tab:all-predictions-1}
+\end{table*}
+
+\begin{table*} % TODO figure out away to make this table more aesthetic
+	\begin{center}	
+\small{
+\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
 	\hline
-% Typological data not available
-appos  &  77   &   \textbf{77}$^{***}$   &   49   &   \textbf{67}$^{***}$   &   \textbf{62}$^{**}$ &  Unknown \\%($>50\%$?
-correlates with \textit{nmod} according to Greenberg U 23)  \\ % \cite{cinque2009greenberg} claims it correlates with OV/VO.
-However, it's not clear to me how closely appos matches the subject of the typological claim
-lifted\_cc  &  71   &   \textbf{83}$^{***}$   &   51   &   \textbf{81}$^{***}$   &   \textbf{71}$^{***}$  & Unknown \\
-expl  &  12   &   36   &   62   &   40   &   53  & Unknown \\
-iobj  &  74   &   49   &   \textbf{78}$^{***}$   &   \textbf{38}$^{*}$   &   62  & Unknown \\
-vocative  &  46   &   \textbf{42}$^{*}$   &   59   &   45   &   \textbf{45}$^{*}$  & Unknown \\
+	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+\hline
+appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\
+lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\
+expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\
+iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\
+vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\
 
@@
 % Uninterpretable
-compound  &  66   &   \textbf{61}$^{*}$   &   48   &   \textbf{58}$^{*}$   &   52  & Uninterpretable \\
-det  &  31   &   \textbf{71}$^{***}$   &   \textbf{56}$^{*}$   &   48   &   54  & Uninterpretable \\%$.50\%$ for articles, $\approx 50\%$ for others, \citep{dryer1992greenbergian} \\
-dislocated  &  47   &   61   &   \textbf{68}$^{**}$   &   55   &   58  & Uninterpretable \\
-dep  &  56   &   \textbf{64}$^{*}$   &   52   &   58   &   \textbf{58}$^{*}$   & Uninterpretable \\
-advmod  &  35   &   51   &   \textbf{18}$^{***}$   &   47   &   \textbf{39}$^{**}$  & Uninterpretable \\
+compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\
+det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\
+dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\
+dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
+advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\
 
@@
 % UD artifacts
-conj  &  72   &   \textbf{82}$^{***}$   &   \textbf{42}$^{**}$   &   \textbf{62}$^{***}$   &   \textbf{59}$^{*}$  & UD artifact  \\
-discourse  &  24   &   \textbf{36}$^{**}$   &   \textbf{41}$^{*}$   &   49   &   46  & UD artifact   \\
-fixed  &  78   &   \textbf{32}$^{***}$   &   45   &   47   &   44  & UD artifact   \\
-flat  &  73   &   \textbf{66}$^{**}$   &   45   &   \textbf{59}$^{**}$   &   54  & UD artifact   \\
-goeswith  &  61   &   57   &   44   &   57   &   52  & UD artifact   \\
-list  &  94   &   59   &   50   &   54   &   52  & UD artifact  \\
-orphan  &  82   &   \textbf{70}$^{**}$   &   50   &   \textbf{55}$^{*}$   &   59  & UD artifact   \\
-parataxis  &  66   &   56   &   43   &   57   &   59  & UD artifact  \\
-reparandum  &  18   &   53   &   57   &   43   &   48  & UD artifact  \\
+conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
+discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\
+fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\
+flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
+goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
+list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
+orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\
+parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\
+reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\
 
@@
 \end{center}
-\caption{Predictions on all UD relations occurring in more than one language.
-Numbers are the fraction of languages where the direction of the given relation agrees in direction with the \emph{obj} relation.
-In the last column, we indicate what direction would be expected typologically.
-The first section correspond to universals from Dryer.
+\caption{Predictions on UD relations for which no predictions are available in the typological literature.
 ``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
 ``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
-\label{tab:all-predictions}
+\label{tab:all-predictions-2}
 \end{table*}
@@
 \end{figure}
-
-\begin{table*}
-	\begin{center}
-%		\setlength\tabcolsep{1.5pt}
-\begin{tabular}{c|ll|c|c|cc|ccccccccccc}
-	\hline
-	%&
-	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Efficiency} &  \multirow{2}{*}{Predictability} &  \multirow{2}{*}{Parseability} & \\ 
-	&	verb & object     & & &   \\ 
-	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
-				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_case.pdf}     }  & \\
-	&		\emph{to}            & \emph{a friend} &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     }
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_cop.pdf}     } 
-& \\
-	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } 
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_aux.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_aux.pdf}     } & \\
-	&	\emph{has}          & \emph{written}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } 
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_nmod.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_nmod.pdf}     } & \\
-	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }
-	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_cop.pdf}     }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_acl.pdf}     } & \\
-	&	\emph{books} & \emph{that you read}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_mark.pdf}     } & \\
-	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_obl.pdf}     } & \\
-	&	\emph{went} & \emph{to school}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_xcomp.pdf}     }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_xcomp.pdf}     } & \\
-	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-\end{tabular}
-	\end{center}
-\caption{OPTION 3: Efficiency optimization accurately predicts the Greenbergian correlations.
-Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
-Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
-%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%Not all correlations are satisfied by every natu
-}\label{table:corr-dryer}
-\end{table*}
 
@@
 
-\section{Joint optimization of grammar, parser, and language model parameters}
+%\section{Joint optimization of grammar, parser, and language model parameters}
+\section{Optimizing Grammars for Efficiency}
 
@@
 \end{document}
+
+\begin{table*} % TODO figure out away to make this table more aesthetic
+	\begin{center}	
+\small{
+\begin{tabular}{|l|l|l|ll|l|l|}
+	\hline
+	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+% From Dryer	
+acl  &  80   &   \textbf{85}$^{***}$   &   43   &   \textbf{82}$^{***}$   &   \textbf{71}$^{***}$  & $>50\%$  \citep{dryer1992greenbergian} \\
+aux  &  12   &   \textbf{26}$^{***}$   &   \textbf{19}$^{***}$   &   49   &   \textbf{35}$^{**}$  & $<50\%$ \citep{dryer1992greenbergian} \\
+lifted\_case  &  86   &   \textbf{81}$^{***}$   &   \textbf{40}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{67}$^{***}$ & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+lifted\_cop  &  94   &   \textbf{80}$^{***}$   &   59   &   \textbf{72}$^{***}$   &   \textbf{63}$^{**}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
+lifted\_mark  &  76   &   \textbf{85}$^{***}$   &   \textbf{57}$^{*}$   &   \textbf{76}$^{***}$   &   \textbf{72}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
+nmod  &  80   &   \textbf{82}$^{***}$   &   49   &   \textbf{71}$^{***}$   &   \textbf{68}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\ 
+obl  &  88   &   \textbf{78}$^{***}$   &   \textbf{71}$^{***}$   &   46   &   \textbf{68}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
+xcomp  &  88   &   \textbf{90}$^{***}$   &   \textbf{76}$^{***}$   &   \textbf{89}$^{***}$   &   \textbf{85}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
+\hdashline
+nsubj  &  33   &   \textbf{29}$^{***}$   &   53   &   \textbf{7}$^{***}$   &   \textbf{23}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
+
+\hline
+% Other universals, not in Dryer
+advcl  &  86   &   \textbf{84}$^{***}$   &   51   &   \textbf{67}$^{***}$   &   \textbf{67}$^{*}$  & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
+ccomp  &  86   &   \textbf{88}$^{***}$   &   \textbf{70}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{77}$^{**}$ & $>50\%$ (cf.
+\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
+csubj  &  79   &   \textbf{78}$^{***}$   &   60   &   \textbf{66}$^{***}$   &   \textbf{68}$^{**}$  & $>50\%$ (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
+amod  &  51   &   \textbf{84}$^{***}$   &   42   &   \textbf{53}$^{*}$   &   51   & $\approx 50\%$ \citep{dryer1992greenbergian} \\
+nummod  &  37   &   \textbf{74}$^{**}$   &   49   &   52   &   52 & $\approx 50\%$ \citep[][89A, 83A]{wals} \\ %
+
+\hline
+% Typological data not available
+appos  &  77   &   \textbf{77}$^{***}$   &   49   &   \textbf{67}$^{***}$   &   \textbf{62}$^{**}$ &  Unknown \\%($>50\%$?
+correlates with \textit{nmod} according to Greenberg U 23)  \\ % \cite{cinque2009greenberg} claims it correlates with OV/VO.
+However, it's not clear to me how closely appos matches the subject of the typological claim
+lifted\_cc  &  71   &   \textbf{83}$^{***}$   &   51   &   \textbf{81}$^{***}$   &   \textbf{71}$^{***}$  & Unknown \\
+expl  &  12   &   36   &   62   &   40   &   53  & Unknown \\
+iobj  &  74   &   49   &   \textbf{78}$^{***}$   &   \textbf{38}$^{*}$   &   62  & Unknown \\
+vocative  &  46   &   \textbf{42}$^{*}$   &   59   &   45   &   \textbf{45}$^{*}$  & Unknown \\
+
+\hline
+% Uninterpretable
+compound  &  66   &   \textbf{61}$^{*}$   &   48   &   \textbf{58}$^{*}$   &   52  & Uninterpretable \\
+det  &  31   &   \textbf{71}$^{***}$   &   \textbf{56}$^{*}$   &   48   &   54  & Uninterpretable \\%$.50\%$ for articles, $\approx 50\%$ for others, \citep{dryer1992greenbergian} \\
+dislocated  &  47   &   61   &   \textbf{68}$^{**}$   &   55   &   58  & Uninterpretable \\
+dep  &  56   &   \textbf{64}$^{*}$   &   52   &   58   &   \textbf{58}$^{*}$   & Uninterpretable \\
+advmod  &  35   &   51   &   \textbf{18}$^{***}$   &   47   &   \textbf{39}$^{**}$  & Uninterpretable \\
+
+\hline
+% UD artifacts
+conj  &  72   &   \textbf{82}$^{***}$   &   \textbf{42}$^{**}$   &   \textbf{62}$^{***}$   &   \textbf{59}$^{*}$  & UD artifact  \\
+discourse  &  24   &   \textbf{36}$^{**}$   &   \textbf{41}$^{*}$   &   49   &   46  & UD artifact   \\
+fixed  &  78   &   \textbf{32}$^{***}$   &   45   &   47   &   44  & UD artifact   \\
+flat  &  73   &   \textbf{66}$^{**}$   &   45   &   \textbf{59}$^{**}$   &   54  & UD artifact   \\
+goeswith  &  61   &   57   &   44   &   57   &   52  & UD artifact   \\
+list  &  94   &   59   &   50   &   54   &   52  & UD artifact  \\
+orphan  &  82   &   \textbf{70}$^{**}$   &   50   &   \textbf{55}$^{*}$   &   59  & UD artifact   \\
+parataxis  &  66   &   56   &   43   &   57   &   59  & UD artifact  \\
+reparandum  &  18   &   53   &   57   &   43   &   48  & UD artifact  \\
+
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+\end{tabular}
+}
+
+\end{center}
+\caption{Predictions on all UD relations occurring in more than one language.
+Numbers are the fraction of languages where the direction of the given relation agrees in direction with the \emph{obj} relation.
+In the last column, we indicate what direction would be expected typologically.
+The first section correspond to universals from Dryer.
+``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
+``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
+\label{tab:all-predictions}
+\end{table*}

### 2019-05-05 results/correlations/createSITable.py

@@ -1,0 +1,5 @@
+with open("../relations.tsv", "r") as inFile:
+   rels = [x.strip() for x in inFile.read().strip().split("\n")]
+for x in rels:
+   print(x+"  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_"+x+".pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_"+x+".pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_"+x+".pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_"+x+".pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_"+x+".pdf}   & ")
+

### 2019-05-05 results/correlations/visualize-posterior-2-large-perRelation.R

@@ -47,8 +47,8 @@ data = merge(data, D, by=c("FileName"))
 
 data = data %>% mutate(Dir = pmax(0, sign(DH_Weight)), DirObj = pmax(0, sign(DH_Weight_Obj))) %>% mutate(Agree = (Dir == DirObj))
 
-reverseOrder = c("aux")
-data[data$Dependency %in% reverseOrder,]$Agree = 1-data[data$Dependency %in% reverseOrder,]$Agree
+#reverseOrder = c("aux")
+#data[data$Dependency %in% reverseOrder,]$Agree = 1-data[data$Dependency %in% reverseOrder,]$Agree
 
 
 D = data %>% group_by(Language, Family, Dependency) %>% summarise(Agree = mean(Agree)) %>% group_by(Dependency) %>% summarise(Agree = mean(Agree)) %>% mutate(Type = "UD Corpora")

### 2019-05-09 second-draft/pnas-word-order-si.tex
revising

@@
 
-TODO
+\tableofcontents
 
-- about the regression
+\section{Study 2: Regression for Predicted Correlations}
 
--- prior, how many samples, how many chains, convergence criteria
+Following (CITE), we used as a very weakly informative prior a Student's $t$ prior with 3 degrees of freedom, mean 0, and scale $\sigma=10$ (i.e., the PDF $p$ is $\frac{1}{\sigma} p_3(x/\sigma)$).
+The covariance matrix is modeled using the LKJ prior, as described in (CITE brms Bayesian multilevel).
+We used four chains, with 5000 samples each (2500 warmup).
+The chains had converged ($\hat{R} < 1.01$).
 
--- robustness to priors
+To verify robustness to more informative priors, we conducted the analysis (REF) with a tighter weakly informative prior.
+Note that an intercept $\alpha \approx 2.19$ in the logistic model corresponds to a correlation that holds in 90\% of cases; we thus selected the prior $\mathcal{N}(0,2)$.
+We obtained a qualitatively equivalent result (posterior probability $0.932$ of predicting all eight correlations).
+We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating the model using \texttt{lme4} (CITE) and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
 
@@
 
-\section{Supporting Analyses for Study 1}
+\section{Study 1: Supporting Analyses for Study 1}
 
-\paragraph{Other tests, maybe som viz}
+- binomial test
 
-\paragraph{Pareto Curve}
+\paragraph{z-Scored Data (Figure (REF))}
+We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section REF), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
 
-It has to be convex, so don't need to do more.
+Within each language, we estimate the Pareto curve by connecting the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section REF), (2) parseability, and (3) predictability.
+We then estimated the overall average Pareto curve, as shown in Figure (REF), by averaging the z-transformed points (1), (2), (3).
 
-maybe per-language curves?
-
-\section{Results on all UD Relations}
+\section{Study 2: Results on all UD Relations}
 In this section, we provide the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
 We also report results for grammars optimized for predictability and parseability individually.
-
-Further, we report results for grammars optimized for dependency length minimization (DLM).
-DLM can be seen as a heuristic approximation to efficiency.
-We also found that DLM is predicted by efficiency optimization.
-Indeed, we show that optimizing for DLM makes predictions similar to efficiency optimization.
 
@@
 Additionally, in Table~\ref{tab:all-predictions-1} we present data for the other UD relations, for which either no typological data is available, or which are not linguistically meaningful.
+
+Further, we report results for grammars optimized for dependency length minimization (DLM).
+Following (CITE), we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
+Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars (CITE).
+Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
+In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
 
@@
 
-\section{Comparing Efficiency to its Components}
+\section{Study 2: Comparing Efficiency to its Components}\label{sec:posterior-number}
+
+In Figure~\ref{fig:posterior}, we plot the posterior number of correlations predicted to hold in most optimized grammars, as obtained from the Bayesian regression.
+In addition to grammars optimized for efficiency, we also report the result for grammars optimized for predictability and for parseability alone.
+Efficiency predicts all eight correlations with high posterior probability; predictability and parseability alone do not.
 
@@
 
-\section{Possible values of $\lambda$}
+\section{Possible values of $\lambda$}\label{sec:lambda}
 
@@
 
-\subsection{Differentiable Ordering Grammars}
+In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
+Our main methodological innovation is to carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
+More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
+This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.
 
-In order to carry out optimization of grammars, we extend the space of grammars to include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
-This makes efficiency a differentiable function of the grammar parameters, and enables optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.
+This method addresses a major challenge noted in previous work optimizing grammars, namely that the predictability (and parseability) of an individual sentence depends on the entire distribution of the language.
+Previously, \citep{gildea2015human} optimized grammars for dependency length and trigram surprisal using a simple hill-climbing method on the grammar parameters, which required reestimating the trigram surprisal model in every iteration.
+Such a method would be computationally prohibitive for efficiency optimization, as it would require reestimating the neural network models after every change to the grammar, which would amount to reestimating them hundreds or thousands of times per grammar.
+Our method, by allowing for the use of stochastic gradient descent, addresses this challenge, as we describe in Section~\ref{sec:optim-eff}.
+
+\subsection{Differentiable Ordering Grammars}\label{sec:diff-gramm}
 
@@
 
-\textbf{rewrite this}
+A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
+More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ is a \emph{probability distribution} over ordered sequences of words $\mathbf{w}$.
+In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $\mathcal{L}_\theta(t)$ concentrates on a single ordering.
 
-For each of our objective functions, we seek parameters $\theta$ of a word order grammar $L_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $L_\theta$.
+Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
+Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
+\begin{equation}
+\label{eq:exp-over-utts}
+J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
+\end{equation}
+where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
+In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
+Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
+
+We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
+\begin{equation*}
+R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
+\end{equation*}
+where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
+\begin{equation*}
+	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+\end{equation*}
+
+We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
+Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
+\begin{align}
+\nonumber
+R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
+\nonumber
+&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
+\end{align}
+where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
+\begin{equation*}
+\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+\end{equation*}
+
+For efficiency, we linearly combine the two terms:
+\begin{equation}
+\label{eq:combined-score}
+R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
+\end{equation}
+
+%\begin{equation}
+%    \label{eq:depl}
+%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+%\end{equation}
+%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost.
+
+For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
 The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
@@
 \begin{align}
-\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
+%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
 \max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
@@
 
-The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
-A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
-We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
-
-The case of predictability, parseability, and efficiency are more challenging, as we need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
+%The case of predictability, parseability, and efficiency are more challenging, as 
+%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
 We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
 Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
-In each step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
+
+We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section (REF)).
+
+In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
 Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
@@
 
-In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
+See Section (REF) for the stopping criterion.
+
+%The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
+A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
+%We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
+
+%In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
 %and $\ell_2$ regularization.
 
-A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
-Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
-We thus do not evaluate this objective function here.
+%A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
+%Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
+%We thus do not evaluate this objective function here.
 
@@
 
-\textbf{TODO study 1 only has 1 set of parameters}
+%\textbf{TODO study 1 only has 1 set of parameters}
 
@@
 
+\paragraph{Optimized Grammars}
+
 Recall that, for each language, we created 8 optimized languages for each optimization criterion.
@@
 This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.
+
+For efficiency optimization, for each grammar, we ran efficiency optimization with three different random seeds, selecting among these the seed that yielded the best overall efficiency value.
+This
 
@@
 
-Here we address this question by demonstrating that our languages optimized for prediction and parseability are also optimal under a range of different language models and parsers.
-These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on a particular prediction and parsing algorithms.
+Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers.
+These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.
 
@@
 
-Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization.
+In this section, we address the idea that the graph-based parser might have a built-in bias toward certain kinds of orderings, and the question whether this might be responsible for our findings.
+In particular, we address the idea that the graph-based parser might have a bias toward parses involving short dependencies, which we call a \key{locality bias}. 
+We address this by changing the order in which the parser sees words, in such a way that the distance between words in the input is not indicative of syntactic relations.
+
+%In particular, this allows us to show tha
+
+%Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization.
 Therefore, we address here the idea that the graph-based parser might have a built-in bias toward parses involving short dependencies, which we call a \key{locality bias}.
-
-A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words.
+%A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words.
 Suppose that a parser $P$ has a bias toward positing a dependency between a word and the immediate previous word.
@@
 
+The result is that distance between words in the input is not indicative at all of the presence of absence of syntactic relations between them.
+
 \paragraph{Experiments}
@@
 
-We constructed a Bigram model with Kneser-Ney smoothing.
+We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
+In previous work, such as \cite{gildea2015human}, predictability has often been measured using n-gram models.
+
+Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
+Specifically, we constructed bigram models with Kneser-Ney smoothing (CITE).
 A bigram model predicts each word taking only the previous word into account.
@@
 
-\subsection{Discussion}
-
-It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.
+%\subsection{Discussion}
+%
+%It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.
 
@@
 
-\textbf{do same for language model, and for the other random grammars}
+\textcolor{red}{do same for language model, and for the other random grammars}
 
@@
     \includegraphics[scale=.4]{figures/learning-parser-czech.pdf} 
-        \caption{Parsing accuracy (measured in UAS) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
+        \caption{Parsing accuracy (measured in UAS) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.\textcolor{red}{font size}}
     \label{fig:learning-czech}

### 2019-05-09 results/correlations/joint-evaluation-efficiency09-large-prior.R

@@ -1,0 +1,61 @@
+data = read.csv("../../grammars/manual_output_funchead_two_coarse_lambda09_best_large/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+best = read.csv("../../grammars/manual_output_funchead_two_coarse_lambda09_best_large/successful-seeds.tsv")
+
+library(forcats)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+data = data %>% mutate(Language = fct_recode(Language, "Ancient_Greek" = "Ancient", "Old_Church_Slavonic" = "Old"))
+
+data = merge(data %>% mutate(FileName = as.character(FileName)), best %>% rename(FileName = Model), by=c("Language", "FileName"))
+
+languages = read.csv("../languages/languages-iso_codes.tsv")
+data  = merge(data, languages, by=c("Language"), all.x=TRUE)
+library("brms")
+dependency = "nmod"
+dependencies = c("acl", "advmod", "aux", "lifted_case", "lifted_cop", "lifted_mark", "nmod", "nsubj", "obl", "xcomp") # "nsubj",  "advmod",
+objs = data %>% filter(CoarseDependency == "obj") %>% mutate(obj = pmax(0, sign(DH_Weight))) %>% select(obj, Language, FileName)
+data = data %>% filter((CoarseDependency %in% dependencies))
+data = merge(data, objs, by=c("Language", "FileName"))
+data = data %>% mutate(dir = pmax(0, sign(DH_Weight)))
+data = data %>% mutate(dir = ifelse(DH_Weight == 0, NA, dir))
+data$agree = (data$dir == data$obj)
+data = unique(data %>% select(Family, Language, FileName, CoarseDependency, agree)) %>% spread(CoarseDependency, agree)
+
+
+get_prior(acl ~ (1|p|Family) + (1|q|Language), family="bernoulli", data=data)
+
+model3 = brm(acl ~ (1|p|Family) + (1|q|Language), family="bernoulli", data=data, iter=5000)
+model3 = brm(acl ~ (1|p|Family) + (1|q|Language), family="bernoulli", data=data, iter=5000, prior=prior(normal(0,2), class=Intercept))
+
+
+# Get the default priors
[…]

### 2019-05-10 second-draft/pnas-word-order-si.tex
SI

@@
 \DeclareMathOperator{\E}{\mathop{\mathbb{E}}}
+
+\usepackage[cm]{fullpage}
 
@@
 
+\section{Study 2: Results for Individual Languages}
+
+\begin{figure} 
+	\begin{center}
+	\includegraphics[width=0.32\textwidth]{../results/correlations/figures/pred-eff-families-2.pdf}
+	\end{center}
+	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
+Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
+Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...) \textcolor{red}{add other functions to this plot}.}\label{fig:per-lang}
+\end{figure}
+
 \section{Study 2: Regression for Predicted Correlations}
 
-Following (CITE), we used as a very weakly informative prior a Student's $t$ prior with 3 degrees of freedom, mean 0, and scale $\sigma=10$ (i.e., the PDF $p$ is $\frac{1}{\sigma} p_3(x/\sigma)$).
-The covariance matrix is modeled using the LKJ prior, as described in (CITE brms Bayesian multilevel).
+%https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
+
+Following the recommendations of \cite{ghosh2018use, burkner2018advanced}, we used as a very weakly informative prior a Student's $t$ prior with $\nu=3$ degrees of freedom, mean 0, and scale $\sigma=10$ (i.e., the PDF $p$ is $\frac{1}{\sigma} p_3(x/\sigma)$, where $p_3$ is the PDF of the $t$-distribution with $\nu=3$).
+Note that a correlation that holds in 90\% of cases corresponds to an intercept $\alpha \approx 2.19$ in the logistic model, well within the main probability mass of the prior.
+The covariance matrix is modeled using the LKJ prior, as described in \cite{burkner2018advanced}.
 We used four chains, with 5000 samples each (2500 warmup).
@@
 
-To verify robustness to more informative priors, we conducted the analysis (REF) with a tighter weakly informative prior.
-Note that an intercept $\alpha \approx 2.19$ in the logistic model corresponds to a correlation that holds in 90\% of cases; we thus selected the prior $\mathcal{N}(0,2)$.
-We obtained a qualitatively equivalent result (posterior probability $0.932$ of predicting all eight correlations).
-We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating the model using \texttt{lme4} (CITE) and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
+%To verify robustness to more informative priors, we conducted the analysis (REF) with a much tighter prior $\mathcal{N}(0,1)$.
+
+%Note that an intercept $\alpha \approx 2.19$ in the logistic model corresponds to a correlation that holds in 90\% of cases; we thus selected the prior $\mathcal{N}(0,2)$.
+%we obtained a qualitatively equivalent result (posterior probability $0.932$ of predicting all eight correlations).
+% N(0,2): 0.9863
+We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
 
@@
 Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{lifted\_mark} relation.
-Third, we both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.
+Third, both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.
 
+The resulting operationalization is shown in Table~\ref{table:greenberg-dryer}.
+
+\begin{table*}[ht]
+	\begin{center}
+\small{
+\begin{tabular}{|l|ll|l|l|}
+	\hline
+&	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{UD Relation}   & \multirow{2}{*}{\cite{greenberg1963universals}}    \\ 
+&	verb & object & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%adp. &
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&adposition    &    NP    &    case   & 3, 4   \\ \hline
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&copula  verb  &    predicate    &    cop   & --    \\\hline
+\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}&tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{aux}   & 16, 13  \\
+&	negative auxiliary    &    VP    &    & -- \\ \hline
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&noun    &    genitive    &    nmod & 2, 23   \\ \hline
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&noun    &    relative clause    &    acl   &  24    \\ \hline
+\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}&complementizer    &    S    &   \multirow{2}{*}{mark}  & --    \\
+&	adverbial subordinator & S &  & -- \\ \hline
+\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}&	adjective & std. of comp. & \multirow{2}{*}{obl} & --\\
+&verb    &    PP    &    & 22   \\\hline
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&`want'    &    VP    &    xcomp   & 15   \\\hline
+ \hline
+\end{tabular}
+}
+	\end{center}
+	\caption{Greenbergian Correlations based on Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies.
+	For reference, we also provide the numbers of the closest corresponding universals stated in Greenberg's original study (if possible).
+	}\label{table:greenberg-dryer}
+\end{table*}
+
+\paragraph{Excluded Correlations}
 We excluded three correlations that are not annotated reliably in UD, and are only relevant to some of the world's languages: Question particles, plural words (i.e., independent plural markers), and articles.
@@
 
-The resulting operationalization is shown in Table~\ref{table:greenberg-dryer}.
-
-\begin{table*}[ht]
-	\begin{center}
-\begin{tabular}{|ll|l|l|}
-	\hline
-	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{UD Relation}   & \multirow{2}{*}{\cite{greenberg1963universals}}    \\ 
-	verb & object &    \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%adp. &
-adposition    &    NP    &    case   & 3, 4   \\ \hline
-copula  verb  &    predicate    &    cop   & --    \\\hline
-`want'    &    VP    &    xcomp   & 15   \\\hline
-tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{aux}   & 16, 13  \\
-	negative auxiliary    &    VP    &    & -- \\ \hline
-complementizer    &    S    &   \multirow{2}{*}{mark}  & --    \\
-	adverbial subordinator & S &  & -- \\ \hline
-noun    &    genitive    &    nmod & 2, 23   \\ \hline
-noun    &    relative clause    &    acl   &  24    \\ \hline
-	adjective & std. of comp. & \multirow{2}{*}{obl} & --\\
-verb    &    PP    &    & 22   \\
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-    \multicolumn{3}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
-\end{tabular}
-	\end{center}
-	\caption{Greenbergian Correlations according to Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies.
-	For reference, we also provide the numbers of the closest corresponding universals stated in Greenberg's original study.
-	}\label{table:greenberg-dryer}
-\end{table*}
-
 %croft2011greenbergian
@@
 
-- binomial test
+\paragraph{Relative Efficiency of Real Grammars}
+In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
+We also conducted the analysis using a Binomial test (one-sided), testing whether the real gramar is more efficient than the \emph{median} of random grammars, avoiding any distributional assumption on the random grammars.
+As before, we used Hochberg's step-up procedure (Note that the tests for different languages are independent, as different random grammars are evaluated for each language), with $\alpha=0.05$.
+In this analysis, real grammars improved in parseability for 80\% of languages, in predictability for 69\% of languages, and in either of both in 92\% of languages ($p<0.05$, with Bonferroni correction).
 
@@
 
+In Figure~\ref{fig:corr-eff-corr}, we plot efficiency, parseability, and predictability (parseability and predictability $z$-scored within language) as a function of the number of satisfied correlations, for the real grammars of the 51 languages.
+
+We found very similar results using Spearman's rank correlation (Efficiency: $\rho=0.59$, $p = 9.8 \cdot 10^{-6}$; Parseability: $\rho=0.55$, $p=4.7 \cdot 10^{-5}$; Predictability: $\rho=0.36$, $p=0.012$).
+
 \begin{figure}[ht]
@@
 	\caption{Correlation between the number of satisfied correlations ($x$-axis) and efficiency, parseability, and predictability ($y$-axis), for the 51 real languages.}
-    \label{fig:todo}
+    \label{fig:corr-eff-corr}
 \end{figure}
@@
 
-Here, we validate our grammars by comparing the dominant orders of six syntactic relations that are also annotated in WALS.
+Here, we validate our grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, CITE).
 Among the eight relations that we were able to test, five are annotated in WALS.
@@
 WALS has some data for all but seven languages.\footnote{Note that Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
-In the table, we compare those with the grammar we extracted for Croatian, noting that it agrees with the Serbian grammar.}
+In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}
 WALS has data for 74\% of the entries, and lists a dominant order for 91\% of these.

### 2019-05-10 second-draft/pnas-word-order-si.tex
SI

@@
 
+We show results on the level of individual languages in Figure~\ref{fig:per-lang}.
+
 \begin{figure} 
 	\begin{center}
-	\includegraphics[width=0.32\textwidth]{../results/correlations/figures/pred-eff-families-2.pdf}
+	\includegraphics[width=0.7\textwidth]{../results/correlations/figures/pred-eff-pred-pars-families-2.pdf} % figures/pred-eff-families-2.pdf
 	\end{center}
-	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
+	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by optimizing for efficiency, predictability, parseability (right).
+Dark blue: Verb patterner \emph{precedes} object patterner (English, Arabic, ...).
+Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).
+White cells indicate that the relation is not annotated in the dataset for the given language.
+\textcolor{red}{say more about how these predictions were generated}}\label{fig:per-lang}
+%	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
 Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
@@
 
-\begin{figure}
-    \centering
-
-    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
-    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
-    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
-
-    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
-    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
-	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
-    \label{fig:posterior-prevalence}
-\end{figure}
+%\begin{figure}
+%    \centering
+%    
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
+%    
+%    
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
+%	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
+%    \label{fig:posterior-prevalence}
+%\end{figure}
 
@@
 `?' indicates that WALS has no data.
-$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
+$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur at similar rates in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
 Finally, `--' indicates that the relation does not occur in the corpus.

### 2019-05-10 second-draft/pnas-word-order-si.tex
si

@@
-\documentclass[12pt]{article}
+\documentclass[10pt,twoside,lineno]{article}
+% Use the lineno option to display guide line numbers if required.
+
+%\templatetype{pnassupportinginfo}
+
+% \readytosubmit %% Uncomment this line before submitting, so that the instruction page is removed.
+
 \usepackage{amsmath}
@@
 
+\textcolor{red}{TODOs}
+
+\textcolor{red}{This idea makes strong predictions, and turns out to overpredict correlations (see SI, Section X).}
+
 \section{Study 2: Results for Individual Languages}
@@
 We show results on the level of individual languages in Figure~\ref{fig:per-lang}.
+
+We obtained predictions for individual languages and each of the eight relations as follows.
+For each language and each of the objective functions (efficiency, predictability, parseability), we considered the optimized grammar that yielded the best value of this objective function among the eight optimized grammars (i.e., the grammar where the optimization procedure had been most successful).
+We interpreted this grammar as verb-object or object-verb depending on the order in the real grammar of the language.
 
@@
 Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).
-White cells indicate that the relation is not annotated in the dataset for the given language.
-\textcolor{red}{say more about how these predictions were generated}}\label{fig:per-lang}
+White cells indicate that the relation is not annotated in the dataset for the given language.}\label{fig:per-lang}
 %	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
@@
 First, tense/aspect and negative auxiliaries are together represented by the \emph{aux} relation in UD.
-Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{lifted\_mark} relation.
+Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{mark} relation.
 Third, both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.
@@
 The resulting operationalization is shown in Table~\ref{table:greenberg-dryer}.
+For each relation, we show the direction of the UD syntactic relation: $\rightarrow$ indicates that the verb patterner is the head; $\leftarrow$ indicates that the object patterner is the head.
+
+Note that we follow \cite{futrell2015largescale} in converting the Universal Dependencies format to a format closer to standard syntactic theory, promoting adpositions, copulas, and complementizers to heads.
+As a consequence, the direction of the relations \emph{case}, \emph{cop}, and \emph{mark} is reversed compared to Universal Dependencies.
+For clarity, we refer to these reversed relations as \emph{lifted\_case}, \emph{lifted\_cop}, and \emph{lifted\_mark}.
 
@@
 %adp. &
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&adposition    &    NP    &    case   & 3, 4   \\ \hline
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&copula  verb  &    predicate    &    cop   & --    \\\hline
-\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}&tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{aux}   & 16, 13  \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&adposition    &    NP    &  $\xrightarrow{lifted\_case}$   & 3, 4   \\ \hline
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&copula  verb  &    predicate    &    $\xrightarrow{lifted\_cop}$   & --    \\\hline
+\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}&tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{$\xleftarrow{aux}$}   & 16, 13  \\
 &	negative auxiliary    &    VP    &    & -- \\ \hline
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&noun    &    genitive    &    nmod & 2, 23   \\ \hline
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&noun    &    relative clause    &    acl   &  24    \\ \hline
-\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}&complementizer    &    S    &   \multirow{2}{*}{mark}  & --    \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&noun    &    genitive    &   $\xrightarrow{nmod}$ & 2, 23   \\ \hline
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&noun    &    relative clause    &    $\xrightarrow{acl}$ &  24    \\ \hline
+\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}&complementizer    &    S    &   \multirow{2}{*}{$\xrightarrow{lifted\_mark}$}  & --    \\
 &	adverbial subordinator & S &  & -- \\ \hline
-\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}&	adjective & std. of comp. & \multirow{2}{*}{obl} & --\\
+\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}&	adjective & std. of comp. & \multirow{2}{*}{$\xrightarrow{obl}$} & --\\
 &verb    &    PP    &    & 22   \\\hline
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&`want'    &    VP    &    xcomp   & 15   \\\hline
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&`want'    &    VP    &    $\xrightarrow{xcomp}$   & 15   \\\hline
  \hline
@@
 	\end{center}
-	\caption{Greenbergian Correlations based on Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies.
+	\caption{Greenbergian Correlations based on Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies using the modified format of \cite{futrell2015largescale} (see text).
 	For reference, we also provide the numbers of the closest corresponding universals stated in Greenberg's original study (if possible).
@@
 \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
 \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
@@
 \end{table*}
+
+\footnotetext{Note that we report an \emph{anti-correlation}.
+The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
+Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
+}
 

### 2019-05-10 second-draft/pnas-word-order-si.tex
SI

@@
 
-\usepackage{natbib}
+\usepackage[numbers]{natbib}
 
@@
 
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+\renewcommand{\thesection}{S\arabic{section}}
+
 \usepackage{mathtools}
@@
 }
-\date{\today}
+%\author{
+%Michael Hahn, Daniel Jurafsky, Richard Futrell}
+
+%\correspondingauthor{Michael Hahn.\\E-mail: mhahn2@stanford.edu}
 
@@
 
-\textcolor{red}{This idea makes strong predictions, and turns out to overpredict correlations (see SI, Section X).}
-
-\section{Study 2: Results for Individual Languages}
-
-We show results on the level of individual languages in Figure~\ref{fig:per-lang}.
-
-We obtained predictions for individual languages and each of the eight relations as follows.
-For each language and each of the objective functions (efficiency, predictability, parseability), we considered the optimized grammar that yielded the best value of this objective function among the eight optimized grammars (i.e., the grammar where the optimization procedure had been most successful).
-We interpreted this grammar as verb-object or object-verb depending on the order in the real grammar of the language.
-
-\begin{figure} 
-	\begin{center}
-	\includegraphics[width=0.7\textwidth]{../results/correlations/figures/pred-eff-pred-pars-families-2.pdf} % figures/pred-eff-families-2.pdf
-	\end{center}
-	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by optimizing for efficiency, predictability, parseability (right).
-Dark blue: Verb patterner \emph{precedes} object patterner (English, Arabic, ...).
-Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).
-White cells indicate that the relation is not annotated in the dataset for the given language.}\label{fig:per-lang}
-%	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
-Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
-Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...) \textcolor{red}{add other functions to this plot}.}\label{fig:per-lang}
-\end{figure}
-
-\section{Study 2: Regression for Predicted Correlations}
-
-%https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
-
-Following the recommendations of \cite{ghosh2018use, burkner2018advanced}, we used as a very weakly informative prior a Student's $t$ prior with $\nu=3$ degrees of freedom, mean 0, and scale $\sigma=10$ (i.e., the PDF $p$ is $\frac{1}{\sigma} p_3(x/\sigma)$, where $p_3$ is the PDF of the $t$-distribution with $\nu=3$).
-Note that a correlation that holds in 90\% of cases corresponds to an intercept $\alpha \approx 2.19$ in the logistic model, well within the main probability mass of the prior.
-The covariance matrix is modeled using the LKJ prior, as described in \cite{burkner2018advanced}.
-We used four chains, with 5000 samples each (2500 warmup).
-The chains had converged ($\hat{R} < 1.01$).
-
-%To verify robustness to more informative priors, we conducted the analysis (REF) with a much tighter prior $\mathcal{N}(0,1)$.
-
-%Note that an intercept $\alpha \approx 2.19$ in the logistic model corresponds to a correlation that holds in 90\% of cases; we thus selected the prior $\mathcal{N}(0,2)$.
-%we obtained a qualitatively equivalent result (posterior probability $0.932$ of predicting all eight correlations).
-% N(0,2): 0.9863
-We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
+\textcolor{red}{TODO This idea makes strong predictions, and turns out to overpredict correlations (see SI, Section X).}
 
@@
 
+\textcolor{red}{TODO At some point say something about \cite{dunn2011evolved}, since that got a lot of attention?
+can refer to follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
+
 Here we describe how we selected the correlations in Table 1 of the main paper, and how we formalized these using syntactic relations defined by Universal Dependencies.
 
-We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding recent objections by~\cite{dunn2011evolved}, see follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
+We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.
 Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
@@
 
-\section{Study 1: Supporting Analyses for Study 1}
+\section{Supplementary Analyses for Study 1}
 
@@
 
-\section{Study 2: Results on all UD Relations}
+\section{Supplementary Analyses for Study 2}
+\subsection{Correlation between Universals and Efficiency}
+
+In Figure~\ref{fig:corr-eff-corr}, we plot efficiency, parseability, and predictability (parseability and predictability $z$-scored within language) as a function of the number of satisfied correlations, for the real grammars of the 51 languages.
+
+We found very similar results using Spearman's rank correlation (Efficiency: $\rho=0.59$, $p = 9.8 \cdot 10^{-6}$; Parseability: $\rho=0.55$, $p=4.7 \cdot 10^{-5}$; Predictability: $\rho=0.36$, $p=0.012$).
+
+\begin{figure}[ht]
+    \centering
+
+    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
+    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
+    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
+
+	\caption{Correlation between the number of satisfied correlations ($x$-axis) and efficiency, parseability, and predictability ($y$-axis), for the 51 real languages.}
+    \label{fig:corr-eff-corr}
+\end{figure}
+
+\subsection{Predictions for Individual Languages}
+
+We show results on the level of individual languages in Figure~\ref{fig:per-lang}.
+
+We obtained predictions for individual languages and each of the eight relations as follows.
+For each language and each of the objective functions (efficiency, predictability, parseability), we considered the optimized grammar that yielded the best value of this objective function among the eight optimized grammars (i.e., the grammar where the optimization procedure had been most successful).
+We interpreted this grammar as verb-object or object-verb depending on the order in the real grammar of the language.
+
+\begin{figure} 
+	\begin{center}
+	\includegraphics[width=0.7\textwidth]{../results/correlations/figures/pred-eff-pred-pars-families-2.pdf} % figures/pred-eff-families-2.pdf
+	\end{center}
+	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by optimizing for efficiency, predictability, parseability (right).
+Dark blue: Verb patterner \emph{precedes} object patterner (English, Arabic, ...).
+Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).
+White cells indicate that the relation is not annotated in the dataset for the given language.}\label{fig:per-lang}
+%	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
+Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
+Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...) \textcolor{red}{add other functions to this plot}.}\label{fig:per-lang}
+\end{figure}
+
+\subsection{Regression for Predicted Correlations}
+
+%https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
+
+Following the recommendations of \cite{ghosh2018use, burkner2018advanced}, we used as a very weakly informative prior a Student's $t$ prior with $\nu=3$ degrees of freedom, mean 0, and scale $\sigma=10$ (i.e., the PDF $p$ is $\frac{1}{\sigma} p_3(x/\sigma)$, where $p_3$ is the PDF of the $t$-distribution with $\nu=3$).
+Note that a correlation that holds in 90\% of cases corresponds to an intercept $\alpha \approx 2.19$ in the logistic model, well within the main probability mass of the prior.
+The covariance matrix is modeled using the LKJ prior, as described in \cite{burkner2018advanced}.
+We used four chains, with 5000 samples each (2500 warmup).
+The chains had converged ($\hat{R} < 1.01$).
+
+%To verify robustness to more informative priors, we conducted the analysis (REF) with a much tighter prior $\mathcal{N}(0,1)$.
+
+%Note that an intercept $\alpha \approx 2.19$ in the logistic model corresponds to a correlation that holds in 90\% of cases; we thus selected the prior $\mathcal{N}(0,2)$.
+%we obtained a qualitatively equivalent result (posterior probability $0.932$ of predicting all eight correlations).
+% N(0,2): 0.9863
+We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
+
+\subsection{Comparing Efficiency to its Components}\label{sec:posterior-number}
+
+In Figure~\ref{fig:posterior}, we plot the posterior number of correlations predicted to hold in most optimized grammars, as obtained from the Bayesian regression.
+In addition to grammars optimized for efficiency, we also report the result for grammars optimized for predictability and for parseability alone.
+Efficiency predicts all eight correlations with high posterior probability; predictability and parseability alone do not.
+
+\begin{figure}[ht]
+	\begin{center}
+	\includegraphics[width=0.98\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
+	\end{center}
+	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
+\end{figure}
+
+%\begin{figure}
+%    \centering
+%    
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
+%    
+%    
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
+%	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
+%    \label{fig:posterior-prevalence}
+%\end{figure}
+
+\subsection{Results on all UD Relations}
 In this section, we provide the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
@@
 
-\section{Study 2: Comparing Efficiency to its Components}\label{sec:posterior-number}
-
-In Figure~\ref{fig:posterior}, we plot the posterior number of correlations predicted to hold in most optimized grammars, as obtained from the Bayesian regression.
-In addition to grammars optimized for efficiency, we also report the result for grammars optimized for predictability and for parseability alone.
-Efficiency predicts all eight correlations with high posterior probability; predictability and parseability alone do not.
-
-\begin{figure}[ht]
-	\begin{center}
-	\includegraphics[width=0.98\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
-	\end{center}
-	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
-\end{figure}
-
-%\begin{figure}
-%    \centering
-%    
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
-%    
-%    
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
-%	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
-%    \label{fig:posterior-prevalence}
-%\end{figure}
-
 \section{Possible values of $\lambda$}\label{sec:lambda}
@@
 %\section{Joint optimization of grammar, parser, and language model parameters}
-\section{Optimizing Grammars for Efficiency}
+\section{Creating Optimized Grammars}
 
 In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
-Our main methodological innovation is to carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
+We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
 More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
@@
 
-\textcolor{red}{do same for language model, and for the other random grammars}
-
 The resulting curves are shown in Figure~\ref{fig:learning-czech}.
-A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.4]{figures/learning-parser-czech.pdf} 
-        \caption{Parsing accuracy (measured in UAS) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.\textcolor{red}{font size}}
-    \label{fig:learning-czech}
-\end{figure}
-
-\section{Study 2, Correlation between Universals and Efficiency}
-
-In Figure~\ref{fig:corr-eff-corr}, we plot efficiency, parseability, and predictability (parseability and predictability $z$-scored within language) as a function of the number of satisfied correlations, for the real grammars of the 51 languages.
-
-We found very similar results using Spearman's rank correlation (Efficiency: $\rho=0.59$, $p = 9.8 \cdot 10^{-6}$; Parseability: $\rho=0.55$, $p=4.7 \cdot 10^{-5}$; Predictability: $\rho=0.36$, $p=0.012$).
+%A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
+A gap in parsing loss of about 0.2 nats appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
 
@@
     \centering
+%    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech.pdf} 
+    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech-logloss.pdf}
 
-    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
-    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
-    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
-
-	\caption{Correlation between the number of satisfied correlations ($x$-axis) and efficiency, parseability, and predictability ($y$-axis), for the 51 real languages.}
-    \label{fig:corr-eff-corr}
+        \caption{Parsing loss ($H[T|\mathcal{L}_\theta(T)]$) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
+    \label{fig:learning-czech}
 \end{figure}

### 2019-05-10 results/learning-curves/learningCurves.R

@@ -4,6 +4,10 @@ library(ggplot2)
 
 plot = ggplot(data=data, aes(x=Fraction, y=UAS, group=Model+RNN_Dim, color=Model, fill = Model)) + 
       geom_line() + theme_bw() + theme(legend.position="none")
+plot = plot + theme(axis.title.x = element_text(size=17))
+plot = plot + theme(axis.title.y = element_text(size=17))
+plot = plot + xlab("Fraction of Data")
+plot = plot + ylab("Parsing Accuracy")
 
 ggsave(plot=plot, filename="figures/learning-parser-czech.pdf") #, height=5, width=5)
 
@@ -11,6 +15,15 @@ plot = ggplot(data=data, aes(x=log(Fraction), y=UAS, group=Model, color=Model, f
       geom_line() + theme_bw()
 
 ggsave(plot=plot, filename="figures/learning-parser-czech-log.pdf") #, height=5, width=5)
+
+plot = ggplot(data=data, aes(x=Fraction, y=Loss, group=Model+RNN_Dim, color=Model, fill = Model)) + 
+      geom_line() + theme_bw() + theme(legend.position="none")
+plot = plot + theme(axis.title.x = element_text(size=17))
+plot = plot + theme(axis.title.y = element_text(size=17))
+plot = plot + xlab("Fraction of Data")
+plot = plot + ylab("Parsing Loss")
+
+ggsave(plot=plot, filename="figures/learning-parser-czech-logloss.pdf") #, height=5, width=5)
 
 
 

### 2019-05-11 second-draft/pnas-word-order-si.tex
SI

@@
 
-\section{Formalization of Correlations}
+\section{Formalization of Correlations}\label{sec:correlations}
 
@@
 	\hline
-&	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{UD Relation}   & \multirow{2}{*}{\cite{greenberg1963universals}}    \\ 
+&	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{UD Relation}   & \multirow{2}{*}{Greenberg \cite{greenberg1963universals}}    \\ 
 &	verb & object & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
@@
 \paragraph{z-Scored Data (Figure (REF))}
-We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section REF), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
+We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section~\ref{sec:lambda}), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
 
-Within each language, we estimate the Pareto curve by connecting the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section REF), (2) parseability, and (3) predictability.
+Within each language, we estimate the Pareto curve by connecting the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
 We then estimated the overall average Pareto curve, as shown in Figure (REF), by averaging the z-transformed points (1), (2), (3).
@@
 
-    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
-    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
-    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
+    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
+    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
+    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
 
@@
 
-%To verify robustness to more informative priors, we conducted the analysis (REF) with a much tighter prior $\mathcal{N}(0,1)$.
-
-%Note that an intercept $\alpha \approx 2.19$ in the logistic model corresponds to a correlation that holds in 90\% of cases; we thus selected the prior $\mathcal{N}(0,2)$.
-%we obtained a qualitatively equivalent result (posterior probability $0.932$ of predicting all eight correlations).
-% N(0,2): 0.9863
 We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
@@
 
+\textcolor{red}{at some place also say the following: Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
+}
+
 Further, we report results for grammars optimized for dependency length minimization (DLM).
-Following (CITE), we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
-Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars (CITE).
+Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
+Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
 Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
@@
 \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
-&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section (REF) \\
+&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section~\ref{sec:correlations} \\
 &amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
@@
 \caption{Predictions on UD relations with predictions from the typological literature.
-The first section contains the eight correlations discussed in the main paper (See Section REF); the second section provides other relations for which predictions are available.
+The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
 In the last column, we indicate what direction would be expected typologically.}
@@
 We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations.
-We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (TODO check).
+We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (Table~\ref{tab:all-predictions-1}).
 
@@
 
-The parameters of such a differentiable word order grammar are as follows. 
+We extended the parameter space of grammars by continuously interpolating between grammars, making efficiency a \emph{differentiable} function of grammar parameters.
+The parameters of such a \key{differentiable word order grammar} are as follows. 
 For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
@@
 
+In this section, we describe how we optimized grammar parameters for efficiency.
 A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
@@
 
-Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
-Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
+\textcolor{red}{TODO I'll clean this up}
+
+Recall the definition of efficiency
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+\end{equation}
+where 
 \begin{equation}
-\label{eq:exp-over-utts}
-J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
+	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
 \end{equation}
-where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
-In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
-Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
-
-We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
-\begin{equation*}
-R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
-\end{equation*}
-where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
-\begin{equation*}
-	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-\end{equation*}
-
-We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
-Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
-\begin{align}
-\nonumber
-R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
-\nonumber
-&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
-\end{align}
-where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
-\begin{equation*}
-\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-\end{equation*}
-
-For efficiency, we linearly combine the two terms:
 \begin{equation}
-\label{eq:combined-score}
-R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
+	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
 \end{equation}
 
+These quatities are estimated using two neural models:
+A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.
+A \key{language model} calculates the word-by-word surprisal of an utterance: $-\log p_\psi(u) = - \sum_{t=1}^{|u|} \log p_\psi(u_t|u_{1\dots t-1})$.
+The parsers and language models appropriate to a tree distribution $\mathcal{T}$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
+\begin{equation}
+	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+\end{equation}
+\begin{equation}
+\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+\end{equation}
+
+Using these models, we can rewrite 
+\begin{equation}
+	R_{Pars}^{\psi(\theta)}= \E_t \E_{u \sim p_\theta(u|t)} R_{Pars}^{\psi(\theta)}(u,t) 
+\end{equation}
+where $R_{Pars}(u,t) = \log \frac{p_\phi(t|u)}{p(t)}$, and $p_\theta(u|t)$ is the probability assigned to sentence $u$ by the distribution $\mathcal{L}_\theta(t)$ defined by the differentiable word order grammar $\mathcal{L}_\theta$,
+and
+\begin{equation}
+	R_{Pred} = \E_t \E_{u \sim p_\theta(u|t)} R_{Pred}(u)
+\end{equation}
+where $R_{Pred}(u) = \log p_\psi(u)$.
+Therefore, we can rewrite efficiency as
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}}^\theta := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi(\theta)}(u,t) + \lambda R_{Pred}^{\psi(\theta)}(u)\right]
+\end{equation}
+Directly applying gradient descent to this expression is hard due to the complicated dependence of $\phi(\theta), \psi(\theta)$ on $\theta$.
+
+However, we can rewrite
+\begin{equation}\label{eq:efficiency}
+\argmax_\theta\	R_{\textit{Eff}}^\theta 
+\end{equation}
+equivalently into
+\begin{equation}\label{eq:efficiency}
+\argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} = \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\end{equation}
+Importantly, $R_{\textit{Eff}}^{\theta, \phi, \psi}$ is differentiable in $\theta, \phi, \psi$: %, and we can apply stochastic gradient descent to carry out this optimization.
+
+\begin{align}
+\partial_\theta R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)} \left[  \left[\partial_\theta \log p_\theta(u|t)\right] \cdot    \left(R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right) \right] \label{eq:dtheta}\\
+\partial_\phi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\partial_\phi R_{Pars}^{\phi}(u,t)\right] \\
+\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}
+\end{align}
+where \ref{eq:dtheta} is derived using the \emph{score-function} or \emph{REINFORCE} trick~\cite{williams1992simple}.
+
+We can therefore apply stochastic gradient descent:
+In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim p_{\theta}(\cdot|t)$.
+Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
+do a gradient descent step using the estimator given by the expressions in the square brackets in \ref{eq:dtheta}-\ref{eq:dpsi}.
+
+\textcolor{red}{fix which one of psi, phi goes with pred, parse}
+
+Optimizing for only parseability (or predictability) is very similar -- in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.
+
+At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
+
+See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
+
+%
+%%(analogous for $R_{Pars}$ instead of $R_{Pred}$)
+%\begin{equation}\label{eq:estimator}
+%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
+%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
+%\end{equation}
+%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
+%for the gradient
+%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
+%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
+%
+
+%
+%
+%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
+Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
 %\begin{equation}
-%    \label{eq:depl}
-%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+%\label{eq:exp-over-utts}
+%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
 %\end{equation}
-%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost.
-
-For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
+%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
+In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
+%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
+%
+%
+%
+%We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
+%\begin{equation*}
+%R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
+%\end{equation*}
+%where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
+%\begin{equation*}
+%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation*}
+%
+%
+%We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
+Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
+%\begin{align}
+%\nonumber
+%R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
+%\nonumber
+%&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
+%\end{align}
+%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
+%\begin{equation*}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation*}
+%
+%For efficiency, we linearly combine the two terms:
+%\begin{equation}
+%\label{eq:combined-score}
+%R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
+%\end{equation}
+%
+%
+%%\begin{equation}
+%%    \label{eq:depl}
+%%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+%%\end{equation}
+%%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
+%
+%For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
 The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
 The optimization problems come out to:
-\begin{align}
-%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
-\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
-\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
-\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
-\end{align}
-Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
-
-%The case of predictability, parseability, and efficiency are more challenging, as 
-%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
-We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
-Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
-
-We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section (REF)).
-
-In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
-Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
-do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
-\begin{equation}\label{eq:estimator}
-{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
-%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
-\end{equation}
-%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
-for the gradient
-$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
-This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
-
-See Section (REF) for the stopping criterion.
+%\begin{align}
+%%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
+%\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
+%\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
+%\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
+%\end{align}
+%Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
+%
+%
+%%The case of predictability, parseability, and efficiency are more challenging, as 
+%%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
+%We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
+%Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
+%
+%We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section (REF)).
+%
+%In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
+%Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
+%do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
+%\begin{equation}\label{eq:estimator}
+%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
+%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
+%\end{equation}
+%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
+%for the gradient
+%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
+%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
+%
+%See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
+%
 
@@
 
-\section{Neural Network Architectures}
+\section{Neural Network Architectures}\label{sec:neural-architectures}
 
@@
 
-\textbf{TODO random seeds for efficiency}
+\textcolor{red}{TODO random seeds for efficiency}
 
@@
 
-We employ two common variance reduction methods to improve the estimator~(\ref{eq:estimator}), while keeping it unbiased.
+We employ two common variance reduction methods to improve the estimator~(\ref{eq:dtheta}), while keeping it unbiased.
 %For dependency length and 
 For predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word (and its head, in the case of dependency length).
-We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:estimator}) with lower variance.
+We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:dtheta}) with lower variance.
 Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple}.
@@
 
-These techniques for improving (\ref{eq:estimator}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
+These techniques for improving (\ref{eq:dtheta}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
 
@@
 Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
-Specifically, we constructed bigram models with Kneser-Ney smoothing (CITE).
+Specifically, we constructed bigram models with Kneser-Ney smoothing~\cite{kneser1995improved, chen1999empirical}.
 A bigram model predicts each word taking only the previous word into account.
@@
 
-Here, we validate our grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, CITE).
+Here, we validate our grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
 Among the eight relations that we were able to test, five are annotated in WALS.

### 2019-05-13 second-draft/pnas-word-order-si.tex
viz

@@
 A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
-More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ is a \emph{probability distribution} over ordered sequences of words $\mathbf{w}$.
-In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $\mathcal{L}_\theta(t)$ concentrates on a single ordering.
-
-\textcolor{red}{TODO I'll clean this up}
+More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $\mathbf{w}$.
+In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $p_{\mathcal{L}_\theta}(u|t)$ concentrates on a single ordering $u$.
 
@@
 \end{equation}
-
-These quatities are estimated using two neural models:
+where $t \sim \mathcal{T}$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
+
+These quatities are estimated using two neural models, as described in Section X:
 A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.
-A \key{language model} calculates the word-by-word surprisal of an utterance: $-\log p_\psi(u) = - \sum_{t=1}^{|u|} \log p_\psi(u_t|u_{1\dots t-1})$.
-The parsers and language models appropriate to a tree distribution $\mathcal{T}$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
+The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is
 \begin{equation}
-	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-\end{equation}
+	R_{Pars}^{\phi}(u,t) =  \log \frac{p_\phi(t|u)}{p(t)}
+\end{equation}
+%and we can write
+%\begin{equation}
+%	R_{Pars}^\phi := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
+%\end{equation}
+%
+A \key{language model}, with some parameters $\psi$, calculates the word-by-word surprisal of an utterance:
 \begin{equation}
-\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-\end{equation}
-
-Using these models, we can rewrite 
+	R_{Pred}^{\psi}(u) = - \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})
+\end{equation}
+Using this, we can rewrite Efficiency~(\ref{eq:efficiency}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
 \begin{equation}
-	R_{Pars}^{\psi(\theta)}= \E_t \E_{u \sim p_\theta(u|t)} R_{Pars}^{\psi(\theta)}(u,t) 
-\end{equation}
-where $R_{Pars}(u,t) = \log \frac{p_\phi(t|u)}{p(t)}$, and $p_\theta(u|t)$ is the probability assigned to sentence $u$ by the distribution $\mathcal{L}_\theta(t)$ defined by the differentiable word order grammar $\mathcal{L}_\theta$,
-and
-\begin{equation}
-	R_{Pred} = \E_t \E_{u \sim p_\theta(u|t)} R_{Pred}(u)
-\end{equation}
-where $R_{Pred}(u) = \log p_\psi(u)$.
-Therefore, we can rewrite efficiency as
+	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi} := \max_{\phi,\psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\end{equation}
+%
+%The parsers and language models appropriate to a tree distribution $\mathcal{T}$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
+%\begin{equation}
+%\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation}
+%\begin{equation}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation}
+%
+%Using these models, we can rewrite 
+%\begin{equation}
+%R_{Pars}^{\psi(\theta)}= \E_t \E_{u \sim p_\theta(u|t)} R_{Pars}^{\psi(\theta)}(u,t) 
+%\end{equation}
+%where $R_{Pars}(u,t) = \log \frac{p_\phi(t|u)}{p(t)}$, and $p_\theta(u|t)$ is the probability assigned to sentence $u$ by the distribution $\mathcal{L}_\theta(t)$ defined by the differentiable word order grammar $\mathcal{L}_\theta$,
+%and
+%\begin{equation}
+%R_{Pred} = \E_t \E_{u \sim p_\theta(u|t)} R_{Pred}(u)
+%\end{equation}
+%where $R_{Pred}(u) = \log p_\psi(u)$.
+%Therefore, we can rewrite efficiency as
+%\begin{equation}\label{eq:efficiency}
+%R_{\textit{Eff}}^\theta := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi(\theta)}(u,t) + \lambda R_{Pred}^{\psi(\theta)}(u)\right]
+%\end{equation}
+%Directly applying gradient descent to this expression is hard due to the complicated dependence of $\phi(\theta), \psi(\theta)$ on $\theta$.
+%
+%However, we can rewrite
+%\begin{equation}\label{eq:efficiency}
+%\argmax_\theta\	R_{\textit{Eff}}^\theta 
+%\end{equation}
+%equivalently into
+In order to find an optimal grammar $\theta$, we thus need to solve
 \begin{equation}\label{eq:efficiency}
-	R_{\textit{Eff}}^\theta := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi(\theta)}(u,t) + \lambda R_{Pred}^{\psi(\theta)}(u)\right]
-\end{equation}
-Directly applying gradient descent to this expression is hard due to the complicated dependence of $\phi(\theta), \psi(\theta)$ on $\theta$.
-
-However, we can rewrite
-\begin{equation}\label{eq:efficiency}
-\argmax_\theta\	R_{\textit{Eff}}^\theta 
-\end{equation}
-equivalently into
-\begin{equation}\label{eq:efficiency}
-\argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} = \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} %= \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
 \end{equation}
 Importantly, $R_{\textit{Eff}}^{\theta, \phi, \psi}$ is differentiable in $\theta, \phi, \psi$: %, and we can apply stochastic gradient descent to carry out this optimization.
-
 \begin{align}
@@
 where \ref{eq:dtheta} is derived using the \emph{score-function} or \emph{REINFORCE} trick~\cite{williams1992simple}.
-
-We can therefore apply stochastic gradient descent:
+Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures (CITE).
+
+We can therefore apply stochastic gradient descent to jointly optimize $\theta, \phi, \psi$:
 In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim p_{\theta}(\cdot|t)$.
@@
 do a gradient descent step using the estimator given by the expressions in the square brackets in \ref{eq:dtheta}-\ref{eq:dpsi}.
-
-\textcolor{red}{fix which one of psi, phi goes with pred, parse}
 
@@
 
-Here, we validate our grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
-Among the eight relations that we were able to test, five are annotated in WALS.
-Here, we compare our grammars with WALS on these five relations, and the verb-object relation.
+We validated our grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
+Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS.
+In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb-object relation.
 
@@
 
-For each of the six syntactic relation, the first column provides the ordered coded in the extracted grammar; the second column provides the order coded in WALS (DH for dependent-head, HD for head-dependent order).
-`?' indicates that WALS has no data.
-$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur at similar rates in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
-Finally, `--' indicates that the relation does not occur in the corpus.
-
+\begin{table}[ht]
 \small{
-\begin{longtable}{l||ll|ll|ll|ll|ll|ll|llllll}
+\begin{center}
+\begin{tabular}{l||ll|ll|ll|ll|ll|ll|llllll}
 %	   for(dep in c("obj", "lifted_mark", "acl", "nmod", "obl")) {
@@
 	\input{../results/grammars/comparison-table.tex}
-\end{longtable}
+\end{tabular}
+\end{center}
 }
+\caption{Comparing grammars extracted from databases to linguistic judgments in the World Atlas of Linguistic Structures.
+For each of the six syntactic relation, the first column provides the ordered coded in the extracted grammar; the second column provides the order coded in WALS (DH for dependent-head, HD for head-dependent order).
+`?' indicates that WALS has no data.
+$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur at similar rates in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
+Finally, `--' indicates that the relation does not occur in the corpus.
+}\label{tab:grammars-wals}
+\end{table}
 

### 2019-05-14 second-draft/pnas-word-order-si.tex
SI

--- 2019-05-10 second-draft/pnas-word-order-si.tex
@@
-\documentclass[10pt,twoside,lineno]{article}
-% Use the lineno option to display guide line numbers if required.
-
-%\templatetype{pnassupportinginfo}
-
-% \readytosubmit %% Uncomment this line before submitting, so that the instruction page is removed.
-
+\documentclass[12pt]{article}
 \usepackage{amsmath}
@@
 
-\usepackage[cm]{fullpage}
-
 \usepackage{longtable}
 
-\usepackage[numbers]{natbib}
+\usepackage{natbib}
 
@@
 
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thesection}{S\arabic{section}}
-
 \usepackage{mathtools}
@@
 }
-%\author{
-%Michael Hahn, Daniel Jurafsky, Richard Futrell}
-
-%\correspondingauthor{Michael Hahn.\\E-mail: mhahn2@stanford.edu}
+\date{\today}
 
@@
 
-\tableofcontents
-
-\textcolor{red}{TODOs}
-
-\textcolor{red}{TODO This idea makes strong predictions, and turns out to overpredict correlations (see SI, Section X).}
-
 \section{Formalization of Correlations}
-
-\textcolor{red}{TODO At some point say something about \cite{dunn2011evolved}, since that got a lot of attention?
-can refer to follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
 
@@
 
-We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.
+We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding recent objections by~\cite{dunn2011evolved}, see follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
 Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
@@
 First, tense/aspect and negative auxiliaries are together represented by the \emph{aux} relation in UD.
-Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{mark} relation.
-Third, both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.
+Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{lifted\_mark} relation.
+Third, we both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.
 
-The resulting operationalization is shown in Table~\ref{table:greenberg-dryer}.
-For each relation, we show the direction of the UD syntactic relation: $\rightarrow$ indicates that the verb patterner is the head; $\leftarrow$ indicates that the object patterner is the head.
-
-Note that we follow \cite{futrell2015largescale} in converting the Universal Dependencies format to a format closer to standard syntactic theory, promoting adpositions, copulas, and complementizers to heads.
-As a consequence, the direction of the relations \emph{case}, \emph{cop}, and \emph{mark} is reversed compared to Universal Dependencies.
-For clarity, we refer to these reversed relations as \emph{lifted\_case}, \emph{lifted\_cop}, and \emph{lifted\_mark}.
-
-\begin{table*}[ht]
-	\begin{center}
-\small{
-\begin{tabular}{|l|ll|l|l|}
-	\hline
-&	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{UD Relation}   & \multirow{2}{*}{\cite{greenberg1963universals}}    \\ 
-&	verb & object & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%adp. &
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&adposition    &    NP    &  $\xrightarrow{lifted\_case}$   & 3, 4   \\ \hline
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&copula  verb  &    predicate    &    $\xrightarrow{lifted\_cop}$   & --    \\\hline
-\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}&tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{$\xleftarrow{aux}$}   & 16, 13  \\
-&	negative auxiliary    &    VP    &    & -- \\ \hline
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&noun    &    genitive    &   $\xrightarrow{nmod}$ & 2, 23   \\ \hline
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&noun    &    relative clause    &    $\xrightarrow{acl}$ &  24    \\ \hline
-\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}&complementizer    &    S    &   \multirow{2}{*}{$\xrightarrow{lifted\_mark}$}  & --    \\
-&	adverbial subordinator & S &  & -- \\ \hline
-\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}&	adjective & std. of comp. & \multirow{2}{*}{$\xrightarrow{obl}$} & --\\
-&verb    &    PP    &    & 22   \\\hline
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&`want'    &    VP    &    $\xrightarrow{xcomp}$   & 15   \\\hline
- \hline
-\end{tabular}
-}
-	\end{center}
-	\caption{Greenbergian Correlations based on Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies using the modified format of \cite{futrell2015largescale} (see text).
-	For reference, we also provide the numbers of the closest corresponding universals stated in Greenberg's original study (if possible).
-	}\label{table:greenberg-dryer}
-\end{table*}
-
-\paragraph{Excluded Correlations}
 We excluded three correlations that are not annotated reliably in UD, and are only relevant to some of the world's languages: Question particles, plural words (i.e., independent plural markers), and articles.
@@
 
-%croft2011greenbergian
-%dryer2011evidence
+The resulting operationalization is shown in Table~\ref{table:greenberg-dryer}.
 
-\section{Supplementary Analyses for Study 1}
-
-\paragraph{Relative Efficiency of Real Grammars}
-In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
-We also conducted the analysis using a Binomial test (one-sided), testing whether the real gramar is more efficient than the \emph{median} of random grammars, avoiding any distributional assumption on the random grammars.
-As before, we used Hochberg's step-up procedure (Note that the tests for different languages are independent, as different random grammars are evaluated for each language), with $\alpha=0.05$.
-In this analysis, real grammars improved in parseability for 80\% of languages, in predictability for 69\% of languages, and in either of both in 92\% of languages ($p<0.05$, with Bonferroni correction).
-
-\paragraph{z-Scored Data (Figure (REF))}
-We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section REF), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
-
-Within each language, we estimate the Pareto curve by connecting the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section REF), (2) parseability, and (3) predictability.
-We then estimated the overall average Pareto curve, as shown in Figure (REF), by averaging the z-transformed points (1), (2), (3).
-
-\section{Supplementary Analyses for Study 2}
-\subsection{Correlation between Universals and Efficiency}
-
-In Figure~\ref{fig:corr-eff-corr}, we plot efficiency, parseability, and predictability (parseability and predictability $z$-scored within language) as a function of the number of satisfied correlations, for the real grammars of the 51 languages.
-
-We found very similar results using Spearman's rank correlation (Efficiency: $\rho=0.59$, $p = 9.8 \cdot 10^{-6}$; Parseability: $\rho=0.55$, $p=4.7 \cdot 10^{-5}$; Predictability: $\rho=0.36$, $p=0.012$).
-
-\begin{figure}[ht]
-    \centering
-
-    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
-    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
-    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
-
-	\caption{Correlation between the number of satisfied correlations ($x$-axis) and efficiency, parseability, and predictability ($y$-axis), for the 51 real languages.}
-    \label{fig:corr-eff-corr}
-\end{figure}
-
-\subsection{Predictions for Individual Languages}
-
-We show results on the level of individual languages in Figure~\ref{fig:per-lang}.
-
-We obtained predictions for individual languages and each of the eight relations as follows.
-For each language and each of the objective functions (efficiency, predictability, parseability), we considered the optimized grammar that yielded the best value of this objective function among the eight optimized grammars (i.e., the grammar where the optimization procedure had been most successful).
-We interpreted this grammar as verb-object or object-verb depending on the order in the real grammar of the language.
-
-\begin{figure} 
+\begin{table*}[ht]
 	\begin{center}
-	\includegraphics[width=0.7\textwidth]{../results/correlations/figures/pred-eff-pred-pars-families-2.pdf} % figures/pred-eff-families-2.pdf
-	\end{center}
-	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by optimizing for efficiency, predictability, parseability (right).
-Dark blue: Verb patterner \emph{precedes} object patterner (English, Arabic, ...).
-Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).
-White cells indicate that the relation is not annotated in the dataset for the given language.}\label{fig:per-lang}
-%	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
-Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
-Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...) \textcolor{red}{add other functions to this plot}.}\label{fig:per-lang}
-\end{figure}
-
-\subsection{Regression for Predicted Correlations}
-
-%https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
-
-Following the recommendations of \cite{ghosh2018use, burkner2018advanced}, we used as a very weakly informative prior a Student's $t$ prior with $\nu=3$ degrees of freedom, mean 0, and scale $\sigma=10$ (i.e., the PDF $p$ is $\frac{1}{\sigma} p_3(x/\sigma)$, where $p_3$ is the PDF of the $t$-distribution with $\nu=3$).
-Note that a correlation that holds in 90\% of cases corresponds to an intercept $\alpha \approx 2.19$ in the logistic model, well within the main probability mass of the prior.
-The covariance matrix is modeled using the LKJ prior, as described in \cite{burkner2018advanced}.
-We used four chains, with 5000 samples each (2500 warmup).
-The chains had converged ($\hat{R} < 1.01$).
-
-%To verify robustness to more informative priors, we conducted the analysis (REF) with a much tighter prior $\mathcal{N}(0,1)$.
-
-%Note that an intercept $\alpha \approx 2.19$ in the logistic model corresponds to a correlation that holds in 90\% of cases; we thus selected the prior $\mathcal{N}(0,2)$.
-%we obtained a qualitatively equivalent result (posterior probability $0.932$ of predicting all eight correlations).
-% N(0,2): 0.9863
-We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
-
-\subsection{Comparing Efficiency to its Components}\label{sec:posterior-number}
-
-In Figure~\ref{fig:posterior}, we plot the posterior number of correlations predicted to hold in most optimized grammars, as obtained from the Bayesian regression.
-In addition to grammars optimized for efficiency, we also report the result for grammars optimized for predictability and for parseability alone.
-Efficiency predicts all eight correlations with high posterior probability; predictability and parseability alone do not.
-
-\begin{figure}[ht]
-	\begin{center}
-	\includegraphics[width=0.98\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
-	\end{center}
-	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
-\end{figure}
-
-%\begin{figure}
-%    \centering
-%    
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
-%    
-%    
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
-%	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
-%    \label{fig:posterior-prevalence}
-%\end{figure}
-
-\subsection{Results on all UD Relations}
-In this section, we provide the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
-We also report results for grammars optimized for predictability and parseability individually.
-
-We considered all UD syntactic relations occurring in at least two of the 51 languages.
-In Table~\ref{tab:all-predictions-1}, we present the data for the eight correlations discussed in the main paper, and for those other relations for which the typological literature provides data.
-Additionally, in Table~\ref{tab:all-predictions-1} we present data for the other UD relations, for which either no typological data is available, or which are not linguistically meaningful.
-
-Further, we report results for grammars optimized for dependency length minimization (DLM).
-Following (CITE), we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
-Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars (CITE).
-Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
-In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
-
-%posterior_perRelation_Real_conj.pdf         posterior_perRelation_Real_goeswith.pdf     posterior_perRelation_Real_nummod.pdf       
-%user@user-G751JY:~/grammar-optim/results/correlations$ evince figures/posteriors/posterior_perRelation_Real_csubj.pdf 
-%user@user-G751JY:~/grammar-optim/results/correlations$
-
-\begin{table*} % TODO figure out away to make this table more aesthetic
-	\begin{center}	
-\small{
-\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
+\begin{tabular}{|ll|l|l|}
 	\hline
-&	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-% From Dryer	
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-%\hdashline
-
-\hline
-% Other universals, not in Dryer
-&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
-&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf.
-\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
-&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf.
-\cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
-&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section (REF) \\
-&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
-&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\
+	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{UD Relation}   & \multirow{2}{*}{\cite{greenberg1963universals}}    \\ 
+	verb & object &    \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%adp. &
+adposition    &    NP    &    case   & 3, 4   \\ \hline
+copula  verb  &    predicate    &    cop   & --    \\\hline
+`want'    &    VP    &    xcomp   & 15   \\\hline
+tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{aux}   & 16, 13  \\
+	negative auxiliary    &    VP    &    & -- \\ \hline
+complementizer    &    S    &   \multirow{2}{*}{mark}  & --    \\
+	adverbial subordinator & S &  & -- \\ \hline
+noun    &    genitive    &    nmod & 2, 23   \\ \hline
+noun    &    relative clause    &    acl   &  24    \\ \hline
+	adjective & std. of comp. & \multirow{2}{*}{obl} & --\\
+verb    &    PP    &    & 22   \\
  \hline
@@
 %	AP &
-    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+    \multicolumn{3}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
 \end{tabular}
-}
-
 	\end{center}
-\caption{Predictions on UD relations with predictions from the typological literature.
-The first section contains the eight correlations discussed in the main paper (See Section REF); the second section provides other relations for which predictions are available.
-In the last column, we indicate what direction would be expected typologically.}
-\label{tab:all-predictions-1}
+	\caption{Greenbergian Correlations according to Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies.
+	For reference, we also provide the numbers of the closest corresponding universals stated in Greenberg's original study.
+	}\label{table:greenberg-dryer}
 \end{table*}
 
-\footnotetext{Note that we report an \emph{anti-correlation}.
-The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
-Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
-}
+%croft2011greenbergian
+%dryer2011evidence
+
+\section{Supporting Analyses for Study 1}
+
+\paragraph{Other tests, maybe som viz}
+
+\paragraph{Pareto Curve}
+
+It has to be convex, so don't need to do more.
+
+\section{Results on all UD Relations}
+Table~\ref{tab:all-predictions} shows the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
+
+We also report results for grammars optimized for predictability and parseability individually.
+
+Further, we report results for grammars optimized for dependency length minimization (DLM).
+DLM can be seen as a heuristic approximation to efficiency.
+We also found that DLM is predicted by efficiency optimization.
+Indeed, we show that optimizing for DLM makes predictions similar to efficiency optimization.
 
@@
 	\begin{center}
-\small{
-\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
-	\hline
-	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-\hline
-appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\
-lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\
-expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\
-iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\
-vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\
 
-\hline
-% Uninterpretable
-compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\
-det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\
-dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\
-dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
-advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\
-
-\hline
-% UD artifacts
-conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
-discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\
-fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\
-flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
-goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
-list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
-orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\
-parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\
-reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\
-
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
-\end{tabular}
-}
-
-\end{center}
-\caption{Predictions on UD relations for which no predictions are available in the typological literature.
-``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
-``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
-\label{tab:all-predictions-2}
-\end{table*}
-
-\section{Possible values of $\lambda$}\label{sec:lambda}
-
-In the efficiency objective~(\ref{eq:efficiency-again}), the value of $\lambda$ is constrained to be in $[0,1)$.
-This means, surprisal must be weighted less strongly than parsability.
-
-Greater values of $\lambda$ mathematically result in degenerate solutions.
-To show this, note that the following inequality always holds:
-\begin{equation}
-I[L(T); T] \leq H[L(T)].
-\end{equation}
-Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
-\begin{equation}
-    J_T(L) = I[L(T); T] - \lambda H[L(T)] \leq 0.
-\end{equation}
-and it takes the maximally possible value zero if $L(T)$ maps all trees to a single utterance, in which case both $I[L(T); T]$ and $H[L(T)]$ are zero.
-This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
-While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
-
-In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
-
-The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
-We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations.
-We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (TODO check).
-
-%\section{Joint optimization of grammar, parser, and language model parameters}
-\section{Creating Optimized Grammars}
-
-In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
-We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
-More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
-This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.
-
-This method addresses a major challenge noted in previous work optimizing grammars, namely that the predictability (and parseability) of an individual sentence depends on the entire distribution of the language.
-Previously, \citep{gildea2015human} optimized grammars for dependency length and trigram surprisal using a simple hill-climbing method on the grammar parameters, which required reestimating the trigram surprisal model in every iteration.
-Such a method would be computationally prohibitive for efficiency optimization, as it would require reestimating the neural network models after every change to the grammar, which would amount to reestimating them hundreds or thousands of times per grammar.
-Our method, by allowing for the use of stochastic gradient descent, addresses this challenge, as we describe in Section~\ref{sec:optim-eff}.
-
-\subsection{Differentiable Ordering Grammars}\label{sec:diff-gramm}
-
-The parameters of such a differentiable word order grammar are as follows. 
-For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
-Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
-Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
-184) without replacement.
-
-If $a_\tau \in \{0, 1\}$, and the distances between values of $b_\tau$ (for different $\tau$) become very large, such a differentiable grammar becomes deterministic, assigning almost full probability to exactly one ordering for each syntactic structure.
-In this case, the grammar can be converted into an equivalent grammar of the form described in Materials and Methods, by extracting a single parameter in $[-1, 1]$ for each relation $\tau$.
-
-\subsection{Extracting Grammars from Datasets}
-We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
-To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
-We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.
-
-\subsection{Optimizing Grammars for Efficiency}\label{sec:optim-eff}
-
-A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
-More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ is a \emph{probability distribution} over ordered sequences of words $\mathbf{w}$.
-In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $\mathcal{L}_\theta(t)$ concentrates on a single ordering.
-
-Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
-Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
-\begin{equation}
-\label{eq:exp-over-utts}
-J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
-\end{equation}
-where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
-In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
-Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
-
-We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
-\begin{equation*}
-R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
-\end{equation*}
-where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
-\begin{equation*}
-	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-\end{equation*}
-
-We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
-Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
-\begin{align}
-\nonumber
-R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
-\nonumber
-&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
-\end{align}
-where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
-\begin{equation*}
-\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-\end{equation*}
-
-For efficiency, we linearly combine the two terms:
-\begin{equation}
-\label{eq:combined-score}
-R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
-\end{equation}
-
-%\begin{equation}
-%    \label{eq:depl}
-%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
-%\end{equation}
-%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost.
-
-For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
-The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
-The optimization problems come out to:
-\begin{align}
-%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
-\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
-\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
-\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
-\end{align}
-Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
-
-%The case of predictability, parseability, and efficiency are more challenging, as 
-%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
-We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
-Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
-
-We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section (REF)).
-
-In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
-Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
-do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
-\begin{equation}\label{eq:estimator}
-{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
-%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
-\end{equation}
-%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
-for the gradient
-$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
-This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
-
-See Section (REF) for the stopping criterion.
-
-%The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
-A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
-%We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
-
-%In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
-%and $\ell_2$ regularization.
-
-%A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
-%Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
-%We thus do not evaluate this objective function here.
-
-%We extract grammars for the actual languages using a Bayesian data analysis method.
-%We specify a simple Bayesian prior over differentiable ordering grammars:
-%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
-%for each $\tau$.
-%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
-%We extract
-
-\section{Neural Network Architectures}
-
-In this section, we describe the details of the neural network architectures.
-All choices described in this section were fixed before evaluating word order properties and the efficiency of real language.
-
-%\paragraph{Locality}
-%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
-
-\paragraph{Estimating Predictability}
-We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
-Given the small size of the corpora, this limit is only attained only for few languages.
-In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
-The model predicts both the next word and its POS tags in each step.
-Using POS tags is intended to prevent overfitting on small corpora.
-These decisions were made before evaluating word order properties.
-
-\paragraph{Estimating Parseability}
-We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}.
-This architecture is remarkably simple: the words of a sentence are encoded into context-sensitive embeddings using bidirectional LSTMs, then a classifier is trained to predict the head for each work.
-The classifier works by calculating a score for every pair of word embeddings $(w_i, w_j)$, indicating the likelihood that the $j$th word is the head of the $i$th word.
-This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
-
-To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags.
-Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged; parsing from POS tags prevents early overfitting.
-This decision was made before evaluating word order properties.
-
-\paragraph{Hyperparameters}
-
-%\textbf{TODO study 1 only has 1 set of parameters}
-
-Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
-For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
-Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (predictability or parseability) on the language.
-We then used this setting for creating optimized word order grammars.
-
-\textbf{TODO random seeds for efficiency}
-
-All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
-We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
-
-%\begin{table}[]
-%    \centering
-%    \begin{tabular}{c|cccc}
-%    Entropy penalty     &  0.0001, 0.001 \\
-%    Learning Rate (Word order grammar)     & 5e-6, 1e-5, 2e-5, 5e-5 \\
-%    Momentum (Word order grammar) & 0.8, 0,9 \\
-%    Dropout (Language model) & 0.0, 0.3, 0.5 \\
-%    
-%    \end{tabular}
-%    \caption{Hyperparameters}
-%    \label{tab:my_label}
-%\end{table}
-
-\paragraph{Optimization details}
-
-We employ two common variance reduction methods to improve the estimator~(\ref{eq:estimator}), while keeping it unbiased.
-%For dependency length and 
-For predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word (and its head, in the case of dependency length).
-We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:estimator}) with lower variance.
-Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple}.
-These two methods reduce the variance of the estimator and thereby increase the speed of optimization and reduce training time, without biasing the results.
-For numerical stability, we represent $a_\tau \in [0,1]$ via its logit $\in \mathbb{R}$.
-Furthermore, to encourage exploration of the parameter space, we add an entropy regularization term \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$.
-The weight of the entropy regularization was chosen together with the other hyperparameters.
-%, adding the negative entropy of the $Bernoulli(a_\tau)$ to the loss $R$. %, weighted with $\alpha$.
-
-These techniques for improving (\ref{eq:estimator}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
-
-We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
-For the language model parameters $\phi$, we use plain SGD. 
-For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
-The learning rates and other optimization hyperparameters were determined together with the other hyperparameters.
-
-For each language and objective function, we apply Early Stopping~\citep{prechelt1998early} as follows.
-We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
-This technique, well known in the machine learning literature, helps prevent overfitting to the specific corpora used.
-
-The choice of optimization methods and the stopping criterion were fixed before we investigated word order correlations.
-
-\paragraph{Optimized Grammars}
-
-Recall that, for each language, we created 8 optimized languages for each optimization criterion.
-We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
-This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.
-
-For efficiency optimization, for each grammar, we ran efficiency optimization with three different random seeds, selecting among these the seed that yielded the best overall efficiency value.
-This
-
- % TODO we could use some language at the top of this section on the general guiding principles for all these decisions.
-Were these implementation decisions made before or after looking at the word order universals results?
-If they were made before, we should highlight that.
-
-\section{Robustness to different language models and parsers}
-
-Here we take up the question of the extent to which our results are dependent on the particular parser and language model used in the optimization process.
-We want to know: when we optimize a word order grammar for efficiency, have we produced a language which is highly efficient \emph{in general}, or one which is highly efficient \emph{for a specific parser}?
-We wish to argue that natural language syntax is optimized for efficiency in general, meaning that syntactic trees are highly recoverable from word orders in principle.
-If it turns out that our optimized languages are only optimal for a certain parser from the NLP literature, then we run the risk of circularity: it may be that the reason this parser was successful in the NLP literature was because it implicitly encoded word order universals in its inductive biases, and thus it would be no surprise that languages which are optimized for parseability also show those universals.
-
-In this connection, we note that the parser and language model architectures we use are highly generic, and do not encode any obvious bias toward natural-language-like word orders.
-The LSTM language model is a generic model of sequence data which is also been used to model financial time series \citep{sirignano2018universal} and purely theoretical chaotic dynamical systems \citep{ogunmolu2016nonlinear}; the neural graph-based parser is simply solving a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
-Nevertheless, it may be the case that a bias toward word order universals is somehow encoded implicitly in the hyperparameters and architectures of these models.
-
-Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers.
-These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.
-
-\subsection{CKY Parsers}
-
-We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of Collins 1997 (Model 1).
-In our PCFGs, each head independently generates a set of left and right dependents.
-We formulate this as a PCFG where each rule has the form:
-\begin{center}
-	POS$_H$ $\rightarrow$ POS$_H$ POS$_D$
-\end{center}
-for head-initial structures, and
-\begin{center}
-	POS$_H$ $\rightarrow$ POS$_D$ POS$_H$
-\end{center}
-for head-final structures, where each symbol is a POS tag.
-Thus, POS tags act both as terminals and as nonterminals.
-
-We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
-For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm.
-
-This parsing strategy is very different from the neural graph-based parser:
-While the graph-based parser solves a minimum spanning tree problem, the CKY algorithm uses dynamic programming to compute the exact probabilities of trees given a sentence, as specified by the generative model encoded in the PCFG.
-Second, while the graph-based neural parser uses machine learning to induce syntactic knowledge from data, the CKY parser performs exact probabilistic inference.
-%In this sense, the CKY algorithm does not have any architectural biases in itself.
-%On the other hand, the PCFG makes severely simplifying independence assumptions, compared to the universal approximation capabilities of neural network-based systems.
-
-We used the CKY algorithm to compute the parsing loss $H[T|L_\theta(T)]$ on the validation partition of the English UD corpora, for random and optimized ordering grammars.
-Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
-% TODO get UAS for compatibility with other parts of SI
-%We then compared parsability of random and optimized word order grammars on English data.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
-	\caption{Parsability loss $H[T|L_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).}
-    \label{fig:cky-parser}
-\end{figure}
-
-\subsection{Distorted graph-based parsers}
-
-In this section, we address the idea that the graph-based parser might have a built-in bias toward certain kinds of orderings, and the question whether this might be responsible for our findings.
-In particular, we address the idea that the graph-based parser might have a bias toward parses involving short dependencies, which we call a \key{locality bias}. 
-We address this by changing the order in which the parser sees words, in such a way that the distance between words in the input is not indicative of syntactic relations.
-
-%In particular, this allows us to show tha
-
-%Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization.
-Therefore, we address here the idea that the graph-based parser might have a built-in bias toward parses involving short dependencies, which we call a \key{locality bias}.
-%A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words.
-Suppose that a parser $P$ has a bias toward positing a dependency between a word and the immediate previous word.
-
-\paragraph{Even--odd order.} A sequence of $n$ words originally ordered as $w_1 w_2 w_3 w_4 \cdots w_n$ is reordered by separating the even and odd indices: $w_2 w_4 w_6 \cdots w_{n-1} w_1 w_3 w_5 \cdots w_n$ (assuming $n$ odd).
-Therefore all words that are adjacent in the original order will be separated by a distance of $\approx n/2$ in the distorted order, while all words of distance 2 in the original order will become adjacent.
-
-\paragraph{Interleaving order.} In interleaving ordering, a sequence originally ordered as $w_1 w_2 w_3 \cdots w_n$ is split in half at the middle (index $m=\ceiling{n/2}$), and the two resulting sequences are interleaved, yielding $w_1 w_m w_2 w_{m+1} w_3 w_{m+3} \cdots w_n$.
-Thus all words that were originally adjacent will have distance 2 in the distorted order, with the intervening word coming from a very distant part of the sentence.
-
-\paragraph{Inwards order.} A sequence originally ordered as $w_1 w_2 w_3 \cdots w_{n-1} w_n$ is ordered from the edges of the string inwards, as $w_1 w_n w_2 w_{n-1} \cdots w_{\ceiling{n/2}}$.
-This corresponds to folding the string in on itself once, or equivalently, splitting the sequence in half at the middle, then interleaving the two resulting sequences after reversing the second one.
-The result is that the most non-local possible dependencies in the original order become the most local dependencies in the distorted order.
-
-\paragraph{Sorted order.} A sequence is reordered by sorting by POS tags, and randomizing the order within each block of identical POS tags.
-To each word, we then add a symbol encoding the original position in the sequence.
-For instance
-\begin{center}
-PRON VERB PRON
-\end{center}
-may be reordered as
-\begin{center}
-PRON 1 PRON 3 VERB 2
-\end{center}
-or
-\begin{center}
-PRON 3 PRON 1 VERB 2
-\end{center}
-The numbers are provided to the parser as atomic symbols from a vocabulary ranging from 1 to 200; numbers greater than 200 (which may occur in extremely long sentences) are replaced by an out-of-range token.
-
-The result is that distance between words in the input is not indicative at all of the presence of absence of syntactic relations between them.
-
-\paragraph{Experiments}
-Using English and Japanese data, we trained parsers for the ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
-Resulting parsing accuracy scores are shown in Figure~\ref{fig:distorted-parser}.
-In all settings, the language optimized for efficiency achieved higher parsing accuracy than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
-%Note that absolute values of parsing accuracy vary between the orderings, being higher for the actual orderings than for distorted orderings, in which trees become nonprojective.
-
-\begin{figure}
-    \centering
-    English
-
-    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-english.pdf}
-
-    Japanese
-
-    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
-	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').}
-    \label{fig:distorted-parser}
-\end{figure}
-
-\subsection{n-gram language models}
-
-We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
-In previous work, such as \cite{gildea2015human}, predictability has often been measured using n-gram models.
-
-Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
-Specifically, we constructed bigram models with Kneser-Ney smoothing (CITE).
-A bigram model predicts each word taking only the previous word into account.
-This contrasts with LSTMs, which take the entire context into consideration.
-Thus, bigram models and LSTMs stand on opposing ends of a spectrum of language models taking more and more aspects of the context into account.
-
-We estimated language models on the training partitions, and used the validation partitions to estimate surprisal.
-We conducted this for the ten random and the best optimized ordering grammars on English and Japanese data.
-Results (Figure~\ref{fig:bigrams}) show that languages optimized for efficiency are optimal for a bigram language model.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.4]{../results/bigrams/bigrams.pdf} 
-	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).}
-    \label{fig:bigrams}
-\end{figure}
-
-%\subsection{Discussion}
-%
-%It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.
-
-\section{Effects of data sparsity}
-
-Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
-If the difference between random and real grammars is due to data sparsity, we expect that it decreases as the amount of training data is increased.
-If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference persists as training data is increased.
-
-We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
-We cosidered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
-For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
-Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.
-
-The resulting curves are shown in Figure~\ref{fig:learning-czech}.
-%A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
-A gap in parsing loss of about 0.2 nats appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
-
-\begin{figure}[ht]
-    \centering
-%    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech.pdf} 
-    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech-logloss.pdf}
-
-        \caption{Parsing loss ($H[T|\mathcal{L}_\theta(T)]$) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
-    \label{fig:learning-czech}
-\end{figure}
-
-\section{Comparing Extracted Grammars to Linguistic Judgments}
-
-% user@user-X510UAR:~/grammar-optim/results/grammars$ git add comparison-table.tex
-
-Here, we validate our grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, CITE).
-Among the eight relations that we were able to test, five are annotated in WALS.
-Here, we compare our grammars with WALS on these five relations, and the verb-object relation.
-
-WALS has some data for all but seven languages.\footnote{Note that Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
-In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}
-WALS has data for 74\% of the entries, and lists a dominant order for 91\% of these.
-The grammars we extracted from the corpora agree with WALS in 96~\% out of those cases where WALS records a dominant word order.
-
-For each of the six syntactic relation, the first column provides the ordered coded in the extracted grammar; the second column provides the order coded in WALS (DH for dependent-head, HD for head-dependent order).
-`?' indicates that WALS has no data.
-$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur at similar rates in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
-Finally, `--' indicates that the relation does not occur in the corpus.
-
-\small{
-\begin{longtable}{l||ll|ll|ll|ll|ll|ll|llllll}
-%	   for(dep in c("obj", "lifted_mark", "acl", "nmod", "obl")) {
-		   Language 
-		   &	\multicolumn{2}{|c|}{Objects} 
-		   &	\multicolumn{2}{|c|}{Adpositions} 
-		   &	\multicolumn{2}{|c|}{Compl.} 
-		   &	\multicolumn{2}{|c|}{Rel.Cl.} 
-		   &	\multicolumn{2}{|c|}{Genitive} 
-		   &	\multicolumn{2}{|c|}{PP}  \\ \hline\hline
-	\input{../results/grammars/comparison-table.tex}
-\end{longtable}
-}
-
-\section{Relation to the Information Bottleneck}
-
-Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
-\begin{equation}
-    \label{eq:efficiency-again}
-    J_T(L) = I[L(T); T] - \lambda H[L(T)].
-\end{equation}
-This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}.
-In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information.
-The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$.
-Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
-\begin{equation}
-    \label{eq:ib}
-    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
-\end{equation}
-In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
-
-In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$.
-To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$.
-We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized.
-Then our objective is parallel to Eq.~\ref{eq:ib}:
-\begin{equation}
-    \label{eq:efficiency-kinda-ib}
-    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
-\end{equation}
-Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$.
-Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
-\begin{equation*}
-    J_T(L) = I[U;T] - \lambda H[U].
-\end{equation*}
-
-Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
-
-\bibliographystyle{unsrtnat}
-\bibliography{everything}
-
-\end{document}
-
-\begin{table*} % TODO figure out away to make this table more aesthetic
-	\begin{center}	
 \small{
@@
 \end{table*}
+
+\section{Comparing Efficiency to its Components}
+
+\begin{figure}[ht]
+	\begin{center}
+	\includegraphics[width=0.98\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
+	\end{center}
+	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
+\end{figure}
+
+\begin{table*}
+	\begin{center}
+%		\setlength\tabcolsep{1.5pt}
+\begin{tabular}{c|ll|c|c|cc|ccccccccccc}
+	\hline
+	%&
+	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Efficiency} &  \multirow{2}{*}{Predictability} &  \multirow{2}{*}{Parseability} & \\ 
+	&	verb & object     & & &   \\ 
+	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
+				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_case.pdf}     }  & \\
+	&		\emph{to}            & \emph{a friend} &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_case.pdf}     }
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_cop.pdf}     } 
+& \\
+	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_aux.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_aux.pdf}     } & \\
+	&	\emph{has}          & \emph{written}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_nmod.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_nmod.pdf}     } & \\
+	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }
+	&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_cop.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_acl.pdf}     } & \\
+	&	\emph{books} & \emph{that you read}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_lifted_mark.pdf}     } & \\
+	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_lifted_mark.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_obl.pdf}     } & \\
+	&	\emph{went} & \emph{to school}  &&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Predictability_xcomp.pdf}     }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Parseability_xcomp.pdf}     } & \\
+	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
+\end{tabular}
+	\end{center}
+\caption{OPTION 3: Efficiency optimization accurately predicts the Greenbergian correlations.
+Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
+For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
+Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
+Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
+%Given the statistical nature of the correlations, not every language satisfies every one of them.
+%Not all correlations are satisfied by every natu
+}\label{table:corr-dryer}
+\end{table*}
+
+\begin{figure}
+    \centering
+
+    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
+    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
+    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
+
+    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
+    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
+	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
+    \label{fig:posterior-prevalence}
+\end{figure}
+
+\section{Possible values of $\lambda$}
+
+In the efficiency objective~(\ref{eq:efficiency-again}), the value of $\lambda$ is constrained to be in $[0,1)$.
+This means, surprisal must be weighted less strongly than parsability.
+
+Greater values of $\lambda$ mathematically result in degenerate solutions.
+To show this, note that the following inequality always holds:
+\begin{equation}
+I[L(T); T] \leq H[L(T)].
+\end{equation}
+Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
+\begin{equation}
+    J_T(L) = I[L(T); T] - \lambda H[L(T)] \leq 0.
+\end{equation}
+and it takes the maximally possible value zero if $L(T)$ maps all trees to a single utterance, in which case both $I[L(T); T]$ and $H[L(T)]$ are zero.
+This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
+While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
+
+In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
+
+The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
+We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations.
+We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (TODO check).
+
+\section{Joint optimization of grammar, parser, and language model parameters}
+
+\subsection{Differentiable Ordering Grammars}
+
+In order to carry out optimization of grammars, we extend the space of grammars to include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
+This makes efficiency a differentiable function of the grammar parameters, and enables optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.
+
+The parameters of such a differentiable word order grammar are as follows. 
+For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
+Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
+Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
+184) without replacement.
+
+If $a_\tau \in \{0, 1\}$, and the distances between values of $b_\tau$ (for different $\tau$) become very large, such a differentiable grammar becomes deterministic, assigning almost full probability to exactly one ordering for each syntactic structure.
+In this case, the grammar can be converted into an equivalent grammar of the form described in Materials and Methods, by extracting a single parameter in $[-1, 1]$ for each relation $\tau$.
+
+\subsection{Extracting Grammars from Datasets}
+We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
+To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
+We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.
+
+\subsection{Optimizing Grammars for Efficiency}\label{sec:optim-eff}
+
+\textbf{rewrite this}
+
+For each of our objective functions, we seek parameters $\theta$ of a word order grammar $L_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $L_\theta$.
+The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
+The optimization problems come out to:
+\begin{align}
+\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
+\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
+\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
+\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
+\end{align}
+Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
+
+The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
+A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
+We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
+
+The case of predictability, parseability, and efficiency are more challenging, as we need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
+We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
+Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
+In each step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
+Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
+do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
+\begin{equation}\label{eq:estimator}
+{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
+%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
+\end{equation}
+%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
+for the gradient
+$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
+This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
+
+In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
+%and $\ell_2$ regularization.
+
+A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
+Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
+We thus do not evaluate this objective function here.
+
+%We extract grammars for the actual languages using a Bayesian data analysis method.
+%We specify a simple Bayesian prior over differentiable ordering grammars:
+%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
+%for each $\tau$.
+%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
+%We extract
+
+\section{Neural Network Architectures}
+
+In this section, we describe the details of the neural network architectures.
+All choices described in this section were fixed before evaluating word order properties and the efficiency of real language.
+
+%\paragraph{Locality}
+%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
+
+\paragraph{Estimating Predictability}
+We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
+Given the small size of the corpora, this limit is only attained only for few languages.
+In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
+The model predicts both the next word and its POS tags in each step.
+Using POS tags is intended to prevent overfitting on small corpora.
+These decisions were made before evaluating word order properties.
+
+\paragraph{Estimating Parseability}
+We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}.
+This architecture is remarkably simple: the words of a sentence are encoded into context-sensitive embeddings using bidirectional LSTMs, then a classifier is trained to predict the head for each work.
+The classifier works by calculating a score for every pair of word embeddings $(w_i, w_j)$, indicating the likelihood that the $j$th word is the head of the $i$th word.
+This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
+
+To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags.
+Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged; parsing from POS tags prevents early overfitting.
+This decision was made before evaluating word order properties.
+
+\paragraph{Hyperparameters}
+
+\textbf{TODO study 1 only has 1 set of parameters}
+
+Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
+For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
+Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (predictability or parseability) on the language.
+We then used this setting for creating optimized word order grammars.
+
+\textbf{TODO random seeds for efficiency}
+
+All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
+We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
+
+%\begin{table}[]
+%    \centering
+%    \begin{tabular}{c|cccc}
+%    Entropy penalty     &  0.0001, 0.001 \\
+%    Learning Rate (Word order grammar)     & 5e-6, 1e-5, 2e-5, 5e-5 \\
+%    Momentum (Word order grammar) & 0.8, 0,9 \\
+%    Dropout (Language model) & 0.0, 0.3, 0.5 \\
+%    
+%    \end{tabular}
+%    \caption{Hyperparameters}
+%    \label{tab:my_label}
+%\end{table}
+
+\paragraph{Optimization details}
+
+We employ two common variance reduction methods to improve the estimator~(\ref{eq:estimator}), while keeping it unbiased.
+%For dependency length and 
+For predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word (and its head, in the case of dependency length).
+We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:estimator}) with lower variance.
+Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple}.
+These two methods reduce the variance of the estimator and thereby increase the speed of optimization and reduce training time, without biasing the results.
+For numerical stability, we represent $a_\tau \in [0,1]$ via its logit $\in \mathbb{R}$.
+Furthermore, to encourage exploration of the parameter space, we add an entropy regularization term \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$.
+The weight of the entropy regularization was chosen together with the other hyperparameters.
+%, adding the negative entropy of the $Bernoulli(a_\tau)$ to the loss $R$. %, weighted with $\alpha$.
+
+These techniques for improving (\ref{eq:estimator}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
+
+We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
+For the language model parameters $\phi$, we use plain SGD. 
+For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
+The learning rates and other optimization hyperparameters were determined together with the other hyperparameters.
+
+For each language and objective function, we apply Early Stopping~\citep{prechelt1998early} as follows.
+We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
+This technique, well known in the machine learning literature, helps prevent overfitting to the specific corpora used.
+
+The choice of optimization methods and the stopping criterion were fixed before we investigated word order correlations.
+
+Recall that, for each language, we created 8 optimized languages for each optimization criterion.
+We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
+This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.
+
+ % TODO we could use some language at the top of this section on the general guiding principles for all these decisions.
+Were these implementation decisions made before or after looking at the word order universals results?
+If they were made before, we should highlight that.
+
+\section{Robustness to different language models and parsers}
+
+Here we take up the question of the extent to which our results are dependent on the particular parser and language model used in the optimization process.
+We want to know: when we optimize a word order grammar for efficiency, have we produced a language which is highly efficient \emph{in general}, or one which is highly efficient \emph{for a specific parser}?
+We wish to argue that natural language syntax is optimized for efficiency in general, meaning that syntactic trees are highly recoverable from word orders in principle.
+If it turns out that our optimized languages are only optimal for a certain parser from the NLP literature, then we run the risk of circularity: it may be that the reason this parser was successful in the NLP literature was because it implicitly encoded word order universals in its inductive biases, and thus it would be no surprise that languages which are optimized for parseability also show those universals.
+
+In this connection, we note that the parser and language model architectures we use are highly generic, and do not encode any obvious bias toward natural-language-like word orders.
+The LSTM language model is a generic model of sequence data which is also been used to model financial time series \citep{sirignano2018universal} and purely theoretical chaotic dynamical systems \citep{ogunmolu2016nonlinear}; the neural graph-based parser is simply solving a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
+Nevertheless, it may be the case that a bias toward word order universals is somehow encoded implicitly in the hyperparameters and architectures of these models.
+
+Here we address this question by demonstrating that our languages optimized for prediction and parseability are also optimal under a range of different language models and parsers.
+These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on a particular prediction and parsing algorithms.
+
+\subsection{CKY Parsers}
+
+We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of Collins 1997 (Model 1).
+In our PCFGs, each head independently generates a set of left and right dependents.
+We formulate this as a PCFG where each rule has the form:
+\begin{center}
+	POS$_H$ $\rightarrow$ POS$_H$ POS$_D$
+\end{center}
+for head-initial structures, and
+\begin{center}
+	POS$_H$ $\rightarrow$ POS$_D$ POS$_H$
+\end{center}
+for head-final structures, where each symbol is a POS tag.
+Thus, POS tags act both as terminals and as nonterminals.
+
+We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
+For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm.
+
+This parsing strategy is very different from the neural graph-based parser:
+While the graph-based parser solves a minimum spanning tree problem, the CKY algorithm uses dynamic programming to compute the exact probabilities of trees given a sentence, as specified by the generative model encoded in the PCFG.
+Second, while the graph-based neural parser uses machine learning to induce syntactic knowledge from data, the CKY parser performs exact probabilistic inference.
+%In this sense, the CKY algorithm does not have any architectural biases in itself.
+%On the other hand, the PCFG makes severely simplifying independence assumptions, compared to the universal approximation capabilities of neural network-based systems.
+
+We used the CKY algorithm to compute the parsing loss $H[T|L_\theta(T)]$ on the validation partition of the English UD corpora, for random and optimized ordering grammars.
+Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
+% TODO get UAS for compatibility with other parts of SI
+%We then compared parsability of random and optimized word order grammars on English data.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
+	\caption{Parsability loss $H[T|L_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).}
+    \label{fig:cky-parser}
+\end{figure}
+
+\subsection{Distorted graph-based parsers}
+
+Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization.
+Therefore, we address here the idea that the graph-based parser might have a built-in bias toward parses involving short dependencies, which we call a \key{locality bias}.
+
+A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words.
+Suppose that a parser $P$ has a bias toward positing a dependency between a word and the immediate previous word.
+
+\paragraph{Even--odd order.} A sequence of $n$ words originally ordered as $w_1 w_2 w_3 w_4 \cdots w_n$ is reordered by separating the even and odd indices: $w_2 w_4 w_6 \cdots w_{n-1} w_1 w_3 w_5 \cdots w_n$ (assuming $n$ odd).
+Therefore all words that are adjacent in the original order will be separated by a distance of $\approx n/2$ in the distorted order, while all words of distance 2 in the original order will become adjacent.
+
+\paragraph{Interleaving order.} In interleaving ordering, a sequence originally ordered as $w_1 w_2 w_3 \cdots w_n$ is split in half at the middle (index $m=\ceiling{n/2}$), and the two resulting sequences are interleaved, yielding $w_1 w_m w_2 w_{m+1} w_3 w_{m+3} \cdots w_n$.
+Thus all words that were originally adjacent will have distance 2 in the distorted order, with the intervening word coming from a very distant part of the sentence.
+
+\paragraph{Inwards order.} A sequence originally ordered as $w_1 w_2 w_3 \cdots w_{n-1} w_n$ is ordered from the edges of the string inwards, as $w_1 w_n w_2 w_{n-1} \cdots w_{\ceiling{n/2}}$.
+This corresponds to folding the string in on itself once, or equivalently, splitting the sequence in half at the middle, then interleaving the two resulting sequences after reversing the second one.
+The result is that the most non-local possible dependencies in the original order become the most local dependencies in the distorted order.
+
+\paragraph{Sorted order.} A sequence is reordered by sorting by POS tags, and randomizing the order within each block of identical POS tags.
+To each word, we then add a symbol encoding the original position in the sequence.
+For instance
+\begin{center}
+PRON VERB PRON
+\end{center}
+may be reordered as
+\begin{center}
+PRON 1 PRON 3 VERB 2
+\end{center}
+or
+\begin{center}
+PRON 3 PRON 1 VERB 2
+\end{center}
+The numbers are provided to the parser as atomic symbols from a vocabulary ranging from 1 to 200; numbers greater than 200 (which may occur in extremely long sentences) are replaced by an out-of-range token.
+
+\paragraph{Experiments}
+Using English and Japanese data, we trained parsers for the ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
+Resulting parsing accuracy scores are shown in Figure~\ref{fig:distorted-parser}.
+In all settings, the language optimized for efficiency achieved higher parsing accuracy than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
+%Note that absolute values of parsing accuracy vary between the orderings, being higher for the actual orderings than for distorted orderings, in which trees become nonprojective.
+
+\begin{figure}
+    \centering
+    English
+
+    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-english.pdf}
+
+    Japanese
+
+    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
+	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').}
+    \label{fig:distorted-parser}
+\end{figure}
+
+\subsection{n-gram language models}
+
+We constructed a Bigram model with Kneser-Ney smoothing.
+A bigram model predicts each word taking only the previous word into account.
+This contrasts with LSTMs, which take the entire context into consideration.
+Thus, bigram models and LSTMs stand on opposing ends of a spectrum of language models taking more and more aspects of the context into account.
+
+We estimated language models on the training partitions, and used the validation partitions to estimate surprisal.
+We conducted this for the ten random and the best optimized ordering grammars on English and Japanese data.
+Results (Figure~\ref{fig:bigrams}) show that languages optimized for efficiency are optimal for a bigram language model.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.4]{../results/bigrams/bigrams.pdf} 
+	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).}
+    \label{fig:bigrams}
+\end{figure}
+
+\subsection{Discussion}
+
+It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.
+
+\section{Effects of data sparsity}
+
+Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
+If the difference between random and real grammars is due to data sparsity, we expect that it decreases as the amount of training data is increased.
+If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference persists as training data is increased.
+
+We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
+We cosidered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
+For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
+Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.
+
+\textbf{do same for language model, and for the other random grammars}
+
+The resulting curves are shown in Figure~\ref{fig:learning-czech}.
+A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.4]{figures/learning-parser-czech.pdf} 
+        \caption{Parsing accuracy (measured in UAS) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
+    \label{fig:learning-czech}
+\end{figure}
+
+\section{Study 2, Correlation between Universals and Efficiency}
+
+\begin{figure}[ht]
+    \centering
+
+    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
+    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
+    \includegraphics[scale=.25]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
+
+	\caption{Correlation between the number of satisfied correlations ($x$-axis) and efficiency, parseability, and predictability ($y$-axis), for the 51 real languages.}
+    \label{fig:todo}
+\end{figure}
+
+\section{Comparing Extracted Grammars to Linguistic Judgments}
+
+% user@user-X510UAR:~/grammar-optim/results/grammars$ git add comparison-table.tex
+
+Here, we validate our grammars by comparing the dominant orders of six syntactic relations that are also annotated in WALS.
+Among the eight relations that we were able to test, five are annotated in WALS.
+Here, we compare our grammars with WALS on these five relations, and the verb-object relation.
+
+WALS has some data for all but seven languages.\footnote{Note that Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
+In the table, we compare those with the grammar we extracted for Croatian, noting that it agrees with the Serbian grammar.}
+WALS has data for 74\% of the entries, and lists a dominant order for 91\% of these.
+The grammars we extracted from the corpora agree with WALS in 96~\% out of those cases where WALS records a dominant word order.
+
+For each of the six syntactic relation, the first column provides the ordered coded in the extracted grammar; the second column provides the order coded in WALS (DH for dependent-head, HD for head-dependent order).
+`?' indicates that WALS has no data.
+$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
+Finally, `--' indicates that the relation does not occur in the corpus.
+
+\small{
+\begin{longtable}{l||ll|ll|ll|ll|ll|ll|llllll}
+%	   for(dep in c("obj", "lifted_mark", "acl", "nmod", "obl")) {
+		   Language 
+		   &	\multicolumn{2}{|c|}{Objects} 
+		   &	\multicolumn{2}{|c|}{Adpositions} 
+		   &	\multicolumn{2}{|c|}{Compl.} 
+		   &	\multicolumn{2}{|c|}{Rel.Cl.} 
+		   &	\multicolumn{2}{|c|}{Genitive} 
+		   &	\multicolumn{2}{|c|}{PP}  \\ \hline\hline
+	\input{../results/grammars/comparison-table.tex}
+\end{longtable}
+}
+
+\section{Efficiency and Prior Formalizations}
+
+Here, we discuss how our formalization of efficiency relates to prior work testing functional explanations in linguistics.
+
+The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
+Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication.
+
+Our objective function is identical to that of \cite{ferreri2003least}, and very similar to the functions in other computational work formalizing Zipf's ideas~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
+
+Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.
+
+All these accounts formalize the amount of information provided as the \key{mutual information} between the forms that a listener receives, and the meanings that the speaker intends to communicate to the speaker.
+In our work, the underlying meanings are the syntactic structures; the forms are the linearized sentences.
+In \cite{ferreri2003least, kemp2012kinship}, forms are words referring to objects.
+In \cite{regier2015word} TODO
+In \cite{zaslavsky2018efficient}, forms are color words, and meanings are TODO.
+
+The accounts
+
+- our function
+
+similar: RSA, there it's utterance-level
+
+\cite{frank2012predicting,goodman2013knowledge}
+
+- number of different forms
+
+- complexity of the mapping \cite{kemp2012kinship}
+
+- MI between
+\cite{zaslavsky2018efficient}
+
+We model the first term as the 
+ degree to which listeners can reconstruct dependency parse structures from an utterance, i.e., the \key{parseability} if the language.
+This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
+\begin{equation}
+	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+\end{equation}
+where the sum runs over all possible pairs of sentences $l$ and syntactic structures $t$ in the language.
+%This quantity describes the degree to which syntactic structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
+
+We formalize the complexity of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
+In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
+Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+%and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+
+In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
+\begin{equation}
+	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+\end{equation}
+where the sum runs over all possible sentences $u$ that belong to the language.
+%In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
+
+The efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+\end{equation}
+with an interpolation weight $\lambda \in [0,1)$.
+In all results from Study 2 reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
+See SI appendix section 5 for mathematical discussion of $\lambda$, and robustness to other choices.
+
+\section{Relation to the Information Bottleneck}
+
+Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
+\begin{equation}
+    \label{eq:efficiency-again}
+    J_T(L) = I[L(T); T] - \lambda H[L(T)].
+\end{equation}
+This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}.
+In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information.
+The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$.
+Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
+\begin{equation}
+    \label{eq:ib}
+    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
+\end{equation}
+In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
+
+In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$.
+To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$.
+We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized.
+Then our objective is parallel to Eq.~\ref{eq:ib}:
+\begin{equation}
+    \label{eq:efficiency-kinda-ib}
+    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
+\end{equation}
+Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$.
+Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
+\begin{equation*}
+    J_T(L) = I[U;T] - \lambda H[U].
+\end{equation*}
+
+Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
+
+\bibliographystyle{unsrtnat}
+\bibliography{everything}
+
+\end{document}

### 2019-05-14 second-draft/pnas-word-order-si.tex

--- 2019-05-13 second-draft/pnas-word-order-si.tex
@@
 
+\section{Efficiency and Prior Formalizations}
+
+Here, we discuss how our formalization of efficiency relates to prior work testing functional explanations in linguistics.
+
+The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
+Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication.
+
+Our objective function is identical to that of \cite{ferreri2003least}, and very similar to the functions in other computational work formalizing Zipf's ideas~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
+
+Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.
+
+All these accounts formalize the amount of information provided as the \key{mutual information} between the forms that a listener receives, and the meanings that the speaker intends to communicate to the speaker.
+In our work, the underlying meanings are the syntactic structures; the forms are the linearized sentences.
+In \cite{ferreri2003least, kemp2012kinship}, forms are words referring to objects.
+In \cite{regier2015word} TODO
+In \cite{zaslavsky2018efficient}, forms are color words, and meanings are TODO.
+
+The accounts
+
+- our function
+
+similar: RSA, there it's utterance-level
+
+\cite{frank2012predicting,goodman2013knowledge}
+
+- number of different forms
+
+- complexity of the mapping \cite{kemp2012kinship}
+
+- MI between
+\cite{zaslavsky2018efficient}
+
+We model the first term as the 
+ degree to which listeners can reconstruct dependency parse structures from an utterance, i.e., the \key{parseability} if the language.
+This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
+\begin{equation}
+	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+\end{equation}
+where the sum runs over all possible pairs of sentences $l$ and syntactic structures $t$ in the language.
+%This quantity describes the degree to which syntactic structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
+
+We formalize the complexity of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
+In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
+Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+%and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+
+In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
+\begin{equation}
+	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+\end{equation}
+where the sum runs over all possible sentences $u$ that belong to the language.
+%In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
+
+The efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+\end{equation}
+with an interpolation weight $\lambda \in [0,1)$.
+In all results from Study 2 reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
+See SI appendix section 5 for mathematical discussion of $\lambda$, and robustness to other choices.
+
 \section{Relation to the Information Bottleneck}

### 2019-05-15 second-draft/pnas-word-order-si.tex
SI

--- 2019-05-13 second-draft/pnas-word-order-si.tex
@@
 
+\section{Efficiency and Prior Formalizations}
+
+Here, we discuss how our formalization of efficiency relates to prior work testing functional explanations in linguistics.
+
+%The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
+%Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication.
+
+%Recall our formalization of efficiency:
+%\begin{equation}
+%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+%\end{equation}
+%\begin{equation}
+%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+%\end{equation}
+%\begin{equation}\label{eq:efficiency}
+%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+%\end{equation}
+
+Our objective function is identical to that of \cite{ferreri2003least}, and very similar to the functions in other computational work formalizing Zipf's ideas~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
+
+Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.
+
+All these accounts formalize the amount of information provided as the \key{mutual information} between the forms that a listener receives, and the meanings that the speaker intends to communicate to the speaker.
+In our work, the underlying meanings are the syntactic structures; the forms are the linearized sentences communicated by the speaker to the listener.
+By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
+The syntactic structure (including the labels on the arcs) specifies how the meanings of words are combined.
+Therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
+In \cite{ferreri2003least, kemp2012kinship}, forms are words referring to objects.
+In \cite{regier2015word} TODO
+In \cite{zaslavsky2018efficient}, forms are color words, and meanings are TODO.
+
+The accounts differ in the formalizations of complexity and diversity:
+Like our model, \cite{ferreri2003least} model the complexity as the \emph{entropy} of the distribution of forms.
+Closely related is the Information Bottleneck approach \cite{zaslavsky2018efficient, zaslavsky2019semantic}, which models complexity as the mutual information between forms (e.g., color words) and meanings (e.g., color chips).
+This mutual information formalization is appropriate to encodes that are nondeterministic \cite{zaslavsky2018efficient}.
+As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoded is deterministic, and which uses the entropy as the complexity measure.
+
+A few models have used other measures that are not direct information-theoretic quantities.
+In \cite{regier2015wordTODO}, the cost function is the number of different forms; in \cite{kemp2012kinship} it is the complexity of the concepts encoded.
+
+%In terms of information theory \cite{shannon1948mathematical}, this corresponds to the notion of the rate-distortion tradeoff.
+%In classical rate-distortion theory, the cost of a code is its rate, the mutual information between codes and underlying messages (as in the Information Bottleneck model).
+% classical models of distortion are the Hamming distortion and 
+%TODO also mention rate-distortion theory
+
+- our function
+
+similar: RSA, there it's utterance-level
+
+\cite{frank2012predicting,goodman2013knowledge}
+
+%- number of different forms 
+%
+%- complexity of the mapping 
+%
+%- MI between
+%
+%
+%
+%We model the first term as the 
+% degree to which listeners can reconstruct dependency parse structures from an utterance, i.e., the \key{parseability} if the language.
+This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
+%\begin{equation}
+%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+%\end{equation}
+%where the sum runs over all possible pairs of sentences $l$ and syntactic structures $t$ in the language.
+%%This quantity describes the degree to which syntactic structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
+%
+%
+%
+%We formalize the complexity of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
+%In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
+%Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+%%and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+%
+%In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
+%\begin{equation}
+%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+%\end{equation}
+%where the sum runs over all possible sentences $u$ that belong to the language.
+%%In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
+%
+%
+%
+%The efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
+%\begin{equation}\label{eq:efficiency}
+%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+%\end{equation}
+%with an interpolation weight $\lambda \in [0,1)$.
+%In all results from Study 2 reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
+%See SI appendix section 5 for mathematical discussion of $\lambda$, and robustness to other choices.
+%
+%
+
 \section{Relation to the Information Bottleneck}

### 2019-05-15 second-draft/pnas-word-order-si.tex
SI

@@
 
-Here, we discuss how our formalization of efficiency relates to prior work testing functional explanations in linguistics.
-
 %The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
@@
 
-Our objective function is identical to that of \cite{ferreri2003least}, and very similar to the functions in other computational work formalizing Zipf's ideas~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
-
+Here we discuss how our formalization of efficiency relates to other information-theoretic work on language.
+
+Our objective function is identical to the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory} as a general efficiency metric for communicative systems, with syntactic structures taking the role of underlying meanings that the speaker wants to convey.
+This function is also equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
+
+Very similar metrics have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
 Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.
@@
 All these accounts formalize the amount of information provided as the \key{mutual information} between the forms that a listener receives, and the meanings that the speaker intends to communicate to the speaker.
+Most previous studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} are concerned with the meanings of individual words; they thus take the forms to be individual words, which refer to individual objects and their properties.
+We instead study how words are composed into sentences.
 In our work, the underlying meanings are the syntactic structures; the forms are the linearized sentences communicated by the speaker to the listener.
@@
 Therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
-In \cite{ferreri2003least, kemp2012kinship}, forms are words referring to objects.
-In \cite{regier2015word} TODO
-In \cite{zaslavsky2018efficient}, forms are color words, and meanings are TODO.
-
-The accounts differ in the formalizations of complexity and diversity:
-Like our model, \cite{ferreri2003least} model the complexity as the \emph{entropy} of the distribution of forms.
-Closely related is the Information Bottleneck approach \cite{zaslavsky2018efficient, zaslavsky2019semantic}, which models complexity as the mutual information between forms (e.g., color words) and meanings (e.g., color chips).
-This mutual information formalization is appropriate to encodes that are nondeterministic \cite{zaslavsky2018efficient}.
-As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoded is deterministic, and which uses the entropy as the complexity measure.
+
+%These information-theoretic accounts differ in the formalizations of complexity and diversity:
+Like our model, \cite{ferreri2003least,ferrericancho2007global,futrell2017memory} model the complexity as the \key{entropy} of the distribution of forms, which arises naturally from the role played by surprisal in determining human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+This model is closely related to the Information Bottleneck approach \cite{zaslavsky2018efficient, zaslavsky2019semantic}, which models complexity as the mutual information between forms (e.g., words) and their referents (e.g., individual objects).
+This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
+As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure.
 
 A few models have used other measures that are not direct information-theoretic quantities.
-In \cite{regier2015wordTODO}, the cost function is the number of different forms; in \cite{kemp2012kinship} it is the complexity of the concepts encoded.
+In \cite{regier2007color, xu2014numeral, xu2016historical}, the cost function is the number of different forms; in \cite{kemp2012kinship} it is the complexity of the concepts encoded.
+The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
+Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the information bottleneck formalism, bringing them even closer to our formalization.
+
+In addition to these models, which quantify the efficiency of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context in models of pragmatic reasoning in reference games \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
+In line with the other models, these assume that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely}.
 
@@
 
-- our function
-
-similar: RSA, there it's utterance-level
-
-\cite{frank2012predicting,goodman2013knowledge}
+%- our function
+
+%similar: RSA, there it's utterance-level
 

### 2019-05-16 second-draft/pnas-word-order-si.tex
si

@@
 
+\paragraph{Per-Language Results}
+Figure~\ref{fig:pareto-per-lang}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
+\caption{per language results.
+Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\textcolor{red}{would be good to have same colors as in main paper} All data are $z$-transformed.}\label{fig:pareto-per-lang}
+\end{figure}
+
 \section{Supplementary Analyses for Study 2}
@@
 
+\section{Other (Preregistered) Experiments}
+
+Here we report the results of prior preregistered simulation runs.
+
+\begin{table}[ht]
+	\begin{center}
+\begin{tabular}{|c|ccc|c}
+	\hline
+	%&
+& $\lambda=1.0$ & $\lambda=0.9$ & $\lambda=0.9$ \\ \hline \hline 
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}		  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_case.pdf}     } 
+ 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_case.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_cop.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_cop.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_aux.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_aux.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_nmod.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_nmod.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	 
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_acl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_acl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_mark.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_mark.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_obl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_obl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_xcomp.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_xcomp.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }  \\
+&&&\\ \hline
+ \hline
+\end{tabular}
+	\end{center}
+\caption{Previous experiment with $\lambda=0.9$ (Preregistration)}\label{table:corr-resu}
+\end{table}
+
 \bibliographystyle{unsrtnat}

### 2019-05-16 results/correlations/joint-evaluation-efficiency09.R

@@ -1,14 +1,17 @@
-data = read.csv("CS_SCR/deps/manual_output_funchead_two_coarse_lambda09_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+# For the first \lambda=0.9 simulation
+
+
+data = read.csv("../../grammars/manual_output_funchead_two_coarse_lambda09_best_balanced/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
 library(forcats)
 library(dplyr)
 library(tidyr)
 library(ggplot2)
 data = data %>% mutate(Language = fct_recode(Language, "Ancient_Greek" = "Ancient", "Old_Church_Slavonic" = "Old"))
-languages = read.csv("languages-iso_codes.tsv")
+languages = read.csv("../languages/languages-iso_codes.tsv")
 data  = merge(data, languages, by=c("Language"), all.x=TRUE)
 library("brms")
 dependency = "nmod"
-dependencies = c("acl", "advmod", "aux", "lifted_case", "lifted_cop", "lifted_mark", "nmod", "nsubj", "obl", "xcomp")
+dependencies = c("acl", "aux", "lifted_case", "lifted_cop", "lifted_mark", "nmod", "obl", "xcomp")
 objs = data %>% filter(CoarseDependency == "obj") %>% mutate(obj = pmax(0, sign(DH_Weight))) %>% select(obj, Language, FileName)
 data = data %>% filter((CoarseDependency %in% dependencies))
 data = merge(data, objs, by=c("Language", "FileName"))
@@ -17,15 +20,15 @@ data = data %>% mutate(dir = ifelse(DH_Weight == 0, NA, dir))
 data$agree = (data$dir == data$obj)
 data = unique(data %>% select(Family, Language, FileName, CoarseDependency, agree)) %>% spread(CoarseDependency, agree)
 
-model3 = brm(cbind(acl, advmod, aux, lifted_case, lifted_cop, lifted_mark, nmod, nsubj, obl, xcomp) ~ (1|p|Family) + (1|q|Language), family="bernoulli", data=data, iter=5000)
[…]

### 2019-05-16 results/plane/plane-2D-best-visualization-large-3.R

@@ -157,49 +157,51 @@ plot = ggplot(data, aes(x=(Pars-MeanPars)/SDPars, y=(Surp-MeanSurp)/SDSurp, colo
 plot = ggplot(data, aes(x=(Pars-MeanPars), y=(Surp-MeanSurp), color=Type, group=Type)) +geom_point() + facet_wrap(~Language)
 
 
+data = data %>% mutate(Pars_z = (Pars-MeanPars)/SDPars, Surp_z=(Surp-MeanSurp)/SDSurp)
+
 subData = data %>% filter(Type %in% c("manual_output_funchead_two_coarse_parser_best_balanced", "manual_output_funchead_two_coarse_lambda09_best_large", "manual_output_funchead_langmod_coarse_best_balanced"))
 
 subData = subData[order(subData$Type),]
 
 
-Surp = c()
-Pars = c()
+Surp_z = c()
+Pars_z = c()
 Language = c()
 for(language in unique(subData$Language)) {
    u = subData %>% filter(Language == language)
-   u = u[order(u$Pars),]
-   lastSurp = max(u$Surp)
-   lastPars = u$Pars[[1]]
-   Surp = c(Surp, lastSurp)
-   Pars = c(Pars, lastPars)
+   u = u[order(u$Pars_z),]
+   lastSurp_z = max(u$Surp_z)
+   lastPars_z = u$Pars_z[[1]]
+   Surp_z = c(Surp_z, lastSurp_z)
+   Pars_z = c(Pars_z, lastPars_z)
    Language = c(Language, language)
    for(i in (1:nrow(u))) {
-      surpHere = u$Surp[[i]]
-      parsHere = u$Pars[[i]]
-      if(surpHere < lastSurp) {
-         Surp = c(Surp, surpHere)
-         Pars = c(Pars, parsHere)
+      surpHere = u$Surp_z[[i]]
+      parsHere = u$Pars_z[[i]]
+      if(surpHere < lastSurp_z) {
+         Surp_z = c(Surp_z, surpHere)
+         Pars_z = c(Pars_z, parsHere)
          Language = c(Language, language)
         
-         lastSurp = surpHere
-         lastPars = parsHere
+         lastSurp_z = surpHere
+         lastPars_z = parsHere
       }
    }
-   Surp = c(Surp, lastSurp)
-   Pars = c(Pars, max(u$Pars))
+   Surp_z = c(Surp_z, lastSurp_z)
[…]

### 2019-05-18 second-draft/pnas-word-order-si.tex
SI

@@
 
-\textcolor{red}{TODO At some point say something about \cite{dunn2011evolved}, since that got a lot of attention?
-can refer to follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
-
 Here we describe how we selected the correlations in Table 1 of the main paper, and how we formalized these using syntactic relations defined by Universal Dependencies.
 
-We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.
+We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding objections by \cite{dunn2011evolved}, we refer to the follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
 Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
@@
 \end{figure}
+
+\textcolor{red}{also provide p value, CI for how many baselines are better in either of the two dimensions}
 
@@
 
+\section{Languages and Corpora}
+
+\begin{table}[ht]
+\begin{tabular}{lllllll}
+Language & ISO Code & Family & Sentences (train/held-out) & Words (train/held-out) \\ \hline
+\input{../results/corpus-size/corpora-table.tex}
+\end{tabular}
+\caption{Languages}
+\end{table}
+
+\section{Dependency Length Minimization}
+
+\textcolor{red}{TODO}
+We have directly implemented parseability and predictability using general-purpose statistical models of language and dependency structures.
+Prior work has considered Dependency Length Minimization (DLM) as an approximate formalization of efficiency optimization~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+This is the idea that languages order words so as to minimize the overall length of dependencies.
+This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
+Indeed, directly modeling efficiency optimization predicts DLM, and appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
+
 \bibliographystyle{unsrtnat}

### 2019-05-18 second-draft/pnas-word-order-si.tex
SI

--- 2019-05-16 second-draft/pnas-word-order-si.tex
@@
 
+\textcolor{blue}{Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
+Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
+The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order, finding 16 correlations to likely be universal.
+Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:corr-dryer}.
+%Due to data availability, we restricted to 11 universals and further collapsed to the eight correlations shown in Table~\ref{table:corr-dryer}.
+Testing efficiency on the excluded correlations would be feasible with additional annotation of data.
+For instance, one of these could be tested by labeling the existing databases for the distinction between manner adverbs from other kinds of adverbs.
+For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).}
+
 \section{Supplementary Analyses for Study 1}
@@
 
-Our objective function is identical to the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory} as a general efficiency metric for communicative systems, with syntactic structures taking the role of underlying meanings that the speaker wants to convey.
+Our objective function is identical to the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems, with syntactic structures taking the role of underlying meanings that the speaker wants to convey.
 This function is also equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
 
-Very similar metrics have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
+Very similar metrics have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
 Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.
@@
 Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
+
+\section{Relation to DLM}
+
+DLM turns out to be predicted by our model
 

### 2019-05-18 second-draft/pnas-word-order-si.tex
SI

--- 2019-05-18 second-draft/pnas-word-order-si.tex
@@
 In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
+%Our method will be applicable more generally to the task of optimizing grammars for a broad class of objective functions.
 We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
@@
 At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
-
-See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
+Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.
+
+See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates used in this optimization scheme.
 

### 2019-05-18 second-draft/pnas-word-order-si.tex
SI

--- 2019-05-18 second-draft/pnas-word-order-si.tex
@@
 
-\textcolor{red}{TODOs}
-
-\textcolor{red}{TODO This idea makes strong predictions, and turns out to overpredict correlations (see SI, Section X).}
-
 \section{Formalization of Correlations}\label{sec:correlations}
@@
 
+%\textcolor{blue}{Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
+%Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
+%The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order, finding 16 correlations to likely be universal.
+%Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:corr-dryer}.
+%Testing efficiency on the excluded correlations would be feasible with additional annotation of data.
+%For instance, one of these could be tested by labeling the existing databases for the distinction between manner adverbs from other kinds of adverbs.
+%For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).}
+
 \section{Supplementary Analyses for Study 1}
@@
 \paragraph{Per-Language Results}
-Figure~\ref{fig:pareto-per-lang}
-
-\begin{figure}[ht]
+In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.\footnote{Note that in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
+Due to use of stochastic gradient descent in optimization, there is some amount of noise in the optimization process.}
+
+\begin{figure}
 \centering
 \includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
-\caption{per language results.
-Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\textcolor{red}{would be good to have same colors as in main paper} All data are $z$-transformed.}\label{fig:pareto-per-lang}
+\caption{Predictability and parseability of 51 languages.
+Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.
+All data are $z$-scored.}\label{fig:pareto-per-lang}
 \end{figure}
@@
 
-\textcolor{red}{at some place also say the following: Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
-}
-
-Further, we report results for grammars optimized for dependency length minimization (DLM).
+Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM}).
 Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
+We optimize grammars using the same stochastic gradient descent method as we use for efficiency optimization (Section REF).
 Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
-Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
-In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
+Indeed, we found that DLM is predicted by efficiency optimization, as discussed in Section REF.
+Results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
 
@@
 	\hline
-&	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+&	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
 % From Dryer	
@@
 	\hline
-	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
 \hline
@@
 While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
+In conclusion, $\lambda$ is constrained to be in $[0,1)$, which values closer to $1$ placing similar weights on both predictability and parsability, whereas values closer to $0$ diminish the role of predictability.
 
 In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
-
 The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
-We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations.
+%We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations, see Table REF.
 We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (Table~\ref{tab:all-predictions-1}).
+%Indeed, the same holds for all of the 38 syntactic relations in UD  (Table~\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})
+%We additionally report results with $\lambda = 0$, i.e., placing all weight on parseability.
+That is, at $\lambda$ close to its maximal value, the predictions of optimizing the two scoring functions individually add up to the predictions of efficiency optimization.
+Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}).
+We also verified the robustness of our results by reporting results with $\lambda=1$ in Table REF>
 
@@
 In this section, we describe the details of the neural network architectures.
-All choices described in this section were fixed before evaluating word order properties and the efficiency of real language.
+Choices follow standard practice and machine learning and were fixed before evaluating word order properties, and the efficiency of real languages.
 
@@
 Using POS tags is intended to prevent overfitting on small corpora.
-These decisions were made before evaluating word order properties.
 
@@
 For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
-Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (predictability or parseability) on the language.
+These parameters are shown in Table~\ref{tab:hyperparameters}.
+Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (efficiency, predictability, parseability) on the language.
 We then used this setting for creating optimized word order grammars.
 
-\textcolor{red}{TODO random seeds for efficiency}
+%\textcolor{red}{TODO random seeds for efficiency}
 
@@
 
-%\begin{table}[]
-%    \centering
-%    \begin{tabular}{c|cccc}
-%    Entropy penalty     &  0.0001, 0.001 \\
-%    Learning Rate (Word order grammar)     & 5e-6, 1e-5, 2e-5, 5e-5 \\
-%    Momentum (Word order grammar) & 0.8, 0,9 \\
-%    Dropout (Language model) & 0.0, 0.3, 0.5 \\
-%    
-%    \end{tabular}
-%    \caption{Hyperparameters}
-%    \label{tab:my_label}
-%\end{table}
-
-\paragraph{Optimization details}
-
+\begin{table}[]
+    \centering
+    \begin{tabular}{c|cccc}
+    Entropy penalty     &  0.0001, 0.001 \\
+    Learning Rate (Word order grammar)     & 5e-6, 1e-5, 2e-5, 5e-5 \\
+    Momentum (Word order grammar) & 0.8, 0,9 \\
+    Dropout (Language model) & 0.0, 0.3, 0.5 \\
+
+    \end{tabular}
+    \caption{Selected Hyperparameters}
+    \label{tab:hyperparameters}
+\end{table}
+
+\paragraph{Improved Unbiased Gradient Estimator}
 We employ two common variance reduction methods to improve the estimator~(\ref{eq:dtheta}), while keeping it unbiased.
 %For dependency length and 
-For predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word (and its head, in the case of dependency length).
+% (and its head, in the case of dependency length)
+For predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word.
 We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:dtheta}) with lower variance.
-Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple}.
+Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple} (the word's surprisal in the case of predictability, and the negative log-probability of the correct head and relation label in the case of parseability).
 These two methods reduce the variance of the estimator and thereby increase the speed of optimization and reduce training time, without biasing the results.
@@
 
+\paragraph{Learning Rate and Stopping Criterion}
 We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
-For the language model parameters $\phi$, we use plain SGD. 
+For the language model parameters $\phi$, we use plain Stochastic Gradient Descent without momentum, as recommended by \citet{merity2018regularizing}. 
 For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
@@
 
-For each language and objective function, we apply Early Stopping~\citep{prechelt1998early} as follows.
+For gramar optimization, we apply Early Stopping~\citep{prechelt1998early} as follows.
 We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
-This technique, well known in the machine learning literature, helps prevent overfitting to the specific corpora used.
+This is a standard technique for optimizing neural models, and helps prevent overfitting to the specific corpora used.
 
@@
 \paragraph{Optimized Grammars}
-
 Recall that, for each language, we created 8 optimized languages for each optimization criterion.
@@
 For efficiency optimization, for each grammar, we ran efficiency optimization with three different random seeds, selecting among these the seed that yielded the best overall efficiency value.
-This
+This helps control for variation across random seeds for the stochastic gradient descent optimization method.
 
@@
 
-\section{Dependency Length Minimization}
+\section{Dependency Length Minimization}\label{sec:DLM}
 
@@
 Indeed, directly modeling efficiency optimization predicts DLM, and appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
+
+\textcolor{red}{at some place also say the following: Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
+}
+
+Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM}).
+Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
+Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
+Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
+In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
 

### 2019-05-18 second-draft/pnas-word-order-si.tex
SI

@@
 % Use the lineno option to display guide line numbers if required.
+
+%\usepackage{lineno}
+%\linenumbers
 
@@
 \DeclareMathOperator{\E}{\mathop{\mathbb{E}}}
+
+\usepackage{footnote}
+\makesavenoteenv{tabular}
+\makesavenoteenv{table}
+
+\usepackage{siunitx}
 
@@
 
-We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding objections by \cite{dunn2011evolved}, we refer to the follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
+We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.
+%\footnote{Regarding objections by \cite{dunn2011evolved}, we refer to the follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
 Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
@@
 
+\section{Formalizing Comunicative Efficiency}
+\subsection{Relation to Other Work}
+
+%The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
+%Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication.
+
+%Recall our formalization of efficiency:
+%\begin{equation}
+%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+%\end{equation}
+%\begin{equation}
+%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+%\end{equation}
+%\begin{equation}\label{eq:efficiency}
+%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+%\end{equation}
+
+Here we discuss how our formalization of communicative efficiency relates to other information-theoretic work on language.
+
+Our formalization of efficiency is identical to the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems, with syntactic structures taking the role of underlying meanings that the speaker wants to convey.
+This function is also equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
+Very similar metrics have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
+%More generally, these approaches instantiate Shannon's Rate Distortion theory.
+Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.
+
+All these accounts formalize the amount of information provided as the \key{mutual information} between the forms that a listener receives, and the meanings that the speaker intends to communicate to the speaker.
+Most previous studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} are concerned with the meanings of individual words; they thus take the forms to be individual words, which refer to individual objects and their properties.
+We instead study how words are composed into sentences.
+In our work, the underlying meanings are the syntactic structures; the forms are the linearized sentences communicated by the speaker to the listener.
+By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
+The syntactic structure specifies how the meanings of words are combined; therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
+
+%These information-theoretic accounts differ in the formalizations of complexity and diversity:
+Like our model, \cite{ferreri2003least,ferrericancho2007global,futrell2017memory} model the complexity as the \key{entropy} of the distribution of forms, which arises naturally from the role played by surprisal in determining human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+This model is closely related to the Information Bottleneck approach \cite{zaslavsky2018efficient, zaslavsky2019semantic}, which models complexity as the mutual information between forms (e.g., words) and their referents (e.g., individual objects).
+This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
+As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure.
+
+A few models have used other complexity measures that are not explicitly information-theoretic quantities.
+In \cite{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \cite{kemp2012kinship} it is the difficulty of defining the concepts that are encoded.
+The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
+Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the information bottleneck formalism, bringing them even closer to our formalization.
+
+In addition to these models, which quantify the efficiency of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context, such as models of pragmatic reasoning in reference games \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
+In line with the other models, these assume that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely}.
+
+\subsection{Possible values of $\lambda$}\label{sec:lambda}
+
+In the efficiency objective
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
+\end{equation}
+the value of $\lambda$ is constrained to be in $[0,1)$.
+This means, surprisal must be weighted less strongly than parsability.
+
+Greater values of $\lambda$ mathematically result in degenerate solutions.
+To show this, note that the following inequality always holds:
+\begin{equation}
+I[L(T); T] \leq H[L(T)].
+\end{equation}
+Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
+\begin{equation}
+    J_T(L) = I[L(T); T] - \lambda H[L(T)] \leq 0.
+\end{equation}
+and it takes the maximally possible value zero if $L(T)$ maps all trees to a single utterance, in which case both $I[L(T); T]$ and $H[L(T)]$ are zero.
+This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
+While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
+In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parsability, whereas values closer to $0$ diminish the role of predictability.
+
+In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
+The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
+%We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations, see Table REF.
+We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (Table~\ref{tab:all-predictions-1}).
+%Indeed, the same holds for all of the 38 syntactic relations in UD  (Table~\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})
+%We additionally report results with $\lambda = 0$, i.e., placing all weight on parseability.
+That is, at $\lambda$ close to its maximal value, the predictions of optimizing the two scoring functions individually add up to the predictions of efficiency optimization.
+Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}).
+We also verified the robustness of our results to small variations of $\lambda$, by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu}.
+
+%In terms of information theory \cite{shannon1948mathematical}, this corresponds to the notion of the rate-distortion tradeoff.
+%In classical rate-distortion theory, the cost of a code is its rate, the mutual information between codes and underlying messages (as in the Information Bottleneck model).
+% classical models of distortion are the Hamming distortion and 
+%TODO also mention rate-distortion theory
+
+%- our function
+
+%similar: RSA, there it's utterance-level
+
+%- number of different forms 
+%
+%- complexity of the mapping 
+%
+%- MI between
+%
+%
+%
+%We model the first term as the 
+% degree to which listeners can reconstruct dependency parse structures from an utterance, i.e., the \key{parseability} if the language.
+This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
+%\begin{equation}
+%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+%\end{equation}
+%where the sum runs over all possible pairs of sentences $l$ and syntactic structures $t$ in the language.
+%%This quantity describes the degree to which syntactic structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
+%
+%
+%
+%We formalize the complexity of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
+%In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
+%Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+%%and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+%
+%In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
+%\begin{equation}
+%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+%\end{equation}
+%where the sum runs over all possible sentences $u$ that belong to the language.
+%%In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
+%
+%
+%
+%The efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
+%\begin{equation}\label{eq:efficiency}
+%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+%\end{equation}
+%with an interpolation weight $\lambda \in [0,1)$.
+%In all results from Study 2 reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
+%See SI appendix section 5 for mathematical discussion of $\lambda$, and robustness to other choices.
+%
+%
+
+%\section{Relation to the Information Bottleneck}
+%
+%Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
+%\begin{equation}
+%    \label{eq:efficiency-again}
+%    J_T(L) = I[L(T); T] - \lambda H[L(T)].
+%\end{equation}
+%This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}.
+In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information.
+The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$.
+Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
+%\begin{equation}
+%    \label{eq:ib}
+%    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
+%\end{equation}
+%In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
+%
+%In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$.
+To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$.
+We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized.
+Then our objective is parallel to Eq.~\ref{eq:ib}:
+%\begin{equation}
+%    \label{eq:efficiency-kinda-ib}
+%    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
+%\end{equation}
+%Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$.
+Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
+%\begin{equation*}
+%    J_T(L) = I[U;T] - \lambda H[U].
+%\end{equation*}
+%
+%Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
+%
+%
+
 \section{Supplementary Analyses for Study 1}
 
-\paragraph{Relative Efficiency of Real Grammars}
-In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
-We also conducted the analysis using a Binomial test (one-sided), testing whether the real gramar is more efficient than the \emph{median} of random grammars, avoiding any distributional assumption on the random grammars.
-As before, we used Hochberg's step-up procedure (Note that the tests for different languages are independent, as different random grammars are evaluated for each language), with $\alpha=0.05$.
-In this analysis, real grammars improved in parseability for 80\% of languages, in predictability for 69\% of languages, and in either of both in 92\% of languages ($p<0.05$, with Bonferroni correction).
-
-\paragraph{z-Scored Data (Figure (REF))}
-We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section~\ref{sec:lambda}), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
-
-Within each language, we estimate the Pareto curve by connecting the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
-We then estimated the overall average Pareto curve, as shown in Figure (REF), by averaging the z-transformed points (1), (2), (3).
-
-\paragraph{Per-Language Results}
 In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.\footnote{Note that in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
 Due to use of stochastic gradient descent in optimization, there is some amount of noise in the optimization process.}
+
+We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section~\ref{sec:lambda}), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
+
+Within each language, we estimate the Pareto curve based on the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
+%We then estimated the overall average Pareto curve, as shown in Figure (REF), by averaging the z-transformed points (1), (2), (3).
+
+Figure (REF) in the main paper shows the average of these per-language plots. \textcolor{red}{unify pareto curve computation in main paper}
 
@@
 
-\textcolor{red}{also provide p value, CI for how many baselines are better in either of the two dimensions}
+In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
+We also conducted the analysis using a Binomial test (one-sided), testing whether the real gramar is more efficient than the \emph{median} of random grammars, avoiding any distributional assumption on the random grammars.
+As before, we used Hochberg's step-up procedure (Note that the tests for different languages are independent, as different random grammars are evaluated for each language), with $\alpha=0.05$.
+In this analysis, real grammars improved in parseability for 80\% of languages, in predictability for 69\% of languages, and in either of both in 92\% of languages ($p<0.05$, with Bonferroni correction).
+In Table~\ref{fig:pareto-per-lang-stats}, we provide per-language results for the $t$-tests and binomial tests.
+
+\begin{table}
+\centering
+\small{
+\begin{tabular}{l||ll|lll|lll}
+Language & Pred. (t) & Parse. (t) & \multicolumn{3}{|c|}{Pred. (Binomial)} & \multicolumn{3}{|c|}{Parseab. (Binomial)} \\ 
+&  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
+\input{../results/plane/per-language-statistics.tex}
+\end{tabular}
+}
+\caption{Per-language results in Study 1.
+For each language, we show the following: (1) $p$-values obtained from a one-sided $t$-test, for the null that the mean predictability/parseability of random grammars is at least as high as that of the real grammar.
+(2) Results from one-sided binomial tests, for the null that the the real grammar is better than at most $50 \%$ of random grammars.
+In addition to the $p$-value, we report point estimate and 95\% confidence interval for the fraction of random grammars that have values below real grammars.}\label{fig:pareto-per-lang-stats}
+\end{table}
 
@@
 
-We show results on the level of individual languages in Figure~\ref{fig:per-lang}.
+We show predictions for the eight correlations on the level of individual languages in Figure~\ref{fig:per-lang}.
 
@@
 
-\section{Possible values of $\lambda$}\label{sec:lambda}
-
-In the efficiency objective~(\ref{eq:efficiency-again}), the value of $\lambda$ is constrained to be in $[0,1)$.
-This means, surprisal must be weighted less strongly than parsability.
-
-Greater values of $\lambda$ mathematically result in degenerate solutions.
-To show this, note that the following inequality always holds:
-\begin{equation}
-I[L(T); T] \leq H[L(T)].
-\end{equation}
-Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
-\begin{equation}
-    J_T(L) = I[L(T); T] - \lambda H[L(T)] \leq 0.
-\end{equation}
-and it takes the maximally possible value zero if $L(T)$ maps all trees to a single utterance, in which case both $I[L(T); T]$ and $H[L(T)]$ are zero.
-This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
-While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
-In conclusion, $\lambda$ is constrained to be in $[0,1)$, which values closer to $1$ placing similar weights on both predictability and parsability, whereas values closer to $0$ diminish the role of predictability.
-
-In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
-The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
-%We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations, see Table REF.
-We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (Table~\ref{tab:all-predictions-1}).
-%Indeed, the same holds for all of the 38 syntactic relations in UD  (Table~\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})
-%We additionally report results with $\lambda = 0$, i.e., placing all weight on parseability.
-That is, at $\lambda$ close to its maximal value, the predictions of optimizing the two scoring functions individually add up to the predictions of efficiency optimization.
-Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}).
-We also verified the robustness of our results by reporting results with $\lambda=1$ in Table REF>
-
-%\section{Joint optimization of grammar, parser, and language model parameters}
-\section{Creating Optimized Grammars}
-
-In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
-%Our method will be applicable more generally to the task of optimizing grammars for a broad class of objective functions.
-We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
-More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
-This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.
-
-This method addresses a major challenge noted in previous work optimizing grammars, namely that the predictability (and parseability) of an individual sentence depends on the entire distribution of the language.
-Previously, \citep{gildea2015human} optimized grammars for dependency length and trigram surprisal using a simple hill-climbing method on the grammar parameters, which required reestimating the trigram surprisal model in every iteration.
-Such a method would be computationally prohibitive for efficiency optimization, as it would require reestimating the neural network models after every change to the grammar, which would amount to reestimating them hundreds or thousands of times per grammar.
-Our method, by allowing for the use of stochastic gradient descent, addresses this challenge, as we describe in Section~\ref{sec:optim-eff}.
-
-\subsection{Differentiable Ordering Grammars}\label{sec:diff-gramm}
-
-We extended the parameter space of grammars by continuously interpolating between grammars, making efficiency a \emph{differentiable} function of grammar parameters.
-The parameters of such a \key{differentiable word order grammar} are as follows. 
-For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
-Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
-Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
-184) without replacement.
-
-If $a_\tau \in \{0, 1\}$, and the distances between values of $b_\tau$ (for different $\tau$) become very large, such a differentiable grammar becomes deterministic, assigning almost full probability to exactly one ordering for each syntactic structure.
-In this case, the grammar can be converted into an equivalent grammar of the form described in Materials and Methods, by extracting a single parameter in $[-1, 1]$ for each relation $\tau$.
-
-\subsection{Extracting Grammars from Datasets}
-We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
-To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
-We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.
-
-\subsection{Optimizing Grammars for Efficiency}\label{sec:optim-eff}
-
-In this section, we describe how we optimized grammar parameters for efficiency.
-A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
-More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $\mathbf{w}$.
-In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $p_{\mathcal{L}_\theta}(u|t)$ concentrates on a single ordering $u$.
-
-Recall the definition of efficiency
-\begin{equation}\label{eq:efficiency}
-	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
-\end{equation}
-where 
-\begin{equation}
-	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
-\end{equation}
-\begin{equation}
-	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
-\end{equation}
-where $t \sim \mathcal{T}$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
-
-These quatities are estimated using two neural models, as described in Section X:
-A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.
-The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is
-\begin{equation}
-	R_{Pars}^{\phi}(u,t) =  \log \frac{p_\phi(t|u)}{p(t)}
-\end{equation}
-%and we can write
-%\begin{equation}
-%	R_{Pars}^\phi := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
-%\end{equation}
-%
-A \key{language model}, with some parameters $\psi$, calculates the word-by-word surprisal of an utterance:
-\begin{equation}
-	R_{Pred}^{\psi}(u) = - \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})
-\end{equation}
-Using this, we can rewrite Efficiency~(\ref{eq:efficiency}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
-\begin{equation}
-	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi} := \max_{\phi,\psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
-\end{equation}
-%
-%The parsers and language models appropriate to a tree distribution $\mathcal{T}$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
-%\begin{equation}
-%\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-%\end{equation}
-%\begin{equation}
-%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-%\end{equation}
-%
-%Using these models, we can rewrite 
-%\begin{equation}
-%R_{Pars}^{\psi(\theta)}= \E_t \E_{u \sim p_\theta(u|t)} R_{Pars}^{\psi(\theta)}(u,t) 
-%\end{equation}
-%where $R_{Pars}(u,t) = \log \frac{p_\phi(t|u)}{p(t)}$, and $p_\theta(u|t)$ is the probability assigned to sentence $u$ by the distribution $\mathcal{L}_\theta(t)$ defined by the differentiable word order grammar $\mathcal{L}_\theta$,
-%and
-%\begin{equation}
-%R_{Pred} = \E_t \E_{u \sim p_\theta(u|t)} R_{Pred}(u)
-%\end{equation}
-%where $R_{Pred}(u) = \log p_\psi(u)$.
-%Therefore, we can rewrite efficiency as
-%\begin{equation}\label{eq:efficiency}
-%R_{\textit{Eff}}^\theta := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi(\theta)}(u,t) + \lambda R_{Pred}^{\psi(\theta)}(u)\right]
-%\end{equation}
-%Directly applying gradient descent to this expression is hard due to the complicated dependence of $\phi(\theta), \psi(\theta)$ on $\theta$.
-%
-%However, we can rewrite
-%\begin{equation}\label{eq:efficiency}
-%\argmax_\theta\	R_{\textit{Eff}}^\theta 
-%\end{equation}
-%equivalently into
-In order to find an optimal grammar $\theta$, we thus need to solve
-\begin{equation}\label{eq:efficiency}
-\argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} %= \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
-\end{equation}
-Importantly, $R_{\textit{Eff}}^{\theta, \phi, \psi}$ is differentiable in $\theta, \phi, \psi$: %, and we can apply stochastic gradient descent to carry out this optimization.
-\begin{align}
-\partial_\theta R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)} \left[  \left[\partial_\theta \log p_\theta(u|t)\right] \cdot    \left(R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right) \right] \label{eq:dtheta}\\
-\partial_\phi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\partial_\phi R_{Pars}^{\phi}(u,t)\right] \\
-\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}
-\end{align}
-where \ref{eq:dtheta} is derived using the \emph{score-function} or \emph{REINFORCE} trick~\cite{williams1992simple}.
-Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures (CITE).
-
-We can therefore apply stochastic gradient descent to jointly optimize $\theta, \phi, \psi$:
-In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim p_{\theta}(\cdot|t)$.
-Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
-do a gradient descent step using the estimator given by the expressions in the square brackets in \ref{eq:dtheta}-\ref{eq:dpsi}.
-
-Optimizing for only parseability (or predictability) is very similar -- in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.
-
-At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
-Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.
-
-See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates used in this optimization scheme.
-
-%
-%%(analogous for $R_{Pars}$ instead of $R_{Pred}$)
-%\begin{equation}\label{eq:estimator}
-%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
-%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
-%\end{equation}
-%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
-%for the gradient
-%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
-%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
-%
-
-%
-%
-%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
-Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
-%\begin{equation}
-%\label{eq:exp-over-utts}
-%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
-%\end{equation}
-%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
-In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
-%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
-%
-%
-%
-%We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
-%\begin{equation*}
-%R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
-%\end{equation*}
-%where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
-%\begin{equation*}
-%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-%\end{equation*}
-%
-%
-%We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
-Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
-%\begin{align}
-%\nonumber
-%R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
-%\nonumber
-%&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
-%\end{align}
-%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
-%\begin{equation*}
-%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-%\end{equation*}
-%
-%For efficiency, we linearly combine the two terms:
-%\begin{equation}
-%\label{eq:combined-score}
-%R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
-%\end{equation}
-%
-%
-%%\begin{equation}
-%%    \label{eq:depl}
-%%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
-%%\end{equation}
-%%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
-%
-%For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
-The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
-The optimization problems come out to:
-%\begin{align}
-%%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
-%\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
-%\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
-%\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
-%\end{align}
-%Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
-%
-%
-%%The case of predictability, parseability, and efficiency are more challenging, as 
-%%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
-%We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
-%Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
-%
-%We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section (REF)).
-%
-%In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
-%Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
-%do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
-%\begin{equation}\label{eq:estimator}
-%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
-%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
-%\end{equation}
-%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
-%for the gradient
-%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
-%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
-%
-%See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
-%
-
-%The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
-A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
-%We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
-
-%In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
-%and $\ell_2$ regularization.
-
-%A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
-%Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
-%We thus do not evaluate this objective function here.
-
-%We extract grammars for the actual languages using a Bayesian data analysis method.
-%We specify a simple Bayesian prior over differentiable ordering grammars:
-%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
-%for each $\tau$.
-%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
-%We extract
-
-\section{Neural Network Architectures}\label{sec:neural-architectures}
-
-In this section, we describe the details of the neural network architectures.
-Choices follow standard practice and machine learning and were fixed before evaluating word order properties, and the efficiency of real languages.
-
-%\paragraph{Locality}
-%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
-
-\paragraph{Estimating Predictability}
-We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
-Given the small size of the corpora, this limit is only attained only for few languages.
-In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
-The model predicts both the next word and its POS tags in each step.
-Using POS tags is intended to prevent overfitting on small corpora.
-
-\paragraph{Estimating Parseability}
-We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}.
-This architecture is remarkably simple: the words of a sentence are encoded into context-sensitive embeddings using bidirectional LSTMs, then a classifier is trained to predict the head for each work.
-The classifier works by calculating a score for every pair of word embeddings $(w_i, w_j)$, indicating the likelihood that the $j$th word is the head of the $i$th word.
-This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
-
-To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags.
-Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged; parsing from POS tags prevents early overfitting.
-This decision was made before evaluating word order properties.
-
-\paragraph{Hyperparameters}
-
-%\textbf{TODO study 1 only has 1 set of parameters}
-
-Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
-For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
-These parameters are shown in Table~\ref{tab:hyperparameters}.
-Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (efficiency, predictability, parseability) on the language.
-We then used this setting for creating optimized word order grammars.
-
-%\textcolor{red}{TODO random seeds for efficiency}
-
-All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
-We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{c|cccc}
-    Entropy penalty     &  0.0001, 0.001 \\
-    Learning Rate (Word order grammar)     & 5e-6, 1e-5, 2e-5, 5e-5 \\
-    Momentum (Word order grammar) & 0.8, 0,9 \\
-    Dropout (Language model) & 0.0, 0.3, 0.5 \\
-
-    \end{tabular}
-    \caption{Selected Hyperparameters}
-    \label{tab:hyperparameters}
-\end{table}
-
-\paragraph{Improved Unbiased Gradient Estimator}
-We employ two common variance reduction methods to improve the estimator~(\ref{eq:dtheta}), while keeping it unbiased.
-%For dependency length and 
-% (and its head, in the case of dependency length)
-For predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word.
-We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:dtheta}) with lower variance.
-Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple} (the word's surprisal in the case of predictability, and the negative log-probability of the correct head and relation label in the case of parseability).
-These two methods reduce the variance of the estimator and thereby increase the speed of optimization and reduce training time, without biasing the results.
-For numerical stability, we represent $a_\tau \in [0,1]$ via its logit $\in \mathbb{R}$.
-Furthermore, to encourage exploration of the parameter space, we add an entropy regularization term \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$.
-The weight of the entropy regularization was chosen together with the other hyperparameters.
-%, adding the negative entropy of the $Bernoulli(a_\tau)$ to the loss $R$. %, weighted with $\alpha$.
-
-These techniques for improving (\ref{eq:dtheta}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
-
-\paragraph{Learning Rate and Stopping Criterion}
-We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
-For the language model parameters $\phi$, we use plain Stochastic Gradient Descent without momentum, as recommended by \citet{merity2018regularizing}. 
-For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
-The learning rates and other optimization hyperparameters were determined together with the other hyperparameters.
-
-For gramar optimization, we apply Early Stopping~\citep{prechelt1998early} as follows.
-We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
-This is a standard technique for optimizing neural models, and helps prevent overfitting to the specific corpora used.
-
-The choice of optimization methods and the stopping criterion were fixed before we investigated word order correlations.
-
-\paragraph{Optimized Grammars}
-Recall that, for each language, we created 8 optimized languages for each optimization criterion.
-We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
-This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.
-
-For efficiency optimization, for each grammar, we ran efficiency optimization with three different random seeds, selecting among these the seed that yielded the best overall efficiency value.
-This helps control for variation across random seeds for the stochastic gradient descent optimization method.
-
- % TODO we could use some language at the top of this section on the general guiding principles for all these decisions.
-Were these implementation decisions made before or after looking at the word order universals results?
-If they were made before, we should highlight that.
-
-\section{Robustness to different language models and parsers}
-
-Here we take up the question of the extent to which our results are dependent on the particular parser and language model used in the optimization process.
-We want to know: when we optimize a word order grammar for efficiency, have we produced a language which is highly efficient \emph{in general}, or one which is highly efficient \emph{for a specific parser}?
-We wish to argue that natural language syntax is optimized for efficiency in general, meaning that syntactic trees are highly recoverable from word orders in principle.
-If it turns out that our optimized languages are only optimal for a certain parser from the NLP literature, then we run the risk of circularity: it may be that the reason this parser was successful in the NLP literature was because it implicitly encoded word order universals in its inductive biases, and thus it would be no surprise that languages which are optimized for parseability also show those universals.
-
-In this connection, we note that the parser and language model architectures we use are highly generic, and do not encode any obvious bias toward natural-language-like word orders.
-The LSTM language model is a generic model of sequence data which is also been used to model financial time series \citep{sirignano2018universal} and purely theoretical chaotic dynamical systems \citep{ogunmolu2016nonlinear}; the neural graph-based parser is simply solving a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
-Nevertheless, it may be the case that a bias toward word order universals is somehow encoded implicitly in the hyperparameters and architectures of these models.
-
-Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers.
-These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.
-
-\subsection{CKY Parsers}
-
-We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of Collins 1997 (Model 1).
-In our PCFGs, each head independently generates a set of left and right dependents.
-We formulate this as a PCFG where each rule has the form:
+\subsection{Other Experiments}
+In Table~\ref{table:corr-resu} we report the results of the simulations we did before the main experiment reported in the main paper (left and center), together with their preregistrations, besides the results from the main experiment (right).
+These experiments all had the same setup described in Section (REF), which was fixed before starting simulations; differences are that (1) the first simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
+Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
+\textcolor{red}{TODO also the aggregate posterior}
+
+\begin{table}[hbt!]
 	\begin{center}
-	POS$_H$ $\rightarrow$ POS$_H$ POS$_D$
-\end{center}
-for head-initial structures, and
-\begin{center}
-	POS$_H$ $\rightarrow$ POS$_D$ POS$_H$
-\end{center}
-for head-final structures, where each symbol is a POS tag.
-Thus, POS tags act both as terminals and as nonterminals.
-
-We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
-For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm.
-
-This parsing strategy is very different from the neural graph-based parser:
-While the graph-based parser solves a minimum spanning tree problem, the CKY algorithm uses dynamic programming to compute the exact probabilities of trees given a sentence, as specified by the generative model encoded in the PCFG.
-Second, while the graph-based neural parser uses machine learning to induce syntactic knowledge from data, the CKY parser performs exact probabilistic inference.
-%In this sense, the CKY algorithm does not have any architectural biases in itself.
-%On the other hand, the PCFG makes severely simplifying independence assumptions, compared to the universal approximation capabilities of neural network-based systems.
-
-We used the CKY algorithm to compute the parsing loss $H[T|L_\theta(T)]$ on the validation partition of the English UD corpora, for random and optimized ordering grammars.
-Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
-% TODO get UAS for compatibility with other parts of SI
-%We then compared parsability of random and optimized word order grammars on English data.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
-	\caption{Parsability loss $H[T|L_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).}
-    \label{fig:cky-parser}
-\end{figure}
-
-\subsection{Distorted graph-based parsers}
-
-In this section, we address the idea that the graph-based parser might have a built-in bias toward certain kinds of orderings, and the question whether this might be responsible for our findings.
-In particular, we address the idea that the graph-based parser might have a bias toward parses involving short dependencies, which we call a \key{locality bias}. 
-We address this by changing the order in which the parser sees words, in such a way that the distance between words in the input is not indicative of syntactic relations.
-
-%In particular, this allows us to show tha
-
-%Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization.
-Therefore, we address here the idea that the graph-based parser might have a built-in bias toward parses involving short dependencies, which we call a \key{locality bias}.
-%A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words.
-Suppose that a parser $P$ has a bias toward positing a dependency between a word and the immediate previous word.
-
-\paragraph{Even--odd order.} A sequence of $n$ words originally ordered as $w_1 w_2 w_3 w_4 \cdots w_n$ is reordered by separating the even and odd indices: $w_2 w_4 w_6 \cdots w_{n-1} w_1 w_3 w_5 \cdots w_n$ (assuming $n$ odd).
-Therefore all words that are adjacent in the original order will be separated by a distance of $\approx n/2$ in the distorted order, while all words of distance 2 in the original order will become adjacent.
-
-\paragraph{Interleaving order.} In interleaving ordering, a sequence originally ordered as $w_1 w_2 w_3 \cdots w_n$ is split in half at the middle (index $m=\ceiling{n/2}$), and the two resulting sequences are interleaved, yielding $w_1 w_m w_2 w_{m+1} w_3 w_{m+3} \cdots w_n$.
-Thus all words that were originally adjacent will have distance 2 in the distorted order, with the intervening word coming from a very distant part of the sentence.
-
-\paragraph{Inwards order.} A sequence originally ordered as $w_1 w_2 w_3 \cdots w_{n-1} w_n$ is ordered from the edges of the string inwards, as $w_1 w_n w_2 w_{n-1} \cdots w_{\ceiling{n/2}}$.
-This corresponds to folding the string in on itself once, or equivalently, splitting the sequence in half at the middle, then interleaving the two resulting sequences after reversing the second one.
-The result is that the most non-local possible dependencies in the original order become the most local dependencies in the distorted order.
-
-\paragraph{Sorted order.} A sequence is reordered by sorting by POS tags, and randomizing the order within each block of identical POS tags.
-To each word, we then add a symbol encoding the original position in the sequence.
-For instance
-\begin{center}
-PRON VERB PRON
-\end{center}
-may be reordered as
-\begin{center}
-PRON 1 PRON 3 VERB 2
-\end{center}
-or
-\begin{center}
-PRON 3 PRON 1 VERB 2
-\end{center}
-The numbers are provided to the parser as atomic symbols from a vocabulary ranging from 1 to 200; numbers greater than 200 (which may occur in extremely long sentences) are replaced by an out-of-range token.
-
-The result is that distance between words in the input is not indicative at all of the presence of absence of syntactic relations between them.
-
-\paragraph{Experiments}
-Using English and Japanese data, we trained parsers for the ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
-Resulting parsing accuracy scores are shown in Figure~\ref{fig:distorted-parser}.
-In all settings, the language optimized for efficiency achieved higher parsing accuracy than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
-%Note that absolute values of parsing accuracy vary between the orderings, being higher for the actual orderings than for distorted orderings, in which trees become nonprojective.
-
-\begin{figure}
-    \centering
-    English
-
-    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-english.pdf}
-
-    Japanese
-
-    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
-	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').}
-    \label{fig:distorted-parser}
-\end{figure}
-
-\subsection{n-gram language models}
-
-We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
-In previous work, such as \cite{gildea2015human}, predictability has often been measured using n-gram models.
-
-Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
-Specifically, we constructed bigram models with Kneser-Ney smoothing~\cite{kneser1995improved, chen1999empirical}.
-A bigram model predicts each word taking only the previous word into account.
-This contrasts with LSTMs, which take the entire context into consideration.
-Thus, bigram models and LSTMs stand on opposing ends of a spectrum of language models taking more and more aspects of the context into account.
-
-We estimated language models on the training partitions, and used the validation partitions to estimate surprisal.
-We conducted this for the ten random and the best optimized ordering grammars on English and Japanese data.
-Results (Figure~\ref{fig:bigrams}) show that languages optimized for efficiency are optimal for a bigram language model.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.4]{../results/bigrams/bigrams.pdf} 
-	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).}
-    \label{fig:bigrams}
-\end{figure}
-
-%\subsection{Discussion}
-%
-%It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.
-
-\section{Effects of data sparsity}
-
-Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
-If the difference between random and real grammars is due to data sparsity, we expect that it decreases as the amount of training data is increased.
-If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference persists as training data is increased.
-
-We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
-We cosidered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
-For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
-Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.
-
-The resulting curves are shown in Figure~\ref{fig:learning-czech}.
-%A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
-A gap in parsing loss of about 0.2 nats appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
-
-\begin{figure}[ht]
-    \centering
-%    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech.pdf} 
-    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech-logloss.pdf}
-
-        \caption{Parsing loss ($H[T|\mathcal{L}_\theta(T)]$) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
-    \label{fig:learning-czech}
-\end{figure}
-
-\section{Comparing Extracted Grammars to Linguistic Judgments}
-
-% user@user-X510UAR:~/grammar-optim/results/grammars$ git add comparison-table.tex
-
-We validated our grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
-Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS.
-In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb-object relation.
-
-WALS has some data for all but seven languages.\footnote{Note that Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
-In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}
-WALS has data for 74\% of the entries, and lists a dominant order for 91\% of these.
-The grammars we extracted from the corpora agree with WALS in 96~\% out of those cases where WALS records a dominant word order.
-
-\begin{table}[ht]
 \small{
-\begin{center}
-\begin{tabular}{l||ll|ll|ll|ll|ll|ll|llllll}
-%	   for(dep in c("obj", "lifted_mark", "acl", "nmod", "obl")) {
-		   Language 
-		   &	\multicolumn{2}{|c|}{Objects} 
-		   &	\multicolumn{2}{|c|}{Adpositions} 
-		   &	\multicolumn{2}{|c|}{Compl.} 
-		   &	\multicolumn{2}{|c|}{Rel.Cl.} 
-		   &	\multicolumn{2}{|c|}{Genitive} 
-		   &	\multicolumn{2}{|c|}{PP}  \\ \hline\hline
-	\input{../results/grammars/comparison-table.tex}
-\end{tabular}
-\end{center}
-}
-\caption{Comparing grammars extracted from databases to linguistic judgments in the World Atlas of Linguistic Structures.
-For each of the six syntactic relation, the first column provides the ordered coded in the extracted grammar; the second column provides the order coded in WALS (DH for dependent-head, HD for head-dependent order).
-`?' indicates that WALS has no data.
-$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur at similar rates in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
-Finally, `--' indicates that the relation does not occur in the corpus.
-}\label{tab:grammars-wals}
-\end{table}
-
-\section{Efficiency and Prior Formalizations}
-
-%The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
-%Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication.
-
-%Recall our formalization of efficiency:
-%\begin{equation}
-%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
-%\end{equation}
-%\begin{equation}
-%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
-%\end{equation}
-%\begin{equation}\label{eq:efficiency}
-%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
-%\end{equation}
-
-Here we discuss how our formalization of efficiency relates to other information-theoretic work on language.
-
-Our objective function is identical to the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems, with syntactic structures taking the role of underlying meanings that the speaker wants to convey.
-This function is also equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
-
-Very similar metrics have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
-Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.
-
-All these accounts formalize the amount of information provided as the \key{mutual information} between the forms that a listener receives, and the meanings that the speaker intends to communicate to the speaker.
-Most previous studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} are concerned with the meanings of individual words; they thus take the forms to be individual words, which refer to individual objects and their properties.
-We instead study how words are composed into sentences.
-In our work, the underlying meanings are the syntactic structures; the forms are the linearized sentences communicated by the speaker to the listener.
-By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
-The syntactic structure (including the labels on the arcs) specifies how the meanings of words are combined.
-Therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
-
-%These information-theoretic accounts differ in the formalizations of complexity and diversity:
-Like our model, \cite{ferreri2003least,ferrericancho2007global,futrell2017memory} model the complexity as the \key{entropy} of the distribution of forms, which arises naturally from the role played by surprisal in determining human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
-This model is closely related to the Information Bottleneck approach \cite{zaslavsky2018efficient, zaslavsky2019semantic}, which models complexity as the mutual information between forms (e.g., words) and their referents (e.g., individual objects).
-This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
-As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure.
-
-A few models have used other measures that are not direct information-theoretic quantities.
-In \cite{regier2007color, xu2014numeral, xu2016historical}, the cost function is the number of different forms; in \cite{kemp2012kinship} it is the complexity of the concepts encoded.
-The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
-Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the information bottleneck formalism, bringing them even closer to our formalization.
-
-In addition to these models, which quantify the efficiency of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context in models of pragmatic reasoning in reference games \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
-In line with the other models, these assume that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely}.
-
-%In terms of information theory \cite{shannon1948mathematical}, this corresponds to the notion of the rate-distortion tradeoff.
-%In classical rate-distortion theory, the cost of a code is its rate, the mutual information between codes and underlying messages (as in the Information Bottleneck model).
-% classical models of distortion are the Hamming distortion and 
-%TODO also mention rate-distortion theory
-
-%- our function
-
-%similar: RSA, there it's utterance-level
-
-%- number of different forms 
-%
-%- complexity of the mapping 
-%
-%- MI between
-%
-%
-%
-%We model the first term as the 
-% degree to which listeners can reconstruct dependency parse structures from an utterance, i.e., the \key{parseability} if the language.
-This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
-%\begin{equation}
-%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
-%\end{equation}
-%where the sum runs over all possible pairs of sentences $l$ and syntactic structures $t$ in the language.
-%%This quantity describes the degree to which syntactic structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
-%
-%
-%
-%We formalize the complexity of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
-%In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
-%Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
-%%and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
-%
-%In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
-%\begin{equation}
-%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
-%\end{equation}
-%where the sum runs over all possible sentences $u$ that belong to the language.
-%%In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
-%
-%
-%
-%The efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
-%\begin{equation}\label{eq:efficiency}
-%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
-%\end{equation}
-%with an interpolation weight $\lambda \in [0,1)$.
-%In all results from Study 2 reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
-%See SI appendix section 5 for mathematical discussion of $\lambda$, and robustness to other choices.
-%
-%
-
-\section{Relation to the Information Bottleneck}
-
-Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
-\begin{equation}
-    \label{eq:efficiency-again}
-    J_T(L) = I[L(T); T] - \lambda H[L(T)].
-\end{equation}
-This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}.
-In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information.
-The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$.
-Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
-\begin{equation}
-    \label{eq:ib}
-    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
-\end{equation}
-In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
-
-In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$.
-To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$.
-We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized.
-Then our objective is parallel to Eq.~\ref{eq:ib}:
-\begin{equation}
-    \label{eq:efficiency-kinda-ib}
-    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
-\end{equation}
-Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$.
-Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
-\begin{equation*}
-    J_T(L) = I[U;T] - \lambda H[U].
-\end{equation*}
-
-Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
-
-\section{Relation to DLM}
-
-DLM turns out to be predicted by our model
-
-\section{Other (Preregistered) Experiments}
-
-Here we report the results of prior preregistered simulation runs.
-
-\begin{table}[ht]
-	\begin{center}
 \begin{tabular}{|c|ccc|c}
@@
 	%&
-& $\lambda=1.0$ & $\lambda=0.9$ & $\lambda=0.9$ \\ \hline \hline 
+& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table REF.} 
+& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
+& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} \\ \hline \hline 
 	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}		  
@@
 \end{tabular}
+}
 	\end{center}
-\caption{Previous experiment with $\lambda=0.9$ (Preregistration)}\label{table:corr-resu}
+\caption{\textcolor{red}{probably should also report parseability results as lambda=0 here}Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
 \end{table}
 
-\section{Languages and Corpora}
+%\footnotetext{}
+%  final experiment
+%  0.9
+%
+
+%\section{Joint optimization of grammar, parser, and language model parameters}
+\section{Creating Optimized Grammars}
+
+In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
+%Our method will be applicable more generally to the task of optimizing grammars for a broad class of objective functions.
+We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
+More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
+This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.
+
+This method addresses a major challenge noted in previous work optimizing grammars, namely that the predictability (and parseability) of an individual sentence depends on the entire distribution of the language.
+Previously, \citet{gildea2015human} optimized grammars for dependency length and trigram surprisal using a simple hill-climbing method on the grammar parameters, which required reestimating the trigram surprisal model in every iteration.
+Such a method would be computationally prohibitive for efficiency optimization, as it would require reestimating the neural network models after every change to the grammar, which would amount to reestimating them hundreds or thousands of times per grammar.
+Our method, by allowing for the use of stochastic gradient descent, addresses this challenge, as we describe in Section~\ref{sec:optim-eff}.
+
+\subsection{Differentiable Ordering Grammars}\label{sec:diff-gramm}
+
+We extended the parameter space of grammars by continuously interpolating between grammars, making efficiency a \emph{differentiable} function of grammar parameters.
+The parameters of such a \key{differentiable word order grammar} are as follows. 
+For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
+Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
+Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
+184) without replacement.
+
+If $a_\tau \in \{0, 1\}$, and the distances between values of $b_\tau$ (for different $\tau$) become very large, such a differentiable grammar becomes deterministic, assigning almost full probability to exactly one ordering for each syntactic structure.
+In this case, the grammar can be converted into an equivalent grammar of the form described in Materials and Methods, by extracting a single parameter in $[-1, 1]$ for each relation $\tau$.
+
+\subsection{Extracting Grammars from Datasets}
+We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
+To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
+We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.
+
+\subsection{Optimizing Grammars for Efficiency}\label{sec:optim-eff}
+
+In this section, we describe how we optimized grammar parameters for efficiency.
+A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
+More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $\mathbf{w}$.
+In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $p_{\mathcal{L}_\theta}(u|t)$ concentrates on a single ordering $u$.
+
+Recall the definition of efficiency
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+\end{equation}
+where 
+\begin{equation}
+	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
+\end{equation}
+\begin{equation}
+	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+\end{equation}
+where $t \sim \mathcal{T}$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
+
+These quatities are estimated using two neural models, as described in Section X:
+A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.
+The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is
+\begin{equation}
+	R_{Pars}^{\phi}(u,t) =  \log \frac{p_\phi(t|u)}{p(t)}
+\end{equation}
+%and we can write
+%\begin{equation}
+%	R_{Pars}^\phi := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
+%\end{equation}
+%
+A \key{language model}, with some parameters $\psi$, calculates the word-by-word surprisal of an utterance:
+\begin{equation}
+	R_{Pred}^{\psi}(u) = - \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})
+\end{equation}
+Using this, we can rewrite Efficiency~(\ref{eq:efficiency}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
+\begin{equation}
+	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi} := \max_{\phi,\psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\end{equation}
+%
+%The parsers and language models appropriate to a tree distribution $\mathcal{T}$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
+%\begin{equation}
+%\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation}
+%\begin{equation}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation}
+%
+%Using these models, we can rewrite 
+%\begin{equation}
+%R_{Pars}^{\psi(\theta)}= \E_t \E_{u \sim p_\theta(u|t)} R_{Pars}^{\psi(\theta)}(u,t) 
+%\end{equation}
+%where $R_{Pars}(u,t) = \log \frac{p_\phi(t|u)}{p(t)}$, and $p_\theta(u|t)$ is the probability assigned to sentence $u$ by the distribution $\mathcal{L}_\theta(t)$ defined by the differentiable word order grammar $\mathcal{L}_\theta$,
+%and
+%\begin{equation}
+%R_{Pred} = \E_t \E_{u \sim p_\theta(u|t)} R_{Pred}(u)
+%\end{equation}
+%where $R_{Pred}(u) = \log p_\psi(u)$.
+%Therefore, we can rewrite efficiency as
+%\begin{equation}\label{eq:efficiency}
+%R_{\textit{Eff}}^\theta := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi(\theta)}(u,t) + \lambda R_{Pred}^{\psi(\theta)}(u)\right]
+%\end{equation}
+%Directly applying gradient descent to this expression is hard due to the complicated dependence of $\phi(\theta), \psi(\theta)$ on $\theta$.
+%
+%However, we can rewrite
+%\begin{equation}\label{eq:efficiency}
+%\argmax_\theta\	R_{\textit{Eff}}^\theta 
+%\end{equation}
+%equivalently into
+In order to find an optimal grammar $\theta$, we thus need to solve
+\begin{equation}\label{eq:efficiency}
+\argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} %= \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\end{equation}
+Importantly, $R_{\textit{Eff}}^{\theta, \phi, \psi}$ is differentiable in $\theta, \phi, \psi$: %, and we can apply stochastic gradient descent to carry out this optimization.
+\begin{align}
+\partial_\theta R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)} \left[  \left[\partial_\theta \log p_\theta(u|t)\right] \cdot    \left(R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right) \right] \label{eq:dtheta}\\
+\partial_\phi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\partial_\phi R_{Pars}^{\phi}(u,t)\right] \\
+\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}
+\end{align}
+where \ref{eq:dtheta} is derived using the \emph{score-function} or \emph{REINFORCE} trick~\cite{williams1992simple}.
+Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures (CITE).
+
+We can therefore apply stochastic gradient descent to jointly optimize $\theta, \phi, \psi$:
+In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim p_{\theta}(\cdot|t)$.
+Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
+do a gradient descent step using the estimator given by the expressions in the square brackets in \ref{eq:dtheta}-\ref{eq:dpsi}.
+
+Optimizing for only parseability (or predictability) is very similar -- in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.
+
+At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
+Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.
+
+See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates used in this optimization scheme.
+
+%
+%%(analogous for $R_{Pars}$ instead of $R_{Pred}$)
+%\begin{equation}\label{eq:estimator}
+%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
+%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
+%\end{equation}
+%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
+%for the gradient
+%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
+%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
+%
+
+%
+%
+%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
+Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
+%\begin{equation}
+%\label{eq:exp-over-utts}
+%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
+%\end{equation}
+%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
+In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
+%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
+%
+%
+%
+%We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
+%\begin{equation*}
+%R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
+%\end{equation*}
+%where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
+%\begin{equation*}
+%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation*}
+%
+%
+%We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
+Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
+%\begin{align}
+%\nonumber
+%R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
+%\nonumber
+%&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
+%\end{align}
+%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
+%\begin{equation*}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation*}
+%
+%For efficiency, we linearly combine the two terms:
+%\begin{equation}
+%\label{eq:combined-score}
+%R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
+%\end{equation}
+%
+%
+%%\begin{equation}
+%%    \label{eq:depl}
+%%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+%%\end{equation}
+%%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
+%
+%For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
+The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
+The optimization problems come out to:
+%\begin{align}
+%%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
+%\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
+%\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
+%\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
+%\end{align}
+%Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
+%
+%
+%%The case of predictability, parseability, and efficiency are more challenging, as 
+%%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
+%We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
+%Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
+%
+%We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section (REF)).
+%
+%In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
+%Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
+%do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
+%\begin{equation}\label{eq:estimator}
+%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
+%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
+%\end{equation}
+%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
+%for the gradient
+%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
+%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
+%
+%See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
+%
+
+%The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
+A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
+%We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
+
+%In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
+%and $\ell_2$ regularization.
+
+%A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
+%Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
+%We thus do not evaluate this objective function here.
+
+%We extract grammars for the actual languages using a Bayesian data analysis method.
+%We specify a simple Bayesian prior over differentiable ordering grammars:
+%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
+%for each $\tau$.
+%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
+%We extract
+
+\section{Neural Network Architectures}\label{sec:neural-architectures}
+
+In this section, we describe the details of the neural network architectures.
+Choices follow standard practice and machine learning and were fixed before evaluating word order properties, and the efficiency of real languages.
+
+%\paragraph{Locality}
+%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
+
+\paragraph{Estimating Predictability}
+We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
+Given the small size of the corpora, this limit is only attained only for few languages.
+In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
+The model predicts both the next word and its POS tags in each step.
+Using POS tags is intended to prevent overfitting on small corpora.
+
+\paragraph{Estimating Parseability}
+We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}.
+This architecture is remarkably simple: the words of a sentence are encoded into context-sensitive embeddings using bidirectional LSTMs, then a classifier is trained to predict the head for each work.
+The classifier works by calculating a score for every pair of word embeddings $(w_i, w_j)$, indicating the likelihood that the $j$th word is the head of the $i$th word.
+This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
+
+To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags.
+Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged; parsing from POS tags prevents early overfitting.
+This decision was made before evaluating word order properties.
+
+\paragraph{Hyperparameters}
+
+%\textbf{TODO study 1 only has 1 set of parameters}
+
+Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
+For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
+These parameters are shown in Table~\ref{tab:hyperparameters}.
+Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (efficiency, predictability, parseability) on the language.
+We then used this setting for creating optimized word order grammars.
+
+%\textcolor{red}{TODO random seeds for efficiency}
+
+All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
+We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{|l|l|l|}
+\hline
+\multirow{2}{*}{Optimization}&    Learning Rate     & 5e-6, 1e-5, 2e-5, 5e-5 \\
+&    Momentum & 0.8, 0.9 \\ \hline
+\multirow{6}{*}{Language Model} &    Learning Rate  & 0.5, 0.1, 0.2 \\
+&    Dropout Rate& 0.0, 0.3, 0.5 \\
+& Embedding Size (Words) & 50 \\
+& Embedding Size (POS) & 20 \\
+&    LSTM Layers  & 2 \\
+&    LSTM Dimensions  & 128 \\
+ \hline
+\multirow{5}{*}{Parser}&    Learning Rate  & 0.001 \\
+&    Dropout Rate  & 0.2 \\
+&    Embedding Size  & 100 \\
+&    LSTM Layers  & 2 \\
+&    LSTM Dimensions  & 200 \\
+\hline
+    \end{tabular}
+    \caption{Hyperparameters}
+    \label{tab:hyperparameters}
+\end{table}
+
+\paragraph{Improved Unbiased Gradient Estimator}
+We employ two common variance reduction methods to improve the estimator~(\ref{eq:dtheta}), while keeping it unbiased.
+%For dependency length and 
+% (and its head, in the case of dependency length)
+For predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word.
+We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:dtheta}) with lower variance.
+Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple} (the word's surprisal in the case of predictability, and the negative log-probability of the correct head and relation label in the case of parseability).
+These two methods reduce the variance of the estimator and thereby increase the speed of optimization and reduce training time, without biasing the results.
+For numerical stability, we represent $a_\tau \in [0,1]$ via its logit $\in \mathbb{R}$.
+Furthermore, to encourage exploration of the parameter space, we add an entropy regularization term \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$.
+The weight of the entropy regularization was chosen together with the other hyperparameters.\footnote{Explored values: 0.0001, 0.001.}
+%, adding the negative entropy of the $Bernoulli(a_\tau)$ to the loss $R$. %, weighted with $\alpha$.
+
+These techniques for improving (\ref{eq:dtheta}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
+
+\paragraph{Learning Rate and Stopping Criterion}
+We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
+For the language model parameters $\phi$, we use plain Stochastic Gradient Descent without momentum, as recommended by \citet{merity2018regularizing}. 
+For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
+The learning rates and other optimization hyperparameters were determined together with the other hyperparameters.
+
+For gramar optimization, we apply Early Stopping~\citep{prechelt1998early} as follows.
+We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
+This is a standard technique for optimizing neural models, and helps prevent overfitting to the specific corpora used.
+
+The choice of optimization methods and the stopping criterion were fixed before we investigated word order correlations.
+
+\paragraph{Optimized Grammars}
+Recall that, for each language, we created 8 optimized languages for each optimization criterion.
+We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
+This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.
+
+For efficiency optimization, for each grammar, we ran efficiency optimization with three different random seeds, selecting among these the seed that yielded the best overall efficiency value.
+This helps control for variation across random seeds for the stochastic gradient descent optimization method.
+
+ % TODO we could use some language at the top of this section on the general guiding principles for all these decisions.
+Were these implementation decisions made before or after looking at the word order universals results?
+If they were made before, we should highlight that.
+
+\section{Robustness to different language models and parsers}
+
+Here we take up the question of the extent to which our results are dependent on the particular parser and language model used in the optimization process.
+We want to know: when we optimize a word order grammar for efficiency, have we produced a language which is highly efficient \emph{in general}, or one which is highly efficient \emph{for a specific parser}?
+We wish to argue that natural language syntax is optimized for efficiency in general, meaning that syntactic trees are highly recoverable from word orders in principle.
+If it turns out that our optimized languages are only optimal for a certain parser from the NLP literature, then we run the risk of circularity: it may be that the reason this parser was successful in the NLP literature was because it implicitly encoded word order universals in its inductive biases, and thus it would be no surprise that languages which are optimized for parseability also show those universals.
+
+In this connection, we note that the parser and language model architectures we use are highly generic, and do not encode any obvious bias toward natural-language-like word orders.
+The LSTM language model is a generic model of sequence data which is also been used to model financial time series \citep{sirignano2018universal} and purely theoretical chaotic dynamical systems \citep{ogunmolu2016nonlinear}; the neural graph-based parser is simply solving a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
+Nevertheless, it may be the case that a bias toward word order universals is somehow encoded implicitly in the hyperparameters and architectures of these models.
+
+Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers.
+These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.
+
+\subsection{CKY Parsers}
+
+We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of \textcolor{red}{CITE Collins 1997 (Model 1).}
+In our PCFGs, each head independently generates a set of left and right dependents.
+We formulate this as a PCFG where each rule has the form:
+\begin{center}
+	POS$_H$ $\rightarrow$ POS$_H$ POS$_D$
+\end{center}
+for head-initial structures, and
+\begin{center}
+	POS$_H$ $\rightarrow$ POS$_D$ POS$_H$
+\end{center}
+for head-final structures, where each symbol is a POS tag.
+Thus, POS tags act both as terminals and as nonterminals.
+
+We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
+For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm \textcolor{red}{CITE}.
+
+This parsing strategy is very different from the neural graph-based parser:
+While the graph-based parser solves a minimum spanning tree problem, the CKY algorithm uses dynamic programming to compute the exact probabilities of trees given a sentence, as specified by the generative model encoded in the PCFG.
+Second, while the graph-based neural parser uses machine learning to induce syntactic knowledge from data, the CKY parser performs exact probabilistic inference.
+%In this sense, the CKY algorithm does not have any architectural biases in itself.
+%On the other hand, the PCFG makes severely simplifying independence assumptions, compared to the universal approximation capabilities of neural network-based systems.
+
+We used the CKY algorithm to compute the parsing loss $H[T|L_\theta(T)]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
+Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
+% TODO get UAS for compatibility with other parts of SI
+%We then compared parsability of random and optimized word order grammars on English data.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
+	\caption{Parsability loss $H[T|L_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).\textcolor{red}{make plot nicer}}
+    \label{fig:cky-parser}
+\end{figure}
+
+\subsection{Distorted graph-based parsers}
+
+In this section, we address the idea that the graph-based parser might have a built-in bias toward certain kinds of orderings, and the question whether this might be responsible for our findings.
+In particular, we address the idea that the graph-based parser might have a bias toward parses involving short dependencies, which we call a \key{locality bias}. 
+We address this by changing the order in which the parser sees words, in such a way that the distance between words in the input is not indicative of syntactic relations.
+
+%In particular, this allows us to show tha
+
+%Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization.
+Therefore, we address here the idea that the graph-based parser might have a built-in bias toward parses involving short dependencies, which we call a \key{locality bias}.
+%A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words.
+Suppose that a parser $P$ has a bias toward positing a dependency between a word and the immediate previous word.
+
+\paragraph{Even--odd order.} A sequence of $n$ words originally ordered as $w_1 w_2 w_3 w_4 \cdots w_n$ is reordered by separating the even and odd indices: $w_2 w_4 w_6 \cdots w_{n-1} w_1 w_3 w_5 \cdots w_n$ (assuming $n$ odd).
+Therefore all words that are adjacent in the original order will be separated by a distance of $\approx n/2$ in the distorted order, while all words of distance 2 in the original order will become adjacent.
+
+\paragraph{Interleaving order.} In interleaving ordering, a sequence originally ordered as $w_1 w_2 w_3 \cdots w_n$ is split in half at the middle (index $m=\ceiling{n/2}$), and the two resulting sequences are interleaved, yielding $w_1 w_m w_2 w_{m+1} w_3 w_{m+3} \cdots w_n$.
+Thus all words that were originally adjacent will have distance 2 in the distorted order, with the intervening word coming from a very distant part of the sentence.
+
+\paragraph{Inwards order.} A sequence originally ordered as $w_1 w_2 w_3 \cdots w_{n-1} w_n$ is ordered from the edges of the string inwards, as $w_1 w_n w_2 w_{n-1} \cdots w_{\ceiling{n/2}}$.
+This corresponds to folding the string in on itself once, or equivalently, splitting the sequence in half at the middle, then interleaving the two resulting sequences after reversing the second one.
+The result is that the most non-local possible dependencies in the original order become the most local dependencies in the distorted order.
+
+\paragraph{Sorted order.} A sequence is reordered by sorting by POS tags, and randomizing the order within each block of identical POS tags.
+To each word, we then add a symbol encoding the original position in the sequence.
+For instance
+\begin{center}
+PRON VERB PRON
+\end{center}
+may be reordered as
+\begin{center}
+PRON 1 PRON 3 VERB 2
+\end{center}
+or
+\begin{center}
+PRON 3 PRON 1 VERB 2
+\end{center}
+The numbers are provided to the parser as atomic symbols from a vocabulary ranging from 1 to 200; numbers greater than 200 (which may occur in extremely long sentences) are replaced by an out-of-range token.
+
+The result is that distance between words in the input is not indicative at all of the presence of absence of syntactic relations between them.
+
+\paragraph{Experiments}
+Using English and Japanese data, we trained parsers for the ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
+Resulting parsing accuracy scores are shown in Figure~\ref{fig:distorted-parser}.
+In all settings, the language optimized for efficiency achieved higher parsing accuracy than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
+%Note that absolute values of parsing accuracy vary between the orderings, being higher for the actual orderings than for distorted orderings, in which trees become nonprojective.
+
+\begin{figure}
+    \centering
+    English
+
+    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-english.pdf}
+
+    Japanese
+
+    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
+	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').}
+    \label{fig:distorted-parser}
+\end{figure}
+
+\subsection{n-gram language models}
+
+We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
+In previous work, such as \cite{gildea2015human}, predictability has often been measured using n-gram models.
+
+Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
+Specifically, we constructed bigram models with Kneser-Ney smoothing~\cite{kneser1995improved, chen1999empirical}.
+A bigram model predicts each word taking only the previous word into account.
+This contrasts with LSTMs, which take the entire context into consideration.
+Thus, bigram models and LSTMs stand on opposing ends of a spectrum of language models taking more and more aspects of the context into account.
+
+We estimated language models on the training partitions, and used the validation partitions to estimate surprisal.
+We conducted this for the ten random and the best optimized ordering grammars on English and Japanese data.
+Results (Figure~\ref{fig:bigrams}) show that languages optimized for efficiency are optimal for a bigram language model.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.4]{../results/bigrams/bigrams.pdf} 
+	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).\textcolor{red}{make fig nicer}}
+    \label{fig:bigrams}
+\end{figure}
+
+%\subsection{Discussion}
+%
+%It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.
+
+\section{Effects of data sparsity}
+
+Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
+If the difference between random and real grammars is due to data sparsity, we expect that it decreases as the amount of training data is increased.
+If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference persists as training data is increased.
+
+We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
+We cosidered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
+For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
+Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.
+
+The resulting curves are shown in Figure~\ref{fig:learning-czech}.
+%A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
+A gap in parsing loss of about 0.2 nats appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
+
+\begin{figure}[ht]
+    \centering
+%    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech.pdf} 
+    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech-logloss.pdf}
+
+        \caption{Parsing loss ($H[T|\mathcal{L}_\theta(T)]$) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
+    \label{fig:learning-czech}
+\end{figure}
+
+\section{Comparing Extracted Grammars to Linguistic Judgments}
+
+% user@user-X510UAR:~/grammar-optim/results/grammars$ git add comparison-table.tex
+
+We validated our grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
+Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS.
+In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb-object relation.
+
+WALS has some data for all but seven languages.\footnote{Note that Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
+In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}
+WALS has data for 74\% of the entries, and lists a dominant order for 91\% of these.
+The grammars we extracted from the corpora agree with WALS in 96~\%  of these cases.
+
+\begin{table}[htb!]
+\small{
+\begin{center}
+\begin{tabular}{l||ll|ll|ll|ll|ll|ll|llllll}
+%	   for(dep in c("obj", "lifted_mark", "acl", "nmod", "obl")) {
+		   Language 
+		   &	\multicolumn{2}{|c|}{Objects} 
+		   &	\multicolumn{2}{|c|}{Adpositions} 
+		   &	\multicolumn{2}{|c|}{Compl.} 
+		   &	\multicolumn{2}{|c|}{Rel.Cl.} 
+		   &	\multicolumn{2}{|c|}{Genitive} 
+		   &	\multicolumn{2}{|c|}{PP}  \\ \hline\hline
+	\input{../results/grammars/comparison-table.tex}
+\end{tabular}
+\end{center}
+}
+\caption{Comparing grammars extracted from databases to linguistic judgments in the World Atlas of Linguistic Structures.
+For each of the six syntactic relation, the first column provides the ordered coded in the extracted grammar; the second column provides the order coded in WALS (DH for dependent-head, HD for head-dependent order).
+`?' indicates that WALS has no data.
+$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur at similar rates in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
+Finally, `--' indicates that the relation does not occur in the corpus.
+}\label{tab:grammars-wals}
+\end{table}
+
+\section{Languages and Corpus Sizes}
+In Table~\ref{tab:langs-iso-sizes}, we list the 51 languages with ISO codes and families, with the size of the available data per language.
+We included all UD 2.1 languages for which a training partition was available.
 
 \begin{table}[ht]
+\small{
 \begin{tabular}{lllllll}
@@
 \end{tabular}
-\caption{Languages}
+}
+\caption{Languages with ISO codes, families (according to \url{https://universaldependencies.org/}), and the number of available sentence and words.}\label{tab:langs-iso-sizes}
 \end{table}
@@
 In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
+
+DLM turns out to be predicted by our model
 

### 2019-05-18 second-draft/pnas-word-order-si.tex
SI

@@
 
-Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM}).
-Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
-We optimize grammars using the same stochastic gradient descent method as we use for efficiency optimization (Section REF).
-Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
-Indeed, we found that DLM is predicted by efficiency optimization, as discussed in Section REF.
-Results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
+Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM} for discussion).
+This has been suggested as a property of efficient word order in prior work, and, indeed, we found that DLM is predicted by efficiency optimization (Section REF).
+Results in Table~\ref{tab:all-predictions-1} also show that optimizing for DLM makes predictions similar to efficiency optimization. See Section REF.
 
@@
 
-\textcolor{red}{TODO}
-We have directly implemented parseability and predictability using general-purpose statistical models of language and dependency structures.
-Prior work has considered Dependency Length Minimization (DLM) as an approximate formalization of efficiency optimization~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-This is the idea that languages order words so as to minimize the overall length of dependencies.
-This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
-Indeed, directly modeling efficiency optimization predicts DLM, and appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
-
-\textcolor{red}{at some place also say the following: Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
-}
-
-Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM}).
-Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
+Prior work has suggested \emph{Dependency Length Minimization} (DLM) as a characteristic of efficient word order~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+This is the idea that word order minimizes the average distance between syntactically related words.
+It is known that human languages reduce dependency length compared to random baselines~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+
 Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
-Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
-In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
-
-DLM turns out to be predicted by our model
+It is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
+Indeed, \cite{futrell2017memory} argues specifically that it emerges from efficiency optimization.
+
+Dependency length is typically quantified as the average distance between all pairs of syntactically related words, measured by the number of intervening words~\cite{futrell2015largescale}.
+Dependency length quantified in this manner is a heuristic measure of complexity:
+The actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
+
+We asked whether efficiency optimization predicts dependency length minimization effects.
+We first computed dependency length for grammars optimized for efficiency.
+We found that 100\% of grammars optimized for efficiency reduce average dependency length compared to baseline grammars ($p < 0.05$, by one-sided $t$-test). \textcolor{red}{TODO}
+This suggests that the reduction of dependency length is predicted by efficiency maximization.
+Next, we constructed grammars that minimize average dependency length, using the same gradient descent method as we used for efficiency optimization (Section REF).
+We expect that such grammars should have shorter dependency length than the real grammars, or grammars optimized for efficiency.
+In Figure~\ref{fig:dlm-avg}, we plot the mean dependency length for optimized, real, and baseline orderings.
+We find that optimizing grammars for efficiency reduces dependency length to a similar degree as found in the actual orderings in the corpora, almost up to the limit given by directly optimizing for dependency length.
+We also plot more detailed results for four languages in Figure~\ref{fig:dlm-4langs}, plotting dependency length as a function of sentence length as reported in prior work \cite{futrell2015largescale}.
+Optiziming grammars for efficiency produces dependency lengths similar to those found in the actual orderings.
+
+\begin{figure}[ht]
+    \centering
+     \includegraphics[width=0.5\textwidth]{../results/dependency-length/figures/depl-violin-all.pdf} 
+        \caption{Average dependency length for grammars optimized to minimize dependency length (DLM, left), optimized for efficiency (second), the real orderings found in corpora (third), and random baseline grammars (right).}
+    \label{fig:dlm-avg}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9\textwidth]{../results/dependency-length/figures/depLength-facet.pdf} 
+        \caption{Total dependency length as a function of sentence length, for four diverse languages.
+We show results for optimized grammars (parseability, predictability, efficiency), for grammars specifically optimized to minimize dependency length, of the actual real orderings, and of the baseline grammars.}
+    \label{fig:dlm-4langs}
+\end{figure}
+
+Next, we examined the word order properties of grammars optimized for DLM.
+In Table~\ref{tab:all-predictions-1}, we report the posterior prevalence of word order correlations in grammars optimized for DLM show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
+We find that these grammars also exhibit the eight correlations, similar to grammars directly optimized for efficiency.
+On other correlations, predictions of DLM also resemble those of efficiency optimization.
+However, it predicts strong correlations with \textit{amod} (adjectival modifiers) and \textit{nummod} (numeral modifiers) (see bottom of Table~\ref{tab:all-predictions-1}), which are not borne out typologically.
+In these cases, efficiency optimization predicts prevalnces closer to 50\%, in line with typological data.
+
+In conclusion, these results suggest that dependency length minimization is a by-product of efficiency optimization, providing support to theoretical arguments from the linguistic literature~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
+Furthermore, optimizing for dependency length correctly predicts a range of word order facts, though it appears to \emph{overpredict} correlations when compared to direct optimization for efficiency.
+
+%Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM}).
+%Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
+%Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
+%Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
+%In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
+
+%Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
+%We optimize grammars using the same stochastic gradient descent method as we use for efficiency optimization (Section REF).
+%Indeed, we found that DLM is predicted by efficiency optimization, as discussed in Section REF.
+%Results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
+
+%\textcolor{red}{TODO}
+
+%DLM turns out to be predicted by our model
 

### 2019-05-18 second-draft/pnas-word-order-si.tex
SI

@@
 The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
-%We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations, see Table REF.
 We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (Table~\ref{tab:all-predictions-1}).
@@
 Within each language, we estimate the Pareto curve based on the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
-%We then estimated the overall average Pareto curve, as shown in Figure (REF), by averaging the z-transformed points (1), (2), (3).
 
@@
 Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).
-White cells indicate that the relation is not annotated in the dataset for the given language.}\label{fig:per-lang}
+White cells indicate that the relation is not annotated in the dataset for the given language.
+\textcolor{red}{TODO fix order of 1-8}}\label{fig:per-lang}
 %	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
@@
 Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM} for discussion).
-This has been suggested as a property of efficient word order in prior work, and, indeed, we found that DLM is predicted by efficiency optimization (Section REF).
-Results in Table~\ref{tab:all-predictions-1} also show that optimizing for DLM makes predictions similar to efficiency optimization. See Section REF.
+This has been suggested as a property of efficient word order in prior work, and, indeed, we found that DLM is predicted by efficiency optimization (Section~\ref{sec:DLM}).
+Results in Table~\ref{tab:all-predictions-1} also show that optimizing for DLM makes predictions similar to efficiency optimization. See Section~\ref{sec:DLM}.
 
@@
 %	AP &
-    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
 \end{tabular}
@@
 In Table~\ref{table:corr-resu} we report the results of the simulations we did before the main experiment reported in the main paper (left and center), together with their preregistrations, besides the results from the main experiment (right).
-These experiments all had the same setup described in Section (REF), which was fixed before starting simulations; differences are that (1) the first simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
+These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) the first simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
 Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
@@
 	%&
-& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table REF.} 
+& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM}.} 
 & $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
@@
 
-\subsection{Optimizing Grammars for Efficiency}\label{sec:optim-eff}
-
-In this section, we describe how we optimized grammar parameters for efficiency.
-A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
-More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $\mathbf{w}$.
-In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $p_{\mathcal{L}_\theta}(u|t)$ concentrates on a single ordering $u$.
-
-Recall the definition of efficiency
-\begin{equation}\label{eq:efficiency}
-	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
-\end{equation}
-where 
-\begin{equation}
-	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
-\end{equation}
-\begin{equation}
-	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
-\end{equation}
-where $t \sim \mathcal{T}$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
-
-These quatities are estimated using two neural models, as described in Section X:
-A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.
-The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is
-\begin{equation}
-	R_{Pars}^{\phi}(u,t) =  \log \frac{p_\phi(t|u)}{p(t)}
-\end{equation}
-%and we can write
-%\begin{equation}
-%	R_{Pars}^\phi := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
-%\end{equation}
-%
-A \key{language model}, with some parameters $\psi$, calculates the word-by-word surprisal of an utterance:
-\begin{equation}
-	R_{Pred}^{\psi}(u) = - \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})
-\end{equation}
-Using this, we can rewrite Efficiency~(\ref{eq:efficiency}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
-\begin{equation}
-	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi} := \max_{\phi,\psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
-\end{equation}
-%
-%The parsers and language models appropriate to a tree distribution $\mathcal{T}$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
-%\begin{equation}
-%\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-%\end{equation}
-%\begin{equation}
-%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-%\end{equation}
-%
-%Using these models, we can rewrite 
-%\begin{equation}
-%R_{Pars}^{\psi(\theta)}= \E_t \E_{u \sim p_\theta(u|t)} R_{Pars}^{\psi(\theta)}(u,t) 
-%\end{equation}
-%where $R_{Pars}(u,t) = \log \frac{p_\phi(t|u)}{p(t)}$, and $p_\theta(u|t)$ is the probability assigned to sentence $u$ by the distribution $\mathcal{L}_\theta(t)$ defined by the differentiable word order grammar $\mathcal{L}_\theta$,
-%and
-%\begin{equation}
-%R_{Pred} = \E_t \E_{u \sim p_\theta(u|t)} R_{Pred}(u)
-%\end{equation}
-%where $R_{Pred}(u) = \log p_\psi(u)$.
-%Therefore, we can rewrite efficiency as
-%\begin{equation}\label{eq:efficiency}
-%R_{\textit{Eff}}^\theta := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi(\theta)}(u,t) + \lambda R_{Pred}^{\psi(\theta)}(u)\right]
-%\end{equation}
-%Directly applying gradient descent to this expression is hard due to the complicated dependence of $\phi(\theta), \psi(\theta)$ on $\theta$.
-%
-%However, we can rewrite
-%\begin{equation}\label{eq:efficiency}
-%\argmax_\theta\	R_{\textit{Eff}}^\theta 
-%\end{equation}
-%equivalently into
-In order to find an optimal grammar $\theta$, we thus need to solve
-\begin{equation}\label{eq:efficiency}
-\argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} %= \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
-\end{equation}
-Importantly, $R_{\textit{Eff}}^{\theta, \phi, \psi}$ is differentiable in $\theta, \phi, \psi$: %, and we can apply stochastic gradient descent to carry out this optimization.
-\begin{align}
-\partial_\theta R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)} \left[  \left[\partial_\theta \log p_\theta(u|t)\right] \cdot    \left(R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right) \right] \label{eq:dtheta}\\
-\partial_\phi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\partial_\phi R_{Pars}^{\phi}(u,t)\right] \\
-\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}
-\end{align}
-where \ref{eq:dtheta} is derived using the \emph{score-function} or \emph{REINFORCE} trick~\cite{williams1992simple}.
-Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures (CITE).
-
-We can therefore apply stochastic gradient descent to jointly optimize $\theta, \phi, \psi$:
-In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim p_{\theta}(\cdot|t)$.
-Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
-do a gradient descent step using the estimator given by the expressions in the square brackets in \ref{eq:dtheta}-\ref{eq:dpsi}.
-
-Optimizing for only parseability (or predictability) is very similar -- in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.
-
-At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
-Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.
-
-See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates used in this optimization scheme.
-
-%
-%%(analogous for $R_{Pars}$ instead of $R_{Pred}$)
-%\begin{equation}\label{eq:estimator}
-%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
-%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
-%\end{equation}
-%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
-%for the gradient
-%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
-%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
-%
-
-%
-%
-%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
-Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
-%\begin{equation}
-%\label{eq:exp-over-utts}
-%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
-%\end{equation}
-%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
-In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
-%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
-%
-%
-%
-%We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
-%\begin{equation*}
-%R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
-%\end{equation*}
-%where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
-%\begin{equation*}
-%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-%\end{equation*}
-%
-%
-%We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
-Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
-%\begin{align}
-%\nonumber
-%R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
-%\nonumber
-%&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
-%\end{align}
-%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
-%\begin{equation*}
-%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-%\end{equation*}
-%
-%For efficiency, we linearly combine the two terms:
-%\begin{equation}
-%\label{eq:combined-score}
-%R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
-%\end{equation}
-%
-%
-%%\begin{equation}
-%%    \label{eq:depl}
-%%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
-%%\end{equation}
-%%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
-%
-%For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
-The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
-The optimization problems come out to:
-%\begin{align}
-%%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
-%\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
-%\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
-%\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
-%\end{align}
-%Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
-%
-%
-%%The case of predictability, parseability, and efficiency are more challenging, as 
-%%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
-%We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
-%Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
-%
-%We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section (REF)).
-%
-%In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
-%Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
-%do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
-%\begin{equation}\label{eq:estimator}
-%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
-%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
-%\end{equation}
-%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
-%for the gradient
-%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
-%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
-%
-%See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
-%
-
-%The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
-A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
-%We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
-
-%In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
-%and $\ell_2$ regularization.
-
-%A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
-%Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
-%We thus do not evaluate this objective function here.
-
-%We extract grammars for the actual languages using a Bayesian data analysis method.
-%We specify a simple Bayesian prior over differentiable ordering grammars:
-%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
-%for each $\tau$.
-%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
-%We extract
-
-\section{Neural Network Architectures}\label{sec:neural-architectures}
-
-In this section, we describe the details of the neural network architectures.
-Choices follow standard practice and machine learning and were fixed before evaluating word order properties, and the efficiency of real languages.
-
-%\paragraph{Locality}
-%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
-
-\paragraph{Estimating Predictability}
-We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
-Given the small size of the corpora, this limit is only attained only for few languages.
-In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
-The model predicts both the next word and its POS tags in each step.
-Using POS tags is intended to prevent overfitting on small corpora.
-
-\paragraph{Estimating Parseability}
-We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}.
-This architecture is remarkably simple: the words of a sentence are encoded into context-sensitive embeddings using bidirectional LSTMs, then a classifier is trained to predict the head for each work.
-The classifier works by calculating a score for every pair of word embeddings $(w_i, w_j)$, indicating the likelihood that the $j$th word is the head of the $i$th word.
-This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
-
-To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags.
-Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged; parsing from POS tags prevents early overfitting.
-This decision was made before evaluating word order properties.
-
-\paragraph{Hyperparameters}
-
-%\textbf{TODO study 1 only has 1 set of parameters}
-
-Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
-For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
-These parameters are shown in Table~\ref{tab:hyperparameters}.
-Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (efficiency, predictability, parseability) on the language.
-We then used this setting for creating optimized word order grammars.
-
-%\textcolor{red}{TODO random seeds for efficiency}
-
-All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
-We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|l|l|l|}
-\hline
-\multirow{2}{*}{Optimization}&    Learning Rate     & 5e-6, 1e-5, 2e-5, 5e-5 \\
-&    Momentum & 0.8, 0.9 \\ \hline
-\multirow{6}{*}{Language Model} &    Learning Rate  & 0.5, 0.1, 0.2 \\
-&    Dropout Rate& 0.0, 0.3, 0.5 \\
-& Embedding Size (Words) & 50 \\
-& Embedding Size (POS) & 20 \\
-&    LSTM Layers  & 2 \\
-&    LSTM Dimensions  & 128 \\
- \hline
-\multirow{5}{*}{Parser}&    Learning Rate  & 0.001 \\
-&    Dropout Rate  & 0.2 \\
-&    Embedding Size  & 100 \\
-&    LSTM Layers  & 2 \\
-&    LSTM Dimensions  & 200 \\
-\hline
-    \end{tabular}
-    \caption{Hyperparameters}
-    \label{tab:hyperparameters}
-\end{table}
-
-\paragraph{Improved Unbiased Gradient Estimator}
-We employ two common variance reduction methods to improve the estimator~(\ref{eq:dtheta}), while keeping it unbiased.
-%For dependency length and 
-% (and its head, in the case of dependency length)
-For predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word.
-We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:dtheta}) with lower variance.
-Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple} (the word's surprisal in the case of predictability, and the negative log-probability of the correct head and relation label in the case of parseability).
-These two methods reduce the variance of the estimator and thereby increase the speed of optimization and reduce training time, without biasing the results.
-For numerical stability, we represent $a_\tau \in [0,1]$ via its logit $\in \mathbb{R}$.
-Furthermore, to encourage exploration of the parameter space, we add an entropy regularization term \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$.
-The weight of the entropy regularization was chosen together with the other hyperparameters.\footnote{Explored values: 0.0001, 0.001.}
-%, adding the negative entropy of the $Bernoulli(a_\tau)$ to the loss $R$. %, weighted with $\alpha$.
-
-These techniques for improving (\ref{eq:dtheta}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
-
-\paragraph{Learning Rate and Stopping Criterion}
-We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
-For the language model parameters $\phi$, we use plain Stochastic Gradient Descent without momentum, as recommended by \citet{merity2018regularizing}. 
-For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
-The learning rates and other optimization hyperparameters were determined together with the other hyperparameters.
-
-For gramar optimization, we apply Early Stopping~\citep{prechelt1998early} as follows.
-We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
-This is a standard technique for optimizing neural models, and helps prevent overfitting to the specific corpora used.
-
-The choice of optimization methods and the stopping criterion were fixed before we investigated word order correlations.
-
-\paragraph{Optimized Grammars}
-Recall that, for each language, we created 8 optimized languages for each optimization criterion.
-We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
-This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.
-
-For efficiency optimization, for each grammar, we ran efficiency optimization with three different random seeds, selecting among these the seed that yielded the best overall efficiency value.
-This helps control for variation across random seeds for the stochastic gradient descent optimization method.
-
- % TODO we could use some language at the top of this section on the general guiding principles for all these decisions.
-Were these implementation decisions made before or after looking at the word order universals results?
-If they were made before, we should highlight that.
-
-\section{Robustness to different language models and parsers}
-
-Here we take up the question of the extent to which our results are dependent on the particular parser and language model used in the optimization process.
-We want to know: when we optimize a word order grammar for efficiency, have we produced a language which is highly efficient \emph{in general}, or one which is highly efficient \emph{for a specific parser}?
-We wish to argue that natural language syntax is optimized for efficiency in general, meaning that syntactic trees are highly recoverable from word orders in principle.
-If it turns out that our optimized languages are only optimal for a certain parser from the NLP literature, then we run the risk of circularity: it may be that the reason this parser was successful in the NLP literature was because it implicitly encoded word order universals in its inductive biases, and thus it would be no surprise that languages which are optimized for parseability also show those universals.
-
-In this connection, we note that the parser and language model architectures we use are highly generic, and do not encode any obvious bias toward natural-language-like word orders.
-The LSTM language model is a generic model of sequence data which is also been used to model financial time series \citep{sirignano2018universal} and purely theoretical chaotic dynamical systems \citep{ogunmolu2016nonlinear}; the neural graph-based parser is simply solving a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
-Nevertheless, it may be the case that a bias toward word order universals is somehow encoded implicitly in the hyperparameters and architectures of these models.
-
-Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers.
-These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.
-
-\subsection{CKY Parsers}
-
-We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of \textcolor{red}{CITE Collins 1997 (Model 1).}
-In our PCFGs, each head independently generates a set of left and right dependents.
-We formulate this as a PCFG where each rule has the form:
-\begin{center}
-	POS$_H$ $\rightarrow$ POS$_H$ POS$_D$
-\end{center}
-for head-initial structures, and
-\begin{center}
-	POS$_H$ $\rightarrow$ POS$_D$ POS$_H$
-\end{center}
-for head-final structures, where each symbol is a POS tag.
-Thus, POS tags act both as terminals and as nonterminals.
-
-We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
-For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm \textcolor{red}{CITE}.
-
-This parsing strategy is very different from the neural graph-based parser:
-While the graph-based parser solves a minimum spanning tree problem, the CKY algorithm uses dynamic programming to compute the exact probabilities of trees given a sentence, as specified by the generative model encoded in the PCFG.
-Second, while the graph-based neural parser uses machine learning to induce syntactic knowledge from data, the CKY parser performs exact probabilistic inference.
-%In this sense, the CKY algorithm does not have any architectural biases in itself.
-%On the other hand, the PCFG makes severely simplifying independence assumptions, compared to the universal approximation capabilities of neural network-based systems.
-
-We used the CKY algorithm to compute the parsing loss $H[T|L_\theta(T)]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
-Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
-% TODO get UAS for compatibility with other parts of SI
-%We then compared parsability of random and optimized word order grammars on English data.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
-	\caption{Parsability loss $H[T|L_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).\textcolor{red}{make plot nicer}}
-    \label{fig:cky-parser}
-\end{figure}
-
-\subsection{Distorted graph-based parsers}
-
-In this section, we address the idea that the graph-based parser might have a built-in bias toward certain kinds of orderings, and the question whether this might be responsible for our findings.
-In particular, we address the idea that the graph-based parser might have a bias toward parses involving short dependencies, which we call a \key{locality bias}. 
-We address this by changing the order in which the parser sees words, in such a way that the distance between words in the input is not indicative of syntactic relations.
-
-%In particular, this allows us to show tha
-
-%Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization.
-Therefore, we address here the idea that the graph-based parser might have a built-in bias toward parses involving short dependencies, which we call a \key{locality bias}.
-%A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words.
-Suppose that a parser $P$ has a bias toward positing a dependency between a word and the immediate previous word.
-
-\paragraph{Even--odd order.} A sequence of $n$ words originally ordered as $w_1 w_2 w_3 w_4 \cdots w_n$ is reordered by separating the even and odd indices: $w_2 w_4 w_6 \cdots w_{n-1} w_1 w_3 w_5 \cdots w_n$ (assuming $n$ odd).
-Therefore all words that are adjacent in the original order will be separated by a distance of $\approx n/2$ in the distorted order, while all words of distance 2 in the original order will become adjacent.
-
-\paragraph{Interleaving order.} In interleaving ordering, a sequence originally ordered as $w_1 w_2 w_3 \cdots w_n$ is split in half at the middle (index $m=\ceiling{n/2}$), and the two resulting sequences are interleaved, yielding $w_1 w_m w_2 w_{m+1} w_3 w_{m+3} \cdots w_n$.
-Thus all words that were originally adjacent will have distance 2 in the distorted order, with the intervening word coming from a very distant part of the sentence.
-
-\paragraph{Inwards order.} A sequence originally ordered as $w_1 w_2 w_3 \cdots w_{n-1} w_n$ is ordered from the edges of the string inwards, as $w_1 w_n w_2 w_{n-1} \cdots w_{\ceiling{n/2}}$.
-This corresponds to folding the string in on itself once, or equivalently, splitting the sequence in half at the middle, then interleaving the two resulting sequences after reversing the second one.
-The result is that the most non-local possible dependencies in the original order become the most local dependencies in the distorted order.
-
-\paragraph{Sorted order.} A sequence is reordered by sorting by POS tags, and randomizing the order within each block of identical POS tags.
-To each word, we then add a symbol encoding the original position in the sequence.
-For instance
-\begin{center}
-PRON VERB PRON
-\end{center}
-may be reordered as
-\begin{center}
-PRON 1 PRON 3 VERB 2
-\end{center}
-or
-\begin{center}
-PRON 3 PRON 1 VERB 2
-\end{center}
-The numbers are provided to the parser as atomic symbols from a vocabulary ranging from 1 to 200; numbers greater than 200 (which may occur in extremely long sentences) are replaced by an out-of-range token.
-
-The result is that distance between words in the input is not indicative at all of the presence of absence of syntactic relations between them.
-
-\paragraph{Experiments}
-Using English and Japanese data, we trained parsers for the ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
-Resulting parsing accuracy scores are shown in Figure~\ref{fig:distorted-parser}.
-In all settings, the language optimized for efficiency achieved higher parsing accuracy than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
-%Note that absolute values of parsing accuracy vary between the orderings, being higher for the actual orderings than for distorted orderings, in which trees become nonprojective.
-
-\begin{figure}
-    \centering
-    English
-
-    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-english.pdf}
-
-    Japanese
-
-    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
-	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').}
-    \label{fig:distorted-parser}
-\end{figure}
-
-\subsection{n-gram language models}
-
-We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
-In previous work, such as \cite{gildea2015human}, predictability has often been measured using n-gram models.
-
-Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
-Specifically, we constructed bigram models with Kneser-Ney smoothing~\cite{kneser1995improved, chen1999empirical}.
-A bigram model predicts each word taking only the previous word into account.
-This contrasts with LSTMs, which take the entire context into consideration.
-Thus, bigram models and LSTMs stand on opposing ends of a spectrum of language models taking more and more aspects of the context into account.
-
-We estimated language models on the training partitions, and used the validation partitions to estimate surprisal.
-We conducted this for the ten random and the best optimized ordering grammars on English and Japanese data.
-Results (Figure~\ref{fig:bigrams}) show that languages optimized for efficiency are optimal for a bigram language model.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.4]{../results/bigrams/bigrams.pdf} 
-	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).\textcolor{red}{make fig nicer}}
-    \label{fig:bigrams}
-\end{figure}
-
-%\subsection{Discussion}
-%
-%It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.
-
-\section{Effects of data sparsity}
-
-Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
-If the difference between random and real grammars is due to data sparsity, we expect that it decreases as the amount of training data is increased.
-If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference persists as training data is increased.
-
-We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
-We cosidered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
-For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
-Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.
-
-The resulting curves are shown in Figure~\ref{fig:learning-czech}.
-%A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
-A gap in parsing loss of about 0.2 nats appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
-
-\begin{figure}[ht]
-    \centering
-%    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech.pdf} 
-    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech-logloss.pdf}
-
-        \caption{Parsing loss ($H[T|\mathcal{L}_\theta(T)]$) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
-    \label{fig:learning-czech}
-\end{figure}
-
-\section{Comparing Extracted Grammars to Linguistic Judgments}
-
-% user@user-X510UAR:~/grammar-optim/results/grammars$ git add comparison-table.tex
-
-We validated our grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
+We validated the extracted grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
 Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS.
@@
 
-\begin{table}[htb!]
+\begin{table}
 \small{
@@
 
+\subsection{Optimizing Grammars for Efficiency}\label{sec:optim-eff}
+
+In this section, we describe how we optimized grammar parameters for efficiency.
+A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
+More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $\mathbf{w}$.
+In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $p_{\mathcal{L}_\theta}(u|t)$ concentrates on a single ordering $u$.
+
+Recall the definition of efficiency
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+\end{equation}
+where 
+\begin{equation}
+	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
+\end{equation}
+\begin{equation}
+	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+\end{equation}
+where $t \sim \mathcal{T}$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
+
+These quatities are estimated using two neural models, as described in Section~\ref{sec:neural-architectures}:
+A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.
+The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is
+\begin{equation}
+	R_{Pars}^{\phi}(u,t) =  \log \frac{p_\phi(t|u)}{p(t)}
+\end{equation}
+%and we can write
+%\begin{equation}
+%	R_{Pars}^\phi := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
+%\end{equation}
+%
+A \key{language model}, with some parameters $\psi$, calculates the word-by-word surprisal of an utterance:
+\begin{equation}
+	R_{Pred}^{\psi}(u) = - \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})
+\end{equation}
+Using this, we can rewrite Efficiency~(\ref{eq:efficiency}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
+\begin{equation}
+	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi} := \max_{\phi,\psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\end{equation}
+%
+%The parsers and language models appropriate to a tree distribution $\mathcal{T}$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
+%\begin{equation}
+%\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation}
+%\begin{equation}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation}
+%
+%Using these models, we can rewrite 
+%\begin{equation}
+%R_{Pars}^{\psi(\theta)}= \E_t \E_{u \sim p_\theta(u|t)} R_{Pars}^{\psi(\theta)}(u,t) 
+%\end{equation}
+%where $R_{Pars}(u,t) = \log \frac{p_\phi(t|u)}{p(t)}$, and $p_\theta(u|t)$ is the probability assigned to sentence $u$ by the distribution $\mathcal{L}_\theta(t)$ defined by the differentiable word order grammar $\mathcal{L}_\theta$,
+%and
+%\begin{equation}
+%R_{Pred} = \E_t \E_{u \sim p_\theta(u|t)} R_{Pred}(u)
+%\end{equation}
+%where $R_{Pred}(u) = \log p_\psi(u)$.
+%Therefore, we can rewrite efficiency as
+%\begin{equation}\label{eq:efficiency}
+%R_{\textit{Eff}}^\theta := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi(\theta)}(u,t) + \lambda R_{Pred}^{\psi(\theta)}(u)\right]
+%\end{equation}
+%Directly applying gradient descent to this expression is hard due to the complicated dependence of $\phi(\theta), \psi(\theta)$ on $\theta$.
+%
+%However, we can rewrite
+%\begin{equation}\label{eq:efficiency}
+%\argmax_\theta\	R_{\textit{Eff}}^\theta 
+%\end{equation}
+%equivalently into
+In order to find an optimal grammar $\theta$, we thus need to solve
+\begin{equation}\label{eq:efficiency}
+\argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} %= \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\end{equation}
+Importantly, $R_{\textit{Eff}}^{\theta, \phi, \psi}$ is differentiable in $\theta, \phi, \psi$: %, and we can apply stochastic gradient descent to carry out this optimization.
+\begin{align}
+\partial_\theta R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)} \left[  \left[\partial_\theta \log p_\theta(u|t)\right] \cdot    \left(R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right) \right] \label{eq:dtheta}\\
+\partial_\phi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\partial_\phi R_{Pars}^{\phi}(u,t)\right] \\
+\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}
+\end{align}
+where \ref{eq:dtheta} is derived using the \emph{score-function} or \emph{REINFORCE} trick~\cite{williams1992simple}.
+Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures (CITE).
+
+We can therefore apply stochastic gradient descent to jointly optimize $\theta, \phi, \psi$:
+In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim p_{\theta}(\cdot|t)$.
+Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
+do a gradient descent step using the estimator given by the expressions in the square brackets in \ref{eq:dtheta}-\ref{eq:dpsi}.
+
+Optimizing for only parseability (or predictability) is very similar -- in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.
+
+At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
+Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.
+
+See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates used in this optimization scheme.
+
+%
+%%(analogous for $R_{Pars}$ instead of $R_{Pred}$)
+%\begin{equation}\label{eq:estimator}
+%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
+%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
+%\end{equation}
+%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
+%for the gradient
+%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
+%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
+%
+
+%
+%
+%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
+Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
+%\begin{equation}
+%\label{eq:exp-over-utts}
+%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
+%\end{equation}
+%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
+In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
+%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
+%
+%
+%
+%We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
+%\begin{equation*}
+%R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
+%\end{equation*}
+%where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
+%\begin{equation*}
+%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation*}
+%
+%
+%We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
+Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
+%\begin{align}
+%\nonumber
+%R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
+%\nonumber
+%&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
+%\end{align}
+%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
+%\begin{equation*}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation*}
+%
+%For efficiency, we linearly combine the two terms:
+%\begin{equation}
+%\label{eq:combined-score}
+%R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
+%\end{equation}
+%
+%
+%%\begin{equation}
+%%    \label{eq:depl}
+%%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+%%\end{equation}
+%%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
+%
+%For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
+The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
+The optimization problems come out to:
+%\begin{align}
+%%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
+%\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
+%\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
+%\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
+%\end{align}
+%Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
+%
+%
+%%The case of predictability, parseability, and efficiency are more challenging, as 
+%%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
+%We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
+%Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
+%
+%We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section).
+%
+%In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
+%Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
+%do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
+%\begin{equation}\label{eq:estimator}
+%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
+%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
+%\end{equation}
+%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
+%for the gradient
+%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
+%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
+%
+%See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
+%
+
+%The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
+A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
+%We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
+
+%In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
+%and $\ell_2$ regularization.
+
+%A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
+%Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
+%We thus do not evaluate this objective function here.
+
+%We extract grammars for the actual languages using a Bayesian data analysis method.
+%We specify a simple Bayesian prior over differentiable ordering grammars:
+%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
+%for each $\tau$.
+%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
+%We extract
+
+\section{Neural Network Architectures}\label{sec:neural-architectures}
+
+In this section, we describe the details of the neural network architectures.
+Choices follow standard practice and machine learning and were fixed before evaluating word order properties, and the efficiency of real languages.
+
+%\paragraph{Locality}
+%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
+
+\paragraph{Estimating Predictability}
+We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
+Given the small size of the corpora, this limit is only attained only for few languages.
+In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
+The model predicts both the next word and its POS tags in each step.
+Using POS tags is intended to prevent overfitting on small corpora.
+
+\paragraph{Estimating Parseability}
+We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}.
+This architecture is remarkably simple: the words of a sentence are encoded into context-sensitive embeddings using bidirectional LSTMs, then a classifier is trained to predict the head for each work.
+The classifier works by calculating a score for every pair of word embeddings $(w_i, w_j)$, indicating the likelihood that the $j$th word is the head of the $i$th word.
+This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
+
+To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags.
+Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged; parsing from POS tags prevents early overfitting.
+This decision was made before evaluating word order properties.
+
+\paragraph{Hyperparameters}
+
+%\textbf{TODO study 1 only has 1 set of parameters}
+
+Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
+For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
+These parameters are shown in Table~\ref{tab:hyperparameters}.
+Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (efficiency, predictability, parseability) on the language.
+We then used this setting for creating optimized word order grammars.
+
+%\textcolor{red}{TODO random seeds for efficiency}
+
+All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
+We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{|l|l|l|}
+\hline
+\multirow{2}{*}{Optimization}&    Learning Rate     & 5e-6, 1e-5, 2e-5, 5e-5 \\
+&    Momentum & 0.8, 0.9 \\ \hline
+\multirow{6}{*}{Language Model} &    Learning Rate  & 0.5, 0.1, 0.2 \\
+&    Dropout Rate& 0.0, 0.3, 0.5 \\
+& Embedding Size (Words) & 50 \\
+& Embedding Size (POS) & 20 \\
+&    LSTM Layers  & 2 \\
+&    LSTM Dimensions  & 128 \\
+ \hline
+\multirow{5}{*}{Parser}&    Learning Rate  & 0.001 \\
+&    Dropout Rate  & 0.2 \\
+&    Embedding Size  & 100 \\
+&    LSTM Layers  & 2 \\
+&    LSTM Dimensions  & 200 \\
+\hline
+    \end{tabular}
+    \caption{Hyperparameters}
+    \label{tab:hyperparameters}
+\end{table}
+
+\paragraph{Improved Unbiased Gradient Estimator}
+We employ two common variance reduction methods to improve the estimator~(\ref{eq:dtheta}), while keeping it unbiased.
+%For dependency length and 
+% (and its head, in the case of dependency length)
+For predictability, note that the surprisal of a specific word only depends on the preceding words (not on the following words), and thus only depends on ordering decisions made up to that word.
+We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:dtheta}) with lower variance.
+Second, we use a word-dependent moving average of recent per-word losses (the word's surprisal in the case of predictability, and the negative log-probability of the correct head and relation label in the case of parseability) as control variate \cite{williams1992simple}.
+These two methods reduce the variance of the estimator and thereby increase the speed of optimization and reduce training time, without biasing the results.
+For numerical stability, we represent $a_\tau \in [0,1]$ via its logit $\in \mathbb{R}$.
+Furthermore, to encourage exploration of the parameter space, we add an entropy regularization term \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$.
+The weight of the entropy regularization was chosen together with the other hyperparameters.\footnote{Explored values: 0.0001, 0.001.}
+
+These techniques for improving (\ref{eq:dtheta}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
+
+\paragraph{Learning Rate and Stopping Criterion}
+We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
+For the language model parameters $\phi$, we use plain Stochastic Gradient Descent without momentum, as recommended by \citet{merity2018regularizing}. 
+For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
+The learning rates and other optimization hyperparameters were determined together with the other hyperparameters.
+
+For gramar optimization, we apply Early Stopping~\citep{prechelt1998early} as follows.
+We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
+This is a standard technique for optimizing neural models, and helps prevent overfitting to the specific corpora used.
+
+The choice of optimization methods and the stopping criterion were fixed before we investigated word order correlations.
+
+\paragraph{Optimized Grammars}
+Recall that, for each language, we created 8 optimized languages for each optimization criterion.
+We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
+This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.
+
+For efficiency optimization, for each grammar, we ran efficiency optimization with three different random seeds, selecting among these the seed that yielded the best overall efficiency value.
+This helps control for variation across random seeds for the stochastic gradient descent optimization method.
+
+ % TODO we could use some language at the top of this section on the general guiding principles for all these decisions.
+Were these implementation decisions made before or after looking at the word order universals results?
+If they were made before, we should highlight that.
+
+\section{Robustness to different language models and parsers}
+
+Here we take up the question of the extent to which our results are dependent on the particular parser and language model used in the optimization process.
+We want to know: when we optimize a word order grammar for efficiency, have we produced a language which is highly efficient \emph{in general}, or one which is highly efficient \emph{for a specific parser}?
+We wish to argue that natural language syntax is optimized for efficiency in general, meaning that syntactic trees are highly recoverable from word orders in principle.
+If it turns out that our optimized languages are only optimal for a certain parser from the NLP literature, then we run the risk of circularity: it may be that the reason this parser was successful in the NLP literature was because it implicitly encoded word order universals in its inductive biases, and thus it would be no surprise that languages which are optimized for parseability also show those universals.
+
+In this connection, we note that the parser and language model architectures we use are highly generic, and do not encode any obvious bias toward natural-language-like word orders.
+The LSTM language model is a generic model of sequence data which is also been used to model financial time series \citep{sirignano2018universal} and purely theoretical chaotic dynamical systems \citep{ogunmolu2016nonlinear}; the neural graph-based parser is simply solving a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
+Nevertheless, it may be the case that a bias toward word order universals is somehow encoded implicitly in the hyperparameters and architectures of these models.
+
+Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers.
+These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.
+
+\subsection{CKY Parsers}
+
+We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of \textcolor{red}{CITE Collins 1997 (Model 1).}
+In our PCFGs, each head independently generates a set of left and right dependents.
+We formulate this as a PCFG where each rule has the form:
+\begin{center}
+	POS$_H$ $\rightarrow$ POS$_H$ POS$_D$
+\end{center}
+for head-initial structures, and
+\begin{center}
+	POS$_H$ $\rightarrow$ POS$_D$ POS$_H$
+\end{center}
+for head-final structures, where each symbol is a POS tag.
+Thus, POS tags act both as terminals and as nonterminals.
+
+We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
+For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm \textcolor{red}{CITE}.
+
+This parsing strategy is very different from the neural graph-based parser:
+While the graph-based parser solves a minimum spanning tree problem, the CKY algorithm uses dynamic programming to compute the exact probabilities of trees given a sentence, as specified by the generative model encoded in the PCFG.
+Second, while the graph-based neural parser uses machine learning to induce syntactic knowledge from data, the CKY parser performs exact probabilistic inference.
+%In this sense, the CKY algorithm does not have any architectural biases in itself.
+%On the other hand, the PCFG makes severely simplifying independence assumptions, compared to the universal approximation capabilities of neural network-based systems.
+
+We used the CKY algorithm to compute the parsing loss $H[T|\mathcal{L}_\theta(T)]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
+Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
+% TODO get UAS for compatibility with other parts of SI
+%We then compared parsability of random and optimized word order grammars on English data.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
+	\caption{Parsability loss $H[T|\mathcal{L}_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).
+We report $H[T|\mathcal{L}_\theta(T)]$ normalized by sentence length.
+\textcolor{red}{make plot nicer}}
+    \label{fig:cky-parser}
+\end{figure}
+
+\subsection{Distorted graph-based parsers}
+
+In this section, we address the idea that the graph-based parser might have a built-in bias toward certain kinds of orderings, and the question whether this might be responsible for our findings.
+In particular, we address the idea that the graph-based parser might have a bias toward parses involving short dependencies, which we call a \key{locality bias}. 
+We address this by changing the order in which the parser sees words, in such a way that the distance between words in the input is not indicative of syntactic relations.
+
+%In particular, this allows us to show tha
+
+%Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization.
+Therefore, we address here the idea that the graph-based parser might have a built-in bias toward parses involving short dependencies, which we call a \key{locality bias}.
+%A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words.
+Suppose that a parser $P$ has a bias toward positing a dependency between a word and the immediate previous word.
+
+\paragraph{Even--odd order.} A sequence of $n$ words originally ordered as $w_1 w_2 w_3 w_4 \cdots w_n$ is reordered by separating the even and odd indices: $w_2 w_4 w_6 \cdots w_{n-1} w_1 w_3 w_5 \cdots w_n$ (assuming $n$ odd).
+Therefore all words that are adjacent in the original order will be separated by a distance of $\approx n/2$ in the distorted order, while all words of distance 2 in the original order will become adjacent.
+
+\paragraph{Interleaving order.} In interleaving ordering, a sequence originally ordered as $w_1 w_2 w_3 \cdots w_n$ is split in half at the middle (index $m=\ceiling{n/2}$), and the two resulting sequences are interleaved, yielding $w_1 w_m w_2 w_{m+1} w_3 w_{m+3} \cdots w_n$.
+Thus all words that were originally adjacent will have distance 2 in the distorted order, with the intervening word coming from a very distant part of the sentence.
+
+\paragraph{Inwards order.} A sequence originally ordered as $w_1 w_2 w_3 \cdots w_{n-1} w_n$ is ordered from the edges of the string inwards, as $w_1 w_n w_2 w_{n-1} \cdots w_{\ceiling{n/2}}$.
+This corresponds to folding the string in on itself once, or equivalently, splitting the sequence in half at the middle, then interleaving the two resulting sequences after reversing the second one.
+The result is that the most non-local possible dependencies in the original order become the most local dependencies in the distorted order.
+
+\paragraph{Sorted order.} A sequence is reordered by sorting by POS tags, and randomizing the order within each block of identical POS tags.
+To each word, we then add a symbol encoding the original position in the sequence.
+For instance
+\begin{center}
+PRON VERB PRON
+\end{center}
+may be reordered as
+\begin{center}
+PRON 1 PRON 3 VERB 2
+\end{center}
+or
+\begin{center}
+PRON 3 PRON 1 VERB 2
+\end{center}
+The numbers are provided to the parser as atomic symbols from a vocabulary ranging from 1 to 200; numbers greater than 200 (which may occur in extremely long sentences) are replaced by an out-of-range token.
+
+The result is that distance between words in the input is not indicative at all of the presence of absence of syntactic relations between them.
+
+\paragraph{Experiments}
+Using English and Japanese data, we trained parsers for the ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
+Resulting parsing accuracy scores are shown in Figure~\ref{fig:distorted-parser}.
+In all settings, the language optimized for efficiency achieved higher parsing accuracy than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
+%Note that absolute values of parsing accuracy vary between the orderings, being higher for the actual orderings than for distorted orderings, in which trees become nonprojective.
+
+\begin{figure}
+    \centering
+    English
+
+    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-english.pdf}
+
+    Japanese
+
+    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
+	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').
+We report $H[T|\mathcal{L}_\theta(T)]$ normalized by sentence length.}
+    \label{fig:distorted-parser}
+\end{figure}
+
+\subsection{n-gram language models}
+
+We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
+In previous work, such as \cite{gildea2015human}, predictability has often been measured using n-gram models.
+
+Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
+Specifically, we constructed bigram models with Kneser-Ney smoothing~\cite{kneser1995improved, chen1999empirical}.
+A bigram model predicts each word taking only the previous word into account.
+This contrasts with LSTMs, which take the entire context into consideration.
+Thus, bigram models and LSTMs stand on opposing ends of a spectrum of language models taking more and more aspects of the context into account.
+
+We estimated language models on the training partitions, and used the validation partitions to estimate surprisal.
+We conducted this for the ten random and the best optimized ordering grammars on English and Japanese data.
+Results (Figure~\ref{fig:bigrams}) show that languages optimized for efficiency are optimal for a bigram language model.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.4]{../results/bigrams/bigrams.pdf} 
+	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).\textcolor{red}{make fig nicer}}
+    \label{fig:bigrams}
+\end{figure}
+
+%\subsection{Discussion}
+%
+%It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.
+
+\section{Effects of data sparsity}
+
+Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
+If the difference between random and real grammars is due to data sparsity, we expect that it decreases as the amount of training data is increased.
+If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference persists as training data is increased.
+
+We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
+We cosidered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
+For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
+Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.
+
+The resulting curves are shown in Figure~\ref{fig:learning-czech}.
+%A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
+A gap in parsing loss of about 0.2 nats appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
+
+\begin{figure}[ht]
+    \centering
+%    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech.pdf} 
+    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech-logloss.pdf}
+
+        \caption{Parsing loss ($H[T|\mathcal{L}_\theta(T)]$, normalized by sentence length) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
+    \label{fig:learning-czech}
+\end{figure}
+
+%\section{Comparing Extracted Grammars to Linguistic Judgments}
+
+% user@user-X510UAR:~/grammar-optim/results/grammars$ git add comparison-table.tex
+
 \section{Languages and Corpus Sizes}
@@
 This suggests that the reduction of dependency length is predicted by efficiency maximization.
-Next, we constructed grammars that minimize average dependency length, using the same gradient descent method as we used for efficiency optimization (Section REF).
+Next, we constructed grammars that minimize average dependency length, using the same gradient descent method as we used for efficiency optimization (Section~\ref{sec:optim-eff}).
 We expect that such grammars should have shorter dependency length than the real grammars, or grammars optimized for efficiency.
@@
 %Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
-%We optimize grammars using the same stochastic gradient descent method as we use for efficiency optimization (Section REF).
-%Indeed, we found that DLM is predicted by efficiency optimization, as discussed in Section REF.
+%We optimize grammars using the same stochastic gradient descent method as we use for efficiency optimization (Section ).
+%Indeed, we found that DLM is predicted by efficiency optimization, as discussed in Section .
 %Results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.

### 2019-05-21 second-draft/pnas-word-order-si.tex
SI

--- 2019-05-18 second-draft/pnas-word-order-si.tex
@@
 % Use the lineno option to display guide line numbers if required.
-
-%\usepackage{lineno}
-%\linenumbers
 
@@
 \DeclareMathOperator{\E}{\mathop{\mathbb{E}}}
-
-\usepackage{footnote}
-\makesavenoteenv{tabular}
-\makesavenoteenv{table}
-
-\usepackage{siunitx}
 
@@
 
+\textcolor{red}{cite rijkhof etc}
+
+\textcolor{red}{TODO This idea makes strong predictions, and turns out to overpredict correlations (see SI, Section X).}
+
 \section{Formalization of Correlations}\label{sec:correlations}
@@
 
-We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.
-%\footnote{Regarding objections by \cite{dunn2011evolved}, we refer to the follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
+We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding objections by \cite{dunn2011evolved}, we refer to the follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
 Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
@@
 
-%\textcolor{blue}{Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
-%Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
-%The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order, finding 16 correlations to likely be universal.
-%Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:corr-dryer}.
-%Testing efficiency on the excluded correlations would be feasible with additional annotation of data.
-%For instance, one of these could be tested by labeling the existing databases for the distinction between manner adverbs from other kinds of adverbs.
-%For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).}
-
-\section{Formalizing Comunicative Efficiency}
-\subsection{Relation to Other Work}
-
-%The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
-%Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication.
-
-%Recall our formalization of efficiency:
-%\begin{equation}
-%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
-%\end{equation}
-%\begin{equation}
-%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
-%\end{equation}
-%\begin{equation}\label{eq:efficiency}
-%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
-%\end{equation}
-
-Here we discuss how our formalization of communicative efficiency relates to other information-theoretic work on language.
-
-Our formalization of efficiency is identical to the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems, with syntactic structures taking the role of underlying meanings that the speaker wants to convey.
-This function is also equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
-Very similar metrics have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
-%More generally, these approaches instantiate Shannon's Rate Distortion theory.
-Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.
-
-All these accounts formalize the amount of information provided as the \key{mutual information} between the forms that a listener receives, and the meanings that the speaker intends to communicate to the speaker.
-Most previous studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} are concerned with the meanings of individual words; they thus take the forms to be individual words, which refer to individual objects and their properties.
-We instead study how words are composed into sentences.
-In our work, the underlying meanings are the syntactic structures; the forms are the linearized sentences communicated by the speaker to the listener.
-By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
-The syntactic structure specifies how the meanings of words are combined; therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
-
-%These information-theoretic accounts differ in the formalizations of complexity and diversity:
-Like our model, \cite{ferreri2003least,ferrericancho2007global,futrell2017memory} model the complexity as the \key{entropy} of the distribution of forms, which arises naturally from the role played by surprisal in determining human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
-This model is closely related to the Information Bottleneck approach \cite{zaslavsky2018efficient, zaslavsky2019semantic}, which models complexity as the mutual information between forms (e.g., words) and their referents (e.g., individual objects).
-This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
-As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure.
-
-A few models have used other complexity measures that are not explicitly information-theoretic quantities.
-In \cite{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \cite{kemp2012kinship} it is the difficulty of defining the concepts that are encoded.
-The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
-Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the information bottleneck formalism, bringing them even closer to our formalization.
-
-In addition to these models, which quantify the efficiency of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context, such as models of pragmatic reasoning in reference games \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
-In line with the other models, these assume that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely}.
-
-\subsection{Possible values of $\lambda$}\label{sec:lambda}
-
-In the efficiency objective
-\begin{equation}\label{eq:efficiency}
-	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
-\end{equation}
-the value of $\lambda$ is constrained to be in $[0,1)$.
+\textcolor{blue}{Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
+Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
+The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order, finding 16 correlations to likely be universal.
+Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:corr-dryer}.
+%Due to data availability, we restricted to 11 universals and further collapsed to the eight correlations shown in Table~\ref{table:corr-dryer}.
+Testing efficiency on the excluded correlations would be feasible with additional annotation of data.
+For instance, one of these could be tested by labeling the existing databases for the distinction between manner adverbs from other kinds of adverbs.
+For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).}
+
+\section{Supplementary Analyses for Study 1}
+
+\paragraph{Relative Efficiency of Real Grammars}
+In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
+We also conducted the analysis using a Binomial test (one-sided), testing whether the real gramar is more efficient than the \emph{median} of random grammars, avoiding any distributional assumption on the random grammars.
+As before, we used Hochberg's step-up procedure (Note that the tests for different languages are independent, as different random grammars are evaluated for each language), with $\alpha=0.05$.
+In this analysis, real grammars improved in parseability for 80\% of languages, in predictability for 69\% of languages, and in either of both in 92\% of languages ($p<0.05$, with Bonferroni correction).
+
+\paragraph{z-Scored Data (Figure (REF))}
+We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section~\ref{sec:lambda}), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
+
+Within each language, we estimate the Pareto curve by connecting the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
+We then estimated the overall average Pareto curve, as shown in Figure (REF), by averaging the z-transformed points (1), (2), (3).
+
+\paragraph{Per-Language Results}
+Figure~\ref{fig:pareto-per-lang}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
+\caption{per language results.
+Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\textcolor{red}{would be good to have same colors as in main paper} All data are $z$-transformed.}\label{fig:pareto-per-lang}
+\end{figure}
+
+\textcolor{red}{also provide p value, CI for how many baselines are better in either of the two dimensions}
+
+\section{Supplementary Analyses for Study 2}
+\subsection{Correlation between Universals and Efficiency}
+
+In Figure~\ref{fig:corr-eff-corr}, we plot efficiency, parseability, and predictability (parseability and predictability $z$-scored within language) as a function of the number of satisfied correlations, for the real grammars of the 51 languages.
+
+We found very similar results using Spearman's rank correlation (Efficiency: $\rho=0.59$, $p = 9.8 \cdot 10^{-6}$; Parseability: $\rho=0.55$, $p=4.7 \cdot 10^{-5}$; Predictability: $\rho=0.36$, $p=0.012$).
+
+\begin{figure}[ht]
+    \centering
+
+    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
+    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
+    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
+
+	\caption{Correlation between the number of satisfied correlations ($x$-axis) and efficiency, parseability, and predictability ($y$-axis), for the 51 real languages.}
+    \label{fig:corr-eff-corr}
+\end{figure}
+
+\subsection{Predictions for Individual Languages}
+
+We show results on the level of individual languages in Figure~\ref{fig:per-lang}.
+
+We obtained predictions for individual languages and each of the eight relations as follows.
+For each language and each of the objective functions (efficiency, predictability, parseability), we considered the optimized grammar that yielded the best value of this objective function among the eight optimized grammars (i.e., the grammar where the optimization procedure had been most successful).
+We interpreted this grammar as verb-object or object-verb depending on the order in the real grammar of the language.
+
+\begin{figure} 
+	\begin{center}
+	\includegraphics[width=0.7\textwidth]{../results/correlations/figures/pred-eff-pred-pars-families-2.pdf} % figures/pred-eff-families-2.pdf
+	\end{center}
+	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by optimizing for efficiency, predictability, parseability (right).
+Dark blue: Verb patterner \emph{precedes} object patterner (English, Arabic, ...).
+Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).
+White cells indicate that the relation is not annotated in the dataset for the given language.}\label{fig:per-lang}
+%	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
+Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
+Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...) \textcolor{red}{add other functions to this plot}.}\label{fig:per-lang}
+\end{figure}
+
+\subsection{Regression for Predicted Correlations}
+
+%https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
+
+Following the recommendations of \cite{ghosh2018use, burkner2018advanced}, we used as a very weakly informative prior a Student's $t$ prior with $\nu=3$ degrees of freedom, mean 0, and scale $\sigma=10$ (i.e., the PDF $p$ is $\frac{1}{\sigma} p_3(x/\sigma)$, where $p_3$ is the PDF of the $t$-distribution with $\nu=3$).
+Note that a correlation that holds in 90\% of cases corresponds to an intercept $\alpha \approx 2.19$ in the logistic model, well within the main probability mass of the prior.
+The covariance matrix is modeled using the LKJ prior, as described in \cite{burkner2018advanced}.
+We used four chains, with 5000 samples each (2500 warmup).
+The chains had converged ($\hat{R} < 1.01$).
+
+We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
+
+\subsection{Comparing Efficiency to its Components}\label{sec:posterior-number}
+
+In Figure~\ref{fig:posterior}, we plot the posterior number of correlations predicted to hold in most optimized grammars, as obtained from the Bayesian regression.
+In addition to grammars optimized for efficiency, we also report the result for grammars optimized for predictability and for parseability alone.
+Efficiency predicts all eight correlations with high posterior probability; predictability and parseability alone do not.
+
+\begin{figure}[ht]
+	\begin{center}
+	\includegraphics[width=0.98\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
+	\end{center}
+	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
+\end{figure}
+
+%\begin{figure}
+%    \centering
+%    
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
+%    
+%    
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
+%	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
+%    \label{fig:posterior-prevalence}
+%\end{figure}
+
+\subsection{Results on all UD Relations}
+In this section, we provide the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
+We also report results for grammars optimized for predictability and parseability individually.
+
+We considered all UD syntactic relations occurring in at least two of the 51 languages.
+In Table~\ref{tab:all-predictions-1}, we present the data for the eight correlations discussed in the main paper, and for those other relations for which the typological literature provides data.
+Additionally, in Table~\ref{tab:all-predictions-1} we present data for the other UD relations, for which either no typological data is available, or which are not linguistically meaningful.
+
+\textcolor{red}{at some place also say the following: Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
+}
+
+Further, we report results for grammars optimized for dependency length minimization (DLM).
+Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
+Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
+Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
+In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
+
+%posterior_perRelation_Real_conj.pdf         posterior_perRelation_Real_goeswith.pdf     posterior_perRelation_Real_nummod.pdf       
+%user@user-G751JY:~/grammar-optim/results/correlations$ evince figures/posteriors/posterior_perRelation_Real_csubj.pdf 
+%user@user-G751JY:~/grammar-optim/results/correlations$
+
+\begin{table*} % TODO figure out away to make this table more aesthetic
+	\begin{center}	
+\small{
+\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
+	\hline
+&	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+% From Dryer	
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+%\hdashline
+
+\hline
+% Other universals, not in Dryer
+&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
+&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf.
+\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
+&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf.
+\cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
+&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section~\ref{sec:correlations} \\
+&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
+&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+\end{tabular}
+}
+
+\end{center}
+\caption{Predictions on UD relations with predictions from the typological literature.
+The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
+In the last column, we indicate what direction would be expected typologically.}
+\label{tab:all-predictions-1}
+\end{table*}
+
+\footnotetext{Note that we report an \emph{anti-correlation}.
+The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
+Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
+}
+
+\begin{table*} % TODO figure out away to make this table more aesthetic
+	\begin{center}	
+\small{
+\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
+	\hline
+	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+\hline
+appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\
+lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\
+expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\
+iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\
+vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\
+
+\hline
+% Uninterpretable
+compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\
+det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\
+dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\
+dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
+advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\
+
+\hline
+% UD artifacts
+conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
+discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\
+fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\
+flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
+goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
+list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
+orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\
+parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\
+reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\
+
+ \hline
+%comp. &
+%\multirow{3}{*}{NP}&
+%	AP &
+    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+\end{tabular}
+}
+
+\end{center}
+\caption{Predictions on UD relations for which no predictions are available in the typological literature.
+``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
+``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
+\label{tab:all-predictions-2}
+\end{table*}
+
+\section{Possible values of $\lambda$}\label{sec:lambda}
+
+In the efficiency objective~(\ref{eq:efficiency-again}), the value of $\lambda$ is constrained to be in $[0,1)$.
 This means, surprisal must be weighted less strongly than parsability.
@@
 While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
-In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parsability, whereas values closer to $0$ diminish the role of predictability.
 
 In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
+
 The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
-%We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations, see Table REF.
+We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations.
 We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (Table~\ref{tab:all-predictions-1}).
-%Indeed, the same holds for all of the 38 syntactic relations in UD  (Table~\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})
-%We additionally report results with $\lambda = 0$, i.e., placing all weight on parseability.
-That is, at $\lambda$ close to its maximal value, the predictions of optimizing the two scoring functions individually add up to the predictions of efficiency optimization.
-Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}).
-We also verified the robustness of our results to small variations of $\lambda$, by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu}.
+
+%\section{Joint optimization of grammar, parser, and language model parameters}
+\section{Creating Optimized Grammars}
+
+In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
+We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
+More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
+This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.
+
+This method addresses a major challenge noted in previous work optimizing grammars, namely that the predictability (and parseability) of an individual sentence depends on the entire distribution of the language.
+Previously, \citep{gildea2015human} optimized grammars for dependency length and trigram surprisal using a simple hill-climbing method on the grammar parameters, which required reestimating the trigram surprisal model in every iteration.
+Such a method would be computationally prohibitive for efficiency optimization, as it would require reestimating the neural network models after every change to the grammar, which would amount to reestimating them hundreds or thousands of times per grammar.
+Our method, by allowing for the use of stochastic gradient descent, addresses this challenge, as we describe in Section~\ref{sec:optim-eff}.
+
+\subsection{Differentiable Ordering Grammars}\label{sec:diff-gramm}
+
+We extended the parameter space of grammars by continuously interpolating between grammars, making efficiency a \emph{differentiable} function of grammar parameters.
+The parameters of such a \key{differentiable word order grammar} are as follows. 
+For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
+Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
+Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
+184) without replacement.
+
+If $a_\tau \in \{0, 1\}$, and the distances between values of $b_\tau$ (for different $\tau$) become very large, such a differentiable grammar becomes deterministic, assigning almost full probability to exactly one ordering for each syntactic structure.
+In this case, the grammar can be converted into an equivalent grammar of the form described in Materials and Methods, by extracting a single parameter in $[-1, 1]$ for each relation $\tau$.
+
+\subsection{Extracting Grammars from Datasets}
+We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
+To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
+We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.
+
+\subsection{Optimizing Grammars for Efficiency}\label{sec:optim-eff}
+
+In this section, we describe how we optimized grammar parameters for efficiency.
+A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
+More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $\mathbf{w}$.
+In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $p_{\mathcal{L}_\theta}(u|t)$ concentrates on a single ordering $u$.
+
+Recall the definition of efficiency
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+\end{equation}
+where 
+\begin{equation}
+	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
+\end{equation}
+\begin{equation}
+	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+\end{equation}
+where $t \sim \mathcal{T}$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
+
+These quatities are estimated using two neural models, as described in Section X:
+A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.
+The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is
+\begin{equation}
+	R_{Pars}^{\phi}(u,t) =  \log \frac{p_\phi(t|u)}{p(t)}
+\end{equation}
+%and we can write
+%\begin{equation}
+%	R_{Pars}^\phi := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
+%\end{equation}
+%
+A \key{language model}, with some parameters $\psi$, calculates the word-by-word surprisal of an utterance:
+\begin{equation}
+	R_{Pred}^{\psi}(u) = - \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})
+\end{equation}
+Using this, we can rewrite Efficiency~(\ref{eq:efficiency}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
+\begin{equation}
+	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi} := \max_{\phi,\psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\end{equation}
+%
+%The parsers and language models appropriate to a tree distribution $\mathcal{T}$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
+%\begin{equation}
+%\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation}
+%\begin{equation}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation}
+%
+%Using these models, we can rewrite 
+%\begin{equation}
+%R_{Pars}^{\psi(\theta)}= \E_t \E_{u \sim p_\theta(u|t)} R_{Pars}^{\psi(\theta)}(u,t) 
+%\end{equation}
+%where $R_{Pars}(u,t) = \log \frac{p_\phi(t|u)}{p(t)}$, and $p_\theta(u|t)$ is the probability assigned to sentence $u$ by the distribution $\mathcal{L}_\theta(t)$ defined by the differentiable word order grammar $\mathcal{L}_\theta$,
+%and
+%\begin{equation}
+%R_{Pred} = \E_t \E_{u \sim p_\theta(u|t)} R_{Pred}(u)
+%\end{equation}
+%where $R_{Pred}(u) = \log p_\psi(u)$.
+%Therefore, we can rewrite efficiency as
+%\begin{equation}\label{eq:efficiency}
+%R_{\textit{Eff}}^\theta := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi(\theta)}(u,t) + \lambda R_{Pred}^{\psi(\theta)}(u)\right]
+%\end{equation}
+%Directly applying gradient descent to this expression is hard due to the complicated dependence of $\phi(\theta), \psi(\theta)$ on $\theta$.
+%
+%However, we can rewrite
+%\begin{equation}\label{eq:efficiency}
+%\argmax_\theta\	R_{\textit{Eff}}^\theta 
+%\end{equation}
+%equivalently into
+In order to find an optimal grammar $\theta$, we thus need to solve
+\begin{equation}\label{eq:efficiency}
+\argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} %= \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\end{equation}
+Importantly, $R_{\textit{Eff}}^{\theta, \phi, \psi}$ is differentiable in $\theta, \phi, \psi$: %, and we can apply stochastic gradient descent to carry out this optimization.
+\begin{align}
+\partial_\theta R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)} \left[  \left[\partial_\theta \log p_\theta(u|t)\right] \cdot    \left(R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right) \right] \label{eq:dtheta}\\
+\partial_\phi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\partial_\phi R_{Pars}^{\phi}(u,t)\right] \\
+\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}
+\end{align}
+where \ref{eq:dtheta} is derived using the \emph{score-function} or \emph{REINFORCE} trick~\cite{williams1992simple}.
+Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures (CITE).
+
+We can therefore apply stochastic gradient descent to jointly optimize $\theta, \phi, \psi$:
+In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim p_{\theta}(\cdot|t)$.
+Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
+do a gradient descent step using the estimator given by the expressions in the square brackets in \ref{eq:dtheta}-\ref{eq:dpsi}.
+
+Optimizing for only parseability (or predictability) is very similar -- in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.
+
+At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
+
+See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
+
+%
+%%(analogous for $R_{Pars}$ instead of $R_{Pred}$)
+%\begin{equation}\label{eq:estimator}
+%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
+%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
+%\end{equation}
+%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
+%for the gradient
+%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
+%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
+%
+
+%
+%
+%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
+Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
+%\begin{equation}
+%\label{eq:exp-over-utts}
+%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
+%\end{equation}
+%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
+In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
+%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
+%
+%
+%
+%We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
+%\begin{equation*}
+%R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
+%\end{equation*}
+%where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
+%\begin{equation*}
+%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation*}
+%
+%
+%We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
+Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
+%\begin{align}
+%\nonumber
+%R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
+%\nonumber
+%&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
+%\end{align}
+%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
+%\begin{equation*}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation*}
+%
+%For efficiency, we linearly combine the two terms:
+%\begin{equation}
+%\label{eq:combined-score}
+%R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
+%\end{equation}
+%
+%
+%%\begin{equation}
+%%    \label{eq:depl}
+%%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+%%\end{equation}
+%%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
+%
+%For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
+The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
+The optimization problems come out to:
+%\begin{align}
+%%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
+%\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
+%\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
+%\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
+%\end{align}
+%Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
+%
+%
+%%The case of predictability, parseability, and efficiency are more challenging, as 
+%%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
+%We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
+%Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
+%
+%We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section (REF)).
+%
+%In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
+%Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
+%do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
+%\begin{equation}\label{eq:estimator}
+%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
+%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
+%\end{equation}
+%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
+%for the gradient
+%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
+%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
+%
+%See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
+%
+
+%The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
+A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
+%We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
+
+%In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
+%and $\ell_2$ regularization.
+
+%A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
+%Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
+%We thus do not evaluate this objective function here.
+
+%We extract grammars for the actual languages using a Bayesian data analysis method.
+%We specify a simple Bayesian prior over differentiable ordering grammars:
+%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
+%for each $\tau$.
+%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
+%We extract
+
+\section{Neural Network Architectures}\label{sec:neural-architectures}
+
+In this section, we describe the details of the neural network architectures.
+All choices described in this section were fixed before evaluating word order properties and the efficiency of real language.
+
+%\paragraph{Locality}
+%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
+
+\paragraph{Estimating Predictability}
+We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
+Given the small size of the corpora, this limit is only attained only for few languages.
+In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
+The model predicts both the next word and its POS tags in each step.
+Using POS tags is intended to prevent overfitting on small corpora.
+These decisions were made before evaluating word order properties.
+
+\paragraph{Estimating Parseability}
+We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}.
+This architecture is remarkably simple: the words of a sentence are encoded into context-sensitive embeddings using bidirectional LSTMs, then a classifier is trained to predict the head for each work.
+The classifier works by calculating a score for every pair of word embeddings $(w_i, w_j)$, indicating the likelihood that the $j$th word is the head of the $i$th word.
+This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
+
+To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags.
+Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged; parsing from POS tags prevents early overfitting.
+This decision was made before evaluating word order properties.
+
+\paragraph{Hyperparameters}
+
+%\textbf{TODO study 1 only has 1 set of parameters}
+
+Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
+For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
+Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (predictability or parseability) on the language.
+We then used this setting for creating optimized word order grammars.
+
+\textcolor{red}{TODO random seeds for efficiency}
+
+All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
+We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
+
+%\begin{table}[]
+%    \centering
+%    \begin{tabular}{c|cccc}
+%    Entropy penalty     &  0.0001, 0.001 \\
+%    Learning Rate (Word order grammar)     & 5e-6, 1e-5, 2e-5, 5e-5 \\
+%    Momentum (Word order grammar) & 0.8, 0,9 \\
+%    Dropout (Language model) & 0.0, 0.3, 0.5 \\
+%    
+%    \end{tabular}
+%    \caption{Hyperparameters}
+%    \label{tab:my_label}
+%\end{table}
+
+\paragraph{Optimization details}
+
+We employ two common variance reduction methods to improve the estimator~(\ref{eq:dtheta}), while keeping it unbiased.
+%For dependency length and 
+For predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word (and its head, in the case of dependency length).
+We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:dtheta}) with lower variance.
+Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple}.
+These two methods reduce the variance of the estimator and thereby increase the speed of optimization and reduce training time, without biasing the results.
+For numerical stability, we represent $a_\tau \in [0,1]$ via its logit $\in \mathbb{R}$.
+Furthermore, to encourage exploration of the parameter space, we add an entropy regularization term \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$.
+The weight of the entropy regularization was chosen together with the other hyperparameters.
+%, adding the negative entropy of the $Bernoulli(a_\tau)$ to the loss $R$. %, weighted with $\alpha$.
+
+These techniques for improving (\ref{eq:dtheta}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
+
+We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
+For the language model parameters $\phi$, we use plain SGD. 
+For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
+The learning rates and other optimization hyperparameters were determined together with the other hyperparameters.
+
+For each language and objective function, we apply Early Stopping~\citep{prechelt1998early} as follows.
+We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
+This technique, well known in the machine learning literature, helps prevent overfitting to the specific corpora used.
+
+The choice of optimization methods and the stopping criterion were fixed before we investigated word order correlations.
+
+\paragraph{Optimized Grammars}
+
+Recall that, for each language, we created 8 optimized languages for each optimization criterion.
+We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
+This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.
+
+For efficiency optimization, for each grammar, we ran efficiency optimization with three different random seeds, selecting among these the seed that yielded the best overall efficiency value.
+This
+
+ % TODO we could use some language at the top of this section on the general guiding principles for all these decisions.
+Were these implementation decisions made before or after looking at the word order universals results?
+If they were made before, we should highlight that.
+
+\section{Robustness to different language models and parsers}
+
+Here we take up the question of the extent to which our results are dependent on the particular parser and language model used in the optimization process.
+We want to know: when we optimize a word order grammar for efficiency, have we produced a language which is highly efficient \emph{in general}, or one which is highly efficient \emph{for a specific parser}?
+We wish to argue that natural language syntax is optimized for efficiency in general, meaning that syntactic trees are highly recoverable from word orders in principle.
+If it turns out that our optimized languages are only optimal for a certain parser from the NLP literature, then we run the risk of circularity: it may be that the reason this parser was successful in the NLP literature was because it implicitly encoded word order universals in its inductive biases, and thus it would be no surprise that languages which are optimized for parseability also show those universals.
+
+In this connection, we note that the parser and language model architectures we use are highly generic, and do not encode any obvious bias toward natural-language-like word orders.
+The LSTM language model is a generic model of sequence data which is also been used to model financial time series \citep{sirignano2018universal} and purely theoretical chaotic dynamical systems \citep{ogunmolu2016nonlinear}; the neural graph-based parser is simply solving a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
+Nevertheless, it may be the case that a bias toward word order universals is somehow encoded implicitly in the hyperparameters and architectures of these models.
+
+Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers.
+These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.
+
+\subsection{CKY Parsers}
+
+We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of Collins 1997 (Model 1).
+In our PCFGs, each head independently generates a set of left and right dependents.
+We formulate this as a PCFG where each rule has the form:
+\begin{center}
+	POS$_H$ $\rightarrow$ POS$_H$ POS$_D$
+\end{center}
+for head-initial structures, and
+\begin{center}
+	POS$_H$ $\rightarrow$ POS$_D$ POS$_H$
+\end{center}
+for head-final structures, where each symbol is a POS tag.
+Thus, POS tags act both as terminals and as nonterminals.
+
+We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
+For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm.
+
+This parsing strategy is very different from the neural graph-based parser:
+While the graph-based parser solves a minimum spanning tree problem, the CKY algorithm uses dynamic programming to compute the exact probabilities of trees given a sentence, as specified by the generative model encoded in the PCFG.
+Second, while the graph-based neural parser uses machine learning to induce syntactic knowledge from data, the CKY parser performs exact probabilistic inference.
+%In this sense, the CKY algorithm does not have any architectural biases in itself.
+%On the other hand, the PCFG makes severely simplifying independence assumptions, compared to the universal approximation capabilities of neural network-based systems.
+
+We used the CKY algorithm to compute the parsing loss $H[T|L_\theta(T)]$ on the validation partition of the English UD corpora, for random and optimized ordering grammars.
+Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
+% TODO get UAS for compatibility with other parts of SI
+%We then compared parsability of random and optimized word order grammars on English data.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
+	\caption{Parsability loss $H[T|L_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).}
+    \label{fig:cky-parser}
+\end{figure}
+
+\subsection{Distorted graph-based parsers}
+
+In this section, we address the idea that the graph-based parser might have a built-in bias toward certain kinds of orderings, and the question whether this might be responsible for our findings.
+In particular, we address the idea that the graph-based parser might have a bias toward parses involving short dependencies, which we call a \key{locality bias}. 
+We address this by changing the order in which the parser sees words, in such a way that the distance between words in the input is not indicative of syntactic relations.
+
+%In particular, this allows us to show tha
+
+%Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization.
+Therefore, we address here the idea that the graph-based parser might have a built-in bias toward parses involving short dependencies, which we call a \key{locality bias}.
+%A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words.
+Suppose that a parser $P$ has a bias toward positing a dependency between a word and the immediate previous word.
+
+\paragraph{Even--odd order.} A sequence of $n$ words originally ordered as $w_1 w_2 w_3 w_4 \cdots w_n$ is reordered by separating the even and odd indices: $w_2 w_4 w_6 \cdots w_{n-1} w_1 w_3 w_5 \cdots w_n$ (assuming $n$ odd).
+Therefore all words that are adjacent in the original order will be separated by a distance of $\approx n/2$ in the distorted order, while all words of distance 2 in the original order will become adjacent.
+
+\paragraph{Interleaving order.} In interleaving ordering, a sequence originally ordered as $w_1 w_2 w_3 \cdots w_n$ is split in half at the middle (index $m=\ceiling{n/2}$), and the two resulting sequences are interleaved, yielding $w_1 w_m w_2 w_{m+1} w_3 w_{m+3} \cdots w_n$.
+Thus all words that were originally adjacent will have distance 2 in the distorted order, with the intervening word coming from a very distant part of the sentence.
+
+\paragraph{Inwards order.} A sequence originally ordered as $w_1 w_2 w_3 \cdots w_{n-1} w_n$ is ordered from the edges of the string inwards, as $w_1 w_n w_2 w_{n-1} \cdots w_{\ceiling{n/2}}$.
+This corresponds to folding the string in on itself once, or equivalently, splitting the sequence in half at the middle, then interleaving the two resulting sequences after reversing the second one.
+The result is that the most non-local possible dependencies in the original order become the most local dependencies in the distorted order.
+
+\paragraph{Sorted order.} A sequence is reordered by sorting by POS tags, and randomizing the order within each block of identical POS tags.
+To each word, we then add a symbol encoding the original position in the sequence.
+For instance
+\begin{center}
+PRON VERB PRON
+\end{center}
+may be reordered as
+\begin{center}
+PRON 1 PRON 3 VERB 2
+\end{center}
+or
+\begin{center}
+PRON 3 PRON 1 VERB 2
+\end{center}
+The numbers are provided to the parser as atomic symbols from a vocabulary ranging from 1 to 200; numbers greater than 200 (which may occur in extremely long sentences) are replaced by an out-of-range token.
+
+The result is that distance between words in the input is not indicative at all of the presence of absence of syntactic relations between them.
+
+\paragraph{Experiments}
+Using English and Japanese data, we trained parsers for the ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
+Resulting parsing accuracy scores are shown in Figure~\ref{fig:distorted-parser}.
+In all settings, the language optimized for efficiency achieved higher parsing accuracy than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
+%Note that absolute values of parsing accuracy vary between the orderings, being higher for the actual orderings than for distorted orderings, in which trees become nonprojective.
+
+\begin{figure}
+    \centering
+    English
+
+    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-english.pdf}
+
+    Japanese
+
+    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
+	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').}
+    \label{fig:distorted-parser}
+\end{figure}
+
+\subsection{n-gram language models}
+
+We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
+In previous work, such as \cite{gildea2015human}, predictability has often been measured using n-gram models.
+
+Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
+Specifically, we constructed bigram models with Kneser-Ney smoothing~\cite{kneser1995improved, chen1999empirical}.
+A bigram model predicts each word taking only the previous word into account.
+This contrasts with LSTMs, which take the entire context into consideration.
+Thus, bigram models and LSTMs stand on opposing ends of a spectrum of language models taking more and more aspects of the context into account.
+
+We estimated language models on the training partitions, and used the validation partitions to estimate surprisal.
+We conducted this for the ten random and the best optimized ordering grammars on English and Japanese data.
+Results (Figure~\ref{fig:bigrams}) show that languages optimized for efficiency are optimal for a bigram language model.
+
+\begin{figure}
+    \centering
+    \includegraphics[scale=.4]{../results/bigrams/bigrams.pdf} 
+	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).}
+    \label{fig:bigrams}
+\end{figure}
+
+%\subsection{Discussion}
+%
+%It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.
+
+\section{Effects of data sparsity}
+
+Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
+If the difference between random and real grammars is due to data sparsity, we expect that it decreases as the amount of training data is increased.
+If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference persists as training data is increased.
+
+We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
+We cosidered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
+For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
+Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.
+
+The resulting curves are shown in Figure~\ref{fig:learning-czech}.
+%A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
+A gap in parsing loss of about 0.2 nats appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
+
+\begin{figure}[ht]
+    \centering
+%    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech.pdf} 
+    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech-logloss.pdf}
+
+        \caption{Parsing loss ($H[T|\mathcal{L}_\theta(T)]$) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
+    \label{fig:learning-czech}
+\end{figure}
+
+\section{Comparing Extracted Grammars to Linguistic Judgments}
+
+% user@user-X510UAR:~/grammar-optim/results/grammars$ git add comparison-table.tex
+
+We validated our grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
+Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS.
+In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb-object relation.
+
+WALS has some data for all but seven languages.\footnote{Note that Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
+In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}
+WALS has data for 74\% of the entries, and lists a dominant order for 91\% of these.
+The grammars we extracted from the corpora agree with WALS in 96~\% out of those cases where WALS records a dominant word order.
+
+\begin{table}[ht]
+\small{
+\begin{center}
+\begin{tabular}{l||ll|ll|ll|ll|ll|ll|llllll}
+%	   for(dep in c("obj", "lifted_mark", "acl", "nmod", "obl")) {
+		   Language 
+		   &	\multicolumn{2}{|c|}{Objects} 
+		   &	\multicolumn{2}{|c|}{Adpositions} 
+		   &	\multicolumn{2}{|c|}{Compl.} 
+		   &	\multicolumn{2}{|c|}{Rel.Cl.} 
+		   &	\multicolumn{2}{|c|}{Genitive} 
+		   &	\multicolumn{2}{|c|}{PP}  \\ \hline\hline
+	\input{../results/grammars/comparison-table.tex}
+\end{tabular}
+\end{center}
+}
+\caption{Comparing grammars extracted from databases to linguistic judgments in the World Atlas of Linguistic Structures.
+For each of the six syntactic relation, the first column provides the ordered coded in the extracted grammar; the second column provides the order coded in WALS (DH for dependent-head, HD for head-dependent order).
+`?' indicates that WALS has no data.
+$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur at similar rates in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
+Finally, `--' indicates that the relation does not occur in the corpus.
+}\label{tab:grammars-wals}
+\end{table}
+
+\section{Efficiency and Prior Formalizations}
+
+%The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
+%Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication.
+
+%Recall our formalization of efficiency:
+%\begin{equation}
+%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+%\end{equation}
+%\begin{equation}
+%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+%\end{equation}
+%\begin{equation}\label{eq:efficiency}
+%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+%\end{equation}
+
+Here we discuss how our formalization of efficiency relates to other information-theoretic work on language.
+
+Our objective function is identical to the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems, with syntactic structures taking the role of underlying meanings that the speaker wants to convey.
+This function is also equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
+
+Very similar metrics have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
+Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.
+
+All these accounts formalize the amount of information provided as the \key{mutual information} between the forms that a listener receives, and the meanings that the speaker intends to communicate to the speaker.
+Most previous studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} are concerned with the meanings of individual words; they thus take the forms to be individual words, which refer to individual objects and their properties.
+We instead study how words are composed into sentences.
+In our work, the underlying meanings are the syntactic structures; the forms are the linearized sentences communicated by the speaker to the listener.
+By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
+The syntactic structure (including the labels on the arcs) specifies how the meanings of words are combined.
+Therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
+
+%These information-theoretic accounts differ in the formalizations of complexity and diversity:
+Like our model, \cite{ferreri2003least,ferrericancho2007global,futrell2017memory} model the complexity as the \key{entropy} of the distribution of forms, which arises naturally from the role played by surprisal in determining human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+This model is closely related to the Information Bottleneck approach \cite{zaslavsky2018efficient, zaslavsky2019semantic}, which models complexity as the mutual information between forms (e.g., words) and their referents (e.g., individual objects).
+This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
+As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure.
+
+A few models have used other measures that are not direct information-theoretic quantities.
+In \cite{regier2007color, xu2014numeral, xu2016historical}, the cost function is the number of different forms; in \cite{kemp2012kinship} it is the complexity of the concepts encoded.
+The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
+Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the information bottleneck formalism, bringing them even closer to our formalization.
+
+In addition to these models, which quantify the efficiency of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context in models of pragmatic reasoning in reference games \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
+In line with the other models, these assume that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely}.
 
@@
 
-%\section{Relation to the Information Bottleneck}
-%
-%Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
-%\begin{equation}
-%    \label{eq:efficiency-again}
-%    J_T(L) = I[L(T); T] - \lambda H[L(T)].
-%\end{equation}
-%This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}.
+\section{Relation to the Information Bottleneck}
+
+Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
+\begin{equation}
+    \label{eq:efficiency-again}
+    J_T(L) = I[L(T); T] - \lambda H[L(T)].
+\end{equation}
+This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}.
 In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information.
@@
 Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
-%\begin{equation}
-%    \label{eq:ib}
-%    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
-%\end{equation}
-%In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
-%
-%In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$.
+\begin{equation}
+    \label{eq:ib}
+    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
+\end{equation}
+In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
+
+In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$.
 To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$.
@@
 Then our objective is parallel to Eq.~\ref{eq:ib}:
-%\begin{equation}
-%    \label{eq:efficiency-kinda-ib}
-%    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
-%\end{equation}
-%Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$.
+\begin{equation}
+    \label{eq:efficiency-kinda-ib}
+    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
+\end{equation}
+Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$.
 Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
-%\begin{equation*}
-%    J_T(L) = I[U;T] - \lambda H[U].
-%\end{equation*}
-%
-%Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
-%
-%
-
-\section{Supplementary Analyses for Study 1}
-
-In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.\footnote{Note that in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
-Due to use of stochastic gradient descent in optimization, there is some amount of noise in the optimization process.}
-
-We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section~\ref{sec:lambda}), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
-
-Within each language, we estimate the Pareto curve based on the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
-%We then estimated the overall average Pareto curve, as shown in Figure (REF), by averaging the z-transformed points (1), (2), (3).
-
-Figure (REF) in the main paper shows the average of these per-language plots. \textcolor{red}{unify pareto curve computation in main paper}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
-\caption{Predictability and parseability of 51 languages.
-Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.
-All data are $z$-scored.}\label{fig:pareto-per-lang}
-\end{figure}
-
-In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
-We also conducted the analysis using a Binomial test (one-sided), testing whether the real gramar is more efficient than the \emph{median} of random grammars, avoiding any distributional assumption on the random grammars.
-As before, we used Hochberg's step-up procedure (Note that the tests for different languages are independent, as different random grammars are evaluated for each language), with $\alpha=0.05$.
-In this analysis, real grammars improved in parseability for 80\% of languages, in predictability for 69\% of languages, and in either of both in 92\% of languages ($p<0.05$, with Bonferroni correction).
-In Table~\ref{fig:pareto-per-lang-stats}, we provide per-language results for the $t$-tests and binomial tests.
-
-\begin{table}
-\centering
-\small{
-\begin{tabular}{l||ll|lll|lll}
-Language & Pred. (t) & Parse. (t) & \multicolumn{3}{|c|}{Pred. (Binomial)} & \multicolumn{3}{|c|}{Parseab. (Binomial)} \\ 
-&  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
-\input{../results/plane/per-language-statistics.tex}
-\end{tabular}
-}
-\caption{Per-language results in Study 1.
-For each language, we show the following: (1) $p$-values obtained from a one-sided $t$-test, for the null that the mean predictability/parseability of random grammars is at least as high as that of the real grammar.
-(2) Results from one-sided binomial tests, for the null that the the real grammar is better than at most $50 \%$ of random grammars.
-In addition to the $p$-value, we report point estimate and 95\% confidence interval for the fraction of random grammars that have values below real grammars.}\label{fig:pareto-per-lang-stats}
-\end{table}
-
-\section{Supplementary Analyses for Study 2}
-\subsection{Correlation between Universals and Efficiency}
-
-In Figure~\ref{fig:corr-eff-corr}, we plot efficiency, parseability, and predictability (parseability and predictability $z$-scored within language) as a function of the number of satisfied correlations, for the real grammars of the 51 languages.
-
-We found very similar results using Spearman's rank correlation (Efficiency: $\rho=0.59$, $p = 9.8 \cdot 10^{-6}$; Parseability: $\rho=0.55$, $p=4.7 \cdot 10^{-5}$; Predictability: $\rho=0.36$, $p=0.012$).
-
-\begin{figure}[ht]
-    \centering
-
-    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
-    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
-    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
-
-	\caption{Correlation between the number of satisfied correlations ($x$-axis) and efficiency, parseability, and predictability ($y$-axis), for the 51 real languages.}
-    \label{fig:corr-eff-corr}
-\end{figure}
-
-\subsection{Predictions for Individual Languages}
-
-We show predictions for the eight correlations on the level of individual languages in Figure~\ref{fig:per-lang}.
-
-We obtained predictions for individual languages and each of the eight relations as follows.
-For each language and each of the objective functions (efficiency, predictability, parseability), we considered the optimized grammar that yielded the best value of this objective function among the eight optimized grammars (i.e., the grammar where the optimization procedure had been most successful).
-We interpreted this grammar as verb-object or object-verb depending on the order in the real grammar of the language.
-
-\begin{figure} 
+\begin{equation*}
+    J_T(L) = I[U;T] - \lambda H[U].
+\end{equation*}
+
+Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
+
+\section{Relation to DLM}
+
+DLM turns out to be predicted by our model
+
+\section{Other (Preregistered) Experiments}
+
+Here we report the results of prior preregistered simulation runs.
+
+\begin{table}[ht]
 	\begin{center}
-	\includegraphics[width=0.7\textwidth]{../results/correlations/figures/pred-eff-pred-pars-families-2.pdf} % figures/pred-eff-families-2.pdf
-	\end{center}
-	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by optimizing for efficiency, predictability, parseability (right).
-Dark blue: Verb patterner \emph{precedes} object patterner (English, Arabic, ...).
-Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).
-White cells indicate that the relation is not annotated in the dataset for the given language.}\label{fig:per-lang}
-%	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
-Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
-Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...) \textcolor{red}{add other functions to this plot}.}\label{fig:per-lang}
-\end{figure}
-
-\subsection{Regression for Predicted Correlations}
-
-%https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
-
-Following the recommendations of \cite{ghosh2018use, burkner2018advanced}, we used as a very weakly informative prior a Student's $t$ prior with $\nu=3$ degrees of freedom, mean 0, and scale $\sigma=10$ (i.e., the PDF $p$ is $\frac{1}{\sigma} p_3(x/\sigma)$, where $p_3$ is the PDF of the $t$-distribution with $\nu=3$).
-Note that a correlation that holds in 90\% of cases corresponds to an intercept $\alpha \approx 2.19$ in the logistic model, well within the main probability mass of the prior.
-The covariance matrix is modeled using the LKJ prior, as described in \cite{burkner2018advanced}.
-We used four chains, with 5000 samples each (2500 warmup).
-The chains had converged ($\hat{R} < 1.01$).
-
-We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
-
-\subsection{Comparing Efficiency to its Components}\label{sec:posterior-number}
-
-In Figure~\ref{fig:posterior}, we plot the posterior number of correlations predicted to hold in most optimized grammars, as obtained from the Bayesian regression.
-In addition to grammars optimized for efficiency, we also report the result for grammars optimized for predictability and for parseability alone.
-Efficiency predicts all eight correlations with high posterior probability; predictability and parseability alone do not.
-
-\begin{figure}[ht]
-	\begin{center}
-	\includegraphics[width=0.98\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
-	\end{center}
-	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
-\end{figure}
-
-%\begin{figure}
-%    \centering
-%    
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
-%    
-%    
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
-%	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
-%    \label{fig:posterior-prevalence}
-%\end{figure}
-
-\subsection{Results on all UD Relations}
-In this section, we provide the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
-We also report results for grammars optimized for predictability and parseability individually.
-
-We considered all UD syntactic relations occurring in at least two of the 51 languages.
-In Table~\ref{tab:all-predictions-1}, we present the data for the eight correlations discussed in the main paper, and for those other relations for which the typological literature provides data.
-Additionally, in Table~\ref{tab:all-predictions-1} we present data for the other UD relations, for which either no typological data is available, or which are not linguistically meaningful.
-
-Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM}).
-Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
-We optimize grammars using the same stochastic gradient descent method as we use for efficiency optimization (Section REF).
-Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
-Indeed, we found that DLM is predicted by efficiency optimization, as discussed in Section REF.
-Results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
-
-%posterior_perRelation_Real_conj.pdf         posterior_perRelation_Real_goeswith.pdf     posterior_perRelation_Real_nummod.pdf       
-%user@user-G751JY:~/grammar-optim/results/correlations$ evince figures/posteriors/posterior_perRelation_Real_csubj.pdf 
-%user@user-G751JY:~/grammar-optim/results/correlations$
-
-\begin{table*} % TODO figure out away to make this table more aesthetic
-	\begin{center}	
-\small{
-\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
-	\hline
-&	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-% From Dryer	
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-%\hdashline
-
-\hline
-% Other universals, not in Dryer
-&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
-&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf.
-\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
-&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf.
-\cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
-&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section~\ref{sec:correlations} \\
-&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
-&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
-\end{tabular}
-}
-
-\end{center}
-\caption{Predictions on UD relations with predictions from the typological literature.
-The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
-In the last column, we indicate what direction would be expected typologically.}
-\label{tab:all-predictions-1}
-\end{table*}
-
-\footnotetext{Note that we report an \emph{anti-correlation}.
-The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
-Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
-}
-
-\begin{table*} % TODO figure out away to make this table more aesthetic
-	\begin{center}	
-\small{
-\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
-	\hline
-	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-\hline
-appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\
-lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\
-expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\
-iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\
-vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\
-
-\hline
-% Uninterpretable
-compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\
-det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\
-dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\
-dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
-advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\
-
-\hline
-% UD artifacts
-conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
-discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\
-fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\
-flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
-goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
-list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
-orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\
-parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\
-reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\
-
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
-\end{tabular}
-}
-
-\end{center}
-\caption{Predictions on UD relations for which no predictions are available in the typological literature.
-``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
-``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
-\label{tab:all-predictions-2}
-\end{table*}
-
-\subsection{Other Experiments}
-In Table~\ref{table:corr-resu} we report the results of the simulations we did before the main experiment reported in the main paper (left and center), together with their preregistrations, besides the results from the main experiment (right).
-These experiments all had the same setup described in Section (REF), which was fixed before starting simulations; differences are that (1) the first simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
-Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
-\textcolor{red}{TODO also the aggregate posterior}
-
-\begin{table}[hbt!]
-	\begin{center}
-\small{
 \begin{tabular}{|c|ccc|c}
@@
 	%&
-& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table REF.} 
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} \\ \hline \hline 
+& $\lambda=1.0$ & $\lambda=0.9$ & $\lambda=0.9$ \\ \hline \hline 
 	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}		  
@@
 \end{tabular}
-}
 	\end{center}
-\caption{\textcolor{red}{probably should also report parseability results as lambda=0 here}Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
+\caption{Previous experiment with $\lambda=0.9$ (Preregistration)}\label{table:corr-resu}
 \end{table}
 
-%\footnotetext{}
-%  final experiment
-%  0.9
-%
-
-%\section{Joint optimization of grammar, parser, and language model parameters}
-\section{Creating Optimized Grammars}
-
-In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
-%Our method will be applicable more generally to the task of optimizing grammars for a broad class of objective functions.
-We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
-More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
-This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.
-
-This method addresses a major challenge noted in previous work optimizing grammars, namely that the predictability (and parseability) of an individual sentence depends on the entire distribution of the language.
-Previously, \citet{gildea2015human} optimized grammars for dependency length and trigram surprisal using a simple hill-climbing method on the grammar parameters, which required reestimating the trigram surprisal model in every iteration.
-Such a method would be computationally prohibitive for efficiency optimization, as it would require reestimating the neural network models after every change to the grammar, which would amount to reestimating them hundreds or thousands of times per grammar.
-Our method, by allowing for the use of stochastic gradient descent, addresses this challenge, as we describe in Section~\ref{sec:optim-eff}.
-
-\subsection{Differentiable Ordering Grammars}\label{sec:diff-gramm}
-
-We extended the parameter space of grammars by continuously interpolating between grammars, making efficiency a \emph{differentiable} function of grammar parameters.
-The parameters of such a \key{differentiable word order grammar} are as follows. 
-For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
-Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
-Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
-184) without replacement.
-
-If $a_\tau \in \{0, 1\}$, and the distances between values of $b_\tau$ (for different $\tau$) become very large, such a differentiable grammar becomes deterministic, assigning almost full probability to exactly one ordering for each syntactic structure.
-In this case, the grammar can be converted into an equivalent grammar of the form described in Materials and Methods, by extracting a single parameter in $[-1, 1]$ for each relation $\tau$.
-
-\subsection{Extracting Grammars from Datasets}
-We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
-To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
-We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.
-
-\subsection{Optimizing Grammars for Efficiency}\label{sec:optim-eff}
-
-In this section, we describe how we optimized grammar parameters for efficiency.
-A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
-More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $\mathbf{w}$.
-In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $p_{\mathcal{L}_\theta}(u|t)$ concentrates on a single ordering $u$.
-
-Recall the definition of efficiency
-\begin{equation}\label{eq:efficiency}
-	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
-\end{equation}
-where 
-\begin{equation}
-	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
-\end{equation}
-\begin{equation}
-	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
-\end{equation}
-where $t \sim \mathcal{T}$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
-
-These quatities are estimated using two neural models, as described in Section X:
-A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.
-The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is
-\begin{equation}
-	R_{Pars}^{\phi}(u,t) =  \log \frac{p_\phi(t|u)}{p(t)}
-\end{equation}
-%and we can write
-%\begin{equation}
-%	R_{Pars}^\phi := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
-%\end{equation}
-%
-A \key{language model}, with some parameters $\psi$, calculates the word-by-word surprisal of an utterance:
-\begin{equation}
-	R_{Pred}^{\psi}(u) = - \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})
-\end{equation}
-Using this, we can rewrite Efficiency~(\ref{eq:efficiency}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
-\begin{equation}
-	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi} := \max_{\phi,\psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
-\end{equation}
-%
-%The parsers and language models appropriate to a tree distribution $\mathcal{T}$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
-%\begin{equation}
-%\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-%\end{equation}
-%\begin{equation}
-%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-%\end{equation}
-%
-%Using these models, we can rewrite 
-%\begin{equation}
-%R_{Pars}^{\psi(\theta)}= \E_t \E_{u \sim p_\theta(u|t)} R_{Pars}^{\psi(\theta)}(u,t) 
-%\end{equation}
-%where $R_{Pars}(u,t) = \log \frac{p_\phi(t|u)}{p(t)}$, and $p_\theta(u|t)$ is the probability assigned to sentence $u$ by the distribution $\mathcal{L}_\theta(t)$ defined by the differentiable word order grammar $\mathcal{L}_\theta$,
-%and
-%\begin{equation}
-%R_{Pred} = \E_t \E_{u \sim p_\theta(u|t)} R_{Pred}(u)
-%\end{equation}
-%where $R_{Pred}(u) = \log p_\psi(u)$.
-%Therefore, we can rewrite efficiency as
-%\begin{equation}\label{eq:efficiency}
-%R_{\textit{Eff}}^\theta := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi(\theta)}(u,t) + \lambda R_{Pred}^{\psi(\theta)}(u)\right]
-%\end{equation}
-%Directly applying gradient descent to this expression is hard due to the complicated dependence of $\phi(\theta), \psi(\theta)$ on $\theta$.
-%
-%However, we can rewrite
-%\begin{equation}\label{eq:efficiency}
-%\argmax_\theta\	R_{\textit{Eff}}^\theta 
-%\end{equation}
-%equivalently into
-In order to find an optimal grammar $\theta$, we thus need to solve
-\begin{equation}\label{eq:efficiency}
-\argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} %= \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
-\end{equation}
-Importantly, $R_{\textit{Eff}}^{\theta, \phi, \psi}$ is differentiable in $\theta, \phi, \psi$: %, and we can apply stochastic gradient descent to carry out this optimization.
-\begin{align}
-\partial_\theta R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)} \left[  \left[\partial_\theta \log p_\theta(u|t)\right] \cdot    \left(R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right) \right] \label{eq:dtheta}\\
-\partial_\phi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\partial_\phi R_{Pars}^{\phi}(u,t)\right] \\
-\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}
-\end{align}
-where \ref{eq:dtheta} is derived using the \emph{score-function} or \emph{REINFORCE} trick~\cite{williams1992simple}.
-Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures (CITE).
-
-We can therefore apply stochastic gradient descent to jointly optimize $\theta, \phi, \psi$:
-In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim p_{\theta}(\cdot|t)$.
-Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
-do a gradient descent step using the estimator given by the expressions in the square brackets in \ref{eq:dtheta}-\ref{eq:dpsi}.
-
-Optimizing for only parseability (or predictability) is very similar -- in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.
-
-At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
-Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.
-
-See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates used in this optimization scheme.
-
-%
-%%(analogous for $R_{Pars}$ instead of $R_{Pred}$)
-%\begin{equation}\label{eq:estimator}
-%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
-%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
-%\end{equation}
-%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
-%for the gradient
-%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
-%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
-%
-
-%
-%
-%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
-Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
-%\begin{equation}
-%\label{eq:exp-over-utts}
-%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
-%\end{equation}
-%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
-In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
-%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
-%
-%
-%
-%We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
-%\begin{equation*}
-%R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
-%\end{equation*}
-%where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
-%\begin{equation*}
-%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-%\end{equation*}
-%
-%
-%We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
-Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
-%\begin{align}
-%\nonumber
-%R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
-%\nonumber
-%&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
-%\end{align}
-%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
-%\begin{equation*}
-%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-%\end{equation*}
-%
-%For efficiency, we linearly combine the two terms:
-%\begin{equation}
-%\label{eq:combined-score}
-%R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
-%\end{equation}
-%
-%
-%%\begin{equation}
-%%    \label{eq:depl}
-%%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
-%%\end{equation}
-%%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
-%
-%For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
-The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
-The optimization problems come out to:
-%\begin{align}
-%%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
-%\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
-%\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
-%\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
-%\end{align}
-%Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
-%
-%
-%%The case of predictability, parseability, and efficiency are more challenging, as 
-%%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
-%We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
-%Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
-%
-%We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section (REF)).
-%
-%In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
-%Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
-%do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
-%\begin{equation}\label{eq:estimator}
-%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
-%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
-%\end{equation}
-%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
-%for the gradient
-%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
-%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
-%
-%See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
-%
-
-%The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
-A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
-%We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
-
-%In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
-%and $\ell_2$ regularization.
-
-%A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
-%Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
-%We thus do not evaluate this objective function here.
-
-%We extract grammars for the actual languages using a Bayesian data analysis method.
-%We specify a simple Bayesian prior over differentiable ordering grammars:
-%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
-%for each $\tau$.
-%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
-%We extract
-
-\section{Neural Network Architectures}\label{sec:neural-architectures}
-
-In this section, we describe the details of the neural network architectures.
-Choices follow standard practice and machine learning and were fixed before evaluating word order properties, and the efficiency of real languages.
-
-%\paragraph{Locality}
-%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
-
-\paragraph{Estimating Predictability}
-We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
-Given the small size of the corpora, this limit is only attained only for few languages.
-In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
-The model predicts both the next word and its POS tags in each step.
-Using POS tags is intended to prevent overfitting on small corpora.
-
-\paragraph{Estimating Parseability}
-We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}.
-This architecture is remarkably simple: the words of a sentence are encoded into context-sensitive embeddings using bidirectional LSTMs, then a classifier is trained to predict the head for each work.
-The classifier works by calculating a score for every pair of word embeddings $(w_i, w_j)$, indicating the likelihood that the $j$th word is the head of the $i$th word.
-This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
-
-To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags.
-Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged; parsing from POS tags prevents early overfitting.
-This decision was made before evaluating word order properties.
-
-\paragraph{Hyperparameters}
-
-%\textbf{TODO study 1 only has 1 set of parameters}
-
-Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
-For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
-These parameters are shown in Table~\ref{tab:hyperparameters}.
-Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (efficiency, predictability, parseability) on the language.
-We then used this setting for creating optimized word order grammars.
-
-%\textcolor{red}{TODO random seeds for efficiency}
-
-All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
-We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{|l|l|l|}
-\hline
-\multirow{2}{*}{Optimization}&    Learning Rate     & 5e-6, 1e-5, 2e-5, 5e-5 \\
-&    Momentum & 0.8, 0.9 \\ \hline
-\multirow{6}{*}{Language Model} &    Learning Rate  & 0.5, 0.1, 0.2 \\
-&    Dropout Rate& 0.0, 0.3, 0.5 \\
-& Embedding Size (Words) & 50 \\
-& Embedding Size (POS) & 20 \\
-&    LSTM Layers  & 2 \\
-&    LSTM Dimensions  & 128 \\
- \hline
-\multirow{5}{*}{Parser}&    Learning Rate  & 0.001 \\
-&    Dropout Rate  & 0.2 \\
-&    Embedding Size  & 100 \\
-&    LSTM Layers  & 2 \\
-&    LSTM Dimensions  & 200 \\
-\hline
-    \end{tabular}
-    \caption{Hyperparameters}
-    \label{tab:hyperparameters}
-\end{table}
-
-\paragraph{Improved Unbiased Gradient Estimator}
-We employ two common variance reduction methods to improve the estimator~(\ref{eq:dtheta}), while keeping it unbiased.
-%For dependency length and 
-% (and its head, in the case of dependency length)
-For predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word.
-We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:dtheta}) with lower variance.
-Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple} (the word's surprisal in the case of predictability, and the negative log-probability of the correct head and relation label in the case of parseability).
-These two methods reduce the variance of the estimator and thereby increase the speed of optimization and reduce training time, without biasing the results.
-For numerical stability, we represent $a_\tau \in [0,1]$ via its logit $\in \mathbb{R}$.
-Furthermore, to encourage exploration of the parameter space, we add an entropy regularization term \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$.
-The weight of the entropy regularization was chosen together with the other hyperparameters.\footnote{Explored values: 0.0001, 0.001.}
-%, adding the negative entropy of the $Bernoulli(a_\tau)$ to the loss $R$. %, weighted with $\alpha$.
-
-These techniques for improving (\ref{eq:dtheta}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
-
-\paragraph{Learning Rate and Stopping Criterion}
-We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
-For the language model parameters $\phi$, we use plain Stochastic Gradient Descent without momentum, as recommended by \citet{merity2018regularizing}. 
-For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
-The learning rates and other optimization hyperparameters were determined together with the other hyperparameters.
-
-For gramar optimization, we apply Early Stopping~\citep{prechelt1998early} as follows.
-We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
-This is a standard technique for optimizing neural models, and helps prevent overfitting to the specific corpora used.
-
-The choice of optimization methods and the stopping criterion were fixed before we investigated word order correlations.
-
-\paragraph{Optimized Grammars}
-Recall that, for each language, we created 8 optimized languages for each optimization criterion.
-We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
-This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.
-
-For efficiency optimization, for each grammar, we ran efficiency optimization with three different random seeds, selecting among these the seed that yielded the best overall efficiency value.
-This helps control for variation across random seeds for the stochastic gradient descent optimization method.
-
- % TODO we could use some language at the top of this section on the general guiding principles for all these decisions.
-Were these implementation decisions made before or after looking at the word order universals results?
-If they were made before, we should highlight that.
-
-\section{Robustness to different language models and parsers}
-
-Here we take up the question of the extent to which our results are dependent on the particular parser and language model used in the optimization process.
-We want to know: when we optimize a word order grammar for efficiency, have we produced a language which is highly efficient \emph{in general}, or one which is highly efficient \emph{for a specific parser}?
-We wish to argue that natural language syntax is optimized for efficiency in general, meaning that syntactic trees are highly recoverable from word orders in principle.
-If it turns out that our optimized languages are only optimal for a certain parser from the NLP literature, then we run the risk of circularity: it may be that the reason this parser was successful in the NLP literature was because it implicitly encoded word order universals in its inductive biases, and thus it would be no surprise that languages which are optimized for parseability also show those universals.
-
-In this connection, we note that the parser and language model architectures we use are highly generic, and do not encode any obvious bias toward natural-language-like word orders.
-The LSTM language model is a generic model of sequence data which is also been used to model financial time series \citep{sirignano2018universal} and purely theoretical chaotic dynamical systems \citep{ogunmolu2016nonlinear}; the neural graph-based parser is simply solving a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
-Nevertheless, it may be the case that a bias toward word order universals is somehow encoded implicitly in the hyperparameters and architectures of these models.
-
-Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers.
-These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.
-
-\subsection{CKY Parsers}
-
-We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of \textcolor{red}{CITE Collins 1997 (Model 1).}
-In our PCFGs, each head independently generates a set of left and right dependents.
-We formulate this as a PCFG where each rule has the form:
-\begin{center}
-	POS$_H$ $\rightarrow$ POS$_H$ POS$_D$
-\end{center}
-for head-initial structures, and
-\begin{center}
-	POS$_H$ $\rightarrow$ POS$_D$ POS$_H$
-\end{center}
-for head-final structures, where each symbol is a POS tag.
-Thus, POS tags act both as terminals and as nonterminals.
-
-We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
-For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm \textcolor{red}{CITE}.
-
-This parsing strategy is very different from the neural graph-based parser:
-While the graph-based parser solves a minimum spanning tree problem, the CKY algorithm uses dynamic programming to compute the exact probabilities of trees given a sentence, as specified by the generative model encoded in the PCFG.
-Second, while the graph-based neural parser uses machine learning to induce syntactic knowledge from data, the CKY parser performs exact probabilistic inference.
-%In this sense, the CKY algorithm does not have any architectural biases in itself.
-%On the other hand, the PCFG makes severely simplifying independence assumptions, compared to the universal approximation capabilities of neural network-based systems.
-
-We used the CKY algorithm to compute the parsing loss $H[T|L_\theta(T)]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
-Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
-% TODO get UAS for compatibility with other parts of SI
-%We then compared parsability of random and optimized word order grammars on English data.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
-	\caption{Parsability loss $H[T|L_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).\textcolor{red}{make plot nicer}}
-    \label{fig:cky-parser}
-\end{figure}
-
-\subsection{Distorted graph-based parsers}
-
-In this section, we address the idea that the graph-based parser might have a built-in bias toward certain kinds of orderings, and the question whether this might be responsible for our findings.
-In particular, we address the idea that the graph-based parser might have a bias toward parses involving short dependencies, which we call a \key{locality bias}. 
-We address this by changing the order in which the parser sees words, in such a way that the distance between words in the input is not indicative of syntactic relations.
-
-%In particular, this allows us to show tha
-
-%Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization.
-Therefore, we address here the idea that the graph-based parser might have a built-in bias toward parses involving short dependencies, which we call a \key{locality bias}.
-%A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words.
-Suppose that a parser $P$ has a bias toward positing a dependency between a word and the immediate previous word.
-
-\paragraph{Even--odd order.} A sequence of $n$ words originally ordered as $w_1 w_2 w_3 w_4 \cdots w_n$ is reordered by separating the even and odd indices: $w_2 w_4 w_6 \cdots w_{n-1} w_1 w_3 w_5 \cdots w_n$ (assuming $n$ odd).
-Therefore all words that are adjacent in the original order will be separated by a distance of $\approx n/2$ in the distorted order, while all words of distance 2 in the original order will become adjacent.
-
-\paragraph{Interleaving order.} In interleaving ordering, a sequence originally ordered as $w_1 w_2 w_3 \cdots w_n$ is split in half at the middle (index $m=\ceiling{n/2}$), and the two resulting sequences are interleaved, yielding $w_1 w_m w_2 w_{m+1} w_3 w_{m+3} \cdots w_n$.
-Thus all words that were originally adjacent will have distance 2 in the distorted order, with the intervening word coming from a very distant part of the sentence.
-
-\paragraph{Inwards order.} A sequence originally ordered as $w_1 w_2 w_3 \cdots w_{n-1} w_n$ is ordered from the edges of the string inwards, as $w_1 w_n w_2 w_{n-1} \cdots w_{\ceiling{n/2}}$.
-This corresponds to folding the string in on itself once, or equivalently, splitting the sequence in half at the middle, then interleaving the two resulting sequences after reversing the second one.
-The result is that the most non-local possible dependencies in the original order become the most local dependencies in the distorted order.
-
-\paragraph{Sorted order.} A sequence is reordered by sorting by POS tags, and randomizing the order within each block of identical POS tags.
-To each word, we then add a symbol encoding the original position in the sequence.
-For instance
-\begin{center}
-PRON VERB PRON
-\end{center}
-may be reordered as
-\begin{center}
-PRON 1 PRON 3 VERB 2
-\end{center}
-or
-\begin{center}
-PRON 3 PRON 1 VERB 2
-\end{center}
-The numbers are provided to the parser as atomic symbols from a vocabulary ranging from 1 to 200; numbers greater than 200 (which may occur in extremely long sentences) are replaced by an out-of-range token.
-
-The result is that distance between words in the input is not indicative at all of the presence of absence of syntactic relations between them.
-
-\paragraph{Experiments}
-Using English and Japanese data, we trained parsers for the ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
-Resulting parsing accuracy scores are shown in Figure~\ref{fig:distorted-parser}.
-In all settings, the language optimized for efficiency achieved higher parsing accuracy than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
-%Note that absolute values of parsing accuracy vary between the orderings, being higher for the actual orderings than for distorted orderings, in which trees become nonprojective.
-
-\begin{figure}
-    \centering
-    English
-
-    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-english.pdf}
-
-    Japanese
-
-    \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
-	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').}
-    \label{fig:distorted-parser}
-\end{figure}
-
-\subsection{n-gram language models}
-
-We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
-In previous work, such as \cite{gildea2015human}, predictability has often been measured using n-gram models.
-
-Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
-Specifically, we constructed bigram models with Kneser-Ney smoothing~\cite{kneser1995improved, chen1999empirical}.
-A bigram model predicts each word taking only the previous word into account.
-This contrasts with LSTMs, which take the entire context into consideration.
-Thus, bigram models and LSTMs stand on opposing ends of a spectrum of language models taking more and more aspects of the context into account.
-
-We estimated language models on the training partitions, and used the validation partitions to estimate surprisal.
-We conducted this for the ten random and the best optimized ordering grammars on English and Japanese data.
-Results (Figure~\ref{fig:bigrams}) show that languages optimized for efficiency are optimal for a bigram language model.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.4]{../results/bigrams/bigrams.pdf} 
-	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).\textcolor{red}{make fig nicer}}
-    \label{fig:bigrams}
-\end{figure}
-
-%\subsection{Discussion}
-%
-%It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.
-
-\section{Effects of data sparsity}
-
-Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
-If the difference between random and real grammars is due to data sparsity, we expect that it decreases as the amount of training data is increased.
-If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference persists as training data is increased.
-
-We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
-We cosidered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
-For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
-Furthermore, for the random grammar, we varied the number of neurons in the BiLSTM (200, 400, 800) to test whether results depend on the capacity of the network.
-
-The resulting curves are shown in Figure~\ref{fig:learning-czech}.
-%A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
-A gap in parsing loss of about 0.2 nats appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
-
-\begin{figure}[ht]
-    \centering
-%    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech.pdf} 
-    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech-logloss.pdf}
-
-        \caption{Parsing loss ($H[T|\mathcal{L}_\theta(T)]$) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
-    \label{fig:learning-czech}
-\end{figure}
-
-\section{Comparing Extracted Grammars to Linguistic Judgments}
-
-% user@user-X510UAR:~/grammar-optim/results/grammars$ git add comparison-table.tex
-
-We validated our grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
-Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS.
-In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb-object relation.
-
-WALS has some data for all but seven languages.\footnote{Note that Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
-In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}
-WALS has data for 74\% of the entries, and lists a dominant order for 91\% of these.
-The grammars we extracted from the corpora agree with WALS in 96~\%  of these cases.
-
-\begin{table}[htb!]
-\small{
-\begin{center}
-\begin{tabular}{l||ll|ll|ll|ll|ll|ll|llllll}
-%	   for(dep in c("obj", "lifted_mark", "acl", "nmod", "obl")) {
-		   Language 
-		   &	\multicolumn{2}{|c|}{Objects} 
-		   &	\multicolumn{2}{|c|}{Adpositions} 
-		   &	\multicolumn{2}{|c|}{Compl.} 
-		   &	\multicolumn{2}{|c|}{Rel.Cl.} 
-		   &	\multicolumn{2}{|c|}{Genitive} 
-		   &	\multicolumn{2}{|c|}{PP}  \\ \hline\hline
-	\input{../results/grammars/comparison-table.tex}
-\end{tabular}
-\end{center}
-}
-\caption{Comparing grammars extracted from databases to linguistic judgments in the World Atlas of Linguistic Structures.
-For each of the six syntactic relation, the first column provides the ordered coded in the extracted grammar; the second column provides the order coded in WALS (DH for dependent-head, HD for head-dependent order).
-`?' indicates that WALS has no data.
-$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur at similar rates in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
-Finally, `--' indicates that the relation does not occur in the corpus.
-}\label{tab:grammars-wals}
-\end{table}
-
-\section{Languages and Corpus Sizes}
-In Table~\ref{tab:langs-iso-sizes}, we list the 51 languages with ISO codes and families, with the size of the available data per language.
-We included all UD 2.1 languages for which a training partition was available.
+\section{Languages and Corpora}
 
 \begin{table}[ht]
-\small{
 \begin{tabular}{lllllll}
@@
 \end{tabular}
-}
-\caption{Languages with ISO codes, families (according to \url{https://universaldependencies.org/}), and the number of available sentence and words.}\label{tab:langs-iso-sizes}
+\caption{Languages}
 \end{table}
 
-\section{Dependency Length Minimization}\label{sec:DLM}
+\section{Dependency Length Minimization}
 
@@
 Indeed, directly modeling efficiency optimization predicts DLM, and appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
-
-\textcolor{red}{at some place also say the following: Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
-}
-
-Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM}).
-Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
-Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
-Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
-In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
-
-DLM turns out to be predicted by our model
 

### 2019-05-21 second-draft/pnas-word-order-si.tex
SI

--- 2019-05-18 second-draft/pnas-word-order-si.tex
@@
 
+\textcolor{red}{cite rijkhof etc}
+
 \section{Formalization of Correlations}\label{sec:correlations}
@@
 
+\textcolor{red}{TODO also report the frequentist things, posteriors, numerically}
+
 %https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
@@
 \textcolor{red}{TODO also the aggregate posterior}
+
+\begin{table}[hbt!]
+	\begin{center}
+\small{
+\begin{tabular}{|c|c|ccc|c}
+	\hline
+	%&
+& $\lambda = 0.0$
+& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM}.} 
+& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
+& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} 
+	\\ \hline \hline 
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}		  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_case.pdf}     } 
+ 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_case.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_cop.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_cop.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_aux.pdf}     }  reverse
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_aux.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_nmod.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_nmod.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}   }
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_acl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_acl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   }
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_mark.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_mark.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   }
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_obl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_obl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   }
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_xcomp.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_xcomp.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }  
+		\\
+&&&\\ \hline
+ \hline
+\end{tabular}
+}
+	\end{center}
+\caption{\textcolor{red}{probably should also report parseability results as lambda=0 here}Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
+\end{table}
 

### 2019-05-21 second-draft/pnas-word-order-si.tex
SI

@@
 The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
-Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the information bottleneck formalism, bringing them even closer to our formalization.
+Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the information bottleneck formalism \cite{zaslavsky2018efficient, zaslavsky2019semantic}, bringing them even closer to our formalization of efficiency.
 

### 2019-05-22 second-draft/pnas-word-order-si.tex
SI

@@
 We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.
-%\footnote{Regarding objections by \cite{dunn2011evolved}, we refer to the follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
+%TODO REF mention this? \footnote{Regarding objections by \cite{dunn2011evolved}, we refer to the follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
 Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
@@
 
+\paragraph{Bayesian Regression}
+We modeled the
+
+We modeled the probabilities $p_{i,j}$ that a grammar $g_i$ satisfies the $j$-th correlation ($j=1,...,8$) using a multilevel logistic model \textcolor{red}{TODO (CITE)}, with random intercepts for the language for whose data the grammar had been optimized, and its language family, annotated according to \url{http://universaldependencies.org/}..
+\textcolor{red}{TODO some more details here}
+%See SI Section S4.3 for details.
+
+As we constructed 4 verb-object and 4 object-verb grammars for each of the 51 languages, TODO...
+
 Following the recommendations of \cite{ghosh2018use, burkner2018advanced}, we used as a very weakly informative prior a Student's $t$ prior with $\nu=3$ degrees of freedom, mean 0, and scale $\sigma=10$ (i.e., the PDF $p$ is $\frac{1}{\sigma} p_3(x/\sigma)$, where $p_3$ is the PDF of the $t$-distribution with $\nu=3$).
@@
 The covariance matrix is modeled using the LKJ prior, as described in \cite{burkner2018advanced}.
-We used four chains, with 5000 samples each (2500 warmup).
-The chains had converged ($\hat{R} < 1.01$).
-
-We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
+We used four chains, with 5000 samples each, of which the first 2500 were discarded as warmup samples.
+We confirmed convergence using $\hat{R}$ and visual inspection of chains (REF).
+
+\paragraph{Robustness}
+To ascertain the robustness of our results, we also conducted a frequentist analysis using \texttt{lme4} \cite{bates2015fitting}.
+For each of the correlations, we conducted a logistic mixed-effects analysis predicting whether a grammar satisfies the correlation, with random effects of language and language family.
+The results are shown in Table~\ref{tab:corr-regression} together with those of the Bayesian analysis.
+The frequentist analysis agrees with the Bayesian model; all eight correlations are predicted to hold in more than half of the optimized grammars ($p < 0.01$ each).
+
+Note that the Bayesian analysis  also estimates a posterior distribution of the number of satisfied correlations (see Figure~\ref{fig:posterior}), providing an elegant solution to the multiple-comparisons problem arising from analysing the eight correlation.
+
+% and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
+
+%We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
+
+\paragraph{Density Plots}
+We obtained the posterior density plots in Figures (REF main, REF SI) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\beta$.
+
+\begin{table}
+\small{
+\begin{center}
+\begin{tabular}{l||l|lll|llll|ll|llllll}
+Correlation & Prevalence & \multicolumn{3}{c}{Bayesian} & \multicolumn{4}{c}{Frequentist} \\ 
+& & Mean & SD & $p(\beta \leq 0)$ & $\beta$ & SE & $z$ & p \\
+\hline\hline
+	\input{../results/correlations/output/efficiency-results-full.tex}
+\end{tabular}
+\end{center}
+}
+\caption{Detailed results for Bayesian and Frequentist mixed-effects analyses for the eight correlations.
+We show (1) the raw prevalence of each correlation in the optimized grammars (8 grammars for each of the 51 languages),
+(2) for the Bayesian analysis, we provide posterior mean and SD of $\beta$, and the posterior probability that $\beta$ has the opposite sign,
+(3) for the Frequentist analysis, we provide the point estimate, SE, $z$, and $p$-values (2-sided).
+The frequentist analysis confirms the results of the Bayesian analysis.
+}\label{tab:corr-regression}
+\end{table}
 

### 2019-05-22 results/plane/plane-2D-best-visualization-large-tests.R

@@ -9,13 +9,19 @@ depl = depl %>% filter(grepl("FuncHead", ModelName)) %>% filter(grepl("Coarse", 
 dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
 dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
 dataS3 = read.csv("../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
-dataS4 = read.csv("~/CS_SCR/deps/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
-dataS = rbind(dataS, dataS2, dataS3, dataS4)
+dataS4 = read.csv("../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS5 = read.csv("../../grammars/plane/plane-fixed-random3.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS6 = read.csv("../../grammars/plane/plane-fixed-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS7 = read.csv("../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS = rbind(dataS, dataS2, dataS3, dataS4, dataS5, dataS6, dataS7)
 dataP = read.csv("../../grammars/plane/plane-parse.tsv", sep="\t") %>% mutate(Model = as.character(Model))
 dataP2 = read.csv("../../grammars/plane/plane-parse-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
 dataP3 = read.csv("../../grammars/plane/plane-parse-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataP4 = read.csv("~/CS_SCR/deps/plane-parse-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model))
[…]

### 2019-05-22 second-draft/pnas-word-order-si.tex
SI

@@
 
-\textcolor{red}{cite rijkhof etc}
-
 \section{Formalization of Correlations}\label{sec:correlations}
@@
 
-%\textcolor{blue}{Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
+%\textcolor{red}{Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
 %Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
@@
 
-%Recall our formalization of efficiency:
-%\begin{equation}
-%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
-%\end{equation}
-%\begin{equation}
-%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
-%\end{equation}
-%\begin{equation}\label{eq:efficiency}
-%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
-%\end{equation}
-
 Here we discuss how our formalization of communicative efficiency relates to other information-theoretic work on language.
-
-Our formalization of efficiency is identical to the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems, with syntactic structures taking the role of underlying meanings that the speaker wants to convey.
-This function is also equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
-Very similar metrics have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
-%More generally, these approaches instantiate Shannon's Rate Distortion theory.
-Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.
-
-All these accounts formalize the amount of information provided as the \key{mutual information} between the forms that a listener receives, and the meanings that the speaker intends to communicate to the speaker.
-Most previous studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} are concerned with the meanings of individual words; they thus take the forms to be individual words, which refer to individual objects and their properties.
-We instead study how words are composed into sentences.
-In our work, the underlying meanings are the syntactic structures; the forms are the linearized sentences communicated by the speaker to the listener.
+Our formalization of efficiency is based on the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems.
+The key idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms.
+%; there are some differences in the formalization of complexity.
+Formally, if $S$ denotes signals (e.g., words, sentences) and $R$ denotes their referents (e.g., objects in a reference game), then this efficiency metric takes the form (notation slightly varies across these publications):
+\begin{equation}\label{eq:eff-general}
+I[S, R] - \lambda H[S]
+\end{equation}
+where $I[S, R]$ describes the \key{informativity} of the signals $S$ about their referents $R$, and $H[S]$ describes the \key{complexity} of the communication system.
+While prior studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} mostly considered settings where the signals $S$ are individual words without further structure, the signals are entire sentences $\mathcal{U}$ in our setting.
+The underlying messages $R$ which the speaker aims to convey are the syntactic structures $\mathcal{T}$.
 By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
@@
 
-%These information-theoretic accounts differ in the formalizations of complexity and diversity:
-Like our model, \cite{ferreri2003least,ferrericancho2007global,futrell2017memory} model the complexity as the \key{entropy} of the distribution of forms, which arises naturally from the role played by surprisal in determining human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
-This model is closely related to the Information Bottleneck approach \cite{zaslavsky2018efficient, zaslavsky2019semantic}, which models complexity as the mutual information between forms (e.g., words) and their referents (e.g., individual objects).
+We therefore arrive at the following efficiency metric:
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+\end{equation}
+where \key{parseability} is the amount of information that utterances provide about their underlying syntactic structures:
+\begin{equation}
+	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+\end{equation}
+and \key{predictability} is the negative entropy or surprisal of the language:
+\begin{equation}
+	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+\end{equation}
+This efficiency metric~(\ref{eq:eff-general}) is also equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
+The Information Bottleneck models complexity using a mutual information term instead of the entropy.
 This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
 As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure.
-
-A few models have used other complexity measures that are not explicitly information-theoretic quantities.
+Entropy also arises naturally as a complexity measure from the role played by surprisal in determining human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+
+A few other approaches share the mutual information term for informativity, while using complexity measures that are not explicitly information-theoretic.
+%A few models have used other complexity measures that are not explicitly information-theoretic quantities.
 In \cite{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \cite{kemp2012kinship} it is the difficulty of defining the concepts that are encoded.
 The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
-Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the information bottleneck formalism \cite{zaslavsky2018efficient, zaslavsky2019semantic}, bringing them even closer to our formalization of efficiency.
+Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the Information Bottleneck formalism \cite{zaslavsky2018efficient, zaslavsky2019semantic}, bringing them even closer to our formalization of efficiency.
 
@@
 In line with the other models, these assume that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely}.
+
+%Other metrics very similar to~(\ref{eq:eff-general}) have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
+%The idea of formalizing informativity using the mutual information $I[S, R]$ is common to all these approaches.
+
+%These information-theoretic accounts vary somewhat in the formalizations of complexity.
+%Like our model, \cite{ferreri2003least,ferrericancho2007global,futrell2017memory} model the complexity as the \key{entropy} of the distribution of forms, which .
+%This model is closely related to the Information Bottleneck approach \cite{zaslavsky2018efficient, zaslavsky2019semantic}, which models complexity as the mutual information between forms (e.g., words) and their referents (e.g., individual objects).
 
@@
 \begin{equation}
-I[L(T); T] \leq H[L(T)].
+I[U; T] \leq H[U].
 \end{equation}
@@
 \begin{equation}
-    J_T(L) = I[L(T); T] - \lambda H[L(T)] \leq 0.
+    R_{Eff} = I[U; T] - \lambda H[U] \leq 0.
 \end{equation}
-and it takes the maximally possible value zero if $L(T)$ maps all trees to a single utterance, in which case both $I[L(T); T]$ and $H[L(T)]$ are zero.
+and it takes the maximally possible value zero if $U$ maps all trees to a single utterance, in which case both $I[U; T]$ and $H[U]$ are zero.
 This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
@@
 \subsection{Regression for Predicted Correlations}
-
-\textcolor{red}{TODO also report the frequentist things, posteriors, numerically}
 
@@
 This has been suggested as a property of efficient word order in prior work, and, indeed, we found that DLM is predicted by efficiency optimization (Section~\ref{sec:DLM}).
-Results in Table~\ref{tab:all-predictions-1} also show that optimizing for DLM makes predictions similar to efficiency optimization. See Section~\ref{sec:DLM}.
+Results in Table~\ref{tab:all-predictions-1} also show that optimizing for DLM makes predictions similar to efficiency optimization.
+See Section~\ref{sec:DLM} for further discussio for further discussionn.
 
@@
 	\end{center}
-\caption{\textcolor{red}{probably should also report parseability results as lambda=0 here}Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
+\caption{\textcolor{red}{parseability color, aux inversion}Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
 \end{table}
@@
 	\end{center}
-\caption{\textcolor{red}{probably should also report parseability results as lambda=0 here}Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
+\caption{Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
 \end{table}
@@
 
-WALS has some data for all but seven languages.\footnote{Note that Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
+WALS has some data for all but seven languages.\footnote{Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
 In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}
@@
 A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.
-The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is
+The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is\footnote{Note that, in the definition of $R_{Pars}$, the term $p(t)$ is a constant independent of $\phi$ and the word order grammar $\mathcal{L}_\theta$; it can therefore be ignored in the optimization process.}
 \begin{equation}
@@
 
-%\textcolor{red}{TODO random seeds for efficiency}
-
 All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
@@
 
-We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of \textcolor{red}{CITE Collins 1997 (Model 1).}
+We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of \cite{collins2003head} (Model 1).
 In our PCFGs, each head independently generates a set of left and right dependents.
@@
 We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
-For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm \textcolor{red}{CITE}.
+For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm \cite{kasami1966efficiency}.
 
@@
     \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
-	\caption{Parsability loss $H[T|\mathcal{L}_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).
-We report $H[T|\mathcal{L}_\theta(T)]$ normalized by sentence length.
-\textcolor{red}{make plot nicer}}
+	\caption{Parsing loss $H[T|\mathcal{L}_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).
+We report $H[T|\mathcal{L}_\theta(T)]$ normalized by sentence length.}
     \label{fig:cky-parser}
@@
     \centering
-    \includegraphics[scale=.4]{../results/bigrams/bigrams.pdf} 
-	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).\textcolor{red}{make fig nicer}}
+    \includegraphics[scale=.6]{../results/bigrams/bigrams.pdf} 
+	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).}
     \label{fig:bigrams}
@@
 %Results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
-
-%\textcolor{red}{TODO}
 

### 2019-05-22 results/correlations/plot_families_num2-allThree.R

@@ -20,6 +20,8 @@ D = merge(D, DFam %>% select(FamilyPrint, yOffset), by=c("FamilyPrint"))
 DLang = unique(D %>% select(Language_Ordered, iso_Ordered, LanguageNumeric, yOffset))
 
 
+D = D %>% mutate(CoarseDependency = recode(CoarseDependency, lifted_case=1, lifted_cop=2, aux=3, nmod=4, acl=5, lifted_mark=6, obl=7, xcomp=8))
+
 plot_orders_real = ggplot(D %>% filter(Type == "Real Languages"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
   geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
 #  scale_color_gradient() + #values=c("blue", "green")) +
@@ -143,6 +145,6 @@ ggsave(plot=plot, "figures/pred-eff-pred-pars-families-2.pdf", width=6, height=8
 D2 = (D %>% select(Family, Language, CoarseDependency, DirB, Type) %>% spread(Type, DirB) %>% rename(Real = 'Real Languages') %>% rename(Predicted = Efficiency))
 
 D2$Agree = (D2$Real == D2$Predicted)
-summary(glmer(Agree ~ (1|CoarseDependency) + (1|Family), data=D2, family="binomial"))
+#summary(glmer(Agree ~ (1|CoarseDependency) + (1|Family), data=D2, family="binomial"))
 mean(D2$Agree)
 

### 2019-05-22 results/correlations/plot_families_num2-dlm.R

@@ -20,6 +20,8 @@ D = merge(D, DFam %>% select(FamilyPrint, yOffset), by=c("FamilyPrint"))
 DLang = unique(D %>% select(Language_Ordered, iso_Ordered, LanguageNumeric, yOffset))
 
 
+D = D %>% mutate(CoarseDependency = recode(CoarseDependency, lifted_case=1, lifted_cop=2, aux=3, nmod=4, acl=5, lifted_mark=6, obl=7, xcomp=8))
+
 plot_orders_real = ggplot(D %>% filter(Type == "Real Languages"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
   geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
 #  scale_color_gradient() + #values=c("blue", "green")) +

### 2019-05-22 results/correlations/plot_families_num2-pars.R

@@ -20,6 +20,7 @@ D = merge(D, DFam %>% select(FamilyPrint, yOffset), by=c("FamilyPrint"))
 DLang = unique(D %>% select(Language_Ordered, iso_Ordered, LanguageNumeric, yOffset))
 
 
+D = D %>% mutate(CoarseDependency = recode(CoarseDependency, lifted_case=1, lifted_cop=2, aux=3, nmod=4, acl=5, lifted_mark=6, obl=7, xcomp=8))
 
 plot_orders_real = ggplot(D %>% filter(Type == "Real Languages"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
   geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +

### 2019-05-22 second-draft/pnas-word-order-si.tex
SI

@@
 The syntactic structure specifies how the meanings of words are combined; therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
+%The idea that efficient languages will permit the listener to infer grammati
 
@@
 
+%\textcolor{red}{also mention \cite{piantadosi2012communicative,fedzechkina2012language}}
+
 %Other metrics very similar to~(\ref{eq:eff-general}) have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
@@
 
-Figure (REF) in the main paper shows the average of these per-language plots. \textcolor{red}{unify pareto curve computation in main paper}
+Figure (REF) in the main paper shows the average of these per-language plots.
 
@@
 Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).
-White cells indicate that the relation is not annotated in the dataset for the given language.
-\textcolor{red}{TODO fix order of 1-8}}\label{fig:per-lang}
+White cells indicate that the relation is not annotated in the dataset for the given language.}\label{fig:per-lang}
 %	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
 Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
-Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...) \textcolor{red}{add other functions to this plot}.}\label{fig:per-lang}
+Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...) .}\label{fig:per-lang}
 \end{figure}
@@
 \paragraph{Bayesian Regression}
-We modeled the
-
-We modeled the probabilities $p_{i,j}$ that a grammar $g_i$ satisfies the $j$-th correlation ($j=1,...,8$) using a multilevel logistic model \textcolor{red}{TODO (CITE)}, with random intercepts for the language for whose data the grammar had been optimized, and its language family, annotated according to \url{http://universaldependencies.org/}..
-\textcolor{red}{TODO some more details here}
-%See SI Section S4.3 for details.
-
-As we constructed 4 verb-object and 4 object-verb grammars for each of the 51 languages, TODO...
+We modeled the probabilities $p_{L,j}$ that a grammar optimized for data from language $L$ satisfies the $j$-th correlation ($j=1,...,8$) using a multilevel logistic model \cite{gelman2013bayesian}, with random intercepts for the language for whose data the grammar had been optimized, and for its language family, annotated according to \url{http://universaldependencies.org/}.
+Formally,
+\begin{equation}\label{eq:mixed-effects}
+logit(p_{L,j}) = \alpha_j + u_{L,j} + v_{f_L,j}
+\end{equation}
+where $f_L$ is the language family of $L$.
+The intercepts $\alpha_j$ ($j=1,...8$) encode the population-level prevalence of the correlations when controlling for differences between datasets from different languages and language families; $u_{L,j}$, $v_{f_L,j}$ encode per-language and per-family deviations from the population-level intercept $\alpha_j$.
+%, and
+%$u_{L,1,...8} \sim \mathcal{N}(0, \Sigma_{lang})$,
+%$v_{f_L,1,...8} \sim \mathcal{N}(0, \Sigma_{fam})$.
+
+%As we constructed 4 verb-object and 4 object-verb grammars for each of the 51 languages, TODO...
 
 Following the recommendations of \cite{ghosh2018use, burkner2018advanced}, we used as a very weakly informative prior a Student's $t$ prior with $\nu=3$ degrees of freedom, mean 0, and scale $\sigma=10$ (i.e., the PDF $p$ is $\frac{1}{\sigma} p_3(x/\sigma)$, where $p_3$ is the PDF of the $t$-distribution with $\nu=3$).
+We used this prior for $\alpha_j, \sigma_{L,j}, \tau_{L,j}$.
 Note that a correlation that holds in 90\% of cases corresponds to an intercept $\alpha \approx 2.19$ in the logistic model, well within the main probability mass of the prior.
-The covariance matrix is modeled using the LKJ prior, as described in \cite{burkner2018advanced}.
-We used four chains, with 5000 samples each, of which the first 2500 were discarded as warmup samples.
-We confirmed convergence using $\hat{R}$ and visual inspection of chains (REF).
+
+We modeled full covariance matrices of per-language and per-family random intercepts over all eight correlations.
+We placed an LKJ prior ($\eta=1$) on these matrices, as described in \cite{burkner2018advanced}.
+We used MCMC sampling implemented in Stan \cite{carpenter2017stan, hoffman2014no} using the R package \texttt{brms} \cite{buerkner2017brms}.
+We ran four chains, with 5000 samples each, of which the first 2500 were discarded as warmup samples.
+We confirmed convergence using $\hat{R}$ and visual inspection of chains \cite{gelman2013bayesian}.
+
+We obtained the posterior density plots in Figures (REF main, \ref{tab:all-predictions-1}) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\alpha_j$ (\ref{eq:mixed-effects}).
 
@@
 %We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
-
-\paragraph{Density Plots}
-We obtained the posterior density plots in Figures (REF main, REF SI) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\beta$.
 
@@
 Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
-\textcolor{red}{TODO also the aggregate posterior}
+
+\textcolor{red}{TODO decide on formatting}
 
@@
 
+\begin{figure}[ht]
+	\begin{center}
+	\includegraphics[width=0.25\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency09.pdf}
+	\includegraphics[width=0.25\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf}
+	\end{center}
+	\caption{(Other experiments) Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior-prior}
+\end{figure}
+
 %\section{Joint optimization of grammar, parser, and language model parameters}
@@
 where \ref{eq:dtheta} is derived using the \emph{score-function} or \emph{REINFORCE} trick~\cite{williams1992simple}.
-Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures (CITE).
+Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures.
 
@@
 \paragraph{Estimating Predictability}
-We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
-Given the small size of the corpora, this limit is only attained only for few languages.
+We choose a standard LSTM language model \citep{goldberg2017neural, hochreiter1997long}.
+We restrict the vocabulary to the most frequent 50,000 words in the treebanks for a given language.
+Given the moderate size of the corpora, this limit is only attained only for few languages.
 In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
@@
 To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags.
-Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged; parsing from POS tags prevents early overfitting.
+Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering grammar had converged; parsing from POS tags prevents early overfitting.
 This decision was made before evaluating word order properties.
@@
 We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
-For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm \cite{kasami1966efficiency}.
+For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm \cite{kasami1966efficient}.
 

### 2019-05-22 results/correlations/efficiency_glmer_table.py

@@ -6,8 +6,10 @@ dependencies = ["acl", "aux", "lifted_case", "lifted_cop", "lifted_mark", "nmod"
 
 brms = [x for x in brms if x[0] in dependencies]
 
-with open("output/efficiency-results-full.tex", "w") as outFile:
- for lme, brm in zip(lme4, brms):
+ordering = ["lifted_case", "lifted_cop", "aux", "nmod", "acl", "lifted_mark", "obl", "xcomp"]
+
+output = []
+for lme, brm in zip(lme4, brms):
    brm[0] = brm[0].replace("_", "\_")
    line = brm + lme[1:]
    if line[0] == "aux":
@@ -30,8 +32,12 @@ with open("output/efficiency-results-full.tex", "w") as outFile:
              line[j] = round(line[j], 3)
          line[j] = str(line[j])
    # acl & 0.793367346938776 & 1.50549627266397 & 0.326852828127335 & 0 & 1.434444 & 0.2716449 & 5.280586 & 1.287717e-07
+   output.append(line)
+   line[0] = ordering.index(line[0])
 
-   print((" & ".join(line)) + "\\\\", file=outFile)
-
-
+output = sorted(output, key=lambda x:x[0])
+with open("output/efficiency-results-full.tex", "w") as outFile:
+  for line in output:
+      line[0] = "\\raisebox{.5pt}{\\textcircled{\\raisebox{-.9pt} {"+str(line[0])+"}}}"
+      print((" & ".join(line)) + "\\\\", file=outFile)
 

### 2019-05-22 results/correlations/joint-evaluation-efficiency09-large.R

@@ -26,6 +26,8 @@ model3 = brm(cbind(acl, advmod, aux, lifted_case, lifted_cop, lifted_mark, nmod,
 # advmod, nsubj, 
 
 
+# TODO there is the significant problem that languages where just one count is missing are fully excluded. A total of 9 languages appear to end up excluded. The per-relation analyses make up for this, still it's a loss of statisical precision.
+
 u = posterior_samples(model3) #cbind(posterior_samples(model3, "acl_Intercept"), posterior_samples(model3, "advmod_Intercept"))
 mean(u$b_acl_Intercept < 0 |  u$b_aux_Intercept > 0  | u$b_liftedcase_Intercept < 0  | u$b_liftedcop_Intercept < 0  | u$b_liftedmark_Intercept < 0  | u$b_nmod_Intercept < 0  | u$b_obl_Intercept < 0  | u$b_xcomp_Intercept < 0 )
 # u$b_advmod_Intercept > 0  |

### 2019-05-22 results/correlations/visualize-posterior-2-prior1.0.R

@@ -98,56 +98,39 @@ library(ggplot2)
 
 cat("\nReading posterior samples\n")
 #parse = read.csv("/home/user/CS_SCR/posteriors/posterior-10-parseability.csv")
-two = read.csv("~/CS_SCR/posteriors/posterior-10-first10-efficiency.csv")
+u = read.csv("~/CS_SCR/posteriors/posterior-10-first10-efficiency.csv")
 #langmod = read.csv("/home/user/CS_SCR/posteriors/posterior-10-langmod.csv")
 
 
 
-data = rbind(two %>% mutate(Type="Efficiency"))
+u$satisfied = 8 - ((u$b_acl_Intercept < 0) + (u$b_aux_Intercept > 0 ) + (u$b_liftedcase_Intercept < 0 ) + (u$b_liftedcop_Intercept < 0 ) + (u$b_liftedmark_Intercept < 0 ) + (u$b_nmod_Intercept < 0 ) + (u$b_obl_Intercept < 0 ) + (u$b_xcomp_Intercept < 0 ))
 
-data2 = data %>% gather(Dependency, Intercept, b_acl_Intercept                             ,b_aux_Intercept                                       ,  b_liftedcase_Intercept,b_liftedcop_Intercept  ,                                 b_liftedmark_Intercept  ,                                b_nmod_Intercept         ,                                                             b_obl_Intercept            ,                             b_xcomp_Intercept           )
 
-data2 = data2 %>% mutate(Prevalence = 1/(1+exp(-Intercept)))
+library(ggplot2)
 
-data2 = data2 %>% mutate(Dependency = substring(Dependency,3, nchar(Dependency)-10))
-data2 = data2 %>% mutate(Dependency = case_when(Dependency == "liftedcase" ~ "lifted_case", Dependency == "liftedcop" ~ "lifted_cop", Dependency == "liftedmark" ~ "lifted_mark", TRUE ~ Dependency))
+u$SamplesNum = NROW(u)
+data2 = u %>% group_by(satisfied) %>% summarise(SamplesNum = mean(SamplesNum), posterior = NROW(satisfied)) %>% mutate(posteriorProb = posterior/SamplesNum)
+u = NULL
 
-data2 = data2 %>% mutate(Prevalence = case_when(Dependency == "aux" ~ 1-Prevalence, TRUE ~ Prevalence))
-
[…]

### 2019-05-22 results/plane/plane-2D-best-visualization-large-2.R

@@ -10,12 +10,18 @@ dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Mo
 dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
 dataS3 = read.csv("../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
 dataS4 = read.csv("../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
-dataS = rbind(dataS, dataS2, dataS3, dataS4)
+dataS5 = read.csv("../../grammars/plane/plane-fixed-random3.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS6 = read.csv("../../grammars/plane/plane-fixed-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS7 = read.csv("../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS = rbind(dataS, dataS2, dataS3, dataS4, dataS5, dataS6, dataS7)
 dataP = read.csv("../../grammars/plane/plane-parse.tsv", sep="\t") %>% mutate(Model = as.character(Model))
 dataP2 = read.csv("../../grammars/plane/plane-parse-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
 dataP3 = read.csv("../../grammars/plane/plane-parse-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model))
 dataP4 = read.csv("../../grammars/plane/plane-parse-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataP = rbind(dataP, dataP2, dataP3, dataP4)
+dataP5 = read.csv("../../grammars/plane/plane-parse-random3.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP6 = read.csv("../../grammars/plane/plane-parse-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model))
[…]

### 2019-05-23 second-draft/pnas-word-order-si.tex
SI

@@
 
-%posterior_perRelation_Real_conj.pdf         posterior_perRelation_Real_goeswith.pdf     posterior_perRelation_Real_nummod.pdf       
-%user@user-G751JY:~/grammar-optim/results/correlations$ evince figures/posteriors/posterior_perRelation_Real_csubj.pdf 
-%user@user-G751JY:~/grammar-optim/results/correlations$
-
 \begin{table*} % TODO figure out away to make this table more aesthetic
 	\begin{center}	
-\small{
-\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
-	\hline
-&	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-% From Dryer	
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-%\hdashline
-
-\hline
-% Other universals, not in Dryer
-&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
-&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf.
-\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
-&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf.
-\cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
-&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section~\ref{sec:correlations} \\
-&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
-&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-\end{tabular}
-}
-
+\includegraphics[width=0.8\textwidth]{si-table-perrel-1.pdf}  
 \end{center}
@@
 
+%
+%%posterior_perRelation_Real_conj.pdf         posterior_perRelation_Real_goeswith.pdf     posterior_perRelation_Real_nummod.pdf       
+%%user@user-G751JY:~/grammar-optim/results/correlations$ evince figures/posteriors/posterior_perRelation_Real_csubj.pdf 
+%%user@user-G751JY:~/grammar-optim/results/correlations$ 
+%
+%\begin{table*} % TODO figure out away to make this table more aesthetic
+%	\begin{center}	
+%\small{
+%\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
+%	\hline
+%&	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+%% From Dryer	
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+%%\hdashline
+%
+%\hline
+%% Other universals, not in Dryer
+%&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
+%&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf.
+\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
+%&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf.
+\cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
+%&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section~\ref{sec:correlations} \\
+%&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
+%&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%\end{tabular}
+%}
+%
+%\end{center}
+%\caption{Predictions on UD relations with predictions from the typological literature.
+The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
+In the last column, we indicate what direction would be expected typologically.}
+%\label{tab:all-predictions-1}
+%\end{table*}
+%
+
 \footnotetext{Note that we report an \emph{anti-correlation}.
@@
 	\begin{center}	
-\small{
-\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
-	\hline
-	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-\hline
-appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\
-lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\
-expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\
-iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\
-vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\
-
-\hline
-% Uninterpretable
-compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\
-det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\
-dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\
-dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
-advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\
-
-\hline
-% UD artifacts
-conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
-discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\
-fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\
-flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
-goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
-list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
-orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\
-parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\
-reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\
-
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-\end{tabular}
-}
-
+\includegraphics[width=0.8\textwidth]{si-table-perrel-2.pdf}  
 \end{center}
@@
 \end{table*}
+
+%
+%\begin{table*} % TODO figure out away to make this table more aesthetic
+%	\begin{center}	
+%\small{
+%\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
+%	\hline
+%	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+%\hline
+%appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\
+%lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\
+%expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\
+%iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\
+%vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\
+%
+%\hline
+%% Uninterpretable
+%compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\
+%det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\
+%dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\
+%dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
+%advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\
+%
+%\hline
+%% UD artifacts
+%conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
+%discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\
+%fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\
+%flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
+%goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
+%list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
+%orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\
+%parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\
+%reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\
+%
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%\end{tabular}
+%}
+%
+%\end{center}
+%\caption{Predictions on UD relations for which no predictions are available in the typological literature.
+``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
+``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
+%\label{tab:all-predictions-2}
+%\end{table*}
+%
+%
 

### 2019-05-23 second-draft/si-table-perrel-1.tex
SI

% https://tex.stackexchange.com/questions/171982/how-to-make-table-floats-work-with-preview-standalone-environment/171983

\documentclass[%
    ,float=false % this is the new default and can be left away.
    ,preview=true
    ,class=scrartcl
%    ,fontsize=20pt
    ]{standalone}

\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{xcite}
\externalcitedocument{pnas-word-order-si}

\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage{siunitx}

\usepackage{longtable}

\usepackage[numbers]{natbib}

\usepackage{hyperref}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{arydshln}

\newcommand{\key}[1]{\textbf{#1}}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}

\usepackage{mathtools}
\DeclarePairedDelimiter{\ceiling}{\lceil}{\rceil}

\begin{document}
\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
	\hline
&	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
% From Dryer	
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
%\hdashline

\hline
% Other universals, not in Dryer
&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf. \cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section~\ref{sec:correlations} \\
&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\ 
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
\end{tabular}
\end{document}

### 2019-05-23 second-draft/si-table-perrel-2.tex
SI

% https://tex.stackexchange.com/questions/171982/how-to-make-table-floats-work-with-preview-standalone-environment/171983

\documentclass[%
    ,float=false % this is the new default and can be left away.
    ,preview=true
    ,class=scrartcl
%    ,fontsize=20pt
    ]{standalone}

\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{xcite}
\externalcitedocument{pnas-word-order-si}

\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage{siunitx}

\usepackage{longtable}

\usepackage[numbers]{natbib}

\usepackage{hyperref}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{arydshln}

\newcommand{\key}[1]{\textbf{#1}}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}

\usepackage{mathtools}
\DeclarePairedDelimiter{\ceiling}{\lceil}{\rceil}

\begin{document}
\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
	\hline
	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
\hline
appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\           
lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\          
expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\         
iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\     
vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\

\hline
% Uninterpretable
compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\        
det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\        
dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\               
dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\

\hline
% UD artifacts
conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\ 
fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\ 
flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\                                        
parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\   
reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\

 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
\end{tabular}
\end{document}

### 2019-05-23 second-draft/pnas-word-order-si.tex
SI

--- 2019-05-22 second-draft/pnas-word-order-si.tex
@@
 \tableofcontents
+
+\textcolor{red}{parseability vs parsability}
 
@@
 
-%\textcolor{red}{Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
-%Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
-%The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order, finding 16 correlations to likely be universal.
-%Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:corr-dryer}.
+\paragraph{Other Greenberg Universals}
+Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
+Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
+The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order; these form the basis of our formalization in Table~\ref{table:greenberg-dryer}.
+%, finding 16 correlations to likely be universal.
+%Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:greenberg-dryer}.
 %Testing efficiency on the excluded correlations would be feasible with additional annotation of data.
 %For instance, one of these could be tested by labeling the existing databases for the distinction between manner adverbs from other kinds of adverbs.
-%For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).}
+%For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).
+There are a few other well-supported word order universals that are not correlations with verb-object order.
+This includes dominance principles~\cite{croft2003typology} such as the strong preference for subjects to precede objects.
+Furthermore, there has been interest in Greenberg's universals 18 and 20, which describe correlations not with verb-object order, but of different elements of noun phrases~\cite{cinque2005deriving, culbertson2014language, dryer2018order}.
+Future work should examine whether these universals can also be linked to efficiency optimization.
 
@@
 the value of $\lambda$ is constrained to be in $[0,1)$.
-This means, surprisal must be weighted less strongly than parsability.
-
-Greater values of $\lambda$ mathematically result in degenerate solutions.
+This means, surprisal must be weighted less strongly than parseability.
+
+Greater values of $\lambda$ can mathematically result in degenerate solutions.
 To show this, note that the following inequality always holds:
@@
 While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
-In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parsability, whereas values closer to $0$ diminish the role of predictability.
-
-In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
+In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parseability, whereas values closer to $0$ diminish the role of predictability.
+
+In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parseability.
 The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
@@
 Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}).
-We also verified the robustness of our results to small variations of $\lambda$, by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu}.
+We also verified the robustness of our results to small variations of $\lambda$, by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu-previous}.
 
@@
 \begin{tabular}{l||l|lll|llll|ll|llllll}
-Correlation & Prevalence & \multicolumn{3}{c}{Bayesian} & \multicolumn{4}{c}{Frequentist} \\ 
+ & Prevalence & \multicolumn{3}{c|}{Bayesian} & \multicolumn{4}{c|}{Frequentist} \\ 
 & & Mean & SD & $p(\beta \leq 0)$ & $\beta$ & SE & $z$ & p \\
@@
 
-%posterior_perRelation_Real_conj.pdf         posterior_perRelation_Real_goeswith.pdf     posterior_perRelation_Real_nummod.pdf       
-%user@user-G751JY:~/grammar-optim/results/correlations$ evince figures/posteriors/posterior_perRelation_Real_csubj.pdf 
-%user@user-G751JY:~/grammar-optim/results/correlations$
-
 \begin{table*} % TODO figure out away to make this table more aesthetic
 	\begin{center}	
-\small{
-\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
-	\hline
-&	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-% From Dryer	
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-%\hdashline
-
-\hline
-% Other universals, not in Dryer
-&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
-&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf.
-\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
-&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf.
-\cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
-&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section~\ref{sec:correlations} \\
-&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
-&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-\end{tabular}
-}
-
+\includegraphics[width=0.8\textwidth]{si-table-perrel-1-1.png}
 \end{center}
@@
 
-\footnotetext{Note that we report an \emph{anti-correlation}.
-The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
+\footnotetext{The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
 Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
-}
+For simplicity, we display this as a corelation in this table.}
 
@@
 	\begin{center}	
-\small{
-\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
-	\hline
-	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-\hline
-appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\
-lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\
-expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\
-iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\
-vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\
-
-\hline
-% Uninterpretable
-compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\
-det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\
-dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\
-dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
-advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\
-
-\hline
-% UD artifacts
-conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
-discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\
-fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\
-flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
-goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
-list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
-orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\
-parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\
-reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\
-
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-\end{tabular}
-}
-
+\includegraphics[width=0.8\textwidth]{si-table-perrel-2-1.png}  
 \end{center}
@@
 \subsection{Other Experiments}
-In Table~\ref{table:corr-resu} we report the results of the simulations we did before the main experiment reported in the main paper (left and center), together with their preregistrations, besides the results from the main experiment (right).
-These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) the first simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
+In Table~\ref{table:corr-resu} we report the results of our two previous, preregistered, simulations\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}, \url{https://aspredicted.org/blind.php?x=bg35x7}.
+For the results of the locality simulations, see the DLM results in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM}.} together with results from the main experiment.
+These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) one simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
 Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
@@
 	\begin{center}
-\small{
-\begin{tabular}{|c|c|ccc|c}
-	\hline
-	%&
-& $\lambda = 0.0$
-& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM}.} 
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} 
-	\\ \hline \hline 
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}		  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_case.pdf}     } 
- 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_case.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_cop.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_cop.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_aux.pdf}     }  reverse
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_aux.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_nmod.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_nmod.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}   }
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_acl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_acl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   }
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_mark.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_mark.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   }
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_obl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_obl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   }
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_xcomp.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_xcomp.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }  
-		\\
-&&&\\ \hline
- \hline
+\includegraphics[width=0.4\textwidth]{si-table-perrel-3-1.png}  
+	\end{center}
+
+	\begin{center}
+\begin{tabular}{cccc}
+$\lambda=0.0$ & $\lambda=0.9$ & $\lambda=0.9$ & $\lambda=1.0$ \\
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency09.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-efficiency-large.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf}
 \end{tabular}
-}
-	\end{center}
-\caption{\textcolor{red}{parseability color, aux inversion}Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
+	\end{center}
+
+\caption{Results from optimization experiments for different values of $\lambda$, including our two previous preregistered experiments.
+For comparison, we also show results for $\lambda=0$, corresponding to optimizing for parseability only (same results as reported in Tables (\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})).
+For $\lambda=0.9$, we report results from one preliminary preregistered experiment (left) and the final experiment (right).
+For $\lambda=1.0$, we report the other preliminary preregistered experiment.
+Giving similar weight to parseability and predictability -- that is, $\lambda$ close to $1$ -- results in more accurate predictions than choosing a small value of $\lambda$ such as $\lambda=0.0$.
+Note that $\lambda$ cannot take values smaller than zero, or greater than one, see Section \ref{sec:lambda}.
+}\label{table:corr-resu-previous}
 \end{table}
 
-\begin{table}[hbt!]
-	\begin{center}
-\small{
-\begin{tabular}{|c|ccc|c}
-	\hline
-	%&
-& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM}.} 
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} \\ \hline \hline 
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}		  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_case.pdf}     } 
- 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_case.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_cop.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_cop.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_aux.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_aux.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_nmod.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_nmod.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	 
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_acl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_acl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_mark.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_mark.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_obl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_obl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_xcomp.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_xcomp.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }  \\
-&&&\\ \hline
- \hline
-\end{tabular}
-}
-	\end{center}
-\caption{Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
-\end{table}
-
-%\footnotetext{}
-%  final experiment
-%  0.9
-%
-
-\begin{figure}[ht]
-	\begin{center}
-	\includegraphics[width=0.25\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency09.pdf}
-	\includegraphics[width=0.25\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf}
-	\end{center}
-	\caption{(Other experiments) Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior-prior}
-\end{figure}
-
-%\section{Joint optimization of grammar, parser, and language model parameters}
 \section{Creating Optimized Grammars}
@@
 % TODO get UAS for compatibility with other parts of SI
-%We then compared parsability of random and optimized word order grammars on English data.
+%We then compared parseability of random and optimized word order grammars on English data.
 
@@
 %DLM turns out to be predicted by our model
+
+\textcolor{red}{phantom cite}\citep{diessel2001ordering, dryer1980positional, wals}
 

### 2019-05-23 second-draft/pnas-word-order-si.tex
SI

@@
 \textcolor{red}{parseability vs parsability}
+
+\textcolor{red}{where do we say that efficiency etc are reported on the held-out set?}
 
@@
 
-In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.\footnote{Note that in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
-Due to use of stochastic gradient descent in optimization, there is some amount of noise in the optimization process.}
-
+\paragraph{Per-Language Results}
+In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.
 We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section~\ref{sec:lambda}), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
-
 Within each language, we estimate the Pareto curve based on the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
-
 Figure (REF) in the main paper shows the average of these per-language plots.
@@
 Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.
-All data are $z$-scored.}\label{fig:pareto-per-lang}
+All data are $z$-scored.
+Note that in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
+This reflects noise in the optimization process due to use of stochastic gradient descent, i.e., the actual Pareto frontier might be somewhat further away from the baselines than the estimated one.}\label{fig:pareto-per-lang}
 \end{figure}
 
+\paragraph{Robustness}
 In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
@@
 
-In Figure~\ref{fig:corr-eff-corr}, we plot efficiency, parseability, and predictability (parseability and predictability $z$-scored within language) as a function of the number of satisfied correlations, for the real grammars of the 51 languages.
+In Figure~\ref{fig:corr-eff-corr}, we plot efficiency, parseability, and predictability (all are $z$-scored within language, as in Study 1) as a function of the number of satisfied correlations, for the real grammars of the 51 languages.
 
@@
 We show predictions for the eight correlations on the level of individual languages in Figure~\ref{fig:per-lang}.
-
-We obtained predictions for individual languages and each of the eight relations as follows.
+We obtained these predictions for individual languages and each of the eight relations as follows.
 For each language and each of the objective functions (efficiency, predictability, parseability), we considered the optimized grammar that yielded the best value of this objective function among the eight optimized grammars (i.e., the grammar where the optimization procedure had been most successful).
@@
 
-Note that the Bayesian analysis  also estimates a posterior distribution of the number of satisfied correlations (see Figure~\ref{fig:posterior}), providing an elegant solution to the multiple-comparisons problem arising from analysing the eight correlation.
+Note that the Bayesian analysis  also estimates a posterior distribution of the number of satisfied correlations (see Figure~\ref{fig:posterior}), providing an elegant solution to the multiple-comparisons problem arising from analysing the eight correlations together.
 
@@
 
-In Figure~\ref{fig:posterior}, we plot the posterior number of correlations predicted to hold in most optimized grammars, as obtained from the Bayesian regression.
+In Figure~\ref{fig:posterior}, we plot the posterior distribution of the number of correlations predicted to hold in most optimized grammars, as obtained from the Bayesian regression.
+For each posterior sample, we say that the $j$-th correlation holds if the value of $\alpha_j$ in that posterior sample is positive.
+In the figure, we plot the fraction of posterior samples in which a given number of correlations is satisfied.
 In addition to grammars optimized for efficiency, we also report the result for grammars optimized for predictability and for parseability alone.
@@
 	\end{center}
-	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
+	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in the Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
 \end{figure}
@@
 We considered all UD syntactic relations occurring in at least two of the 51 languages.
-In Table~\ref{tab:all-predictions-1}, we present the data for the eight correlations discussed in the main paper, and for those other relations for which the typological literature provides data.
-Additionally, in Table~\ref{tab:all-predictions-1} we present data for the other UD relations, for which either no typological data is available, or which are not linguistically meaningful.
+In Table~\ref{tab:all-predictions-1}, we present the data for the eight correlations discussed in the main paper, and for those other relations for which the typological literature provides data.\footnote{The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
+Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
+For simplicity, we display this as a corelation in this table.}
+Additionally, in Table~\ref{tab:all-predictions-2} we present data for the other UD relations, for which either no typological data is available, or which are not linguistically meaningful.
 
@@
 Results in Table~\ref{tab:all-predictions-1} also show that optimizing for DLM makes predictions similar to efficiency optimization.
-See Section~\ref{sec:DLM} for further discussio for further discussionn.
+See Section~\ref{sec:DLM} for further discussion for further discussionn.
 
@@
 
-\footnotetext{The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
-Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
-For simplicity, we display this as a corelation in this table.}
-
 \begin{table*} % TODO figure out away to make this table more aesthetic
@@
 \subsection{Other Experiments}
-In Table~\ref{table:corr-resu} we report the results of our two previous, preregistered, simulations\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}, \url{https://aspredicted.org/blind.php?x=bg35x7}.
-For the results of the locality simulations, see the DLM results in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM}.} together with results from the main experiment.
+In Table~\ref{table:corr-resu-previous} we report the results of our two previous, preregistered, simulations\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}, \url{https://aspredicted.org/blind.php?x=bg35x7}.
+For the results of the locality simulations described in the first preregistration, see the DLM results in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM} for discussion.} together with results from the main experiment.
 These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) one simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
 Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
-
-\textcolor{red}{TODO decide on formatting}
 
@@
 For $\lambda=1.0$, we report the other preliminary preregistered experiment.
-Giving similar weight to parseability and predictability -- that is, $\lambda$ close to $1$ -- results in more accurate predictions than choosing a small value of $\lambda$ such as $\lambda=0.0$.
+Giving similar weight to parseability and predictability -- that is, $\lambda$ close to $1$ -- results in more accurate word order predictions than choosing a small value of $\lambda$ such as $\lambda=0.0$.
 Note that $\lambda$ cannot take values smaller than zero, or greater than one, see Section \ref{sec:lambda}.
@@
 In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb-object relation.
-
-WALS has some data for all but seven languages.\footnote{Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
-In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}
-WALS has data for 74\% of the entries, and lists a dominant order for 91\% of these.
+WALS has data for 74\% of the entries\footnote{Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
+In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}, and lists a dominant order for 91\% of these.
 The grammars we extracted from the corpora agree with WALS in 96~\%  of these cases.
@@
 Recall the definition of efficiency
-\begin{equation}\label{eq:efficiency}
+\begin{equation}\label{eq:efficiency-recall}
 	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
@@
 \end{equation}
-Using this, we can rewrite Efficiency~(\ref{eq:efficiency}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
-\begin{equation}
+Using this, we can rewrite Efficiency~(\ref{eq:efficiency-recall}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
+\begin{equation}\label{eq:efficiency-rewrite}
 	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi} := \max_{\phi,\psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
@@
 
-For gramar optimization, we apply Early Stopping~\citep{prechelt1998early} as follows.
-We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
-This is a standard technique for optimizing neural models, and helps prevent overfitting to the specific corpora used.
-
-The choice of optimization methods and the stopping criterion were fixed before we investigated word order correlations.
+All corpora have a predefined split in training and held-out (development) sets.
+We use the training set for optimizing parameters, and apply Early Stopping~\citep{prechelt1998early} using the held-out set.
+
+For \key{estimating the parseability or predictability} of a given grammar, we optimize the neural model on data ordered according to this grammar, and report the parseability/predictability on the held-out set to avoid overfitting to the training set.
+For Early Stopping, we evaluate on the held-out set at the end of every epoch. %run through the training set.
+
+For \key{optimizing grammars}, we jointly apply gradient descent to the grammar parameters and the neural models, using the gradient estimator (\ref{eq:dtheta}-\ref{eq:dpsi}).
+For Early Stopping, we evaluate on the held-out set in intervals of 50,000 sentences, using a Monte-Carlo estimate of $R_{\textit{Eff}}^{\theta, \phi, \psi}$ (\ref{eq:efficiency-rewrite}), sampling a single linearized sentence for each syntactic structure in the held-out set.
+When reporting the parseability/predictability of an optimized grammar, we evaluate these values for its fully deterministic version (Section~\ref{sec:diff-gramm}) to allow fair comparison with baseline grammars.
+
+%We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
+%This is a standard technique for optimizing neural models, and helps prevent overfitting to the specific corpora used.
+
+The choice of optimization methods and the stopping criterion were fixed before we investigated language efficiency or word order correlations.
 
 \paragraph{Optimized Grammars}
-Recall that, for each language, we created 8 optimized languages for each optimization criterion.
+As described in the main paper, for each language, we created 8 optimized languages for each optimization criterion.
 We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
@@
 
-\textcolor{red}{phantom cite}\citep{diessel2001ordering, dryer1980positional, wals}
+\nocite{diessel2001ordering, dryer1980positional, wals}
 

### 2019-05-25 second-draft/pnas-word-order-si.tex
SI

@@
 \textcolor{red}{where do we say that efficiency etc are reported on the held-out set?}
+
+\textcolor{red}{notation U, L(T)}
+
+\textcolor{red}{run czech parser}
+
+\textcolor{red}{run controls on additional random grammars}
 
@@
 The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
-We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (Table~\ref{tab:all-predictions-1}).
+To tease apart the relative contributions of the two components, we also examine word order predictions for grammars optimized only to optimize parseability, or only predictability.
+As shown in Table~\ref{tab:all-predictions-1}, each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions.
 %Indeed, the same holds for all of the 38 syntactic relations in UD  (Table~\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})
@@
 That is, at $\lambda$ close to its maximal value, the predictions of optimizing the two scoring functions individually add up to the predictions of efficiency optimization.
-Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}).
+Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}), in which case not all correlations are predicted (Figure~\ref{fig:posterior}).
 We also verified the robustness of our results to small variations of $\lambda$, by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu-previous}.
@@
 In this section, we describe the details of the neural network architectures.
-Choices follow standard practice and machine learning and were fixed before evaluating word order properties, and the efficiency of real languages.
+Choices follow standard practice in machine learning and were fixed before evaluating word order properties, and the efficiency of real languages.
 

### 2019-05-25 results/correlations/correlationsByObjOrder.R

@@ -158,7 +158,17 @@ plot = plot + theme(        legend.margin=margin(0,0,0,0), legend.box.margin=mar
 ggsave(plot, file="figures/correlations-curve.pdf", height=3, width=7)
 
 
+plot = ggplot(u, aes(x=Dir+4, y=..scaled.., color=DirObj)) 
+plot = plot + geom_density(size=1.5, linetype="dashed")
+plot = plot + facet_wrap(~ Type, nrow=1, scales="free_y") + xlim(-0.5,8.5) + theme_classic() + ylab("Grammars")   + xlab("Object Patterners Preceding Verb Patterners")
+plot = plot + theme(axis.text.y = element_blank())
+plot = plot + theme(legend.position="bottom") +  theme(legend.title = element_blank())
+plot = plot + theme(        legend.margin=margin(0,0,0,0), legend.box.margin=margin(-5,-10,2,-10))
+plot = plot + geom_segment(data=data.frame(x=c(1), xend=c(7), y=c(0), Type=c("Baselines")), aes(x=x, xend=xend, y=0, yend=0), color="white", size=1.8)
+plot = plot + geom_segment(data=data.frame(x=c(0), xend=c(8), y=c(0), Type=c("Optimized")), aes(x=x, xend=xend, y=0, yend=0), color="white", size=1.8)
+plot = plot + geom_segment(data=data.frame(x=c(0), xend=c(3.5), y=c(0), Type=c("Real")), aes(x=x, xend=xend, y=0, yend=0), color="white", size=1.8)
 
+ggsave(plot, file="figures/correlations-curve-whiteaxis.pdf", height=2.2, width=5.5)
 
 
 

### 2019-05-29 second-draft/pnas-word-order-si.tex
code

--- 2019-05-23 second-draft/pnas-word-order-si.tex
@@
 \renewcommand{\thesection}{S\arabic{section}}
+
+\newcommand{\utterance}{\mathcal{U}}
+\newcommand{\tree}{\mathcal{T}}
 
@@
 
-\textcolor{red}{parseability vs parsability}
+\textcolor{red}{word limit}
+
+\textcolor{red}{grammar example in SI}
+
+\textcolor{red}{cover letter goes to broad person}
+
+\textcolor{red}{Czech parser}
+
+\textcolor{red}{random all 50}
 
@@
 
-\paragraph{Other Greenberg Universals}
-Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
-Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
-The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order; these form the basis of our formalization in Table~\ref{table:greenberg-dryer}.
-%, finding 16 correlations to likely be universal.
-%Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:greenberg-dryer}.
+%\textcolor{red}{Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
+%Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
+%The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order, finding 16 correlations to likely be universal.
+%Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:corr-dryer}.
 %Testing efficiency on the excluded correlations would be feasible with additional annotation of data.
 %For instance, one of these could be tested by labeling the existing databases for the distinction between manner adverbs from other kinds of adverbs.
-%For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).
-There are a few other well-supported word order universals that are not correlations with verb-object order.
-This includes dominance principles~\cite{croft2003typology} such as the strong preference for subjects to precede objects.
-Furthermore, there has been interest in Greenberg's universals 18 and 20, which describe correlations not with verb-object order, but of different elements of noun phrases~\cite{cinque2005deriving, culbertson2014language, dryer2018order}.
-Future work should examine whether these universals can also be linked to efficiency optimization.
+%For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).}
 
@@
 where $I[S, R]$ describes the \key{informativity} of the signals $S$ about their referents $R$, and $H[S]$ describes the \key{complexity} of the communication system.
-While prior studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} mostly considered settings where the signals $S$ are individual words without further structure, the signals are entire sentences $\mathcal{U}$ in our setting.
-The underlying messages $R$ which the speaker aims to convey are the syntactic structures $\mathcal{T}$.
+While prior studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} mostly considered settings where the signals $S$ are individual words without further structure, the signals are entire sentences $\utterance$ in our setting.
+The underlying messages $R$ which the speaker aims to convey are the syntactic structures $\tree$.
 By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
@@
 \begin{equation}
-	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
 \end{equation}
@@
 \begin{equation}
-	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
 \end{equation}
@@
 the value of $\lambda$ is constrained to be in $[0,1)$.
-This means, surprisal must be weighted less strongly than parseability.
-
-Greater values of $\lambda$ can mathematically result in degenerate solutions.
+This means, surprisal must be weighted less strongly than parsability.
+
+Greater values of $\lambda$ mathematically result in degenerate solutions.
 To show this, note that the following inequality always holds:
 \begin{equation}
-I[U; T] \leq H[U].
+I[\utterance; \tree] \leq H[\utterance].
 \end{equation}
@@
 \begin{equation}
-    R_{Eff} = I[U; T] - \lambda H[U] \leq 0.
+    R_{Eff} = I[\utterance; \tree] - \lambda H[\utterance] \leq 0.
 \end{equation}
-and it takes the maximally possible value zero if $U$ maps all trees to a single utterance, in which case both $I[U; T]$ and $H[U]$ are zero.
+and it takes the maximally possible value zero if there is only a single utterance $\utterance$, in which case both $I[\utterance; \tree]$ and $H[\utterance]$ are zero.
 This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
 While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
-In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parseability, whereas values closer to $0$ diminish the role of predictability.
-
-In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parseability.
+In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parsability, whereas values closer to $0$ diminish the role of predictability.
+
+In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
 The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
@@
 Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}).
-We also verified the robustness of our results to small variations of $\lambda$, by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu-previous}.
+We also verified the robustness of our results to small variations of $\lambda$, by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu}.
 
@@
 %\begin{equation}
-%	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+%	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
 %\end{equation}
@@
 %\begin{equation}
-%	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+%	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
 %\end{equation}
@@
 \begin{tabular}{l||l|lll|llll|ll|llllll}
- & Prevalence & \multicolumn{3}{c|}{Bayesian} & \multicolumn{4}{c|}{Frequentist} \\ 
+Correlation & Prevalence & \multicolumn{3}{c}{Bayesian} & \multicolumn{4}{c}{Frequentist} \\ 
 & & Mean & SD & $p(\beta \leq 0)$ & $\beta$ & SE & $z$ & p \\
@@
 	\begin{center}	
-\includegraphics[width=0.8\textwidth]{si-table-perrel-1-1.png}
+\includegraphics[width=0.8\textwidth]{si-table-perrel-1.pdf}  
 \end{center}
@@
 
-\footnotetext{The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
+%
+%%posterior_perRelation_Real_conj.pdf         posterior_perRelation_Real_goeswith.pdf     posterior_perRelation_Real_nummod.pdf       
+%%user@user-G751JY:~/grammar-optim/results/correlations$ evince figures/posteriors/posterior_perRelation_Real_csubj.pdf 
+%%user@user-G751JY:~/grammar-optim/results/correlations$ 
+%
+%\begin{table*} % TODO figure out away to make this table more aesthetic
+%	\begin{center}	
+%\small{
+%\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
+%	\hline
+%&	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+%% From Dryer	
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+%%\hdashline
+%
+%\hline
+%% Other universals, not in Dryer
+%&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
+%&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf.
+\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
+%&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf.
+\cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
+%&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section~\ref{sec:correlations} \\
+%&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
+%&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%\end{tabular}
+%}
+%
+%\end{center}
+%\caption{Predictions on UD relations with predictions from the typological literature.
+The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
+In the last column, we indicate what direction would be expected typologically.}
+%\label{tab:all-predictions-1}
+%\end{table*}
+%
+
+\footnotetext{Note that we report an \emph{anti-correlation}.
+The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
 Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
-For simplicity, we display this as a corelation in this table.}
+}
 
@@
 	\begin{center}	
-\includegraphics[width=0.8\textwidth]{si-table-perrel-2-1.png}  
+\includegraphics[width=0.8\textwidth]{si-table-perrel-2.pdf}  
 \end{center}
@@
 
+%
+%\begin{table*} % TODO figure out away to make this table more aesthetic
+%	\begin{center}	
+%\small{
+%\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
+%	\hline
+%	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+%\hline
+%appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\
+%lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\
+%expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\
+%iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\
+%vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\
+%
+%\hline
+%% Uninterpretable
+%compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\
+%det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\
+%dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\
+%dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
+%advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\
+%
+%\hline
+%% UD artifacts
+%conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
+%discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\
+%fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\
+%flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
+%goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
+%list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
+%orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\
+%parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\
+%reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\
+%
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%\end{tabular}
+%}
+%
+%\end{center}
+%\caption{Predictions on UD relations for which no predictions are available in the typological literature.
+``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
+``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
+%\label{tab:all-predictions-2}
+%\end{table*}
+%
+%
+
 \subsection{Other Experiments}
-In Table~\ref{table:corr-resu} we report the results of our two previous, preregistered, simulations\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}, \url{https://aspredicted.org/blind.php?x=bg35x7}.
-For the results of the locality simulations, see the DLM results in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM}.} together with results from the main experiment.
-These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) one simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
+In Table~\ref{table:corr-resu} we report the results of the simulations we did before the main experiment reported in the main paper (left and center), together with their preregistrations, besides the results from the main experiment (right).
+These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) the first simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
 Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
@@
 	\begin{center}
-\includegraphics[width=0.4\textwidth]{si-table-perrel-3-1.png}  
+\small{
+\begin{tabular}{|c|c|ccc|c}
+	\hline
+	%&
+& $\lambda = 0.0$
+& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM}.} 
+& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
+& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} 
+	\\ \hline \hline 
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}		  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_case.pdf}     } 
+ 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_case.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_cop.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_cop.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_aux.pdf}     }  reverse
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_aux.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   }
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_nmod.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_nmod.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}   }
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_acl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_acl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   }
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_mark.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_mark.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   }
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_obl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_obl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     }  
+		\\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   }
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_xcomp.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_xcomp.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }  
+		\\
+&&&\\ \hline
+ \hline
+\end{tabular}
+}
 	\end{center}
-
+\caption{\textcolor{red}{parseability color, aux inversion}Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
+\end{table}
+
+\begin{table}[hbt!]
 	\begin{center}
-\begin{tabular}{cccc}
-$\lambda=0.0$ & $\lambda=0.9$ & $\lambda=0.9$ & $\lambda=1.0$ \\
-	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}&
-	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency09.pdf}&
-	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-efficiency-large.pdf}&
-	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf}
+\small{
+\begin{tabular}{|c|ccc|c}
+	\hline
+	%&
+& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM}.} 
+& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
+& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} \\ \hline \hline 
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}		  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_case.pdf}     } 
+ 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_case.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_cop.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_cop.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	 
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_aux.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_aux.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_nmod.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_nmod.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	 
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_acl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_acl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_mark.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_mark.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_obl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_obl.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     }  \\
+&&&\\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	
+			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_xcomp.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_xcomp.pdf}     }  
+		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }  \\
+&&&\\ \hline
+ \hline
 \end{tabular}
+}
 	\end{center}
-
-\caption{Results from optimization experiments for different values of $\lambda$, including our two previous preregistered experiments.
-For comparison, we also show results for $\lambda=0$, corresponding to optimizing for parseability only (same results as reported in Tables (\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})).
-For $\lambda=0.9$, we report results from one preliminary preregistered experiment (left) and the final experiment (right).
-For $\lambda=1.0$, we report the other preliminary preregistered experiment.
-Giving similar weight to parseability and predictability -- that is, $\lambda$ close to $1$ -- results in more accurate predictions than choosing a small value of $\lambda$ such as $\lambda=0.0$.
-Note that $\lambda$ cannot take values smaller than zero, or greater than one, see Section \ref{sec:lambda}.
-}\label{table:corr-resu-previous}
+\caption{Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
 \end{table}
 
+%\footnotetext{}
+%  final experiment
+%  0.9
+%
+
+\begin{figure}[ht]
+	\begin{center}
+	\includegraphics[width=0.25\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency09.pdf}
+	\includegraphics[width=0.25\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf}
+	\end{center}
+	\caption{(Other experiments) Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior-prior}
+\end{figure}
+
+%\section{Joint optimization of grammar, parser, and language model parameters}
 \section{Creating Optimized Grammars}
@@
 In this case, the grammar can be converted into an equivalent grammar of the form described in Materials and Methods, by extracting a single parameter in $[-1, 1]$ for each relation $\tau$.
+
+\paragraph{Grammar Example}
+
+\begin{tabular}{llllllllll}
+	                             & \multicolumn{3}{English}    &   \\
+	Relation                     &     & $a_\tau$ & $b_\tau$   &   \\
+	object (\textit{obj})        &     &          &            &   \\
+	oblique (\textit{obl})       &     &          &            &   \\
+	case (\textit{lifted\_case}) &     &          &            &   \\
+\end{tabular}
 
@@
 In this section, we describe how we optimized grammar parameters for efficiency.
-A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $\mathcal{L}_\theta(t) = \mathbf{w}$ linearizing the tree.
-More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $\mathbf{w}$.
+A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $u = \mathcal{L}_\theta(t)$ linearizing the tree.
+More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $u$.
 In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $p_{\mathcal{L}_\theta}(u|t)$ concentrates on a single ordering $u$.
@@
 \begin{equation}
-	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
+	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
 \end{equation}
 \begin{equation}
-	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
+	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
 \end{equation}
-where $t \sim \mathcal{T}$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
+where $t \sim \tree$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
 
@@
 %\begin{equation}
-%	R_{Pars}^\phi := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
+%	R_{Pars}^\phi := I[\utterance,\tree] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
 %\end{equation}
@@
 %
-%The parsers and language models appropriate to a tree distribution $\mathcal{T}$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
+%The parsers and language models appropriate to a tree distribution $\tree$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
 %\begin{equation}
@@
 At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
+The neural parser and language model are also randomly initialized at the beginning of optimization.
 Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.
@@
 
-We used the CKY algorithm to compute the parsing loss $H[T|\mathcal{L}_\theta(T)]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
+We used the CKY algorithm to compute the syntactic ambiguity $H[\tree|\utterance]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
 Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
 % TODO get UAS for compatibility with other parts of SI
-%We then compared parseability of random and optimized word order grammars on English data.
+%We then compared parsability of random and optimized word order grammars on English data.
 
@@
     \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
-	\caption{Parsing loss $H[T|\mathcal{L}_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).
-We report $H[T|\mathcal{L}_\theta(T)]$ normalized by sentence length.}
+	\caption{Parsing loss $H[\tree|\utterance]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).
+We report $H[\tree|\utterance]$ normalized by sentence length.}
     \label{fig:cky-parser}
@@
     \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
-	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').
-We report $H[T|\mathcal{L}_\theta(T)]$ normalized by sentence length.}
+	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[\tree|\utterance]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').
+We report $H[\tree|\utterance]$ normalized by sentence length.}
     \label{fig:distorted-parser}
@@
 
-        \caption{Parsing loss ($H[T|\mathcal{L}_\theta(T)]$, normalized by sentence length) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
+	\caption{Parsing loss ($H[\tree|\utterance]$, normalized by sentence length) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
     \label{fig:learning-czech}
@@
 %DLM turns out to be predicted by our model
-
-\textcolor{red}{phantom cite}\citep{diessel2001ordering, dryer1980positional, wals}
 

### 2019-05-29 second-draft/pnas-word-order-si.tex

--- 2019-05-25 second-draft/pnas-word-order-si.tex
@@
 
+\textcolor{red}{word limit}
+
+\textcolor{red}{grammar example in SI}
+
+\textcolor{red}{cover letter goes to broad person}
 \textcolor{red}{parseability vs parsability}
@@
 \textcolor{red}{run controls on additional random grammars}
+
+\textcolor{red}{Czech parser}
+
+\textcolor{red}{random all 50}
 
@@
 \begin{equation}
-I[U; T] \leq H[U].
+I[\utterance; \tree] \leq H[\utterance].
 \end{equation}
@@
 \begin{equation}
-    R_{Eff} = I[U; T] - \lambda H[U] \leq 0.
-\end{equation}
-and it takes the maximally possible value zero if $U$ maps all trees to a single utterance, in which case both $I[U; T]$ and $H[U]$ are zero.
+    R_{Eff} = I[\utterance; \tree] - \lambda H[\utterance] \leq 0.
+\end{equation}
+and it takes the maximally possible value zero if there is only a single utterance $\utterance$, in which case both $I[\utterance; \tree]$ and $H[\utterance]$ are zero.
 This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
@@
 
+\paragraph{Grammar Example}
+
+\begin{tabular}{llllllllll}
+	                             & \multicolumn{3}{English}    &   \\
+	Relation                     &     & $a_\tau$ & $b_\tau$   &   \\
+	object (\textit{obj})        &     &          &            &   \\
+	oblique (\textit{obl})       &     &          &            &   \\
+	case (\textit{lifted\_case}) &     &          &            &   \\
+\end{tabular}
+
 \subsection{Extracting Grammars from Datasets}
@@
 \begin{equation}
-	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
+	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
 \end{equation}
 \begin{equation}
-	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
-\end{equation}
-where $t \sim \mathcal{T}$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
+	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
+\end{equation}
+where $t \sim \tree$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
 
@@
 
-We used the CKY algorithm to compute the parsing loss $H[T|\mathcal{L}_\theta(T)]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
+We used the CKY algorithm to compute the syntactic ambiguity $H[\tree|\utterance]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
 Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.

### 2019-05-29 second-draft/si-table-perrel-3.tex

% https://tex.stackexchange.com/questions/171982/how-to-make-table-floats-work-with-preview-standalone-environment/171983

\documentclass[%
    ,float=false % this is the new default and can be left away.
    ,preview=true
    ,class=scrartcl
%    ,fontsize=20pt
    ]{standalone}

\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{xcite}
\externalcitedocument{pnas-word-order-si}

\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage{siunitx}

\usepackage{longtable}

\usepackage[numbers]{natbib}

\usepackage{hyperref}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{arydshln}

\newcommand{\key}[1]{\textbf{#1}}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}

\usepackage{mathtools}
\DeclarePairedDelimiter{\ceiling}{\lceil}{\rceil}

\begin{document}
\small{
\begin{tabular}{|c||c|cc|c|c}
	\hline
	%&
& $\lambda = 0.0$
& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} 
& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}} 
	\\ \hline \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}		  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   }
 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_case.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_case.pdf}     } 
		\\
&&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   }
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_cop.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_cop.pdf}     }  
		\\
&&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   }
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_aux.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_aux.pdf}     }  
		\\
&&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   }
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_nmod.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_nmod.pdf}     }  
		\\
&&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}   }
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_acl.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     }  
			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_acl.pdf}     }  
		\\
&&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   }
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_mark.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     }  
			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_mark.pdf}     }  
		\\
&&&\\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   }
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_obl.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } 
			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_obl.pdf}     }  
		\\
&&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   }
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_xcomp.pdf}     }  
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }  
			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_xcomp.pdf}     }  
		\\
&&&\\ \hline
 \hline
\end{tabular}}
\end{document}

### 2019-05-29 second-draft/pnas-word-order-si.tex
SI

--- 2019-05-29 second-draft/pnas-word-order-si.tex
@@
 
-\paragraph{Grammar Example}
-
-\begin{tabular}{llllllllll}
-	                             & \multicolumn{3}{English}    &   \\
-	Relation                     &     & $a_\tau$ & $b_\tau$   &   \\
-	object (\textit{obj})        &     &          &            &   \\
-	oblique (\textit{obl})       &     &          &            &   \\
-	case (\textit{lifted\_case}) &     &          &            &   \\
+\begin{figure}
+	\begin{center}
+\begin{tabular}{l||l|ll||l|lllll}
+	& \multicolumn{3}{c||}{English}    &  \multicolumn{3}{c}{Japanese}  \\ 
+	Relation                     &  Par.   & $a_\tau$ & $b_\tau$            & Par.   & $a_\tau$ & $b_\tau$     \\ \hline \hline
+	object (\textit{obj})        &  $0.1$   &   $0.04$   & $-1.46$            & $-0.1$  & $0.99$ & $-0.7$2 \\
+	oblique (\textit{obl})       &  $0.3$   &   $0.13$     & $1.25$             & $-0.3$  & $0.99$ & $0.73$ \\
+	case (\textit{lifted\_case}) &  $0.2$   &   $0.07$       &   $-0.89$         & $-0.2$  &  $0.92$ & $0.02$  \\
 \end{tabular}
+	\end{center}
+	\caption{Sample Coefficients from grammars extracted from the real English and Japanese orderings, for the relations occurring in Figure 3 (Main Paper).
+We show parameters in $[-1,1]$ for deterministic word order grammars, and the coefficients ($a_\tau, b_\tau$) for corresponding differentiable ordering grammars.
+For the deterministic grammars (`Par.'), positive coefficients indicate that the dependent will be placed after the head.
+For the differentiable grammars, $a_\tau > 0.5$ indicates ordering of dependents before heads, and larger $b_\tau$ indicates greater distance between head and dependent.}\label{fig:grammar-sample}
+\end{figure}
 
@@
 We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.
+
+We provide an example in Figure~\ref{fig:grammar-sample}, illustrating grammar parameters for the relations in Figure 3 of the main paper.
 

### 2019-05-30 second-draft/pnas-word-order-si.tex
SI

@@
 
+\textcolor{red}{parseability vs parsability}
+
+\textcolor{red}{where do we say that efficiency etc are reported on the held-out set?}
+
+\textcolor{red}{notation U, L(T)}
+
+\textcolor{red}{run czech parser}
+
+\textcolor{red}{run controls on additional random grammars}
+
 \textcolor{red}{word limit}
@@
 \textcolor{red}{random all 50}
+
+\textcolor{red}{keep in mind: pareto plot unify with SI plot}
 
@@
 While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
-In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parsability, whereas values closer to $0$ diminish the role of predictability.
-
-In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
+In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parseability, whereas values closer to $0$ diminish the role of predictability.
+
+In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parseability.
 The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
-We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (Table~\ref{tab:all-predictions-1}).
+To tease apart the relative contributions of the two components, we also examine word order predictions for grammars optimized only to optimize parseability, or only predictability.
+As shown in Table~\ref{tab:all-predictions-1}, each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions.
 %Indeed, the same holds for all of the 38 syntactic relations in UD  (Table~\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})
@@
 That is, at $\lambda$ close to its maximal value, the predictions of optimizing the two scoring functions individually add up to the predictions of efficiency optimization.
-Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}).
-We also verified the robustness of our results to small variations of $\lambda$, by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu}.
+Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}), in which case not all correlations are predicted (Figure~\ref{fig:posterior}).
+We also verified the robustness of our results to small variations of $\lambda$, by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu-previous}.
 
@@
 
-In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.\footnote{Note that in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
+\paragraph{Per-Language Results}
+In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.
+%\footnote{Note that in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
 Due to use of stochastic gradient descent in optimization, there is some amount of noise in the optimization process.}
@@
 This has been suggested as a property of efficient word order in prior work, and, indeed, we found that DLM is predicted by efficiency optimization (Section~\ref{sec:DLM}).
-Results in Table~\ref{tab:all-predictions-1} also show that optimizing for DLM makes predictions similar to efficiency optimization.
-See Section~\ref{sec:DLM} for further discussio for further discussionn.
+Results in Table~\ref{tab:all-predictions-1} also show that optimizing for DLM makes predictions similar to efficiency optimization. See Section~\ref{sec:DLM} for further discussion.
 
@@
 	\begin{center}	
-\includegraphics[width=0.8\textwidth]{si-table-perrel-1.pdf}  
+\includegraphics[width=0.8\textwidth]{si-table-perrel-1-1.png}
 \end{center}
@@
 	\begin{center}	
-\includegraphics[width=0.8\textwidth]{si-table-perrel-2.pdf}  
+\includegraphics[width=0.8\textwidth]{si-table-perrel-2-1.png}  
 \end{center}
@@
 \subsection{Other Experiments}
-In Table~\ref{table:corr-resu} we report the results of the simulations we did before the main experiment reported in the main paper (left and center), together with their preregistrations, besides the results from the main experiment (right).
-These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) the first simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
+In Table~\ref{table:corr-resu-previous} we report the results of our two previous, preregistered, simulations\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}, \url{https://aspredicted.org/blind.php?x=bg35x7}.
+For the results of the locality simulations described in the first preregistration, see the DLM results in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM} for discussion.} together with results from the main experiment.
+These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) one simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
 Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
@@
 %DLM turns out to be predicted by our model
+
+\nocite{diessel2001ordering, dryer1980positional, wals}
 

### 2019-05-30 second-draft/pnas-word-order-si.tex
SI

@@
 
-\textcolor{red}{TODO decide on formatting}
-
 \begin{table}[hbt!]
-	\begin{center}
-\small{
-\begin{tabular}{|c|c|ccc|c}
-	\hline
-	%&
-& $\lambda = 0.0$
-& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM}.} 
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} 
-	\\ \hline \hline 
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}		  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_case.pdf}     } 
- 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_case.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_cop.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_cop.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_aux.pdf}     }  reverse
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_aux.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   }
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_nmod.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_nmod.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}   }
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_acl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_acl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   }
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_mark.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_mark.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   }
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_obl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_obl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     }  
-		\\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   }
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_xcomp.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_xcomp.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }  
-		\\
-&&&\\ \hline
- \hline
+
+	\begin{center}
+\includegraphics[width=0.4\textwidth]{si-table-perrel-3-1.png}  
+\end{center}
+
+	\begin{center}
+\begin{tabular}{cccc}
+$\lambda=0.0$ & $\lambda=0.9$ & $\lambda=0.9$ & $\lambda=1.0$ \\
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency09.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-efficiency-large.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf}
 \end{tabular}
-}
-	\end{center}
-\caption{\textcolor{red}{parseability color, aux inversion}Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
+	\end{center}
+\caption{Results from optimization experiments for different values of $\lambda$, including our two previous preregistered experiments.
+For comparison, we also show results for $\lambda=0$, corresponding to optimizing for parseability only (same results as reported in Tables (\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})).
+For $\lambda=0.9$, we report results from one preliminary preregistered experiment (left) and the final experiment (right).
+For $\lambda=1.0$, we report the other preliminary preregistered experiment.
+Giving similar weight to parseability and predictability -- that is, $\lambda$ close to $1$ -- results in more accurate word order predictions than choosing a small value of $\lambda$ such as $\lambda=0.0$.
+Note that $\lambda$ cannot take values smaller than zero, or greater than one, see Section \ref{sec:lambda}.
+}\label{table:corr-resu-previous}
 \end{table}
-
-\begin{table}[hbt!]
-	\begin{center}
-\small{
-\begin{tabular}{|c|ccc|c}
-	\hline
-	%&
-& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM}.} 
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} \\ \hline \hline 
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}		  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_case.pdf}     } 
- 		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_case.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_cop.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_cop.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_aux.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_aux.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_nmod.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_nmod.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	 
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_acl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_acl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_lifted_mark.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_lifted_mark.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_obl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_obl.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     }  \\
-&&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	
-			&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first10_Efficiency_xcomp.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_first09_Efficiency_xcomp.pdf}     }  
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }  \\
-&&&\\ \hline
- \hline
-\end{tabular}
-}
-	\end{center}
-\caption{Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}
-\end{table}
-
-%\footnotetext{}
-%  final experiment
-%  0.9
-%
-
-\begin{figure}[ht]
-	\begin{center}
-	\includegraphics[width=0.25\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency09.pdf}
-	\includegraphics[width=0.25\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf}
-	\end{center}
-	\caption{(Other experiments) Posterior of the number of correlations correctly predicted by efficiency and its components, in a Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior-prior}
-\end{figure}
 

### 2019-05-30 second-draft/pnas-word-order-si.tex
SI

@@
 \end{equation}
-Using this, we can rewrite Efficiency~(\ref{eq:efficiency}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
+Using this, we can rewrite Efficiency~(\ref{eq:efficiency-recall}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
 \begin{equation}
 	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi} := \max_{\phi,\psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
-\end{equation}
+\end{equation}\label{eq:efficiency-rewrite}
 %

### 2019-05-31 second-draft/pnas-word-order-si.tex
SI

@@
 
-\textcolor{red}{parseability vs parsability}
-
-\textcolor{red}{where do we say that efficiency etc are reported on the held-out set?}
-
-\textcolor{red}{notation U, L(T)}
-
 \textcolor{red}{run czech parser}
 
-\textcolor{red}{run controls on additional random grammars}
-
-\textcolor{red}{word limit}
-
-\textcolor{red}{grammar example in SI}
-
-\textcolor{red}{cover letter goes to broad person}
-
-\textcolor{red}{Czech parser}
-
-\textcolor{red}{random all 50}
-
 \textcolor{red}{keep in mind: pareto plot unify with SI plot}
 
-\section{Formalization of Correlations}\label{sec:correlations}
-
-Here we describe how we selected the correlations in Table 1 of the main paper, and how we formalized these using syntactic relations defined by Universal Dependencies.
-
-We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.
-%TODO REF mention this? \footnote{Regarding objections by \cite{dunn2011evolved}, we refer to the follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
+\section{Formalization of Greenberg Correlation Universals}\label{sec:correlations}
+
+Here we describe how we selected the word order correlations in Table 1 of the main paper, and how we formalized these using syntactic relations defined by Universal Dependencies.
+
+We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding objections by \citet{dunn2011evolved}, we refer to the follow-ups by \citet{levy2011computational}, and \citet{croft2011greenbergian}.}
 Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
@@
 
-Note that we follow \cite{futrell2015largescale} in converting the Universal Dependencies format to a format closer to standard syntactic theory, promoting adpositions, copulas, and complementizers to heads.
+As described in \emph{Materials and Methods}, we follow \citet{futrell2015largescale} in converting the Universal Dependencies format to a format closer to standard syntactic theory, promoting adpositions, copulas, and complementizers to heads.
 As a consequence, the direction of the relations \emph{case}, \emph{cop}, and \emph{mark} is reversed compared to Universal Dependencies.
@@
 	\hline
-&	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{UD Relation}   & \multirow{2}{*}{Greenberg \cite{greenberg1963universals}}    \\ 
+&	\multicolumn{2}{c|}{Correlates with...}   &          \multirow{2}{*}{UD Relation}   & \multirow{2}{*}{\citet{greenberg1963universals}}    \\ 
 &	verb & object & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
@@
 \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&`want'    &    VP    &    $\xrightarrow{xcomp}$   & 15   \\\hline
- \hline
 \end{tabular}
@@
 	\caption{Greenbergian Correlations based on Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies using the modified format of \cite{futrell2015largescale} (see text).
-	For reference, we also provide the numbers of the closest corresponding universals stated in Greenberg's original study (if possible).
+	For reference, we also provide the numbers of the closest corresponding universals stated in Greenberg's original study, to the extent that this is possible.
 	}\label{table:greenberg-dryer}
@@
 
-We further excluded the verb-subject correlation, which is not satisfied by much more than half of the world's languages (51 \% among those with annotation in \cite{wals-81}, with clear violation in 35.4 \%).
+We further excluded the verb-subject correlation, which is not satisfied by much more than half of the world's languages (51 \% among those with annotation in the \emph{World Atlas of Language Structures} \cite{wals-81}, with clear violation in 35.4 \%).
 It is satisfied only in 33\% of our sample of 51 UD languages, as quantified using the grammars we extracted.
@@
 
-Here we discuss how our formalization of communicative efficiency relates to other information-theoretic work on language.
-Our formalization of efficiency is based on the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems.
-The key idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms.
+Here we discuss how our formalization of communicative efficiency relates to formalizations that have been proposed in the information-theoretic literature on language.
+Across the literature, the core idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms, with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global, kemp2012kinship,xu2014numeral, regier2015word,xu2016historical, zaslavsky2018efficient,futrell2017memory,peloquin2019interactions, zaslavsky2019semantic, frank2012predicting,goodman2013knowledge, kao2014nonliteral, bennett2018extremely}.
+
+The basis for our precise formalization is the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems.
+%The key idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms.
 %; there are some differences in the formalization of complexity.
-Formally, if $S$ denotes signals (e.g., words, sentences) and $R$ denotes their referents (e.g., objects in a reference game), then this efficiency metric takes the form (notation slightly varies across these publications):
+If $S$ denotes signals (e.g., words, sentences) and $R$ denotes their referents (e.g., objects in a reference game), then this efficiency metric takes the form (notation slightly varies across these publications):
 \begin{equation}\label{eq:eff-general}
@@
 We therefore arrive at the following efficiency metric:
-\begin{equation}\label{eq:efficiency}
+\begin{equation}\label{eq:efficiency-derived}
 	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
@@
 \end{equation}
-This efficiency metric~(\ref{eq:eff-general}) is also equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
-The Information Bottleneck models complexity using a mutual information term instead of the entropy.
+
+We now discuss how this metric relates to other variants that have been proposed in the information-theoretic literature on language, and describe why~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is particularly appropriate to our setting.
+This efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
+The Information Bottleneck models complexity using a mutual information term instead of the entropy appearing in~(\ref{eq:eff-general}).
 This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
 As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure.
-Entropy also arises naturally as a complexity measure from the role played by surprisal in determining human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
-
-A few other approaches share the mutual information term for informativity, while using complexity measures that are not explicitly information-theoretic.
+Furthermore, entropy also arises naturally as a complexity measure from the role played by surprisal in determining human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+
+A few other approaches share the mutual information term for informativity in (\ref{eq:eff-general}), while using complexity measures that are not explicitly information-theoretic.
 %A few models have used other complexity measures that are not explicitly information-theoretic quantities.
-In \cite{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \cite{kemp2012kinship} it is the difficulty of defining the concepts that are encoded.
+In \cite{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \cite{kemp2012kinship} it is the complexity of defining the concepts that are encoded.
 The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
@@
 This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
-While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
+While the design of our word order grammars (see \emph{Materials and Methods}) precludes a collapse of all syntactic structures to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable description of the efficiency of communication systems.
 In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parseability, whereas values closer to $0$ diminish the role of predictability.
@@
 In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parseability.
-The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
-To tease apart the relative contributions of the two components, we also examine word order predictions for grammars optimized only to optimize parseability, or only predictability.
+While the computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$, we also examined word order predictions for grammars optimized for only parseability or only predictability, in order to tease apart predictions made by these two components.
 As shown in Table~\ref{tab:all-predictions-1}, each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions.
@@
 %We additionally report results with $\lambda = 0$, i.e., placing all weight on parseability.
-That is, at $\lambda$ close to its maximal value, the predictions of optimizing the two scoring functions individually add up to the predictions of efficiency optimization.
+That is, at $\lambda$ close to its maximal value, the predictions of optimizing the two scoring functions individually add up to the predictions of efficiency optimization.\footnote{Results from one of the preliminary experiments reported in Figure~\ref{table:corr-resu-previous} show that results are stable to small variation of $\lambda$: Essentially equivalent predictions are obtained for $\lambda=1.0$.
+While $\lambda=1.0$ is not a valid choice for communicative efficiency in general due to the possibility of collapse to a single utterance, our setting does not allow such a collapse, as the syntactic structure already determines which words are present in the sentence.}
 Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}), in which case not all correlations are predicted (Figure~\ref{fig:posterior}).
-We also verified the robustness of our results to small variations of $\lambda$, by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu-previous}.
+%Finally, the robustness of our results to small variations of $\lambda$ is confime by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu-previous}.
 
@@
 
-\paragraph{Per-Language Results}
 In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.
+
 %\footnote{Note that in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
@@
 
-Figure (REF) in the main paper shows the average of these per-language plots.
+Figure 4 in the main paper shows the average of these per-language plots.
 
@@
 
-\paragraph{Robustness}
 In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
@@
 
-\subsection{Other Experiments}
+\subsection{Previous Experiments}\label{sec:previous-exps}
 In Table~\ref{table:corr-resu-previous} we report the results of our two previous, preregistered, simulations\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}, \url{https://aspredicted.org/blind.php?x=bg35x7}.
@@
 
-In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
+In this section, we describe the method we employ for creating grammars that are optimized for efficiency, and how we extract grammars describing the actual ordering rules of languages.
 %Our method will be applicable more generally to the task of optimizing grammars for a broad class of objective functions.
-We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
+We carry out grammar optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
 More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
@@
 In this case, the grammar can be converted into an equivalent grammar of the form described in Materials and Methods, by extracting a single parameter in $[-1, 1]$ for each relation $\tau$.
+
+We provide an example in Figure~\ref{fig:grammar-sample}, illustrating grammar parameters for the relations in Figure 3 of the main paper.
 
@@
 	\end{center}
-	\caption{Sample Coefficients from grammars extracted from the real English and Japanese orderings, for the relations occurring in Figure 3 (Main Paper).
-We show parameters in $[-1,1]$ for deterministic word order grammars, and the coefficients ($a_\tau, b_\tau$) for corresponding differentiable ordering grammars.
+	\caption{Sample Coefficients from grammars extracted from the real English and Japanese orderings (Section~\ref{sec:extract-grammars}), for the relations occurring in Figure 3 (Main Paper).
+We show parameters in $[-1,1]$ for deterministic word order grammars as described in \emph{Materials and Methods}, and the coefficients ($a_\tau, b_\tau$) for corresponding differentiable ordering grammars.
 For the deterministic grammars (`Par.'), positive coefficients indicate that the dependent will be placed after the head.
-For the differentiable grammars, $a_\tau > 0.5$ indicates ordering of dependents before heads, and larger $b_\tau$ indicates greater distance between head and dependent.}\label{fig:grammar-sample}
+For the differentiable grammars, $a_\tau > 0.5$ indicates predominance of ordering of dependents before heads, and larger $b_\tau$ indicates greater distance between head and dependent.}\label{fig:grammar-sample}
 \end{figure}
 
-\subsection{Extracting Grammars from Datasets}
+\subsection{Extracting Grammars from Datasets}\label{sec:extract-grammars}
 We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
@@
 We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.
-
-We provide an example in Figure~\ref{fig:grammar-sample}, illustrating grammar parameters for the relations in Figure 3 of the main paper.
 
@@
 The grammars we extracted from the corpora agree with WALS in 96~\%  of these cases.
+
+Note that the grammatical formalism simplifies some aspects of the word order regularities of natural languages.
+For instance, it does not represent cases where ordering varies between main and embedded clauses, as it does not condition ordering decisions on the larger context.
+More complex and powerful ordering grammar models have been proposed \cite{futrell2015experiments, wang2016galactic}; however, they have similar limitations, and for our purposes, the model adopted here has the advantage of being simple and interpretable.
+%There currently is no grammatical formalism that is at the same time capable of representing the full range of word order regularities that are found, 
+%\cite{futrell2015experiments}
 
@@
 In this section, we describe the details of the neural network architectures.
-Choices follow standard practice in machine learning and were fixed before evaluating word order properties, and the efficiency of real languages.
+Choices follow standard practice in machine learning.
+All choices, except where explicitly noted otherwise, were made before evaluating word order properties, and the efficiency of real grammars.
 
@@
 We choose a standard LSTM language model \citep{goldberg2017neural, hochreiter1997long}.
-We restrict the vocabulary to the most frequent 50,000 words in the treebanks for a given language.
+For computational reasons, we restrict the vocabulary to the most frequent 50,000 words in the treebanks for a given language.
 Given the moderate size of the corpora, this limit is only attained only for few languages.
@@
 \paragraph{Hyperparameters}
-
-%\textbf{TODO study 1 only has 1 set of parameters}
-
 Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
@@
 We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
-This maximizes statistical precision in detecting and quantifying correlations between the \textit{obj} relation and other relations.
+This maximizes statistical precision in detecting and quantifying correlations between the verb-object relation and other relations.
 
 For efficiency optimization, for each grammar, we ran efficiency optimization with three different random seeds, selecting among these the seed that yielded the best overall efficiency value.
-This helps control for variation across random seeds for the stochastic gradient descent optimization method.
+We did this in order to control for possible variation across random seeds for the stochastic gradient descent optimization method.
+As described in our preregistration \url{http://aspredicted.org/blind.php?x=ya4qf8}, this choice was made after conducting a preliminary version of Study 2 reported in Section~\ref{sec:previous-exps}; results reported there show essentially identical results.
 

### 2019-05-31 second-draft/compile.sh
moving files

@@ -1,12 +1,0 @@
-pdflatex pnas-word-order-si.tex
-bibtex pnas-word-order-si
-pdflatex si-table-perrel-1.tex
-pdflatex si-table-perrel-2.tex
-pdflatex si-table-perrel-3.tex
-
-pdftoppm si-table-perrel-1.pdf si-table-perrel-1 -png
-pdftoppm si-table-perrel-2.pdf si-table-perrel-2 -png
-pdftoppm si-table-perrel-3.pdf si-table-perrel-3 -png
-
-pdflatex pnas-word-order-si.tex
-

### 2019-05-31 writeup/pnas-word-order-si.tex

@@
-\documentclass[12pt]{article}
+\documentclass[10pt,twoside,lineno]{article}
+% Use the lineno option to display guide line numbers if required.
+
+%\usepackage{lineno}
+%\linenumbers
+
+%\templatetype{pnassupportinginfo}
+
+% \readytosubmit %% Uncomment this line before submitting, so that the instruction page is removed.
+
 \usepackage{amsmath}
@@
 \DeclareMathOperator{\E}{\mathop{\mathbb{E}}}
+
+\usepackage{footnote}
+\makesavenoteenv{tabular}
+\makesavenoteenv{table}
+
+\usepackage{siunitx}
+
+\usepackage[cm]{fullpage}
+
+\usepackage{longtable}
+
+\usepackage[numbers]{natbib}
 
@@
 
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+\renewcommand{\thesection}{S\arabic{section}}
+
+\newcommand{\utterance}{\mathcal{U}}
+\newcommand{\tree}{\mathcal{T}}
+
 \usepackage{mathtools}
@@
 
-\section{Results on all UD Relations}
-Table~\ref{tab:all-predictions} shows the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
+\tableofcontents
 
-\begin{table*} % TODO figure out away to make this table more aesthetic
+\textcolor{red}{run czech parser}
+
+\textcolor{red}{keep in mind: pareto plot unify with SI plot}
+
+\section{Formalization of Greenberg Correlation Universals}\label{sec:correlations}
+
+Here we describe how we selected the word order correlations in Table 1 of the main paper, and how we formalized these using syntactic relations defined by Universal Dependencies.
+
+We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding objections by \citet{dunn2011evolved}, we refer to the follow-ups by \citet{levy2011computational}, and \citet{croft2011greenbergian}.}
+Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
+\cite{dryer1992greenbergian} formulated these universals as correlations between the order of objects and verbs and the orders of other syntactic relations.
+We test our ordering grammars for these correlations by testing whether the coefficients for these syntactic relations have the same sign as the coefficient of the verb-object relation.
+Testing correlations is therefore constrained by the degree to which these relations are annotated in UD.
+The verb-object relation corresponds to the  \emph{obj} relation defined by UD.
+While most of the other relations also correspond to UD relations, some are not annotated reliably.
+We were able formalize eleven out of Dryer's seventeen correlations in UD.
+Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations:
+First, tense/aspect and negative auxiliaries are together represented by the \emph{aux} relation in UD.
+Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{mark} relation.
+Third, both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.
+
+The resulting operationalization is shown in Table~\ref{table:greenberg-dryer}.
+For each relation, we show the direction of the UD syntactic relation: $\rightarrow$ indicates that the verb patterner is the head; $\leftarrow$ indicates that the object patterner is the head.
+
+As described in \emph{Materials and Methods}, we follow \citet{futrell2015largescale} in converting the Universal Dependencies format to a format closer to standard syntactic theory, promoting adpositions, copulas, and complementizers to heads.
+As a consequence, the direction of the relations \emph{case}, \emph{cop}, and \emph{mark} is reversed compared to Universal Dependencies.
+For clarity, we refer to these reversed relations as \emph{lifted\_case}, \emph{lifted\_cop}, and \emph{lifted\_mark}.
+
+\begin{table*}[ht]
 	\begin{center}
-
-\tiny{
-\begin{tabular}{|l|l|l|ll|l|l|}
+\small{
+\begin{tabular}{|l|ll|l|l|}
 	\hline
-	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-% From Dryer	
-acl  &  80   &   \textbf{85}$^{***}$   &   43   &   \textbf{82}$^{***}$   &   \textbf{71}$^{***}$  & $>50\%$  \citep{dryer1992greenbergian} \\
-aux  &  12   &   \textbf{26}$^{***}$   &   \textbf{19}$^{***}$   &   49   &   \textbf{35}$^{**}$  & $<50\%$ \citep{dryer1992greenbergian} \\
-lifted\_case  &  86   &   \textbf{81}$^{***}$   &   \textbf{40}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{67}$^{***}$ & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-lifted\_cop  &  94   &   \textbf{80}$^{***}$   &   59   &   \textbf{72}$^{***}$   &   \textbf{63}$^{**}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
-lifted\_mark  &  76   &   \textbf{85}$^{***}$   &   \textbf{57}$^{*}$   &   \textbf{76}$^{***}$   &   \textbf{72}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
-nmod  &  80   &   \textbf{82}$^{***}$   &   49   &   \textbf{71}$^{***}$   &   \textbf{68}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\ 
-obl  &  88   &   \textbf{78}$^{***}$   &   \textbf{71}$^{***}$   &   46   &   \textbf{68}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
-xcomp  &  88   &   \textbf{90}$^{***}$   &   \textbf{76}$^{***}$   &   \textbf{89}$^{***}$   &   \textbf{85}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
-\hdashline
-nsubj  &  33   &   \textbf{29}$^{***}$   &   53   &   \textbf{7}$^{***}$   &   \textbf{23}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
-advmod  &  35   &   51   &   \textbf{18}$^{***}$   &   47   &   \textbf{39}$^{**}$  & $>50\%$ for manner adverbs; $\approx 50\%$ for negation, TAM, intensifiers \citep{dryer1992greenbergian} \\
-
-\hline
-% Other universals, not in Dryer
-advcl  &  86   &   \textbf{84}$^{***}$   &   51   &   \textbf{67}$^{***}$   &   \textbf{67}$^{*}$  & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
-amod  &  51   &   \textbf{84}$^{***}$   &   42   &   \textbf{53}$^{*}$   &   51   & $\approx 50\%$ \citep{dryer1992greenbergian} \\
-nummod  &  37   &   \textbf{74}$^{**}$   &   49   &   52   &   52 & $\approx 50\%$ \citep[][89A vs. 83A]{wals} \\
-
-\hline
-% Typological data not available
-appos  &  77   &   \textbf{77}$^{***}$   &   49   &   \textbf{67}$^{***}$   &   \textbf{62}$^{**}$ &  Unknown \\%($>50\%$?
-correlates with \textit{nmod} according to Greenberg U 23)  \\ % \cite{cinque2009greenberg} claims it correlates with OV/VO.
-However, it's not clear to me how closely appos matches the subject of the typological claim
-lifted\_cc  &  71   &   \textbf{83}$^{***}$   &   51   &   \textbf{81}$^{***}$   &   \textbf{71}$^{***}$  & Unknown \\
-ccomp  &  86   &   \textbf{88}$^{***}$   &   \textbf{70}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{77}$^{**}$ & Unknown (cf.
-\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
-csubj  &  79   &   \textbf{78}$^{***}$   &   60   &   \textbf{66}$^{***}$   &   \textbf{68}$^{**}$  & Unknown (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
-expl  &  12   &   36   &   62   &   40   &   53  & Unknown \\
-iobj  &  74   &   49   &   \textbf{78}$^{***}$   &   \textbf{38}$^{*}$   &   62  & Unknown \\
-vocative  &  46   &   \textbf{42}$^{*}$   &   59   &   45   &   \textbf{45}$^{*}$  & Unknown \\
-
-\hline
-% Uninterpretable
-compound  &  66   &   \textbf{61}$^{*}$   &   48   &   \textbf{58}$^{*}$   &   52  & Uninterpretable \\
-det  &  31   &   \textbf{71}$^{***}$   &   \textbf{56}$^{*}$   &   48   &   54  & Uninterpretable (conflates articles and others) \\%$.50\%$ for articles, $\approx 50\%$ for others, \citep{dryer1992greenbergian} \\ %TODO what else is conflated in det?
-why do we keep advmod but not det?
-dislocated  &  47   &   61   &   \textbf{68}$^{**}$   &   55   &   58  & Uninterpretable \\
-dep  &  56   &   \textbf{64}$^{*}$   &   52   &   58   &   \textbf{58}$^{*}$   & Uninterpretable \\
-
-\hline
-% UD artifacts
-conj  &  72   &   \textbf{82}$^{***}$   &   \textbf{42}$^{**}$   &   \textbf{62}$^{***}$   &   \textbf{59}$^{*}$  & UD artifact  \\
-discourse  &  24   &   \textbf{36}$^{**}$   &   \textbf{41}$^{*}$   &   49   &   46  & UD artifact   \\
-fixed  &  78   &   \textbf{32}$^{***}$   &   45   &   47   &   44  & UD artifact   \\
-flat  &  73   &   \textbf{66}$^{**}$   &   45   &   \textbf{59}$^{**}$   &   54  & UD artifact   \\
-goeswith  &  61   &   57   &   44   &   57   &   52  & UD artifact   \\
-list  &  94   &   59   &   50   &   54   &   52  & UD artifact  \\
-orphan  &  82   &   \textbf{70}$^{**}$   &   50   &   \textbf{55}$^{*}$   &   59  & UD artifact   \\
-parataxis  &  66   &   56   &   43   &   57   &   59  & UD artifact  \\
-reparandum  &  18   &   53   &   57   &   43   &   48  & UD artifact  \\
-
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
+&	\multicolumn{2}{c|}{Correlates with...}   &          \multirow{2}{*}{UD Relation}   & \multirow{2}{*}{\citet{greenberg1963universals}}    \\ 
+&	verb & object & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
+%adp. &
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&adposition    &    NP    &  $\xrightarrow{lifted\_case}$   & 3, 4   \\ \hline
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&copula  verb  &    predicate    &    $\xrightarrow{lifted\_cop}$   & --    \\\hline
+\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}&tense/aspect auxiliary    &    VP    &    \multirow{2}{*}{$\xleftarrow{aux}$}   & 16, 13  \\
+&	negative auxiliary    &    VP    &    & -- \\ \hline
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&noun    &    genitive    &   $\xrightarrow{nmod}$ & 2, 23   \\ \hline
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&noun    &    relative clause    &    $\xrightarrow{acl}$ &  24    \\ \hline
+\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}&complementizer    &    S    &   \multirow{2}{*}{$\xrightarrow{lifted\_mark}$}  & --    \\
+&	adverbial subordinator & S &  & -- \\ \hline
+\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}&	adjective & std. of comp. & \multirow{2}{*}{$\xrightarrow{obl}$} & --\\
+&verb    &    PP    &    & 22   \\\hline
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&`want'    &    VP    &    $\xrightarrow{xcomp}$   & 15   \\\hline
 \end{tabular}
 }
-
 	\end{center}
-\caption{Predictions on all UD relations occurring in more than one language.
-Numbers are the fraction of languages where the direction of the given relation agrees in direction with the \emph{obj} relation.
-In the last column, we indicate what direction would be expected typologically.
-``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
-``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
-\label{tab:all-predictions}
+	\caption{Greenbergian Correlations based on Dryer \cite{dryer1992greenbergian}, with operationalizations with Universal Dependencies using the modified format of \cite{futrell2015largescale} (see text).
+	For reference, we also provide the numbers of the closest corresponding universals stated in Greenberg's original study, to the extent that this is possible.
+	}\label{table:greenberg-dryer}
 \end{table*}
 
-\section{Results on Best Optimized Grammars}
+\paragraph{Excluded Correlations}
+We excluded three correlations that are not annotated reliably in UD, and are only relevant to some of the world's languages: Question particles, plural words (i.e., independent plural markers), and articles.
+All three types of elements occur at most in parts of the 51 UD languages, and none of them is annotated reliably in those languages where they occur.
+Among these three types of elements, the one most prominent in our sample of 51 languages is articles (occurring in many European languages); UD subsumes them under the \emph{det} relation, which is also used for other frequent elements, such as demonstratives and quantifiers.
 
-In the main part, we reported the percentage of optimized grammars supporting correlations.
-Here we report these numbers for the grammars that optimize each objective most successfully.
-That is, for each language and objective, we chose among the eight optimized grammars the one that achieves the best value for the objective function on the validation partition of the treebanks -- that is, the one where the optimization procedure was most successful.
+We also excluded the verb-manner adverb correlation.
+UD does not distinguish manner adverbs from other elements labeled as adverbs, such as sentence-level adverbs and negation markers, whose ordering is very different from manner adverbs.
+All types of adverbs are unified under the \emph{advmod} relation.
+In the real orderings in our sample of 51 UD languages, the dominant ordering of \emph{advmod} almost always matches that of subjects -- that is, \emph{advmod} dependents are ordered after the verb only in VSO languages.
+This ordering behavior is very different from that documented for manner adverbs by Dryer.
 
-Numbers for these languages are shown in Table~\ref{table:results-dryer-best-only}.
-%Confirming the results from the main paper, 
-As expected, the correlations mostly have stronger prevalences than when considering all optimized languages.
-For Dependency Length, the correlations that it predicts mostly have prevalences close to 100 \%, whereas both the UD languages and efficiency-optimized languages show somewhat lower percentages.
-That is, dependency length mostly overshoots for the correlations that it predicts correctly.
+We further excluded the verb-subject correlation, which is not satisfied by much more than half of the world's languages (51 \% among those with annotation in the \emph{World Atlas of Language Structures} \cite{wals-81}, with clear violation in 35.4 \%).
+It is satisfied only in 33\% of our sample of 51 UD languages, as quantified using the grammars we extracted.
+Dryer \cite{dryer1992greenbergian} counts this as a correlation since he describes the distribution of subject order as an interaction between a weak correlation with object order, and a very strong dominance principle favoring SV orderings.
+We focus on the modeling of correlations, and leave dominance principles to future research.
+We therefore excluded this correlation here.
+%We note that efficiency makes predictions compatible with the pattern observed in our sample of 51 languages, potentially in line with the noisy-channel explanation of SVO described by~\cite{gibson2013noisy} (see Table~\ref{tab:all-predictions}).
 
-\begin{table*}
-	\begin{center}
-\begin{tabular}{|ll|l|l|l|ll|l|}
-	\hline
-	%&
-	\multicolumn{2}{|c|}{Correlates with...}   &          \multirow{2}{*}{Operationalization}       &     \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\
-	verb & object &      & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%adp. &
-adposition    &    NP    &    lifted\_case    &    86    &    98    &    69    &    88    &    83   \\
-copula    &    NP    &    lifted\_cop    &    94    &    98    &    58    &    85    &    74   \\
-auxiliary    &    VP    &    aux    &    88    &    84    &    90    &    61    &    73   \\
-noun    &    genitive    &    nmod    &    80    &    98    &    75    &    88    &    91   \\
-noun    &    relative clause    &    acl    &    80    &    100    &    69    &    88    &    90   \\
-complementizer    &    S    &    lifted\_mark    &    76    &    100    &    76    &    82    &    87   \\
-verb    &    PP    &    obl    &    88    &    98    &    88    &    75    &    91   \\
-want    &    VP    &    xcomp    &    88    &    100    &    92    &    94    &    96   \\
-verb    &    subject    &    nsubj    &    33    &    25    &    53    &    10    &    21   \\
-verb    &    manner adverb    &    advmod    &    35    &    61    &    27    &    55    &    36   \\
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-\end{tabular}
-	\end{center}
-	\caption{Percentages of grammars supporting correlations, among grammars that best optimize each objective on treebanks of a given language.}\label{table:results-dryer-best-only}
-\end{table*}
+%croft2011greenbergian
+%dryer2011evidence
+
+\paragraph{Other Greenberg Universals}
+Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
+Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
+The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order; these form the basis of our formalization in Table~\ref{table:greenberg-dryer}.
+%, finding 16 correlations to likely be universal.
+%Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:greenberg-dryer}.
+%Testing efficiency on the excluded correlations would be feasible with additional annotation of data.
+%For instance, one of these could be tested by labeling the existing databases for the distinction between manner adverbs from other kinds of adverbs.
+%For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).
+There are a few other well-supported word order universals that are not correlations with verb-object order.
+This includes dominance principles~\cite{croft2003typology} such as the strong preference for subjects to precede objects.
+Furthermore, there has been interest in Greenberg's universals 18 and 20, which describe correlations not with verb-object order, but of different elements of noun phrases~\cite{cinque2005deriving, culbertson2014language, dryer2018order}.
+Future work should examine whether these universals can also be linked to efficiency optimization.
+
+\section{Formalizing Comunicative Efficiency}
+\subsection{Relation to Other Work}
+
+%The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
+%Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication.
+
+Here we discuss how our formalization of communicative efficiency relates to formalizations that have been proposed in the information-theoretic literature on language.
+Across the literature, the core idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms, with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global, kemp2012kinship,xu2014numeral, regier2015word,xu2016historical, zaslavsky2018efficient,futrell2017memory,peloquin2019interactions, zaslavsky2019semantic, frank2012predicting,goodman2013knowledge, kao2014nonliteral, bennett2018extremely}.
+
+The basis for our precise formalization is the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems.
+%The key idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms.
+%; there are some differences in the formalization of complexity.
+If $S$ denotes signals (e.g., words, sentences) and $R$ denotes their referents (e.g., objects in a reference game), then this efficiency metric takes the form (notation slightly varies across these publications):
+\begin{equation}\label{eq:eff-general}
+I[S, R] - \lambda H[S]
+\end{equation}
+where $I[S, R]$ describes the \key{informativity} of the signals $S$ about their referents $R$, and $H[S]$ describes the \key{complexity} of the communication system.
+While prior studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} mostly considered settings where the signals $S$ are individual words without further structure, the signals are entire sentences $\utterance$ in our setting.
+The underlying messages $R$ which the speaker aims to convey are the syntactic structures $\tree$.
+By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
+The syntactic structure specifies how the meanings of words are combined; therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.
+%The idea that efficient languages will permit the listener to infer grammati
+
+We therefore arrive at the following efficiency metric:
+\begin{equation}\label{eq:efficiency-derived}
+	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+\end{equation}
+where \key{parseability} is the amount of information that utterances provide about their underlying syntactic structures:
+\begin{equation}
+	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+\end{equation}
+and \key{predictability} is the negative entropy or surprisal of the language:
+\begin{equation}
+	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
+\end{equation}
+
+We now discuss how this metric relates to other variants that have been proposed in the information-theoretic literature on language, and describe why~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is particularly appropriate to our setting.
+This efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
+The Information Bottleneck models complexity using a mutual information term instead of the entropy appearing in~(\ref{eq:eff-general}).
+This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
+As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure.
+Furthermore, entropy also arises naturally as a complexity measure from the role played by surprisal in determining human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+
+A few other approaches share the mutual information term for informativity in (\ref{eq:eff-general}), while using complexity measures that are not explicitly information-theoretic.
+%A few models have used other complexity measures that are not explicitly information-theoretic quantities.
+In \cite{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \cite{kemp2012kinship} it is the complexity of defining the concepts that are encoded.
+The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
+Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the Information Bottleneck formalism \cite{zaslavsky2018efficient, zaslavsky2019semantic}, bringing them even closer to our formalization of efficiency.
+
+In addition to these models, which quantify the efficiency of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context, such as models of pragmatic reasoning in reference games \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
+In line with the other models, these assume that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely}.
+
+%\textcolor{red}{also mention \cite{piantadosi2012communicative,fedzechkina2012language}}
+
+%Other metrics very similar to~(\ref{eq:eff-general}) have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
+%The idea of formalizing informativity using the mutual information $I[S, R]$ is common to all these approaches.
+
+%These information-theoretic accounts vary somewhat in the formalizations of complexity.
+%Like our model, \cite{ferreri2003least,ferrericancho2007global,futrell2017memory} model the complexity as the \key{entropy} of the distribution of forms, which .
+%This model is closely related to the Information Bottleneck approach \cite{zaslavsky2018efficient, zaslavsky2019semantic}, which models complexity as the mutual information between forms (e.g., words) and their referents (e.g., individual objects).
+
+\subsection{Possible values of $\lambda$}\label{sec:lambda}
+
+In the efficiency objective
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
+\end{equation}
+the value of $\lambda$ is constrained to be in $[0,1)$.
+This means, surprisal must be weighted less strongly than parseability.
+
+Greater values of $\lambda$ can mathematically result in degenerate solutions.
+To show this, note that the following inequality always holds:
+\begin{equation}
+I[\utterance; \tree] \leq H[\utterance].
+\end{equation}
+Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
+\begin{equation}
+    R_{Eff} = I[\utterance; \tree] - \lambda H[\utterance] \leq 0.
+\end{equation}
+and it takes the maximally possible value zero if there is only a single utterance $\utterance$, in which case both $I[\utterance; \tree]$ and $H[\utterance]$ are zero.
+This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
+While the design of our word order grammars (see \emph{Materials and Methods}) precludes a collapse of all syntactic structures to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable description of the efficiency of communication systems.
+In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parseability, whereas values closer to $0$ diminish the role of predictability.
+
+In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parseability.
+While the computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$, we also examined word order predictions for grammars optimized for only parseability or only predictability, in order to tease apart predictions made by these two components.
+As shown in Table~\ref{tab:all-predictions-1}, each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions.
+%Indeed, the same holds for all of the 38 syntactic relations in UD  (Table~\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})
+%We additionally report results with $\lambda = 0$, i.e., placing all weight on parseability.
+That is, at $\lambda$ close to its maximal value, the predictions of optimizing the two scoring functions individually add up to the predictions of efficiency optimization.\footnote{Results from one of the preliminary experiments reported in Figure~\ref{table:corr-resu-previous} show that results are stable to small variation of $\lambda$: Essentially equivalent predictions are obtained for $\lambda=1.0$.
+While $\lambda=1.0$ is not a valid choice for communicative efficiency in general due to the possibility of collapse to a single utterance, our setting does not allow such a collapse, as the syntactic structure already determines which words are present in the sentence.}
+Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}), in which case not all correlations are predicted (Figure~\ref{fig:posterior}).
+%Finally, the robustness of our results to small variations of $\lambda$ is confime by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu-previous}.
+
+%In terms of information theory \cite{shannon1948mathematical}, this corresponds to the notion of the rate-distortion tradeoff.
+%In classical rate-distortion theory, the cost of a code is its rate, the mutual information between codes and underlying messages (as in the Information Bottleneck model).
+% classical models of distortion are the Hamming distortion and 
+%TODO also mention rate-distortion theory
+
+%- our function
+
+%similar: RSA, there it's utterance-level
+
+%- number of different forms 
+%
+%- complexity of the mapping 
+%
+%- MI between
+%
+%
+%
+%We model the first term as the 
+% degree to which listeners can reconstruct dependency parse structures from an utterance, i.e., the \key{parseability} if the language.
+This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
+%\begin{equation}
+%	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+%\end{equation}
+%where the sum runs over all possible pairs of sentences $l$ and syntactic structures $t$ in the language.
+%%This quantity describes the degree to which syntactic structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
+%
+%
+%
+%We formalize the complexity of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
+%In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
+%Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+%%and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
+%
+%In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
+%\begin{equation}
+%	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
+%\end{equation}
+%where the sum runs over all possible sentences $u$ that belong to the language.
+%%In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
+%
+%
+%
+%The efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
+%\begin{equation}\label{eq:efficiency}
+%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+%\end{equation}
+%with an interpolation weight $\lambda \in [0,1)$.
+%In all results from Study 2 reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
+%See SI appendix section 5 for mathematical discussion of $\lambda$, and robustness to other choices.
+%
+%
+
+%\section{Relation to the Information Bottleneck}
+%
+%Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
+%\begin{equation}
+%    \label{eq:efficiency-again}
+%    J_T(L) = I[L(T); T] - \lambda H[L(T)].
+%\end{equation}
+%This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}.
+In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information.
+The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$.
+Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
+%\begin{equation}
+%    \label{eq:ib}
+%    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
+%\end{equation}
+%In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
+%
+%In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$.
+To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$.
+We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized.
+Then our objective is parallel to Eq.~\ref{eq:ib}:
+%\begin{equation}
+%    \label{eq:efficiency-kinda-ib}
+%    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
+%\end{equation}
+%Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$.
+Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
+%\begin{equation*}
+%    J_T(L) = I[U;T] - \lambda H[U].
+%\end{equation*}
+%
+%Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
+%
+%
+
+\section{Supplementary Analyses for Study 1}
+
+In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.
+
+We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section~\ref{sec:lambda}), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
+
+Within each language, we estimate the Pareto curve based on the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
+
+Figure 4 in the main paper shows the average of these per-language plots.
 
@@
 \centering
-
-    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
-    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
-    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
-
-    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
-    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
-	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
-    \label{fig:posterior-prevalence}
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
+\caption[Predictability and Parseability]{Predictability and parseability of 51 languages.
+Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.}\label{fig:pareto-per-lang}
 \end{figure}
 
-\section{Further Statistical Analyses}
+\footnotetext{Note that, in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
+This can be attributed to two reasons: First, stochastic gradient descent introduces noise due to its stochasticity and will only approximately find an optimal solution; second, for some corpora, there may be some degree of distributional mismatch between the training partitions (on which grammars are optimized) and held-out partitions (on which efficiency is estimated).
+This in particular applies to very small corpora such as Irish (121 training sentences).}
 
-In the main results table, we reported the fraction of optimized languages satisfying a correlation.
-We show that essentially the same predicted fractions are obtained when controlling for the imbalance in distribution of language families.
+In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
+We also conducted the analysis using a Binomial test (one-sided), testing whether the real gramar is more efficient than the \emph{median} of random grammars, avoiding any distributional assumption on the random grammars.
+As before, we used Hochberg's step-up procedure (Note that the tests for different languages are independent, as different random grammars are evaluated for each language), with $\alpha=0.05$.
+In this analysis, real grammars improved in parseability for 80\% of languages, in predictability for 69\% of languages, and in either of both in 92\% of languages ($p<0.05$, with Bonferroni correction).
+In Table~\ref{fig:pareto-per-lang-stats}, we provide per-language results for the $t$-tests and binomial tests.
 
-To this end, we used the logistic mixed-effects regressions that were used to estimate significance to compute the expected prevalence of a correlation controlling for language families.
-To obtain predictions controlling for languages and language families, we plot the logit-transformed posterior of $\alpha$ in Figure \ref{fig:posterior-prevalence}.
+\begin{table}
+\centering
+\small{
+\begin{tabular}{l||ll|lll|lll}
+Language & Pred. (t) & Parse. (t) & \multicolumn{3}{|c|}{Pred. (Binomial)} & \multicolumn{3}{|c|}{Parseab. (Binomial)} \\ 
+&  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
+\input{../results/plane/per-language-statistics.tex}
+\end{tabular}
+}
+\caption{Per-language results in Study 1.
+For each language, we show the following: (1) $p$-values obtained from a one-sided $t$-test, for the null that the mean predictability/parseability of random grammars is at least as high as that of the real grammar.
+(2) Results from one-sided binomial tests, for the null that the the real grammar is better than at most $50 \%$ of random grammars.
+In addition to the $p$-value, we report point estimate and 95\% confidence interval for the fraction of random grammars that have values below real grammars.}\label{fig:pareto-per-lang-stats}
+\end{table}
+
+\section{Supplementary Analyses for Study 2}
+\subsection{Correlation between Universals and Efficiency}
+
+In Figure~\ref{fig:corr-eff-corr}, we plot efficiency, parseability, and predictability (all are $z$-scored within language, as in Study 1) as a function of the number of satisfied correlations, for the real grammars of the 51 languages.
+
+We found very similar results using Spearman's rank correlation (Efficiency: $\rho=0.59$, $p = 9.8 \cdot 10^{-6}$; Parseability: $\rho=0.55$, $p=4.7 \cdot 10^{-5}$; Predictability: $\rho=0.36$, $p=0.012$).
+
+\begin{figure}[ht]
+    \centering
+
+    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
+    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
+    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
+
+	\caption{Correlation between the number of satisfied correlations ($x$-axis) and efficiency, parseability, and predictability ($y$-axis), for the 51 real languages.}
+    \label{fig:corr-eff-corr}
+\end{figure}
+
+\subsection{Predictions for Individual Languages}
+
+We show predictions for the eight correlations on the level of individual languages in Figure~\ref{fig:per-lang}.
+We obtained these predictions for individual languages and each of the eight relations as follows.
+For each language and each of the objective functions (efficiency, predictability, parseability), we considered the optimized grammar that yielded the best value of this objective function among the eight optimized grammars (i.e., the grammar where the optimization procedure had been most successful).
+We interpreted this grammar as verb-object or object-verb depending on the order in the real grammar of the language.
 
 \begin{figure} 
-    \centering
-    \includegraphics[scale=.7]{../results/correlations/figures/posterior_joint_with_ud.pdf}
-	\caption{Posteriors of the prevalences of the correlations among languages optimized for Predictability (red), parseability (green), and efficiency (blue), controlling for languages and language families, with prevalence in the UD languages.}
-    \label{fig:posterior-prevalence}
+	\begin{center}
+	\includegraphics[width=0.7\textwidth]{../results/correlations/figures/pred-eff-pred-pars-families-2.pdf} % figures/pred-eff-families-2.pdf
+	\end{center}
+	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by optimizing for efficiency, predictability, parseability (right).
+Dark blue: Verb patterner \emph{precedes} object patterner (English, Arabic, ...).
+Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).
+White cells indicate that the relation is not annotated in the dataset for the given language.}\label{fig:per-lang}
+%	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
+Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
+Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...) .}\label{fig:per-lang}
+\end{figure}
+
+\subsection{Regression for Predicted Correlations}
+
+%https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
+
+\paragraph{Bayesian Regression}
+We modeled the probabilities $p_{L,j}$ that a grammar optimized for data from language $L$ satisfies the $j$-th correlation ($j=1,...,8$) using a multilevel logistic model \cite{gelman2013bayesian}, with random intercepts for the language for whose data the grammar had been optimized, and for its language family, annotated according to \url{http://universaldependencies.org/}.
+Formally,
+\begin{equation}\label{eq:mixed-effects}
+logit(p_{L,j}) = \alpha_j + u_{L,j} + v_{f_L,j}
+\end{equation}
+where $f_L$ is the language family of $L$.
+The intercepts $\alpha_j$ ($j=1,...8$) encode the population-level prevalence of the correlations when controlling for differences between datasets from different languages and language families; $u_{L,j}$, $v_{f_L,j}$ encode per-language and per-family deviations from the population-level intercept $\alpha_j$.
+%, and
+%$u_{L,1,...8} \sim \mathcal{N}(0, \Sigma_{lang})$,
+%$v_{f_L,1,...8} \sim \mathcal{N}(0, \Sigma_{fam})$.
+
+%As we constructed 4 verb-object and 4 object-verb grammars for each of the 51 languages, TODO...
+
+Following the recommendations of \cite{ghosh2018use, burkner2018advanced}, we used as a very weakly informative prior a Student's $t$ prior with $\nu=3$ degrees of freedom, mean 0, and scale $\sigma=10$ (i.e., the PDF $p$ is $\frac{1}{\sigma} p_3(x/\sigma)$, where $p_3$ is the PDF of the $t$-distribution with $\nu=3$).
+We used this prior for $\alpha_j, \sigma_{L,j}, \tau_{L,j}$.
+Note that a correlation that holds in 90\% of cases corresponds to an intercept $\alpha \approx 2.19$ in the logistic model, well within the main probability mass of the prior.
+
+We modeled full covariance matrices of per-language and per-family random intercepts over all eight correlations.
+We placed an LKJ prior ($\eta=1$) on these matrices, as described in \cite{burkner2018advanced}.
+We used MCMC sampling implemented in Stan \cite{carpenter2017stan, hoffman2014no} using the R package \texttt{brms} \cite{buerkner2017brms}.
+We ran four chains, with 5000 samples each, of which the first 2500 were discarded as warmup samples.
+We confirmed convergence using $\hat{R}$ and visual inspection of chains \cite{gelman2013bayesian}.
+
+We obtained the posterior density plots in Table 2 (Main Paper) and in Figure \ref{tab:all-predictions-1}) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\alpha_j$ (\ref{eq:mixed-effects}).
+
+\paragraph{Robustness}
+To ascertain the robustness of our results, we also conducted a frequentist analysis using \texttt{lme4} \cite{bates2015fitting}.
+For each of the correlations, we conducted a logistic mixed-effects analysis predicting whether a grammar satisfies the correlation, with random effects of language and language family.
+The results are shown in Table~\ref{tab:corr-regression} together with those of the Bayesian analysis.
+The frequentist analysis agrees with the Bayesian model; all eight correlations are predicted to hold in more than half of the optimized grammars ($p < 0.01$ each).
+
+Note that the Bayesian analysis  also estimates a posterior distribution of the number of satisfied correlations (see Figure~\ref{fig:posterior}), providing an elegant solution to the multiple-comparisons problem arising from analysing the eight correlation.
+
+% and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
+
+%We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
+
+\begin{table}
+\small{
+\begin{center}
+\begin{tabular}{|l||l|lll|llll|ll|llllll}
+\hline
+ & Prevalence & \multicolumn{3}{c|}{Bayesian} & \multicolumn{4}{c|}{Frequentist} \\ 
+& & Mean & SD & $p(\beta \leq 0)$ & $\beta$ & SE & $z$ & p \\
+\hline\hline
+	\input{../results/correlations/output/efficiency-results-full.tex}
+\hline
+\end{tabular}
+\end{center}
+}
+	\caption{Detailed results for Bayesian and Frequentist mixed-effects analyses for the eight correlations. 
+We show (1) the raw prevalence of each correlation in the optimized grammars (8 grammars for each of the 51 languages),
+(2) for the Bayesian analysis, we provide posterior mean and SD of $\beta$, and the posterior probability that $\beta$ has the opposite sign,
+(3) for the Frequentist analysis, we provide the point estimate, SE, $z$, and $p$-values (2-sided).
+The frequentist analysis confirms the results of the Bayesian analysis.
+}\label{tab:corr-regression}
+\end{table}
+
+\subsection{Comparing Efficiency to its Components}\label{sec:posterior-number}
+
+In Figure~\ref{fig:posterior}, we plot the posterior distribution of the number of correlations predicted to hold in most optimized grammars, as obtained from the Bayesian regression.
+For each posterior sample, we say that the $j$-th correlation holds if the value of $\alpha_j$ in that posterior sample is positive.
+In the figure, we plot the fraction of posterior samples in which a given number of correlations is satisfied.
+In addition to grammars optimized for efficiency, we also report the result for grammars optimized for predictability and for parseability alone.
+Efficiency predicts all eight correlations with high posterior probability; predictability and parseability alone do not.
+
+\begin{figure}[ht]
+	\begin{center}
+	\includegraphics[width=0.98\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
+	\end{center}
+	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in the Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
 \end{figure}
@@
 %    \centering
-%    \includegraphics[scale=.3]{figures/posterior-satisfied-universals-efficiency.pdf}
-%     \includegraphics[scale=.3]{figures/posterior-satisfied-universals-langmod.pdf}
-%     \includegraphics[scale=.3]{figures/posterior-satisfied-universals-parseability.pdf}
-%    \includegraphics[scale=.3]{figures/posterior-satisfied-universals-depl.pdf}
-%	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks.}
-%    \label{fig:posterior-satisfied-count}
+%    
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
+%    
+%    
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
+%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
+%	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
+%    \label{fig:posterior-prevalence}
 %\end{figure}
 
-%We argued that efficiency predicts all 10 correlation universals in the direction manifested in the UD corpora, as for each universal, the posterior probability that less than $50 \%$ of the optimized grammars satisfy the correlation was less than $0.05$.
-%To address potential multiple comparisons problems, we use a single logistic model to show that efficiency predicts all ten correlations with high posterior probability.
+\subsection{Results on all UD Relations}
+In this section, we provide the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
+We also report results for grammars optimized for predictability and parseability individually.
 
-%For each objective function, we constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency, with random effects for language and language family, correlated across the ten dependencies.
+We considered all UD syntactic relations occurring in at least two of the 51 languages.
+In Table~\ref{tab:all-predictions-1}, we present the data for the eight correlations discussed in the main paper, and for those other relations for which the typological literature provides data.\footnote{The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
+Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
+For simplicity, we display this as a corelation in this table.}
+Additionally, in Table~\ref{tab:all-predictions-2} we present data for the other UD relations, for which either no typological data is available, or which are not linguistically meaningful.
 
-%We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
-%The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
-%The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$.
-The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.
+Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM} for discussion).
+This has been suggested as a property of efficient word order in prior work, and, indeed, we found that DLM is predicted by efficiency optimization (Section~\ref{sec:DLM}).
+Results in Table~\ref{tab:all-predictions-1} also show that optimizing for DLM makes predictions similar to efficiency optimization. See Section~\ref{sec:DLM} for further discussion.
 
-\section{Relation to the Information Bottleneck}
+\begin{table*} % TODO figure out away to make this table more aesthetic
+	\begin{center}	
+\includegraphics[width=0.8\textwidth]{si-table-perrel-1-1.png}
+\end{center}
+\caption{Predictions on UD relations with predictions from the typological literature.
+The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
+In the last column, we indicate what direction would be expected typologically.}
+\label{tab:all-predictions-1}
+\end{table*}
 
-Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
+%
+%%posterior_perRelation_Real_conj.pdf         posterior_perRelation_Real_goeswith.pdf     posterior_perRelation_Real_nummod.pdf       
+%%user@user-G751JY:~/grammar-optim/results/correlations$ evince figures/posteriors/posterior_perRelation_Real_csubj.pdf 
+%%user@user-G751JY:~/grammar-optim/results/correlations$ 
+%
+%\begin{table*} % TODO figure out away to make this table more aesthetic
+%	\begin{center}	
+%\small{
+%\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
+%	\hline
+%&	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+%% From Dryer	
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
+%%\hdashline
+%
+%\hline
+%% Other universals, not in Dryer
+%&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
+%&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf.
+\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
+%&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf.
+\cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
+%&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section~\ref{sec:correlations} \\
+%&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
+%&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%\end{tabular}
+%}
+%
+%\end{center}
+%\caption{Predictions on UD relations with predictions from the typological literature.
+The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
+In the last column, we indicate what direction would be expected typologically.}
+%\label{tab:all-predictions-1}
+%\end{table*}
+%
+
+%\footnotetext{Note that we report an \emph{anti-correlation}.
+The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
+Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
+}
+
+\begin{table*} % TODO figure out away to make this table more aesthetic
+	\begin{center}	
+\includegraphics[width=0.8\textwidth]{si-table-perrel-2-1.png}  
+\end{center}
+\caption{Predictions on UD relations for which no predictions are available in the typological literature.
+``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
+``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
+\label{tab:all-predictions-2}
+\end{table*}
+
+%
+%\begin{table*} % TODO figure out away to make this table more aesthetic
+%	\begin{center}	
+%\small{
+%\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
+%	\hline
+%	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
+%\hline
+%appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\
+%lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\
+%expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\
+%iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\
+%vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\
+%
+%\hline
+%% Uninterpretable
+%compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\
+%det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\
+%dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\
+%dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
+%advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\
+%
+%\hline
+%% UD artifacts
+%conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
+%discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\
+%fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\
+%flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
+%goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
+%list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
+%orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\
+%parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\
+%reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\
+%
+% \hline
+%%comp. &
+%%\multirow{3}{*}{NP}&
+%%	AP &
+%\end{tabular}
+%}
+%
+%\end{center}
+%\caption{Predictions on UD relations for which no predictions are available in the typological literature.
+``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
+``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
+%\label{tab:all-predictions-2}
+%\end{table*}
+%
+%
+
+\subsection{Previous Experiments}\label{sec:previous-exps}
+In Table~\ref{table:corr-resu-previous} we report the results of our two previous, preregistered, simulations\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}, \url{https://aspredicted.org/blind.php?x=bg35x7}.
+For the results of the locality simulations described in the first preregistration, see the DLM results in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM} for discussion.} together with results from the main experiment.
+These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) one simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
+Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
+
+\begin{table}[hbt!]
+
+	\begin{center}
+\includegraphics[width=0.4\textwidth]{si-table-perrel-3-1.png}  
+\end{center}
+
+	\begin{center}
+\begin{tabular}{cccc}
+$\lambda=0.0$ & $\lambda=0.9$ & $\lambda=0.9$ & $\lambda=1.0$ \\
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency09.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-efficiency-large.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf}
+\end{tabular}
+	\end{center}
+\caption{Results from optimization experiments for different values of $\lambda$, including our two previous preregistered experiments.
+For comparison, we also show results for $\lambda=0$, corresponding to optimizing for parseability only (same results as reported in Tables (\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})).
+For $\lambda=0.9$, we report results from one preliminary preregistered experiment (left) and the final experiment (right).
+For $\lambda=1.0$, we report the other preliminary preregistered experiment.
+Giving similar weight to parseability and predictability -- that is, $\lambda$ close to $1$ -- results in more accurate word order predictions than choosing a small value of $\lambda$ such as $\lambda=0.0$.
+Note that $\lambda$ cannot take values smaller than zero, or greater than one, see Section \ref{sec:lambda}.
+}\label{table:corr-resu-previous}
+\end{table}
+
+%\section{Joint optimization of grammar, parser, and language model parameters}
+\section{Creating Optimized Grammars}
+
+In this section, we describe the method we employ for creating grammars that are optimized for efficiency, and how we extract grammars describing the actual ordering rules of languages.
+%Our method will be applicable more generally to the task of optimizing grammars for a broad class of objective functions.
+We carry out grammar optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
+More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
+This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.
+
+This method addresses a major challenge noted in previous work optimizing grammars, namely that the predictability (and parseability) of an individual sentence depends on the entire distribution of the language.
+Previously, \citet{gildea2015human} optimized grammars for dependency length and trigram surprisal using a simple hill-climbing method on the grammar parameters, which required reestimating the trigram surprisal model in every iteration.
+Such a method would be computationally prohibitive for efficiency optimization, as it would require reestimating the neural network models after every change to the grammar, which would amount to reestimating them hundreds or thousands of times per grammar.
+Our method, by allowing for the use of stochastic gradient descent, addresses this challenge, as we describe in Section~\ref{sec:optim-eff}.
+
+\subsection{Differentiable Ordering Grammars}\label{sec:diff-gramm}
+
+We extended the parameter space of grammars by continuously interpolating between grammars, making efficiency a \emph{differentiable} function of grammar parameters.
+The parameters of such a \key{differentiable word order grammar} are as follows. 
+For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
+Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
+Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
+184) without replacement.
+
+If $a_\tau \in \{0, 1\}$, and the distances between values of $b_\tau$ (for different $\tau$) become very large, such a differentiable grammar becomes deterministic, assigning almost full probability to exactly one ordering for each syntactic structure.
+In this case, the grammar can be converted into an equivalent grammar of the form described in Materials and Methods, by extracting a single parameter in $[-1, 1]$ for each relation $\tau$.
+
+We provide an example in Figure~\ref{fig:grammar-sample}, illustrating grammar parameters for the relations in Figure 3 of the main paper.
+
+\begin{figure}
+	\begin{center}
+\begin{tabular}{l||l|ll||l|lllll}
+	& \multicolumn{3}{c||}{English}    &  \multicolumn{3}{c}{Japanese}  \\ 
+	Relation                     &  Par.   & $a_\tau$ & $b_\tau$            & Par.   & $a_\tau$ & $b_\tau$     \\ \hline \hline
+	object (\textit{obj})        &  $0.1$   &   $0.04$   & $-1.46$            & $-0.1$  & $0.99$ & $-0.7$2 \\
+	oblique (\textit{obl})       &  $0.3$   &   $0.13$     & $1.25$             & $-0.3$  & $0.99$ & $0.73$ \\
+	case (\textit{lifted\_case}) &  $0.2$   &   $0.07$       &   $-0.89$         & $-0.2$  &  $0.92$ & $0.02$  \\
+\end{tabular}
+	\end{center}
+	\caption{Sample Coefficients from grammars extracted from the real English and Japanese orderings (Section~\ref{sec:extract-grammars}), for the relations occurring in Figure 3 (Main Paper).
+We show parameters in $[-1,1]$ for deterministic word order grammars as described in \emph{Materials and Methods}, and the coefficients ($a_\tau, b_\tau$) for corresponding differentiable ordering grammars.
+For the deterministic grammars (`Par.'), positive coefficients indicate that the dependent will be placed after the head.
+For the differentiable grammars, $a_\tau > 0.5$ indicates predominance of ordering of dependents before heads, and larger $b_\tau$ indicates greater distance between head and dependent.}\label{fig:grammar-sample}
+\end{figure}
+
+\subsection{Extracting Grammars from Datasets}\label{sec:extract-grammars}
+We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
+To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
+We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.
+
+We validated the extracted grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
+Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS.
+In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb-object relation.
+WALS has data for 74\% of the entries\footnote{Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
+In the table, we compare those with the grammar we extracted for Croatian, noting that it fully agrees with the Serbian grammar.}, and lists a dominant order for 91\% of these.
+The grammars we extracted from the corpora agree with WALS in 96~\%  of these cases.
+
+Note that the grammatical formalism simplifies some aspects of the word order regularities of natural languages.
+For instance, it does not represent cases where ordering varies between main and embedded clauses, as it does not condition ordering decisions on the larger context.
+%Our method extracts the most frequent ordering and applies it throughout all 
+More complex and powerful ordering grammar models have been proposed \cite{futrell2015experiments, wang2016galactic}; however, they have similar limitations, and for our purposes, the model adopted here has the advantage of being simple and interpretable.
+%There currently is no grammatical formalism that is at the same time capable of representing the full range of word order regularities that are found, 
+%\cite{futrell2015experiments}
+
+\begin{table}
+\small{
+\begin{center}
+\begin{tabular}{l||ll|ll|ll|ll|ll|ll|llllll}
+%	   for(dep in c("obj", "lifted_mark", "acl", "nmod", "obl")) {
+		   Language 
+		   &	\multicolumn{2}{|c|}{Objects} 
+		   &	\multicolumn{2}{|c|}{Adpositions} 
+		   &	\multicolumn{2}{|c|}{Compl.} 
+		   &	\multicolumn{2}{|c|}{Rel.Cl.} 
+		   &	\multicolumn{2}{|c|}{Genitive} 
+		   &	\multicolumn{2}{|c|}{PP}  \\ \hline\hline
+	\input{../results/grammars/comparison-table.tex}
+\end{tabular}
+\end{center}
+}
+\caption{Comparing grammars extracted from databases to linguistic judgments in the World Atlas of Linguistic Structures.
+For each of the six syntactic relation, the first column provides the ordered coded in the extracted grammar; the second column provides the order coded in WALS (DH for dependent-head, HD for head-dependent order).
+`?' indicates that WALS has no data.
+$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur at similar rates in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
+Finally, `--' indicates that the relation does not occur in the corpus.
+}\label{tab:grammars-wals}
+\end{table}
+
+\subsection{Optimizing Grammars for Efficiency}\label{sec:optim-eff}
+
+In this section, we describe how we optimized grammar parameters for efficiency.
+A word order grammar can be viewed as a function $\mathcal{L}_\theta$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output an ordered sequence of words $u = \mathcal{L}_\theta(t)$ linearizing the tree.
+More generally, if $\mathcal{L}_\theta$ is a differentiable ordering grammar (Section~\ref{sec:diff-gramm}), then $\mathcal{L}_\theta(t)$ defines a \emph{probability distribution} $p_{\mathcal{L}_\theta}(u|t)$ over ordered sequences of words $u$.
+In the limit where $\mathcal{L}_\theta$ becomes deterministic, the distribution $p_{\mathcal{L}_\theta}(u|t)$ concentrates on a single ordering $u$.
+
+Recall the definition of efficiency
+\begin{equation}\label{eq:efficiency-recall}
+	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+\end{equation}
+where
 \begin{equation}
-    \label{eq:efficiency-again}
-    J_T(L) = I[L(T); T] - \lambda H[L(T)].
+	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
 \end{equation}
-This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}.
-In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information.
-The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$.
-Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
 \begin{equation}
-    \label{eq:ib}
-    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
+	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
 \end{equation}
-In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
+where $t \sim \tree$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
 
-In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$.
-To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$.
-We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized.
-Then our objective is parallel to Eq.~\ref{eq:ib}:
+These quatities are estimated using two neural models, as described in Section~\ref{sec:neural-architectures}:
+A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.
+The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is\footnote{Note that, in the definition of $R_{Pars}$, the term $p(t)$ is a constant independent of $\phi$ and the word order grammar $\mathcal{L}_\theta$; it can therefore be ignored in the optimization process.}
 \begin{equation}
-    \label{eq:efficiency-kinda-ib}
-    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
+	R_{Pars}^{\phi}(u,t) =  \log \frac{p_\phi(t|u)}{p(t)}
 \end{equation}
-Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$.
-Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
-\begin{equation*}
-    J_T(L) = I[U;T] - \lambda H[U].
-\end{equation*}
+%and we can write
+%\begin{equation}
+%	R_{Pars}^\phi := I[\utterance,\tree] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
+%\end{equation}
+%
+A \key{language model}, with some parameters $\psi$, calculates the word-by-word surprisal of an utterance:
+\begin{equation}
+	R_{Pred}^{\psi}(u) = - \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})
+\end{equation}
+Using this, we can rewrite Efficiency~(\ref{eq:efficiency-recall}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
+\begin{equation}
+	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi} := \max_{\phi,\psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\end{equation}\label{eq:efficiency-rewrite}
+%
+%The parsers and language models appropriate to a tree distribution $\tree$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
+%\begin{equation}
+%\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation}
+%\begin{equation}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation}
+%
+%Using these models, we can rewrite 
+%\begin{equation}
+%R_{Pars}^{\psi(\theta)}= \E_t \E_{u \sim p_\theta(u|t)} R_{Pars}^{\psi(\theta)}(u,t) 
+%\end{equation}
+%where $R_{Pars}(u,t) = \log \frac{p_\phi(t|u)}{p(t)}$, and $p_\theta(u|t)$ is the probability assigned to sentence $u$ by the distribution $\mathcal{L}_\theta(t)$ defined by the differentiable word order grammar $\mathcal{L}_\theta$,
+%and
+%\begin{equation}
+%R_{Pred} = \E_t \E_{u \sim p_\theta(u|t)} R_{Pred}(u)
+%\end{equation}
+%where $R_{Pred}(u) = \log p_\psi(u)$.
+%Therefore, we can rewrite efficiency as
+%\begin{equation}\label{eq:efficiency}
+%R_{\textit{Eff}}^\theta := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi(\theta)}(u,t) + \lambda R_{Pred}^{\psi(\theta)}(u)\right]
+%\end{equation}
+%Directly applying gradient descent to this expression is hard due to the complicated dependence of $\phi(\theta), \psi(\theta)$ on $\theta$.
+%
+%However, we can rewrite
+%\begin{equation}\label{eq:efficiency}
+%\argmax_\theta\	R_{\textit{Eff}}^\theta 
+%\end{equation}
+%equivalently into
+In order to find an optimal grammar $\theta$, we thus need to solve
+\begin{equation}\label{eq:efficiency}
+\argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} %= \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\end{equation}
+Importantly, $R_{\textit{Eff}}^{\theta, \phi, \psi}$ is differentiable in $\theta, \phi, \psi$: %, and we can apply stochastic gradient descent to carry out this optimization.
+\begin{align}
+\partial_\theta R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)} \left[  \left[\partial_\theta \log p_\theta(u|t)\right] \cdot    \left(R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right) \right] \label{eq:dtheta}\\
+\partial_\phi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\partial_\phi R_{Pars}^{\phi}(u,t)\right] \\
+\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}
+\end{align}
+where \ref{eq:dtheta} is derived using the \emph{score-function} or \emph{REINFORCE} trick~\cite{williams1992simple}.
+Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures.
 
-Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
+We can therefore apply stochastic gradient descent to jointly optimize $\theta, \phi, \psi$:
+In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim p_{\theta}(\cdot|t)$.
+Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
+do a gradient descent step using the estimator given by the expressions in the square brackets in \ref{eq:dtheta}-\ref{eq:dpsi}.
 
-\section{Possible values of $\lambda$}
+Optimizing for only parseability (or predictability) is very similar -- in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.
 
-In the efficiency objective~(\ref{eq:efficiency-again}), the value of $\lambda$ is constrained to be in $[0,1)$.
-This means, surprisal must be weighted less strongly than parsability.
+At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
+The neural parser and language model are also randomly initialized at the beginning of optimization.
+Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.
 
-Greater values of $\lambda$ mathematically result in degenerate solutions.
-To show this, note that the following inequality always holds:
-\begin{equation}
-I[L(T); T] \leq H[L(T)].
-\end{equation}
-Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
-\begin{equation}
-    J_T(L) = I[L(T); T] - \lambda H[L(T)] \leq 0.
-\end{equation}
-and it takes the maximally possible value zero if $L(T)$ maps all trees to a single utterance, in which case both $I[L(T); T]$ and $H[L(T)]$ are zero.
-This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
-While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
+See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates used in this optimization scheme.
 
-In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
-We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order     correlations and relative optimality of language.
+%
+%%(analogous for $R_{Pars}$ instead of $R_{Pred}$)
+%\begin{equation}\label{eq:estimator}
+%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
+%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
+%\end{equation}
+%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
+%for the gradient
+%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
+%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
+%
 
-\section{Joint optimization of grammar, parser, and language model parameters}
-
-For each of our objective functions, we seek parameters $\theta$ of a word order grammar $L_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $L_\theta$.
+%
+%
+%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
+Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
+%\begin{equation}
+%\label{eq:exp-over-utts}
+%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
+%\end{equation}
+%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
+In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
+%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
+%
+%
+%
+%We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
+%\begin{equation*}
+%R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
+%\end{equation*}
+%where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
+%\begin{equation*}
+%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
+%\end{equation*}
+%
+%
+%We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
+Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
+%\begin{align}
+%\nonumber
+%R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
+%\nonumber
+%&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
+%\end{align}
+%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
+%\begin{equation*}
+%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
+%\end{equation*}
+%
+%For efficiency, we linearly combine the two terms:
+%\begin{equation}
+%\label{eq:combined-score}
+%R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
+%\end{equation}
+%
+%
+%%\begin{equation}
+%%    \label{eq:depl}
+%%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
+%%\end{equation}
+%%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
+%
+%For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
 The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
 The optimization problems come out to:
-\begin{align}
-\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
-\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
-\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
-\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
-\end{align}
-Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
+%\begin{align}
+%%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
+%\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
+%\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
+%\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
+%\end{align}
+%Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
+%
+%
+%%The case of predictability, parseability, and efficiency are more challenging, as 
+%%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
+%We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
+%Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
+%
+%We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section).
+%
+%In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
+%Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
+%do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
+%\begin{equation}\label{eq:estimator}
+%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
+%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
+%\end{equation}
+%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
+%for the gradient
+%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
+%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
+%
+%See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
+%
 
-The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
+%The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
 A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
-We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
+%We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
 
-The case of predictability, parseability, and efficiency are more challenging, as we need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
-We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
-Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
-In each step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
-Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
-do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
-\begin{equation}\label{eq:estimator}
-{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
-%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
-\end{equation}
-%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
-for the gradient
-$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
-This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
-
-In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
+%In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
 %and $\ell_2$ regularization.
 
-A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
-Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
-We thus do not evaluate this objective function here.
+%A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
+%Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
+%We thus do not evaluate this objective function here.
 
-\section{Implementation Details}
+%We extract grammars for the actual languages using a Bayesian data analysis method.
+%We specify a simple Bayesian prior over differentiable ordering grammars:
+%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
+%for each $\tau$.
+%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
+%We extract
 
-\paragraph{Locality}
-In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
+\section{Neural Network Architectures}\label{sec:neural-architectures}
 
-\paragraph{Language Model}
-We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
-Given the small size of the corpora, this limit is only attained only for few languages.
+In this section, we describe the details of the neural network architectures.
+Choices follow standard practice in machine learning.
+All choices, except where explicitly noted otherwise, were made before evaluating word order properties, and the efficiency of real grammars.
+
+%\paragraph{Locality}
+%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
+
+\paragraph{Estimating Predictability}
+We choose a standard LSTM language model \citep{goldberg2017neural, hochreiter1997long}.
+For computational reasons, we restrict the vocabulary to the most frequent 50,000 words in the treebanks for a given language.
+Given the moderate size of the corpora, this limit is only attained only for few languages.
 In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
@@
 Using POS tags is intended to prevent overfitting on small corpora.
-These decisions were made before evaluating word order properties.
 
-\paragraph{Parser}
+\paragraph{Estimating Parseability}
 We use a biaffine attention parser architecture \citep{kiperwasser2016simple,zhang2017dependency,dozat2017stanford}.
@@
 To reduce overfitting on small corpora, we choose a delexicalized setup, parsing only from POS tags.
-Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering model had converged.
-Parsing from POS tags reduces early overfitting.
+Preliminary experiments showed that a parser incorporating word forms overfitted long before the ordering grammar had converged; parsing from POS tags prevents early overfitting.
 This decision was made before evaluating word order properties.
@@
 \paragraph{Hyperparameters}
-For locality, predictability, and parseability, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
-Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function on the language.
+Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
+For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
+These parameters are shown in Table~\ref{tab:hyperparameters}.
+Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (efficiency, predictability, parseability) on the language.
 We then used this setting for creating optimized word order grammars.
-% TODO more detail and maybe an example here
-% TODO all design changes were made before examining word order universal results
-
-% TODO Discuss iconicity somewhere somehow
 
 All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
-We do not use pretrained embeddings \citep{peters2018deep}: these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
+We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
 
-%\begin{table}[]
-%    \centering
-%    \begin{tabular}{c|cccc}
-%    Entropy penalty     &  0.0001, 0.001 \\
-%    Learning Rate (Word order grammar)     & 5e-6, 1e-5, 2e-5, 5e-5 \\
-%    Momentum (Word order grammar) & 0.8, 0,9 \\
-%    Dropout (Language model) & 0.0, 0.3, 0.5 \\
-%    
-%    \end{tabular}
-%    \caption{Hyperparameters}
-%    \label{tab:my_label}
-%\end{table}
+\begin{table}[]
+    \centering
+    \begin{tabular}{|l|l|l|}
+\hline
+\multirow{2}{*}{Optimization}&    Learning Rate     & 5e-6, 1e-5, 2e-5, 5e-5 \\
+&    Momentum & 0.8, 0.9 \\ \hline
+\multirow{6}{*}{Language Model} &    Learning Rate  & 0.5, 0.1, 0.2 \\
+&    Dropout Rate& 0.0, 0.3, 0.5 \\
+& Embedding Size (Words) & 50 \\
+& Embedding Size (POS) & 20 \\
+&    LSTM Layers  & 2 \\
+&    LSTM Dimensions  & 128 \\
+ \hline
+\multirow{5}{*}{Parser}&    Learning Rate  & 0.001 \\
+&    Dropout Rate  & 0.2 \\
+&    Embedding Size  & 100 \\
+&    LSTM Layers  & 2 \\
+&    LSTM Dimensions  & 200 \\
+\hline
+    \end{tabular}
+    \caption{Hyperparameters}
+    \label{tab:hyperparameters}
+\end{table}
 
-\paragraph{Optimization details}
-
-We employ two common variance reduction methods to improve the estimator~(\ref{eq:estimator}), while keeping it unbiased.
-For dependency length and predictability, note that the loss incurred for a specific word only depends on ordering decisions made up to that word (and its head, in the case of dependency length).
-We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:estimator}) with lower variance.
-Second, we use a word-dependent moving average of recent per-word losses as control variate \cite{williams1992simple}.
+\paragraph{Improved Unbiased Gradient Estimator}
+We employ two common variance reduction methods to improve the estimator~(\ref{eq:dtheta}), while keeping it unbiased.
+%For dependency length and 
+% (and its head, in the case of dependency length)
+For predictability, note that the surprisal of a specific word only depends on the preceding words (not on the following words), and thus only depends on ordering decisions made up to that word.
+We represent the process of linearizing a tree as a dynamic stochastic computation graph, and use these independence properties to apply the method described in \citet{schulman2015gradient} to obtain a version of~(\ref{eq:dtheta}) with lower variance.
+Second, we use a word-dependent moving average of recent per-word losses (the word's surprisal in the case of predictability, and the negative log-probability of the correct head and relation label in the case of parseability) as control variate \cite{williams1992simple}.
 These two methods reduce the variance of the estimator and thereby increase the speed of optimization and reduce training time, without biasing the results.
@@
 Furthermore, to encourage exploration of the parameter space, we add an entropy regularization term \citep{xu2015show} for each Direction Parameter $a_\tau$, which penalizes $a_\tau$ values near $0$ or $1$.
-The weight of the entropy regularization was chosen together with the other hyperparameters.
-%, adding the negative entropy of the $Bernoulli(a_\tau)$ to the loss $R$. %, weighted with $\alpha$.
+The weight of the entropy regularization was chosen together with the other hyperparameters.\footnote{Explored values: 0.0001, 0.001.}
 
-These techniques for improving (\ref{eq:estimator}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
+These techniques for improving (\ref{eq:dtheta}) are well-known in the machine learning literature, and we fixed these before evaluating optimized grammars for word order properties.
 
-We update word order grammar parameters $\theta$ using SGD with momentum.
-For the language model parameters $\phi$, we use plain SGD. 
+\paragraph{Learning Rate and Stopping Criterion}
+We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.
+For the language model parameters $\phi$, we use plain Stochastic Gradient Descent without momentum, as recommended by \citet{merity2018regularizing}. 
 For the parser parameters $\psi$, we use Adam \citep{kingma2014adam}, following \citet{dozat2017stanford}.
@@
 
-For each language and objective function, we apply Early Stopping~\citep{prechelt1998early} as follows.
-We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
-This technique, well known in the machine learning literature, helps prevent overfitting to the specific corpora used.
+All corpora have a predefined split in training and held-out (development) sets.
+We use the training set for optimizing parameters, and apply Early Stopping~\citep{prechelt1998early} using the held-out set.
 
-The choice of optimization methods and the stopping criterion were fixed before we investigated word order correlations.
+For \key{estimating the parseability or predictability} of a given grammar, we optimize the neural model on data ordered according to this grammar, and report the parseability/predictability on the held-out set to avoid overfitting to the training set.
+For Early Stopping, we evaluate on the held-out set at the end of every epoch. %run through the training set.
 
-Recall that, for each language, we created 8 optimized languages for each optimization criterion.
+For \key{optimizing grammars}, we jointly apply gradient descent to the grammar parameters and the neural models, using the gradient estimator (\ref{eq:dtheta}-\ref{eq:dpsi}).
+For Early Stopping, we evaluate on the held-out set in intervals of 50,000 sentences, using a Monte-Carlo estimate of $R_{\textit{Eff}}^{\theta, \phi, \psi}$ (\ref{eq:efficiency-rewrite}), sampling a single linearized sentence for each syntactic structure in the held-out set.
+When reporting the parseability/predictability of an optimized grammar, we evaluate these values for its fully deterministic version (Section~\ref{sec:diff-gramm}) to allow fair comparison with baseline grammars.
+
+%We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
+%This is a standard technique for optimizing neural models, and helps prevent overfitting to the specific corpora used.
+
+The choice of optimization methods and the stopping criterion were fixed before we investigated language efficiency or word order correlations.
+
+\paragraph{Optimized Grammars}
+As described in the main paper, for each language, we created 8 optimized languages for each optimization criterion.
 We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
-This maximizes statistical power for detecting and quantifying correlations between the \textit{obj} relation and other relations.
+This maximizes statistical precision in detecting and quantifying correlations between the verb-object relation and other relations.
+
+For efficiency optimization, for each grammar, we ran efficiency optimization with three different random seeds, selecting among these the seed that yielded the best overall efficiency value.
+We did this in order to control for possible variation across random seeds for the stochastic gradient descent optimization method.
+As described in our preregistration \url{http://aspredicted.org/blind.php?x=ya4qf8}, this choice was made after conducting a preliminary version of Study 2 reported in Section~\ref{sec:previous-exps}; results reported there show qualitatively identical results regarding the prediction of the eight word order correlations by efficiency optimization.
 
@@
 
-Here we address this question by demonstrating that our languages optimized for prediction and parseability are also optimal under a range of different language models and parsers.
-These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on a particular prediction and parsing algorithms.
+Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers.
+These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.
 
-\subsection{CYK Parsers}
+\subsection{CKY Parsers}
 
-We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of Collins 1997 (Model 1).
+We constructed simple Probabilistic Context-Free Grammars (PCFGs) from corpora and word order grammars, using a simplified version of the models of \cite{collins2003head} (Model 1).
 In our PCFGs, each head independently generates a set of left and right dependents.
@@
 We estimated probabilities by taking counts in the training partition, and performing Laplace smoothing with a pseudocount $\alpha=1$ for each possible rule of this form.
-For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CYK algorithm.
+For such a PCFG, exact parsing is possible using Dynamic Programming, and specifically the CKY algorithm \cite{kasami1966efficient}.
 
@@
 
-We used the CKY algorithm to compute the parsing loss $H[T|L_\theta(T)]$ on the validation partition of the English UD corpora, for random and optimized ordering grammars.
+We used the CKY algorithm to compute the syntactic ambiguity $H[\tree|\utterance]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
 Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
@@
     \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
-	\caption{Parsability loss $H[T|L_\theta(T)]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).}
+	\caption{Parsing loss $H[\tree|\utterance]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).
+We report $H[\tree|\utterance]$ normalized by sentence length.}
     \label{fig:cky-parser}
@@
 
+The result is that distance between words in the input is not indicative at all of the presence of absence of syntactic relations between them.
+
 \paragraph{Experiments}
 Using English and Japanese data, we trained parsers for the ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
-Resulting parsing accuracy scores are shown in Figure~\ref{fig:distorted-parser}.
-In all settings, the language optimized for efficiency achieved higher parsing accuracy than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
+Resulting parsing scores are shown in Figure~\ref{fig:distorted-parser}.
+In all settings, the language optimized for efficiency achieved lower parsing loss (i.e., higher parseability) than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
 %Note that absolute values of parsing accuracy vary between the orderings, being higher for the actual orderings than for distorted orderings, in which trees become nonprojective.
@@
     \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
-	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[T|L_\theta(T)]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').}
+	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[\tree|\utterance]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').
+We report $H[\tree|\utterance]$ normalized by sentence length.}
     \label{fig:distorted-parser}
@@
 
-We constructed a Bigram model with Kneser-Ney smoothing.
+We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
+In previous work, such as \cite{gildea2015human}, predictability has often been measured using n-gram models.
+
+Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
+Specifically, we constructed bigram models with Kneser-Ney smoothing~\cite{kneser1995improved, chen1999empirical}.
 A bigram model predicts each word taking only the previous word into account.
@@
     \centering
-    \includegraphics[scale=.4]{../results/bigrams/bigrams.pdf} 
+    \includegraphics[scale=.6]{../results/bigrams/bigrams.pdf} 
 	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).}
@@
 The resulting curves are shown in Figure~\ref{fig:learning-czech}.
-A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
+%A gap in parsing accuracy of about 0.07-0.1 appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
+A gap in parsing loss of about 0.2 nats appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
 
-\begin{figure}
+\begin{figure}[ht]
     \centering
-    \includegraphics[scale=.4]{figures/learning-parser-czech.pdf} 
-        \caption{Parsing accuracy (measured in UAS) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
+%    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech.pdf} 
+    \includegraphics[scale=.4]{../results/learning-curves/figures/learning-parser-czech-logloss.pdf}
+
+	\caption{Parsing loss ($H[\tree|\utterance]$, normalized by sentence length) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
     \label{fig:learning-czech}
@@
 
-\section{}
+%\section{Comparing Extracted Grammars to Linguistic Judgments}
 
-\bibliographystyle{acl_natbib}
+% user@user-X510UAR:~/grammar-optim/results/grammars$ git add comparison-table.tex
+
+\section{Languages and Corpus Sizes}
+In Table~\ref{tab:langs-iso-sizes}, we list the 51 languages with ISO codes and families, with the size of the available data per language.
+We included all UD 2.1 languages for which a training partition was available.
+
+\begin{table}[ht]
+\small{
+\begin{tabular}{lllllll}
+Language & ISO Code & Family & Sentences (train/held-out) & Words (train/held-out) \\ \hline
+\input{../results/corpus-size/corpora-table.tex}
+\end{tabular}
+}
+\caption{Languages with ISO codes, families (according to \url{https://universaldependencies.org/}), and the number of available sentence and words.}\label{tab:langs-iso-sizes}
+\end{table}
+
+\section{Dependency Length Minimization}\label{sec:DLM}
+
+Prior work has suggested \emph{Dependency Length Minimization} (DLM) as a characteristic of efficient word order~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+This is the idea that word order minimizes the average distance between syntactically related words.
+It is known that human languages reduce dependency length compared to random baselines~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+
+Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
+It is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
+Indeed, \cite{futrell2017memory} argues specifically that it emerges from efficiency optimization.
+
+Dependency length is typically quantified as the average distance between all pairs of syntactically related words, measured by the number of intervening words~\cite{futrell2015largescale}.
+Dependency length quantified in this manner is a heuristic measure of complexity:
+The actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
+
+We asked whether efficiency optimization predicts dependency length minimization effects.
+We first computed dependency length for grammars optimized for efficiency.
+We found that 100\% of grammars optimized for efficiency reduce average dependency length compared to baseline grammars ($p < 0.05$, by one-sided $t$-test). \textcolor{red}{TODO}
+This suggests that the reduction of dependency length is predicted by efficiency maximization.
+Next, we constructed grammars that minimize average dependency length, using the same gradient descent method as we used for efficiency optimization (Section~\ref{sec:optim-eff}).
+We expect that such grammars should have shorter dependency length than the real grammars, or grammars optimized for efficiency.
+In Figure~\ref{fig:dlm-avg}, we plot the mean dependency length for optimized, real, and baseline orderings.
+We find that optimizing grammars for efficiency reduces dependency length to a similar degree as found in the actual orderings in the corpora, almost up to the limit given by directly optimizing for dependency length.
+We also plot more detailed results for four languages in Figure~\ref{fig:dlm-4langs}, plotting dependency length as a function of sentence length as reported in prior work \cite{futrell2015largescale}.
+Optiziming grammars for efficiency produces dependency lengths similar to those found in the actual orderings.
+
+\begin{figure}[ht]
+    \centering
+     \includegraphics[width=0.5\textwidth]{../results/dependency-length/figures/depl-violin-all.pdf} 
+        \caption{Average dependency length for grammars optimized to minimize dependency length (DLM, left), optimized for efficiency (second), the real orderings found in corpora (third), and random baseline grammars (right).
+The lines connect the means points for each of the 51 languages in our sample.}
+    \label{fig:dlm-avg}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9\textwidth]{../results/dependency-length/figures/depLength-facet.pdf} 
+        \caption{Total dependency length as a function of sentence length, for four diverse languages.
+We show results for optimized grammars (parseability, predictability, efficiency), for grammars specifically optimized to minimize dependency length, of the actual real orderings, and of the baseline grammars.}
+    \label{fig:dlm-4langs}
+\end{figure}
+
+Next, we examined the word order properties of grammars optimized for DLM.
+In Table~\ref{tab:all-predictions-1}, we report the posterior prevalence of word order correlations in grammars optimized for DLM show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
+We find that these grammars also exhibit the eight correlations, similar to grammars directly optimized for efficiency.
+On other correlations, predictions of DLM also resemble those of efficiency optimization.
+However, it predicts strong correlations with \textit{amod} (adjectival modifiers) and \textit{nummod} (numeral modifiers) (see bottom of Table~\ref{tab:all-predictions-1}), which are not borne out typologically.
+In these cases, efficiency optimization predicts prevalences closer to 50\%, in line with typological data.
+
+In conclusion, these results suggest that dependency length minimization is a by-product of efficiency optimization, providing support to theoretical arguments from the linguistic literature~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
+Furthermore, optimizing for dependency length correctly predicts a range of word order facts, though it appears to \emph{overpredict} correlations when compared to direct optimization for efficiency.
+
+%Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM}).
+%Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
+%Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
+%Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
+%In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
+
+%Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
+%We optimize grammars using the same stochastic gradient descent method as we use for efficiency optimization (Section ).
+%Indeed, we found that DLM is predicted by efficiency optimization, as discussed in Section .
+%Results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
+
+%DLM turns out to be predicted by our model
+
+\nocite{diessel2001ordering, dryer1980positional, wals}
+
+\bibliographystyle{unsrtnat}
 \bibliography{everything}

### 2019-05-31 writeup/si-table-perrel-1.tex

--- 2019-05-23 second-draft/si-table-perrel-1.tex
@@
 
-\usepackage{hyperref}
-
 \usepackage{amssymb}% http://ctan.org/pkg/amssymb
@@
 \begin{document}
-\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
+\begin{tabular}{|l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
 	\hline
@@
 \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
 \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
@@
 \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
-&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section~\ref{sec:correlations} \\
+&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section S1 \\
 &amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\

### 2019-05-31 results/correlations/createSITable.py
cleaning code

@@ -1,3 +1,5 @@
+# Creates source for per-relation results tables in the SI.
+
 with open("../relations.tsv", "r") as inFile:
    rels = [x.strip() for x in inFile.read().strip().split("\n")]
 for x in rels:

### 2019-05-31 results/correlations/plot_byLanguage/plot_families_num2-all8.R

@@ -1,4 +1,4 @@
-# tiles-real_eff_large-byObj-restricted-viz-pred-allThree.R
+source("../tiles-real_eff_large-byObj-restricted-viz-pred-large-all8.R")
 
 
 
@@ -47,35 +47,6 @@ plot_orders_eff = ggplot(D %>% filter(Type == "Efficiency"), aes(x = 1, y = Lang
   scale_x_continuous(breaks = NULL) +
   scale_y_continuous(breaks = NULL)
 
-plot_orders_surp = ggplot(D %>% filter(Type == "Predictability"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
-  geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
-#  scale_color_gradient() + #values=c("blue", "green")) +
-  theme_classic() +
- theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-                     axis.text.y=element_blank(),axis.ticks=element_blank(),
-                     plot.title=element_text(size=11)) +
-  theme(axis.title=element_blank()) + 
-  theme(legend.position="none") + labs(x=NULL) +
-  scale_x_continuous(breaks = NULL) +
-  scale_y_continuous(breaks = NULL)
-
-plot_orders_pars = ggplot(D %>% filter(Type == "Parseability"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
-  geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
-#  scale_color_gradient() + #values=c("blue", "green")) +
-  theme_classic() +
- theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-                     axis.text.y=element_blank(),axis.ticks=element_blank(),
-                     plot.title=element_text(size=11)) +
-  theme(axis.title=element_blank()) + 
-  theme(legend.position="none") + labs(x=NULL) +
-  scale_x_continuous(breaks = NULL) +
-  scale_y_continuous(breaks = NULL)
-
-
-
-
-
-
 plot_langs = ggplot(DLang) 
 plot_langs = plot_langs +  theme_classic() 
[…]

### 2019-05-31 results/correlations/plot_byLanguage/plot_families_num2-allThree.R

@@ -1,4 +1,4 @@
-# tiles-real_eff_large-byObj-restricted-viz-pred-large.R
+source("./tiles-real_eff_large-byObj-restricted-viz-pred-allThree.R")
 
 
 
@@ -19,6 +19,8 @@ D = merge(D, DFam %>% select(FamilyPrint, yOffset), by=c("FamilyPrint"))
 
 DLang = unique(D %>% select(Language_Ordered, iso_Ordered, LanguageNumeric, yOffset))
 
+
+D = D %>% mutate(CoarseDependency = recode(CoarseDependency, lifted_case=1, lifted_cop=2, aux=3, nmod=4, acl=5, lifted_mark=6, obl=7, xcomp=8))
 
 plot_orders_real = ggplot(D %>% filter(Type == "Real Languages"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
   geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
@@ -45,6 +47,35 @@ plot_orders_eff = ggplot(D %>% filter(Type == "Efficiency"), aes(x = 1, y = Lang
   scale_x_continuous(breaks = NULL) +
   scale_y_continuous(breaks = NULL)
 
+plot_orders_surp = ggplot(D %>% filter(Type == "Predictability"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
+  geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
+#  scale_color_gradient() + #values=c("blue", "green")) +
+  theme_classic() +
+ theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
+                     axis.text.y=element_blank(),axis.ticks=element_blank(),
+                     plot.title=element_text(size=11)) +
+  theme(axis.title=element_blank()) + 
+  theme(legend.position="none") + labs(x=NULL) +
+  scale_x_continuous(breaks = NULL) +
+  scale_y_continuous(breaks = NULL)
+
+plot_orders_pars = ggplot(D %>% filter(Type == "Parseability"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
[…]

### 2019-05-31 results/correlations/plot_byLanguage/plot_families_num2.R

@@ -1,4 +1,4 @@
-# tiles-real_eff_large-byObj-restricted-viz-pred-large-dlm.R
+source("tiles-real_eff_large-byObj-restricted-viz-pred-large.R")
 
 
 
@@ -19,8 +19,6 @@ D = merge(D, DFam %>% select(FamilyPrint, yOffset), by=c("FamilyPrint"))
 
 DLang = unique(D %>% select(Language_Ordered, iso_Ordered, LanguageNumeric, yOffset))
 
-
-D = D %>% mutate(CoarseDependency = recode(CoarseDependency, lifted_case=1, lifted_cop=2, aux=3, nmod=4, acl=5, lifted_mark=6, obl=7, xcomp=8))
 
 plot_orders_real = ggplot(D %>% filter(Type == "Real Languages"), aes(x = 1, y = LanguageNumeric+yOffset, group=CoarseDependency)) + 
   geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
@@ -69,13 +67,7 @@ plot_orders_eff = plot_orders_eff + theme(                    plot.margin=unit(c
 
 
 plot = grid.arrange(plot_langs, plot_orders_real, plot_orders_eff, nrow=1, widths=c(1, 1.2, 1.2))
-ggsave(plot=plot, "figures/pred-eff-families-dlm.pdf", width=4, height=8)
-
-#plot1 = ggplot(DLang)
-#plot2 = ggplot(DLang) + annotate("text", label="Real")
-#plot3 = ggplot(DLang) + annotate("text", label="Optimized")
-
-
+ggsave(plot=plot, "../figures/pred-eff-families.pdf", width=4, height=8)
 
 
 
@@ -104,19 +96,19 @@ plot_orders_eff2
 plot = grid.arrange(plot_langs2, plot_orders_real2, plot_orders_eff2, nrow=1, widths=c(1, 1.2, 1.2))
 plot
 
-ggsave(plot=plot, "figures/pred-eff-families-2-dlm.pdf", width=4, height=8)
+
+
+
+
+
+
+
+
+ggsave(plot=plot, "../figures/pred-eff-families-2.pdf", width=4, height=8)
 
 D2 = (D %>% select(Family, Language, CoarseDependency, DirB, Type) %>% spread(Type, DirB) %>% rename(Real = 'Real Languages') %>% rename(Predicted = Efficiency))
 
[…]

### 2019-05-31 results/correlations/tiles-real_eff_large-byObj-restricted-viz-pred-allThree.R

@@ -159,47 +159,4 @@ D$iso_Ordered = factor(D$iso_code, levels=unique(D_Ground_Mean[order(D_Ground_Me
 
 D = D[order(D$Language_Ordered),]
 
-#plot = ggplot(D %>% filter(Type == "Efficiency"), aes(x = Family, y = Language_Ordered)) + 
-#  geom_point(aes(fill=DirB, colour = DirB, size =1)) +
-##  scale_color_gradient() + #values=c("blue", "green")) +
-#  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-#                     axis.text.y=element_text(size=9),
-#                     plot.title=element_text(size=11)) +
-#  theme(axis.title=element_blank()) + 
-#  theme(legend.position="none") #+ facet_grid(~Family) 
-#ggsave(file="test1.pdf")
-#
-#
-#plot = ggplot(D %>% filter(Type == "Real Languages"), aes(x = Family, y = Language_Ordered)) + 
-#  geom_point(aes(fill=DirB, colour = DirB, size =1)) +
-##  scale_color_gradient() + #values=c("blue", "green")) +
-#  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-#                     axis.text.y=element_text(size=9),
-#                     plot.title=element_text(size=11)) +
-#  theme(axis.title=element_blank()) + 
-#  theme(legend.position="none") #+ facet_grid(~Family) 
-#ggsave(file="test2.pdf")
-#
-#plot = ggplot(D %>% filter(Type == "Real Languages"), aes(x = Family, y = Language_Ordered)) + 
-#  geom_point(aes(fill=DirObj, colour = DirObj, size =1)) +
-##  scale_color_gradient() + #values=c("blue", "green")) +
-#  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-#                     axis.text.y=element_text(size=9),
-#                     plot.title=element_text(size=11)) +
-#  theme(axis.title=element_blank()) + 
-#  theme(legend.position="none") #+ facet_grid(~Family) 
-#ggsave(file="test3.pdf")
-#
-#plot = ggplot(D, aes(x = 1, y = Language_Ordered, group=CoarseDependency)) + 
-#  geom_point(aes(fill=DirB, colour = DirB, size =1), position = position_dodge(width=2.0)) +
[…]

### 2019-05-31 results/correlations/tiles-real_eff_large-byObj-restricted-viz-pred-large-all8.R

@@ -94,47 +94,5 @@ D$iso_Ordered = factor(D$iso_code, levels=unique(D_Ground_Mean[order(D_Ground_Me
 
 D = D[order(D$Language_Ordered),]
 
-#plot = ggplot(D %>% filter(Type == "Efficiency"), aes(x = Family, y = Language_Ordered)) + 
-#  geom_point(aes(fill=DirB, colour = DirB, size =1)) +
-##  scale_color_gradient() + #values=c("blue", "green")) +
-#  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-#                     axis.text.y=element_text(size=9),
-#                     plot.title=element_text(size=11)) +
-#  theme(axis.title=element_blank()) + 
-#  theme(legend.position="none") #+ facet_grid(~Family) 
-#ggsave(file="test1.pdf")
-#
-#
-#plot = ggplot(D %>% filter(Type == "Real Languages"), aes(x = Family, y = Language_Ordered)) + 
-#  geom_point(aes(fill=DirB, colour = DirB, size =1)) +
-##  scale_color_gradient() + #values=c("blue", "green")) +
-#  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-#                     axis.text.y=element_text(size=9),
-#                     plot.title=element_text(size=11)) +
-#  theme(axis.title=element_blank()) + 
-#  theme(legend.position="none") #+ facet_grid(~Family) 
-#ggsave(file="test2.pdf")
-#
-#plot = ggplot(D %>% filter(Type == "Real Languages"), aes(x = Family, y = Language_Ordered)) + 
-#  geom_point(aes(fill=DirObj, colour = DirObj, size =1)) +
-##  scale_color_gradient() + #values=c("blue", "green")) +
-#  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-#                     axis.text.y=element_text(size=9),
-#                     plot.title=element_text(size=11)) +
-#  theme(axis.title=element_blank()) + 
-#  theme(legend.position="none") #+ facet_grid(~Family) 
-#ggsave(file="test3.pdf")
-#
-#plot = ggplot(D, aes(x = 1, y = Language_Ordered, group=CoarseDependency)) + 
[…]

### 2019-05-31 results/correlations/tiles-real_eff_large-byObj-restricted-viz-pred-large-pars.R

@@ -94,47 +94,5 @@ D$iso_Ordered = factor(D$iso_code, levels=unique(D_Ground_Mean[order(D_Ground_Me
 
 D = D[order(D$Language_Ordered),]
 
-#plot = ggplot(D %>% filter(Type == "Efficiency"), aes(x = Family, y = Language_Ordered)) + 
-#  geom_point(aes(fill=DirB, colour = DirB, size =1)) +
-##  scale_color_gradient() + #values=c("blue", "green")) +
-#  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-#                     axis.text.y=element_text(size=9),
-#                     plot.title=element_text(size=11)) +
-#  theme(axis.title=element_blank()) + 
-#  theme(legend.position="none") #+ facet_grid(~Family) 
-#ggsave(file="test1.pdf")
-#
-#
-#plot = ggplot(D %>% filter(Type == "Real Languages"), aes(x = Family, y = Language_Ordered)) + 
-#  geom_point(aes(fill=DirB, colour = DirB, size =1)) +
-##  scale_color_gradient() + #values=c("blue", "green")) +
-#  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-#                     axis.text.y=element_text(size=9),
-#                     plot.title=element_text(size=11)) +
-#  theme(axis.title=element_blank()) + 
-#  theme(legend.position="none") #+ facet_grid(~Family) 
-#ggsave(file="test2.pdf")
-#
-#plot = ggplot(D %>% filter(Type == "Real Languages"), aes(x = Family, y = Language_Ordered)) + 
-#  geom_point(aes(fill=DirObj, colour = DirObj, size =1)) +
-##  scale_color_gradient() + #values=c("blue", "green")) +
-#  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-#                     axis.text.y=element_text(size=9),
-#                     plot.title=element_text(size=11)) +
-#  theme(axis.title=element_blank()) + 
-#  theme(legend.position="none") #+ facet_grid(~Family) 
-#ggsave(file="test3.pdf")
-#
-#plot = ggplot(D, aes(x = 1, y = Language_Ordered, group=CoarseDependency)) + 
[…]

### 2019-05-31 results/correlations/tiles-real_eff_large-byObj-restricted-viz-pred.R

@@ -76,27 +76,4 @@ D$Language_Ordered = factor(D$Language, levels=unique(D_Ground_Mean[order(D_Grou
 
 D = D[order(D$Language_Ordered),]
 
-#plot = ggplot(D %>% filter(Type == "Efficiency"), aes(x = CoarseDependency, y = Language_Ordered)) + 
-#  geom_point(aes(fill=DirB, colour = DirB, size =1)) +
-##  scale_color_gradient() + #values=c("blue", "green")) +
-#  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-#                     axis.text.y=element_text(size=9),
-#                     plot.title=element_text(size=11)) +
-#  theme(axis.title=element_blank()) + 
-#  theme(legend.position="none") #+ facet_grid(~Family) 
-#ggsave(file="test1.pdf")
-#
-#
-#plot = ggplot(D %>% filter(Type == "Real Languages"), aes(x = CoarseDependency, y = Language_Ordered)) + 
-#  geom_point(aes(fill=DirB, colour = DirB, size =1)) +
-##  scale_color_gradient() + #values=c("blue", "green")) +
-#  theme_bw() + theme(axis.text.x=element_blank(), #element_text(size=9, angle=0, vjust=0.3),
-#                     axis.text.y=element_text(size=9),
-#                     plot.title=element_text(size=11)) +
-#  theme(axis.title=element_blank()) + 
-#  theme(legend.position="none") #+ facet_grid(~Family) 
-#ggsave(file="test2.pdf")
-#
-##ggsave(file="figures/coverage-ground_eff-byObj-circles-viz.pdf", plot=plot)
-#
-#
+

### 2019-05-31 results/correlations/visualize_posterior/visualize-posterior-2-large.R

@@ -3,7 +3,7 @@ dependencies = c("acl", "advcl", "advmod", "amod", "appos", "aux", "ccomp", "com
 
 dependencies =  c("acl", "advmod", "aux", "lifted_case", "lifted_cop", "lifted_mark", "nmod", "nsubj", "obl", "xcomp")
 
-data = read.csv("../../grammars/manual_output_funchead_ground_coarse_final/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+data = read.csv("../../../grammars/manual_output_funchead_ground_coarse_final/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
 
 
 
@@ -17,7 +17,7 @@ data = data %>% mutate(DH_Weight = DH_Mean_NoPunct)
 library(forcats)
 data = data %>% mutate(Language = fct_recode(Language, "Old_Church_Slavonic" = "Old", "Ancient_Greek" = "Ancient"))
 
-languages = read.csv("../languages/languages-iso_codes.tsv", sep=",")
+languages = read.csv("../../languages/languages-iso_codes.tsv", sep=",")
 data  = merge(data, languages, by=c("Language"), all.x=TRUE)
 
 
@@ -112,7 +112,7 @@ data2 = data2 %>% mutate(Dependency = case_when(Dependency == "liftedcase" ~ "li
 data2 = data2 %>% mutate(Prevalence = case_when(Dependency == "aux" ~ 1-Prevalence, TRUE ~ Prevalence))
 
   plot = ggplot(data=data2) + geom_density(aes(x=Prevalence, y=..scaled.., fill=Type, group=Type), alpha=.5) + xlim(0,1) +  facet_wrap( ~ Dependency, ncol=1) 
-  ggsave(plot, file=paste("figures/posterior_joint_large" , ".pdf", sep=""))
+  ggsave(plot, file=paste("../figures/posterior_joint_large" , ".pdf", sep=""))
 
 
 
@@ -123,7 +123,7 @@ data2 = data2 %>% mutate(Prevalence = case_when(Dependency == "aux" ~ 1-Prevalen
 
 
   plot = ggplot(data=data2) + geom_density(aes(x=Prevalence, y=..scaled.., fill=Type, group=Type), alpha=.5) + xlim(0,1) + geom_bar(data=D, stat="identity", width = 0.01, aes(x=Agree, y=1, fill=Type, group=Type)) + facet_wrap( ~ Dependency, ncol=1) 
[…]

### 2019-05-31 results/correlations/visualize_posterior/visualize-posterior-2-prior1.0.R

@@ -6,7 +6,7 @@ dependencies = c("acl", "advcl", "advmod", "amod", "appos", "aux", "ccomp", "com
 
 dependencies =  c("acl", "aux", "lifted_case", "lifted_cop", "lifted_mark", "nmod", "obl", "xcomp")
 
-data = read.csv("../../grammars/manual_output_funchead_ground_coarse_final/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+data = read.csv("../../../grammars/manual_output_funchead_ground_coarse_final/auto-summary-lstm.tsv", sep="\t")# %>% rename(Quality=AverageLength)
 
 
 
@@ -20,7 +20,7 @@ data = data %>% mutate(DH_Weight = DH_Mean_NoPunct)
 library(forcats)
 data = data %>% mutate(Language = fct_recode(Language, "Old_Church_Slavonic" = "Old", "Ancient_Greek" = "Ancient"))
 
-languages = read.csv("../languages/languages-iso_codes.tsv", sep=",")
+languages = read.csv("../../languages/languages-iso_codes.tsv", sep=",")
 data  = merge(data, languages, by=c("Language"), all.x=TRUE)
 
 
@@ -126,7 +126,7 @@ plot = plot + theme(text = element_text(size=14),
         axis.text.x = element_text(angle=90, hjust=1)) 
 plot = plot + theme(legend.position="none")
 
-ggsave(plot=plot, filename="figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf", width=3, height=3)
+ggsave(plot=plot, filename="../figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf", width=3, height=3)
 
 
 

### 2019-06-01 writeup/si-table-perrel-1.tex
code

@@
 \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
+\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
 \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\

### 2019-06-01 writeup/pnas-word-order-si.tex
SI

@@
 
+\textcolor{red}{TODOs}
+
 \textcolor{red}{run czech parser}
@@
 Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
-\cite{dryer1992greenbergian} formulated these universals as correlations between the order of objects and verbs and the orders of other syntactic relations.
+\citet{dryer1992greenbergian} formulated these universals as correlations between the order of objects and verbs and the orders of other syntactic relations.
 We test our ordering grammars for these correlations by testing whether the coefficients for these syntactic relations have the same sign as the coefficient of the verb-object relation.
@@
 Here we discuss how our formalization of communicative efficiency relates to formalizations that have been proposed in the information-theoretic literature on language.
-Across the literature, the core idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms, with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global, kemp2012kinship,xu2014numeral, regier2015word,xu2016historical, zaslavsky2018efficient,futrell2017memory,peloquin2019interactions, zaslavsky2019semantic, frank2012predicting,goodman2013knowledge, kao2014nonliteral, bennett2018extremely}.
+Across the literature, the core idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms, with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global,kemp2012kinship,frank2012predicting,goodman2013knowledge,kao2014nonliteral,xu2014numeral,kirby2015compression,regier2015word,xu2016historical,futrell2017memory,zaslavsky2018efficient,bennett2018extremely, peloquin2019interactions,zaslavsky2019semantic}.
 
@@
 Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}), in which case not all correlations are predicted (Figure~\ref{fig:posterior}).
+See also Figure~\ref{table:corr-resu-previous} for converging evidence from our preregistered preliminary experiments.
 %Finally, the robustness of our results to small variations of $\lambda$ is confime by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu-previous}.
@@
 `?' indicates that WALS has no data.
-$*$ indicates that WALS does not list a dominant order, which can mean that both orders occur at similar rates in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
+$*$ indicates that WALS does not list a dominant order, which can mean that neither order is dominant in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
 Finally, `--' indicates that the relation does not occur in the corpus.
@@
 where
-\begin{equation}
+\begin{equation}\label{eq:rpars}
 	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
@@
 A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.
-The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is\footnote{Note that, in the definition of $R_{Pars}$, the term $p(t)$ is a constant independent of $\phi$ and the word order grammar $\mathcal{L}_\theta$; it can therefore be ignored in the optimization process.}
+The degree to which a parser with parameters $\phi$ succeeds in parsing a sentence $u$ with structure $t$ is\footnote{Note that, in the definition of $R_{Pars}$ (\ref{eq:rpars}), the term $p(t)$ is a constant independent of $\phi$ and the word order grammar $\mathcal{L}_\theta$; it can therefore be ignored in the optimization process.}
 \begin{equation}
-	R_{Pars}^{\phi}(u,t) =  \log \frac{p_\phi(t|u)}{p(t)}
+	R_{Pars}^{\phi}(u,t) =  \log p_\phi(t|u)
 \end{equation}
@@
 \end{equation}
-Using this, we can rewrite Efficiency~(\ref{eq:efficiency-recall}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
+Using this and Gibbs' inequality~\cite{cover2006elements}, we can rewrite Efficiency~(\ref{eq:efficiency-recall}), for a given grammar $\theta$, equivalently into the parseability and predictability achieved with the best parser and language models:
 \begin{equation}
-	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi} := \max_{\phi,\psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi}
 \end{equation}\label{eq:efficiency-rewrite}
+where we have written
+\begin{equation}
+R_{\textit{Eff}}^{\theta, \phi, \psi} := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\end{equation}
 %
@@
 %equivalently into
-In order to find an optimal grammar $\theta$, we thus need to solve
+In order to find an optimal grammar $\theta$, we thus need to compute 
 \begin{equation}\label{eq:efficiency}
-\argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} %= \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+\argmax_\theta\	R_{\textit{Eff}}^{\theta} = \argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} 
+%= \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
 \end{equation}
@@
 \end{align}
-where \ref{eq:dtheta} is derived using the \emph{score-function} or \emph{REINFORCE} trick~\cite{williams1992simple}.
+where (\ref{eq:dtheta}) is derived using the \emph{score-function} or \emph{REINFORCE} theorem~\cite{williams1992simple}.
 Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures.
@@
 Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
-do a gradient descent step using the estimator given by the expressions in the square brackets in \ref{eq:dtheta}-\ref{eq:dpsi}.
+do a gradient descent step using the estimator given by the expressions in the square brackets in (\ref{eq:dtheta}-\ref{eq:dpsi}).
 

### 2019-06-02 writeup/pnas-word-order-si.tex
SI

@@
 All three types of elements occur at most in parts of the 51 UD languages, and none of them is annotated reliably in those languages where they occur.
-Among these three types of elements, the one most prominent in our sample of 51 languages is articles (occurring in many European languages); UD subsumes them under the \emph{det} relation, which is also used for other frequent elements, such as demonstratives and quantifiers.
+Among these three types of elements, the one most prominent in our sample of 51 languages is articles, which occur in many European languages.
+However, UD subsumes them under the \emph{det} relation, which is also used for other highly frequent elements, such as demonstratives and quantifiers.
+%Testing this correlation would require augmenting these corpora with detailed annotation of articles, which is beyond the scope of this study.
+The other elements (question particles and plural words) are found at most in a handful of UD languages, and are not specifically annotated in these either.
 
@@
 All types of adverbs are unified under the \emph{advmod} relation.
-In the real orderings in our sample of 51 UD languages, the dominant ordering of \emph{advmod} almost always matches that of subjects -- that is, \emph{advmod} dependents are ordered after the verb only in VSO languages.
-This ordering behavior is very different from that documented for manner adverbs by Dryer.
+In the real orderings in our sample of 51 UD languages, the dominant ordering of \emph{advmod} almost always matches that of subjects -- that is, \emph{advmod} dependents are predominantly ordered after the verb only in VSO languages.
+This observed ordering behavior in the 51 languages is very different from that documented for manner adverbs by Dryer, showing that a large part of \emph{advmod} dependencies as annotated in UD consists of elements that are not manner adverbs.
 
@@
 As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure.
-Furthermore, entropy also arises naturally as a complexity measure from the role played by surprisal in determining human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+Entropy also corresponds to the average of surprisal, which is a strong and linear predictor of human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
 
@@
 The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
-In the last column, we indicate what direction would be expected typologically.}
+In the last column, we indicate what direction would be expected typologically.
+See Section~\ref{sec:DLM} for the `DLM' column.}
 \label{tab:all-predictions-1}
@@
 We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
-To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
+To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.\footnote{We implemented this regularized optimization as mean-field ELBO variational inference in Pyro~\cite{bingham2018pyro}.}
 We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.

### 2019-06-03 writeup/pnas-word-order-si.tex
SI

@@
 In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parseability.
-While the computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$, we also examined word order predictions for grammars optimized for only parseability or only predictability, in order to tease apart predictions made by these two components.
+The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
+However, we also examined word order predictions for grammars optimized for only parseability or only predictability, in order to tease apart predictions made by these two components.
 As shown in Table~\ref{tab:all-predictions-1}, each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions.
-%Indeed, the same holds for all of the 38 syntactic relations in UD  (Table~\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})
-%We additionally report results with $\lambda = 0$, i.e., placing all weight on parseability.
 That is, at $\lambda$ close to its maximal value, the predictions of optimizing the two scoring functions individually add up to the predictions of efficiency optimization.\footnote{Results from one of the preliminary experiments reported in Figure~\ref{table:corr-resu-previous} show that results are stable to small variation of $\lambda$: Essentially equivalent predictions are obtained for $\lambda=1.0$.
@@
 Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}), in which case not all correlations are predicted (Figure~\ref{fig:posterior}).
-See also Figure~\ref{table:corr-resu-previous} for converging evidence from our preregistered preliminary experiments.
+This is confirmed by converging evidence from our preregistered preliminary experiments in Figure~\ref{table:corr-resu-previous}.
+
 %Finally, the robustness of our results to small variations of $\lambda$ is confime by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu-previous}.
@@
 
+\textcolor{red}{have to improve DLM coverage}
 Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM} for discussion).
@@
 In the last column, we indicate what direction would be expected typologically. 
-See Section~\ref{sec:DLM} for the `DLM' column.}
+We additionally report results when optimizing for Dependency Length Minimization (DLM), a heuristic approximation to efficiency that has been suggested in previous work.
+See Section~\ref{sec:DLM} for further discussion.
+\textcolor{red}{have to improve DLM coverage}   }
 \label{tab:all-predictions-1}
@@
 %In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
+
+\textcolor{red}{TODO put language modeling and parsing formulas to make this more explicit}
 

### 2019-06-03 writeup/pnas-word-order-si.tex
SI

@@
 
-We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding objections by \citet{dunn2011evolved}, we refer to the follow-ups by \citet{levy2011computational}, and \citet{croft2011greenbergian}.}
+We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding the objections by \citet{dunn2011evolved}, we refer to the follow-ups by \citet{levy2011computational}, and \citet{croft2011greenbergian}.}
 Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
@@
 
-\textcolor{red}{have to improve DLM coverage}
-Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM} for discussion).
-This has been suggested as a property of efficient word order in prior work, and, indeed, we found that DLM is predicted by efficiency optimization (Section~\ref{sec:DLM}).
-Results in Table~\ref{tab:all-predictions-1} also show that optimizing for DLM makes predictions similar to efficiency optimization. See Section~\ref{sec:DLM} for further discussion.
+%We also report results for grammars optimized for Dependency Length Minimization (DLM), which has been argued to be a property of efficient word order, and which we find to be predicted by efficiency optimization (see Section~\ref{sec:DLM}).
+%In line with this idea, our results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization.
+%These results suggest that it is in part through favoring short dependencies that efficiency predicts word order universals, an idea that has been proposed in prior theoretical studies \citep{hawkins1994performance, hawkins2004efficiency}.
+%%Previous theoretical studies arguing for efficiency explanations of word order universals have already suggested that principles akin to DLM explain the Greenberg correlations ; our results provide the first large-scale empirical confirmation of this idea, and show that this can be reduced to efficiency optimization.
+%See Section~\ref{sec:DLM} for further discussion.
+%
 
@@
 	\begin{center}	
-\includegraphics[width=0.8\textwidth]{si-table-perrel-1-1.png}
+\includegraphics[width=0.8\textwidth]{si-table-perrel-1a-1.png}
 \end{center}
@@
 In the last column, we indicate what direction would be expected typologically. 
-We additionally report results when optimizing for Dependency Length Minimization (DLM), a heuristic approximation to efficiency that has been suggested in previous work.
-See Section~\ref{sec:DLM} for further discussion.
-\textcolor{red}{have to improve DLM coverage}   }
+%We additionally report results obtained when optimizing for Dependency Length Minimization (DLM), a heuristic approximation to efficiency that has been suggested in previous work.
+See Section~\ref{sec:DLM} for further discussion of DLM and its relation to efficiency.
+}
 \label{tab:all-predictions-1}
@@
 	\begin{center}	
-\includegraphics[width=0.8\textwidth]{si-table-perrel-2-1.png}  
+\includegraphics[width=0.8\textwidth]{si-table-perrel-2a-1.png}  
 \end{center}
@@
 In Table~\ref{table:corr-resu-previous} we report the results of our two previous, preregistered, simulations\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}, \url{https://aspredicted.org/blind.php?x=bg35x7}.
-For the results of the locality simulations described in the first preregistration, see the DLM results in Table~\ref{tab:all-predictions-1}, see also Section~\ref{sec:DLM} for discussion.} together with results from the main experiment.
+For the results of the locality simulations described in the first preregistration, see the Dependency Length Minimization results in Table~\ref{tab:all-predictions-1a}, with discussion in Section~\ref{sec:DLM}.} together with results from the main experiment.
 These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) one simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
@@
 It is known that human languages reduce dependency length compared to random baselines~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-
-Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
+Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,hawkins2004efficiency,futrell2017memory, futrell2017generalizing}.
 It is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
@@
 We first computed dependency length for grammars optimized for efficiency.
-We found that 100\% of grammars optimized for efficiency reduce average dependency length compared to baseline grammars ($p < 0.05$, by one-sided $t$-test). \textcolor{red}{TODO}
-This suggests that the reduction of dependency length is predicted by efficiency maximization.
+We found that 100\% of grammars optimized for efficiency reduce average dependency length compared to baseline grammars ($p < 0.05$, by one-sided $t$-test). 
+This suggests that the reduction of dependency length observed in natural language is indeed predicted by efficiency maximization, confirming theoretical arguments made in prior work~\cite{hawkins1994performance,hawkins2004efficiency,futrell2017memory, futrell2017generalizing}.
 Next, we constructed grammars that minimize average dependency length, using the same gradient descent method as we used for efficiency optimization (Section~\ref{sec:optim-eff}).
@@
 Next, we examined the word order properties of grammars optimized for DLM.
-In Table~\ref{tab:all-predictions-1}, we report the posterior prevalence of word order correlations in grammars optimized for DLM show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
+In Table~\ref{tab:all-predictions-1b}, we report the posterior prevalence of word order correlations in grammars optimized for DLM; our results show that optimizing for DLM makes predictions similar to efficiency optimization.
+This is itself a novel result, suggesting that it is in part through favoring short dependencies that efficiency predicts word order universals, an idea that has been proposed in prior theoretical studies \citep{hawkins1994performance, hawkins2004efficiency}.
 We find that these grammars also exhibit the eight correlations, similar to grammars directly optimized for efficiency.
 On other correlations, predictions of DLM also resemble those of efficiency optimization.
-However, it predicts strong correlations with \textit{amod} (adjectival modifiers) and \textit{nummod} (numeral modifiers) (see bottom of Table~\ref{tab:all-predictions-1}), which are not borne out typologically.
+However, it predicts strong correlations with \textit{amod} (adjectival modifiers) and \textit{nummod} (numeral modifiers) (see bottom of Table~\ref{tab:all-predictions-1b}), which are not borne out typologically.
 In these cases, efficiency optimization predicts prevalences closer to 50\%, in line with typological data.
 
-In conclusion, these results suggest that dependency length minimization is a by-product of efficiency optimization, providing support to theoretical arguments from the linguistic literature~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
-Furthermore, optimizing for dependency length correctly predicts a range of word order facts, though it appears to \emph{overpredict} correlations when compared to direct optimization for efficiency.
+\begin{table}[ht] % TODO figure out away to make this table more aesthetic
+	\begin{center}	
+\includegraphics[width=0.7\textwidth]{si-table-perrel-1b-1.png}
+\end{center}
+\caption{Predictions on UD relations with predictions from the typological literature (compare Table~\ref{tab:all-predictions-1}), for languages optimized for Efficiency and Dependency Length Minimization.
+}
+\label{tab:all-predictions-1b}
+\end{table}
+
+In conclusion, these results suggest that the phenomenon of dependency length minimization is a by-product of efficiency optimization, providing support to theoretical arguments from the linguistic literature~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
+Furthermore, optimizing for dependency length correctly predicts a range of word order facts, though it appears to \emph{overpredict} correlations when compared to direct optimization for communicative efficiency.
+
+%
+%We also report results for grammars optimized for Dependency Length Minimization (DLM), which has been argued to be a property of efficient word order, and which we find to be predicted by efficiency optimization (see Section~\ref{sec:DLM}).
+%In line with this idea, our results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization.
+%These results suggest that it is in part through favoring short dependencies that efficiency predicts word order universals, an idea that has been proposed in prior theoretical studies \citep{hawkins1994performance, hawkins2004efficiency}.
+%%Previous theoretical studies arguing for efficiency explanations of word order universals have already suggested that principles akin to DLM explain the Greenberg correlations ; our results provide the first large-scale empirical confirmation of this idea, and show that this can be reduced to efficiency optimization.
+%See Section~\ref{sec:DLM} for further discussion.
+%
+%
+%
+%
+%\begin{table*} % TODO figure out away to make this table more aesthetic
+%        \begin{center}
+%\includegraphics[width=0.8\textwidth]{si-table-perrel-1a-1.png}
+%\end{center}
+%\caption{Predictions on UD relations with predictions from the typological literature.
+%The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
+%In the last column, we indicate what direction would be expected typologically.
+%We additionally report results obtained when optimizing for Dependency Length Minimization (DLM), a heuristic approximation to efficiency that has been suggested in previous work.
+See Section~\ref{sec:DLM} for further discussion of DLM and its relation to efficiency.
+}
+%\label{tab:all-predictions-1}
+%\end{table*}
+%
+%
+%
+%
 

### 2019-06-03 writeup/si-table-perrel-2a.tex
SI

% https://tex.stackexchange.com/questions/171982/how-to-make-table-floats-work-with-preview-standalone-environment/171983

\documentclass[%
    ,float=false % this is the new default and can be left away.
    ,preview=true
    ,class=scrartcl
%    ,fontsize=20pt
    ]{standalone}

\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{xcite}
\externalcitedocument{pnas-word-order-si}

\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage{siunitx}

\usepackage{longtable}

\usepackage[numbers]{natbib}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{arydshln}

\newcommand{\key}[1]{\textbf{#1}}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}

\usepackage{mathtools}
\DeclarePairedDelimiter{\ceiling}{\lceil}{\rceil}

\begin{document}
\begin{tabular}{|l|l|ll|l|l|llllllllllllllllllllllllllll}
	\hline
	Relation & Real & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
\hline
appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\           
lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}    &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\          
expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\         
iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\     
vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}    &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\

\hline
% Uninterpretable
compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}    &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\        
det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\        
dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}    &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\               
dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}     &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}    &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\

\hline
% UD artifacts
conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\ 
fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\ 
flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}     &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}    &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\                                        
parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}    &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\   
reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}     &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\

 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
\end{tabular}
\end{document}

### 2019-06-03 writeup/si-table-perrel-1a.tex

% https://tex.stackexchange.com/questions/171982/how-to-make-table-floats-work-with-preview-standalone-environment/171983

\documentclass[%
    ,float=false % this is the new default and can be left away.
    ,preview=true
    ,class=scrartcl
%    ,fontsize=20pt
    ]{standalone}

\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{xcite}
\externalcitedocument{pnas-word-order-si}

\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage{siunitx}

\usepackage{longtable}

\usepackage[numbers]{natbib}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{arydshln}

\newcommand{\key}[1]{\textbf{#1}}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}

\usepackage{mathtools}
\DeclarePairedDelimiter{\ceiling}{\lceil}{\rceil}

\begin{document}
\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
	\hline
&	Relation & Real &  Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
% From Dryer	
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}     &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}    &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}    &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}     &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
%\hdashline

\hline
% Other universals, not in Dryer
&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf. \cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section S1 \\
&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}     &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\ 
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
\end{tabular}
\end{document}

### 2019-06-03 writeup/si-table-perrel-1b.tex

% https://tex.stackexchange.com/questions/171982/how-to-make-table-floats-work-with-preview-standalone-environment/171983

\documentclass[%
    ,float=false % this is the new default and can be left away.
    ,preview=true
    ,class=scrartcl
%    ,fontsize=20pt
    ]{standalone}

\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{xcite}
\externalcitedocument{pnas-word-order-si}

\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage{siunitx}

\usepackage{longtable}

\usepackage[numbers]{natbib}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{arydshln}

\newcommand{\key}[1]{\textbf{#1}}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}

\usepackage{mathtools}
\DeclarePairedDelimiter{\ceiling}{\lceil}{\rceil}

\begin{document}
\begin{tabular}{|l|l|l|l|l|l|llllllllllllllllllllllllllll}
	\hline
&	Relation & Real & DLM & Efficiency & Expected Prevalence  \\ \hline
% From Dryer	
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}    &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}       &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &     \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &    \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}     &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
%\hdashline

\hline
% Other universals, not in Dryer
&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &    \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}     &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf. \cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}     &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}      &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section S1 \\
&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}     &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &    \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\ 
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
\end{tabular}
\end{document}

### 2019-06-04 writeup/pnas-word-order-si.tex

@@
 
-See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates used in this optimization scheme.
+See Section~\ref{sec:neural-architectures}, paragraph `Optimization Details' for the stopping criterion and learning rates used in this optimization scheme.
 
@@
 
-\textcolor{red}{TODO put language modeling and parsing formulas to make this more explicit}
-
 \paragraph{Estimating Predictability}
 We choose a standard LSTM language model \citep{goldberg2017neural, hochreiter1997long}.
+This model uses a recurrent neural network to compute the predictability of a sentence $u = u_1...u_n$\footnote{Technically, $u_1...u_{n-1}$ are words, and $u_n$ is an end-of-sentence token, to ensure the probability distribution over all sentences is normalized.}:
+\begin{equation}
+\log p_\psi(u) = \sum_{i=1}^n \log p_\psi(u_i|u_{1\dots i-1})
+\end{equation}
+where $\psi$ are the parameters of the recurrent LSTM network, optimized on training data (see paragraph `Optimization Details').
+We estimate the average predictability of a language as a Monte Carlo estimate on held-out data:
+\begin{equation}
+	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p_\psi(u) \approx \frac{1}{|\text{Heldout Data}|} \sum_{u \in \text{Heldout Data}} \log p_\psi(u)
+\end{equation}
+by averaging over all sentences $u$ occurring in the corpus.
+
 For computational reasons, we restrict the vocabulary to the most frequent 50,000 words in the treebanks for a given language.
@@
 
+%\begin{equation}\label{eq:rpars}
+%	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
+%\end{equation}
+%\begin{equation}
+%	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
+%\end{equation}
+
 \paragraph{Estimating Parseability}
@@
 This is a highly generic architecture for recovering graph structures from strings, and is a simplification of graph-based parsers which reduce the parsing problem to a minimal spanning tree problem \citep{mcdonald2005nonprojective}.
+The parseability of a sentence $u = u_1\dots u_n$ with syntactic structure $t$ is computed as
+\begin{equation}
+	\log p_\phi(t|u) = \sum_{i=1}^n \log p_\phi(\text{head}_i, \text{label}_i | u, i)
+\end{equation}
+where $\text{head}_i \in \{\textsc{root}, 1,\dots,n\}$ is the index of the head of $u_i$ in the syntactic structure, and $\text{label}_i$ is its syntactic relation as formalized in UD; $\phi$ denotes the parameters estimated on the training data (see paragraph `Optimization Details').
+The overall parseability is estimated as a Monte Carlo estimate on held-out data:
+\begin{equation}\label{eq:rpars}
+	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p_\phi(t|u)}{p(t)} = \frac{1}{|\text{Heldout Data}|} \sum_{t,u \in \text{Heldout Data}} \log \frac{p_\phi(t|u)}{p(t)}
+\end{equation}
+The constant $p(t)$ only depends on the language (but not on the word order rules), and can thus be ignored when comparing different grammars applied to the same language, and when optimizing grammars for a given language; we therefore do not attempt to explicitly estimate it.
 
@@
 
-\paragraph{Learning Rate and Stopping Criterion}
+\paragraph{Optimization Details}
 We update word order grammar parameters $\theta$ using Stochastic Gradient Descent with momentum.

### 2019-06-21 writeup/pnas-word-order-si.tex
small fixes

@@
 Here, we investigate whether the difference between real and baseline grammars is affected by the size of available datasets.
-If the difference between random and real grammars is due to data sparsity, we expect that it decreases as the amount of training data is increased.
-If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference persists as training data is increased.
+We are addressing the following confound: It is conceivable that with enough data, our neural network language models and parsers would do equally well on real grammars and baseline grammars.
+If the difference between random and real grammars is due to data sparsity in this way, then we expect that the difference will decrease as the amount of training data is increased.
+If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference will persists as training data is increased.
 
 We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
-We cosidered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
+We considered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.
 For both of these ordering grammars, we trained the parser on successively larger portions of the training data (0.1 \%, 1 \%, 5\%, 10\%, 20 \%, ..., 90 \%, 100 \%) and recorded parsing accuracy.
@@
 Dependency length quantified in this manner is a heuristic measure of complexity:
-The actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
+The actual empirically-measured processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
 

### 2019-06-25 writeup/archive/3-paper-names.tex
code

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

%https://tex.stackexchange.com/questions/219676/how-to-change-spacing-before-and-after-paragraph-command
%\usepackage{xpatch}

%\usepackage{pslatex}
%\usepackage{latexsym}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
%\usepackage[colorinlistoftodos]{todonotes}
\usepackage{rotating}
\usepackage{multirow}

\newcommand{\R}[0]{\mathbb{R}}
\newcommand{\Ff}[0]{\mathcal{F}}
\newcommand{\key}[1]{\textbf{#1}}

\newcommand{\soft}[1]{}
\newcommand{\nopreview}[1]{}
\newcommand\comment[1]{{\color{red}#1}}
\newcommand\mhahn[1]{{\color{red}(#1)}}
\newcommand{\rljf}[1]{{\color{blue}[rljf: #1]}}

\newcommand{\thetad}[0]{{\theta_d}}
\newcommand{\thetal}[0]{{\theta_{LM}}}
\newcommand{\thetap}[0]{{\theta_{P}}}

\renewcommand{\baselinestretch}{0.99}
\addtolength{\textfloatsep}{-0.18in}
\setlength{\abovecaptionskip}{5pt plus 1pt minus 0pt} % Chosen fairly arbitrarily
%\setlength{\belowcaptionskip}{-5pt}
%https://tex.stackexchange.com/questions/23313/how-can-i-reduce-padding-after-figure

\title{Universals of word order result from optimization of grammars for efficient communication}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Michael Hahn}
\author[a]{Dan Jurafsky}
\author[b]{Richard Futrell}

\affil[a]{Stanford University}
\affil[b]{University of California, Irvine}

% Please give the surname of the lead author for the running footer
\leadauthor{Hahn}

% Please add here a significance statement to explain the relevance of your work
% limit 120 words 
\significancestatement{
%What explains the universal properties of human languages? We present evidence that a major subset of these properties can be explained by viewing languages as codes for efficient communication among agents with highly generic cognitive constraints. In doing so, we provide the first full formalization and computational implementation of ideas which have been stated informally in the functional linguistics literature for decades. The success of this approach suggests a new way to conceptualize human language in quantitative and computational work, as an information-theoretic code dynamically shaped by communicative and cognitive pressures. Our results argue against the idea that the distinctive properties of human language result from essentially arbitrary genetic constraints.
Human languages share many grammatical properties.
We show that some of these properties 
can be explained by the need for languages to offer
efficient communication between humans given our cognitive constraints. Grammars of languages seem to find a balance between two communicative pressures: to be simple enough to allow the speaker to easily produce sentences, but complex enough to be unambiguous to the hearer, and this balance explains well-known word order generalizations across our sample of 51 varied languages.   Our results offer new quantitative and computational evidence that language structure, rather than being an outcome of arbitrary genetic constraints, is dynamically shaped by communicative and cognitive pressures.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{MH and RF designed research. MH implemented models and experiments. MH, DJ, and RF wrote the paper.}
\authordeclaration{The authors declare no conflict of interest.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: mhahn2@stanford.edu}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{language universals $|$ language processing $|$ computational linguistics}

\begin{abstract}
The universal properties of human languages have been the subject
of intense study across the language sciences.
We report novel computational and corpus evidence for
the hypothesis that a prominent subset of these universal
properties---those related to word order---result from a
process of optimization for efficient communication among humans,
trading off the need to reduce complexity with the need to reduce ambiguity.
We formalize these two pressures with information-theoretic and neural network models
of complexity and ambiguity, and simulate grammars optimizing their
word order parameters on data from 51 languages. 
Evolution of grammars towards efficiency results in word order patterns that
predict a large subset of the major word order correlations 
across languages.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}
\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

%\textcolor{red}{A length estimate is not required for initial submissions, but 6-page articles should be under 49,000 characters (including spaces, figures, and tables)  An online submission tool estimates whether the manuscript fits within PNAS length requirements (see Length Estimate Guidelines and FAQ). When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for 6-page articles and 65,000 for 10-page articles.}

%\textcolor{red}{Provide these separately from figures, after the references in the manuscript. For figures with multiple panels, the first sentence of the legend should be a brief overview of the entire figure; each panel must be explicitly referenced and described at least once in the figure legend. Graphs should include clearly labeled error bars described in the figure legend. Authors must state whether a number that follows the +- sign is a standard error (SEM) or a standard deviation (SD). The P value, magnification, or scale bar information should be provided when applicable. The number of independent data points (N) represented in a graph must be indicated in the legend. Numerical axes on graphs should go to 0, except for log axes.}

%\textcolor{red}{High-resolution figure files are not required for initial submissions. Resolution of at least 300 dpi for all figures is required. EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main text. Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format. (See SI below for supplementary material.) Color images must be in RGB (red, green, blue) mode. Include the font files for any text.}

%\textcolor{red}{Images must be final size, preferably 1 column width (8.7 cm). Figures wider than 1 column should be sized to 11.4 cm or 17.8 cm wide. Numbers, letters, and symbols should be no smaller than 6 points (2 mm) and no larger than 12 points (6 mm) after reduction and must be consistent. Composite figures must be preassembled. Figures must be submitted as separate files, not embedded in manuscript text. See the PNAS Digital Art Guidelines. Figures and tables may be enlarged to improve legibility of text.}

%\textcolor{red}{Statistical Analysis: Authors should include the source and version of any software used, }

\dropcap{U}nderstanding what is universal and what varies across human languages is a central goal of linguistics.
Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic, nowak1999evolution, nowak2001towards, nowak2002computational}.
But formalizing  how these pressures explain specific grammatical universals has proved difficult.
Here we pair new computational models that measure the communicative efficiency of grammars with a new simulation framework for finding optimal grammars, and show that the most efficient grammars also exhibit a large class of language universals. % information-theoretic

The language universals we study are the well-known Greenberg universals of word order \cite{greenberg1963universals}. 
Human languages vary in the order in which they express information.
Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
Both sentences contain a verb meaning `wrote', a noun expressing  the object `letter', and a phrase translating to `to a friend'.
Yet, the order of these words are entirely different in the two languages:
The verb stands at the beginning in Arabic, and at the end in Japanese.
Arabic expresses `to' by a \emph{preposition} (\emph{preceding} the noun `friend'); Japanese uses a \emph{postposition} (\emph{following} it).

Yet this variation reflects a deep and stable regularity:
While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
this is strongly correlated with the occurrence of pre- or postpositions (Figure~\ref{fig:corr-table}, top):
Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{figures/corr-full.png}
\caption{ One word order correlation: Languages can order the object after (Arabic) or before (Japanese) the verb, and have prepositions (Arabic) or postpositions (Japanese). For each combination, we indicate how many languages satisfy it, as documented in the World Atlas of Language Structures~\cite{wals}. Combinations on the diagonal are vastly more common than off-diagonal ones. 
	}\label{fig:arabic-japanese-simple}	\label{fig:corr-table}
\end{figure}

This generalization lies in a group of language universals originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}.
These describe correlations between the relative positions of different types of expressions across languages.
The example above documents that the position of the object (`letter') relative to the verb is \key{correlated} with the position of the adposition (`to'). %adpositional phrases (`to a friend') relative to the verb.
Greenberg also found that the order of verb and object is correlated with other aspects of a language's word order (Table~\ref{table:corr-dryer}), such the order of verb and adpositional phrase (`wrote -- to friend' in Arabic vs `friend to -- wrote' in Japanese), and that of noun and genitive (`book -- of friend' in Arabic, `friend of -- book' in Japanese).

\begin{table}[ht]
	\begin{center}
%		\setlength\tabcolsep{1.5pt}
\begin{tabular}{|c|ll|ll|}
	\hline
	%&
	&	\multicolumn{2}{c|}{\textbf{Arabic} (English, ...)}   &        \multicolumn{2}{c|}{\textbf{Japanese} (Turkish, ...)} \\ \hline\hline
	& \multicolumn{2}{c}{Correlates with...} & \multicolumn{2}{|c|}{Correlates with...}  \\
	&	Verb & Object     & Object & Verb    \\ \hline\hline
	& kataba & rasaa'il 	&tegami-o & kaita \\
	&	\emph{wrote} & \emph{letters} & 	\emph{letter} & \emph{wrote} \\ \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    sadiiq       &	tomodachi & ni \\ % G3, G4
	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&kaana    &    sadiiq         &	tomodachi & datta \\ % not in Greenberg
	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak & udesho \\ % G16, G13 TODO this should only be inflected auxiliaries, replace Arabic
	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& sadiiq  &    John    & 	John no & tomodachi \\ % G2, G23
	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uha       & 	anata-ga yonda & hon \\ % G24
	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tusil        & 	toochaku suru & koto \\ % 
	&	\emph{that} & \emph{she arrives}  &	\emph{has.arrived} & \emph{that} \\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'ila lmadrasa     & 	gakkoo ni & ittekimashita \\ % G22
	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'uriid    &    'an 'ughaadir         & 	yom & itai \\ % G15
	& \emph{wants}   &  \emph{to leave}  &	\emph{to leave} & \emph{want} \\ \hline \hline
\end{tabular}
	\end{center}
	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples: Across the world, the orders of different constituents are strikingly correlated with that of verb and object.
	Selection is based on a more recent typological study by Dryer~\cite{dryer1992greenbergian}, restricted to those correlations that are annotated in available corpus data. See SI Section S1 for more on Greenberg correlations.
	}\label{table:corr-dryer}
\end{table}

Supported by languages on all continents, these correlations are among the language universals with the strongest empirical support.
Importantly, their validity is also independent from specific assumptions about theories of grammar.

Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study~\cite{lehmann1973structural, vennemann1974theoretical,jackendoff1977x,frazier1985syntactic,chomsky1988language, dryer1992greenbergian, hawkins1994performance}.
Prominent among this research is the argument that
%A prominent line of research has argued that 
language universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,pinker1990natural,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.

In an early and  influential functional framework,
Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between two pressures: to reduce complexity and to reduce ambiguity.
%what he called the Forces of \emph{Unification} and %\emph{Diversification}.
What Zipf called the Force of Unification is a pressure to reduce the complexity of the language to make production and processing as easy as possible.
His Force of Diversification favors languages that provide different utterances for different meanings, so that the listener can unambiguously identify the meaning from the utterance.
These two forces act in opposing directions:
Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required  to provide enough information.
The idea that many properties of language arise from the tension between these two pressures has a long and fruitful history in linguistics~\cite{gabelentz1901sprachwissenschaft,horn1984toward, lindblom1990explaining, schwartz1997dispersion, hawkins2004efficiency, haspelmath2006against}.

Recent work has drawn on information theory  to computationally test this ``dual pressures'' idea in various domains of language, showing that it predicts
both basic statistical properties of languages~\cite{ferreri2003least} and language development~\cite{smith2013linguistic},
and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.
While it has been suggested that the dual pressure  should also apply to grammar~\cite{hawkins2004efficiency}, testing these accounts is more difficult, as this requires  large amounts of data representative of language use across languages,  computational methods for estimating the efficiency of entire languages, and a simulation methodology for comparing different possible grammars.

In this work, we address these challenges by combining large-scale text data from 51 languages with machine learning techniques to estimate both aspects of the communicative efficiency of grammar: complexity and ambiguity.
We use machine learning models based on neural networks to model the evolution of grammars towards efficiency.
We apply this approach to the problem of explaining Greenberg's word order correlation universals.

In Study 1 we compare the word order of actual grammars of 51 languages with alternative ``counterfactual'' grammars parameterized by different word orders. We use our model to measure the communicative efficiency of each possible grammar, showing  that the grammars of real languages are more efficient than alternative  grammars. The fact that real grammars lie at the Pareto frontier of the efficiency space of possible  grammars suggests that the word order of languages has evolved to optimize communicative efficiency.

In Study 2 we test whether efficiency optimization accounts for the Greenbergian word order correlations.  For each of the 51 languages, we create hypothetical grammars optimized for efficiency.  We then test statistically whether these optimized grammars exhibit the Greenberg correlations, using a Bayesian mixed-effects logistic regression to control for language and language family.  Efficiency optimization indeed predicts all eight Greenberg collections.
%
%and makes fine-grained per-language predictions about individual syntactic patterns.
Our results show how general properties of efficient communication give rise to these universal word order properties of human language.

%\section*{Paradigm:   Random Grammars}
%
%In order to test whether

%idea of counterfactual grammars.  hypothetical grammars with different orders, so can study properties.

%\begin{figure}[ht]
%    \centering
%    \includegraphics[width=0.45\textwidth]{figures/communication-two-langs.png}
%	\caption{Grammars define cAommunication channels: A speaker (left) expresses a syntactic structure into a string of words forming a sentence. The order of words is determined by the grammar of the language. A listener recovers a syntactic structure. Communication is successful if the listener recovers the correct depenency structure.}
%	\label{fig:comm}
%\end{figure}

%We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
%In this model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
%Applying this model to word order, we take the message to consist of people, objects, events, and a set of syntactic and semantic relations that hold between them~\cite{fodor1975language,jackendoff1983semantics,pinker1990natural}.

%TODO also need to mention Principle of Compositionality, as additional justification for why listener needs to recover syntactic structures

%Consider Figure~\ref{fig:comm}.
%Imagine the speaker wants to communicate that Mary saw John.
%The message (left) consists of a syntactic structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
%When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
%In human language, this order is largely determined by the \key{grammar} of the language (CITE).
%The speaker will order it as \emph{Mary saw John} (Figure~\ref{fig:comm}, top) in English, and as \emph{Mary John saw} (Figure~\ref{fig:comm}, bottom) in Japanese.
%The listener receives a sentence -- a linear string of words -- and decodes the underlying message in accordance with the grammar of the language (Figure~\ref{fig:comm}, right).

\section*{Grammars and Grammar Data}

Following a long tradition in theoretical and computational linguistics, we formalize  the grammatical structure of languages using dependency trees~\cite{hays1964dependency,melcuk1988dependency,corbett1993heads,tesniere2015elements}.
%This linguistic formalism draws directed arcs between syntactically related words, annotated with grammatical relations like subject or object.
This linguistic formalism represents grammatical dependencies as directed arcs between syntactically related words, annotated with grammatical relations like subject or object (Figure~\ref{fig:sent-dep}).
While syntactic formalisms vary, the dependency grammar community has an agreed representation format  which has been used to  annotate corpora of text from dozens of languages \cite{ud2.1}, and there are 
computational methods for mapping syntactic structures in this representation to other standard linguistic formalisms \cite{boston2009dependency}.

\begin{figure}[ht]
    \centering
{
\centering{
\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
          She \& met \& friends  \\
   \end{deptext}
        %   \deproot{3}{ROOT}
   \depedge{2}{1}{subject}
   \depedge{2}{3}{object}
   %\depedge[arc angle=50]{7}{6}{ATT}
\end{dependency}
}
}
	\caption{An English sentence with annotated syntactic relations.}
	\label{fig:sent-dep}
\end{figure}

Our models require a sample of  syntactic structures as actually used by speakers across different languages, for which we draw on the recent
Universal Dependencies project~\cite{ud2.1}, which has collected and created syntactic annotations for several dozens of different languages.
51 languages had sufficient data for our purposes. 
These corpora represent a typologically and genetically diverse group of languages. %  (see Figure~\ref{fig:per-lang})
We obtained a total of 11.7M words in 700K sentences annotated with syntactic structures; with a median of 117K words and 7K sentences for each individual language.

\section*{Study 1: Relative efficiency of languages}
\label{sec:relative-efficiency}

We first ask whether the grammars of human languages evolve towards optimizing efficiency of communication.  To do this we
%We first demonstrate that the grammars of real languages are relatively efficient compared to random baselines, providing evidence that grammars evolve towards efficiency.
compare the efficiency of the actual grammars of the 51 languages from the Universal Dependencies datasets to randomly constructed baseline grammars.

The grammars of natural languages specify how the different words in a syntactic structure are ordered into a sentence, i.e., a string of words  \cite{adger2015syntax}.
%Grammars do this largely by providing consistent ordering rules for all syntactic relations (such as subjects and objects).
This is illustrated in Figure~\ref{fig:grammars}:
We show how four different grammars order objects, adpositional phrases, and adpositions.
For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
%Grammars 2 and 4 are rare or impossible.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.48\textwidth]{figures/grammar-langs-fontsize-friends.png}
	\caption{Grammars define consistent ordering rules for syntactic structures. Here, Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb. Grammars 1 and 3 conform to the Greenberg correlations and are common around the world; Grammars 2 and 4 are rare or impossible.}
	\label{fig:grammars}
\end{figure}

Beyond the syntactic relations exemplified in Figure~\ref{fig:grammars}, human languages have further types of syntactic relations.
The Universal Dependencies project, the source of our data, defines a total of 38 syntactic relations.
We adopt the grammar model of Gildea and colleagues~\cite{gildea2007optimizing, gildea2010grammars, gildea2015human}; a grammar assigns a weight from $[-1,1]$ to each of these 38 syntactic relations, and orders words according to the weights assigned to their relations
%Words that are syntactic dependents of the same word are ordered according to the weights assigned to the syntactic relations that link each of them to that word 
(See \textit{Materials and Methods} for details).

Given a database of syntactic structures (such as those at the top of Figure~\ref{fig:grammars}), obtained from a corpus of some real language L , we can apply a grammar to order the structures in the database into a dataset of counterfactual sentences belonging to a hypothetical language defined by that grammar (Figure~\ref{fig:grammars}).
This hypothetical language has identical syntactic structures and grammatical relations as the true language L, but different word order.

We create baseline grammars by randomly sampling the weights for each syntactic relation.
These baseline grammars have systematic word order rules similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations.
All four grammars in Figure~\ref{fig:grammars} are equally likely under this baseline distribution.

%For a language L, we obtain counterfactual versions

%Each of these random grammars for a language L has the identical syntactic structure and grammatical relations as the true grammar of L, but have different word orders.

%We therefore built a minigrammar for each language L with the same syntactic structures but with different orderings.

For every one of the 51 languages, we construct 50 counterfactual versions by randomly creating 50 baseline grammars, applying these to obtain counterfactual orderings for all syntactic structures that were available for that language.
%Each grammar specifies a weight from [-1,1] to each of these 38 syntactic relations.
%Words that are syntactically related to the same word are ordered according to these weights \cite{gildea2007optimizing}.

Following the information-theoretical literature on language processing, 
we formalize the  efficiency of a language as a weighted combination of two terms:  the 
amount of information that utterances contain about the underlying messages, and the  cost or difficulty of communication~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
We model the informativity term as the 
 degree to which listeners can reconstruct syntactic structures from an utterance, i.e., the \key{parseability} of the language. We model the cost or complexity term as the 
 \key{predictability}, or negative entropy, of the utterances. %, since entropy is a standard measure of the complexity of any message.
 We then estimate predictability and parseability using standard neural network models, and combine them linearly to produce a single model for efficiency.
 See \textit{Materials and Methods} for details.

For every one of the 51 languages, we computationally construct grammars that are \emph{optimized} for parseability, predictability, and efficiency (see \textit{Materials and Methods}).
This optimization problem is challenging because both the parseability and predictability of a sentence can only evaluated globally, in the context of an entire language.
We address this challenge by using neural network methods to estimate parseability and predictability, and by introducing a simple, differentiable computational formalism for describing grammatical regularities. 
This setup means that it is possible to find optimal grammars by standard methods such as stochastic gradient descent (see SI section S5).
For each grammar, we report predictability and parseability as estimated on the data resulting from ordering the syntactic structures from the corpus according to the grammar.

\begin{figure}
    \centering
    \includegraphics[scale=.5]{../results/plane/pareto-plane-iso-best-balanced-legend-viz-10-fontsize.pdf} % mention how many independent N's there are?
    \caption{Predictability and parseability of 51 languages (red), indicated by ISO codes, compared to baseline word order grammars (blue distribution). Predictability and parseability scores are $z$-scored within language. The gray curve indicates the Pareto frontier of computationally optimized grammars, averaged over the 51 languages.} 
    \label{fig:pareto-plane}
\end{figure}

In Figure~\ref{fig:pareto-plane}, we plot predictability and parseability of the grammars of 51 languages, together with the distribution of random baseline grammars, and the Pareto frontier defined by computationally optimized grammars.
Languages are attracted towards the Pareto frontier and away from the region of the baseline languages.
The majority of real languages are above and to the right of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
100~\% of languages improve on either predictability or parseability ($p<0.05$, by one-sided $t$-test, with Bonferroni correction and Hochberg's step-up procedure \cite{hochberg1988sharper}).
90~\% of languages improve over the baselines in parseability ($p < 0.05$), 80~\% improve in predictability ($p < 0.05$).
See SI Section S3 for additional analyses.

\section*{Study 2: Greenberg Word Order Correlations}
\label{sec:greenberg}

%\begin{figure} 
%	\begin{center}
%	\includegraphics[width=0.32\textwidth]{../results/correlations/figures/pred-eff-families-2.pdf}
%	\end{center}
%	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right). Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...). Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).}\label{fig:per-lang}
%\end{figure}

We have found that the grammars of human languages concentrate along the Pareto frontier of parseability and predictability.
This does not yet, however, say anything about what grammatical properties characterize Pareto-optimal languages in general, or which properties of human languages make them efficient.

We show that all languages close to the Pareto frontier -- both real and counterfactual ones -- are highly likely to satisfy Greenberg's correlation universals.
That is, optimizing for efficiency produces languages that satisfy these correlations.
In contrast, the baseline grammars are constructed without any correlations between the ordering of different syntactic relations, and will therefore not satisfy any of those universals.

We first considered this question for the 51 real languages.
Among the real grammars of the 51 languages, the number of satisfied correlations is strongly correlated with efficiency ($\rho = 0.61$, $p<0.0001$). %, and also with parseability ($\rho = 0.57$, $p<0.0001$) and predictability ($\rho =0.33$, $p = 0.024$) individually.
%(Parseability and predictability were $z$-scored within language as in Figure~\ref{fig:pareto-plane}. \textcolor{red}{TODO})

We next examine those grammars from Study 1 that we had computationally optimized for efficiency.
We controlled for variation across different optima by creating eight optimized grammars for each of the 51 datasets of syntactic structures from real languages.
For each language, we created four grammars with verb-object order, and four object-verb grammars.
We test whether the process of efficiency optimization produces the Greenberg correlations.

For each grammar (baseline, optimized, and real), we computed how many of the eight relations in Table~\ref{table:corr-dryer} had the same order as  Japanese (in contrast to Arabic).
Figure~\ref{fig:joint} shows the results, separately for grammars with verb-object and object-verb orders.
In optimized grammars, the order of the eight relations is strongly correlated with the placement of the object, similar to the 51 real languages in our sample.
In contrast, baseline languages show no correlation.

\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{../results/correlations/figures/correlations-curve-whiteaxis.pdf}
    %{figures/joint-results-corrs-curves.png} % ~/grammar-optim/results/correlations/figures/correlations-curve.pdf
    \caption{
    Efficiency optimization produces grammars where the orders of the eight relations in Table~\ref{table:corr-dryer} are strongly correlated with the order of verb and object. We arrange grammars (baseline, optimized, real) by the number of relations where the language patterns with Japanese (as opposed to with Arabic), and plot a kernel density estimate. Object-Verb order leads to grammars where object patterners precede (like Japanese); verb-object order leads to verb patterners preceding (like Arabic). Baseline grammars show no such correlation.
    }
    \label{fig:joint}
\end{figure}

%We compare grammars optimized for efficiency with the actual grammars of the 51 languages, as extracted from the orderings attested in the original corpora and validated against expert judgments (see \textit{Materials and Methods}).

So far, our results show that languages are efficient in part because they mostly satisfy the Greenberg correlations.
We asked whether efficiency optimization predicts every one of the eight correlations to hold in most languages.
We constructed a Bayesian multivariate logistic regression model predicting which of the eight correlations an optimized grammar satisfies.
%We first evaluated which of the eight correlations emerge from efficiency optimization across language families.
We controlled for variation between the syntactic structures used in different languages and language families by entering the language and language family as random effects.
See SI Section S4.3 for details.

\begin{table}
	\begin{center}
\begin{tabular}{|c|ll|c|cc|ccc}
	\hline
	%&
	&	\multicolumn{2}{c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Optimized} & \\ 
	&	verb & object     & & &   \\ 
	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
	&		\emph{to}            & \emph{a friend} &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
	&	\emph{has}          & \emph{written}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
	&	\emph{books} & \emph{that you read}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
	&	\emph{went} & \emph{to school}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
 \hline
\end{tabular}
	\end{center}
\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
For each correlation, we provide the prevalence among the actual grammars of the 51 languages (left), and the posterior distribution of the prevalence among grammars optimized for efficiency (right).
Efficiency optimization predicts all eight correlations to hold in the majority of grammars. \textcolor{red}{make visually clear that it's 51 languages}
}\label{table:corr-resu}
\end{table}

In Table~\ref{table:corr-resu}, we compare the prevalence of the eight correlations in real and optimized languages.
For the real languages, we indicate how many of the 51 languages satisfy a correlation.
For the optimized languages, we indicate the posterior distribution of the proportion of satisfying languages, obtained from the mixed-effects analysis. % (see \textit{Materials and Methods}).
Grammars optimized for efficiency accurately predict all eight correlations to hold at prevalences significantly greater than 50\%, similar to actual human languages.
In the multivariate mixed-effects analysis, efficiency optimization predicts all eight correlations to hold across languages (posterior probability $0.9911$). 
Optimizing for only predicability or only parseability does not predict all of the correlations (See SI Section S4.4).

%We confirmed these results by evaluating predictions of efficiency optimization on the level of individual languages (Figure~\ref{fig:per-lang}).
%For every one of the 51 languages, we show orders predicted by efficiency, for the eight correlating pairs.
%For each language, we consider the optimized grammar that had the best efficiency value, among those with the direction of the verb-object relation fixed to correspond to the actual language.
%Predictions of efficiency optimization agree with the actual word order patterns in 82~\% of the cases.
%Note that perfect agreement is not to be expected, as grammars will be impacted by factors other than efficiency optimization, such as their historical development.
%For instance, Dutch (nld) and Afrikaans (afr) have a prevalence of object-verb order in corpus data, and efficiency optimization accordingly predicts them to have corresponding orders (light blue).
%However, Dutch and Afrikaans have orders similar to the other Germanic languages (dark blue), which they are historically closely related to.

%We have directly implemented parseability and predictability using general-purpose statistical models of language and syntactic structures.
%Prior work has considered Dependency Length Minimization (DLM) as an approximate formalization of efficiency optimization~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
%This is the idea that languages order words so as to minimize the overall length of syntactic relations.
%This principle is a heuristic formalization of the idea that long syntactic relations should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
%Indeed, directly modeling efficiency optimization predicts DLM, and appears to predict word order universals with greater accuracy than optimizing for DLM (see  \emph{Relation to Dependency Length Minimization}).

\section*{Discussion}

We found that the grammars of natural languages are more efficient than baseline grammars, and that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication. %, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
%\textcolor{red}{[TODO adapt the next sentence to framing]} %We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals.

Our work makes crucial use of neural network models for estimating the efficiency of languages.
This method currently requires large computational resources; it still takes about three weeks to create optimized grammars for 51 languages, even with specialized hardware.
We believe that further advances in machine learning will reduce the computational cost, making this approach more widely applicable.

%This is the idea that languages order words so as to minimize the overall length of syntactic relations.
%This principle is a heuristic formalization of the idea that long syntactic relations should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.

What makes the grammars of human languages efficient?
Study 2 shows that Greenberg's correlations are one key property that real languages share with optimal grammars.
Prior work has suggested \emph{Dependency Length Minimization} as another characteristic of efficient word order.
This is the idea that word order minimizes the average distance between syntactically related words.
%Prior work has suggested some other properties as characteristics of efficient word order.
%One such property is Dependency Length Minimization, the idea that word order minimizes the distance between syntactically related words.
It is known that human languages reduce this distance compared to random baselines~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
Indeed, 100\% of grammars optimized for efficiency also reduce average distance between related words  compared to baselines ($p < 0.05$, by one-sided $t$-test).
To some extent, the Greenberg correlations help reduce the distance between related words. Consider again the sentence `I wrote letters to friends' (cf. Figures~\ref{fig:corr-table} and \ref{fig:grammars}). Both real and optimized grammars of English linearize its syntactic structure as follows:
\begin{center}
\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
          I \& wrote \& letters \& to \& friends  \\
   \end{deptext}
        %   \deproot{3}{ROOT}
   \depedge{2}{1}{subject}
   \depedge{2}{3}{object}
   \depedge{2}{4}{oblique}
   \depedge{4}{5}{case}
   %\depedge[arc angle=50]{7}{6}{ATT}
\end{dependency}
\end{center}
Note that this ordering exhibits correlations (1) and (7) from Table~\ref{table:corr-dryer}.
Among all possible ways of ordering this syntactic structure, this ordering also minimizes the average distance between any two syntactically related words: E.g., inverting `to' and `friends' would increase the distance between `wrote' and `to'.
%\textcolor{red}{I'm thinking about removing this part, I think it's a bit close to giving away the main result of another paper.}
%\textcolor{orange}{A second suggested property of efficient word order is Information Locality~\cite{gildea2015human,futrell2017noisycontext}, the idea that words that predict each other should be close.
%94~\% of the real grammars, and 90~\% of optimized grammars have significantly higher degrees of information locality than baseline grammars (at $p < 0.05$, one-sided $t$-test).}
Understanding mathematically why dependency length minimization  improves efficiency is an important problem for future research.
%and information locality

%\textcolor{red}{TODO revise this and make this about simplicity of the grammar, also harmonize with succeeding paragraph}
An idea related to, but distinct from, functional optimization is the idea that grammars are biased towards simplicity \cite{smith2013linguistic}. %\textcolor{red}{(TODO what to CITE here)} 
It has been proposed that languages have a single head-directionality parameter and that this accounts for the Greenberg correlations \cite{chomsky1988language}.
%We note that, for the eight syntactic relations in Table~\ref{table:corr-dryer}, the element patterning with the verb is often classified as the \emph{head} of the relation in syntactic theories.
%One way to capture the generalization would have been to stipulate that language has a single head directionality parameter, and assume that languages have identical head direction across different syntactic relations.
As an explanation of correlations, this idea makes strong predictions, and turns out to overpredict correlations \cite{dryer1992greenbergian}.
Nevertheless, future research should examine whether there are more principled connections between communicative efficiency and grammar simplicity.
%Our results, however, show that the tendency towards consistent head directionality is explainable from communicative principles.

A major question for functional explanations for linguistic universals is: \emph{how} do languages end up optimized? Do speakers actively seek out new communicative conventions that allow better efficiency? Or do languages change in response to biases that come into play during language acquisition \cite{fedzechkina2012language,culbertson2012learning}? Our work is neutral toward such questions. To the extent that language universals arise from biases in learning or in the representational capacity of the human brain, our results suggest that those biases tilt toward communicative efficiency. Our work does provide evidence against the idea that word order universals are best explained in terms of learning biases that are irreducibly arbitrary and genetic in nature \cite{chomsky2010some}.

While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into trees of syntactic relations. The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication.

%\textcolor{red}{TODO unify the next two paragraphs}
Beyond those universals that we examined, Greenberg~\cite{greenberg1963universals} described other universals; some of those are syntactic and could be tested with more specific annotation of datasets; others are not syntactic and are beyond the scope of this study.
Beyond Greenberg's work, there are remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
Our models do not capture this trade-off because the grammar model assumes fixed word order.
Future work can investigate whether these and other remaining universals can be explained using more sophisticated models of how syntactic structures are transduced into strings of words, based on larger databases with richer annotations.

These questions notwithstanding, our work provides strong evidence that the grammatical structure of languages is shaped by the need to support efficient communication.
Beyond our present results, we provide a computational framework in which theories of the efficiency optimization of languages can be tested.
While our study has focused on syntax, our results suggest that this method can be fruitfully applied to testing efficiency explanations in other domains of language structure.
%Despite to date only syntax, ...
%Other objective functions could be hypothesized and tested within our framework
%Furthermore, future advances in machine learning will enable the optimization of richer and richer models of grammar.

\matmethods{
\unskip
\subsection*{Corpus Data}
We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{ud2.1}. % \footnote{\url{http://universaldependencies.org/}}
We use all languages for which at least one treebank with a training partition was available, a total of 51 languages.
For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.
Punctuation was removed.

%Universal dependencies deviates from standard syntactic theory in the representation of certain syntactic relations.
Universal dependencies represents as dependents some words that are typically classified as heads in syntactic theory.
This particularly applies to the \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark} dependencies.
Following prior work studying dependency length minimization \cite{futrell2015largescale}, we applied automated conversion to a more standard formalism, modifying each treebank by inverting these dependencies, and promoting the dependent to the head position.
%We report results on this modified version of UD.

\subsection*{Word Order Grammars}
We adapt the grammar model of~\cite{gildea2007optimizing} to UD.
A grammar assigns a parameter $x_\tau \in [-1,1]$ to every relation  $\tau$ belonging to the 38 universal syntactic relations defined by UD.
A syntactic structure, consisting of a set of words and syntactic relations between them, is then ordered into a string of words recursively starting from the root; the dependents of a word then are ordered around the head according to the values $x_\tau$ corresponding to their syntactic relations; those dependents where $x_\tau < 0$ are ordered before the head; the others after the head.
See SI Section S5.2 for the methodology used to extract the languages' actual grammars from datasets, and for validation against expert judgments.

\subsection*{Formalizing Efficiency}
We adopt the formalization of language efficiency of \cite{ferreri2003least}, equivalent to a deterministic version of the Information Bottleneck \cite{zaslavsky2018efficient}.
Very similar formalizations of Zipf's ideas have been proposed across the information-theoretic literature on language~\cite{frank2012predicting,kemp2012kinship,regier2015word}. 
See SI Section S2 for discussion.
%The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
% both information theory~\cite{shannon1948mathematical}, and
%Following prior computational studies of functional explanations in linguistics~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge},

In this framework, the overall efficiency of language is a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication  \cite{ferreri2003least,frank2012predicting,kemp2012kinship,regier2015word,zaslavsky2018efficient}. We model the first term as the 
 degree to which listeners can reconstruct syntactic structures from an utterance, i.e., the \key{parseability} if the language.
 This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
\begin{equation}
	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
\end{equation}
where the sum runs over all possible pairs of utterances $u$ and syntactic structures $t$ in the language.
%This quantity describes the degree to which syntactic structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).

Again following \cite{ferreri2003least}, we formalize the complexity of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
This corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
%and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
\begin{equation}
	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
\end{equation}
where the sum runs over all possible sentences $u$ that belong to the language.
%In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.

%TODO
%An efficient language will balance these pressures by maximizing 
Maximizing one of the two scoring functions under a constraint on the other function (e.g., maximizing parseability under a constraint on the minimal predictability) amounts to maximizing a weighted combination of the two scoring functions~\cite{ferreri2003least,zaslavsky2018efficient}:
\begin{equation}\label{eq:efficiency}
	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
\end{equation}
with an interpolation weight $\lambda \in [0,1)$ that controls the relative strength of the two pressures.
When optimizing grammars for efficiency, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency} in order to give approximately equal weight to both components.
See SI section S2.2 for mathematical discussion of $\lambda$, and robustness to other choices.

%\subsection*{Measuring Efficiency and Optimizing Grammars}
%To estimate predictability and parseability we draw on recent machine learning methods in natural language processing.
We estimate predictability using LSTM recurrent neural networks~\citep{hochreiter1997long}, general sequence models that are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
We estimate parseability using a generic neural network architecture that casts recovery of syntactic structures as a minimum spanning tree problem \citep{dozat2017stanford, zhang2017dependency}.
All parseability and predictability values are reported on the held-out (\emph{dev}) partitions from the predefined split for each UD corpus.
Grammars are optimized for efficiency by simultaneous gradient descent on the parameters of the grammar and these neural models.
See SI Sections S5-S7 for details and for robustness of our results to these choices.

%\subsection*{Mixed-Effects Analysis in Study 2}
%We modeled the probabilities $p_{i,j}$ that a grammar $g_i$ satisfies the $j$-th correlation ($j=1,...,8$) using a multilevel logistic model \textcolor{red}{TODO (CITE)}, with random intercepts for the language for whose data the grammar had been optimized, and its language family, annotated according to \url{http://universaldependencies.org/}..
%\textcolor{red}{TODO some more details here}
%See SI Section S4.3 for details.

%\textcolor{red}{how much detail necessary here?}
%We modeled the probability $p_{i,j}$ that a grammar $g_i$ satisfies the $j$-th correlation ($j=1,...,8$) as
%$$logit(p_{i,j}) = \alpha_j + \sigma_{l(i),j} + \tau_{f(i),j}$$
%where $l(i)$ is the language for whose data $g_i$ was optimized, and $f(i)$ is its language family.
%We used a flat uninformative prior for $\alpha_j$.
%TODO add the priors, CITE BRMS

%

%See the SI for details on the neural language models and parsers used, and on the optimization procedures.

%\paragraph{Information Locality}
%We quantified information locality by the mutual information between adjacent words, which is high if words predicting each other are very close together.
%We estimated  the mutual information as the difference between the cross-entropies of a unigram and a bigram language model.
%We applied Kneser-Ney smoothing to the bigram model.
%Counts were constructed on the training partitions; cross-entropies were evaluated on the held-out partitions.

%As a simple measure of information locality, we computed the mutual information between adjacent words, which is high if words predicting each other are very close together.
%By this measure,

\subsection*{Data Availability}
The efficiency optimization results from Table~\ref{table:corr-resu} were preregistered: \url{ http://aspredicted.org/blind.php?x=ya4qf8}.

Code and results are available at \url{https://github.com/m-hahn/grammar-optim}.

}

\showmatmethods{} % Display the Materials and Methods section

\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Paul Kiparsky, and audiences at CAMP 2018 and CUNY 2019 for helpful discussion.}

\showacknow{} % Display the acknowledgments section

%\bibliographystyle{acl_natbib}
%\bibliography{extracted}
\bibliography{everything}

\end{document}

%Ray Jackendoff. 2002. Foundations of Language:
%Brain, Meaning, Grammar, Evolution. Oxford
%University Press, Oxford, UK

%William Croft and Alan Cruse. 2004. Cognitive
%Linguistics. Cambridge University Press, Cam-
%bridge, UK.
% --> emphasize language use

%Adele Goldberg. 2005. Constructions at work:
%The nature of generalization in language. Ox-
%ford University Press, Oxford, UK.
% --> emphasize usage-based grammar
% we are all born with the same basic conceptual apparatus, with the
%same basic communicative demands, and we all live in the physical world with
%forces of gravity, bodies, and night and day (cf. also LakoV 1987; Webelhuth and
%Ackerman 1998). Certain semantic/pragmatic functions are so relevant and
%useful that they can be found in language after language (Croft 2001).
%Cross-linguistic generalizations that relate form and function are explained
%by appeal to general cognitive constraints together with the functions of the
%constructions involved. Constructionists turn to grammar-external explan-
%ations such as universal functional pressures, iconic principles, and processing
%and learning constraints to explain such empirically observable cross-linguis-
%tic generalizations.
%For example, the fact
%that that most languages seem to have noun and verb categories may be
%explained by the existence of corresponding basic semantic categories (Baker
%2004). Hauser, Chomsky, and Fitch go so far as to suggest that the only innate
%ability speciWc to language that may be required are the recursive interfaces to
%phonology and semantics, and they allow that even these may turn out not to
%be speciWc to language (Hauser, Chomsky, and Fitch 2002).

% Haspelmath 1999 talks about types of explanations

%Joan Bresnan, Ash Asudeh, Ida Toivonen, and
%Stephen Wechsler. 2016. Lexical-Functional
%Syntax, 2nd ed. Blackwell, Malden, MA.

%Andrew Radford. 2006. Minimalist syntax re-
%visited. http://www.public.asu.edu/
%~gelderen/Radford2009.pdf.

%Givon

% TODO optimality theory and syntax citation?
%Blutner R. (2000) Some aspects of optimality in natural language interpretation. Journal of Semantics 17: 189–216

% ,lin2017critical

 % \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
% behaghel1909beziehungen,
% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
% TODO mention universals here

%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only

% TODO change wording so it's clear that this is a very major thing

% Things TODO
% - make clear that efficiency is something both functional and generative linguists have been appealing to
% - at some point address head direction idea
%TODO at some point address head direction idea
%    - Vennemann 1972, 1974ab, 1976
%    -Hawkins 1980 limitations of head direction, also Hawkins 1982

%foreshadow functional explanations?
%
%- these linguistic patterns reflect general principles of communication
%
%- big idea has been around for a long time
%
%- here for the first time direct computational test
%
%- we can predict broad patterns, and make language-specific predictions based on these general principles 
%

% TODO add a cogsci person to citation chain at the beginning

%\dropcap{F}or decades, researchers in fields ranging from philology to cognitive science to statistical physics have been involved in documenting and trying to explain the universal syntactic and statistical properties of human language \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical}. An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.

%Berwick and Chomsky ‘gene rative process  is optimal  ’, ‘efficient computa  tion ’(71), ‘this newly  emerged  computationalsystem forthought  ...ispe rfe ct, in sofar asSMT is corr ect ’(80).  ‘op -tim al’or ‘effici ent ’or ‘per fect ’.

%Chomsky Three Factors, 2003:
% --> emphasizes 'computational efficiency'

%\textbf{TODO maybe should be more general in the beginning, as in the first version. Also make sure the reader doesn't think 'it's just headedness, why bother'.}

%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.

\begin{table}
	\begin{center}
\begin{tabular}{|c|ll|c|cc|ccc}
	\hline
	%&
	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Optimized} & \\ 
	&	verb & object     & & &   \\ 
	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
	&		\emph{to}            & \emph{a friend} &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
	&	\emph{has}          & \emph{written}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
	&	\emph{books} & \emph{that you read}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
	&	\emph{went} & \emph{to school}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
 \hline
\end{tabular}
	\end{center}
\caption{Efficiency optimization accurately predicts the Greenbergian correlations.
For each correlation, we provide the prevalence among the actual grammars of the 51 languages (left), and the posterior distribution of the prevalence among grammars optimized for efficiency (right).
Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
}\label{table:corr-resu}
\end{table}

\begin{table}
	\begin{center}
\begin{tabular}{|c|cc||c|cc|ccccccc}
	\hline
	%&
	&        Real   &  Optimized &  & Real   &  Optimized  \\ \hline \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	
				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     }  \\
	&	 &  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & 	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     }\\
	&	 &  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     }\\
	&	 &  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     }\\
	&	 &  &&&\\ \hline
 \hline
\end{tabular}
	\end{center}
\caption{\textcolor{red}{choose this or the other version}Efficiency optimization accurately predicts the Greenbergian correlations.
For each of the eight correlations (Table~\ref{table:corr-dryer}), we provide the prevalence among the actual grammars of the 51 languages (left), and the posterior distribution of the prevalence among grammars optimized for efficiency (right).
Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
}\label{table:corr-resu}
\end{table}

### 2019-06-25 writeup/pnas-word-order-si.tex
code

@@
 \tableofcontents
-
-\textcolor{red}{TODOs}
-
-\textcolor{red}{run czech parser}
-
-\textcolor{red}{keep in mind: pareto plot unify with SI plot}
 
@@
 \begin{equation}\label{eq:eff-general}
-I[S, R] - \lambda H[S]
-\end{equation}
-where $I[S, R]$ describes the \key{informativity} of the signals $S$ about their referents $R$, and $H[S]$ describes the \key{complexity} of the communication system.
+I[S, R] - \lambda \cdot H[S]
+\end{equation}
+where $I[S, R]$ describes the \key{informativity} of the signals $S$ about their referents $R$, and $H[S]$ describes the \key{complexity} of the communication system, and $\lambda > 0$ trades off the two aspects of efficiency.
 While prior studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} mostly considered settings where the signals $S$ are individual words without further structure, the signals are entire sentences $\utterance$ in our setting.
@@
 %The idea that efficient languages will permit the listener to infer grammati
-
-We therefore arrive at the following efficiency metric:
+Hence, substituting utterances $\utterance$ for signals $S$, and syntactic structures $\tree$ for underlying messages $R$, into (\ref{eq:eff-general}), we  arrive at the following efficiency metric for word order:
 \begin{equation}\label{eq:efficiency-derived}
-	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda \cdot R_\textit{Pred}
 \end{equation}
@@
 
-We now discuss how this metric relates to other variants that have been proposed in the information-theoretic literature on language, and describe why~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is particularly appropriate to our setting.
+We now discuss how this metric relates to other variants that have been proposed in the information-theoretic literature on language, and describe why~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is most appropriate to our setting.
 This efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
@@
 
+Beyond the study of natural language, the eficiency metric (\ref{eq:eff-general}) is also closely related to information-theoretic models in other disciplines.
+The tradeoff between informativity and complexity of communication systems is studied by Rate-distortion theory \cite{shannon1948mathematical, berger1971rate}.
+The \emph{Infomax Principle} derives parsimonious data representations by maximizing the mutual information between data and representations, subject to constraints on the representations~\cite{linsker1988self}; a constraint on the representation entropy leads to a metric equivalent to (\ref{eq:eff-general}) \cite{friston2010free}.
+
 %\textcolor{red}{also mention \cite{piantadosi2012communicative,fedzechkina2012language}}
@@
 
-In the efficiency objective
+In the efficiency objective (\ref{eq:efficiency-derived})
 \begin{equation}\label{eq:efficiency}
@@
 In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parseability.
-The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
-However, we also examined word order predictions for grammars optimized for only parseability or only predictability, in order to tease apart predictions made by these two components.
+While the computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$, we also examined word order predictions for grammars optimized for only parseability or only predictability, in order to tease apart predictions made by these two components.
 As shown in Table~\ref{tab:all-predictions-1}, each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions.
@@
 
-Figure 4 in the main paper shows the average of these per-language plots.
+Figure 4 in the main paper shows the average of these per-language plots, with a kernel density estimate of the distribution of baseline grammars.
 
@@
 In Table~\ref{table:corr-resu-previous} we report the results of our two previous, preregistered, simulations\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}, \url{https://aspredicted.org/blind.php?x=bg35x7}.
-For the results of the locality simulations described in the first preregistration, see the Dependency Length Minimization results in Table~\ref{tab:all-predictions-1a}, with discussion in Section~\ref{sec:DLM}.} together with results from the main experiment.
+For the results of the Locality simulations described in the first preregistration, see the Dependency Length Minimization results in Table~\ref{tab:all-predictions-1b}, with discussion in Section~\ref{sec:DLM}.} together with results from the main experiment.
 These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) one simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
@@
 We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
-To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.\footnote{We implemented this regularized optimization as mean-field ELBO variational inference in Pyro~\cite{bingham2018pyro}.}
+To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
+We implemented this regularized optimization as mean-field ELBO variational inference in Pyro~\cite{bingham2018pyro}.
 We then extract the posterior means for each parameter $a_\tau, b_\tau$, and convert the resulting differentiable grammar into an ordinary ordering grammar.

### 2019-06-26 writeup/pnas-word-order-si.tex
code

--- 2019-06-21 writeup/pnas-word-order-si.tex
@@
     \centering
-     \includegraphics[width=0.5\textwidth]{../results/dependency-length/figures/depl-violin-all.pdf} 
+     \includegraphics[width=0.5\textwidth]{depl-violin-all-1.png} % ../results/dependency-length/figures/depl-violin-all.pdf 
         \caption{Average dependency length for grammars optimized to minimize dependency length (DLM, left), optimized for efficiency (second), the real orderings found in corpora (third), and random baseline grammars (right).
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{../results/dependency-length/figures/depLength-facet.pdf} 
+	\includegraphics[width=0.9\textwidth]{depLength-facet-1.png} %../results/dependency-length/figures/depLength-facet.pdf} 
         \caption{Total dependency length as a function of sentence length, for four diverse languages.

### 2019-06-26 writeup/compile.sh

@@ -17,3 +17,8 @@ bibtex pnas-word-order-si
 pdflatex pnas-word-order-si.tex
 pdflatex pnas-word-order-si.tex
 
+
+pdftoppm ../results/dependency-length/figures/depl-violin-all.pdf depl-violin-all -png
+pdftoppm ../results/dependency-length/figures/depLength-facet.pdf depLength-facet -png
+
+

### 2019-06-26 writeup/pnas-word-order-si.tex
code

@@
 While most of the other relations also correspond to UD relations, some are not annotated reliably.
-We were able formalize eleven out of Dryer's seventeen correlations in UD.
+We were able formalize eleven out of Dryer's sixteen correlations in UD.
 Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations:
@@
 \paragraph{Excluded Correlations}
-We excluded three correlations that are not annotated reliably in UD, and are only relevant to some of the world's languages: Question particles, plural words (i.e., independent plural markers), and articles.
+Here, we discuss in more detail the five correlations from Dryer's study that we had to exclude.
+First, we excluded three correlations that are not annotated reliably in UD, and are only relevant to some of the world's languages: Question particles, plural words (i.e., independent plural markers), and articles.
 All three types of elements occur at most in parts of the 51 UD languages, and none of them is annotated reliably in those languages where they occur.

### 2019-06-26 writeup/pnas-word-order-si.tex
code

@@
 	\end{center}
-\caption{Results from optimization experiments for different values of $\lambda$, including our two previous preregistered experiments.
+	\caption{Results from optimization experiments for different values of $\lambda$, including our two previous preregistered experiments (Section~\ref{sec:previous-exps}).
 For comparison, we also show results for $\lambda=0$, corresponding to optimizing for parseability only (same results as reported in Tables (\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})).

### 2019-06-26 writeup/pnas-word-order-si.tex
si

--- 2019-06-26 writeup/pnas-word-order-si.tex
@@
 % Use the lineno option to display guide line numbers if required.
+
+% pdftk grammar_optim_paper.pdf pnas-word-order-si.pdf cat output submission.pdf
 
@@
     \centering
-     \includegraphics[width=0.5\textwidth]{depl-violin-all-1.png} % ../results/dependency-length/figures/depl-violin-all.pdf 
+     \includegraphics[width=0.5\textwidth]{../results/dependency-length/figures/depl-violin-all.pdf} 
         \caption{Average dependency length for grammars optimized to minimize dependency length (DLM, left), optimized for efficiency (second), the real orderings found in corpora (third), and random baseline grammars (right).
@@
     \centering
-	\includegraphics[width=0.9\textwidth]{depLength-facet-1.png} %../results/dependency-length/figures/depLength-facet.pdf} 
+    \includegraphics[width=0.9\textwidth]{../results/dependency-length/figures/depLength-facet.pdf} 
         \caption{Total dependency length as a function of sentence length, for four diverse languages.

### 2019-06-27 writeup/pnas-word-order-si.tex
SI

--- 2019-06-26 writeup/pnas-word-order-si.tex
@@
 The tradeoff between informativity and complexity of communication systems is studied by Rate-distortion theory \cite{shannon1948mathematical, berger1971rate}.
-The \emph{Infomax Principle} derives parsimonious data representations by maximizing the mutual information between data and representations, subject to constraints on the representations~\cite{linsker1988self}; a constraint on the representation entropy leads to a metric equivalent to (\ref{eq:eff-general}) \cite{friston2010free}.
+The \emph{Infomax Principle} derives parsimonious data representations by maximizing the mutual information between data and representations, subject to constraints on the representations~\cite{linsker1988self}; a constraint on the representation entropy leads to a metric equivalent to (\ref{eq:eff-general}) and to a version of the Free-Energy principle (see Section S3 in \citet{friston2010free}).
 
@@
 
-\subsection{Possible values of $\lambda$}\label{sec:lambda}
+\subsection{Choice of $\lambda$}\label{sec:lambda}
 
@@
 
-Greater values of $\lambda$ can mathematically result in degenerate solutions.
+The reason is that greater values of $\lambda$ can mathematically result in degenerate solutions.
 To show this, note that the following inequality always holds:
@@
 Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
-\begin{equation}
-    R_{Eff} = I[\utterance; \tree] - \lambda H[\utterance] \leq 0.
-\end{equation}
+%\begin{equation}
+$R_{\textit{Eff}} = I[\utterance; \tree] - \lambda H[\utterance] \leq 0$,
+%\end{equation}
 and it takes the maximally possible value zero if there is only a single utterance $\utterance$, in which case both $I[\utterance; \tree]$ and $H[\utterance]$ are zero.
@@
 	\begin{center}	
-\includegraphics[width=0.8\textwidth]{si-table-perrel-2a-1.png}  
+\includegraphics[width=0.7\textwidth]{si-table-perrel-2a-1.png}  
 \end{center}
@@
 \subsection{Extracting Grammars from Datasets}\label{sec:extract-grammars}
-We extract grammars for the actual languages by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
+We extract grammars for each actual language by fitting a differentiable ordering grammar maximizing the likelihood of the observed orderings.
 To prevent overfitting, we regularize each $a_\tau$, $b_\tau$ with a simple Bayesian prior $logit(a_\tau) \sim \mathcal{N}(0,1)$, $b_\tau \sim \mathcal{N}(0,1)$.
@@
 \begin{equation}\label{eq:efficiency-recall}
-	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
+	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
 \end{equation}
@@
 \end{equation}
-where $t \sim \tree$ is a distribution over syntactic structures, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
+where $t \sim \tree$ is the distribution over syntactic structures as found in databases of the language, and $u \sim p_{\mathcal{L}_\theta}(u|t)$ denotes the corresponding linearized sentences.
 
@@
 \begin{equation}
-	R_{Pred}^{\psi}(u) = - \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})
+	R_{Pred}^{\psi}(u) = \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})
 \end{equation}
@@
 \begin{equation}
-R_{\textit{Eff}}^{\theta, \phi, \psi} := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+R_{\textit{Eff}}^{\theta, \phi, \psi} := \E_{t \sim \mathcal{T}} \E_{u \sim p_{\mathcal{L}_\theta}(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
 \end{equation}
@@
 \begin{align}
-\partial_\theta R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)} \left[  \left[\partial_\theta \log p_\theta(u|t)\right] \cdot    \left(R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right) \right] \label{eq:dtheta}\\
-\partial_\phi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\partial_\phi R_{Pars}^{\phi}(u,t)\right] \\
-\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_\theta(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}
+\partial_\theta R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_{\mathcal{L}_\theta}(u|t)} \left[  \left[\partial_\theta \log p_{\mathcal{L}_\theta}(u|t)\right] \cdot    \left(R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right) \right] \label{eq:dtheta}\\
+\partial_\phi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_{\mathcal{L}_\theta}(u|t)}  \left[\partial_\phi R_{Pars}^{\phi}(u,t)\right] \\
+\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_{\mathcal{L}_\theta}(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}
 \end{align}
@@
 We can therefore apply stochastic gradient descent to jointly optimize $\theta, \phi, \psi$:
-In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim p_{\theta}(\cdot|t)$.
+In each optimization step, we sample a dependency tree $t$ from the database, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim p_{\theta}(\cdot|t)$.
 Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
@@
 We used the CKY algorithm to compute the syntactic ambiguity $H[\tree|\utterance]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
-Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable for exact parsing of a simple PCFG.
+Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable than baseline grammars, for exact parsing of a simple PCFG.
 % TODO get UAS for compatibility with other parts of SI
@@
 \paragraph{Experiments}
-Using English and Japanese data, we trained parsers for the ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
+Using English and Japanese data, we trained parsers for ten random word order grammars and for the best grammar optimized for efficiency, with the input presented in each of the distorted orderings.
 Resulting parsing scores are shown in Figure~\ref{fig:distorted-parser}.
@@
 We estimated language models on the training partitions, and used the validation partitions to estimate surprisal.
-We conducted this for the ten random and the best optimized ordering grammars on English and Japanese data.
-Results (Figure~\ref{fig:bigrams}) show that languages optimized for efficiency are optimal for a bigram language model.
+We conducted this for ten random and the best optimized ordering grammars on English and Japanese data.
+Results (Figure~\ref{fig:bigrams}) show that languages optimized for efficiency are also optimal for a bigram language model.
 
@@
 A gap in parsing loss of about 0.2 nats appears already at 0.01 \% of the training data (2000 words), and persists for larger amounts of training data.
+This shows that the observed efficiency differences between grammars cannot be attributed to data sparsity.
 
@@
 }
-\caption{Languages with ISO codes, families (according to \url{https://universaldependencies.org/}), and the number of available sentence and words.}\label{tab:langs-iso-sizes}
+\caption{Languages with ISO codes, families (according to \url{https://universaldependencies.org/}), and the number of available sentences and words.}\label{tab:langs-iso-sizes}
 \end{table}

### 2019-06-27 writeup/pnas-word-order-si.tex
SI
si

@@
 \section{Formalizing Comunicative Efficiency}
-\subsection{Relation to Other Work}
+\subsection{Derivation and Relation to Other Work}
 
@@
 Here we discuss how our formalization of communicative efficiency relates to formalizations that have been proposed in the information-theoretic literature on language.
-Across the literature, the core idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms, with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global,kemp2012kinship,frank2012predicting,goodman2013knowledge,kao2014nonliteral,xu2014numeral,kirby2015compression,regier2015word,xu2016historical,futrell2017memory,zaslavsky2018efficient,bennett2018extremely, peloquin2019interactions,zaslavsky2019semantic}.
-
+Across the literature, the core idea is to maximize the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms:
+\begin{equation}\label{eq:eff-basic}
+\text{Informativity} - \lambda \cdot \text{Complexity}
+\end{equation}
+with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global,kemp2012kinship,frank2012predicting,goodman2013knowledge,kao2014nonliteral,xu2014numeral,kirby2015compression,regier2015word,xu2016historical,futrell2017memory,zaslavsky2018efficient,bennett2018extremely, peloquin2019interactions,zaslavsky2019semantic}.
+
+\paragraph{Derivation of our Formalization}
 The basis for our precise formalization is the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems.
@@
 \begin{equation}\label{eq:eff-general}
-I[S, R] - \lambda \cdot H[S]
-\end{equation}
-where $I[S, R]$ describes the \key{informativity} of the signals $S$ about their referents $R$, and $H[S]$ describes the \key{complexity} of the communication system, and $\lambda > 0$ trades off the two aspects of efficiency.
+\operatorname{I}[S, R] - \lambda \cdot \operatorname{H}[S]
+\end{equation}
+where $\operatorname{I}[S, R]$ describes the \key{informativity} of the signals $S$ about their referents $R$, and $\operatorname{H}[S]$ describes the \key{complexity} of the communication system, and $\lambda \geq 0$ trades off the two aspects of efficiency.
 While prior studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} mostly considered settings where the signals $S$ are individual words without further structure, the signals are entire sentences $\utterance$ in our setting.
@@
 \begin{equation}
-	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
 \end{equation}
@@
 \begin{equation}
-	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
-\end{equation}
-
+	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
+\end{equation}
+Parseability $\operatorname{I}[\utterance,\tree]$ is higher if utterances provide more information about their underlying syntactic structure.
+Due to the identity $\operatorname{I}[\utterance, \tree] = \operatorname{H}[\tree] - \operatorname{H}[\tree|\utterance]$, parseability is maximized if every utterance can be parsed unambiguously -- that is, if the listener's uncertainty about syntactic structures given received utterances, $\operatorname{H}[\tree|\utterance]$, is zero.
+Predictability $- \operatorname{H}[\utterance]$ is higher if the distribution over utterances is concentrated on a few utterances, and is maximized if there is just a single utterance.
+It is also equal to the negative average of surprisal, which is a strong and linear predictor of human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+
+\paragraph{Relation to Formalizations in Semantic Typology}
 We now discuss how this metric relates to other variants that have been proposed in the information-theoretic literature on language, and describe why~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is most appropriate to our setting.
-This efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
+This efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains by \citet{zaslavsky2018efficient} and \citet{zaslavsky2019semantic}.
 The Information Bottleneck models complexity using a mutual information term instead of the entropy appearing in~(\ref{eq:eff-general}).
 This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
-As we assume deterministic grammars that transduce every underlying syntactic structure into one surface order, our model corresponds to the Deterministic Information Bottleneck~\cite{strouse2017deterministic}, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure.
-Entropy also corresponds to the average of surprisal, which is a strong and linear predictor of human language processing effort~\cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+We assume deterministic grammars that transduce every underlying syntactic structure into one surface order; our model therefore corresponds to the deterministic version of the Information Bottleneck, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure~\cite{strouse2017deterministic}.
 
 A few other approaches share the mutual information term for informativity in (\ref{eq:eff-general}), while using complexity measures that are not explicitly information-theoretic.
-%A few models have used other complexity measures that are not explicitly information-theoretic quantities.
-In \cite{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \cite{kemp2012kinship} it is the complexity of defining the concepts that are encoded.
-The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
-Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the Information Bottleneck formalism \cite{zaslavsky2018efficient, zaslavsky2019semantic}, bringing them even closer to our formalization of efficiency.
-
-In addition to these models, which quantify the efficiency of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context, such as models of pragmatic reasoning in reference games \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
-In line with the other models, these assume that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely}.
-
+In studies of semantic typology by \citet{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \citet{kemp2012kinship} it is the complexity of defining the concepts that are encoded.
+%The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
+Notably, the models of \citet{regier2007color} and \citet{xu2016historical} have since been reformulated successfully in the Information Bottleneck formalism \cite{zaslavsky2018efficient, zaslavsky2019semantic}, bringing them even closer to our formalization of efficiency.
+
+\paragraph{Relation to Formalizations of Language Evolution}
+There are also related models of the evolution of languages.
+Most related to our work, \citet{kirby2015compression} describe a model of language evolution balancing the pressure for languages to be informative about the intended meanings, while minimizing complexity.
+Formally, their model studies a Bayesian language learner who infers a language $h$ from data $d$ according to $P(h|d) \propto P(d|h) P(h)$, where $P(h)$ defines a prior distribution over languages, and $P(d|h)$ is the likelihood of observed data $d$ under the grammar $h$, assuming that speakers produce utterances pragmatically.
+The prior $P(h)$ favors less complex languages; the likelihood $P(d|h)$ favors languages that communicate meanings unambiguously.
+We now show that this model instantiates the basic objective~(\ref{eq:eff-basic}). 
+If the dataset $d$ consists of observed pairs $(t,f)$ of meanings $t$ and forms $f$, and the language $h$ defines a set of possible pairs $(t,f)$, then the log-likelihood as defined by their model can be written as follows (up to constants)\footnote{We assume for simplicity that the error probability $\epsilon$ in the model is equal to $0$.}
+\begin{equation}
+\log P(d|h) = \sum_{(t,f) \in d} \log P(f|h,t) = \sum_{(t,f) \in d} \log P(f|h,t) \propto \sum_{(t,f) \in d} \log \frac{1}{|\{t' : (t', f) \in h\}|} = \sum_{(t,f) \in d} \log P(t|f)
+\end{equation}
+where $P(t|f)$ is the probability that the observed form $f$ referred to meaning $t$, as the model assumes uniform meaning distributions and uniform choice of appropriate forms.
+Replacing the sum over the dataset $d$ by the expectation over the idealized full distribution over meaning-form pairs, this can be rewritten as 
+\begin{equation}
+- \operatorname{H}[t|f] = \operatorname{I}[t, f] - \operatorname{H}[t]
+\end{equation}
+where the first term is the mutual information between forms and meanings, as in our efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}).
+The second term, the entropy of meanings, is a constant independent of the form-meaning mapping.
+The overall log probablity assigned by the Bayesian learner thus comes out to (up to constants)
+\begin{equation}
+\log P(h|d) = \operatorname{I}[t, f] + \lambda \log P(h)
+\end{equation}
+where the prior $P(h)$ favors simpler languages.
+This result shows that the model of \citet{kirby2015compression} predicts that language evolution favors languages optimizing a function of the form~(\ref{eq:eff-basic}), with a informativity term identical to that of our model~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}).
+
+\paragraph{Relation to Formalizations of Pragmatics}
+In addition to these models, which model the efficiency and evolution of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context.
+Our work is most closely related to the Rational Speech Acts model of pragmatic reasoning \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
+In line with the other models discussed here, it assumes that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely, peloquin2019interactions}.
+\citet{peloquin2019interactions} provide further discussion of the links between pragmatics and the efficiency metric (\ref{eq:eff-general}-\ref{eq:efficiency-derived}).
+
+\paragraph{Relation to Models in Other Disciplines}
 Beyond the study of natural language, the eficiency metric (\ref{eq:eff-general}) is also closely related to information-theoretic models in other disciplines.
@@
 %Other metrics very similar to~(\ref{eq:eff-general}) have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
-%The idea of formalizing informativity using the mutual information $I[S, R]$ is common to all these approaches.
+%The idea of formalizing informativity using the mutual information $\operatorname{I}[S, R]$ is common to all these approaches.
 
@@
 \begin{equation}
-I[\utterance; \tree] \leq H[\utterance].
+\operatorname{I}[\utterance; \tree] \leq \operatorname{H}[\utterance].
 \end{equation}
@@
 %\begin{equation}
-$R_{\textit{Eff}} = I[\utterance; \tree] - \lambda H[\utterance] \leq 0$,
-%\end{equation}
-and it takes the maximally possible value zero if there is only a single utterance $\utterance$, in which case both $I[\utterance; \tree]$ and $H[\utterance]$ are zero.
+$R_{\textit{Eff}} = \operatorname{I}[\utterance; \tree] - \lambda \operatorname{H}[\utterance] \leq 0$,
+%\end{equation}
+and it takes the maximally possible value zero if there is only a single utterance $\utterance$, in which case both $\operatorname{I}[\utterance; \tree]$ and $\operatorname{H}[\utterance]$ are zero.
 This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
@@
 %\begin{equation}
-%	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+%	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
 %\end{equation}
@@
 %\begin{equation}
-%	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
+%	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
 %\end{equation}
@@
 %    \label{eq:efficiency-again}
-%    J_T(L) = I[L(T); T] - \lambda H[L(T)].
+%    J_T(L) = \operatorname{I}[L(T); T] - \lambda \operatorname{H}[L(T)].
 %\end{equation}
@@
 %    \label{eq:ib}
-%    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
+%    J_{X,Y}(L) = \operatorname{I}[\hat{X}; Y] - \lambda \operatorname{I}[\hat{X}; X].
 %\end{equation}
@@
 %    \label{eq:efficiency-kinda-ib}
-%    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
-%\end{equation}
-%Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$.
+%    J_T(L) = \operatorname{I}[U;T] - \lambda \operatorname{I}[U;\hat{U}]. % TODO should this be \operatorname{I}[\hat{U};T]?
+%\end{equation}
+%Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $\operatorname{I}[U;\hat{U}]=\operatorname{I}[U;U] = \operatorname{H}[U]$.
 Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
 %\begin{equation*}
-%    J_T(L) = I[U;T] - \lambda H[U].
+%    J_T(L) = \operatorname{I}[U;T] - \lambda \operatorname{H}[U].
 %\end{equation*}
@@
 \begin{tabular}{l||ll|lll|lll}
-Language & Pred. (t) & Parse. (t) & \multicolumn{3}{|c|}{Pred. (Binomial)} & \multicolumn{3}{|c|}{Parseab. (Binomial)} \\ 
+Language & Pred. (t) & Parse. (t) & \multicolumn{3}{c|}{Pred. (Binomial)} & \multicolumn{3}{c}{Parseab. (Binomial)} \\ 
 &  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
@@
 
-We obtained the posterior density plots in Table 2 (Main Paper) and in Figure \ref{tab:all-predictions-1}) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\alpha_j$ (\ref{eq:mixed-effects}).
+We obtained the posterior density plots in Table 2 (Main Paper) and in Figure (\ref{tab:all-predictions-1}) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\alpha_j$ (\ref{eq:mixed-effects}).
 
@@
 We provide an example in Figure~\ref{fig:grammar-sample}, illustrating grammar parameters for the relations in Figure 3 of the main paper.
+Note that the grammatical formalism simplifies some aspects of the word order regularities of natural languages.
+For instance, it does not represent cases where ordering varies between main and embedded clauses, as it does not condition ordering decisions on the larger context.
+%Our method extracts the most frequent ordering and applies it throughout all 
+More complex and powerful ordering grammar models have been proposed \cite{futrell2015experiments, wang2016galactic}; however, they have similar limitations, and for our purposes, the model adopted here has the advantage of being simple and interpretable.
+%There currently is no grammatical formalism that is at the same time capable of representing the full range of word order regularities that are found, 
+%\cite{futrell2015experiments}
 
@@
 We validated the extracted grammars by comparing the dominant orders of six syntactic relations that are also annotated in the World Atlas of Linguistic Structures (WALS, \cite{haspelmath2005world}).
-Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS.
+Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS: adpositions, complementizers, relative clauses, genitives, and oblique PPs.
 In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb-object relation.
@@
 
-Note that the grammatical formalism simplifies some aspects of the word order regularities of natural languages.
-For instance, it does not represent cases where ordering varies between main and embedded clauses, as it does not condition ordering decisions on the larger context.
-%Our method extracts the most frequent ordering and applies it throughout all 
-More complex and powerful ordering grammar models have been proposed \cite{futrell2015experiments, wang2016galactic}; however, they have similar limitations, and for our purposes, the model adopted here has the advantage of being simple and interpretable.
-%There currently is no grammatical formalism that is at the same time capable of representing the full range of word order regularities that are found, 
-%\cite{futrell2015experiments}
-
 \begin{table}
@@
 		   Language 
-		   &	\multicolumn{2}{|c|}{Objects} 
-		   &	\multicolumn{2}{|c|}{Adpositions} 
-		   &	\multicolumn{2}{|c|}{Compl.} 
-		   &	\multicolumn{2}{|c|}{Rel.Cl.} 
-		   &	\multicolumn{2}{|c|}{Genitive} 
-		   &	\multicolumn{2}{|c|}{PP}  \\ \hline\hline
+		   &	\multicolumn{2}{c|}{Objects} 
+		   &	\multicolumn{2}{c|}{Adpositions} 
+		   &	\multicolumn{2}{c|}{Compl.} 
+		   &	\multicolumn{2}{c|}{Rel.Cl.} 
+		   &	\multicolumn{2}{c|}{Genitive} 
+		   &	\multicolumn{2}{c|}{PP}  \\ \hline\hline
 	\input{../results/grammars/comparison-table.tex}
@@
 `?' indicates that WALS has no data.
-$*$ indicates that WALS does not list a dominant order, which can mean that neither order is dominant in the language, or that insufficient data was available when compiling WALS \citep{dryer2011evidence}.
-Finally, `--' indicates that the relation does not occur in the corpus.
+$*$ indicates that WALS does not list a dominant order; as \citet{dryer2011evidence} describes, this can mean that neither order is dominant in the language, or that insufficient data was available when compiling WALS.
+Finally, `--' indicates that the relation does not occur in the Universal Dependencies corpus.
 }\label{tab:grammars-wals}
@@
 \begin{equation}\label{eq:rpars}
-	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
+	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
 \end{equation}
 \begin{equation}
-	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
+	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
 \end{equation}
@@
 %\begin{equation}
-%	R_{Pars}^\phi := I[\utterance,\tree] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
+%	R_{Pars}^\phi := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
 %\end{equation}
@@
 \begin{equation}
-	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p_\psi(u) \approx \frac{1}{|\text{Heldout Data}|} \sum_{u \in \text{Heldout Data}} \log p_\psi(u)
+	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p_\psi(u) \approx \frac{1}{|\text{Heldout Data}|} \sum_{u \in \text{Heldout Data}} \log p_\psi(u)
 \end{equation}
@@
 %\begin{equation}\label{eq:rpars}
-%	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
+%	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
 %\end{equation}
 %\begin{equation}
-%	R_{Pred} := - H[\utterance] = \sum_{u} p(u) \log p(u)
+%	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
 %\end{equation}
@@
 \begin{equation}\label{eq:rpars}
-	R_{Pars} := I[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p_\phi(t|u)}{p(t)} = \frac{1}{|\text{Heldout Data}|} \sum_{t,u \in \text{Heldout Data}} \log \frac{p_\phi(t|u)}{p(t)}
+	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p_\phi(t|u)}{p(t)} = \frac{1}{|\text{Heldout Data}|} \sum_{t,u \in \text{Heldout Data}} \log \frac{p_\phi(t|u)}{p(t)}
 \end{equation}
@@
 
-We used the CKY algorithm to compute the syntactic ambiguity $H[\tree|\utterance]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
+We used the CKY algorithm to compute the syntactic ambiguity $\operatorname{H}[\tree|\utterance]$ on the validation partition of the English and Japanese UD corpora, for random and optimized ordering grammars.
 Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable than baseline grammars, for exact parsing of a simple PCFG.
@@
     \includegraphics[scale=.7]{../results/cky/cky-parse.pdf} 
-	\caption{Parsing loss $H[\tree|\utterance]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).
-We report $H[\tree|\utterance]$ normalized by sentence length.}
+	\caption{Parsing loss $\operatorname{H}[\tree|\utterance]$ (lower is better) computed by a simple CKY parser, for random word order grammars (red) and word order grammars optimized for efficiency (blue).
+We report $\operatorname{H}[\tree|\utterance]$ normalized by sentence length.}
     \label{fig:cky-parser}
@@
 
-\paragraph{Sorted order.} A sequence is reordered by sorting by POS tags, and randomizing the order within each block of identical POS tags.
+\paragraph{Lexicographic order.} A sequence is reordered by sorting by POS tags, and randomizing the order within each block of identical POS tags.
 To each word, we then add a symbol encoding the original position in the sequence.
@@
     \includegraphics[scale=.5]{../results/permuted/adversarial-parse-loss-japanese.pdf}
-	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $H[\tree|\utterance]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').
-We report $H[\tree|\utterance]$ normalized by sentence length.}
+	\caption{Parseability of baseline grammars and grammars optimized for efficiency, in English (top) and Japanese (bottom), measured by parsing loss $\operatorname{H}[\tree|\utterance]$ (lower is better), for the four distorted orderings, and the actual orderings (`real').
+We report $\operatorname{H}[\tree|\utterance]$ normalized by sentence length.}
     \label{fig:distorted-parser}
@@
     \includegraphics[scale=.6]{../results/bigrams/bigrams.pdf} 
-	\caption{Surprisal (lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).}
+	\caption{Surprisal (i.e., negative predictability, lower is better) computed from Bigram model, on English and Japanese data ordered according to random ordering grammars (red) and ordering grammars optimized for efficiency (blue).}
     \label{fig:bigrams}
@@
 
-	\caption{Parsing loss ($H[\tree|\utterance]$, normalized by sentence length) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
+	\caption{Parsing loss ($\operatorname{H}[\tree|\utterance]$, normalized by sentence length) for optimized (light blue) and random (black) ordering grammar on Czech data, as a function of the fraction of total training data provided.}
     \label{fig:learning-czech}

### 2019-06-27 writeup/pnas-word-order-si.tex
SI

@@
 \end{equation}
-with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global,kemp2012kinship,frank2012predicting,goodman2013knowledge,kao2014nonliteral,xu2014numeral,kirby2015compression,regier2015word,xu2016historical,futrell2017memory,zaslavsky2018efficient,bennett2018extremely, peloquin2019interactions,zaslavsky2019semantic}.
+with some differences in the precise formalization of these two quantities \cite{ferreri2003least,ferrericancho2007global,kemp2012kinship,frank2012predicting,goodman2013knowledge,kao2014nonliteral,xu2014numeral,kirby2015compression,regier2015word,xu2016historical,futrell2017memory,zaslavsky2018efficient,bennett2018extremely,hahn2018information, peloquin2019interactions,zaslavsky2019semantic}.
 
@@
 
-\paragraph{Relation to Formalizations in Semantic Typology}
-We now discuss how this metric relates to other variants that have been proposed in the information-theoretic literature on language, and describe why~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is most appropriate to our setting.
-This efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains by \citet{zaslavsky2018efficient} and \citet{zaslavsky2019semantic}.
+\paragraph{Relation to Models of Semantic Typology}
+Our model of language efficiency is closely related to models of semantic typology that quantify the efficiency of mappings between concepts and individual words, applied with great success across different domains such color words, container names, and kinship terms~\cite{kemp2012kinship,xu2014numeral,regier2015word,xu2016historical,zaslavsky2018efficient,zaslavsky2019semantic}.
+We discuss how our metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) relates to metrics assumed in this literature, and describe why~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is most appropriate to our setting.
+
+This efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been applied to modeling word meaning across different domains by \citet{zaslavsky2018efficient} and \citet{zaslavsky2019semantic}.
 The Information Bottleneck models complexity using a mutual information term instead of the entropy appearing in~(\ref{eq:eff-general}).
@@
 A few other approaches share the mutual information term for informativity in (\ref{eq:eff-general}), while using complexity measures that are not explicitly information-theoretic.
-In studies of semantic typology by \citet{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \citet{kemp2012kinship} it is the complexity of defining the concepts that are encoded.
-%The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
+In studies of semantic typology by \citet{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms.
+As the entropy of a finite and uniform distribution is the logarithm of the number of objects, this complexity function arises from the entropy measure $\operatorname{H}[S]$~(\ref{eq:eff-general}) in the special case where all forms are used at equal frequency.
+%In \citet{kemp2012kinship} it is the complexity of defining the concepts that are encoded.
+%The number of forms is not applicable as a complexity metric in our case, as languages will typically have infinitely many sentences.
+%As the logarithm of the number of forms upper-bounds the entropy, entropy $\operatorname{H}[S]$ can nevertheless be seen as a generalization of 
 Notably, the models of \citet{regier2007color} and \citet{xu2016historical} have since been reformulated successfully in the Information Bottleneck formalism \cite{zaslavsky2018efficient, zaslavsky2019semantic}, bringing them even closer to our formalization of efficiency.
 
-\paragraph{Relation to Formalizations of Language Evolution}
-There are also related models of the evolution of languages.
-Most related to our work, \citet{kirby2015compression} describe a model of language evolution balancing the pressure for languages to be informative about the intended meanings, while minimizing complexity.
+\paragraph{Relation to Models of Language Evolution}
+Our model is also related to models of language evolution.
+Most closely related to our work, \citet{kirby2015compression} model language evolution as balancing the pressure towards simple languages with the pressure for languages to be informative about the intended meanings.
 Formally, their model studies a Bayesian language learner who infers a language $h$ from data $d$ according to $P(h|d) \propto P(d|h) P(h)$, where $P(h)$ defines a prior distribution over languages, and $P(d|h)$ is the likelihood of observed data $d$ under the grammar $h$, assuming that speakers produce utterances pragmatically.
@@
 where the prior $P(h)$ favors simpler languages.
-This result shows that the model of \citet{kirby2015compression} predicts that language evolution favors languages optimizing a function of the form~(\ref{eq:eff-basic}), with a informativity term identical to that of our model~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}).
+This result shows that the model of \citet{kirby2015compression} predicts that language evolution favors languages optimizing a function of the form~(\ref{eq:eff-basic}), with an informativity term identical to that of our model~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}).
 
@@
 Our work is most closely related to the Rational Speech Acts model of pragmatic reasoning \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
-In line with the other models discussed here, it assumes that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely, peloquin2019interactions}.
+In line with the other models discussed here, it assumes that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely, hahn2018information, peloquin2019interactions}.
 \citet{peloquin2019interactions} provide further discussion of the links between pragmatics and the efficiency metric (\ref{eq:eff-general}-\ref{eq:efficiency-derived}).

### 2019-06-29 writeup/pnas-word-order-si.tex
SI

@@
 
-\paragraph{Bayesian Regression}
+\paragraph{Bayesian Mixed-Effects Regression}
 We modeled the probabilities $p_{L,j}$ that a grammar optimized for data from language $L$ satisfies the $j$-th correlation ($j=1,...,8$) using a multilevel logistic model \cite{gelman2013bayesian}, with random intercepts for the language for whose data the grammar had been optimized, and for its language family, annotated according to \url{http://universaldependencies.org/}.
@@
 We used this prior for $\alpha_j, \sigma_{L,j}, \tau_{L,j}$.
-Note that a correlation that holds in 90\% of cases corresponds to an intercept $\alpha \approx 2.19$ in the logistic model, well within the main probability mass of the prior.
+A correlation that holds in 90\% of cases would correspond to an intercept $\alpha \approx 2.19$ in the logistic model, well within the main probability mass of the prior.
 
@@
 We obtained the posterior density plots in Table 2 (Main Paper) and in Figure (\ref{tab:all-predictions-1}) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\alpha_j$ (\ref{eq:mixed-effects}).
+As the logistic transformation is inverse to the logit transform (\ref{eq:mixed-effects}), this corresponds to the posterior distribution of the prevalence (between 0.0 and 1.0) of each correlation, controlling for languages and language families.
 
@@
 	\end{center}
-	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in the Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.}\label{fig:posterior}
+	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in the Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.
+We show results for grammars optimized for only Predictability (left), only Parseability (center), and full Efficiency (right).}\label{fig:posterior}
 \end{figure}
@@
 The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
-In the last column, we indicate what direction would be expected typologically. 
+The `Real' column provides the prevalence among the 51 languages in the Universal Dependencies data.
+We provide posterior prevalences for grammars optimized for Efficiency, and for grammars optimized for Pars(eability) and Pred(ictability) alone, obtained from the Bayesian mixed-effects analysis controlling for languages and language families (as in Table 2 of the main paper).
+In the last column, we indicate what prevalence is expected according to the typological literature. 
 %We additionally report results obtained when optimizing for Dependency Length Minimization (DLM), a heuristic approximation to efficiency that has been suggested in previous work.
@@
 ``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
-``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
+``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).
+We provide results for grammars optimized for Efficiency, and for grammars optimized for Pars(eability) and Pred(ictability) alone.
+}
 \label{tab:all-predictions-2}
@@
 \paragraph{Estimating Predictability}
-We choose a standard LSTM language model \citep{goldberg2017neural, hochreiter1997long}.
+We choose a standard LSTM language model \citep{goldberg2017neural, hochreiter1997long}, as such recurrent neural models are the strongest known predictors of the surprisal effect on human processing effort~\cite{frank2011insensitivity,goodkind2018predictive}.
 This model uses a recurrent neural network to compute the predictability of a sentence $u = u_1...u_n$\footnote{Technically, $u_1...u_{n-1}$ are words, and $u_n$ is an end-of-sentence token, to ensure the probability distribution over all sentences is normalized.}:
@@
 where $\psi$ are the parameters of the recurrent LSTM network, optimized on training data (see paragraph `Optimization Details').
+
 We estimate the average predictability of a language as a Monte Carlo estimate on held-out data:
@@
         \caption{Average dependency length for grammars optimized to minimize dependency length (DLM, left), optimized for efficiency (second), the real orderings found in corpora (third), and random baseline grammars (right).
-The lines connect the means points for each of the 51 languages in our sample.}
+The lines connect the mean points for each of the 51 languages in our sample.}
     \label{fig:dlm-avg}

### 2019-09-01 writeup/pnas-word-order-si.tex

@@
 
+\paragraph{Evaluating Accuracy of Formalization}
+An anonymous reviewer notes that the mapping between Dryer's relations and UD is not perfect, since some of the UD relations subsume other relations.
+Here we show that this is not a problem, since the ordering of the various relations subsumed under the UD label typologically strongly agree.
+
+\begin{enumerate}
+\item aux also covers modal, ... auxiliaries
+
+Universals Archive 501 provides references, other people have made the statement for auxiliaires in general.
+
+We also collected, for each language, all auxiliaires, that is, all dependents of the \emph{aux} dependency, occurring at least 10 times, and compared their dominant orders.
+In X languages, all extracted auxiliaries had the same dominant order.
+In X languages, there were auxiliaires with divergent order. In Y cases, these were modal auxiliaries, or uninflected elements. In Z cases, these were tense-aspect auxiliaries.
+Taken together, this shows that the dominant order of the \emph{aux} relation strongly coincides with that of inflected tense-aspect or negation auxiliaries.
+
+\item nmod covers all noun-modifying NPs and PPs:
+WALS evaluation, 37 entries, agreement in all but 2 cases (Danish, Swedish).
+
+\item acl not  just relative clauses: In the WALS evaluation, acl agrees with the WALS entry for relative clauses: WALS has an entry in 36 languages where acl occurs in the treebank, which agrees with the extracted grammars in all but 3 languages (Estonian, Finnish, Tamil).
+Also, UD provides a specific acl:relcl label in 21 of the languages.
+In all but 3 languages, the majority order is the same for the general acl label as for the specific acl:relcl one (exceptions: Estonian, Finnish, Hindi).
+
+\item obl covers adjunct NPs, ...: WALS evaluation: has 18 entries, no disagreement with extracted grammars
+
+\item xcomp covers other control verbs, not just volition.
+
+For each language, we extracted all lemmas of words heading an \emph{xcomp} dependency, occurring at least 2 times.
+In X languages, all extracted words had the same dominant order.
+
+In Ancient Greek, Latin, and Old Church Slavonic, verbs of volition agree with the extracted order, while several other verbs that do not indicate volition show opposite order.
+In Gothic, the verb of volition (wiljan) has opposite order, resulting in an apparent violation of Correlation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}.
+
+In four Germanic languages (Afrikaans, Danish, Dutch, and German), the verb of volition (Afrikaans \emph{wil}, Danish \emph{ville}, Dutch \emph{willen}, German \emph{wollen}) is mostly annotated with the \emph{aux} relation due to UD annotation guidelines, but in all languages, its dominant order (verb of volition before its complement) is compatible with the dominant order of the \emph{xcomp} dependency (head-initial).
+
+Taken together, the order of the xcomp dependency tends to agree with that of most other xcomp dependencies, with the sole exception of Gothic.
+\end{enumerate}
+
 \section{Formalizing Comunicative Efficiency}
@@
 \paragraph{Relation to Models in Other Disciplines}
-Beyond the study of natural language, the eficiency metric (\ref{eq:eff-general}) is also closely related to information-theoretic models in other disciplines.
+Beyond the study of natural language, the efficiency metric (\ref{eq:eff-general}) is also closely related to information-theoretic models in other disciplines.
 The tradeoff between informativity and complexity of communication systems is studied by Rate-distortion theory \cite{shannon1948mathematical, berger1971rate}.
@@
 In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.
+We also provide the raw numerical values, before $z$-scoring, in Figure~\ref{fig:pareto-per-lang-raw}.
 
@@
 Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.}\label{fig:pareto-per-lang}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-raw.pdf}
+	\caption[Predictability and Parseability]{Raw numerical values estimated for Predictability (negative surprisal), and negative syntactic ambiguity $-H[T|U]$, before $z$-scoring.
+For more meaningful comparison, both quantities are normalized by the number of words in the corpus, i.e., we plot per-word negative surprisal and per-word difficulty in recovering syntactic structures.
+Note that the negative syntactic ambiguity $-H[T|U]$ equals parseability $I[T,U] = H[T] - H[T|U]$ up to a per-language constant $H[T]$, which we do not attempt to estimate.}\label{fig:pareto-per-lang-raw}
 \end{figure}
@@
 \end{table}
+
+\paragraph{Comparison to other Formalizations}
+
+Greenberg's original correlations/implications
+
+\cite{justeson1990explanation} conducted a log-linear analysis on typological judgments of 147 languages, constructing an undirected graphical model.
+For the relations examined here, beyond correlations with verb-object order, they found additional correlations between G-N, N-Adp and between Rel-N, N-Adp.
+With their model, they found no evidence for a direct correlation V-O, G-N.
 
@@
 
-These quatities are estimated using two neural models, as described in Section~\ref{sec:neural-architectures}:
+These quantities are estimated using two neural models, as described in Section~\ref{sec:neural-architectures}:
 A \key{parser} recovers syntactic structures from utterances by computing a distribution $p_\phi(t|u)$, parameterized via parser parameters $\phi$.

### 2019-09-09 models/revision/grammar_formalism/computeLikelihood.py
code

@@ -1,0 +1,325 @@
+
+
+
+
+import random
+import sys
+
+objectiveName = "LM"
+
+language = sys.argv[1]
+
+batchSize = 1
+
+myID = random.randint(0,10000000)
+
+FILENAME = "dotdotdot_COARSE_PLANE_noise.py"
+
+#with open("../../../raw-results/LOG"+language+"_"+FILENAME+"_model_"+str(myID)+".txt", "w") as outFile:
+#    print >> outFile, " ".join(sys.argv)
+
+
+posUni = set()
+posFine = set()
+
+
+
+from math import log, exp, sqrt
+from random import random, shuffle, randint, choice
+import os
+
+
+from corpusIterator_FuncHead import CorpusIteratorFuncHead
+
+originalDistanceWeights = {}
+
+
+def makeCoarse(x):
+   if ":" in x:
+      return x[:x.index(":")]
+   return x
+
+def initializeOrderTable():
+   orderTable = {}
+   keys = set()
+   vocab = {}
+   distanceSum = {}
+   distanceCounts = {}
+   depsVocab = set()
+   for partition in ["train", "dev"]:
+     for sentence in CorpusIteratorFuncHead(language,partition).iterator():
+      for line in sentence:
+          vocab[line["word"]] = vocab.get(line["word"], 0) + 1
+          line["coarse_dep"] = makeCoarse(line["dep"])
+          depsVocab.add(line["coarse_dep"])
+          posFine.add(line["posFine"])
+          posUni.add(line["posUni"])
+  
+          if line["coarse_dep"] == "root":
+             continue
+          posHere = line["posUni"]
+          posHead = sentence[line["head"]-1]["posUni"]
+          dep = line["coarse_dep"]
+          direction = "HD" if line["head"] < line["index"] else "DH"
+          key = dep
+          keyWithDir = (dep, direction)
+          orderTable[keyWithDir] = orderTable.get(keyWithDir, 0) + 1
+          keys.add(key)
+          distanceCounts[key] = distanceCounts.get(key,0.0) + 1.0
+          distanceSum[key] = distanceSum.get(key,0.0) + abs(line["index"] - line["head"])
+   dhLogits = {}
+   for key in keys:
+      hd = orderTable.get((key, "HD"), 0) + 1.0
+      dh = orderTable.get((key, "DH"), 0) + 1.0
+      dhLogit = log(dh) - log(hd)
+      dhLogits[key] = dhLogit
[…]

### 2019-09-09 models/revision/grammar_formalism/computeLikelihoodAll.py

@@ -1,0 +1,11 @@
+languages = ["Hindi", "Swedish", "German", "Urdu", "English", "Spanish", "Chinese", "Slovenian", "Estonian", "Norwegian", "Serbian", "Croatian", "Finnish", "Portuguese", "Catalan", "Russian", "Arabic", "Czech", "Japanese", "French", "Latvian", "Basque", "Danish", "Dutch", "Ukrainian", "Gothic", "Hebrew", "Hungarian", "Latin", "Persian", "Bulgarian", "Romanian", "Indonesian", "Greek", "Turkish", "Slovak", "Belarusian", "Galician", "Italian", "Lithuanian", "Polish", "Vietnamese", "Korean", "Tamil", "Irish", "Marathi", "Afrikaans", "Telugu", "Coptic", "Old_Church_Slavonic", "Ancient_Greek"]
+
+import random
+random.shuffle(languages)
+import os
+import subprocess
+for language in languages:
+   for typ in ["computeLikelihood.py", "computeLikelihood_PureUD.py"]:
+      if language+"_"+typ+".txt" in os.listdir("../../../raw-results/treebank_likelihood/"):
+         continue
+      subprocess.call(["/u/nlp/anaconda/ubuntu_16/envs/py27-mhahn/bin/python2.7", typ, language])

### 2019-09-09 models/revision/grammar_formalism/computeLikelihood_PureUD.py

@@ -1,0 +1,325 @@
+
+
+
+
+import random
+import sys
+
+objectiveName = "LM"
+
+language = sys.argv[1]
+
+batchSize = 1
+
+myID = random.randint(0,10000000)
+
+FILENAME = "dotdotdot_COARSE_PLANE_noise.py"
+
+#with open("../../../raw-results/LOG"+language+"_"+FILENAME+"_model_"+str(myID)+".txt", "w") as outFile:
+#    print >> outFile, " ".join(sys.argv)
+
+
+posUni = set()
+posFine = set()
+
+
+
+from math import log, exp, sqrt
+from random import random, shuffle, randint, choice
+import os
+
+
+from corpusIterator import CorpusIterator
+
+originalDistanceWeights = {}
+
+
+def makeCoarse(x):
+   if ":" in x:
+      return x[:x.index(":")]
+   return x
+
+def initializeOrderTable():
+   orderTable = {}
+   keys = set()
+   vocab = {}
+   distanceSum = {}
+   distanceCounts = {}
+   depsVocab = set()
+   for partition in ["train", "dev"]:
+     for sentence in CorpusIterator(language,partition).iterator():
+      for line in sentence:
+          vocab[line["word"]] = vocab.get(line["word"], 0) + 1
+          line["coarse_dep"] = makeCoarse(line["dep"])
+          depsVocab.add(line["coarse_dep"])
+          posFine.add(line["posFine"])
+          posUni.add(line["posUni"])
+  
+          if line["coarse_dep"] == "root":
+             continue
+          posHere = line["posUni"]
+          posHead = sentence[line["head"]-1]["posUni"]
+          dep = line["coarse_dep"]
+          direction = "HD" if line["head"] < line["index"] else "DH"
+          key = dep
+          keyWithDir = (dep, direction)
+          orderTable[keyWithDir] = orderTable.get(keyWithDir, 0) + 1
+          keys.add(key)
+          distanceCounts[key] = distanceCounts.get(key,0.0) + 1.0
+          distanceSum[key] = distanceSum.get(key,0.0) + abs(line["index"] - line["head"])
+   dhLogits = {}
+   for key in keys:
+      hd = orderTable.get((key, "HD"), 0) + 1.0
+      dh = orderTable.get((key, "DH"), 0) + 1.0
+      dhLogit = log(dh) - log(hd)
+      dhLogits[key] = dhLogit
+   return dhLogits, vocab, keys, depsVocab
+
[…]

### 2019-09-09 models/revision/grammar_formalism/computeLikelihood_Symm.py

@@ -1,0 +1,323 @@
+
+
+
+
+import random
+import sys
+
+objectiveName = "LM"
+
+language = sys.argv[1]
+
+batchSize = 1
+
+myID = random.randint(0,10000000)
+
+FILENAME = "dotdotdot_COARSE_PLANE_noise.py"
+
+#with open("../../../raw-results/LOG"+language+"_"+FILENAME+"_model_"+str(myID)+".txt", "w") as outFile:
+#    print >> outFile, " ".join(sys.argv)
+
+
+posUni = set()
+posFine = set()
+
+
+
+from math import log, exp, sqrt
+from random import random, shuffle, randint, choice
+import os
+
+
+from corpusIterator_FuncHead import CorpusIteratorFuncHead
+
+originalDistanceWeights = {}
+
+
+def makeCoarse(x):
+   if ":" in x:
+      return x[:x.index(":")]
+   return x
+
+def initializeOrderTable():
+   orderTable = {}
+   keys = set()
+   vocab = {}
+   distanceSum = {}
+   distanceCounts = {}
+   depsVocab = set()
+   for partition in ["train", "dev"]:
+     for sentence in CorpusIteratorFuncHead(language,partition).iterator():
+      for line in sentence:
+          vocab[line["word"]] = vocab.get(line["word"], 0) + 1
+          line["coarse_dep"] = makeCoarse(line["dep"])
+          depsVocab.add(line["coarse_dep"])
+          posFine.add(line["posFine"])
+          posUni.add(line["posUni"])
+  
+          if line["coarse_dep"] == "root":
+             continue
+          posHere = line["posUni"]
+          posHead = sentence[line["head"]-1]["posUni"]
+          dep = line["coarse_dep"]
+          direction = "HD" if line["head"] < line["index"] else "DH"
+          key = dep
+          keyWithDir = (dep, direction)
+          orderTable[keyWithDir] = orderTable.get(keyWithDir, 0) + 1
+          keys.add(key)
+          distanceCounts[key] = distanceCounts.get(key,0.0) + 1.0
+          distanceSum[key] = distanceSum.get(key,0.0) + abs(line["index"] - line["head"])
+   dhLogits = {}
+   for key in keys:
+      hd = orderTable.get((key, "HD"), 0) + 1.0
+      dh = orderTable.get((key, "DH"), 0) + 1.0
+      dhLogit = log(dh) - log(hd)
+      dhLogits[key] = dhLogit
[…]

### 2019-09-09 results/plane/unlexicalized/plane-2D-best-visualization-large-tests-unlexicalized.R

@@ -1,0 +1,178 @@
+library(lme4)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(dplyr)
+dataS =  read.csv("../../../grammars/plane/controls/plane-fixed-withoutPOS.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS3 = read.csv("../../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+#dataS4 = read.csv("../../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+#dataS5 = read.csv("../../../grammars/plane/plane-fixed-random3.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+#dataS6 = read.csv("../../../grammars/plane/plane-fixed-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+#dataS7 = read.csv("../../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS = rbind(dataS, dataS2, dataS3)#, dataS4, dataS5, dataS6, dataS7)
+dataP =  read.csv("../../../grammars/plane/controls/plane-parse-lexicalized.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP2 = read.csv("../../../grammars/plane/plane-parse-best.tsv", sep="\t") %>% mutate(Model = as.character(Model), LAS=NA)
+dataP3 = read.csv("../../../grammars/plane/plane-parse-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model), LAS=NA)
+#dataP4 = read.csv("../../../grammars/plane/plane-parse-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model))
[…]

### 2019-09-10 writeup/pnas-word-order-si.tex

@@
 
-We also collected, for each language, all auxiliaires, that is, all dependents of the \emph{aux} dependency, occurring at least 10 times, and compared their dominant orders.
+We also collected, for each language, all auxiliaires, that is, all dependents of the \emph{aux} dependency, occurring at least 10 times, and compared their dominant orders, which we operationalized as their more common order in the treebank (auxiliary-head or head-auxiliary).
 In X languages, all extracted auxiliaries had the same dominant order.
-In X languages, there were auxiliaires with divergent order. In Y cases, these were modal auxiliaries, or uninflected elements. In Z cases, these were tense-aspect auxiliaries.
+In X languages, there were auxiliaires with different dominant orders.
+		In Y cases, these were modal or passive auxiliaries, or uninflected elements.
+		In Z cases, these were tense-aspect auxiliaries:
+		In Persian and Afrikaans, not all tense-aspect auxiliaries 
+		(1) In Persian, not all tense-aspect auxiliaries have the same order: E.g.
+budan as a perfect auxiliary follows the main verb (Mace p.
+117), whereas xaastan xaah- as a future auxiliary precedes it (Mace p.
+122).
+		(1) In Afrikaans, the past auxiliary \emph{het} shows dominant aux-head order, whereas the overall most common order of auxiliaries is head-aux (e.g. future auxiliary \emph{sal}).
+		(2) In Old Church Slavonic, most aux dependencies have the reflexive \emph{sebe} which tends to precede the head, whereas inflected verbal auxiliaries (primarily, forms of \emph{byti}) have opposite order, leading to a seeming violation of the correlation  \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
+
 Taken together, this shows that the dominant order of the \emph{aux} relation strongly coincides with that of inflected tense-aspect or negation auxiliaries.
@@
 Note that the negative syntactic ambiguity $-H[T|U]$ equals parseability $I[T,U] = H[T] - H[T|U]$ up to a per-language constant $H[T]$, which we do not attempt to estimate.}\label{fig:pareto-per-lang-raw}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized.pdf}
+	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang}.}
 \end{figure}
@@
 In addition to the $p$-value, we report point estimate and 95\% confidence interval for the fraction of random grammars that have values below real grammars.}\label{fig:pareto-per-lang-stats}
+\end{table}
+
+\begin{table}
+\centering
+\small{
+\begin{tabular}{l||ll|lll|lll}
+Language & Pred. (t) & Parse. (t) & \multicolumn{3}{c|}{Pred. (Binomial)} & \multicolumn{3}{c}{Parseab. (Binomial)} \\ 
+&  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
+\input{../results/plane/unlexicalized/per-language-statistics-unlexicalized.tex}
+\end{tabular}
+}
+	\caption{Per-language results in Study 1, with \emph{lexicalized} parsers and word-level-only language models. Compare Table~\ref{fig:pareto-per-lang-stats}}
 \end{table}

### 2019-09-10 writeup/pnas-word-order-si.tex

--- 2019-06-29 writeup/pnas-word-order-si.tex
@@
 
+\paragraph{Evaluating Accuracy of Formalization}
+An anonymous reviewer notes that the mapping between Dryer's relations and UD is not perfect, since some of the UD relations subsume other relations.
+Here we show that this is not a problem, since the ordering of the various relations subsumed under the UD label typologically strongly agree.
+
+\begin{enumerate}
+	\item Correlation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} captures correlations with inflected tense, aspect, and negation auxiliaries as stated by \citet{dryer1992greenbergian}; however, \emph{aux} aso encompasses other types of auxiliaires, such as modals.
+We note that other authors, including \citet{greenberg1963universals}, have stated the correlation for all inflected auxiliaries; for further references, we refer to \citet[Number 501]{plank2000universals}.
+
+We used the UD treebanks to confirm that different auxiliaires tend to pattern together, and that the most frequent order of the \emph{aux} relation coincides with that of inflected tense-aspect or negation auxiliaries.
+
+We collected, for each UD language, all dependents of the \emph{aux} dependency, occurring at least 10 times, and compared their dominant orders, which we operationalized as their more common order in the treebank (auxiliary-head or head-auxiliary).
+The dependency occurs in all but two very small treebanks (Telugu and Irish).
+In 43 languages, all extracted auxiliaries had the same dominant order, with the possible exception of uninflected particles labeled \emph{aux} (Croatian, German, Polish, Ukrainian).
+In three languages (Ancient Greek, Russian), there were other auxiliaries with different dominant order, but these were modal or passive auxiliaries.
+Finally, in three languages (Afrikaans, Old Church Slavonic, and Persian), not all tense-aspect auxiliaries showed the same dominant order as the \emph{aux} dependency overall.
+For instance, in Persian, the perfect auxiliary \emph{budan} follows the main verb, whereas the future auxiliary \emph{xaastan xaah-} precedes it~\citep[pp. 117, 121]{mace2015persian}.
+%		(1) In Afrikaans, the past auxiliary \emph{het} shows dominant aux-head order, whereas the overall most common order of auxiliaries is head-aux (e.g. future auxiliary \emph{sal}).
+%		Finally, in Old Church Slavonic, most aux dependencies have the reflexive \emph{sebe} which tends to precede the head, whereas inflected verbal auxiliaries (primarily, forms of \emph{byti}) have opposite order, leading to a seeming violation of the correlation  \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
+
+Taken together, this shows that the dominant order of the \emph{aux} relation strongly coincides with that of inflected tense-aspect auxiliaries, except for a small number of languages where different tense-aspect auxiliaries show different orders.
+
+\item Correlation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}} is formalized using \emph{nmod} which covers not only genitives, but also all other noun-modifying NPs and PPs.
+	The evaluation of extracted grammars against WALS (Table~\ref{tab:grammars-wals}) shows that, among the 37 languages where WALS has an entry, the dominant direction of \emph{nmod} agrees with that of genitives, with two exceptions (Danish and Swedish).
+
+TODO figure out why
+
+\item Correlation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}} is formalized using \emph{acl}, which covers not just relative clauses, but also other adnominal clauses.
+	In the WALS evaluation (Table~\ref{tab:grammars-wals}), the dominant order of \emph{acl} agrees with the WALS entry for relative clauses in all but three languages (Estonian, Finnish, Tamil) out of the 36 languages for which WALS has an entry.
+		Also, UD provides a specific \emph{acl:relcl} sub-label for relative clauses in 21 of the languages.
+		In all but three languages, the dominant order is the same for the general \emph{acl} label as for the specific \emph{acl:relcl} one (exceptions: Estonian, Finnish, Hindi).
+
+TODO figure out why
+
+	\item Correlation \multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}} is formalized using \emph{nmod}, which covers not only PPs and standards of comparison, but also adjunct NPs.
+		In the WALS evaluation (Table~\ref{tab:grammars-wals}), the dominant order of \emph{nmod} agrees with that annotated for obliques in all 18 languages for which WALS has an entry.
+
+	\item Correlation \multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}} is formalized using \emph{xcomp}, which covers other control verbs, not just verbs of volition (`want').
+
+		We used the UD treebanks to investigate whether there are differences in the ordering of `want' and other verbs using the \emph{xcomp} dependency.
+
+		The dependency is annotated in all but two languages (Japanese and Turkish).
+
+		For each language, we extracted all lemmas of words heading an \emph{xcomp} dependency, occurring at least 2 times.
+In 39 languages, all extracted words had the same dominant order.
+Additionally, in four Germanic languages (Afrikaans, Danish, Dutch, and German), the verb of volition (Afrikaans \emph{wil}, Danish \emph{ville}, Dutch \emph{willen}, German \emph{wollen}) is mostly annotated with the \emph{aux} relation due to UD annotation guidelines, but in all languages, its dominant order (verb of volition before its complement) agrees with the dominant order of the \emph{xcomp} dependency (head-initial).
+		In three historical languages (Ancient Greek, Latin, and Old Church Slavonic), verbs of volition agree with the dominant order of \emph{xcomp}, while several other verbs that do not indicate volition show opposite dominant order.
+Finally, in Gothic, the verb of volition (\emph{wiljan}) has opposite order, resulting in an apparent violation of Correlation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}.
+
+		Taken together, the order of the xcomp dependency tends to agree with that of most other \emph{xcomp} dependencies, with the sole exception of Gothic.
+\end{enumerate}
+
 \section{Formalizing Comunicative Efficiency}
@@
 Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.}\label{fig:pareto-per-lang}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-raw.pdf}
+	\caption[Predictability and Parseability]{Raw numerical values estimated for Predictability (negative surprisal), and negative syntactic ambiguity $-H[T|U]$, before $z$-scoring.
+For more meaningful comparison, both quantities are normalized by the number of words in the corpus, i.e., we plot per-word negative surprisal and per-word difficulty in recovering syntactic structures.
+Note that the negative syntactic ambiguity $-H[T|U]$ equals parseability $I[T,U] = H[T] - H[T|U]$ up to a per-language constant $H[T]$, which we do not attempt to estimate.}\label{fig:pareto-per-lang-raw}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized.pdf}
+	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang}.}
 \end{figure}

### 2019-09-12 models/revision/toy_simulations/estimateParseability_All5.py
code

@@ -1,0 +1,19 @@
+import subprocess
+import random
+
+
+val = {}
+for probVPBranching in [0.5]: #sorted([0.1, 0.2, 0.3, 0.4, 0.5], key=lambda k: random.random()):
+    val["probVPBranching"] = probVPBranching
+    for probObj in [0.3]: #sorted([0.3, 0.5, 0.7, 0.9], key=lambda k: random.random()):
+      val["probObj"] = probObj
+      val["probNPBranching"] = 0.0
+      for i in [1,2]:
+         val["correlation_xcomp"] = True if i ==1 else False
+         val["dlm_xcomp"] = True if i == 1 else False
+         val["correlation_acl"] = True if i == 1 else False
+         command = ["/u/nlp/anaconda/ubuntu_16/envs/py27-mhahn/bin/python2.7", "estimateParseability_CKY.py"]
+         for a, b in val.iteritems():
+           command.append("--"+a+"="+str(b))
+         subprocess.call(command)   
+      

### 2019-09-12 models/revision/toy_simulations/estimateParseability_All6.py

@@ -1,0 +1,19 @@
+import subprocess
+import random
+
+
+val = {}
+for probVPBranching in sorted([0.1, 0.2, 0.3, 0.4], key=lambda k: random.random()): # 0.1, 0.2, 0.3, 0.4, 
+    val["probVPBranching"] = probVPBranching
+    for probObj in sorted([0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9], key=lambda k: random.random()):
+      val["probObj"] = probObj
+      val["probNPBranching"] = 0.3
+      for i in [1,2]:
+         val["correlation_xcomp"] = True if i ==1 else False
+         val["dlm_xcomp"] = True if i == 1 else False
+         val["correlation_acl"] = True if i == 1 else False
+         command = ["/u/nlp/anaconda/ubuntu_16/envs/py27-mhahn/bin/python2.7", "estimateParseability_CKY.py"]
+         for a, b in val.iteritems():
+           command.append("--"+a+"="+str(b))
+         subprocess.call(command)   
+      

### 2019-09-12 writeup/pnas-word-order-si.tex
code
SI

@@
 Note that the negative syntactic ambiguity $-H[T|U]$ equals parseability $I[T,U] = H[T] - H[T|U]$ up to a per-language constant $H[T]$, which we do not attempt to estimate.}\label{fig:pareto-per-lang-raw}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized.pdf}
-	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang}.}
 \end{figure}
@@
 
+\section{Other Methods of Estimating Efficiency and Efficiency Optimization}
+
+\subsection{Lexicalized Models}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized.pdf}
+	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang}.}
+\end{figure}
+
+\subsection{Original UD Format}
+
+\subsection{Nondeterministic Baseline Grammars}
+
 %\subsection{Discussion}
@@
 %DLM turns out to be predicted by our model
+
+\section{Simulation for Parseability and Correlating Orders}
+
+%VP -> V
+%VP -> VP[head] VP[xcomp]
+%VP -> VP[head] NP[obj]
+%
+%NP -> N
+%NP -> NP[head] VP[acl]
+
+We consider the following head-outward generative model.
+Verbs generate verb dependents (\emph{xcomp}) and noun dependents (\emph{obj}), independently.
+The overall number $N$ of dependents is $NB(1, p_{branching})$, the number of \emph{obj} dependents is $Binom(p_{obj}, N)$.
+Nouns can generate verb dependents (\emph{acl}), of number $NB(1, p_{acl})$.
+
+Trees are linearized using one of two grammars: One (`Correlating') places \emph{obj}, \emph{xcomp}, and \emph{acl} dependents on the same side of the head, and (in accordance with crosslinguistic tendencies, which are predicted by DLM) places the \emph{obj} dependents closer to the head than \emph{xcomp} dependents.
+The other grammar (`Anti-Correlating') places \emph{xcomp} and \emph{acl} dependents opposite to \emph{obj} dependents.
+
+We formulated this model as a binary-branching PCFG, and used a CKY parser to estimate $I[\mathcal{T}, \mathcal{U}]$ from 10,000 random sample sentences.
+
+We computed this for different settings of $p_{branching} \in [0, 0.5]$ and $p_{obj} \in [0,1]$, at $p_{acl} \in \{0, 0.3\}$.\footnote{For $p_{acl} = 0.3$, we only computed values for $p_{branching} \leq 0.4$, due to high computational cost on long sentences resulting from $p_{branching} = 0.5$ and $p_{acl} = 0.3$.}
+For these settings, we computed the difference in $I[\mathcal{T}, \mathcal{U}]$ between the two grammars.
+Results are shown in Figure~\ref{fig:toy-parseability}.
+
+\begin{figure}
+	\begin{center}	
+\includegraphics[width=0.4\textwidth]{../models/revision/toy_simulations/result_NPBranching0_0.pdf}
+\includegraphics[width=0.4\textwidth]{../models/revision/toy_simulations/result_NPBranching0_3.pdf}
+\end{center}
+
+	\caption{Difference in parseability $I[\mathcal{T}, \mathcal{U}]$ (normalized by sentence length) between the correlating and anti-correlating grammars, for $p_{acl} = 0.0$ (left) and $p_{acl} = 0.3$ (right).
+Positive values indicate greater values of $I[\mathcal{T}, \mathcal{U}]$ for correlating grammars.}\label{fig:toy-parseability}
+\end{figure}
 

### 2019-09-12 writeup/pnas-word-order-si.tex
add a bit to si

@@
 
-This efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been applied to modeling word meaning across different domains by \citet{zaslavsky2018efficient} and \citet{zaslavsky2019semantic}.
-The Information Bottleneck models complexity using a mutual information term instead of the entropy appearing in~(\ref{eq:eff-general}).
-This mutual information formalization is appropriate to codes that are nondeterministic \cite{zaslavsky2018efficient}.
-We assume deterministic grammars that transduce every underlying syntactic structure into one surface order; our model therefore corresponds to the deterministic version of the Information Bottleneck, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure~\cite{strouse2017deterministic}.
-
-A few other approaches share the mutual information term for informativity in (\ref{eq:eff-general}), while using complexity measures that are not explicitly information-theoretic.
+This efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}) is part of the Information Bottleneck family of models.
+The Information Bottleneck was introduced by \citet{tishby1999information} and has recently been applied to modeling word meaning across different domains by \citet{zaslavsky2018efficient} and \citet{zaslavsky2019semantic}.
+In the standard Information Bottleneck, complexity is modeled using a mutual information term, instead of the entropy term appearing in~(\ref{eq:eff-general}).
+The setting for the standard Information Bottleneck is a case where there is a random variable $X$ which contains information about some underlying variable of interest $Y$; the goal of the Information Bottleneck is to find a representation $\hat{X}$ of $X$ which maximizes $\operatorname{I}[\hat{X},Y]$ while minimizing $\operatorname{I}[\hat{X},X]$.
+One key property of the standard Information Bottleneck is that it results in codes $\hat{X}$ that are nondeterministic.
+
+The variant of the Information Bottleneck that we use has been explored in the machine learning literature by \citet{strouse2017deterministic} and dubbed the ``Deterministic Information Bottleneck'' because, in the setting studied by \citet{strouse2017efficient}, it results in codes that are a deterministic function of the information to be expressed.
+We use this version of the Information Bottleneck because (1) it has been proposed in previous literature as a generic formalization of efficiency \citep{ferreri2003least}, and (2) it is not clear what would count as the three variables $Y$, $X$, and $\hat{X}$ in our setting.
+In our setting we have unordered tree structures $T$ to be conveyed, and utterances $U$ representing them.
+It is not currently clear what would count as a third variable for the application of the standard Information Bottleneck, although we believe such formulations may be fruitful in the future.
+
+%We assume deterministic grammars that transduce every underlying syntactic structure into one surface order; our model therefore corresponds to the deterministic version of the Information Bottleneck, which arises when the encoder is deterministic, and which uses the entropy as the complexity measure~\cite{strouse2017deterministic}.
+
+A few other approaches to formalizing efficiency share the mutual information term for informativity in (\ref{eq:eff-general}), while using complexity measures that are not explicitly information-theoretic.
 In studies of semantic typology by \citet{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms.
@@
 Beyond the study of natural language, the efficiency metric (\ref{eq:eff-general}) is also closely related to information-theoretic models in other disciplines.
-The tradeoff between informativity and complexity of communication systems is studied by Rate-distortion theory \cite{shannon1948mathematical, berger1971rate}.
-The \emph{Infomax Principle} derives parsimonious data representations by maximizing the mutual information between data and representations, subject to constraints on the representations~\cite{linsker1988self}; a constraint on the representation entropy leads to a metric equivalent to (\ref{eq:eff-general}) and to a version of the Free-Energy principle (see Section S3 in \citet{friston2010free}).
+The tradeoff between informativity and complexity of communication systems is studied extensively in rate--distortion theory \cite{berger1971rate}.
+Our efficiency metric is closely related to the the \emph{Infomax principle} from theoretical neuroscience, which is a theory of how information is encoded in neuronal signals.
+The Infomax principle derives parsimonious data representations by maximizing the mutual information between data and representations, subject to constraints on the representations~\cite{linsker1988self}; a constraint on the representation entropy leads to a metric equivalent to (\ref{eq:eff-general}) and to a version of the Free-Energy principle (see Section S3 in \citet{friston2010free}).
+A family of Infomax models called ``Coherent Infomax'' has been proposed by \citet{kay2011coherent}; our efficiency metric is a special case within this framework.
+
+% J.
+W.
+Kay, W.
+A.
+Phillips, Coherent Infomax as a computational goal for neural systems., Bull Math Biol 73 (2) (2011) 344–372.
+doi:10.1007/s11538-010-9564-x.
+URL http://dx.doi.org/10.1007/s11538-010-9564-x
 

### 2019-09-12 writeup/pnas-word-order-si.tex
add prose

@@
 We use this version of the Information Bottleneck because (1) it has been proposed in previous literature as a generic formalization of efficiency \citep{ferreri2003least}, and (2) it is not clear what would count as the three variables $Y$, $X$, and $\hat{X}$ in our setting.
-In our setting we have unordered tree structures $T$ to be conveyed, and utterances $U$ representing them.
+In our setting we have unordered tree structures $\mathcal{T}$ to be conveyed, and utterances $\mathcal{U}$ representing them.
 It is not currently clear what would count as a third variable for the application of the standard Information Bottleneck, although we believe such formulations may be fruitful in the future.
@@
 \subsection{Distorted graph-based parsers}
+\label{sec:distorted}
 
@@
 
-We consider the following head-outward generative model.
+When we optimize grammars for efficiency, we find that the optimized grammars exhibit dependency length minimization and the Greenbergian word order correlations.
+To some extent, this result is surprising, because previous functional explanations for DLM (and the Greenbergian correlations, which are predicted by DLM) have been based on the idea of limitations in working memory, and yet our models do not instantiate any explicit working memory pressures.
+Nor is it the case that our parsers have implicit locality biases, as demonstrated by the experiments in Section~\ref{sec:distorted} above.
+Our results therefore suggest that DLM and word order correlations might arise purely because they enable tree structures to be better recovered from trees, and/or they make sequences more predictable.
+
+Here we perform some simulation studies to bolster the argument that DLM and word order correlations can enhance the recoverability of tree structures in a generic sense, without any appeal to memory limitations.
+To do so, we experiment with toy grammars that can be defined to either (1) exhibit word order correlations or (2) not, and we test whether the grammars of type (1) are more or less parseable than the grammars of type (2).
+We measure parseability using a CYK PCFG parser, thus removing any potential confounds arising from the neural network parsing model.
+
+Our toy grammar consists of the following head-outward generative model.
 Verbs generate verb dependents (\emph{xcomp}) and noun dependents (\emph{obj}), independently.
-The overall number $N$ of dependents is $NB(1, p_{branching})$, the number of \emph{obj} dependents is $Binom(p_{obj}, N)$.
-Nouns can generate verb dependents (\emph{acl}), of number $NB(1, p_{acl})$.
+The overall number $N$ of dependents is $NB(1, p_\text{branching})$, the number of \emph{obj} dependents is $Binom(p_{obj}, N)$.
+Nouns can generate verb dependents (\emph{acl}), of number $NB(1, p_\text{acl})$.
 
@@
 
-We computed this for different settings of $p_{branching} \in [0, 0.5]$ and $p_{obj} \in [0,1]$, at $p_{acl} \in \{0, 0.3\}$.\footnote{For $p_{acl} = 0.3$, we only computed values for $p_{branching} \leq 0.4$, due to high computational cost on long sentences resulting from $p_{branching} = 0.5$ and $p_{acl} = 0.3$.}
+We computed this for different settings of $p_\text{branching} \in [0, 0.5]$ and $p_{obj} \in [0,1]$, at $p_\text{acl} \in \{0, 0.3\}$.\footnote{For $p_\text{acl} = 0.3$, we only computed values for $p_\text{branching} \leq 0.4$, due to high computational cost on long sentences resulting from $p_\text{branching} = 0.5$ and $p_\text{acl} = 0.3$.}
 For these settings, we computed the difference in $I[\mathcal{T}, \mathcal{U}]$ between the two grammars.
-Results are shown in Figure~\ref{fig:toy-parseability}.
 
@@
 
-	\caption{Difference in parseability $I[\mathcal{T}, \mathcal{U}]$ (normalized by sentence length) between the correlating and anti-correlating grammars, for $p_{acl} = 0.0$ (left) and $p_{acl} = 0.3$ (right).
+	\caption{Difference in parseability $I[\mathcal{T}, \mathcal{U}]$ (normalized by sentence length) between the correlating and anti-correlating grammars, for $p_\text{acl} = 0.0$ (left) and $p_\text{acl} = 0.3$ (right).
 Positive values indicate greater values of $I[\mathcal{T}, \mathcal{U}]$ for correlating grammars.}\label{fig:toy-parseability}
 \end{figure}
+
+Results are shown in Figure~\ref{fig:toy-parseability}.
+For all parameter regimes, the correlating grammars have better parseability than the anti-correlating grammars.
+This is especially the case for grammars with high $p_\text{branching}$.
 

### 2019-09-13 writeup/pnas-word-order-si.tex
code

--- 2019-09-12 writeup/pnas-word-order-si.tex
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-arrows-smoothed.pdf}
+\caption[Predictability and Parseability]{Smoothed version with arrows. Is this useful?}\label{fig:pareto-per-lang-arrows}
+\end{figure}
+
 \footnotetext{Note that, in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
@@
 
-Greenberg's original correlations/implications
-
-\cite{justeson1990explanation} conducted a log-linear analysis on typological judgments of 147 languages, constructing an undirected graphical model.
-For the relations examined here, beyond correlations with verb-object order, they found additional correlations between G-N, N-Adp and between Rel-N, N-Adp.
-With their model, they found no evidence for a direct correlation V-O, G-N.
+We followed \citet{dryer1992greenbergian} in treating Greenberg's correlations as pairwise correlations with verb-object order.
+While Greenberg's original study \cite{greenberg1963universals} also formalized most of these as correlations with verb-object order, a few were formalized as correlations between other relations that are only indirectly related to verb-object order (e.g., Universal 22 linking the position of the standard of comparison to the order of adpositions).
+
+\citet{justeson1990explanation} conducted a log-linear analysis on typological judgments of 147 languages, constructing an undirected graphical model.
+%While they found evidence for correlations similar to those found by \citet{dryer1992greenbergian}, they found that some correlations are me
+For the relations examined here, beyond correlations with verb-object order, they found additional correlations between (1) the noun-genitive and adposition-noun dependencies, and (2) between the noun-relative clause and adposition-noun dependencies, \emph{beyond} the correlation mediated through the individual correlations with the verb-object dependency.
+%With their model, they found no evidence for a direct correlation V-O, G-N beyond that mediated through other correlations.
+We ran the same Bayesian logistic mixed-effects analysis for these two correlations.
+Results are shown in Figure~\ref{tab:corr-regression-implication}.
+Both correlations are very strongly supported by grammars optimized for efficiency.
+Future research might apply the graphical model approach of \citet{justeson1990explanation} to optimized grammars, and test whether grammar optimization predicts not just the presence of correlations, the structure their method found on typological data.
+
+\begin{table}
+\small{
+\begin{center}
+\begin{tabular}{|ll||l|lll|llll|ll|llllll}
+\hline
+	& &Prevalence & Mean & SD & $p(\beta \leq 0)$  \\
+\hline\hline
+	G-N (nmod) & N-Adp (lifted\_case)  & 0.919 & 4.482 & 1.058 & $<$ \num{1e-4} \\
+	Rel-N (acl) & N-Adp (lifted\_case) & 0.898 & 4.653 & 1.286 & $<$ \num{1e-4} \\
+\hline
+\end{tabular}
+\end{center}
+}
+	\caption{Detailed results for the two correlations found by \citet{justeson1990explanation} that do not involve the verb-object dependency, for grammars optimized for efficiency.
+Both correlations are strongly supported by optimized grammars, holding in about 90\% of optimized grammars.
+Compare Table~\ref{tab:corr-regression}.
+}\label{tab:corr-regression-implication}
+\end{table}
 

### 2019-09-13 writeup/pnas-word-order-si.tex

--- 2019-09-12 writeup/pnas-word-order-si.tex
@@
 Testing correlations is therefore constrained by the degree to which these relations are annotated in UD.
-The verb-object relation corresponds to the  \emph{obj} relation defined by UD.
+The verb--object relation corresponds to the  \emph{obj} relation defined by UD.
 While most of the other relations also correspond to UD relations, some are not annotated reliably.
@@
 An anonymous reviewer notes that the mapping between Dryer's relations and UD is not perfect, since some of the UD relations subsume other relations.
-Here we show that this is not a problem, since the ordering of the various relations subsumed under the UD label typologically strongly agree.
+Here we show that this is not a problem, since the ordering of the various relations subsumed under the UD label strongly agree typologically.
 
@@
 
-We used the UD treebanks to confirm that different auxiliaires tend to pattern together, and that the most frequent order of the \emph{aux} relation coincides with that of inflected tense-aspect or negation auxiliaries.
-
-We collected, for each UD language, all dependents of the \emph{aux} dependency, occurring at least 10 times, and compared their dominant orders, which we operationalized as their more common order in the treebank (auxiliary-head or head-auxiliary).
+We used the UD treebanks to confirm that different auxiliaries tend to pattern together, and that the most frequent order of the \emph{aux} relation coincides with that of inflected tense-aspect or negation auxiliaries.
+
+We collected, for each UD language, all dependents of the \emph{aux} dependency, occurring at least 10 times, and compared their dominant orders, which we operationalized as their more common order in the treebank (auxiliary--head or head--auxiliary).
 The dependency occurs in all but two very small treebanks (Telugu and Irish).
@@
 
-\section{Formalizing Comunicative Efficiency}
+\section{Formalizing Communicative Efficiency}
 \subsection{Derivation and Relation to Other Work}
@@
 \begin{equation}\label{eq:eff-basic}
-\text{Informativity} - \lambda \cdot \text{Complexity}
+\text{Informativity} - \lambda \cdot \text{Complexity},
 \end{equation}
@@
 \begin{equation}\label{eq:eff-general}
-\operatorname{I}[S, R] - \lambda \cdot \operatorname{H}[S]
+\operatorname{I}[S, R] - \lambda \cdot \operatorname{H}[S],
 \end{equation}
@@
 \begin{equation}\label{eq:efficiency-derived}
-	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda \cdot R_\textit{Pred}
+	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda \cdot R_\textit{Pred},
 \end{equation}
@@
 \begin{equation}
-	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
+	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)},
 \end{equation}
@@
 \begin{equation}
-	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
+	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u).
 \end{equation}
 Parseability $\operatorname{I}[\utterance,\tree]$ is higher if utterances provide more information about their underlying syntactic structure.
-Due to the identity $\operatorname{I}[\utterance, \tree] = \operatorname{H}[\tree] - \operatorname{H}[\tree|\utterance]$, parseability is maximized if every utterance can be parsed unambiguously -- that is, if the listener's uncertainty about syntactic structures given received utterances, $\operatorname{H}[\tree|\utterance]$, is zero.
+Due to the identity $\operatorname{I}[\utterance, \tree] = \operatorname{H}[\tree] - \operatorname{H}[\tree|\utterance]$, parseability is maximized if every utterance can be parsed unambiguously---that is, if the listener's uncertainty about syntactic structures given received utterances, $\operatorname{H}[\tree|\utterance]$, is zero.
 Predictability $- \operatorname{H}[\utterance]$ is higher if the distribution over utterances is concentrated on a few utterances, and is maximized if there is just a single utterance.
@@
 \begin{equation}\label{eq:efficiency}
-	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
+	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred},
 \end{equation}
@@
 %\end{equation}
-and it takes the maximally possible value zero if there is only a single utterance $\utterance$, in which case both $\operatorname{I}[\utterance; \tree]$ and $\operatorname{H}[\utterance]$ are zero.
+and it takes the maximal possible value of zero if there is only a single utterance $\utterance$, in which case both $\operatorname{I}[\utterance; \tree]$ and $\operatorname{H}[\utterance]$ are zero.
 This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
@@
 
-Greenberg's original correlations/implications
+Greenberg's original correlations/implications % ?? TODO
 
@@
 \begin{equation}\label{eq:efficiency-recall}
-	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
+	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred},
 \end{equation}
@@
 \begin{equation}
-	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
+	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u),
 \end{equation}
@@
 \begin{equation}
-	R_{Pars}^{\phi}(u,t) =  \log p_\phi(t|u)
+	R_{Pars}^{\phi}(u,t) =  \log p_\phi(t|u).
 \end{equation}
@@
 \begin{equation}
-	R_{Pred}^{\psi}(u) = \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1})
+	R_{Pred}^{\psi}(u) = \sum_{i=1}^{|u|} \log p_\psi(u_i|u_{1\dots i-1}).
 \end{equation}
@@
 \begin{equation}
-	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi}
+	R_{\textit{Eff}}^{\theta} := \max_{\phi,\psi} R_{\textit{Eff}}^{\theta, \phi, \psi},
 \end{equation}\label{eq:efficiency-rewrite}
@@
 \begin{equation}
-R_{\textit{Eff}}^{\theta, \phi, \psi} := \E_{t \sim \mathcal{T}} \E_{u \sim p_{\mathcal{L}_\theta}(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
+R_{\textit{Eff}}^{\theta, \phi, \psi} := \E_{t \sim \mathcal{T}} \E_{u \sim p_{\mathcal{L}_\theta}(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right].
 \end{equation}
@@
 \partial_\phi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_{\mathcal{L}_\theta}(u|t)}  \left[\partial_\phi R_{Pars}^{\phi}(u,t)\right] \\
-\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_{\mathcal{L}_\theta}(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi}
+\partial_\psi R_{\textit{Eff}}^{\theta, \phi, \psi} &= \E_t \E_{u \sim p_{\mathcal{L}_\theta}(u|t)}  \left[\lambda \cdot \partial_\psi R_{Pred}^{\psi}(u)\right] \label{eq:dpsi},
 \end{align}
 where (\ref{eq:dtheta}) is derived using the \emph{score-function} or \emph{REINFORCE} theorem~\cite{williams1992simple}.
-Note that the the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures.
+Note that the derivatives inside the expectations on the right hand sides can all be computed using backpropagation for our neural network architectures.
 
@@
 
-Optimizing for only parseability (or predictability) is very similar -- in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.
+Optimizing for only parseability (or predictability) is very similar---in this case, the terms involving $R_{Pred}^\phi$ (or $R_{Pars}^\psi$) are removed.
 
@@
 All word and POS embeddings are randomly initialized with uniform values from $[-0.01, 0.01]$.
-We do not use pretrained embeddings \citep{peters2018deep}: While these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
+We do not use pretrained embeddings \citep{peters2018deep}; while these could improve performance of language models and parsers, they would introduce confounds from the languages' actual word orders as found in the unlabeled data.
 
@@
 As described in the main paper, for each language, we created 8 optimized languages for each optimization criterion.
-We enforced balanced distribution of object-verb and verb-object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
+We enforced balanced distribution of object--verb and verb--object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
 This maximizes statistical precision in detecting and quantifying correlations between the verb-object relation and other relations.
@@
 Here we address this question by demonstrating that our languages optimized for efficiency are also optimal under a range of different language models and parsers.
-These results show that our optimization process creates language in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.
+These results show that our optimization process creates languages in which strings are generally predictable and informative about trees, without dependence on particular prediction and parsing algorithms.
 
@@
 
-\subsection{n-gram language models}
+\subsection{$n$-gram language models}
 
 We model predictability using LSTM language models, which are are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
-In previous work, such as \cite{gildea2015human}, predictability has often been measured using n-gram models.
-
-Here, we show that languages optimized for LSTM predictability are also optimal for n-gram predictability.
+In previous work, such as \cite{gildea2015human}, predictability has often been measured using $n$-gram models.
+
+Here, we show that languages optimized for LSTM predictability are also optimal for $n$-gram predictability.
 Specifically, we constructed bigram models with Kneser-Ney smoothing~\cite{kneser1995improved, chen1999empirical}.
@@
 If the difference between random and real grammars is due to data sparsity in this way, then we expect that the difference will decrease as the amount of training data is increased.
-If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference will persists as training data is increased.
-
-We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 Million words), up to $\approx$ 300 times more data than is available for some other UD languages.
+If, on the other hand, there is an inherent difference in efficiency between random and real grammars, we expect that the difference will persist as training data is increased.
+
+We considered Czech, the UD language with the largest amount of available treebank data (approx. 2.2 million words), up to $\approx$ 300 times more data than is available for some other UD languages.
 We considered both a random ordering grammar, and the best ordering grammar optimized for parseabaility.

### 2019-09-13 writeup/pnas-word-order-si.tex

@@
 
-Greenberg's original correlations/implications % ?? TODO
-
-\cite{justeson1990explanation} conducted a log-linear analysis on typological judgments of 147 languages, constructing an undirected graphical model.
-For the relations examined here, beyond correlations with verb-object order, they found additional correlations between G-N, N-Adp and between Rel-N, N-Adp.
-With their model, they found no evidence for a direct correlation V-O, G-N.
+We followed \citet{dryer1992greenbergian} in treating Greenberg's correlations as pairwise correlations with verb-object order.
+While Greenberg's original study \cite{greenberg1963universals} also formalized most of these as correlations with verb-object order, a few were formalized as correlations between other relations that are only indirectly related to verb-object order (e.g., Universal 22 linking the position of the standard of comparison to the order of adpositions).
+
+\citet{justeson1990explanation} conducted a log-linear analysis on typological judgments of 147 languages, constructing an undirected graphical model.
+%While they found evidence for correlations similar to those found by \citet{dryer1992greenbergian}, they found that some correlations are me
+For the relations examined here, beyond correlations with verb-object order, they found additional correlations between (1) the noun-genitive and adposition-noun dependencies, and (2) between the noun-relative clause and adposition-noun dependencies, \emph{beyond} the correlation mediated through the individual correlations with the verb-object dependency.
+%With their model, they found no evidence for a direct correlation V-O, G-N beyond that mediated through other correlations.
+We ran the same Bayesian logistic mixed-effects analysis for these two correlations.
+Results are shown in Figure~\ref{tab:corr-regression-implication}.
+Both correlations are very strongly supported by grammars optimized for efficiency.
+Future research might apply the graphical model approach of \citet{justeson1990explanation} to optimized grammars, and test whether grammar optimization predicts not just the presence of correlations, the structure their method found on typological data.
+
+\begin{table}
+\small{
+\begin{center}
+\begin{tabular}{|ll||l|lll|llll|ll|llllll}
+\hline
+	& &Prevalence & Mean & SD & $p(\beta \leq 0)$  \\
+\hline\hline
+	G-N (nmod) & N-Adp (lifted\_case)  & 0.919 & 4.482 & 1.058 & $<$ \num{1e-4} \\
+	Rel-N (acl) & N-Adp (lifted\_case) & 0.898 & 4.653 & 1.286 & $<$ \num{1e-4} \\
+\hline
+\end{tabular}
+\end{center}
+}
+	\caption{Detailed results for the two correlations found by \citet{justeson1990explanation} that do not involve the verb-object dependency, for grammars optimized for efficiency.
+Both correlations are strongly supported by optimized grammars, holding in about 90\% of optimized grammars.
+Compare Table~\ref{tab:corr-regression}.
+}\label{tab:corr-regression-implication}
+\end{table}
 
@@
 Results are shown in Figure~\ref{fig:toy-parseability}.
-For all parameter regimes, the correlating grammars have better parseability than the anti-correlating grammars.
+For almost all parameter regimes, the correlating grammars have better parseability than the anti-correlating grammars.
 This is especially the case for grammars with high $p_\text{branching}$.

### 2019-09-16 writeup/pnas-word-order-si.tex
SI

@@
 \item Correlation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}} is formalized using \emph{nmod} which covers not only genitives, but also all other noun-modifying NPs and PPs.
-	The evaluation of extracted grammars against WALS (Table~\ref{tab:grammars-wals}) shows that, among the 37 languages where WALS has an entry, the dominant direction of \emph{nmod} agrees with that of genitives, with two exceptions (Danish and Swedish).
-
-TODO figure out why
+	The evaluation of extracted grammars against WALS (Table~\ref{tab:grammars-wals}) shows that, among the 37 languages where WALS has an entry, the dominant direction of \emph{nmod} agrees with that of genitives, with two exceptions (Danish and Swedish\footnote{Danish and Swedish have genitives preceding the head marked with -\emph{s} similar to English (reflected in the WALS entry), while noun-modifying PPs, including phrases similar to English \emph{of} phrases, follow the head.
+In these two languages, the order of adnominal PPs, agreeing with the more frequent order of \emph{nmod} relations, agrees with the verb-object relation, whereas prenominal -s genitives show the opposite ordering.}.
 
@@
 
+\subsection{Per-Language Results}
 In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.
@@
 \end{table}
+
+\subsection{Analysis controlling for Families}
+
+The UD treebanks overrepresent certain language families.
+Does this impact the results of Study 1?
+We estimate the overall degree of optimization of languages for efficiency, controlling for differences between families, we constructed a Bayesian logistic mixed-effects model estimating, for each language $L$ among the 51 UD languages, the rate $q_L \in [0,1]$ of random baseline grammars that have lower efficiency (parseability, predictability) than the real grammar.
+We entered languages and language families as random effects:
+\begin{equation}\label{eq:mixed-effects-study1}
+logit(q_{L}) = \beta + u_{L} + v_{f_L}
+\end{equation}
+where $f_L$ is the language family of $L$.
+
+Results for the posterior of $\beta$ are shown in Table~\ref{tab:study1}.
+Across language families, real grammars tend to be more efficient than almost all baseline grammars.
+
+\begin{table}
+\small{
+\begin{center}
+\begin{tabular}{|l||l|lll|llll|ll|llllll}
+\hline
+	& Mean $\beta$ & SD & Lower 95\% CrI & Upper 95\% CrI \\
+\hline\hline
+	Efficiency ($\lambda=0.9$) &	-5.88   &   1.08  &  -8.28  &  -3.97  \\
+	Surp & -3.48   &   0.88  &  -5.42  &  -1.85   \\
+	Pars & -5.55   &   1.08  &  -7.80  &  -3.67    \\
+\hline
+\end{tabular}
+\end{center}
+}
+	\caption{TODO}\label{tab:study1}
+\end{table}
+% paretoByFamily.R
+
+\subsection{Quantifying Pareto-Optimality}
+
+\paragraph{Inferring $\lambda$}
+We choose $\lambda$ to minimize the number of baselines that have better values of $R_\lambda := R_{Pars} + \lambda R_{Pred}$ than the real language:
+\begin{equation}
+	O := \min_\lambda \# \{b \in B: R_{eff}(b) > R_{eff}(r)\}
+\end{equation}
+where $r$ is the real language, and $B$ is the set of baselines.
+Naively choosing $\lambda$ to minimize this quantity from the observed $50$ baseline samples will bias $O$ towards overestimating the optimality of the real language.
+
+TODO maybe add a figure explaining the Bayesian story a bit more -- how we're using a continuous distribution to smooth out the observed samples?
+
+We eliminate this anti-conservative bias by conducting joint Bayesian inference of $\lambda$ of $O$, using the Dirichlet Gaussian Process, a Bayesian version of Kernel Density Smoothing.
+The 50 samples represent an underlying baseline distribution in the predictability-parseability plane.
+We approximate this underlying distribution as a continuous distribution from the Dirichlet Gaussian Process, and infer its parameters using Bayesian inference.
+We use this method to obtain a Bayesian estimate of the distribution of baseline distributions in the Surprisal-Parseability plane.
+The Dirichlet Gaussian Process (CITE) defines a prior over possible baseline distributions.
+Under this prior, each possible distribution is parameterized by a number $n$ of components, and $n$ Gaussians $\mu_n, \sigma^2_n$.
+Due to the flexible number of components, this nonparametric prior can accomodate a wide range of continuous probability distributions, including multimodal ones.
+We use the $50$ observed baseline samples to infer a posterior distribution over such smooth distributions.
+We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
+We used the MCMC implementation provided by the \texttt{R} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples.
+
+For every sampled distribution -- consisting of a component number $n$ and $n$ Gaussians, we then computed 
+\begin{equation}
+	O := \min_\lambda P(R_{eff}(b) > R_{eff}(r))\}
+\end{equation}
+where we considered $\lambda \in (0,1)$ in steps of $0.5$.
+
+By taking the values $O, \lambda$ for the 1,000 posterior samples, we obtained an estimate of the posterior of $O$ and $\lambda$.
+
+\paragraph{Analysis across Values of $\lambda$}
+
+We quantify the Pareto-optimality of the language $r$ as the fraction of baselines that are better than $r$ on \emph{both} parseability and predictability:
+\begin{equation}
+	PO := \frac{\# \{b \in B: R_{Pars}(b) > R_{eff}(r) \text{ and }  R_{Pred}(b) > R_{Pred}(r) \}}{\#B}
+\end{equation}
+This quantifies how many hypothetical grammars are more efficient than $r$ \emph{irrespective} of $\lambda$.
+
+Again, we use the Dirichlet Gaussian Process to obtain an estimate of the posterior of $PO$.
+We used the same procedure to also estimate the curve connecting points $x$ in the predictability-parseability that have the same value $PO(x) = PO(r)$ -- this curve represents which baselines 
+Again, due to the smoothing implied by the Dirichlet Gaussian Process model, this curve connects not observed baseline samples, but hypothetical other baselines in the predictability-parseability plane.
+
+- identify lambda and quantify what fraction of the baselines are less efficient at this lambda (this is essentially the same as what Zaslavsky et al 2018 do)
+
+- also to make sure this isn't just because lambda can be chosen, quantify what fraction of baselines Pareto-dominate the same fraction
 

### 2019-09-17 writeup/pnas-word-order-si.tex
code

@@
 
+TODO - say that, beyond just what Dryer reported, we see this distribution in the UD languages, and reproduce this
+
 \begin{enumerate}
@@
 
-	\item Correlation \multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}} is formalized using \emph{nmod}, which covers not only PPs and standards of comparison, but also adjunct NPs.
+	\item Correlation \multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}} is formalized using \emph{obl}, which covers not only PPs and standards of comparison, but also adjunct NPs.
 		In the WALS evaluation (Table~\ref{tab:grammars-wals}), the dominant order of \emph{nmod} agrees with that annotated for obliques in all 18 languages for which WALS has an entry.
@@
 
+% say this somewhere
+Note that the predictability scores are considerably worse than the state-of-the art in Natural Language Processing; this is because most of the corpora are much smaller than those typically used in language modeling work.
+
 \begin{figure}
@@
 For more meaningful comparison, both quantities are normalized by the number of words in the corpus, i.e., we plot per-word negative surprisal and per-word difficulty in recovering syntactic structures.
-Note that the negative syntactic ambiguity $-H[T|U]$ equals parseability $I[T,U] = H[T] - H[T|U]$ up to a per-language constant $H[T]$, which we do not attempt to estimate.}\label{fig:pareto-per-lang-raw}
+Note that the negative syntactic ambiguity $-H[T|U]$ equals parseability $I[T,U] = H[T] - H[T|U]$ up to a per-language constant $H[T]$, which we do not attempt to estimate.
+Further note that we use different scales in the different panels.
+}\label{fig:pareto-per-lang-raw}
 \end{figure}
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/figures/quantileByLambda.pdf}
+	\caption[]{The $x$-axis shows $\lambda \in [0,1)$, the $y$-axis shows the fraction of baselines that have lower efficiency at this value of $\lambda$, with 95\% confidence bands obtained from a binomial test.}\label{fig:lambda-quantile}
+\end{figure}
+
+\paragraph{Optimality as a Function of $\lambda$}
+
+Figure~\ref{fig:lambda-quantile}
+
 \paragraph{Inferring $\lambda$}
@@
 We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
-We used the MCMC implementation provided by the \texttt{R} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples.
+We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples.
 
@@
 \end{equation}
-This quantifies how many hypothetical grammars are more efficient than $r$ \emph{irrespective} of $\lambda$.
+This quantifies how many hypothetical grammars are more efficient than $r$ for \emph{all} choices of $\lambda$.
 

### 2019-09-17 writeup/pnas-word-order-si.tex
SI

@@
 
-TODO figure out why
-
-	\item Correlation \multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}} is formalized using \emph{obl}, which covers not only PPs and standards of comparison, but also adjunct NPs.
+The exceptions mainly arise because some languages have multiple common word orders for relativization:
+Hindi uses correlatives that can precede or follow the coreferent noun \citep[3.1.3]{montaut2004grammar} and relatives following the noun \citep[4.3]{montaut2004grammar}.
+Estonian and Finnish have finite relative clauses following nouns (\citep[p.
+176]{karlsson2013finnish}, \citep[p.
+256]{tauli1983standard}) and nonfinite participial modifiers preceding it \citep[Chapter 18]{karlsson2013finnish}.
+
+		Finally, in Tamil, the divergence is caused by the treebank annotation convention for Tamil, where the label \emph{acl} is used to mark certain elements of compounds whose positioning differs from the participial phrases that correspond most closely to relative clauses of other languages.\footnote{In the original HamleDT \citep{zeman2012hamledt, Ramasamy:2011:TDP:1964799.1964808} version of the Tamil treebank, these relations were labeled as CC, marking compounds (\url{http://ufal.mff.cuni.cz/~ramasamy/tamiltb/0.1/dependency_annotation.html}).
+We did not attempt to modify this labeling convention.}
+
+\item Correlation {\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}} is formalized using \emph{obl}, which covers not only PPs and standards of comparison, but also adjunct NPs.
 In the WALS evaluation (Table~\ref{tab:grammars-wals}), the dominant order of \emph{nmod} agrees with that annotated for obliques in all 18 languages for which WALS has an entry.
 
-	\item Correlation \multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}} is formalized using \emph{xcomp}, which covers other control verbs, not just verbs of volition (`want').
+\item Correlation {\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}} is formalized using \emph{xcomp}, which covers other control verbs, not just verbs of volition (`want').
 
@@
 In three historical languages (Ancient Greek, Latin, and Old Church Slavonic), verbs of volition agree with the dominant order of \emph{xcomp}, while several other verbs that do not indicate volition show opposite dominant order.
-Finally, in Gothic, the verb of volition (\emph{wiljan}) has opposite order, resulting in an apparent violation of Correlation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}.
+Finally, in Gothic, the verb of volition (\emph{wiljan}) has opposite dominant order, resulting in an apparent violation of Correlation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}.
 
@@
 \caption[Predictability and Parseability]{Predictability and parseability of 51 languages.
-Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.}\label{fig:pareto-per-lang}
+Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.
+TODO adapt for colorblindness.}\label{fig:pareto-per-lang}
 \end{figure}
@@
 \paragraph{Analysis across Values of $\lambda$}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-isoCurve.pdf}
+\caption{Per-language results.
+Orange curves connect points estimated to be dominated by the same number of baselines as the real language, obtained by Bayesian inference using a Dirichlet Gaussian Process model.}
+\end{figure}
 

### 2019-09-24 writeup/pnas-word-order-si.tex

@@
 
+Two ways of quatifying optimality
+
 \paragraph{Optimality as a Function of $\lambda$}
@@
 
-\paragraph{Inferring $\lambda$}
-We choose $\lambda$ to minimize the number of baselines that have better values of $R_\lambda := R_{Pars} + \lambda R_{Pred}$ than the real language:
-\begin{equation}
-	O := \min_\lambda \# \{b \in B: R_{eff}(b) > R_{eff}(r)\}
-\end{equation}
-where $r$ is the real language, and $B$ is the set of baselines.
-Naively choosing $\lambda$ to minimize this quantity from the observed $50$ baseline samples will bias $O$ towards overestimating the optimality of the real language.
-
-TODO maybe add a figure explaining the Bayesian story a bit more -- how we're using a continuous distribution to smooth out the observed samples?
-
-We eliminate this anti-conservative bias by conducting joint Bayesian inference of $\lambda$ of $O$, using the Dirichlet Gaussian Process, a Bayesian version of Kernel Density Smoothing.
-The 50 samples represent an underlying baseline distribution in the predictability-parseability plane.
-We approximate this underlying distribution as a continuous distribution from the Dirichlet Gaussian Process, and infer its parameters using Bayesian inference.
-We use this method to obtain a Bayesian estimate of the distribution of baseline distributions in the Surprisal-Parseability plane.
-The Dirichlet Gaussian Process (CITE) defines a prior over possible baseline distributions.
-Under this prior, each possible distribution is parameterized by a number $n$ of components, and $n$ Gaussians $\mu_n, \sigma^2_n$.
-Due to the flexible number of components, this nonparametric prior can accomodate a wide range of continuous probability distributions, including multimodal ones.
-We use the $50$ observed baseline samples to infer a posterior distribution over such smooth distributions.
-We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
-We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples.
-
-For every sampled distribution -- consisting of a component number $n$ and $n$ Gaussians, we then computed 
-\begin{equation}
-	O := \min_\lambda P(R_{eff}(b) > R_{eff}(r))\}
-\end{equation}
-where we considered $\lambda \in (0,1)$ in steps of $0.5$.
-
-By taking the values $O, \lambda$ for the 1,000 posterior samples, we obtained an estimate of the posterior of $O$ and $\lambda$.
+%\paragraph{Inferring $\lambda$}
+%We choose $\lambda$ to minimize the number of baselines that have better values of $R_\lambda := R_{Pars} + \lambda R_{Pred}$ than the real language:
+%\begin{equation}
+%	O := \min_\lambda \# \{b \in B: R_{eff}(b) > R_{eff}(r)\}
+%\end{equation}
+%where $r$ is the real language, and $B$ is the set of baselines.
+%Naively choosing $\lambda$ to minimize this quantity from the observed $50$ baseline samples will bias $O$ towards overestimating the optimality of the real language.
+%
+%TODO maybe add a figure explaining the Bayesian story a bit more -- how we're using a continuous distribution to smooth out the observed samples?
+%
+%We eliminate this anti-conservative bias by conducting joint Bayesian inference of $\lambda$ of $O$, using the Dirichlet Gaussian Process, a Bayesian version of Kernel Density Smoothing.
+%The 50 samples represent an underlying baseline distribution in the predictability-parseability plane.
+%We approximate this underlying distribution as a continuous distribution from the Dirichlet Gaussian Process, and infer its parameters using Bayesian inference.
+%We use this method to obtain a Bayesian estimate of the distribution of baseline distributions in the Surprisal-Parseability plane.
+%The Dirichlet Gaussian Process (CITE) defines a prior over possible baseline distributions.
+%Under this prior, each possible distribution is parameterized by a number $n$ of components, and $n$ Gaussians $\mu_n, \sigma^2_n$.
+%Due to the flexible number of components, this nonparametric prior can accomodate a wide range of continuous probability distributions, including multimodal ones.
+%We use the $50$ observed baseline samples to infer a posterior distribution over such smooth distributions.
+%We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
+%We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples.
+%
+%For every sampled distribution -- consisting of a component number $n$ and $n$ Gaussians, we then computed 
+%\begin{equation}
+%	O := \min_\lambda P(R_{eff}(b) > R_{eff}(r))\}
+%\end{equation}
+%where we considered $\lambda \in (0,1)$ in steps of $0.5$.
+%
+%By taking the values $O, \lambda$ for the 1,000 posterior samples, we obtained an estimate of the posterior of $O$ and $\lambda$.
 

### 2019-09-24 writeup/pnas-word-order-si.tex
SI

@@
 
-Two ways of quatifying optimality
-
-\paragraph{Optimality as a Function of $\lambda$}
-
-Figure~\ref{fig:lambda-quantile}
+We consider two methods of quantifying the \key{degee of Pareto optimality}.
+First, we quantify, for each value of $\lambda$ trading off parseability and predictability, how many baseline grammars are less efficient than the real language.
+Second, we quantify how many baseline languages are more efficient across all values of $\lambda$.
+
+Reecall the efficiency metric $R_\lambda := R_{Pars} + \lambda R_{Pred}$ with the tradeoff parameter $\lambda \in [0,1)$.
+For each $\lambda$, we can quantify the fraction of baselines that have better efficiency than the real language
+\begin{equation}
+	O_\lambda := \frac{\# \{b \in B: R_{\lambda}(b) > R_{\lambda}(r)\}}{\# B}
+\end{equation}
+where $B$ is the set of baseline languages (make figure).
+
+The results are plotted in Figure~\ref{fig:lambda-quantile}.
 
@@
 \centering
-\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-isoCurve.pdf}
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-isoCurve-smooth.pdf}
 	\caption{Per-language results.
-Orange curves connect points estimated to be dominated by the same number of baselines as the real language, obtained by Bayesian inference using a Dirichlet Gaussian Process model.}
+Baseline distributions are estimated using Kernel Density estimation.
+Orange curves connect points estimated to be dominated by the same number of baselines as the real language, obtained by Bayesian inference using a Dirichlet Gaussian Process model.}\label{fig:isoline}
 \end{figure}
+
+We also considered a second method of quantifying the degree of Pareto optimality.
 
@@
 \end{equation}
-This quantifies how many hypothetical grammars are more efficient than $r$ for \emph{all} choices of $\lambda$.
-
-Again, we use the Dirichlet Gaussian Process to obtain an estimate of the posterior of $PO$.
-We used the same procedure to also estimate the curve connecting points $x$ in the predictability-parseability that have the same value $PO(x) = PO(r)$ -- this curve represents which baselines 
-Again, due to the smoothing implied by the Dirichlet Gaussian Process model, this curve connects not observed baseline samples, but hypothetical other baselines in the predictability-parseability plane.
-
-- identify lambda and quantify what fraction of the baselines are less efficient at this lambda (this is essentially the same as what Zaslavsky et al 2018 do)
-
-- also to make sure this isn't just because lambda can be chosen, quantify what fraction of baselines Pareto-dominate the same fraction
+This quantifies how many hypothetical grammars are more efficient than $r$ for \emph{all} choices of $\lambda$ (make a figure).
+
+This quantity is also known as the joint (or bivariate) exceedance probability.
+
+If we have access to the full distribution over baseline grammars, then the Pareto frontier is formed by exactly those grammars where $PO=0$ (make a figure).
+In practice, we only have a limited sample of $50$ baseline grammar samples; so only an approximation of $PO$ is actually observable, and this approximation will be $0$ in a larger region of the plane (make a figure).
+
+We aim to estimate $PO$, provide uncertainty intervals, and compare with the values of $PO$ attained by baseline languages, and estimate the isoline (cotour line)  of grammars having the same $PO$ value as the real language.
+
+In order to estimate $PO$ from the 50 observed baseline samples, we did the following.
+
+The baseline languages form a distribution in the parseability-predictability plane, from which 50 independent samples are known.
+We model the underlying distribution using the Dirichlet Process Gaussian Mixture Model (DPGMM), a flexible nonparametric model that can accomodate multimodal and other non-normal distributions \citep{rasmussen2000infinite, escobar1995bayesian, muller1996bayesian}.
+This model is a Gaussian mixture model where the number of components itself is a latent parameter inferred from the data.
+DPGMMs can be seen as a Bayesian version of Kernel Density Smoothing.
+In our context here, the reason for using DPGMMs instead of Kernel Density Smoothing is that Bayesian inference using the DPGMM provides a straightforward method of deriving estimates and credible intervals for $PO$ and the isoline.
+
+% http://mlg.eng.cam.ac.uk/pub/pdf/GoeRas10.pdf
+% TODO report prior
+
+The DPGMM defines a prior over possible baseline distributions in the parseability-predictability plane.
+%Under this prior, each possible distribution is parameterized by a number $n$ of components, $n$ bivariate Gaussians $\mu_i, \Sigma_i$, and $n$ weights $w_i$ for these components ($\sum_{i=1}^n w_i=1$).
+Each possible distribution is parameterized as $\sum_{i=1}^n w_i p_{\mu_i, \Sigma_i}(x)$ where $p_{\mu_i, \Sigma_i}$ is the bivariate Gaussian density.
+The number $n$ of components is inferred from the data together with the weights $w_i$, the means $\mu_i$, and the covariance matrices $\Sigma_i$.
+A Dirichlet process prior (TODO) is placed on $n$, favoring models with smaller numbers of components unless the data provides evidence for more components.
+We use the $50$ observed baseline samples to infer a posterior distribution over such mixture distributions.
+We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
+We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples.
+
+For each of these samples, we then computed $PO$ of the real language, and the isoline.
+The collection of these results provides a set of samples from the posterior distribution over the true value of $PO$ and the true isoline.
+We plot the posterior means of $PO$ and the isoline in Figure~\ref{fig:isoline}.
+
+%We eliminate this anti-conservative bias by conducting joint Bayesian inference of $\lambda$ of $O$, using the Dirichlet Gaussian Process, a Bayesian version of Kernel Density Smoothing.
+%The 50 samples represent an underlying baseline distribution in the predictability-parseability plane.
+%We approximate this underlying distribution as a continuous distribution from the Dirichlet Gaussian Process, and infer its parameters using Bayesian inference.
+%We use this method to obtain a Bayesian estimate of the distribution of baseline distributions in the Surprisal-Parseability plane.
+%The Dirichlet Gaussian Process (CITE) defines a prior over possible baseline distributions.
+%Under this prior, each possible distribution is parameterized by a number $n$ of components, and $n$ Gaussians $\mu_n, \sigma^2_n$.
+
+%We assumed that the distribution
+
+%Again, we use the Dirichlet Gaussian Process to obtain an estimate of the posterior of $PO$.
+%We used the same procedure to also estimate the curve connecting points $x$ in the predictability-parseability that have the same value $PO(x) = PO(r)$ -- this curve represents which baselines 
+%Again, due to the smoothing implied by the Dirichlet Gaussian Process model, this curve connects not observed baseline samples, but hypothetical other baselines in the predictability-parseability plane.
+
+%- identify lambda and quantify what fraction of the baselines are less efficient at this lambda (this is essentially the same as what Zaslavsky et al 2018 do)
+
+%- also to make sure this isn't just because lambda can be chosen, quantify what fraction of baselines Pareto-dominate the same fraction
 
@@
 For comparison, we also show results for $\lambda=0$, corresponding to optimizing for parseability only (same results as reported in Tables (\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})).
-For $\lambda=0.9$, we report results from one preliminary preregistered experiment (left) and the final experiment (right).
+For $\lambda=0.9$, we report results from one preliminary preregistered experiment (center left) and the final experiment (center right).
 For $\lambda=1.0$, we report the other preliminary preregistered experiment.

### 2019-09-24 results/plane/analyze_pareto_optimality/bayesianAnalysisIsoQuantile2.R

@@ -29,7 +29,7 @@ for(language in unique(data$Language)) {
     
     library(dirichletprocess)
     dp = DirichletProcessMvnormal(X)
-    dp = Fit(dp, 2000)
+    dp = Fit(dp, 1105)
     
     bestLambdas = c()
     bestQuantiles = c()
@@ -57,7 +57,7 @@ for(language in unique(data$Language)) {
        GRID_SIZEs = length(surpGrid_)
 
 
-    for(i in (1:3)) {
+    for(i in (1:100)) {
        cat(language, i, "\n")
        j = 1000 + 50 * i
        weight = dp$weightsChain[[j]]
@@ -95,11 +95,11 @@ for(language in unique(data$Language)) {
       totalSurpGridBySample = rbind(totalSurpGridBySample, curves[[i]] %>% mutate(sample=i))
     }
     library(ggplot2)
-#    isoCurve = data.frame(x=parsGrid_, y=totalSurpGrid/length(curves))
- #   boundary = nrow(isoCurve %>% filter(y > surpOffset)) + 1
-  #  isoCurve = isoCurve[(1:boundary),]
-   # plot = ggplot(isoCurve, aes(x=-x, y=-y)) + geom_line() + geom_point(data=data.frame(pars=pars, surp=surp), aes(x=-pars, y=-surp)) + geom_point(data=data.frame(pars=c(parsGround), surp=c(surpGround)), aes(x=-pars, y=-surp), color="red")
-    #ggsave(plot, file=paste("figures/isoCurve_",language,".pdf", sep=""))
+    isoCurve = data.frame(x=parsGrid_, y=totalSurpGrid/length(curves))
+    boundary = nrow(isoCurve %>% filter(y > surpOffset)) + 1
+    isoCurve = isoCurve[(1:boundary),]
+    plot = ggplot(isoCurve, aes(x=-x, y=-y)) + geom_line() + geom_point(data=data.frame(pars=pars, surp=surp), aes(x=-pars, y=-surp)) + geom_point(data=data.frame(pars=c(parsGround), surp=c(surpGround)), aes(x=-pars, y=-surp), color="red")
+    ggsave(plot, file=paste("figures/isoCurve_",language,".pdf", sep=""))
 
     write.csv(totalSurpGridBySample, file=paste("../../../grammars/pareto-curves/pareto-smooth/iso-pareto-full-", language, sep=""))
 

### 2019-09-24 results/plane/analyze_pareto_optimality/plane-2D-best-visualization-large-6.R

@@ -1,0 +1,297 @@
+# per language
+
+
+library(lme4)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+depl = read.csv("../../../grammars/dependency_length/total_summary_funchead_coarse.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+library(tidyr)
+library(dplyr)
+depl = depl %>% filter(grepl("FuncHead", ModelName)) %>% filter(grepl("Coarse", ModelName))
+dataS  = read.csv("../../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS3 = read.csv("../../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS4 = read.csv("../../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS5 = read.csv("../../../grammars/plane/plane-fixed-random3.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS6 = read.csv("../../../grammars/plane/plane-fixed-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS7 = read.csv("../../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS = rbind(dataS, dataS2, dataS3, dataS4, dataS5, dataS6, dataS7)
+dataP  = read.csv("../../../grammars/plane/plane-parse.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP2 = read.csv("../../../grammars/plane/plane-parse-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP3 = read.csv("../../../grammars/plane/plane-parse-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model))
[…]

### 2019-09-25 writeup/pnas-word-order-si.tex

@@
 
-\subsection{Quantifying Pareto-Optimality}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/figures/quantileByLambda.pdf}
-	\caption[]{The $x$-axis shows $\lambda \in [0,1)$, the $y$-axis shows the fraction of baselines that have lower efficiency at this value of $\lambda$, with 95\% confidence bands obtained from a binomial test.}\label{fig:lambda-quantile}
-\end{figure}
-
-We consider two methods of quantifying the \key{degee of Pareto optimality}.
-First, we quantify, for each value of $\lambda$ trading off parseability and predictability, how many baseline grammars are less efficient than the real language.
-Second, we quantify how many baseline languages are more efficient across all values of $\lambda$.
-
-Reecall the efficiency metric $R_\lambda := R_{Pars} + \lambda R_{Pred}$ with the tradeoff parameter $\lambda \in [0,1)$.
-For each $\lambda$, we can quantify the fraction of baselines that have better efficiency than the real language
-\begin{equation}
-	O_\lambda := \frac{\# \{b \in B: R_{\lambda}(b) > R_{\lambda}(r)\}}{\# B}
-\end{equation}
-where $B$ is the set of baseline languages (make figure).
-
-The results are plotted in Figure~\ref{fig:lambda-quantile}.
-
 %\paragraph{Inferring $\lambda$}
@@
 
-\paragraph{Analysis across Values of $\lambda$}
+\subsection{Quantifying Degree of Optimality}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/figures/quantileByLambda.pdf}
+	\caption[]{The $x$-axis shows $\lambda \in [0,1)$, the $y$-axis shows the fraction of baselines that have lower efficiency at this value of $\lambda$, with 95\% confidence bands obtained from a binomial test.}\label{fig:lambda-quantile}
+\end{figure}
+
+Efficiency combines predictability and parseability.
+In the main paper (Study 1), we showed that languages tend to be optimized for these two quantities.
+In this section, we show that languages are also optimized for the \emph{combination} of the two objectives.
+
+We quantify the degree of optimality in two ways.
+In Analysis 1, we consider the efficiency metric $R_\lambda := R_{Pars} + \lambda R_{Pred}$ with the tradeoff parameter $\lambda \in [0,1)$.
+For each possible value of $\lambda \in [0,1)$ trading off parseability and predictability, we quantify how many baseline grammars are less efficient than the real language.
+In Analysis 2, we quantify what fraction of baseline languages \emph{Pareto-dominates} the real language, showing that at most a very small fraction of grammars are more efficient than the real language.
+
+\paragraph{Analysis 1: Optimality by $\lambda$}
+%Second, we quantify how many baseline languages are more efficient across all values of $\lambda$.
+
+Recall the efficiency metric $R_\lambda := R_{Pars} + \lambda R_{Pred}$ with the tradeoff parameter $\lambda \in [0,1)$.
+For each $\lambda$, we can quantify the fraction of baselines that have better efficiency than the real language
+\begin{equation}
+	O_\lambda := \frac{\# \{b \in B: R_{\lambda}(b) > R_{\lambda}(r)\}}{\# B}
+\end{equation}
+where $B$ is the set of baseline languages (make figure).
+
+The results are plotted in Figure~\ref{fig:lambda-quantile}.
+
+TODO also visualize and explain
+
+\paragraph{Analysis 2: Quantifying Degree of Pareto Optimality}
+
+\begin{figure}
+\centering
+	\begin{tabular}{llll}
+		(a) & (b) & (c) & (d) \\
+		\includegraphics[width=0.22\textwidth]{figures/isocline1.png}&
+		\includegraphics[width=0.22\textwidth]{figures/isocline2.png}&
+		\includegraphics[width=0.22\textwidth]{figures/isocline3.png}&
+\includegraphics[width=0.22\textwidth]{figures/isocline4.png}
+	\end{tabular}
+
+	\caption{Isolines: (a) The set of all possible baseline grammars forms a distribution in the parseability-predictability space.
+	The Pareto frontier consists of those points that are not dominated by any baseline grammar -- that is, those for which $P=0$: For points on this curve, no baseline grammar is both to their top and their right.
+	(b) Other grammars are dominated by some fractions $P > 0$ of baseline grammars.
+The two green crosses mark grammars that both are dominated by $P=0.05$ of baseline grammars.
+A blue curve connects all grammars dominated by exactly $P=0.05$ of baseline grammars.
+	(c) We observe a sample of $N=50$ baseline grammars, sampled uniformly from the set of baseline grammars. Samples will tend to concentrate in the high-density regions of the baseline distribution.
+	We also observe the efficiency of the real language (red cross) and an estimate of the Pareto frontier (blue line).
+	(d) We use the $50$ samples from the baseline grammar distribution to construct an estimate of the full abseline distibution, using the Dirichlet Process Gaussian Mixture Model.
+Using this inferred distribution, we estimate what fraction $P$ of baseline grammars dominate the real language, and the curve connecting all grammars dominated by the same fraction $P$ of baselines.
+	}\label{fig:isoline-explanation}
+\end{figure}
 
@@
 
-We also considered a second method of quantifying the degree of Pareto optimality.
-
-We quantify the Pareto-optimality of the language $r$ as the fraction of baselines that are better than $r$ on \emph{both} parseability and predictability:
+%The Pareto curve is formed by those grammars such that no grammar simultaneously achieves higher parseability and higher predictability.
+%This motivates
+
+We quantify the Pareto-optimality of the language $r$ as the fraction of baselines that Pareto-dominate $r$ -- that is, baselines that are at least as good as $r$ on \emph{both} parseability and predictability:
 \begin{equation}
-PO := \frac{\# \{b \in B: R_{Pars}(b) > R_{eff}(r) \text{ and }  R_{Pred}(b) > R_{Pred}(r) \}}{\#B}
-\end{equation}
-This quantifies how many hypothetical grammars are more efficient than $r$ for \emph{all} choices of $\lambda$ (make a figure).
-
-This quantity is also known as the joint (or bivariate) exceedance probability.
-
-If we have access to the full distribution over baseline grammars, then the Pareto frontier is formed by exactly those grammars where $PO=0$ (make a figure).
-In practice, we only have a limited sample of $50$ baseline grammar samples; so only an approximation of $PO$ is actually observable, and this approximation will be $0$ in a larger region of the plane (make a figure).
-
-We aim to estimate $PO$, provide uncertainty intervals, and compare with the values of $PO$ attained by baseline languages, and estimate the isoline (cotour line)  of grammars having the same $PO$ value as the real language.
-
-In order to estimate $PO$ from the 50 observed baseline samples, we did the following.
-
+P := \frac{\# \{b \in B: R_{Pars}(b) \geq R_{eff}(r) \text{ and }  R_{Pred}(b) \geq R_{Pred}(r) \}}{\#B}
+\end{equation}
+%This quantifies how many hypothetical grammars are more efficient than $r$ for \emph{all} choices of $\lambda$ (make a figure).
+%This quantity is also known as the joint (or bivariate) exceedance probability.
+If we have access to the full distribution over baseline grammars, then the Pareto frontier is formed by exactly those grammars where $P=0$ (make a figure).
+(TODO slight subtlety: $r$ itself is part of $B$, so $P > 0$ TODO change definition)
+Small values of $P$ indicate that only few baseline grammars are better than $r$.
+%In practice, we only have a limited sample of $50$ baseline grammar samples; so only an approximation of $P$ is actually observable, and this approximation will be $0$ in a larger region of the plane (make a figure).
+
+In this section, we aim to estimate $P$, provide uncertainty intervals.
+In order to also compare with the values of $P$ attained by baseline languages, we also estimated the isoline (cotour line) of grammars having the same $P$ value as the real language.
+
+%$P$ is also known as the \emph{joint} or \emph{bivariate} \emph{exceedance probability}, and there are a range of methods for estimating it.
+
+We used a simple nonparametric Bayesian approach to estimate $P$ from the 50 observed baseline samples.
 The baseline languages form a distribution in the parseability-predictability plane, from which 50 independent samples are known.
@@
 This model is a Gaussian mixture model where the number of components itself is a latent parameter inferred from the data.
-DPGMMs can be seen as a Bayesian version of Kernel Density Smoothing.
+DPGMMs can be seen as a Bayesian version of Kernel Density Smoothing.\footnote{A potential objection to using such Kernel Density Estimation is that it models a continuous distribution with support everywhere, while the distribution of grammars is discrete and finite, in particular bounded by a Pareto frontier.
+We believe that this does not cause problems for this analysis.
+First, the number of possible grammars is extremey large, exceeding $37!$, which is more than $10^{43}$, suggesting that the distribution in the parseability-predictability plane is closely approximated by a continuous distribution.
+Second, the approximate Pareto frontier tends to be clearly separated from the high-density regions of the baseline distribution, in which most sampled baselines fall, showing that the baseline distribution has very small density around and beyond the Pareto frontier.}
 In our context here, the reason for using DPGMMs instead of Kernel Density Smoothing is that Bayesian inference using the DPGMM provides a straightforward method of deriving estimates and credible intervals for $PO$ and the isoline.
@@
 We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
-We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples.
-
-For each of these samples, we then computed $PO$ of the real language, and the isoline.
-The collection of these results provides a set of samples from the posterior distribution over the true value of $PO$ and the true isoline.
-We plot the posterior means of $PO$ and the isoline in Figure~\ref{fig:isoline}.
+We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples, from which we selected an evenly spaced subset of 100 samples.
+
+For each of these samples, we then computed $P$ of the real language, and the isoline.
+The collection of these results provides a set of samples from the posterior distribution over the true value of $P$ and the true isoline.
+We plot the posterior means of $P$ and the isoline in Figure~\ref{fig:isoline}.
 
@@
 
+%Among UD 2.1 corpora, three treebanks require a special license (Arabic-NYUAD, French-FTB).
+%For a few UD corpora, only the syntactic annotation is freely available, while the individual words require a license.
+%For the lexicalized experiment, we had access to the individual words for
+
 \begin{table}[ht]

### 2019-09-25 writeup/si-table-perrel-3.tex

--- 2019-05-29 second-draft/si-table-perrel-3.tex
@@
 \usepackage[numbers]{natbib}
-
-\usepackage{hyperref}
 
@@
 & $\lambda = 0.0$
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
-& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} 
-& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}} 
+& $\lambda=0.9$%\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
+& $\lambda=0.9$%\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} 
+& $\lambda=1.0$%\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}} 
 	\\ \hline \hline 
@@
 		\\
-&&&\\ \hline
+&&&&\\ \hline
 	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	 
@@
 		\\
-&&&\\ \hline
+&&&&\\ \hline
 	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	 
@@
 		\\
-&&&\\ \hline
+&&&&\\ \hline
 	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	
@@
 		\\
-&&&\\ \hline
+&&&&\\ \hline
 	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	 
@@
 		\\
-&&&\\ \hline
+&&&&\\ \hline
 	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	
@@
 		\\
-&&&\\ \hline
+&&&&\\ \hline
 	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	
@@
 		\\
-&&&\\ \hline
+&&&&\\ \hline
 	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	
@@
 		\\
-&&&\\ \hline
+&&&&\\ \hline
  \hline

### 2019-09-29 writeup/pnas-word-order-si.tex
code

@@
 
-\title{Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''}
+\title{Supplemental Materials for ``Universals of word order reflect optimization of grammars for efficient communication''}
 \author{
@@
 Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
-%\begin{equation}
 $R_{\textit{Eff}} = \operatorname{I}[\utterance; \tree] - \lambda \operatorname{H}[\utterance] \leq 0$,
-%\end{equation}
 and it takes the maximal possible value of zero if there is only a single utterance $\utterance$, in which case both $\operatorname{I}[\utterance; \tree]$ and $\operatorname{H}[\utterance]$ are zero.
@@
 This is confirmed by converging evidence from our preregistered preliminary experiments in Figure~\ref{table:corr-resu-previous}.
-
-%Finally, the robustness of our results to small variations of $\lambda$ is confime by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu-previous}.
-
-%In terms of information theory \cite{shannon1948mathematical}, this corresponds to the notion of the rate-distortion tradeoff.
-%In classical rate-distortion theory, the cost of a code is its rate, the mutual information between codes and underlying messages (as in the Information Bottleneck model).
-% classical models of distortion are the Hamming distortion and 
-%TODO also mention rate-distortion theory
-
-%- our function
-
-%similar: RSA, there it's utterance-level
-
-%- number of different forms 
-%
-%- complexity of the mapping 
-%
-%- MI between
-%
-%
-%
-%We model the first term as the 
-% degree to which listeners can reconstruct dependency parse structures from an utterance, i.e., the \key{parseability} if the language.
-This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
-%\begin{equation}
-%	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
-%\end{equation}
-%where the sum runs over all possible pairs of sentences $l$ and syntactic structures $t$ in the language.
-%%This quantity describes the degree to which syntactic structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
-%
-%
-%
-%We formalize the complexity of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
-%In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
-%Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
-%%and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
-%
-%In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
-%\begin{equation}
-%	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
-%\end{equation}
-%where the sum runs over all possible sentences $u$ that belong to the language.
-%%In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
-%
-%
-%
-%The efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
-%\begin{equation}\label{eq:efficiency}
-%	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
-%\end{equation}
-%with an interpolation weight $\lambda \in [0,1)$.
-%In all results from Study 2 reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
-%See SI appendix section 5 for mathematical discussion of $\lambda$, and robustness to other choices.
-%
-%
-
-%\section{Relation to the Information Bottleneck}
-%
-%Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
-%\begin{equation}
-%    \label{eq:efficiency-again}
-%    J_T(L) = \operatorname{I}[L(T); T] - \lambda \operatorname{H}[L(T)].
-%\end{equation}
-%This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}.
-In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information.
-The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$.
-Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
-%\begin{equation}
-%    \label{eq:ib}
-%    J_{X,Y}(L) = \operatorname{I}[\hat{X}; Y] - \lambda \operatorname{I}[\hat{X}; X].
-%\end{equation}
-%In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
-%
-%In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$.
-To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$.
-We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized.
-Then our objective is parallel to Eq.~\ref{eq:ib}:
-%\begin{equation}
-%    \label{eq:efficiency-kinda-ib}
-%    J_T(L) = \operatorname{I}[U;T] - \lambda \operatorname{I}[U;\hat{U}]. % TODO should this be \operatorname{I}[\hat{U};T]?
-%\end{equation}
-%Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $\operatorname{I}[U;\hat{U}]=\operatorname{I}[U;U] = \operatorname{H}[U]$.
-Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
-%\begin{equation*}
-%    J_T(L) = \operatorname{I}[U;T] - \lambda \operatorname{H}[U].
-%\end{equation*}
-%
-%Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
-%
-%
 
@@
 In addition to the $p$-value, we report point estimate and 95\% confidence interval for the fraction of random grammars that have values below real grammars.}\label{fig:pareto-per-lang-stats}
-\end{table}
-
-\begin{table}
-\centering
-\small{
-\begin{tabular}{l||ll|lll|lll}
-Language & Pred. (t) & Parse. (t) & \multicolumn{3}{c|}{Pred. (Binomial)} & \multicolumn{3}{c}{Parseab. (Binomial)} \\ 
-&  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
-\input{../results/plane/unlexicalized/per-language-statistics-unlexicalized.tex}
-\end{tabular}
-}
-	\caption{Per-language results in Study 1, with \emph{lexicalized} parsers and word-level-only language models. Compare Table~\ref{fig:pareto-per-lang-stats}}
 \end{table}
@@
 \centering
+		\includegraphics[width=0.5\textwidth]{lambda-explanation.png}
+
+	\caption{TODO For $\lambda=0$ and $\lambda=0.9$, we provide lines connecting points that have the same efficiency value as the real language (red cross).
+Points to the bottom right of the lines have lower efficiency under the respective values of $\lambda$.
+In this example, all baseline samples (green dots) have lower efficiency than the real language, even though the real language does not outperform every baselines on both components of efficiency.}\label{fig:opt-lambda-explanation}
+\end{figure}
+
+\begin{figure}
+\centering
 \includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/figures/quantileByLambda.pdf}
@@
 
-Efficiency combines predictability and parseability.
-In the main paper (Study 1), we showed that languages tend to be optimized for these two quantities.
-In this section, we show that languages are also optimized for the \emph{combination} of the two objectives.
-
-We quantify the degree of optimality in two ways.
+In the main paper (Study 1), we showed that languages tend to be optimized for parseability and/or predictability.
+Efficiency is a combination of both components; in this section we address the question whether languages are generally optimized for efficiency, i.e., the multi-objective optimization problem of optimizing for parseability and predictability.
+%Note that the numerical value of efficiency depends on the parameter $\lambda \in [0,1)$.
+
+We quantify degree of optimality in two ways.
 In Analysis 1, we consider the efficiency metric $R_\lambda := R_{Pars} + \lambda R_{Pred}$ with the tradeoff parameter $\lambda \in [0,1)$.
 For each possible value of $\lambda \in [0,1)$ trading off parseability and predictability, we quantify how many baseline grammars are less efficient than the real language.
-In Analysis 2, we quantify what fraction of baseline languages \emph{Pareto-dominates} the real language, showing that at most a very small fraction of grammars are more efficient than the real language.
+In Analysis 2, we quantify what fraction of baseline languages \emph{Pareto-dominates} the real language, showing that -- while not all real languages may be strongly optimized for both objectives individually -- at most a very small fraction of baseline grammars improve on the real language on both components.
 
@@
 
-Recall the efficiency metric $R_\lambda := R_{Pars} + \lambda R_{Pred}$ with the tradeoff parameter $\lambda \in [0,1)$.
-For each $\lambda$, we can quantify the fraction of baselines that have better efficiency than the real language
+Recall the efficiency metric
+\begin{equation}
+	R_\lambda := R_{Pars} + \lambda R_{Pred}
+\end{equation}
+ with the tradeoff parameter $\lambda \in [0,1)$.
+For each $\lambda$, we quantify the fraction of baselines that have better efficiency than the real language:
 \begin{equation}
@@
 \end{equation}
-where $B$ is the set of baseline languages (make figure).
+where $B$ is the set of baseline languages.
+We illustrate this in Figure~\ref{fig:opt-lambda-explanation}.
 
@@
 
-TODO also visualize and explain
+For all languages, there are some values of $\lambda$ where the real language improves at least on more than the median of the baseline grammars, i.e., $O_\lambda >0$.
+In about 40 of the languages, $O_\lambda$ is even close to one across all values of $\lambda$.
+
+This shows that, while many languages do not improve over \emph{all} baselines on \emph{both} individual components, they mostly improve over the large majority of baselines on the combined objective of efficiency, even across different values of $\lambda$.
 
 \paragraph{Analysis 2: Quantifying Degree of Pareto Optimality}
+
+TODO maybe we can scrap this part as it is harder to explain well, if Analysis 1 sufficient.
 
@@
 \centering
-\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-isoCurve-smooth.pdf}
-	\caption{Per-language results.
-Baseline distributions are estimated using Kernel Density estimation.
-Orange curves connect points estimated to be dominated by the same number of baselines as the real language, obtained by Bayesian inference using a Dirichlet Gaussian Process model.}\label{fig:isoline}
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-isoCurve-smooth-interval.pdf}
+	\caption{Real languages, even those that are not at the estimated Pareto frontier, are Pareto-dominated by very few other hypothetical grammars.
+Green curves: Baseline distribution, estimated using Kernel Density estimation.
+Black numbers indicate the fraction of baselines dominating the real language, with a 95\% credible interval, estimated using a nonparametric Bayesian method (see text).
+Orange curves connect points estimated to be dominated by the same fraction of baselines as the real language, estimated using the same Bayesian method (see text).
+}\label{fig:isoline}
 \end{figure}
@@
 
-%We eliminate this anti-conservative bias by conducting joint Bayesian inference of $\lambda$ of $O$, using the Dirichlet Gaussian Process, a Bayesian version of Kernel Density Smoothing.
-%The 50 samples represent an underlying baseline distribution in the predictability-parseability plane.
-%We approximate this underlying distribution as a continuous distribution from the Dirichlet Gaussian Process, and infer its parameters using Bayesian inference.
-%We use this method to obtain a Bayesian estimate of the distribution of baseline distributions in the Surprisal-Parseability plane.
-%The Dirichlet Gaussian Process (CITE) defines a prior over possible baseline distributions.
-%Under this prior, each possible distribution is parameterized by a number $n$ of components, and $n$ Gaussians $\mu_n, \sigma^2_n$.
-
-%We assumed that the distribution
-
-%Again, we use the Dirichlet Gaussian Process to obtain an estimate of the posterior of $PO$.
-%We used the same procedure to also estimate the curve connecting points $x$ in the predictability-parseability that have the same value $PO(x) = PO(r)$ -- this curve represents which baselines 
-%Again, due to the smoothing implied by the Dirichlet Gaussian Process model, this curve connects not observed baseline samples, but hypothetical other baselines in the predictability-parseability plane.
-
-%- identify lambda and quantify what fraction of the baselines are less efficient at this lambda (this is essentially the same as what Zaslavsky et al 2018 do)
-
-%- also to make sure this isn't just because lambda can be chosen, quantify what fraction of baselines Pareto-dominate the same fraction
-
 \section{Supplementary Analyses for Study 2}
@@
 	\begin{center}
-	\includegraphics[width=0.7\textwidth]{../results/correlations/figures/pred-eff-pred-pars-families-2.pdf} % figures/pred-eff-families-2.pdf
+	\includegraphics[width=0.7\textwidth]{../results/correlations/figures/pred-eff-pred-pars-families-2.pdf} 
 	\end{center}
@@
 White cells indicate that the relation is not annotated in the dataset for the given language.}\label{fig:per-lang}
-%	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
-Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
-Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...) .}\label{fig:per-lang}
 \end{figure}
@@
 Note that the Bayesian analysis  also estimates a posterior distribution of the number of satisfied correlations (see Figure~\ref{fig:posterior}), providing an elegant solution to the multiple-comparisons problem arising from analysing the eight correlation.
-
-% and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
-
-%We obtained qualitatively equivalent results for analyses on the levels of individual relations when estimating frequentist models individually for each correlation using \texttt{lme4} \cite{bates2015fitting} and obtaining $p$-values from the resulting fit: For all eight correlations, the sign of $\alpha$ was positive ($p < 0.05$).
 
@@
 
-%\begin{figure}
-%    \centering
-%    
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-langmod-best.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-parse-best.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-two-lambda09-best.pdf}
-%    
-%    
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-ground.pdf}
-%    \includegraphics[scale=.25]{../results/correlations/figures/coverage-depl.pdf}
-%	\caption{Correlations satisfied by the best word order grammars for predictability, parseability, efficiency, by the UD languages, and by the best grammars for dependency length.}
-%    \label{fig:posterior-prevalence}
-%\end{figure}
-
 \subsection{Results on all UD Relations}
@@
 
-%We also report results for grammars optimized for Dependency Length Minimization (DLM), which has been argued to be a property of efficient word order, and which we find to be predicted by efficiency optimization (see Section~\ref{sec:DLM}).
-%In line with this idea, our results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization.
-%These results suggest that it is in part through favoring short dependencies that efficiency predicts word order universals, an idea that has been proposed in prior theoretical studies \citep{hawkins1994performance, hawkins2004efficiency}.
-%%Previous theoretical studies arguing for efficiency explanations of word order universals have already suggested that principles akin to DLM explain the Greenberg correlations ; our results provide the first large-scale empirical confirmation of this idea, and show that this can be reduced to efficiency optimization.
-%See Section~\ref{sec:DLM} for further discussion.
-%
-
-\begin{table*} % TODO figure out away to make this table more aesthetic
+\begin{table*} 
 	\begin{center}	
@@
 In the last column, we indicate what prevalence is expected according to the typological literature. 
-%We additionally report results obtained when optimizing for Dependency Length Minimization (DLM), a heuristic approximation to efficiency that has been suggested in previous work.
-See Section~\ref{sec:DLM} for further discussion of DLM and its relation to efficiency.
 }
@@
 
-%
-%%posterior_perRelation_Real_conj.pdf         posterior_perRelation_Real_goeswith.pdf     posterior_perRelation_Real_nummod.pdf       
-%%user@user-G751JY:~/grammar-optim/results/correlations$ evince figures/posteriors/posterior_perRelation_Real_csubj.pdf 
-%%user@user-G751JY:~/grammar-optim/results/correlations$ 
-%
-%\begin{table*} % TODO figure out away to make this table more aesthetic
-%	\begin{center}	
-%\small{
-%\begin{tabular}{l|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
-%	\hline
-%&	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-%% From Dryer	
-%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}&	lifted\_case  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_case.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_case.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}&	lifted\_cop  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cop.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cop.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}&	aux\footnotemark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_aux.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_aux.pdf}  & $<50\%$ \citep{dryer1992greenbergian} \\
-%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}&	nmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nmod.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}&	acl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_acl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_acl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_acl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_acl.pdf}   & $>50\%$  \citep{dryer1992greenbergian} \\
-%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}&	lifted\_mark  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_mark.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_mark.pdf} & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}&	obl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_obl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_obl.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-%\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}&	xcomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_xcomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_xcomp.pdf}  & $>50\%$ \citep{dryer1992greenbergian} \\
-%%\hdashline
-%
-%\hline
-%% Other universals, not in Dryer
-%&advcl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advcl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advcl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advcl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advcl.pdf}   & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
-%&ccomp  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_ccomp.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_ccomp.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_ccomp.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_ccomp.pdf}   & $>50\%$ (cf.
-\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
-%&csubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_csubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_csubj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_csubj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_csubj.pdf}   & $>50\%$ (cf.
-\cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
-%&nsubj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nsubj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nsubj.pdf}  & See Section~\ref{sec:correlations} \\
-%&amod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_amod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_amod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_amod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_amod.pdf}   &  $\approx 50\%$ \citep{dryer1992greenbergian} \\
-%&nummod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_nummod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_nummod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_nummod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_nummod.pdf}   &$\approx 50\%$ \citep[][89A, 83A]{wals} \\
-% \hline
-%%comp. &
-%%\multirow{3}{*}{NP}&
-%%	AP &
-%\end{tabular}
-%}
-%
-%\end{center}
-%\caption{Predictions on UD relations with predictions from the typological literature.
-The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
-In the last column, we indicate what direction would be expected typologically.}
-%\label{tab:all-predictions-1}
-%\end{table*}
-%
-
-%\footnotetext{Note that we report an \emph{anti-correlation}.
-The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
-Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
-}
-
-\begin{table*} % TODO figure out away to make this table more aesthetic
+\begin{table*} 
 	\begin{center}	
@@
 \end{table*}
-
-%
-%\begin{table*} % TODO figure out away to make this table more aesthetic
-%	\begin{center}	
-%\small{
-%\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
-%	\hline
-%	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-%\hline
-%appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\
-%lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\
-%expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\
-%iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\
-%vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\
-%
-%\hline
-%% Uninterpretable
-%compound  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_compound.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_compound.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_compound.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_compound.pdf}   &                       Uninterpretable \\
-%det  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_det.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_det.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_det.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_det.pdf}   &                                                    Uninterpretable \\
-%dislocated  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dislocated.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dislocated.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dislocated.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dislocated.pdf}   &             Uninterpretable \\
-%dep  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_dep.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_dep.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_dep.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_dep.pdf}   &                                                 Uninterpretable \\
-%advmod  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_advmod.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_advmod.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_advmod.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_advmod.pdf}   &                                         Uninterpretable \\
-%
-%\hline
-%% UD artifacts
-%conj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_conj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_conj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_conj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_conj.pdf}   &                                                      UD Artifact \\
-%discourse  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_discourse.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_discourse.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_discourse.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_discourse.pdf}   &                 UD Artifact \\
-%fixed  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_fixed.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_fixed.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_fixed.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_fixed.pdf}   &                                             UD Artifact \\
-%flat  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_flat.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_flat.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_flat.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_flat.pdf}   &                                                      UD Artifact \\
-%goeswith  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_goeswith.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_goeswith.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_goeswith.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_goeswith.pdf}   &                          UD Artifact \\
-%list  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_list.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_list.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_list.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_list.pdf}   &                                                      UD Artifact \\
-%orphan  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_orphan.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_orphan.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_orphan.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_orphan.pdf}   & UD Artifact \\
-%parataxis  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_parataxis.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_parataxis.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_parataxis.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_parataxis.pdf}   &                 UD Artifact \\
-%reparandum  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_reparandum.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_reparandum.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_reparandum.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_reparandum.pdf}   &                UD Artifact \\
-%
-% \hline
-%%comp. &
-%%\multirow{3}{*}{NP}&
-%%	AP &
-%\end{tabular}
-%}
-%
-%\end{center}
-%\caption{Predictions on UD relations for which no predictions are available in the typological literature.
-``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
-``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
-%\label{tab:all-predictions-2}
-%\end{table*}
-%
-%
 
@@
 
-%\section{Joint optimization of grammar, parser, and language model parameters}
 \section{Creating Optimized Grammars}
@@
 For instance, it does not represent cases where ordering varies between main and embedded clauses, as it does not condition ordering decisions on the larger context.
-%Our method extracts the most frequent ordering and applies it throughout all 
 More complex and powerful ordering grammar models have been proposed \cite{futrell2015experiments, wang2016galactic}; however, they have similar limitations, and for our purposes, the model adopted here has the advantage of being simple and interpretable.
-%There currently is no grammatical formalism that is at the same time capable of representing the full range of word order regularities that are found, 
-%\cite{futrell2015experiments}
 
@@
 \begin{tabular}{l||ll|ll|ll|ll|ll|ll|llllll}
-%	   for(dep in c("obj", "lifted_mark", "acl", "nmod", "obl")) {
 		   Language 
@@
 \end{equation}
-%and we can write
-%\begin{equation}
-%	R_{Pars}^\phi := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) R_{Pars}^{\phi}(u,t)
-%\end{equation}
-%
 A \key{language model}, with some parameters $\psi$, calculates the word-by-word surprisal of an utterance:
@@
 \end{equation}
-%
-%The parsers and language models appropriate to a tree distribution $\tree$ and a differentiable ordering grammar $\mathcal{L}_\theta$ are the ones minimizing parsing loss and surprisal:
-%\begin{equation}
-%\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-%\end{equation}
-%\begin{equation}
-%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-%\end{equation}
-%
-%Using these models, we can rewrite 
-%\begin{equation}
-%R_{Pars}^{\psi(\theta)}= \E_t \E_{u \sim p_\theta(u|t)} R_{Pars}^{\psi(\theta)}(u,t) 
-%\end{equation}
-%where $R_{Pars}(u,t) = \log \frac{p_\phi(t|u)}{p(t)}$, and $p_\theta(u|t)$ is the probability assigned to sentence $u$ by the distribution $\mathcal{L}_\theta(t)$ defined by the differentiable word order grammar $\mathcal{L}_\theta$,
-%and
-%\begin{equation}
-%R_{Pred} = \E_t \E_{u \sim p_\theta(u|t)} R_{Pred}(u)
-%\end{equation}
-%where $R_{Pred}(u) = \log p_\psi(u)$.
-%Therefore, we can rewrite efficiency as
-%\begin{equation}\label{eq:efficiency}
-%R_{\textit{Eff}}^\theta := \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi(\theta)}(u,t) + \lambda R_{Pred}^{\psi(\theta)}(u)\right]
-%\end{equation}
-%Directly applying gradient descent to this expression is hard due to the complicated dependence of $\phi(\theta), \psi(\theta)$ on $\theta$.
-%
-%However, we can rewrite
-%\begin{equation}\label{eq:efficiency}
-%\argmax_\theta\	R_{\textit{Eff}}^\theta 
-%\end{equation}
-%equivalently into
 In order to find an optimal grammar $\theta$, we thus need to compute 
@@
 
-%
-%%(analogous for $R_{Pars}$ instead of $R_{Pred}$)
-%\begin{equation}\label{eq:estimator}
-%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
-%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
-%\end{equation}
-%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
-%for the gradient
-%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
-%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
-%
-
-%
-%
-%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a differentiable word order grammar.
-Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
-%\begin{equation}
-%\label{eq:exp-over-utts}
-%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim \mathcal{L}_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
-%\end{equation}
-%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
-In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
-%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
-%
-%
-%
-%We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
-%\begin{equation*}
-%R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
-%\end{equation*}
-%where the language model parameters $\phi$ are chosen predict the language induced by tree distribution $T$ and word order grammar $\mathcal{L}_\theta$ as accurately as possible (maximizing predictability):
-%\begin{equation*}
-%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-%\end{equation*}
-%
-%
-%We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
-Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
-%\begin{align}
-%\nonumber
-%R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
-%\nonumber
-%&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
-%\end{align}
-%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
-%\begin{equation*}
-%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-%\end{equation*}
-%
-%For efficiency, we linearly combine the two terms:
-%\begin{equation}
-%\label{eq:combined-score}
-%R_{Efficiency}^{\phi,\psi}(t; {\bf w}) = R_{Pars}^\psi(t; {\bf w}) + \lambda R_{Pred}^\phi(t; {\bf w}),
-%\end{equation}
-%
-%
-%%\begin{equation}
-%%    \label{eq:depl}
-%%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
-%%\end{equation}
-%%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
-%
-%For each of our objective functions, we seek parameters $\theta$ of a word order grammar $\mathcal{L}_\theta$ that maximize the average utility of sentences in the language obtained by linearizing unordered trees $t \sim T$ according to $\mathcal{L}_\theta$.
-The calculation of the predictability and parseability objectives also requires fitting language models parameterized by $\phi$ and parsers parameterized by $\psi$, respectively; and the calculation of the efficiency objective requires joint optimization of the word order grammar, language models, and parsers.
-The optimization problems come out to:
-%\begin{align}
-%%\max_\theta &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{DepL}(t; {\bf w}) \label{eq:maximize-depl}\\
-%\max_{\theta,\phi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pred}^{\phi(\theta)}(t; {\bf w}) \label{eq:maximize-pred}\\
-%\max_{\theta,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Pars}^{\psi(\theta)}(t; {\bf w}) \label{Eq:maximize-pars} \\
-%\max_{\theta,\phi,\psi} &\E_{t \sim T}\E_{{\bf w} \sim L_\theta(t)} R_{Efficiency}^{\psi(\theta), \phi(\theta)}(t; {\bf w}). \label{eq:maximize-efficiency} 
-%\end{align}
-%Note that the original word orders from the treebank never enter this objective---in particular, they do not influence the language model and the parser, which are entirely determined by $\theta$.
-%
-%
-%%The case of predictability, parseability, and efficiency are more challenging, as 
-%%e need to simultaneously look for an ordering model optimized for the loss of a language model and/or a parser, and for a language model and/or parser that are optimized for that given ordering model.
-%We solve Eqs.~\ref{eq:maximize-pred}--\ref{eq:maximize-efficiency} by performing stochastic gradient descent over all sets of parameters $\theta, \phi$, and/or $\psi$ simultaneously.
-%Below we discuss the process only for the predictability objective (Eq.~\ref{eq:maximize-pred}); the other objectives are optimized analogously.
-%
-%We initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section).
-%
-%In each optimization step, we sample a dependency tree $t$ from the treebank, then sample an ordering from the current setting of $\theta$ to obtain a linearized sentence ${\bf w} \sim P_{L_\theta}(\cdot|t)$.
-%Then we %update $\thetal$ with ordinary gradient descent, and $\thetad$ with the REINFORCE estimator~\ref{williams-simple-1992}.
-%do a gradient descent step using the estimator %(analogous for $R_{Pars}$ instead of $R_{Pred}$)
-%\begin{equation}\label{eq:estimator}
-%{ \partial_\theta \left(\log P_{L_\theta}({\bf w}|t)\right) \cdot R^\phi_{Pred} (t; {\bf w}) \choose  \partial_\phi R_{Pred}^\phi(t; {\bf w})}
-%%{- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot R_{Pred}(\bf{w}, \thetal) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}
-%\end{equation}
-%%$${- \partial_\thetad \left(\log P_\thetad({\bf w})\right) \cdot \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i}) \choose  - \partial_\thetal \sum_{i=1}^{\#{\bf w}} \log P_\thetal(w_{i+1}|w_{1\dots i})}$$
-%for the gradient
-%$${ \partial_\theta \choose \partial_\phi} \E_{{\bf w'} \sim L_\theta(t)} R_{Pred}^\phi(t; {\bf w'}).$$
-%This unbiased estimator is the ordinary gradient estimator for $\phi$, and the REINFORCE estimator for $\theta$ \citep{williams1992simple}.
-%
-%See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.
-%
-
-%The case of dependency length in Eq.~\ref{maximize-depl} is the simplest, requiring only coordinate descent on the parameters $a_\tau, b_\tau \in \theta$.
-A similar optimization problem was solved by hill-climbing in \citet{gildea2007optimizing} and \citet{gildea2015human}.
-%We optimize this objective by stochastic gradient descent with a batch size of one unordered dependency tree.
-
-%In addition, we estimated maximum-likelihood ordering grammars on the original ordered dependency trees with SGD. For nonprojective trees, we ignore discontinuities.
-%and $\ell_2$ regularization.
-
-%A different commonly considered information-theoretic objective function for word order is Uniform Information Density~\citep{jaeger2010redundancy}, formalized as minimizing a superlinear function of per-word surprisal~\citep{levy:2018cogsci}.
-%Note that optimizing grammars for this objective faces additional challenges, as the optimization problem cannot be rewritten as joint optimization in the form (\ref{eq:maximize-pred}).
-%We thus do not evaluate this objective function here.
-
-%We extract grammars for the actual languages using a Bayesian data analysis method.
-%We specify a simple Bayesian prior over differentiable ordering grammars:
-%$$logit(a_\tau) \sim \mathcal{N}(0,1),\ \ \ \ b_\tau \sim \mathcal{N}(0,1)$$
-%for each $\tau$.
-%Recall that, given a syntactic structure, a differentiable ordering grammar assigns a probability distribution over possible orderings.
-%We extract
-
 \section{Neural Network Architectures}\label{sec:neural-architectures}
@@
 All choices, except where explicitly noted otherwise, were made before evaluating word order properties, and the efficiency of real grammars.
-
-%\paragraph{Locality}
-%In Equation~\ref{eq:depl}, we assume that $head(w_i) = i$ iff $w_i$ is the root of the sentence, so that the root does not contribute to dependency length.
 
@@
 Using POS tags is intended to prevent overfitting on small corpora.
-
-%\begin{equation}\label{eq:rpars}
-%	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)} 
-%\end{equation}
-%\begin{equation}
-%	R_{Pred} := - \operatorname{H}[\utterance] = \sum_{u} p(u) \log p(u)
-%\end{equation}
 
@@
 We employ two common variance reduction methods to improve the estimator~(\ref{eq:dtheta}), while keeping it unbiased.
-%For dependency length and 
-% (and its head, in the case of dependency length)
 For predictability, note that the surprisal of a specific word only depends on the preceding words (not on the following words), and thus only depends on ordering decisions made up to that word.
@@
 
-%We compute a Monte-Carlo estimate of the objective on the UD development set in intervals of 50,000 training steps, and stop optimization once the value of the objective deteriorates compared to the last previous estimate.
-%This is a standard technique for optimizing neural models, and helps prevent overfitting to the specific corpora used.
-
 The choice of optimization methods and the stopping criterion were fixed before we investigated language efficiency or word order correlations.
@@
 As described in our preregistration \url{http://aspredicted.org/blind.php?x=ya4qf8}, this choice was made after conducting a preliminary version of Study 2 reported in Section~\ref{sec:previous-exps}; results reported there show qualitatively identical results regarding the prediction of the eight word order correlations by efficiency optimization.
-
- % TODO we could use some language at the top of this section on the general guiding principles for all these decisions.
-Were these implementation decisions made before or after looking at the word order universals results?
-If they were made before, we should highlight that.
 
@@
 Results (Figure~\ref{fig:cky-parser}) show that optimized grammars are more parseable than baseline grammars, for exact parsing of a simple PCFG.
-% TODO get UAS for compatibility with other parts of SI
-%We then compared parsability of random and optimized word order grammars on English data.
 
@@
 
-%In particular, this allows us to show tha
-
-%Many of the word order biases explained by parseability in Table~\ref{tab:result-dryer} are also explained by dependency length minimization.
-Therefore, we address here the idea that the graph-based parser might have a built-in bias toward parses involving short dependencies, which we call a \key{locality bias}.
-%A simple way to eliminate or alter a locality bias for any parser would be to change the order in which the parser sees words.
-Suppose that a parser $P$ has a bias toward positing a dependency between a word and the immediate previous word.
-
 \paragraph{Even--odd order.} A sequence of $n$ words originally ordered as $w_1 w_2 w_3 w_4 \cdots w_n$ is reordered by separating the even and odd indices: $w_2 w_4 w_6 \cdots w_{n-1} w_1 w_3 w_5 \cdots w_n$ (assuming $n$ odd).
@@
 In all settings, the language optimized for efficiency achieved lower parsing loss (i.e., higher parseability) than random ordering grammars, showing that the parser's preference for optimized languages cannot be attributed to a locality bias.
-%Note that absolute values of parsing accuracy vary between the orderings, being higher for the actual orderings than for distorted orderings, in which trees become nonprojective.
 
@@
 
+TODO explain.
+Here we replicate Study 1.
+We
+
 \begin{figure}
 \centering
-\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized.pdf}
-	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang}.}
+\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized-mle.pdf}
+	\caption[Predictability and Parseability]{Study 1, replication with lexicalized models: Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang}.}
+\end{figure}
+
+\begin{table}
+\centering
+\small{
+\begin{tabular}{l||ll|lll|lll}
+Language & Pred. (t) & Parse. (t) & \multicolumn{3}{c|}{Pred. (Binomial)} & \multicolumn{3}{c}{Parseab. (Binomial)} \\ 
+&  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
+\input{../results/plane/unlexicalized/per-language-statistics-unlexicalized.tex}
+\end{tabular}
+}
+	\caption{Study 1, replication with lexicalized models: Per-language results in Study 1, with \emph{lexicalized} parsers and word-level-only language models.
+Compare Table~\ref{fig:pareto-per-lang-stats}}
+\end{table}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{../results/plane/unlexicalized/analyze_pareto_optimality/figures/quantileByLambda-mle.pdf}
+	\caption{Study 1, replication with lexicalized models: Optimality by $\lambda$ (lexicalized models).}
 \end{figure}
@@
 
+We replicate the comparison between real and random grammars in Study 1.
+
 \subsection{Nondeterministic Baseline Grammars}
-
-%\subsection{Discussion}
-%
-%It is always a logical possibility in studies such as this one that the results are dependent on the particular probabilistic models used.
 
@@
 \end{figure}
-
-%\section{Comparing Extracted Grammars to Linguistic Judgments}
-
-% user@user-X510UAR:~/grammar-optim/results/grammars$ git add comparison-table.tex
 
@@
     \centering
-     \includegraphics[width=0.5\textwidth]{depl-violin-all-1.png} % ../results/dependency-length/figures/depl-violin-all.pdf 
+     \includegraphics[width=0.5\textwidth]{depl-violin-all-1.png} 
         \caption{Average dependency length for grammars optimized to minimize dependency length (DLM, left), optimized for efficiency (second), the real orderings found in corpora (third), and random baseline grammars (right).
@@
     \centering
-	\includegraphics[width=0.9\textwidth]{depLength-facet-1.png} %../results/dependency-length/figures/depLength-facet.pdf} 
+	\includegraphics[width=0.9\textwidth]{depLength-facet-1.png} 
         \caption{Total dependency length as a function of sentence length, for four diverse languages.
@@
 
-\begin{table}[ht] % TODO figure out away to make this table more aesthetic
+\begin{table}[ht] 
 	\begin{center}	
@@
 Furthermore, optimizing for dependency length correctly predicts a range of word order facts, though it appears to \emph{overpredict} correlations when compared to direct optimization for communicative efficiency.
-
-%
-%We also report results for grammars optimized for Dependency Length Minimization (DLM), which has been argued to be a property of efficient word order, and which we find to be predicted by efficiency optimization (see Section~\ref{sec:DLM}).
-%In line with this idea, our results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization.
-%These results suggest that it is in part through favoring short dependencies that efficiency predicts word order universals, an idea that has been proposed in prior theoretical studies \citep{hawkins1994performance, hawkins2004efficiency}.
-%%Previous theoretical studies arguing for efficiency explanations of word order universals have already suggested that principles akin to DLM explain the Greenberg correlations ; our results provide the first large-scale empirical confirmation of this idea, and show that this can be reduced to efficiency optimization.
-%See Section~\ref{sec:DLM} for further discussion.
-%
-%
-%
-%
-%\begin{table*} % TODO figure out away to make this table more aesthetic
-%        \begin{center}
-%\includegraphics[width=0.8\textwidth]{si-table-perrel-1a-1.png}
-%\end{center}
-%\caption{Predictions on UD relations with predictions from the typological literature.
-%The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
-%In the last column, we indicate what direction would be expected typologically.
-%We additionally report results obtained when optimizing for Dependency Length Minimization (DLM), a heuristic approximation to efficiency that has been suggested in previous work.
-See Section~\ref{sec:DLM} for further discussion of DLM and its relation to efficiency.
-}
-%\label{tab:all-predictions-1}
-%\end{table*}
-%
-%
-%
-%
-
-%Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM}).
-%Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
-%Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
-%Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
-%In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
-
-%Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
-%We optimize grammars using the same stochastic gradient descent method as we use for efficiency optimization (Section ).
-%Indeed, we found that DLM is predicted by efficiency optimization, as discussed in Section .
-%Results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.
-
-%DLM turns out to be predicted by our model
 
@@
 \end{document}
-
-\begin{table*} % TODO figure out away to make this table more aesthetic
-	\begin{center}	
-\small{
-\begin{tabular}{|l|l|l|ll|l|l|}
-	\hline
-	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
-% From Dryer	
-acl  &  80   &   \textbf{85}$^{***}$   &   43   &   \textbf{82}$^{***}$   &   \textbf{71}$^{***}$  & $>50\%$  \citep{dryer1992greenbergian} \\
-aux  &  12   &   \textbf{26}$^{***}$   &   \textbf{19}$^{***}$   &   49   &   \textbf{35}$^{**}$  & $<50\%$ \citep{dryer1992greenbergian} \\
-lifted\_case  &  86   &   \textbf{81}$^{***}$   &   \textbf{40}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{67}$^{***}$ & $> 50 \%$ \citep{dryer1992greenbergian}  \\
-lifted\_cop  &  94   &   \textbf{80}$^{***}$   &   59   &   \textbf{72}$^{***}$   &   \textbf{63}$^{**}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
-lifted\_mark  &  76   &   \textbf{85}$^{***}$   &   \textbf{57}$^{*}$   &   \textbf{76}$^{***}$   &   \textbf{72}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
-nmod  &  80   &   \textbf{82}$^{***}$   &   49   &   \textbf{71}$^{***}$   &   \textbf{68}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\ 
-obl  &  88   &   \textbf{78}$^{***}$   &   \textbf{71}$^{***}$   &   46   &   \textbf{68}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
-xcomp  &  88   &   \textbf{90}$^{***}$   &   \textbf{76}$^{***}$   &   \textbf{89}$^{***}$   &   \textbf{85}$^{***}$  &  $> 50 \%$ \citep{dryer1992greenbergian} \\
-\hdashline
-nsubj  &  33   &   \textbf{29}$^{***}$   &   53   &   \textbf{7}$^{***}$   &   \textbf{23}$^{***}$  & $> 50 \%$ \citep{dryer1992greenbergian} \\
-
-\hline
-% Other universals, not in Dryer
-advcl  &  86   &   \textbf{84}$^{***}$   &   51   &   \textbf{67}$^{***}$   &   \textbf{67}$^{*}$  & $>50\%$ \citep{greenberg1963universals,diessel2001ordering} \\
-ccomp  &  86   &   \textbf{88}$^{***}$   &   \textbf{70}$^{**}$   &   \textbf{77}$^{***}$   &   \textbf{77}$^{**}$ & $>50\%$ (cf.
-\cite{dryer1980positional}) \\ % TODO According to Dryer 1980, there is a correlation
-csubj  &  79   &   \textbf{78}$^{***}$   &   60   &   \textbf{66}$^{***}$   &   \textbf{68}$^{**}$  & $>50\%$ (cf. \cite{dryer1980positional}) \\% TODO According to Dryer 1980, there is a correlation
-amod  &  51   &   \textbf{84}$^{***}$   &   42   &   \textbf{53}$^{*}$   &   51   & $\approx 50\%$ \citep{dryer1992greenbergian} \\
-nummod  &  37   &   \textbf{74}$^{**}$   &   49   &   52   &   52 & $\approx 50\%$ \citep[][89A, 83A]{wals} \\ %
-
-\hline
-% Typological data not available
-appos  &  77   &   \textbf{77}$^{***}$   &   49   &   \textbf{67}$^{***}$   &   \textbf{62}$^{**}$ &  Unknown \\%($>50\%$?
-correlates with \textit{nmod} according to Greenberg U 23)  \\ % \cite{cinque2009greenberg} claims it correlates with OV/VO.
-However, it's not clear to me how closely appos matches the subject of the typological claim
-lifted\_cc  &  71   &   \textbf{83}$^{***}$   &   51   &   \textbf{81}$^{***}$   &   \textbf{71}$^{***}$  & Unknown \\
-expl  &  12   &   36   &   62   &   40   &   53  & Unknown \\
-iobj  &  74   &   49   &   \textbf{78}$^{***}$   &   \textbf{38}$^{*}$   &   62  & Unknown \\
-vocative  &  46   &   \textbf{42}$^{*}$   &   59   &   45   &   \textbf{45}$^{*}$  & Unknown \\
-
-\hline
-% Uninterpretable
-compound  &  66   &   \textbf{61}$^{*}$   &   48   &   \textbf{58}$^{*}$   &   52  & Uninterpretable \\
-det  &  31   &   \textbf{71}$^{***}$   &   \textbf{56}$^{*}$   &   48   &   54  & Uninterpretable \\%$.50\%$ for articles, $\approx 50\%$ for others, \citep{dryer1992greenbergian} \\
-dislocated  &  47   &   61   &   \textbf{68}$^{**}$   &   55   &   58  & Uninterpretable \\
-dep  &  56   &   \textbf{64}$^{*}$   &   52   &   58   &   \textbf{58}$^{*}$   & Uninterpretable \\
-advmod  &  35   &   51   &   \textbf{18}$^{***}$   &   47   &   \textbf{39}$^{**}$  & Uninterpretable \\
-
-\hline
-% UD artifacts
-conj  &  72   &   \textbf{82}$^{***}$   &   \textbf{42}$^{**}$   &   \textbf{62}$^{***}$   &   \textbf{59}$^{*}$  & UD artifact  \\
-discourse  &  24   &   \textbf{36}$^{**}$   &   \textbf{41}$^{*}$   &   49   &   46  & UD artifact   \\
-fixed  &  78   &   \textbf{32}$^{***}$   &   45   &   47   &   44  & UD artifact   \\
-flat  &  73   &   \textbf{66}$^{**}$   &   45   &   \textbf{59}$^{**}$   &   54  & UD artifact   \\
-goeswith  &  61   &   57   &   44   &   57   &   52  & UD artifact   \\
-list  &  94   &   59   &   50   &   54   &   52  & UD artifact  \\
-orphan  &  82   &   \textbf{70}$^{**}$   &   50   &   \textbf{55}$^{*}$   &   59  & UD artifact   \\
-parataxis  &  66   &   56   &   43   &   57   &   59  & UD artifact  \\
-reparandum  &  18   &   53   &   57   &   43   &   48  & UD artifact  \\
-
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-    \multicolumn{7}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
-\end{tabular}
-}
-
-\end{center}
-\caption{Predictions on all UD relations occurring in more than one language.
-Numbers are the fraction of languages where the direction of the given relation agrees in direction with the \emph{obj} relation.
-In the last column, we indicate what direction would be expected typologically.
-The first section correspond to universals from Dryer.
-``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
-``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).}
-\label{tab:all-predictions}
-\end{table*}

### 2019-09-29 results/plane/optimalityTests_perLanguage.R

@@ -1,0 +1,169 @@
+library(lme4)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+depl = read.csv("../../grammars/dependency_length/total_summary_funchead_coarse.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+library(tidyr)
+library(dplyr)
+depl = depl %>% filter(grepl("FuncHead", ModelName)) %>% filter(grepl("Coarse", ModelName))
+dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS3 = read.csv("../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS4 = read.csv("../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS5 = read.csv("../../grammars/plane/plane-fixed-random3.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS6 = read.csv("../../grammars/plane/plane-fixed-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS7 = read.csv("../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS = rbind(dataS, dataS2, dataS3, dataS4, dataS5, dataS6, dataS7)
+dataP = read.csv("../../grammars/plane/plane-parse-redone.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS = dataS %>% group_by(Language, Type, Model) %>% summarise(Surprisal = mean(Surp, na.rm=TRUE))
+dataP = dataP %>% group_by(Language, Type, Model) %>% summarise(Pars = mean(ParsU, na.rm=TRUE))
+dataS = as.data.frame(dataS)
+dataP = as.data.frame(dataP)
+library(lme4)
+dataS = dataS %>% mutate(Type = as.character(Type))
[…]

### 2019-09-29 results/plane/recordBaselinePositions.R

@@ -1,0 +1,47 @@
+library(lme4)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+depl = read.csv("../../grammars/dependency_length/total_summary_funchead_coarse.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+library(tidyr)
+library(dplyr)
+depl = depl %>% filter(grepl("FuncHead", ModelName)) %>% filter(grepl("Coarse", ModelName))
+dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS3 = read.csv("../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS4 = read.csv("../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS5 = read.csv("../../grammars/plane/plane-fixed-random3.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS6 = read.csv("../../grammars/plane/plane-fixed-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS7 = read.csv("../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS = rbind(dataS, dataS2, dataS3, dataS4, dataS5, dataS6, dataS7)
+dataP = read.csv("../../grammars/plane/plane-parse-redone.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS = dataS %>% group_by(Language, Type, Model) %>% summarise(Surprisal = mean(Surp, na.rm=TRUE))
+dataP = dataP %>% group_by(Language, Type, Model) %>% summarise(Pars = mean(ParsU, na.rm=TRUE))
+dataS = as.data.frame(dataS)
+dataP = as.data.frame(dataP)
+library(lme4)
+summary(lmer(Surprisal ~ Type + (1|Language), data=dataS %>% filter(grepl("langm", Type))))
[…]

### 2019-09-29 results/plane/visualizePlane_perLanguage.R

@@ -1,0 +1,190 @@
+# per language
+
+
+library(lme4)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+depl = read.csv("../../grammars/dependency_length/total_summary_funchead_coarse.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+library(tidyr)
+library(dplyr)
+depl = depl %>% filter(grepl("FuncHead", ModelName)) %>% filter(grepl("Coarse", ModelName))
+dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS3 = read.csv("../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS4 = read.csv("../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS5 = read.csv("../../grammars/plane/plane-fixed-random3.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS6 = read.csv("../../grammars/plane/plane-fixed-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS7 = read.csv("../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS = rbind(dataS, dataS2, dataS3, dataS4, dataS5, dataS6, dataS7)
+dataP = read.csv("../../grammars/plane/plane-parse-redone.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS = dataS %>% group_by(Language, Type, Model) %>% summarise(Surprisal = mean(Surp, na.rm=TRUE))
+dataP = dataP %>% group_by(Language, Type, Model) %>% summarise(Pars = mean(ParsU, na.rm=TRUE))
+dataS = as.data.frame(dataS)
+dataP = as.data.frame(dataP)
+library(lme4)
+summary(lmer(Surprisal ~ Type + (1|Language), data=dataS %>% filter(grepl("langm", Type))))
[…]

### 2019-09-30 writeup/pnas-word-order-si.tex

@@
 % say this somewhere
-Note that the predictability scores are considerably worse than the state-of-the art in Natural Language Processing; this is because most of the corpora are much smaller than those typically used in language modeling work.
+Note that the predictability scores are considerably worse than the state-of-the art in Natural Language Processing; this is because most of the corpora are much smaller than those typically used in language modeling work (CITE a few things on large corpora; also cite language modeling work on many small corpora).
+
+%TODO
+%$\sum_{t \in C} \log p(t|u) \approx |C| H[P|T]  $
+%$\sum_{t \in C} \sum_{i} \log p(t|u) \approx |C| H[P|T]  $
 
@@
 \end{figure}
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
+%\caption[Predictability and Parseability]{Predictability and parseability of 51 languages.
+Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.
+TODO adapt for colorblindness.}\label{fig:pareto-per-lang}
+%\end{figure}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-arrows-smoothed.pdf}
-\caption[Predictability and Parseability]{Smoothed version with arrows. Is this useful?}\label{fig:pareto-per-lang-arrows}
-\end{figure}
+%\begin{figure}
+%\centering
+%\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-arrows-smoothed.pdf}
+%\caption[Predictability and Parseability]{Smoothed version with arrows. Is this useful?}\label{fig:pareto-per-lang-arrows}
+%\end{figure}
+%
 
@@
 
+%
+%
+%\begin{table}
+%\centering
+%\small{
+%\begin{tabular}{l||ll|lll|lll}
+%Language & Pred. (t) & Parse. (t) & \multicolumn{3}{c|}{Pred. (Binomial)} & \multicolumn{3}{c}{Parseab. (Binomial)} \\ 
+%&  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
+%\input{../results/plane/per-language-statistics.tex}
+%\end{tabular}
+%}
+%\caption{Per-language results in Study 1.
+For each language, we show the following: (1) $p$-values obtained from a one-sided $t$-test, for the null that the mean predictability/parseability of random grammars is at least as high as that of the real grammar.
+(2) Results from one-sided binomial tests, for the null that the the real grammar is better than at most $50 \%$ of random grammars.
+In addition to the $p$-value, we report point estimate and 95\% confidence interval for the fraction of random grammars that have values below real grammars.}\label{fig:pareto-per-lang-stats}
+%\end{table}
+%
+%
+
 \subsection{Analysis controlling for Families}
@@
 
-\paragraph{Analysis 2: Quantifying Degree of Pareto Optimality}
-
-TODO maybe we can scrap this part as it is harder to explain well, if Analysis 1 sufficient.
+TODO also make a plot with the 0.0 ad 0.9 lines
 
@@
 \centering
-	\begin{tabular}{llll}
-		(a) & (b) & (c) & (d) \\
-		\includegraphics[width=0.22\textwidth]{figures/isocline1.png}&
-		\includegraphics[width=0.22\textwidth]{figures/isocline2.png}&
-		\includegraphics[width=0.22\textwidth]{figures/isocline3.png}&
-\includegraphics[width=0.22\textwidth]{figures/isocline4.png}
-	\end{tabular}
-
-	\caption{Isolines: (a) The set of all possible baseline grammars forms a distribution in the parseability-predictability space.
-	The Pareto frontier consists of those points that are not dominated by any baseline grammar -- that is, those for which $P=0$: For points on this curve, no baseline grammar is both to their top and their right.
-	(b) Other grammars are dominated by some fractions $P > 0$ of baseline grammars.
+\includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/pareto-plane-perLanguage-arrows-smoothed-halfspace-untransformed.pdf}
+	\caption[]{TODO}\label{fig:lambda-halfplane-09}
+\end{figure}
+
+%\paragraph{Analysis 2: Quantifying Degree of Pareto Optimality}
+%
+%TODO maybe we can scrap this part as it is harder to explain well, if Analysis 1 sufficient.
+%
+%
+%\begin{figure}
+%\centering
+%	\begin{tabular}{llll}
+%		(a) & (b) & (c) & (d) \\
+%		\includegraphics[width=0.22\textwidth]{figures/isocline1.png}&
+%		\includegraphics[width=0.22\textwidth]{figures/isocline2.png}&
+%		\includegraphics[width=0.22\textwidth]{figures/isocline3.png}&
+%\includegraphics[width=0.22\textwidth]{figures/isocline4.png}
+%	\end{tabular}
+%
+%	\caption{Isolines: (a) The set of all possible baseline grammars forms a distribution in the parseability-predictability space.
+%	The Pareto frontier consists of those points that are not dominated by any baseline grammar -- that is, those for which $P=0$: For points on this curve, no baseline grammar is both to their top and their right.
+%	(b) Other grammars are dominated by some fractions $P > 0$ of baseline grammars.
 The two green crosses mark grammars that both are dominated by $P=0.05$ of baseline grammars.
 A blue curve connects all grammars dominated by exactly $P=0.05$ of baseline grammars.
-	(c) We observe a sample of $N=50$ baseline grammars, sampled uniformly from the set of baseline grammars. Samples will tend to concentrate in the high-density regions of the baseline distribution.
-	We also observe the efficiency of the real language (red cross) and an estimate of the Pareto frontier (blue line).
-	(d) We use the $50$ samples from the baseline grammar distribution to construct an estimate of the full abseline distibution, using the Dirichlet Process Gaussian Mixture Model.
+%	(c) We observe a sample of $N=50$ baseline grammars, sampled uniformly from the set of baseline grammars. Samples will tend to concentrate in the high-density regions of the baseline distribution.
+%	We also observe the efficiency of the real language (red cross) and an estimate of the Pareto frontier (blue line).
+%	(d) We use the $50$ samples from the baseline grammar distribution to construct an estimate of the full abseline distibution, using the Dirichlet Process Gaussian Mixture Model.
 Using this inferred distribution, we estimate what fraction $P$ of baseline grammars dominate the real language, and the curve connecting all grammars dominated by the same fraction $P$ of baselines.
-	}\label{fig:isoline-explanation}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-isoCurve-smooth-interval.pdf}
-	\caption{Real languages, even those that are not at the estimated Pareto frontier, are Pareto-dominated by very few other hypothetical grammars.
+%	}\label{fig:isoline-explanation}
+%\end{figure}
+%
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-isoCurve-smooth-interval.pdf}
+%	\caption{Real languages, even those that are not at the estimated Pareto frontier, are Pareto-dominated by very few other hypothetical grammars.
 Green curves: Baseline distribution, estimated using Kernel Density estimation.
@@
 }\label{fig:isoline}
-\end{figure}
-
-%The Pareto curve is formed by those grammars such that no grammar simultaneously achieves higher parseability and higher predictability.
-%This motivates
-
-We quantify the Pareto-optimality of the language $r$ as the fraction of baselines that Pareto-dominate $r$ -- that is, baselines that are at least as good as $r$ on \emph{both} parseability and predictability:
-\begin{equation}
-P := \frac{\# \{b \in B: R_{Pars}(b) \geq R_{eff}(r) \text{ and }  R_{Pred}(b) \geq R_{Pred}(r) \}}{\#B}
-\end{equation}
-%This quantifies how many hypothetical grammars are more efficient than $r$ for \emph{all} choices of $\lambda$ (make a figure).
-%This quantity is also known as the joint (or bivariate) exceedance probability.
-If we have access to the full distribution over baseline grammars, then the Pareto frontier is formed by exactly those grammars where $P=0$ (make a figure).
-(TODO slight subtlety: $r$ itself is part of $B$, so $P > 0$ TODO change definition)
-Small values of $P$ indicate that only few baseline grammars are better than $r$.
-%In practice, we only have a limited sample of $50$ baseline grammar samples; so only an approximation of $P$ is actually observable, and this approximation will be $0$ in a larger region of the plane (make a figure).
-
-In this section, we aim to estimate $P$, provide uncertainty intervals.
-In order to also compare with the values of $P$ attained by baseline languages, we also estimated the isoline (cotour line) of grammars having the same $P$ value as the real language.
-
-%$P$ is also known as the \emph{joint} or \emph{bivariate} \emph{exceedance probability}, and there are a range of methods for estimating it.
-
-We used a simple nonparametric Bayesian approach to estimate $P$ from the 50 observed baseline samples.
-The baseline languages form a distribution in the parseability-predictability plane, from which 50 independent samples are known.
-We model the underlying distribution using the Dirichlet Process Gaussian Mixture Model (DPGMM), a flexible nonparametric model that can accomodate multimodal and other non-normal distributions \citep{rasmussen2000infinite, escobar1995bayesian, muller1996bayesian}.
-This model is a Gaussian mixture model where the number of components itself is a latent parameter inferred from the data.
-DPGMMs can be seen as a Bayesian version of Kernel Density Smoothing.\footnote{A potential objection to using such Kernel Density Estimation is that it models a continuous distribution with support everywhere, while the distribution of grammars is discrete and finite, in particular bounded by a Pareto frontier.
+%\end{figure}
+%
+%
+%%The Pareto curve is formed by those grammars such that no grammar simultaneously achieves higher parseability and higher predictability.
+%%This motivates 
+%
+%We quantify the Pareto-optimality of the language $r$ as the fraction of baselines that Pareto-dominate $r$ -- that is, baselines that are at least as good as $r$ on \emph{both} parseability and predictability:
+%\begin{equation}
+%P := \frac{\# \{b \in B: R_{Pars}(b) \geq R_{eff}(r) \text{ and }  R_{Pred}(b) \geq R_{Pred}(r) \}}{\#B}
+%\end{equation}
+%%This quantifies how many hypothetical grammars are more efficient than $r$ for \emph{all} choices of $\lambda$ (make a figure).
+%%This quantity is also known as the joint (or bivariate) exceedance probability.
+%If we have access to the full distribution over baseline grammars, then the Pareto frontier is formed by exactly those grammars where $P=0$ (make a figure).
+%(TODO slight subtlety: $r$ itself is part of $B$, so $P > 0$ TODO change definition)
+%Small values of $P$ indicate that only few baseline grammars are better than $r$.
+%%In practice, we only have a limited sample of $50$ baseline grammar samples; so only an approximation of $P$ is actually observable, and this approximation will be $0$ in a larger region of the plane (make a figure).
+%
+%In this section, we aim to estimate $P$, provide uncertainty intervals.
+%In order to also compare with the values of $P$ attained by baseline languages, we also estimated the isoline (cotour line) of grammars having the same $P$ value as the real language.
+%
+%%$P$ is also known as the \emph{joint} or \emph{bivariate} \emph{exceedance probability}, and there are a range of methods for estimating it.
+%
+%We used a simple nonparametric Bayesian approach to estimate $P$ from the 50 observed baseline samples.
+%The baseline languages form a distribution in the parseability-predictability plane, from which 50 independent samples are known.
+%We model the underlying distribution using the Dirichlet Process Gaussian Mixture Model (DPGMM), a flexible nonparametric model that can accomodate multimodal and other non-normal distributions \citep{rasmussen2000infinite, escobar1995bayesian, muller1996bayesian}.
+%This model is a Gaussian mixture model where the number of components itself is a latent parameter inferred from the data.
+%DPGMMs can be seen as a Bayesian version of Kernel Density Smoothing.\footnote{A potential objection to using such Kernel Density Estimation is that it models a continuous distribution with support everywhere, while the distribution of grammars is discrete and finite, in particular bounded by a Pareto frontier.
 We believe that this does not cause problems for this analysis.
@@
 Second, the approximate Pareto frontier tends to be clearly separated from the high-density regions of the baseline distribution, in which most sampled baselines fall, showing that the baseline distribution has very small density around and beyond the Pareto frontier.}
-In our context here, the reason for using DPGMMs instead of Kernel Density Smoothing is that Bayesian inference using the DPGMM provides a straightforward method of deriving estimates and credible intervals for $PO$ and the isoline.
-
-% http://mlg.eng.cam.ac.uk/pub/pdf/GoeRas10.pdf
-% TODO report prior
-
-The DPGMM defines a prior over possible baseline distributions in the parseability-predictability plane.
-%Under this prior, each possible distribution is parameterized by a number $n$ of components, $n$ bivariate Gaussians $\mu_i, \Sigma_i$, and $n$ weights $w_i$ for these components ($\sum_{i=1}^n w_i=1$).
-Each possible distribution is parameterized as $\sum_{i=1}^n w_i p_{\mu_i, \Sigma_i}(x)$ where $p_{\mu_i, \Sigma_i}$ is the bivariate Gaussian density.
-The number $n$ of components is inferred from the data together with the weights $w_i$, the means $\mu_i$, and the covariance matrices $\Sigma_i$.
-A Dirichlet process prior (TODO) is placed on $n$, favoring models with smaller numbers of components unless the data provides evidence for more components.
-We use the $50$ observed baseline samples to infer a posterior distribution over such mixture distributions.
-We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
-We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples, from which we selected an evenly spaced subset of 100 samples.
-
-For each of these samples, we then computed $P$ of the real language, and the isoline.
-The collection of these results provides a set of samples from the posterior distribution over the true value of $P$ and the true isoline.
-We plot the posterior means of $P$ and the isoline in Figure~\ref{fig:isoline}.
+%In our context here, the reason for using DPGMMs instead of Kernel Density Smoothing is that Bayesian inference using the DPGMM provides a straightforward method of deriving estimates and credible intervals for $PO$ and the isoline.
+%
+%% http://mlg.eng.cam.ac.uk/pub/pdf/GoeRas10.pdf
+%% TODO report prior
+%
+%The DPGMM defines a prior over possible baseline distributions in the parseability-predictability plane.
+%%Under this prior, each possible distribution is parameterized by a number $n$ of components, $n$ bivariate Gaussians $\mu_i, \Sigma_i$, and $n$ weights $w_i$ for these components ($\sum_{i=1}^n w_i=1$).
+%Each possible distribution is parameterized as $\sum_{i=1}^n w_i p_{\mu_i, \Sigma_i}(x)$ where $p_{\mu_i, \Sigma_i}$ is the bivariate Gaussian density.
+%The number $n$ of components is inferred from the data together with the weights $w_i$, the means $\mu_i$, and the covariance matrices $\Sigma_i$.
+%A Dirichlet process prior (TODO) is placed on $n$, favoring models with smaller numbers of components unless the data provides evidence for more components.
+%We use the $50$ observed baseline samples to infer a posterior distribution over such mixture distributions.
+%We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
+%We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples, from which we selected an evenly spaced subset of 100 samples.
+%
+%For each of these samples, we then computed $P$ of the real language, and the isoline.
+%The collection of these results provides a set of samples from the posterior distribution over the true value of $P$ and the true isoline.
+%We plot the posterior means of $P$ and the isoline in Figure~\ref{fig:isoline}.
+%
+%
+%
+%
+%
 
@@
 
-    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
-    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
-    \includegraphics[scale=.55]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
+    \includegraphics[width=.3\textwidth]{../results/correlations/correlations-by-grammar/ground-corrs-efficiency.pdf}
+    \includegraphics[width=.3\textwidth]{../results/correlations/correlations-by-grammar/ground-corrs-parseability.pdf}
+    \includegraphics[width=.3\textwidth]{../results/correlations/correlations-by-grammar/ground-corrs-predictability.pdf}
 
@@
 We provide an example in Figure~\ref{fig:grammar-sample}, illustrating grammar parameters for the relations in Figure 3 of the main paper.
+
 Note that the grammatical formalism simplifies some aspects of the word order regularities of natural languages.
 For instance, it does not represent cases where ordering varies between main and embedded clauses, as it does not condition ordering decisions on the larger context.
+It also does not nonprojective orderings, which -- while generally rare -- do occur in many languages (CITE).
 More complex and powerful ordering grammar models have been proposed \cite{futrell2015experiments, wang2016galactic}; however, they have similar limitations, and for our purposes, the model adopted here has the advantage of being simple and interpretable.
@@
 
+TODO also mention: As we estimate parseability and predictability using cross-entropies, our models constitute variational approximations to the true parseability and predictability values.
+A similar approach has been applied to quantifying morphological complexity by \citet{cotterell2019complexity}.
+
 \subsection{CKY Parsers}
@@
 Indeed, \cite{futrell2017memory} argues specifically that it emerges from efficiency optimization.
+[TODO also cite rijkhof, ferrer i cancho]
 
@@
 We find that optimizing grammars for efficiency reduces dependency length to a similar degree as found in the actual orderings in the corpora, almost up to the limit given by directly optimizing for dependency length.
+[TODO acknowledge  Ferrer i Cancho R, Liu H (2014) The risks of mixing dependency lengths from se-quences of different length.Glottotheory5(2):143–155.]
 We also plot more detailed results for four languages in Figure~\ref{fig:dlm-4langs}, plotting dependency length as a function of sentence length as reported in prior work \cite{futrell2015largescale}.

### 2019-10-02 writeup/pnas-word-order-si.tex
code

@@
 \tableofcontents
+
+TODO CITE \cite{mielke-et-al-2019} % However,  a lan-guage’s vocabulary size and the length in charactersof its sentences were well-correlated with difficultyon our large set of languages.
+\cite{gerz2018language}
+% TODO maybe also some recent stuff in data-driven induction of typology https://www.mitpressjournals.org/doi/abs/10.1162/COLI_a_00357
 
@@
 Within each language, we estimate the Pareto curve based on the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
+% The Pareto frontier is the set of all Pareto efficient allocations, conventionally shown graphically.
+This frontier is approximate:
+The actual curve may not be piecewise linear, and it may also extend beyond the estimated curve, as the grammar optimization method is approximate.
 
@@
 Across language families, real grammars tend to be more efficient than almost all baseline grammars.
+According to this analysis, exceptions among the UD corpora can be attributed to language-specific effects.
+
+TODO also report glmer, posterior plots, transformed probabilities
 
@@
 \end{figure}
+
+TODO discuss difference from semantic typology:
+TODO mention that, in semantic typology, random baselines might tend to be closer to the frontier, whereas, in our study, there tends to be a low-density region between the baseline grammars and the optimum (as also observed in previous DLM and surprisal studies, cite Futrell and Gildea/Jaeger).
 
@@
 
-TODO also mention: As we estimate parseability and predictability using cross-entropies, our models constitute variational approximations to the true parseability and predictability values.
-A similar approach has been applied to quantifying morphological complexity by \citet{cotterell2019complexity}.
+TODO also mention: As we estimate parseability and predictability using cross-entropies, our models constitute variational approximations to the true parseability and predictability values, similar to previous work quantifying other aspects of linguistic complexity \citep{cotterell2019complexity}.
+TODO
 
@@
 Second, while the graph-based neural parser uses machine learning to induce syntactic knowledge from data, the CKY parser performs exact probabilistic inference.
-%In this sense, the CKY algorithm does not have any architectural biases in itself.
-%On the other hand, the PCFG makes severely simplifying independence assumptions, compared to the universal approximation capabilities of neural network-based systems.
+In this sense, the CKY algorithm does not have any architectural biases in itself.
+On the other hand, the PCFG makes severely simplifying independence assumptions, compared to the universal approximation capabilities of neural network-based systems.
 

### 2019-10-08 writeup/pnas-word-order-si.tex

@@
 Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
-Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
-184) without replacement.
+Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order from left to right is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (for dependents preceding the head) or $\operatorname{softmax}(-b_{\tau_1}, \dots, -b_{\tau_n})$ (for dependents following the head) (for the definition of $\operatorname{Softmax}$, see \cite[p.
+184]{goodfellow2016deep}) without replacement.
 

### 2019-10-08 results/correlations/analysis/efficiency-control-lme4.R
code

@@ -16,13 +16,13 @@ data  = merge(data, languages, by=c("Language"), all.x=TRUE)
 dependency = "nmod"
 dependencies = c("acl", "aux", "lifted_case", "lifted_cop", "lifted_mark", "nmod", "obl", "xcomp")
 objs = data %>% filter(CoarseDependency == "obj") %>% mutate(obj = pmax(0, sign(DH_Weight))) %>% select(obj, Language, FileName)
+
 data = data %>% filter((CoarseDependency %in% dependencies))
 data = merge(data, objs, by=c("Language", "FileName"))
 data = data %>% mutate(dir = pmax(0, sign(DH_Weight)))
 data = data %>% mutate(dir = ifelse(DH_Weight == 0, NA, dir))
 data$agree = (data$dir == data$obj)
 data = unique(data %>% select(Family, Language, FileName, CoarseDependency, agree)) %>% spread(CoarseDependency, agree)
-
 
 library(lme4)
 
@@ -38,4 +38,9 @@ for(dependency in dependencies) {
 sink()
 
 
+
+##################################
+
+#data = merge(data, objs, by=c("Language", "FileName"))
+#summary(glmer(paste(dependency, " ~ obj + (1|Family) + (1|Language)"), family="binomial", data=data))
 

### 2019-10-08 results/plane/visualizePlane_perLanguage.R

@@ -19,7 +19,7 @@ dataS7 = read.csv("../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% 
 dataS = rbind(dataS, dataS2, dataS3, dataS4, dataS5, dataS6, dataS7)
 dataP = read.csv("../../grammars/plane/plane-parse-redone.tsv", sep="\t") %>% mutate(Model = as.character(Model))
 dataS = dataS %>% group_by(Language, Type, Model) %>% summarise(Surprisal = mean(Surp, na.rm=TRUE))
-dataP = dataP %>% group_by(Language, Type, Model) %>% summarise(Pars = mean(ParsU, na.rm=TRUE))
+dataP = dataP %>% group_by(Language, Type, Model) %>% summarise(Pars = mean(Pars, na.rm=TRUE))
 dataS = as.data.frame(dataS)
 dataP = as.data.frame(dataP)
 library(lme4)

### 2019-10-11 writeup/pnas-word-order-si.tex
code

@@
 
-For each language, we extracted all lemmas of words heading an \emph{xcomp} dependency, occurring at least 2 times.
+For each language, we extracted all lemmas of words heading an \emph{xcomp} dependency, occurring at least 10 times.
 In 39 languages, all extracted words had the same dominant order.
@@
 
-%TODO
 %$\sum_{t \in C} \log p(t|u) \approx |C| H[P|T]  $
@@
 
-TODO also make a plot with the 0.0 ad 0.9 lines
-
 \begin{figure}
@@
 \includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/pareto-plane-perLanguage-arrows-smoothed-halfspace-untransformed.pdf}
-	\caption[]{TODO}\label{fig:lambda-halfplane-09}
+	\caption[]{Per-Language results as in Figure~\ref{fig:pareto-per-lang-raw}, with lines connecting those points that have the same efficiency as the real grammar at $\lambda = 0.0$ (dotted) and $\lambda = 0.9$ (dashed).
+Points to the bottom/left of these line have lower efficiency than the real grammar, at the given value of $\lambda$.}\label{fig:lambda-halfplane-09}
 \end{figure}
@@
 TODO mention that, in semantic typology, random baselines might tend to be closer to the frontier, whereas, in our study, there tends to be a low-density region between the baseline grammars and the optimum (as also observed in previous DLM and surprisal studies, cite Futrell and Gildea/Jaeger).
-
-%\paragraph{Analysis 2: Quantifying Degree of Pareto Optimality}
-%
-%TODO maybe we can scrap this part as it is harder to explain well, if Analysis 1 sufficient.
-%
-%
-%\begin{figure}
-%\centering
-%	\begin{tabular}{llll}
-%		(a) & (b) & (c) & (d) \\
-%		\includegraphics[width=0.22\textwidth]{figures/isocline1.png}&
-%		\includegraphics[width=0.22\textwidth]{figures/isocline2.png}&
-%		\includegraphics[width=0.22\textwidth]{figures/isocline3.png}&
-%\includegraphics[width=0.22\textwidth]{figures/isocline4.png}
-%	\end{tabular}
-%
-%	\caption{Isolines: (a) The set of all possible baseline grammars forms a distribution in the parseability-predictability space.
-%	The Pareto frontier consists of those points that are not dominated by any baseline grammar -- that is, those for which $P=0$: For points on this curve, no baseline grammar is both to their top and their right.
-%	(b) Other grammars are dominated by some fractions $P > 0$ of baseline grammars.
-The two green crosses mark grammars that both are dominated by $P=0.05$ of baseline grammars.
-A blue curve connects all grammars dominated by exactly $P=0.05$ of baseline grammars.
-%	(c) We observe a sample of $N=50$ baseline grammars, sampled uniformly from the set of baseline grammars. Samples will tend to concentrate in the high-density regions of the baseline distribution.
-%	We also observe the efficiency of the real language (red cross) and an estimate of the Pareto frontier (blue line).
-%	(d) We use the $50$ samples from the baseline grammar distribution to construct an estimate of the full abseline distibution, using the Dirichlet Process Gaussian Mixture Model.
-Using this inferred distribution, we estimate what fraction $P$ of baseline grammars dominate the real language, and the curve connecting all grammars dominated by the same fraction $P$ of baselines.
-%	}\label{fig:isoline-explanation}
-%\end{figure}
-%
-%
-%\begin{figure}
-%\centering
-%\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-isoCurve-smooth-interval.pdf}
-%	\caption{Real languages, even those that are not at the estimated Pareto frontier, are Pareto-dominated by very few other hypothetical grammars.
-Green curves: Baseline distribution, estimated using Kernel Density estimation.
-Black numbers indicate the fraction of baselines dominating the real language, with a 95\% credible interval, estimated using a nonparametric Bayesian method (see text).
-Orange curves connect points estimated to be dominated by the same fraction of baselines as the real language, estimated using the same Bayesian method (see text).
-}\label{fig:isoline}
-%\end{figure}
-%
-%
-%%The Pareto curve is formed by those grammars such that no grammar simultaneously achieves higher parseability and higher predictability.
-%%This motivates 
-%
-%We quantify the Pareto-optimality of the language $r$ as the fraction of baselines that Pareto-dominate $r$ -- that is, baselines that are at least as good as $r$ on \emph{both} parseability and predictability:
-%\begin{equation}
-%P := \frac{\# \{b \in B: R_{Pars}(b) \geq R_{eff}(r) \text{ and }  R_{Pred}(b) \geq R_{Pred}(r) \}}{\#B}
-%\end{equation}
-%%This quantifies how many hypothetical grammars are more efficient than $r$ for \emph{all} choices of $\lambda$ (make a figure).
-%%This quantity is also known as the joint (or bivariate) exceedance probability.
-%If we have access to the full distribution over baseline grammars, then the Pareto frontier is formed by exactly those grammars where $P=0$ (make a figure).
-%(TODO slight subtlety: $r$ itself is part of $B$, so $P > 0$ TODO change definition)
-%Small values of $P$ indicate that only few baseline grammars are better than $r$.
-%%In practice, we only have a limited sample of $50$ baseline grammar samples; so only an approximation of $P$ is actually observable, and this approximation will be $0$ in a larger region of the plane (make a figure).
-%
-%In this section, we aim to estimate $P$, provide uncertainty intervals.
-%In order to also compare with the values of $P$ attained by baseline languages, we also estimated the isoline (cotour line) of grammars having the same $P$ value as the real language.
-%
-%%$P$ is also known as the \emph{joint} or \emph{bivariate} \emph{exceedance probability}, and there are a range of methods for estimating it.
-%
-%We used a simple nonparametric Bayesian approach to estimate $P$ from the 50 observed baseline samples.
-%The baseline languages form a distribution in the parseability-predictability plane, from which 50 independent samples are known.
-%We model the underlying distribution using the Dirichlet Process Gaussian Mixture Model (DPGMM), a flexible nonparametric model that can accomodate multimodal and other non-normal distributions \citep{rasmussen2000infinite, escobar1995bayesian, muller1996bayesian}.
-%This model is a Gaussian mixture model where the number of components itself is a latent parameter inferred from the data.
-%DPGMMs can be seen as a Bayesian version of Kernel Density Smoothing.\footnote{A potential objection to using such Kernel Density Estimation is that it models a continuous distribution with support everywhere, while the distribution of grammars is discrete and finite, in particular bounded by a Pareto frontier.
-We believe that this does not cause problems for this analysis.
-First, the number of possible grammars is extremey large, exceeding $37!$, which is more than $10^{43}$, suggesting that the distribution in the parseability-predictability plane is closely approximated by a continuous distribution.
-Second, the approximate Pareto frontier tends to be clearly separated from the high-density regions of the baseline distribution, in which most sampled baselines fall, showing that the baseline distribution has very small density around and beyond the Pareto frontier.}
-%In our context here, the reason for using DPGMMs instead of Kernel Density Smoothing is that Bayesian inference using the DPGMM provides a straightforward method of deriving estimates and credible intervals for $PO$ and the isoline.
-%
-%% http://mlg.eng.cam.ac.uk/pub/pdf/GoeRas10.pdf
-%% TODO report prior
-%
-%The DPGMM defines a prior over possible baseline distributions in the parseability-predictability plane.
-%%Under this prior, each possible distribution is parameterized by a number $n$ of components, $n$ bivariate Gaussians $\mu_i, \Sigma_i$, and $n$ weights $w_i$ for these components ($\sum_{i=1}^n w_i=1$).
-%Each possible distribution is parameterized as $\sum_{i=1}^n w_i p_{\mu_i, \Sigma_i}(x)$ where $p_{\mu_i, \Sigma_i}$ is the bivariate Gaussian density.
-%The number $n$ of components is inferred from the data together with the weights $w_i$, the means $\mu_i$, and the covariance matrices $\Sigma_i$.
-%A Dirichlet process prior (TODO) is placed on $n$, favoring models with smaller numbers of components unless the data provides evidence for more components.
-%We use the $50$ observed baseline samples to infer a posterior distribution over such mixture distributions.
-%We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
-%We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples, from which we selected an evenly spaced subset of 100 samples.
-%
-%For each of these samples, we then computed $P$ of the real language, and the isoline.
-%The collection of these results provides a set of samples from the posterior distribution over the true value of $P$ and the true isoline.
-%We plot the posterior means of $P$ and the isoline in Figure~\ref{fig:isoline}.
-%
-%
-%
-%
-%
 
@@
 In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
-The model predicts both the next word and its POS tags in each step.
+The model predicts both the next word and its language-specific POS tag in each step.
 Using POS tags is intended to prevent overfitting on small corpora.
+This choice was made before evaluating the efficiency of real grammars, and before evaluating word order properties.
 
@@
 
-TODO explain.
-Here we replicate Study 1.
-We
+Here we replicate Study 1 with both parsers and language models operating entirely on word forms, without POS tags.
+Results are shown in Figures~\ref{fig:pareto-plane-noPOS}.
+As shown in Figure~\ref{fig:byLambda-noPOS}, most real grammars have higher efficiency than most baselines across permissible values of $\lambda$.
 
@@
 \includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized-mle.pdf}
-	\caption[Predictability and Parseability]{Study 1, replication with lexicalized models: Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang}.}
+	\caption[Predictability and Parseability]{Study 1, replication with lexicalized models: Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang}.}\label{fig:pareto-plane-noPOS}
 \end{figure}
@@
 	\caption{Study 1, replication with lexicalized models: Per-language results in Study 1, with \emph{lexicalized} parsers and word-level-only language models.
-Compare Table~\ref{fig:pareto-per-lang-stats}}
+Compare Table~\ref{fig:pareto-per-lang-stats}}\label{tab:pareto-plane-noPOS}
 \end{table}
@@
 	\includegraphics[width=\textwidth]{../results/plane/unlexicalized/analyze_pareto_optimality/figures/quantileByLambda-mle.pdf}
-	\caption{Study 1, replication with lexicalized models: Optimality by $\lambda$ (lexicalized models).}
+	\caption{Study 1, replication with lexicalized models: Optimality by $\lambda$ (lexicalized models).}\label{fig:byLambda-noPOS}
 \end{figure}
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../results/plane/pureUD/pareto-plane-perLanguage-pureUD-mle.pdf}
+	\caption[Predictability and Parseability]{Study 1, replication with the original UD format models: Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang-pureUD}.}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{../results/plane/pureUD/analyze_pareto_optimality/figures/quantileByLambda-mle.pdf}
+	\caption{Study 1, replication with original UD format: Optimality by $\lambda$ (lexicalized models).}
+\end{figure}
+
 \subsection{Nondeterministic Baseline Grammars}
+
+In Study 1, we considered deterministic ordering grammars, even though real languages have some degree of word order freedom.
+Here, we show that deterministic baseline grammars provide conservative baselines: They have higher efficiency than baseline grammars with word order freedom comparable to real languages.
+
+In order to obtain baselines whose freedom is comparable to that of real languages, we constructed baselines that have the same Branching Direction Entropy \cite{futrell2015quantifying}:
+For every one of the 37 syntactic relations, we chose $a_\tau$ the direction entropy to be the same as observed in the actual orderings found in the UD corpus.
+We kept the relative ordering of siblings on one side deterministic, i.e., we considered the limit where the values $b_\tau$ for different $\tau$ are far apart.
+
+Results are shown in Figure~\ref{fig:nondeterministic-english}.
+For every one of the baseline grammars, we show both its deterministic and its nondeterministic version.
+Nondeterministic grammars are strongly less efficient than deterministic grammars, showing that deterministic grammars constitute stronger baselines.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.4\textwidth]{../results/plane/nondeterministic/pareto-plane-perLanguage-nondeterministic-mle.pdf}
+	\caption{Parseability and predictability, including nondeterministic versions of the 50 baseline grammars.}\label{fig:nondeterministic-english}
+\end{figure}
 
@@
 \end{document}
+
+%\paragraph{Analysis 2: Quantifying Degree of Pareto Optimality}
+%
+%TODO maybe we can scrap this part as it is harder to explain well, if Analysis 1 sufficient.
+%
+%
+%\begin{figure}
+%\centering
+%	\begin{tabular}{llll}
+%		(a) & (b) & (c) & (d) \\
+%		\includegraphics[width=0.22\textwidth]{figures/isocline1.png}&
+%		\includegraphics[width=0.22\textwidth]{figures/isocline2.png}&
+%		\includegraphics[width=0.22\textwidth]{figures/isocline3.png}&
+%\includegraphics[width=0.22\textwidth]{figures/isocline4.png}
+%	\end{tabular}
+%
+%	\caption{Isolines: (a) The set of all possible baseline grammars forms a distribution in the parseability-predictability space.
+%	The Pareto frontier consists of those points that are not dominated by any baseline grammar -- that is, those for which $P=0$: For points on this curve, no baseline grammar is both to their top and their right.
+%	(b) Other grammars are dominated by some fractions $P > 0$ of baseline grammars.
+The two green crosses mark grammars that both are dominated by $P=0.05$ of baseline grammars.
+A blue curve connects all grammars dominated by exactly $P=0.05$ of baseline grammars.
+%	(c) We observe a sample of $N=50$ baseline grammars, sampled uniformly from the set of baseline grammars. Samples will tend to concentrate in the high-density regions of the baseline distribution.
+%	We also observe the efficiency of the real language (red cross) and an estimate of the Pareto frontier (blue line).
+%	(d) We use the $50$ samples from the baseline grammar distribution to construct an estimate of the full abseline distibution, using the Dirichlet Process Gaussian Mixture Model.
+Using this inferred distribution, we estimate what fraction $P$ of baseline grammars dominate the real language, and the curve connecting all grammars dominated by the same fraction $P$ of baselines.
+%	}\label{fig:isoline-explanation}
+%\end{figure}
+%
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-isoCurve-smooth-interval.pdf}
+%	\caption{Real languages, even those that are not at the estimated Pareto frontier, are Pareto-dominated by very few other hypothetical grammars.
+Green curves: Baseline distribution, estimated using Kernel Density estimation.
+Black numbers indicate the fraction of baselines dominating the real language, with a 95\% credible interval, estimated using a nonparametric Bayesian method (see text).
+Orange curves connect points estimated to be dominated by the same fraction of baselines as the real language, estimated using the same Bayesian method (see text).
+}\label{fig:isoline}
+%\end{figure}
+%
+%
+%%The Pareto curve is formed by those grammars such that no grammar simultaneously achieves higher parseability and higher predictability.
+%%This motivates 
+%
+%We quantify the Pareto-optimality of the language $r$ as the fraction of baselines that Pareto-dominate $r$ -- that is, baselines that are at least as good as $r$ on \emph{both} parseability and predictability:
+%\begin{equation}
+%P := \frac{\# \{b \in B: R_{Pars}(b) \geq R_{eff}(r) \text{ and }  R_{Pred}(b) \geq R_{Pred}(r) \}}{\#B}
+%\end{equation}
+%%This quantifies how many hypothetical grammars are more efficient than $r$ for \emph{all} choices of $\lambda$ (make a figure).
+%%This quantity is also known as the joint (or bivariate) exceedance probability.
+%If we have access to the full distribution over baseline grammars, then the Pareto frontier is formed by exactly those grammars where $P=0$ (make a figure).
+%(TODO slight subtlety: $r$ itself is part of $B$, so $P > 0$ TODO change definition)
+%Small values of $P$ indicate that only few baseline grammars are better than $r$.
+%%In practice, we only have a limited sample of $50$ baseline grammar samples; so only an approximation of $P$ is actually observable, and this approximation will be $0$ in a larger region of the plane (make a figure).
+%
+%In this section, we aim to estimate $P$, provide uncertainty intervals.
+%In order to also compare with the values of $P$ attained by baseline languages, we also estimated the isoline (cotour line) of grammars having the same $P$ value as the real language.
+%
+%%$P$ is also known as the \emph{joint} or \emph{bivariate} \emph{exceedance probability}, and there are a range of methods for estimating it.
+%
+%We used a simple nonparametric Bayesian approach to estimate $P$ from the 50 observed baseline samples.
+%The baseline languages form a distribution in the parseability-predictability plane, from which 50 independent samples are known.
+%We model the underlying distribution using the Dirichlet Process Gaussian Mixture Model (DPGMM), a flexible nonparametric model that can accomodate multimodal and other non-normal distributions \citep{rasmussen2000infinite, escobar1995bayesian, muller1996bayesian}.
+%This model is a Gaussian mixture model where the number of components itself is a latent parameter inferred from the data.
+%DPGMMs can be seen as a Bayesian version of Kernel Density Smoothing.\footnote{A potential objection to using such Kernel Density Estimation is that it models a continuous distribution with support everywhere, while the distribution of grammars is discrete and finite, in particular bounded by a Pareto frontier.
+We believe that this does not cause problems for this analysis.
+First, the number of possible grammars is extremey large, exceeding $37!$, which is more than $10^{43}$, suggesting that the distribution in the parseability-predictability plane is closely approximated by a continuous distribution.
+Second, the approximate Pareto frontier tends to be clearly separated from the high-density regions of the baseline distribution, in which most sampled baselines fall, showing that the baseline distribution has very small density around and beyond the Pareto frontier.}
+%In our context here, the reason for using DPGMMs instead of Kernel Density Smoothing is that Bayesian inference using the DPGMM provides a straightforward method of deriving estimates and credible intervals for $PO$ and the isoline.
+%
+%% http://mlg.eng.cam.ac.uk/pub/pdf/GoeRas10.pdf
+%% TODO report prior
+%
+%The DPGMM defines a prior over possible baseline distributions in the parseability-predictability plane.
+%%Under this prior, each possible distribution is parameterized by a number $n$ of components, $n$ bivariate Gaussians $\mu_i, \Sigma_i$, and $n$ weights $w_i$ for these components ($\sum_{i=1}^n w_i=1$).
+%Each possible distribution is parameterized as $\sum_{i=1}^n w_i p_{\mu_i, \Sigma_i}(x)$ where $p_{\mu_i, \Sigma_i}$ is the bivariate Gaussian density.
+%The number $n$ of components is inferred from the data together with the weights $w_i$, the means $\mu_i$, and the covariance matrices $\Sigma_i$.
+%A Dirichlet process prior (TODO) is placed on $n$, favoring models with smaller numbers of components unless the data provides evidence for more components.
+%We use the $50$ observed baseline samples to infer a posterior distribution over such mixture distributions.
+%We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
+%We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples, from which we selected an evenly spaced subset of 100 samples.
+%
+%For each of these samples, we then computed $P$ of the real language, and the isoline.
+%The collection of these results provides a set of samples from the posterior distribution over the true value of $P$ and the true isoline.
+%We plot the posterior means of $P$ and the isoline in Figure~\ref{fig:isoline}.
+%
+%
+%
+%
+%

### 2019-10-12 writeup/pnas-word-order-si.tex
code

@@
 
-TODO CITE \cite{mielke-et-al-2019} % However,  a lan-guage’s vocabulary size and the length in charactersof its sentences were well-correlated with difficultyon our large set of languages.
-\cite{gerz2018language}
+%TODO CITE \cite{mielke-et-al-2019} % However,  a lan-guage’s vocabulary size and the length in charactersof its sentences were well-correlated with difficultyon our large set of languages.
+%\cite{gerz2018language}
 % TODO maybe also some recent stuff in data-driven induction of typology https://www.mitpressjournals.org/doi/abs/10.1162/COLI_a_00357
@@
 \caption[Predictability and Parseability]{Predictability and parseability of 51 languages.
-Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.
-TODO adapt for colorblindness.}\label{fig:pareto-per-lang}
-\end{figure}
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
-%\caption[Predictability and Parseability]{Predictability and parseability of 51 languages.
-Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.
-TODO adapt for colorblindness.}\label{fig:pareto-per-lang}
-%\end{figure}
+Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.}\label{fig:pareto-per-lang}
+\end{figure}
 
@@
 }
-	\caption{TODO}\label{tab:study1}
+	\caption{Models estimating the log-odds of a random baseline grammar improving over a real grammar in efficiency ($\lambda=0.9$), surprisal, or parseability, with random effects for languages and language families.
+The strongly negative estimates of $\beta$ confirm that, across languages and language families, real grammars improve over most baselines.
+TODO}\label{tab:study1-glmer}
 \end{table}
-% paretoByFamily.R
-
-%\paragraph{Inferring $\lambda$}
-%We choose $\lambda$ to minimize the number of baselines that have better values of $R_\lambda := R_{Pars} + \lambda R_{Pred}$ than the real language:
-%\begin{equation}
-%	O := \min_\lambda \# \{b \in B: R_{eff}(b) > R_{eff}(r)\}
-%\end{equation}
-%where $r$ is the real language, and $B$ is the set of baselines.
-%Naively choosing $\lambda$ to minimize this quantity from the observed $50$ baseline samples will bias $O$ towards overestimating the optimality of the real language.
-%
-%TODO maybe add a figure explaining the Bayesian story a bit more -- how we're using a continuous distribution to smooth out the observed samples?
-%
-%We eliminate this anti-conservative bias by conducting joint Bayesian inference of $\lambda$ of $O$, using the Dirichlet Gaussian Process, a Bayesian version of Kernel Density Smoothing.
-%The 50 samples represent an underlying baseline distribution in the predictability-parseability plane.
-%We approximate this underlying distribution as a continuous distribution from the Dirichlet Gaussian Process, and infer its parameters using Bayesian inference.
-%We use this method to obtain a Bayesian estimate of the distribution of baseline distributions in the Surprisal-Parseability plane.
-%The Dirichlet Gaussian Process (CITE) defines a prior over possible baseline distributions.
-%Under this prior, each possible distribution is parameterized by a number $n$ of components, and $n$ Gaussians $\mu_n, \sigma^2_n$.
-%Due to the flexible number of components, this nonparametric prior can accomodate a wide range of continuous probability distributions, including multimodal ones.
-%We use the $50$ observed baseline samples to infer a posterior distribution over such smooth distributions.
-%We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
-%We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples.
-%
-%For every sampled distribution -- consisting of a component number $n$ and $n$ Gaussians, we then computed 
-%\begin{equation}
-%	O := \min_\lambda P(R_{eff}(b) > R_{eff}(r))\}
-%\end{equation}
-%where we considered $\lambda \in (0,1)$ in steps of $0.5$.
-%
-%By taking the values $O, \lambda$ for the 1,000 posterior samples, we obtained an estimate of the posterior of $O$ and $\lambda$.
 
@@
 
-	\caption{TODO For $\lambda=0$ and $\lambda=0.9$, we provide lines connecting points that have the same efficiency value as the real language (red cross).
+	\caption{For $\lambda=0$ and $\lambda=0.9$, we provide lines connecting points that have the same efficiency value as the real language (red cross).
 Points to the bottom right of the lines have lower efficiency under the respective values of $\lambda$.
@@
 
+TODO
+We note that not all real grammars are on the estimated Pareto frontier.
 TODO discuss difference from semantic typology:
@@
 The intercepts $\alpha_j$ ($j=1,...8$) encode the population-level prevalence of the correlations when controlling for differences between datasets from different languages and language families; $u_{L,j}$, $v_{f_L,j}$ encode per-language and per-family deviations from the population-level intercept $\alpha_j$.
-%, and
-%$u_{L,1,...8} \sim \mathcal{N}(0, \Sigma_{lang})$,
-%$v_{f_L,1,...8} \sim \mathcal{N}(0, \Sigma_{fam})$.
-
-%As we constructed 4 verb-object and 4 object-verb grammars for each of the 51 languages, TODO...
 
@@
 
-TODO also mention: As we estimate parseability and predictability using cross-entropies, our models constitute variational approximations to the true parseability and predictability values, similar to previous work quantifying other aspects of linguistic complexity \citep{cotterell2019complexity}.
-TODO
+%TODO also mention: As we estimate parseability and predictability using cross-entropies, our models constitute variational approximations to the true parseability and predictability values, similar to previous work quantifying other aspects of linguistic complexity \citep{cotterell2019complexity}.
+%TODO
 
@@
 
-Prior work has suggested \emph{Dependency Length Minimization} (DLM) as a characteristic of efficient word order~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+Prior work has suggested \emph{Dependency Length Minimization} (DLM) as a characteristic of efficient word order~\cite{futrell2015largescale,ferrericancho2004euclidean,liu2017dependency,temperley2018minimizing}.
 This is the idea that word order minimizes the average distance between syntactically related words.
-It is known that human languages reduce dependency length compared to random baselines~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+It is known that human languages reduce dependency length compared to random baselines~\cite{liu2010dependency,futrell2015largescale,liu2017dependency,temperley2018minimizing}.
 Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,hawkins2004efficiency,futrell2017memory, futrell2017generalizing}.
@@
 Indeed, \cite{futrell2017memory} argues specifically that it emerges from efficiency optimization.
-[TODO also cite rijkhof, ferrer i cancho]
+%[TODO also cite rijkhof, ferrer i cancho]
 
@@
 We found that 100\% of grammars optimized for efficiency reduce average dependency length compared to baseline grammars ($p < 0.05$, by one-sided $t$-test). 
-This suggests that the reduction of dependency length observed in natural language is indeed predicted by efficiency maximization, confirming theoretical arguments made in prior work~\cite{hawkins1994performance,hawkins2004efficiency,futrell2017memory, futrell2017generalizing}.
+This suggests that the reduction of dependency length observed in natural language is indeed predicted by efficiency maximization, confirming theoretical arguments made in prior work~\cite{rijkhoff1986word,rijkhoff1990explaining,hawkins1994performance,hawkins2004efficiency,futrell2017memory, futrell2017generalizing}.
 Next, we constructed grammars that minimize average dependency length, using the same gradient descent method as we used for efficiency optimization (Section~\ref{sec:optim-eff}).
@@
 We find that optimizing grammars for efficiency reduces dependency length to a similar degree as found in the actual orderings in the corpora, almost up to the limit given by directly optimizing for dependency length.
-[TODO acknowledge  Ferrer i Cancho R, Liu H (2014) The risks of mixing dependency lengths from se-quences of different length.Glottotheory5(2):143–155.]
+%[TODO acknowledge  Ferrer i Cancho R, Liu H (2014) The risks of mixing dependency lengths from se-quences of different length.Glottotheory5(2):143–155.]
 We also plot more detailed results for four languages in Figure~\ref{fig:dlm-4langs}, plotting dependency length as a function of sentence length as reported in prior work \cite{futrell2015largescale}.
@@
 %
+
+% paretoByFamily.R
+
+%\paragraph{Inferring $\lambda$}
+%We choose $\lambda$ to minimize the number of baselines that have better values of $R_\lambda := R_{Pars} + \lambda R_{Pred}$ than the real language:
+%\begin{equation}
+%	O := \min_\lambda \# \{b \in B: R_{eff}(b) > R_{eff}(r)\}
+%\end{equation}
+%where $r$ is the real language, and $B$ is the set of baselines.
+%Naively choosing $\lambda$ to minimize this quantity from the observed $50$ baseline samples will bias $O$ towards overestimating the optimality of the real language.
+%
+%TODO maybe add a figure explaining the Bayesian story a bit more -- how we're using a continuous distribution to smooth out the observed samples?
+%
+%We eliminate this anti-conservative bias by conducting joint Bayesian inference of $\lambda$ of $O$, using the Dirichlet Gaussian Process, a Bayesian version of Kernel Density Smoothing.
+%The 50 samples represent an underlying baseline distribution in the predictability-parseability plane.
+%We approximate this underlying distribution as a continuous distribution from the Dirichlet Gaussian Process, and infer its parameters using Bayesian inference.
+%We use this method to obtain a Bayesian estimate of the distribution of baseline distributions in the Surprisal-Parseability plane.
+%The Dirichlet Gaussian Process (CITE) defines a prior over possible baseline distributions.
+%Under this prior, each possible distribution is parameterized by a number $n$ of components, and $n$ Gaussians $\mu_n, \sigma^2_n$.
+%Due to the flexible number of components, this nonparametric prior can accomodate a wide range of continuous probability distributions, including multimodal ones.
+%We use the $50$ observed baseline samples to infer a posterior distribution over such smooth distributions.
+%We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
+%We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples.
+%
+%For every sampled distribution -- consisting of a component number $n$ and $n$ Gaussians, we then computed 
+%\begin{equation}
+%	O := \min_\lambda P(R_{eff}(b) > R_{eff}(r))\}
+%\end{equation}
+%where we considered $\lambda \in (0,1)$ in steps of $0.5$.
+%
+%By taking the values $O, \lambda$ for the 1,000 posterior samples, we obtained an estimate of the posterior of $O$ and $\lambda$.

### 2019-10-14 writeup/pnas-word-order-si.tex

@@
 }
-%\author{
-%Michael Hahn, Daniel Jurafsky, Richard Futrell}
-
-%\correspondingauthor{Michael Hahn.\\E-mail: mhahn2@stanford.edu}
 
@@
 However, UD subsumes them under the \emph{det} relation, which is also used for other highly frequent elements, such as demonstratives and quantifiers.
-%Testing this correlation would require augmenting these corpora with detailed annotation of articles, which is beyond the scope of this study.
 The other elements (question particles and plural words) are found at most in a handful of UD languages, and are not specifically annotated in these either.
@@
 We therefore excluded this correlation here.
-%We note that efficiency makes predictions compatible with the pattern observed in our sample of 51 languages, potentially in line with the noisy-channel explanation of SVO described by~\cite{gibson2013noisy} (see Table~\ref{tab:all-predictions}).
 
@@
 The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order; these form the basis of our formalization in Table~\ref{table:greenberg-dryer}.
-%, finding 16 correlations to likely be universal.
-%Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:greenberg-dryer}.
-%Testing efficiency on the excluded correlations would be feasible with additional annotation of data.
-%For instance, one of these could be tested by labeling the existing databases for the distinction between manner adverbs from other kinds of adverbs.
-%For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).
 There are a few other well-supported word order universals that are not correlations with verb-object order.
@@
 	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages.
-Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.}\label{fig:pareto-per-lang}
+Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.
+Languages are ordered by corpus size, measured by the number of sentences in the training partition, from largest (Czech) to smallest (Irish).}\label{fig:pareto-per-lang}
 \end{figure}
@@
 
-Results for the posterior of $\beta$ are shown in Table~\ref{tab:study1}.
+Results for the posterior of $\beta$ are shown in Table~\ref{tab:study1-glmer}.
 Across language families, real grammars tend to be more efficient than almost all baseline grammars.
@@
 
+\begin{figure}[ht]
+    \centering
+{
+\centering{
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+	   V \& N \& V \& N  \\
+   \end{deptext}
+   \depedge{1}{2}{obj}
+   \depedge{1}{3}{xcomp}
+   \depedge{3}{4}{obj}
+\end{dependency}
+
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+	   N \& V \& N \& V  \\
+   \end{deptext}
+   \depedge{2}{1}{obj}
+   \depedge{2}{4}{xcomp}
+   \depedge{4}{3}{obj}
+\end{dependency}
+
+}
+}
+	\caption{An English sentence with annotated syntactic relations.}
+	\label{fig:sent-dep}
+\end{figure}
+
 \begin{figure}

### 2019-10-14 writeup/pnas-word-order-si.tex
code

@@
 
-Trees are linearized using one of two grammars: One (`Correlating') places \emph{obj}, \emph{xcomp}, and \emph{acl} dependents on the same side of the head, and (in accordance with crosslinguistic tendencies, which are predicted by DLM) places the \emph{obj} dependents closer to the head than \emph{xcomp} dependents.
+Trees are linearized using one of two grammars: One (`Correlating') places \emph{obj}, \emph{xcomp}, and \emph{acl} dependents on the same side of the head, and (in accordance with crosslinguistic tendencies) places the \emph{obj} dependents closer to the head than \emph{xcomp} dependents.
 The other grammar (`Anti-Correlating') places \emph{xcomp} and \emph{acl} dependents opposite to \emph{obj} dependents.
 
-We formulated this model as a binary-branching PCFG, and used a CKY parser to estimate $I[\mathcal{T}, \mathcal{U}]$ from 10,000 random sample sentences.
-
-We computed this for different settings of $p_\text{branching} \in [0, 0.5]$ and $p_{obj} \in [0,1]$, at $p_\text{acl} \in \{0, 0.3\}$.\footnote{For $p_\text{acl} = 0.3$, we only computed values for $p_\text{branching} \leq 0.4$, due to high computational cost on long sentences resulting from $p_\text{branching} = 0.5$ and $p_\text{acl} = 0.3$.}
-For these settings, we computed the difference in $I[\mathcal{T}, \mathcal{U}]$ between the two grammars.
+An example is provided in Figure~\ref{fig:sent-dep}.
+We show how the two grammars linearize the same syntactic dependency structure:
+The correlating grammar (left) linearizes the three relation types towards the right of the head; the anti-correlating one places \emph{obj} dependencies on the left and the other dependencies on the right.
+This example provides some intuitive idea of why the correlating grammar might lead to improved parseability:
+Note that the red token labeled `N' occupies the same structural position in both versions.
+In the anti-correlating version (right), when given only the token sequence, without the syntactic structure, this word could a priori be an \emph{obj} dependent of any of the three verbs occurring to its right.
+In the correlating version (left), this `N' token can only possibly be a dependent of the verb occurring to its left.
 
@@
    \begin{deptext}[column sep=1em]
-	   V \& N \& V \& N  \\
+	   V \& \textcolor{red}{N} \& V \& N \& V  \\
    \end{deptext}
    \depedge{1}{2}{obj}
-   \depedge{1}{3}{xcomp}
+   \depedge{2}{3}{acl}
    \depedge{3}{4}{obj}
+   \depedge{1}{5}{xcomp}
 \end{dependency}
-
 \begin{dependency}[theme = simple]
    \begin{deptext}[column sep=1em]
-	   N \& V \& N \& V  \\
+	   \textcolor{red}{N} \& N \& V \&  V \&  V   \\
    \end{deptext}
-   \depedge{2}{1}{obj}
-   \depedge{2}{4}{xcomp}
-   \depedge{4}{3}{obj}
+   \depedge{4}{1}{obj}
+   \depedge{1}{3}{acl}
+   \depedge{3}{2}{obj}
+   \depedge{4}{5}{xcomp}
 \end{dependency}
@@
 }
-	\caption{An English sentence with annotated syntactic relations.}
+	\caption{Linearizations of a syntactic dependency structure, under a correlating grammar (left) and an anti-correlating grammar (right).}
 	\label{fig:sent-dep}
 \end{figure}
+
+In order to test this intuition on the level of the entire tree distribution, we formulated this model as a binary-branching PCFG, and used a CKY parser to estimate $I[\mathcal{T}, \mathcal{U}]$ from 10,000 random sample sentences.
+
+We computed this for different settings of $p_\text{branching} \in [0, 0.5]$ and $p_{obj} \in [0,1]$, at $p_\text{acl} \in \{0, 0.3\}$.\footnote{For $p_\text{acl} = 0.3$, we only computed values for $p_\text{branching} \leq 0.4$, due to high computational cost on long sentences resulting from $p_\text{branching} = 0.5$ and $p_\text{acl} = 0.3$.}
+For these settings, we computed the difference in $I[\mathcal{T}, \mathcal{U}]$ between the two grammars.
 
@@
 This is especially the case for grammars with high $p_\text{branching}$.
+
+This simulation shows that correlations can in principle improve parseability in the controlled setting of such a model, without appeal to memory limitations; we leave a full graph-theoretical understanding to future work.
 

### 2019-10-14 results/plane/visualizePlane_perLanguage.R

@@ -17,7 +17,7 @@ dataS5 = read.csv("../../grammars/plane/plane-fixed-random3.tsv", sep="\t") %>% 
 dataS6 = read.csv("../../grammars/plane/plane-fixed-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
 dataS7 = read.csv("../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
 dataS = rbind(dataS, dataS2, dataS3, dataS4, dataS5, dataS6, dataS7)
-dataP = read.csv("../../grammars/plane/plane-parse-redone.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP = read.csv("../../grammars/plane/plane-parse-unified.tsv", sep="\t") %>% mutate(Model = as.character(Model))
 dataS = dataS %>% group_by(Language, Type, Model) %>% summarise(Surprisal = mean(Surp, na.rm=TRUE))
 dataP = dataP %>% group_by(Language, Type, Model) %>% summarise(Pars = mean(Pars, na.rm=TRUE))
 dataS = as.data.frame(dataS)
@@ -161,6 +161,16 @@ for(language in unique(subData$Language)) {
 
 subData = data.frame(Language=Language, Surprisal_z=Surprisal_z, Pars_z=Pars_z, Step=Step)
 subData$Type = "Pareto"
+
+
+
+corpusSize = read.csv("../corpus-size/corpus-sizes.tsv", sep="\t")
+languagesOrdered = corpusSize$language[order(-corpusSize$sents_train)]
+
+data$Language = factor(data$Language, levels=languagesOrdered)
+subData$Language = factor(subData$Language, levels=languagesOrdered)
+
+
 
 plot = ggplot(data %>% filter(Type %in% c("manual_output_funchead_RANDOM")) %>% filter(Surprisal_z < 3), aes(x=-Pars_z, y=-Surprisal_z, color=Type, group=Type))
 plot = plot + geom_point()

### 2019-10-14 writeup/pnas-word-order-si.tex
code

@@
 \centering
-\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage_withReal.pdf}
 	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages.
@@
 \centering
-\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized-mle.pdf}
+\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized-real.pdf}
 	\caption[Predictability and Parseability]{Study 1, replication with lexicalized models: Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang}.}\label{fig:pareto-plane-noPOS}

### 2019-10-15 writeup/pnas-word-order-si.tex
code

--- 2019-10-14 writeup/pnas-word-order-si.tex
@@
 
+\subsection{Parseability and Surprisal Metrics for Observed Orders and Extracted Grammars}
+
+We note that the numbers are below the state of the art in NLP.
+
+\begin{table}
+\centering
+\small{
+\begin{tabular}{l||ll|ll|l||ll|ll|l|lllll}
+	Language & \multicolumn{5}{c||}{Observed Orders} & \multicolumn{5}{c|}{Extracted Grammars} \\ 
+	& UAS & LAS & U.Pars. & L.Pars. & Surp. & UAS & LAS & U.Pars. & L.Pars. & Surp.\\ \hline \hline
+\input{../results/plane/results-observed-and-mle.tex }
+\end{tabular}
+}
+	\caption{Parsing and Surprisal metrics for observed orders (left), and for corpora ordered according to extracted real grammars (right).
+UAS and LAS refer to \emph{Unlabeled} and \emph{Labeled Attachment Scores}, respectively, indicating what fraction of words is assigned the correct head (UAS) or the correct head and relation label (LAS) when choosing heads and labels assigned the highest probability $p_\phi(\text{head}_i, \text{label}_i | u, i)$ (Equation~\ref{eq:pars-obj}) by the parser.
+	U.Pars refers to the average value of $- \log p_\phi(\text{head}_i| u, i)$, which is a measure of the difficulty of recovering the raw tree structure, without relation labels. 
+	L.Pars refers to the average value of $- \log p_\phi(\text{head}_i, \text{label}_i| u, i)$, measuring the difficulty of recoovering tree structures including relation labels.
+	Note that L.Pars corresponds to $\operatorname{H}[\tree|\utterance]$ normalized by the number of words.
+	Finally, Surp. refers to the average word-by-word surprisal, which corresponds to the predictability measure $\operatorname{H}[\utterance]$ normalized by the number of words.
+	}\label{tab:observed-and-extracted}
+\end{table}
+
 \section{Supplementary Analyses for Study 2}
@@
 	\log p_\phi(t|u) = \sum_{i=1}^n \log p_\phi(\text{head}_i, \text{label}_i | u, i)
-\end{equation}
+\end{equation}\label{eq:pars-obj}
 where $\text{head}_i \in \{\textsc{root}, 1,\dots,n\}$ is the index of the head of $u_i$ in the syntactic structure, and $\text{label}_i$ is its syntactic relation as formalized in UD; $\phi$ denotes the parameters estimated on the training data (see paragraph `Optimization Details').
@@
 \centering
-\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized-mle.pdf}
+\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized.pdf}
 	\caption[Predictability and Parseability]{Study 1, replication with lexicalized models: Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang}.}\label{fig:pareto-plane-noPOS}

### 2019-10-15 results/plane/reportParsingAndPredictionLosses.R

@@ -1,0 +1,125 @@
+# per language
+
+
+library(lme4)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+depl = read.csv("../../grammars/dependency_length/total_summary_funchead_coarse.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+library(tidyr)
+library(dplyr)
+depl = depl %>% filter(grepl("FuncHead", ModelName)) %>% filter(grepl("Coarse", ModelName))
+dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS3 = read.csv("../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS4 = read.csv("../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS5 = read.csv("../../grammars/plane/plane-fixed-random3.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS6 = read.csv("../../grammars/plane/plane-fixed-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS7 = read.csv("../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS = rbind(dataS, dataS2, dataS3, dataS4, dataS5, dataS6, dataS7)
+#dataP = read.csv("../../grammars/plane/plane-parse-unified.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP = read.csv("../../grammars/plane/plane-parse-redone.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS = dataS %>% group_by(Language, Type, Model) %>% summarise(Surprisal = mean(Surp, na.rm=TRUE))
[…]

### 2019-10-16 writeup/pnas-word-order-si.tex
loglinear analysis

@@
 There are a few other well-supported word order universals that are not correlations with verb-object order.
-This includes dominance principles~\cite{croft2003typology} such as the strong preference for subjects to precede objects.
+This includes dominance principles~\cite{greenberg1963universals,croft2003typology} such as the strong preference for subjects to precede objects.
 Furthermore, there has been interest in Greenberg's universals 18 and 20, which describe correlations not with verb-object order, but of different elements of noun phrases~\cite{cinque2005deriving, culbertson2014language, dryer2018order}.
@@
 
-\paragraph{Comparison to other Formalizations}
+\subsection{Comparing Efficiency to its Components}\label{sec:posterior-number}
+
+In Figure~\ref{fig:posterior}, we plot the posterior distribution of the number of correlations predicted to hold in most optimized grammars, as obtained from the Bayesian regression.
+For each posterior sample, we say that the $j$-th correlation holds if the value of $\alpha_j$ in that posterior sample is positive.
+In the figure, we plot the fraction of posterior samples in which a given number of correlations is satisfied.
+In addition to grammars optimized for efficiency, we also report the result for grammars optimized for predictability and for parseability alone.
+Efficiency predicts all eight correlations with high posterior probability; predictability and parseability alone do not.
+
+\begin{figure}[ht]
+	\begin{center}
+	\includegraphics[width=0.98\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
+	\end{center}
+	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in the Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.
+We show results for grammars optimized for only Predictability (left), only Parseability (center), and full Efficiency (right).}\label{fig:posterior}
+\end{figure}
+
+\subsection{Results on all UD Relations}
+In this section, we provide the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
+We also report results for grammars optimized for predictability and parseability individually.
+
+We considered all UD syntactic relations occurring in at least two of the 51 languages.
+In Table~\ref{tab:all-predictions-1}, we present the data for the eight correlations discussed in the main paper, and for those other relations for which the typological literature provides data.\footnote{The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
+Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
+For simplicity, we display this as a corelation in this table.}
+Additionally, in Table~\ref{tab:all-predictions-2} we present data for the other UD relations, for which either no typological data is available, or which are not linguistically meaningful.
+
+\begin{table*} 
+	\begin{center}	
+\includegraphics[width=0.8\textwidth]{si-table-perrel-1a-1.png}
+\end{center}
+\caption{Predictions on UD relations with predictions from the typological literature.
+The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
+The `Real' column provides the prevalence among the 51 languages in the Universal Dependencies data.
+We provide posterior prevalences for grammars optimized for Efficiency, and for grammars optimized for Pars(eability) and Pred(ictability) alone, obtained from the Bayesian mixed-effects analysis controlling for languages and language families (as in Table 2 of the main paper).
+In the last column, we indicate what prevalence is expected according to the typological literature. 
+}
+\label{tab:all-predictions-1}
+\end{table*}
+
+\begin{table*} 
+	\begin{center}	
+\includegraphics[width=0.7\textwidth]{si-table-perrel-2a-1.png}  
+\end{center}
+\caption{Predictions on UD relations for which no predictions are available in the typological literature.
+``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
+``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).
+We provide results for grammars optimized for Efficiency, and for grammars optimized for Pars(eability) and Pred(ictability) alone.
+}
+\label{tab:all-predictions-2}
+\end{table*}
+
+\subsection{Previous Experiments}\label{sec:previous-exps}
+In Table~\ref{table:corr-resu-previous} we report the results of our two previous, preregistered, simulations\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}, \url{https://aspredicted.org/blind.php?x=bg35x7}.
+For the results of the Locality simulations described in the first preregistration, see the Dependency Length Minimization results in Table~\ref{tab:all-predictions-1b}, with discussion in Section~\ref{sec:DLM}.} together with results from the main experiment.
+These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) one simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
+Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
+
+\begin{table}[hbt!]
+
+	\begin{center}
+\includegraphics[width=0.4\textwidth]{si-table-perrel-3-1.png}  
+\end{center}
+
+	\begin{center}
+\begin{tabular}{cccc}
+$\lambda=0.0$ & $\lambda=0.9$ & $\lambda=0.9$ & $\lambda=1.0$ \\
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency09.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-efficiency-large.pdf}&
+	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf}
+\end{tabular}
+	\end{center}
+	\caption{Results from optimization experiments for different values of $\lambda$, including our two previous preregistered experiments (Section~\ref{sec:previous-exps}).
+For comparison, we also show results for $\lambda=0$, corresponding to optimizing for parseability only (same results as reported in Tables (\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})).
+For $\lambda=0.9$, we report results from one preliminary preregistered experiment (center left) and the final experiment (center right).
+For $\lambda=1.0$, we report the other preliminary preregistered experiment.
+Giving similar weight to parseability and predictability -- that is, $\lambda$ close to $1$ -- results in more accurate word order predictions than choosing a small value of $\lambda$ such as $\lambda=0.0$.
+Note that $\lambda$ cannot take values smaller than zero, or greater than one, see Section \ref{sec:lambda}.
+}\label{table:corr-resu-previous}
+\end{table}
+
+\subsection{Comparison to other Formalizations of Greenberg's Correlations}
 
@@
 
-\citet{justeson1990explanation} conducted a log-linear analysis on typological judgments of 147 languages, constructing an undirected graphical model.
-%While they found evidence for correlations similar to those found by \citet{dryer1992greenbergian}, they found that some correlations are me
+\citet{justeson1990explanation} conducted a log-linear analysis on typological judgments of 147 languages, constructing an undirected graphical model modeling correlations among any pair of six syntactic relations (verb-object, adposition-noun, noun-genitive, noun-relative clause, noun-adjective, verb-subject).
+Results from their analysis suggested that, while some relations are directly correlated with the verb-object order, whereas other relations are only indirectly correlated with it.
+In particular, in their analysis, the noun-genitive relation (corresponding to Correlation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}} here) was not directly correlated with the verb-object correlation; instead, the typologically observed correlation was explained through correlations between the noun-genitive relation and other relations (such as the adposition-noun relation) that directly correlate with the verb-object relation.
+Note that this does not contradict the statement that verb-object and noun-genitive relations correlate; it shows that the observed correlation can be explained through a chain of other correlations.
+
+Since the set of syntactic relations examined here is different from that examined by \citet{justeson1990explanation}, we cannot directly compare the predictions of efficiency optimization with their results.
+Nonetheless, we can show that efficiency optimization is compatible with a picture of Greenberg's correlation as a network of pairwise correlations among different syntactic relations, and in particular the result that the correlation between the verb-object and noun-genitive relations is mediated through other correlations.
+
+First, we directly test the optimized grammars for two additional correlations found by \citet{justeson1990explanation}:
 For the relations examined here, beyond correlations with verb-object order, they found additional correlations between (1) the noun-genitive and adposition-noun dependencies, and (2) between the noun-relative clause and adposition-noun dependencies, \emph{beyond} the correlation mediated through the individual correlations with the verb-object dependency.
@@
 Both correlations are very strongly supported by grammars optimized for efficiency.
-Future research might apply the graphical model approach of \citet{justeson1990explanation} to optimized grammars, and test whether grammar optimization predicts not just the presence of correlations, the structure their method found on typological data.
+
+Second, we directly applied the log-linear analysis described by \citet{justeson1990explanation} to optimized grammars.
+We represent each grammar via the directions $v_1, \dots v_9$ of the nine relations indicated in Table 1 of the main paper (verb-object, and \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}), we coded these as $-0.5$ for Japanese-like order, and $+0.5$ for Arabic-like order.
+This analysis models the relative frequency $p_{(v_1, \dots, v_9)}$ of a particular configuration of such a configuration $(v_1, \dots, v_9)$ by a log-linear model:
+\begin{equation}
+	\log p_{(v_1, \dots, v_9)} = u_0 + \sum_{i=1}^9 u_i v_i + \sum_{i, j \in C} u_{i,j} v_{i} v_{j}
+\end{equation}
+where $C$ is some set of (unordered) pairs of relations $\in \{1, \dots, 9\}$, modeling those pairs of relations that directly correlate with each other, and where $u_0, i_i, i_{i,j}$ are real-valued parameters.
+For instance, if all relations directly correlate with the verb-object order, and not with any other relation, $C$ would contain all the unordered pairs containing the verb-object relation.
+
+We inferred the best-fitting such model by selecting the pairs in $C$ via forward-selection using AIC.
+The best-fitting model includes a set $C$ of 13 correlating pairs, with $AIC=274.18$.
+This resulting model is shown in Figure~\ref{fig:loglinear}; following \cite{justeson1990explanation}, we show those links between nodes that are included in this selected model.
+In agreement with the results of \cite{justeson1990explanation}, a network is identified in which all relations are connected at least indirectly, but several relations are not directly connected to the verb-object relation:
+In particular, in accordance with the typological data analysed by \cite{justeson1990explanation}, the observed correlation between the verb-object and noun-genitive relation is entirely mediated through correlations with other relations (adposition-noun and verb-adpositional phrase) that directly correlate with the verb-object relation.
+A difference is that, in our analysis and unlike the analysis by \cite{justeson1990explanation}, the noun-relative clause dependency is not directly linked to the verb-object relation; this might be because our analysis takes a different set of relations into account compared to \cite{justeson1990explanation}.
+
+We also note that, unlike our mixed-effects models, this log-linear model does not have random effects, as we found that adding random effects to the log-linear model led to nonconvergence.
+This means that it does not account for differences in the tree structures between languages and language families; as a result, the mixed-effects analyses for individual correlation pairs may be more conservative than this log-linear model.
+Future work should replicate the analysis of \cite{justeson1990explanation} on a larger typological database and with more relations, to enable a direct comparison with the network structure predicted by  efficiency optimization.
 
@@
 
-\subsection{Comparing Efficiency to its Components}\label{sec:posterior-number}
-
-In Figure~\ref{fig:posterior}, we plot the posterior distribution of the number of correlations predicted to hold in most optimized grammars, as obtained from the Bayesian regression.
-For each posterior sample, we say that the $j$-th correlation holds if the value of $\alpha_j$ in that posterior sample is positive.
-In the figure, we plot the fraction of posterior samples in which a given number of correlations is satisfied.
-In addition to grammars optimized for efficiency, we also report the result for grammars optimized for predictability and for parseability alone.
-Efficiency predicts all eight correlations with high posterior probability; predictability and parseability alone do not.
-
-\begin{figure}[ht]
-	\begin{center}
-	\includegraphics[width=0.98\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-three.pdf}
-	\end{center}
-	\caption{Posterior of the number of correlations correctly predicted by efficiency and its components, in the Bayesian multivariate mixed-effects logistic regression with random effects for languages and language families.
-We show results for grammars optimized for only Predictability (left), only Parseability (center), and full Efficiency (right).}\label{fig:posterior}
-\end{figure}
-
-\subsection{Results on all UD Relations}
-In this section, we provide the predicted prevalence of correlations between the \emph{obj} dependency and all UD dependency types, along with the expected prevalence according to typological studies.
-We also report results for grammars optimized for predictability and parseability individually.
-
-We considered all UD syntactic relations occurring in at least two of the 51 languages.
-In Table~\ref{tab:all-predictions-1}, we present the data for the eight correlations discussed in the main paper, and for those other relations for which the typological literature provides data.\footnote{The \emph{aux} syntactic relation in UD has the auxiliary (verb-patterner) as its dependent, and has direction \emph{opposite} to the auxiliary-verb relation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}.
-Therefore, this relation is \emph{anti-correlated} with the verb-object relation, while \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} is \emph{correlated}.
-For simplicity, we display this as a corelation in this table.}
-Additionally, in Table~\ref{tab:all-predictions-2} we present data for the other UD relations, for which either no typological data is available, or which are not linguistically meaningful.
-
-\begin{table*} 
-	\begin{center}	
-\includegraphics[width=0.8\textwidth]{si-table-perrel-1a-1.png}
-\end{center}
-\caption{Predictions on UD relations with predictions from the typological literature.
-The first section contains the eight correlations discussed in the main paper (See Section~\ref{sec:correlations}); the second section provides other relations for which predictions are available.
-The `Real' column provides the prevalence among the 51 languages in the Universal Dependencies data.
-We provide posterior prevalences for grammars optimized for Efficiency, and for grammars optimized for Pars(eability) and Pred(ictability) alone, obtained from the Bayesian mixed-effects analysis controlling for languages and language families (as in Table 2 of the main paper).
-In the last column, we indicate what prevalence is expected according to the typological literature. 
-}
-\label{tab:all-predictions-1}
-\end{table*}
-
-\begin{table*} 
-	\begin{center}	
-\includegraphics[width=0.7\textwidth]{si-table-perrel-2a-1.png}  
-\end{center}
-\caption{Predictions on UD relations for which no predictions are available in the typological literature.
-``Uninterpretable'' UD relations are those which collapse so many different linguistic relationships that they are not linguistically meaningful.
-``UD artifact'' relations are those whose order is determined strictly by UD parsing standards, such that their order is not linguistically meaningful: these include dependencies such as the connection between two parts of a word that have been separated by whitespace inserted as a typo (\emph{goeswith}).
-We provide results for grammars optimized for Efficiency, and for grammars optimized for Pars(eability) and Pred(ictability) alone.
-}
-\label{tab:all-predictions-2}
-\end{table*}
-
-\subsection{Previous Experiments}\label{sec:previous-exps}
-In Table~\ref{table:corr-resu-previous} we report the results of our two previous, preregistered, simulations\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}, \url{https://aspredicted.org/blind.php?x=bg35x7}.
-For the results of the Locality simulations described in the first preregistration, see the Dependency Length Minimization results in Table~\ref{tab:all-predictions-1b}, with discussion in Section~\ref{sec:DLM}.} together with results from the main experiment.
-These experiments all had the same setup described in Section~\ref{sec:neural-architectures}, which was fixed before starting simulations; differences are that (1) one simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
-Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
-
-\begin{table}[hbt!]
-
-	\begin{center}
-\includegraphics[width=0.4\textwidth]{si-table-perrel-3-1.png}  
-\end{center}
-
-	\begin{center}
-\begin{tabular}{cccc}
-$\lambda=0.0$ & $\lambda=0.9$ & $\lambda=0.9$ & $\lambda=1.0$ \\
-	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}&
-	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency09.pdf}&
-	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-efficiency-large.pdf}&
-	\includegraphics[width=0.15\textwidth]{../results/correlations/figures/posterior-satisfied-universals-together-large-prior-efficiency10.pdf}
-\end{tabular}
-	\end{center}
-	\caption{Results from optimization experiments for different values of $\lambda$, including our two previous preregistered experiments (Section~\ref{sec:previous-exps}).
-For comparison, we also show results for $\lambda=0$, corresponding to optimizing for parseability only (same results as reported in Tables (\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})).
-For $\lambda=0.9$, we report results from one preliminary preregistered experiment (center left) and the final experiment (center right).
-For $\lambda=1.0$, we report the other preliminary preregistered experiment.
-Giving similar weight to parseability and predictability -- that is, $\lambda$ close to $1$ -- results in more accurate word order predictions than choosing a small value of $\lambda$ such as $\lambda=0.0$.
-Note that $\lambda$ cannot take values smaller than zero, or greater than one, see Section \ref{sec:lambda}.
-}\label{table:corr-resu-previous}
-\end{table}
+\begin{figure}
+	\begin{center}
+	\includegraphics[width=0.8\textwidth]{../results/correlations/analysis/controls/loglinear-pairwise-correlations.pdf}
+	\end{center}
+	\caption{Network of pairwise correlations among the nine syntactic relations examined in this study, estimated from grammars optimized, identified using a log-linear model following \citet{justeson1990explanation}.
+The verb-object relation is at the center of the network.
+Relations between verbs and their dependents are colored in blue; relations between nouns and their dependents are colored in red; other relations are colored in green.}\label{fig:loglinear}
+\end{figure}
 
@@
 We found that 100\% of grammars optimized for efficiency reduce average dependency length compared to baseline grammars ($p < 0.05$, by one-sided $t$-test). 
-This suggests that the reduction of dependency length observed in natural language is indeed predicted by efficiency maximization, confirming theoretical arguments made in prior work~\cite{rijkhoff1986word,rijkhoff1990explaining,hawkins1994performance,hawkins2004efficiency,futrell2017memory, futrell2017generalizing}.
+This suggests that the reduction of dependency length observed in natural language is indeed predicted by efficiency maximization, confirming theoretical arguments made in prior work~\cite{hawkins1994performance,hawkins2004efficiency,futrell2017memory, futrell2017generalizing}.
 Next, we constructed grammars that minimize average dependency length, using the same gradient descent method as we used for efficiency optimization (Section~\ref{sec:optim-eff}).
@@
 In Table~\ref{tab:all-predictions-1b}, we report the posterior prevalence of word order correlations in grammars optimized for DLM; our results show that optimizing for DLM makes predictions similar to efficiency optimization.
-This is itself a novel result, suggesting that it is in part through favoring short dependencies that efficiency predicts word order universals, an idea that has been proposed in prior theoretical studies \citep{hawkins1994performance, hawkins2004efficiency}.
 We find that these grammars also exhibit the eight correlations, similar to grammars directly optimized for efficiency.
+This is itself a novel result, suggesting that it is in part through favoring short dependencies that efficiency predicts word order universals, an idea that has been proposed in prior theoretical studies, though never tested computationally on large-scale text data \citep{kuno1974position,rijkhoff1986word,frazier1985syntactic,rijkhoff1990explaining,hawkins1994performance, hawkins2004efficiency}.
 On other correlations, predictions of DLM also resemble those of efficiency optimization.

### 2019-10-16 writeup/pnas-word-order-si.tex
code

@@
 % TODO maybe also some recent stuff in data-driven induction of typology https://www.mitpressjournals.org/doi/abs/10.1162/COLI_a_00357
+
+TODO specifically have a discussion about differences to the semantic typology studies?
 
@@
 
-We note that the numbers are below the state of the art in NLP.
+In Table~\ref{tab:observed-and-extracted}, we report parsing and surprisal metrics that are commonly used in the NLP literature, both for the originally observed orders in the corpora, and the corpora ordered according to the real grammars as extracted and expressed in our grammar formalism.
+We observe similar performance on observed orders and the extracted grammars, across all metrics.
+We note that, while our parsing model is based on the strongest available dependency parsing method from the NLP literature \cite{dozat2017stanford,zeman2018proceedings,che2018towards}, the parsing metrics here are mostly below the best numbers reported with this architecture \cite{zeman2018proceedings} due to the use of an unlexicalized parsing model.
 

### 2019-10-17 writeup/pnas-word-order-si.tex

@@
 TODO specifically have a discussion about differences to the semantic typology studies?
+
+TODO
+We note that not all real grammars are on the estimated Pareto frontier.
+TODO discuss difference from semantic typology:
+TODO mention that, in semantic typology, random baselines might tend to be closer to the frontier, whereas, in our study, there tends to be a low-density region between the baseline grammars and the optimum (as also observed in previous DLM and surprisal studies, cite Futrell and Gildea/Jaeger).
 
@@
 The UD treebanks overrepresent certain language families.
-Does this impact the results of Study 1?
-We estimate the overall degree of optimization of languages for efficiency, controlling for differences between families, we constructed a Bayesian logistic mixed-effects model estimating, for each language $L$ among the 51 UD languages, the rate $q_L \in [0,1]$ of random baseline grammars that have lower efficiency (parseability, predictability) than the real grammar.
+This raises the question of whether the relative optimality of real grammars observed in Study 1 could be due to family-specific effects.
+We address this question in this section, by estimating the overall degree of optimization of languages for efficiency, controlling for differences between families.
+To this end, we constructed a Bayesian logistic mixed-effects model estimating, for each language $L$ among the 51 UD languages, the rate $q_L \in [0,1]$ of random baseline grammars that have \emph{higher} efficiency (parseability, predictability) than the real grammar.
 We entered languages and language families as random effects:
@@
 where $f_L$ is the language family of $L$.
+Here, $\beta$ models the overall probability $logit(q_{L})$ of a baseline grammar having higher efficiency than the real grammar, controlling for differences in the tree structures and real grammars of different languages and language families.
+If optimality of real grammars holds generally across families, and exceptions are due to to language- or family-specific effects, we expect $\beta$ to be $<0$ significantly.
+On the other hand, if optimality of real grammars does not generally hold across families, and the observed optimality is due to family-specific effects, then we expect $\beta \geq 0$.
+
+We estimated the mixed-effects model (\ref{eq:mixed-effects-study1}) from the 50 baseline grammars for each language, using the same priors and sampling method as in the analysis in Study 2 (reported in Section~\ref{sec:study2-regression}).
 
 Results for the posterior of $\beta$ are shown in Table~\ref{tab:study1-glmer}.
-Across language families, real grammars tend to be more efficient than almost all baseline grammars.
-According to this analysis, exceptions among the UD corpora can be attributed to language-specific effects.
-
-TODO also report glmer, posterior plots, transformed probabilities
+For all three models, $\beta$ is estimated to be $< 0$, showing that the observed optimality of real grammars holds across families, and exceptions are due to language- or family-specific effects.
+For instance, for efficiency, the posterior mean estimate $\beta=-5.88$ corresponds to  less than $1\%$ of baseline grammars showing higher efficiency than the real grammar, when controlling for language- and family-specific effects.
+Similar results hold for predictability and parseability individually.
+
+%TODO also report glmer, posterior plots, transformed probabilities
 
@@
 	Efficiency ($\lambda=0.9$) &	-5.88   &   1.08  &  -8.28  &  -3.97  \\
-	Surp & -3.48   &   0.88  &  -5.42  &  -1.85   \\
-	Pars & -5.55   &   1.08  &  -7.80  &  -3.67    \\
+	Predictability & -3.48   &   0.88  &  -5.42  &  -1.85   \\
+	Parseability & -5.55   &   1.08  &  -7.80  &  -3.67    \\
 \hline
@@
 	\caption{Models estimating the log-odds of a random baseline grammar improving over a real grammar in efficiency ($\lambda=0.9$), surprisal, or parseability, with random effects for languages and language families.
-The strongly negative estimates of $\beta$ confirm that, across languages and language families, real grammars improve over most baselines.
-TODO}\label{tab:study1-glmer}
+The strongly negative estimates of $\beta$ confirm that, across languages and language families, real grammars improve over most baselines.}\label{tab:study1-glmer}
 \end{table}
@@
 
-\begin{figure}
-\centering
-		\includegraphics[width=0.5\textwidth]{lambda-explanation.png}
-
-	\caption{For $\lambda=0$ and $\lambda=0.9$, we provide lines connecting points that have the same efficiency value as the real language (red cross).
+%\begin{figure}
+%\centering
+%		\includegraphics[width=0.5\textwidth]{lambda-explanation.png}
+%
+%	\caption{For $\lambda=0$ and $\lambda=0.9$, we provide lines connecting points that have the same efficiency value as the real language (red cross).
 Points to the bottom right of the lines have lower efficiency under the respective values of $\lambda$.
 In this example, all baseline samples (green dots) have lower efficiency than the real language, even though the real language does not outperform every baselines on both components of efficiency.}\label{fig:opt-lambda-explanation}
-\end{figure}
+%\end{figure}
 
@@
 \includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/figures/quantileByLambda.pdf}
-	\caption[]{The $x$-axis shows $\lambda \in [0,1)$, the $y$-axis shows the fraction of baselines that have lower efficiency at this value of $\lambda$, with 95\% confidence bands obtained from a binomial test.}\label{fig:lambda-quantile}
-\end{figure}
+	\caption[]{Optimality of real grammars for efficiency, compared to baselines, across values of $\lambda$: The $x$-axis shows $\lambda \in [0,1)$, the $y$-axis shows the fraction of baselines that have lower efficiency than the real grammar at this value of $\lambda$, with 95\% confidence bands obtained from a two-sided binomial test.}\label{fig:lambda-quantile}
+\end{figure}
+
+%In Analysis 2, we quantify what fraction of baseline languages \emph{Pareto-dominates} the real language, showing that -- while not all real languages may be strongly optimized for both objectives individually -- at most a very small fraction of baseline grammars improve on the real language on both components.
+%\paragraph{Analysis 1: Optimality by $\lambda$}
+%Second, we quantify how many baseline languages are more efficient across all values of $\lambda$.
+%Recall the efficiency metric
+%\begin{equation}
+%	R_\lambda := R_{Pars} + \lambda R_{Pred}
+%\end{equation}
+% with the tradeoff parameter $\lambda \in [0,1)$.
+%For each $\lambda$, we quantify the fraction of baselines that have better efficiency than the real language:
+%\begin{equation}
+%	O_\lambda := \frac{\# \{b \in B: R_{\lambda}(b) > R_{\lambda}(r)\}}{\# B}
+%\end{equation}
+%where $B$ is the set of baseline languages.
+%We illustrate this in Figure~\ref{fig:opt-lambda-explanation}.
+
+%E.g., the real grammar of Czech does not improve over all baselines in predictability, but it is higher than the vast majority of baselines in efficiency, for all values $\lambda \in [0,1]$.
+%This highlights that, even when the real grammar does not appear strongly optimized at all for one individual component, it may still be more efficient than all baselines.
 
@@
 Efficiency is a combination of both components; in this section we address the question whether languages are generally optimized for efficiency, i.e., the multi-objective optimization problem of optimizing for parseability and predictability.
-%Note that the numerical value of efficiency depends on the parameter $\lambda \in [0,1)$.
-
-We quantify degree of optimality in two ways.
-In Analysis 1, we consider the efficiency metric $R_\lambda := R_{Pars} + \lambda R_{Pred}$ with the tradeoff parameter $\lambda \in [0,1)$.
-For each possible value of $\lambda \in [0,1)$ trading off parseability and predictability, we quantify how many baseline grammars are less efficient than the real language.
-In Analysis 2, we quantify what fraction of baseline languages \emph{Pareto-dominates} the real language, showing that -- while not all real languages may be strongly optimized for both objectives individually -- at most a very small fraction of baseline grammars improve on the real language on both components.
-
-\paragraph{Analysis 1: Optimality by $\lambda$}
-%Second, we quantify how many baseline languages are more efficient across all values of $\lambda$.
-
 Recall the efficiency metric
@@
 with the tradeoff parameter $\lambda \in [0,1)$.
-For each $\lambda$, we quantify the fraction of baselines that have better efficiency than the real language:
-\begin{equation}
-	O_\lambda := \frac{\# \{b \in B: R_{\lambda}(b) > R_{\lambda}(r)\}}{\# B}
-\end{equation}
-where $B$ is the set of baseline languages.
-We illustrate this in Figure~\ref{fig:opt-lambda-explanation}.
+For each possible value of $\lambda \in [0,1)$ trading off parseability and predictability, we quantify how many baseline grammars are less efficient than the real language.
 
 The results are plotted in Figure~\ref{fig:lambda-quantile}.
-
-For all languages, there are some values of $\lambda$ where the real language improves at least on more than the median of the baseline grammars, i.e., $O_\lambda >0$.
-In about 40 of the languages, $O_\lambda$ is even close to one across all values of $\lambda$.
-
+For all languages, there are some values of $\lambda$ where the real grammar improves at least half of the baseline grammars.
+In about 40 of the languages, the real grammar improves over almost all baseline grammars and for all values of $\lambda$.
 This shows that, while many languages do not improve over \emph{all} baselines on \emph{both} individual components, they mostly improve over the large majority of baselines on the combined objective of efficiency, even across different values of $\lambda$.
+For instance, the real grammar of Czech does not improve over all baselines in predictability (see Figure~\ref{fig:pareto-per-lang}), but it has higher overall efficiency than the vast majority of baselines in efficiency, for all values $\lambda \in [0,1)$.
+
+We further confirm this in Figure~\ref{fig:lambda-halfplane-09}:
+We plot the rean and optimized grammars together with a Kernel density estimate of the distribution of baseline grammars.
+We add lines connecting those points that have the same efficiency $R_\lambda$ as the real grammar, at very low ($\lambda = 0.0$, dotted line) and very high ($\lambda = 0.9$, dashed line) values of $\lambda$.
+Grammars to the bottom/left of this lines have lower efficiency than the real grammar, at these two given values of $\lambda$.
+The distribution of baseline grammars is largely to the bottom/left of at least one of the two lines, and often to the bottom/left of both lines.
+This highlights that, even when the real grammar does not appear strongly optimized at all for one individual component, it may still be more efficient than all baselines.
 
@@
 \includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/pareto-plane-perLanguage-arrows-smoothed-halfspace-untransformed.pdf}
-	\caption[]{Per-Language results as in Figure~\ref{fig:pareto-per-lang-raw}, with lines connecting those points that have the same efficiency as the real grammar at $\lambda = 0.0$ (dotted) and $\lambda = 0.9$ (dashed).
+	\caption[]{Per-Language results as in Figure~\ref{fig:pareto-per-lang-raw}, representing the distribution of baseline grammars with a Kernel density estimate.
+We add lines connecting those points that have the same efficiency as the real grammar at $\lambda = 0.0$ (dotted) and $\lambda = 0.9$ (dashed).
 Points to the bottom/left of these line have lower efficiency than the real grammar, at the given value of $\lambda$.}\label{fig:lambda-halfplane-09}
 \end{figure}
-
-TODO
-We note that not all real grammars are on the estimated Pareto frontier.
-TODO discuss difference from semantic typology:
-TODO mention that, in semantic typology, random baselines might tend to be closer to the frontier, whereas, in our study, there tends to be a low-density region between the baseline grammars and the optimum (as also observed in previous DLM and surprisal studies, cite Futrell and Gildea/Jaeger).
 
@@
 
-\subsection{Regression for Predicted Correlations}
+\subsection{Regression for Predicted Correlations}\label{sec:study2-regression}
 
@@
 Indeed, \cite{futrell2017memory} argues specifically that it emerges from efficiency optimization.
-%[TODO also cite rijkhof, ferrer i cancho]
 
@@
 We expect that such grammars should have shorter dependency length than the real grammars, or grammars optimized for efficiency.
-In Figure~\ref{fig:dlm-avg}, we plot the mean dependency length for optimized, real, and baseline orderings.
+In Figure~\ref{fig:dlm-avg}, we plot the mean dependency length for optimized, real, and baseline orderings.\footnote{We show results for the actually observed orderings, not for corpora ordered according to extracted grammars as in Study 1; results are similar for those extracted grammars.}
 We find that optimizing grammars for efficiency reduces dependency length to a similar degree as found in the actual orderings in the corpora, almost up to the limit given by directly optimizing for dependency length.
-%[TODO acknowledge  Ferrer i Cancho R, Liu H (2014) The risks of mixing dependency lengths from se-quences of different length.Glottotheory5(2):143–155.]
-We also plot more detailed results for four languages in Figure~\ref{fig:dlm-4langs}, plotting dependency length as a function of sentence length as reported in prior work \cite{futrell2015largescale}.
+We also plot more detailed results for four languages in Figure~\ref{fig:dlm-4langs}, plotting dependency length as a function of sentence length as reported in prior work \cite{ferrericancho2014risks,futrell2015largescale}.
 Optiziming grammars for efficiency produces dependency lengths similar to those found in the actual orderings.

### 2019-10-17 writeup/pnas-word-order-si.tex
code

@@
 
-Here we replicate Study 1 with both parsers and language models operating entirely on word forms, without POS tags.
-Results are shown in Figures~\ref{fig:pareto-plane-noPOS}.
-As shown in Figure~\ref{fig:byLambda-noPOS}, most real grammars have higher efficiency than most baselines across permissible values of $\lambda$.
+In Study 1, we calculate parseability on the part-of-speech level, and also add part-of-speech tags when calculating predictability.
+These choices are intended to prevent early overfitting during the grammar optimization process (Section~\ref{sec:neural-architectures}).
+However, such unlexicalized parsers are less accurate than parsers taking acual word-forms into account, and adding part-of-speech tags might provide additional disambiguation that is absent in the original word-level input.
+Here, we show that these limitations do not affect conclusions from Study 1, by replicating Study 1 with both parsers and language models operating entirely on word forms, without POS tags.
+Results are shown in Figure~\ref{fig:pareto-plane-noPOS} and Table~\ref{tab:pareto-plane-noPOS}.
+In agreement with the previous results (Figure~\ref{fig:pareto-per-lang}), real grammars are mostly to the top right of their corresponding baselines.
+We further confirm this in Figure~\ref{fig:lambda-quantile-noPOS}, which shows that most real grammars have higher efficiency than most baselines across permissible values of $\lambda$.
+In fact, comparing Figure~\ref{fig:lambda-quantile-noPOS} to Figure~\ref{fig:lambda-quantile} suggests that optimality of real grammars is \emph{more} pronounced when modeling predictability and parseability fully on the level of word forms.
 
@@
 \centering
-\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized.pdf}
+\includegraphics[width=\textwidth]{../results/plane/unlexicalized/pareto-plane-perLanguage-lexicalized-mle.pdf}
 	\caption[Predictability and Parseability]{Study 1, replication with lexicalized models: Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang}.}\label{fig:pareto-plane-noPOS}
@@
 &  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
-\input{../results/plane/unlexicalized/per-language-statistics-unlexicalized.tex}
+\input{../results/plane/unlexicalized/per-language-statistics-unlexicalized-ground.tex}
 \end{tabular}
@@
 	\includegraphics[width=\textwidth]{../results/plane/unlexicalized/analyze_pareto_optimality/figures/quantileByLambda-mle.pdf}
-	\caption{Study 1, replication with lexicalized models: Optimality by $\lambda$ (lexicalized models).}\label{fig:byLambda-noPOS}
+	\caption{Study 1, replication with lexicalized models: Optimality of real grammars for efficiency, compared to baselines, across values of $\lambda$: The $x$-axis shows $\lambda \in [0,1)$, the $y$-axis shows the fraction of baselines that have lower efficiency than the real grammar at this value of $\lambda$, with 95\% confidence bands obtained from a two-sided binomial test.
+Compare Figure~\ref{fig:lambda-quantile}.}\label{fig:lambda-quantile-noPOS}
 \end{figure}
@@
 In Study 1, we considered deterministic ordering grammars, even though real languages have some degree of word order freedom.
-Here, we show that deterministic baseline grammars provide conservative baselines: They have higher efficiency than baseline grammars with word order freedom comparable to real languages.
-
-In order to obtain baselines whose freedom is comparable to that of real languages, we constructed baselines that have the same Branching Direction Entropy \cite{futrell2015quantifying}:
-For every one of the 37 syntactic relations, we chose $a_\tau$ the direction entropy to be the same as observed in the actual orderings found in the UD corpus.
-We kept the relative ordering of siblings on one side deterministic, i.e., we considered the limit where the values $b_\tau$ for different $\tau$ are far apart.
-
-Results are shown in Figure~\ref{fig:nondeterministic-english}.
+Here, we show that deterministic baseline grammars provide conservative baselines: They have higher efficiency than baseline grammars with word order freedom comparable to the orders found in the original corpora, and thus provide conservative baselines for comparison with other grammars.
+
+In order to obtain baselines whose freedom is comparable to that of real languages, we constructed baselines that have the same Branching Direction Entropy \cite{futrell2015quantifying} as observed in the original corpora.
+The Branching Direction Entropy measures the extent of freedom in choice between head-final and head-initial orderings, and it is a corpus-based quantitative measure of word order freedom \cite{futrell2015quantifying}.
+For a given syntactic relation, the \emph{branching direction entropy} measures the entropy of the Bernoulli variable that is $1$ whenever the head is ordered before the dependent, and $0$ if the dependent is ordered before the head.
+The branching direction entropy is $0$ if only one of the two orders can occur, and it is $\log 2$ if both orders are equally frequent.
+
+We constructed baseline grammars that match the branching direction entropies found in the original orders found in the corpora.
+To this end, we converted the baseline grammars into differentiable ordering grammars (Section~\ref{sec:diff-gramm}).
+Such grammars have parameters $a_\tau, b_\tau$ for each relation $\tau$.
+For every one of the 37 syntactic relations, we chose $a_\tau$ so as to match the the direction entropy to that observed in the actual orderings found in the UD corpus.
+For $b_\tau$, we considered the limit where the values $b_\tau$ for different relations $\tau$ are very far apart, making the relative ordering of siblings on the same side of the head fully deterministic.
+That is, these ordering grammars match word order freedom as quantified by Branching Direction Entropy, and show no additional degrees of order freedom.
+
+We compared these nondeterministic baseline grammars to their deterministic versions, for one language with relatively free order (Czech), and for two languages with relatively fixed order (English and Japanese).
+Results are shown in Figure~\ref{fig:nondeterministic}.
 For every one of the baseline grammars, we show both its deterministic and its nondeterministic version.
-Nondeterministic grammars are strongly less efficient than deterministic grammars, showing that deterministic grammars constitute stronger baselines.
+Nondeterministic grammars are less efficient than deterministic grammars, in particular in language with greater degrees of word order freedom (Czech).
+This shows that deterministic grammars constitute stronger baselines.
 
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{../results/plane/nondeterministic/pareto-plane-perLanguage-nondeterministic-mle.pdf}
-	\caption{Parseability and predictability, including nondeterministic versions of the 50 baseline grammars.}\label{fig:nondeterministic-english}
+	\includegraphics[width=0.95\textwidth]{../results/plane/nondeterministic/pareto-plane-perLanguage-nondeterministic-mle.pdf}
+	\caption{Parseability and predictability for three languages, including both deterministic (green) and nondeterministic  (blue) versions of the 50 baseline grammars.}\label{fig:nondeterministic}
 \end{figure}

### 2019-10-17 results/plane/nondeterministic/visualizePlane_perLanguage_Nondeterministic.R

@@ -1,0 +1,227 @@
+# per language
+
+
+library(lme4)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+#depl = read.csv("../../../grammars/dependency_length/total_summary_funchead_coarse.tsv", sep="\t")# %>% rename(Quality=AverageLength)
+
+# Read predictability estimates
+dataS =  read.csv("../../../grammars/plane/controls/plane-fixed-nondeterministic.tsv", sep="\t") %>% mutate(Model = as.character(Model), Type=as.character(Type)) %>% mutate(Type = ifelse(grepl("RANDOM", Type), "manual_output_funchead_RANDOM_nondet", Type)) %>% mutate(Model = paste(Model, "nondet", sep="_"))
+dataS1 = read.csv("../../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model), Type=as.character(Type)) %>% mutate(Type = ifelse(grepl("RANDOM", Type), "manual_output_funchead_RANDOM_det", Type)) %>% mutate(Model = paste(Model, "det", sep="_"))
+dataS2 = read.csv("../../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model), Type=as.character(Type)) %>% mutate(Type = ifelse(grepl("RANDOM", Type), "manual_output_funchead_RANDOM_det", Type)) %>% mutate(Model = paste(Model, "det", sep="_"))
+dataS3 = read.csv("../../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model), Type=as.character(Type)) %>% mutate(Type = ifelse(grepl("RANDOM", Type), "manual_output_funchead_RANDOM_det", Type)) %>% mutate(FullSurp=NULL) %>% mutate(Model = paste(Model, "det", sep="_"))
+dataS4 = read.csv("../../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model), Type=as.character(Type)) %>% mutate(Type = ifelse(grepl("RANDOM", Type), "manual_output_funchead_RANDOM_det", Type)) %>% mutate(FullSurp=NULL) %>% mutate(Model = paste(Model, "det", sep="_"))
[…]

### 2019-10-17 writeup/pnas-word-order-si.tex
code

@@
 
-		Finally, in Tamil, the divergence is caused by the treebank annotation convention for Tamil, where the label \emph{acl} is used to mark certain elements of compounds whose positioning differs from the participial phrases that correspond most closely to relative clauses of other languages.\footnote{In the original HamleDT \citep{zeman2012hamledt, Ramasamy:2011:TDP:1964799.1964808} version of the Tamil treebank, these relations were labeled as CC, marking compounds (\url{http://ufal.mff.cuni.cz/~ramasamy/tamiltb/0.1/dependency_annotation.html}).
+		Finally, in Tamil, the divergence is caused by the treebank annotation convention for Tamil, where the label \emph{acl} is used to mark certain elements of compounds, not for the participial phrases that correspond most closely to relative clauses of other languages.\footnote{In the original HamleDT \citep{zeman2012hamledt, Ramasamy:2011:TDP:1964799.1964808} version of the Tamil treebank, these relations were labeled as CC, marking compounds (\url{http://ufal.mff.cuni.cz/~ramasamy/tamiltb/0.1/dependency_annotation.html}).
 We did not attempt to modify this labeling convention.}
@@
 
-\subsection{Per-Language Results}
+\subsection{Details and Additional Analyses}
+
 In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.
-We also provide the raw numerical values, before $z$-scoring, in Figure~\ref{fig:pareto-per-lang-raw}.
-
-We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section~\ref{sec:lambda}), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
-
-Within each language, we estimate the Pareto curve based on the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
-% The Pareto frontier is the set of all Pareto efficient allocations, conventionally shown graphically.
-This frontier is approximate:
-The actual curve may not be piecewise linear, and it may also extend beyond the estimated curve, as the grammar optimization method is approximate.
-
 Figure 4 in the main paper shows the average of these per-language plots, with a kernel density estimate of the distribution of baseline grammars. 
-
-% say this somewhere
-Note that the predictability scores are considerably worse than the state-of-the art in Natural Language Processing; this is because most of the corpora are much smaller than those typically used in language modeling work (CITE a few things on large corpora; also cite language modeling work on many small corpora).
+In addition to the $z$-scored values in Figure~\ref{fig:pareto-per-lang} and the main paper, we also provide the raw numerical values, before $z$-scoring, in Figure~\ref{fig:pareto-per-lang-raw}.
+
+Note that, in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
+This can be attributed to two reasons: First, stochastic gradient descent introduces noise due to its stochasticity and will only approximately find an optimal solution; second, for some corpora, there may be some degree of distributional mismatch between the training partitions (on which grammars are optimized) and held-out partitions (on which efficiency is estimated).
+This in particular applies to very small corpora such as Irish (121 training sentences).
+
+\paragraph{Method applied for $z$-transforming and for estimating Pareto frontier}
+We $z$-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section~\ref{sec:lambda}), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.
+For (3-5), we choose the grammar, among all eight optimized grammars, that has the highest estimated efficiency (paresability, predictability) value.
+
+We define the \emph{Pareto frontier} as the boundary of the \emph{set of Pareto-efficient points}, that is, of those points such that no grammar (expressible in our formalism) has both higher predictability and higher parseability.
+We approximately estimate this frontier based on optimized grammars, by constructing a \emph{lower bound} on this curve from the optimized grammars:
+Among the eight grammars optimized for efficiency (at $\lambda=0.9$), we select the one with the highest estimated efficiency value; similarly for grammars optimized for parseability and predictability.
+Connecting these three grammars results in a piecewise linear curve that is guaranteed to be a lower bound on the true Pareto frontier (meaning that the true Pareto frontier can only lie above to the right of this curve).
+In cases where the grammar optimized for predictability (similarly parseability) has lower predictability (and parseability) than the grammar optimized for efficiency, we can replace its predictability value by that of the grammar optimized for efficiency: This is guaranteed to result in a point that is still Pareto-dominated by the grammar optimized for efficiency, and provides a tighter bound on the true curve.
+Th resulting frontier is guaranteed to provide a lower bound on the true Pareto frontier, but is nonetheless approximate: The actual curve may not be piecewise linear, and it may also extend beyond the estimated curve, as the grammar optimization method is approximate.
+
+%Note that the predictability scores are considerably worse than the state-of-the art in Natural Language Processing; this is because most of the corpora are much smaller than those typically used in language modeling work (CITE a few things on large corpora; also cite language modeling work on many small corpora).
 
@@
 \includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
-	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages.
-Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\footnotemark All data are $z$-scored.
-Languages are ordered by corpus size, measured by the number of sentences in the training partition, from largest (Czech) to smallest (Irish).}\label{fig:pareto-per-lang}
+	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages, ordered by corpus size, measured by the number of sentences in the training partition, from largest (Czech) to smallest (Irish).
+Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.
+All data are $z$-scored.}\label{fig:pareto-per-lang}
 \end{figure}
@@
 
-\footnotetext{Note that, in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier.
-This can be attributed to two reasons: First, stochastic gradient descent introduces noise due to its stochasticity and will only approximately find an optimal solution; second, for some corpora, there may be some degree of distributional mismatch between the training partitions (on which grammars are optimized) and held-out partitions (on which efficiency is estimated).
-This in particular applies to very small corpora such as Irish (121 training sentences).}
-
-In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
-We also conducted the analysis using a Binomial test (one-sided), testing whether the real gramar is more efficient than the \emph{median} of random grammars, avoiding any distributional assumption on the random grammars.
-As before, we used Hochberg's step-up procedure (Note that the tests for different languages are independent, as different random grammars are evaluated for each language), with $\alpha=0.05$.
+%\footnotetext{}
+
+\paragraph{Further Analysis of Optimality}
+In the main paper, we tested whether real grammars are more efficient than the mean of baseline grammars, using a $t$-test.
+We also conducted the analysis using a Binomial test (one-sided), testing whether the real gramar is more efficient than the \emph{median} of baseline grammars, avoiding any distributional assumption on the baseline grammars.
+As before, we used Hochberg's step-up procedure (Note that the tests for different languages are independent, as different baseline grammars are evaluated for each language), with $\alpha=0.05$.
 In this analysis, real grammars improved in parseability for 80\% of languages, in predictability for 69\% of languages, and in either of both in 92\% of languages ($p<0.05$, with Bonferroni correction).
@@
 
+\subsection{Impact of Tree Structures on Estimated Frontier}
+
+An anonymous reviewer notes that the shape of the estimated Pareto frontier (Figure~\ref{fig:pareto-per-lang}) seems to vary between languages, and asks how the tree structure distributions impact the shape of the estimated frontier, and whether deriving a language-independent bound on efficiency is possible.
+
+We conducted a series of linear regressions predicting (1) the predictability and parseability of the best grammar optimzied for efficiency, (2) the parseability and predictability difference between this end and the end optimized for predictability, (3) the difference between this end and the end optimized for parseability.
+For more meaningful comparison, we analyzed values normalized for sentence length as reported in Figure~\ref{fig:pareto-per-lang-raw}.
+
+We considered as independent variables the following quantities, computed on the training set: (1) median sentence length, (2) median tree depth, (3) mean arity, i.e., the mean number of dependents of each word\footnote{The \emph{median} is always 0 or 1 in the available corpora, we thus chose the mean as a more granular measure.}, (4) the unigram entropy, and (5) the logarithm of the overall number of sentences.
+
+These independent variables measure the complexity of syntactic strutures (1-3), the diversity of the vocabulary (4), and the amount of data available for constructing the neural network models (5).
+The resulting regressions are shown in Table~\ref{tab:corpus-frontier-regressions}.
+
+Among factors measuring the complexity of syntactic structures (predictors (1)-(3)), the strongest effect is a positive effect of arity on predictability ($\beta=7.76$, $SE=1.51$, $p < 0.001$), suggesting that structures with more dependents per head lead to higher achievable predictability.
+In contrast, we observe little evidence for an impact of sentence length or tree depth.
+We also observe an effect of unigram entropy (4), showing that datasets with more diverse vocabulary reduce both predictability and parseability.\footnote{For predictability, a similar result about vocabulary size and estimated surprisal across many languages is reported by \cite{mielke-et-al-2019}.}
+Finally, larger amounts of training data (5) lead to higher estimated predictability and parseability -- this is expected, as more training data enables better statistical estimation of the distribution of sentences and syntactic structures.
+More training data also increases the difference between the efficiency-optimal and the predictability-optimal ends of the estimated curve, suggesting that more training data enables more precise estimation of the different extremes of the curve.
+
+These results show that general quantitative properties of the available syntactic structures partially account for variation in the achievable parseability and predictability values.
+Note that at least some of these quantitative properties are impacted by factors external to the syntax of a language, e.g., the unigram entropy may be impacted by the genre of available texts.
+This result suggests that deriving a language-independent bound on efficiency, as is possible in studies of semantic typology where there is a language-invariant parameterization of the possible meanings (e.g., \cite{kemp2012kinship,xu2014numeral,zaslavsky2018efficient}), is not possible in general for syntactic typology.
+
+\begin{table}
+\centering
+\small{
+\begin{tabular}{l||lll|lll|lll|lll}
+	& \multicolumn{3}{c|}{Optimized for Efficiency} & \multicolumn{3}{c|}{Distance to Pred. End}   & \multicolumn{3}{c|}{Distance to Pars. End}\\
+	Predictor & $\beta$ & SE & $t$ & $\beta$ & SE & $t$  & $\beta$ & SE & $t$ \\ \hline \hline
+\input{../results/plane/corpus-frontier-regressions.tex}
+	\multicolumn{10}{l}{Significance levels: $^{*}$ : $p < 0.05$, $^{**}$ : $p < 0.01$, $^{***}$ : $p<0.001$}
+\end{tabular}
+}
+	\caption{Linear regression models predicting the position of the estimated Pareto frontier, from quantitative properties of the available syntactic tree structures.
+The top half provides models predicting predictability values, the bottom half provides models predicting parseability values.
+Columns correspond to the three pairs of independent variables defined in the text: predictability/parseability for the best grammar optimized for efficiency, the predictability/parseability distance to the end optimized for predictability, and the predictability/parseability distance to the end optimized for parseability.}\label{tab:corpus-frontier-regressions}
+\end{table}
+
+%summary(lm(A2Minus1 ~ MedianSentenceLength + UnigramEntropy + MedianTreeDepth + log(SentenceCount) +MeanArity , data=pars__))
+
 \section{Supplementary Analyses for Study 2}
@@
 	\end{center}
-	\caption{Network of pairwise correlations among the nine syntactic relations examined in this study, estimated from grammars optimized, identified using a log-linear model following \citet{justeson1990explanation}.
+	\caption{Network of pairwise correlations among the nine syntactic relations examined in this study, estimated from grammars optimized for efficiency, identified using a log-linear model following \citet{justeson1990explanation}.
 The verb-object relation is at the center of the network.
@@
 
-\section{Other Methods of Estimating Efficiency and Efficiency Optimization}
+\section{Other Methods of Estimating Efficiency and Constructing Baselines in Study 1}
 
@@
 Results are shown in Figure~\ref{fig:pareto-plane-noPOS} and Table~\ref{tab:pareto-plane-noPOS}.
+We compare real and baseline grammars; here, we do not have an estimate of the Pareto frontier, as the grammar optimization process uses part-of-speech tags (Section~\ref{sec:neural-architectures}).
 In agreement with the previous results (Figure~\ref{fig:pareto-per-lang}), real grammars are mostly to the top right of their corresponding baselines.
@@
 
-We replicate the comparison between real and random grammars in Study 1.
+As described in \emph{Materials and Methods}, we follow \cite{futrell2015largescale} in applying automated conversion of tree structures to a more standard formalism, modifying each treebank by inverting dependencies of types \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark}.
+This converted version is intended to more closely reflect assumptions about syntactic structure shared across a wide range of linguistic theories, addressing criticism of the Universal Dependencies representation \cite{osborne2019status}.
+
+In this section, we provide evidence that this conversion does not affect our results by replicating the comparison between real and baseline grammars in Study 1 using the original Universal Dependencies (UD) representation.
+As in Study 1, we represented the real grammars by extracting grammars from the observed orderings; for each language, we constructed a new set of 50 baseline grammars.
+Results are shown in Figures~\ref{fig:pareto-per-lang-pureUD} and~\ref{fig:lambda-quantile-pureUD}.
+The results agree with those found on the converted versions; across languages, real grammars are at the top-right of the baseline distributions, and (with the exception of Telugu, a language with a small corpus)
+
+%we also computed held-out likelihood for the extracted ordering grammars for the original UD structures and the converted version.
+Likelihood is higher for the converted version in 47 out of 51 languages (exceptions: Japanese, Korean, Hindi, Urdu).
 
@@
 \includegraphics[width=\textwidth]{../results/plane/pureUD/pareto-plane-perLanguage-pureUD-mle.pdf}
-	\caption[Predictability and Parseability]{Study 1, replication with the original UD format models: Predictability and parseability of 51 languages, for \emph{lexicalized} models, compare Figure~\ref{fig:pareto-per-lang-pureUD}.}
+	\caption[Predictability and Parseability]{Study 1, replication with the original UD format: Predictability and parseability of real and baseline grammars in 51 languages, compare Figure~\ref{fig:pareto-per-lang}.}\label{fig:pareto-per-lang-pureUD}
 \end{figure}
@@
 	\includegraphics[width=\textwidth]{../results/plane/pureUD/analyze_pareto_optimality/figures/quantileByLambda-mle.pdf}
-	\caption{Study 1, replication with original UD format: Optimality by $\lambda$ (lexicalized models).}
+	\caption{Study 1, replication with the original UD format: Optimality of real grammars for efficiency, compared to baselines, across values of $\lambda$: The $x$-axis shows $\lambda \in [0,1)$, the $y$-axis shows the fraction of baselines that have lower efficiency than the real grammar at this value of $\lambda$, with 95\% confidence bands obtained from a two-sided binomial test.
+Compare Figure~\ref{fig:lambda-quantile}.}\label{fig:lambda-quantile-pureUD}
 \end{figure}

### 2019-10-17 results/plane/corpusPropertiesCorrelations_NoZ.R

@@ -1,0 +1,270 @@
+
+
+library(lme4)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(dplyr)
+dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS3 = read.csv("../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS4 = read.csv("../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS5 = read.csv("../../grammars/plane/plane-fixed-random3.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS6 = read.csv("../../grammars/plane/plane-fixed-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS7 = read.csv("../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS = rbind(dataS, dataS2, dataS3, dataS4, dataS5, dataS6, dataS7)
+dataP = read.csv("../../grammars/plane/plane-parse-unified.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS = dataS %>% group_by(Language, Type, Model) %>% summarise(Surprisal = mean(Surp, na.rm=TRUE))
+dataP = dataP %>% group_by(Language, Type, Model) %>% summarise(Pars = mean(Pars, na.rm=TRUE))
+dataS = as.data.frame(dataS)
+dataP = as.data.frame(dataP)
+library(lme4)
+summary(lmer(Surprisal ~ Type + (1|Language), data=dataS %>% filter(grepl("langm", Type))))
+summary(lmer(Surprisal ~ Type + (1|Language), data=dataS %>% filter(grepl("RL", Type) | grepl("ground", Type))))
+dataS = dataS %>% mutate(Type = as.character(Type))
[…]

### 2019-10-17 writeup/pnas-word-order-si.tex
code

@@
 
-TODO specifically have a discussion about differences to the semantic typology studies?
-
-TODO
-We note that not all real grammars are on the estimated Pareto frontier.
-TODO discuss difference from semantic typology:
-TODO mention that, in semantic typology, random baselines might tend to be closer to the frontier, whereas, in our study, there tends to be a low-density region between the baseline grammars and the optimum (as also observed in previous DLM and surprisal studies, cite Futrell and Gildea/Jaeger).
+%TODO specifically have a discussion about differences to the semantic typology studies?
+
+%TODO
+%We note that not all real grammars are on the estimated Pareto frontier.
+%TODO discuss difference from semantic typology:
+%TODO mention that, in semantic typology, random baselines might tend to be closer to the frontier, whereas, in our study, there tends to be a low-density region between the baseline grammars and the optimum (as also observed in previous DLM and surprisal studies, cite Futrell and Gildea/Jaeger).
 
@@
 
-\subsection{Impact of Tree Structures on Estimated Frontier}
-
-An anonymous reviewer notes that the shape of the estimated Pareto frontier (Figure~\ref{fig:pareto-per-lang}) seems to vary between languages, and asks how the tree structure distributions impact the shape of the estimated frontier, and whether deriving a language-independent bound on efficiency is possible.
+\subsection{Impact of Tree Structures on Optimality and Estimated Frontier}
+
+\begin{figure}
+\centering
+	\includegraphics[width=0.7\textwidth]{../results/plane/pareto-plane-other-language.pdf}
+	\caption[]{Languages $L$ have grammars optimized specifically for the tree structure distributions of $L$: We show the real (cross) and baseline (dots) grammars for English and Japanese, together with the estimated Pareto frontier.
+Additionally, we plot the efficiency values obtained when applying the Japanese grammar to English tree structures (purple square, left), and when applying the English grammar to Japanese tree structures (purple square, right).
+In both languages, the respective real grammars (crosses) are more efficient than grammars from the other language (squares).
+This suggests that the grammars of languages are specifically optimized for their individual tree structure distributions.}\label{fig:english-japanese-crossed}
+\end{figure}
+
+\paragraph{Language-Dependence of Tree Structure Distribution}
+Unlike similar efficiency studies in the domain of lexical semantics~\cite{kemp2012kinship,regier2015word,zaslavsky2018efficient}, we did not derive a single universal bound for the efficiency across all 51 languages in Study 1; instead, we constructed optimized grammars individually for each language.
+In this section, we show why this is necessary:
+The efficiency of a grammar crucially depends on the tree structure distribution, and this tree structure distribution is language-specific.
+%Each language $L$ has its own distribution of tree structures; our results show that the grammar of each language $L$ is optimized to reflect the tree structures used in $L$ (see SI Section TODO).
+%Nonetheless, exploring ways of deriving universal bounds is an interesting problem for future work.
+%Figure~\ref{fig:english-japanese-crossed}
+%Here we show that different languages have different tree distributions, and provide evidence that languages are specifically optimized for their individual tree structure distributions.
+To show this, we compared the efficiency of the real grammar of English and Japanese with that obtained when applying the real grammar of \emph{the other} language.
+The results are shown in Figure~\ref{fig:english-japanese-crossed}.
+In both languages, the respective real grammars (crosses) are more efficient than grammars from the other language (squares).
+This suggests that the grammars of languages are specifically optimized for their individual tree structure distributions.
+Furthermore, it demonstrates that the tree structure distribution, and therefore the optimality of a given grammar, is language-specific.
+
+%This result suggests that deriving a language-independent bound on efficiency, as is possible in studies of semantic typology where there is a language-invariant parameterization of the possible meanings (e.g., \cite{kemp2012kinship,xu2014numeral,zaslavsky2018efficient}), is not possible in general for syntactic typology.
+
+\paragraph{Estimated Frontier and Corpus Properties}
+An anonymous reviewer notes that the shape of the estimated Pareto frontier (Figure~\ref{fig:pareto-per-lang}) seems to vary between languages, and asks how the tree structure distributions impact the shape of the estimated frontier.
 
@@
 Note that at least some of these quantitative properties are impacted by factors external to the syntax of a language, e.g., the unigram entropy may be impacted by the genre of available texts.
-This result suggests that deriving a language-independent bound on efficiency, as is possible in studies of semantic typology where there is a language-invariant parameterization of the possible meanings (e.g., \cite{kemp2012kinship,xu2014numeral,zaslavsky2018efficient}), is not possible in general for syntactic typology.
+This result again suggests that deriving a language-independent bound on efficiency, as is possible in studies of semantic typology where there is a language-invariant parameterization of the possible meanings (e.g., \cite{kemp2012kinship,xu2014numeral,zaslavsky2018efficient}), is not possible in general for syntactic typology.
 

### 2019-10-17 results/plane/visualizePlane_acrossLanguages.R

@@ -1,0 +1,196 @@
+# per language
+
+
+library(lme4)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(dplyr)
+dataS = read.csv("../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS2 = read.csv("../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS3 = read.csv("../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS4 = read.csv("../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS5 = read.csv("../../grammars/plane/plane-fixed-random3.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS6 = read.csv("../../grammars/plane/plane-fixed-random4.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS7 = read.csv("../../grammars/plane/plane-fixed-random5.tsv", sep="\t") %>% mutate(Model = as.character(Model)) %>% mutate(FullSurp = NULL)
+dataS = rbind(dataS, dataS2, dataS3, dataS4, dataS5, dataS6, dataS7)
+dataP = read.csv("../../grammars/plane/plane-parse-unified.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS = dataS %>% group_by(Language, Type, Model) %>% summarise(Surprisal = mean(Surp, na.rm=TRUE))
+dataP = dataP %>% group_by(Language, Type, Model) %>% summarise(Pars = mean(Pars, na.rm=TRUE))
+dataS = as.data.frame(dataS)
+dataP = as.data.frame(dataP)
+library(lme4)
+summary(lmer(Surprisal ~ Type + (1|Language), data=dataS %>% filter(grepl("langm", Type))))
+summary(lmer(Surprisal ~ Type + (1|Language), data=dataS %>% filter(grepl("RL", Type) | grepl("ground", Type))))
+dataS = dataS %>% mutate(Type = as.character(Type))
+dataP = dataP %>% mutate(Type = as.character(Type))
[…]

### 2019-10-24 results/plane/analyze_pareto_optimality/paretoByFamily.R
cleaning up

@@ -10,7 +10,7 @@ library(ggplot2)
 
 lambda = 0.9
 
-data = data %>% mutate(Better = (Pars + lambda*Surp) < (ParsGround + lambda*SurpGround))
+data = data %>% mutate(Better = (Pars + lambda*Surprisal) < (ParsGround + lambda*SurprisalGround))
 
 library(brms)
 
@@ -22,28 +22,28 @@ model = brm(Better ~ (1|Family) + (1|Language), family="bernoulli", data=data)
 samples = posterior_samples(model, "b_Intercept")[,]
 posteriorOpposite = ecdf(samples)(0.0)
 
-for(family in unique(languages$Family)) {
-
-   samples_ = (samples + posterior_samples(model)[[paste("r_Family[",sub(" ", ".", family),",Intercept]", sep="")]])
-
-   plot = ggplot(data.frame(BetterThan = 1/(1+exp(samples_))), aes(x=BetterThan, y=..scaled..)) + geom_density(fill="blue")
-   plot = plot + xlim(0,1)
-   plot = plot + theme_classic()
-   plot = plot + theme_void()
-   plot = plot + theme(legend.position="none")
-   plot = plot + geom_segment(aes(x=0.5, xend=0.5, y=0, yend=1), linetype=2)
-#   plot = plot + theme(axis.line.x = element_line(colour = "black"))
-   ggsave(paste("figures/quantile_posterior_", lambda, "_", family, ".pdf", sep=""), plot=plot, height=1, width=2)
-
-
-
-}
+#for(family in unique(languages$Family)) {
+#
+#   samples_ = (samples + posterior_samples(model)[[paste("r_Family[",sub(" ", ".", family),",Intercept]", sep="")]])
+#
+#   plot = ggplot(data.frame(BetterThan = 1/(1+exp(samples_))), aes(x=BetterThan, y=..scaled..)) + geom_density(fill="blue")
+#   plot = plot + xlim(0,1)
+#   plot = plot + theme_classic()
+#   plot = plot + theme_void()
+#   plot = plot + theme(legend.position="none")
+#   plot = plot + geom_segment(aes(x=0.5, xend=0.5, y=0, yend=1), linetype=2)
+##   plot = plot + theme(axis.line.x = element_line(colour = "black"))
+#   ggsave(paste("figures/quantile_posterior_", lambda, "_", family, ".pdf", sep=""), plot=plot, height=1, width=2)
+#
+#
+#
+#}
 
[…]

### 2019-10-24 writeup/pnas-word-order-si.tex
cleaning up
updates to SI

@@
 
-The variant of the Information Bottleneck that we use has been explored in the machine learning literature by \citet{strouse2017deterministic} and dubbed the ``Deterministic Information Bottleneck'' because, in the setting studied by \citet{strouse2017efficient}, it results in codes that are a deterministic function of the information to be expressed.
+The variant of the Information Bottleneck that we use has been explored in the machine learning literature by \citet{strouse2017deterministic} and dubbed the ``Deterministic Information Bottleneck'' because, in the setting studied by \citet{strouse2017deterministic}, it results in codes that are a deterministic function of the information to be expressed.
 We use this version of the Information Bottleneck because (1) it has been proposed in previous literature as a generic formalization of efficiency \citep{ferreri2003least}, and (2) it is not clear what would count as the three variables $Y$, $X$, and $\hat{X}$ in our setting.
@@
 We now show that this model instantiates the basic objective~(\ref{eq:eff-basic}). 
-If the dataset $d$ consists of observed pairs $(t,f)$ of meanings $t$ and forms $f$, and the language $h$ defines a set of possible pairs $(t,f)$, then the log-likelihood as defined by their model can be written as follows (up to constants)\footnote{We assume for simplicity that the error probability $\epsilon$ in the model is equal to $0$.}
-\begin{equation}
-\log P(d|h) = \sum_{(t,f) \in d} \log P(f|h,t) = \sum_{(t,f) \in d} \log P(f|h,t) \propto \sum_{(t,f) \in d} \log \frac{1}{|\{t' : (t', f) \in h\}|} = \sum_{(t,f) \in d} \log P(t|f)
-\end{equation}
+If the dataset $d$ consists of observed pairs $(t,f)$ of meanings $t$ and forms $f$, and the language $h$ defines a set of possible pairs $(t,f)$, then the log-likelihood as defined by their model can be written as follows (up to constants):\footnote{We assume for simplicity that the error probability $\epsilon$ in the model is equal to $0$.}
+\begin{align}
+  \nonumber
+  \log P(d|h) &= \sum_{(t,f) \in d} \log P(f|h,t) \\
+  \nonumber
+  & = \sum_{(t,f) \in d} \log P(f|h,t) \\
+  \nonumber
+  & \propto \sum_{(t,f) \in d} \log \frac{1}{|\{t' : (t', f) \in h\}|} \\
+  \nonumber
+  & = \sum_{(t,f) \in d} \log P(t|f),
+\end{align}
 where $P(t|f)$ is the probability that the observed form $f$ referred to meaning $t$, as the model assumes uniform meaning distributions and uniform choice of appropriate forms.
@@
 \begin{equation}
-- \operatorname{H}[t|f] = \operatorname{I}[t, f] - \operatorname{H}[t]
+- \operatorname{H}[t|f] = \operatorname{I}[t, f] - \operatorname{H}[t],
 \end{equation}
 where the first term is the mutual information between forms and meanings, as in our efficiency metric~(\ref{eq:eff-general}-\ref{eq:efficiency-derived}).
-The second term, the entropy of meanings, is a constant independent of the form-meaning mapping.
-The overall log probablity assigned by the Bayesian learner thus comes out to (up to constants)
+The second term, the entropy of meanings, is a constant independent of the form--meaning mapping.
+The overall log probability assigned by the Bayesian learner thus comes out to (up to constants)
 \begin{equation}
-\log P(h|d) = \operatorname{I}[t, f] + \lambda \log P(h)
+\log P(h|d) = \operatorname{I}[t, f] + \lambda \log P(h),
 \end{equation}
@@
 \paragraph{Relation to Formalizations of Pragmatics}
-In addition to these models, which model the efficiency and evolution of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context.
+In addition to these models, which  the concernefficiency and evolution of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context.
 Our work is most closely related to the Rational Speech Acts model of pragmatic reasoning \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
@@
 In cases where the grammar optimized for predictability (similarly parseability) has lower predictability (and parseability) than the grammar optimized for efficiency, we can replace its predictability value by that of the grammar optimized for efficiency: This is guaranteed to result in a point that is still Pareto-dominated by the grammar optimized for efficiency, and provides a tighter bound on the true curve.
-Th resulting frontier is guaranteed to provide a lower bound on the true Pareto frontier, but is nonetheless approximate: The actual curve may not be piecewise linear, and it may also extend beyond the estimated curve, as the grammar optimization method is approximate.
+The resulting frontier is guaranteed to provide a lower bound on the true Pareto frontier, but is nonetheless approximate: the actual curve may not be piecewise linear, and it may also extend beyond the estimated curve, as the grammar optimization method is approximate.
 
@@
 We further confirm this in Figure~\ref{fig:lambda-halfplane-09}:
-We plot the rean and optimized grammars together with a Kernel density estimate of the distribution of baseline grammars.
+We plot the real and optimized grammars together with a kernel density estimate of the distribution of baseline grammars.
 We add lines connecting those points that have the same efficiency $R_\lambda$ as the real grammar, at very low ($\lambda = 0.0$, dotted line) and very high ($\lambda = 0.9$, dashed line) values of $\lambda$.
@@
 \includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/pareto-plane-perLanguage-arrows-smoothed-halfspace-untransformed.pdf}
-	\caption[]{Per-Language results as in Figure~\ref{fig:pareto-per-lang-raw}, representing the distribution of baseline grammars with a Kernel density estimate.
+	\caption[]{Per-language results as in Figure~\ref{fig:pareto-per-lang-raw}, representing the distribution of baseline grammars with a Kernel density estimate.
 We add lines connecting those points that have the same efficiency as the real grammar at $\lambda = 0.0$ (dotted) and $\lambda = 0.9$ (dashed).
@@
 
-We conducted a series of linear regressions predicting (1) the predictability and parseability of the best grammar optimzied for efficiency, (2) the parseability and predictability difference between this end and the end optimized for predictability, (3) the difference between this end and the end optimized for parseability.
+We conducted a series of linear regressions predicting (1) the predictability and parseability of the best grammar optimized for efficiency, (2) the parseability and predictability difference between this end and the end optimized for predictability, (3) the difference between this end and the end optimized for parseability.
 For more meaningful comparison, we analyzed values normalized for sentence length as reported in Figure~\ref{fig:pareto-per-lang-raw}.
@@
 We also observe an effect of unigram entropy (4), showing that datasets with more diverse vocabulary reduce both predictability and parseability.\footnote{For predictability, a similar result about vocabulary size and estimated surprisal across many languages is reported by \cite{mielke-et-al-2019}.}
-Finally, larger amounts of training data (5) lead to higher estimated predictability and parseability -- this is expected, as more training data enables better statistical estimation of the distribution of sentences and syntactic structures.
+Finally, larger amounts of training data (5) lead to higher estimated predictability and parseability---this is expected, as more training data enables better statistical estimation of the distribution of sentences and syntactic structures.
 More training data also increases the difference between the efficiency-optimal and the predictability-optimal ends of the estimated curve, suggesting that more training data enables more precise estimation of the different extremes of the curve.
@@
 Note that at least some of these quantitative properties are impacted by factors external to the syntax of a language, e.g., the unigram entropy may be impacted by the genre of available texts.
-This result again suggests that deriving a language-independent bound on efficiency, as is possible in studies of semantic typology where there is a language-invariant parameterization of the possible meanings (e.g., \cite{kemp2012kinship,xu2014numeral,zaslavsky2018efficient}), is not possible in general for syntactic typology.
+This result again suggests that it may not be possible to derive a language-independent bound on syntactic efficiency, in contrast with studies of semantic typology where there is a language-invariant parameterization of the possible meanings (e.g., \cite{kemp2012kinship,xu2014numeral,zaslavsky2018efficient}).
 
@@
 
-We followed \citet{dryer1992greenbergian} in treating Greenberg's correlations as pairwise correlations with verb-object order.
-While Greenberg's original study \cite{greenberg1963universals} also formalized most of these as correlations with verb-object order, a few were formalized as correlations between other relations that are only indirectly related to verb-object order (e.g., Universal 22 linking the position of the standard of comparison to the order of adpositions).
-
-\citet{justeson1990explanation} conducted a log-linear analysis on typological judgments of 147 languages, constructing an undirected graphical model modeling correlations among any pair of six syntactic relations (verb-object, adposition-noun, noun-genitive, noun-relative clause, noun-adjective, verb-subject).
-Results from their analysis suggested that, while some relations are directly correlated with the verb-object order, whereas other relations are only indirectly correlated with it.
-In particular, in their analysis, the noun-genitive relation (corresponding to Correlation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}} here) was not directly correlated with the verb-object correlation; instead, the typologically observed correlation was explained through correlations between the noun-genitive relation and other relations (such as the adposition-noun relation) that directly correlate with the verb-object relation.
-Note that this does not contradict the statement that verb-object and noun-genitive relations correlate; it shows that the observed correlation can be explained through a chain of other correlations.
+We followed \citet{dryer1992greenbergian} in treating Greenberg's correlations as pairwise correlations with verb--object order.
+While Greenberg's original study \cite{greenberg1963universals} also formalized most of these as correlations with verb--object order, a few were formalized as correlations between other relations that are only indirectly related to verb-object order (e.g., Universal 22 linking the position of the standard of comparison to the order of adpositions).
+
+\citet{justeson1990explanation} conducted a log-linear analysis on typological judgments of 147 languages, constructing an undirected graphical model modeling correlations among any pair of six syntactic relations (verb--object, adposition--noun, noun--genitive, noun--relative clause, noun--adjective, verb--subject).
+Results from their analysis suggested that some relations are directly correlated with the verb--object order, whereas other relations are only indirectly correlated with it.
+In particular, in their analysis, the noun--genitive relation (corresponding to Correlation \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}} here) was not directly correlated with the verb--object correlation; instead, the typologically observed correlation was explained through correlations between the noun--genitive relation and other relations (such as the adposition--noun relation) that directly correlate with the verb--object relation.
+Note that this does not contradict the statement that verb--object and noun--genitive relations correlate; it shows that the observed correlation can be explained through a chain of other correlations.
 
 Since the set of syntactic relations examined here is different from that examined by \citet{justeson1990explanation}, we cannot directly compare the predictions of efficiency optimization with their results.
-Nonetheless, we can show that efficiency optimization is compatible with a picture of Greenberg's correlation as a network of pairwise correlations among different syntactic relations, and in particular the result that the correlation between the verb-object and noun-genitive relations is mediated through other correlations.
+Nonetheless, we can show that efficiency optimization is compatible with a picture of Greenberg's correlation as a network of pairwise correlations among different syntactic relations, and in particular the result that the correlation between the verb--object and noun--genitive relations is mediated through other correlations.
 
 First, we directly test the optimized grammars for two additional correlations found by \citet{justeson1990explanation}:
-For the relations examined here, beyond correlations with verb-object order, they found additional correlations between (1) the noun-genitive and adposition-noun dependencies, and (2) between the noun-relative clause and adposition-noun dependencies, \emph{beyond} the correlation mediated through the individual correlations with the verb-object dependency.
+For the relations examined here, beyond correlations with verb--object order, they found additional correlations between (1) the noun--genitive and adposition--noun dependencies, and (2) between the noun--relative clause and adposition--noun dependencies, \emph{beyond} the correlation mediated through the individual correlations with the verb--object dependency.
 %With their model, they found no evidence for a direct correlation V-O, G-N beyond that mediated through other correlations.
@@
 Second, we directly applied the log-linear analysis described by \citet{justeson1990explanation} to optimized grammars.
-We represent each grammar via the directions $v_1, \dots v_9$ of the nine relations indicated in Table 1 of the main paper (verb-object, and \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}), we coded these as $-0.5$ for Japanese-like order, and $+0.5$ for Arabic-like order.
+We represent each grammar via the directions $v_1, \dots v_9$ of the nine relations indicated in Table 1 of the main paper (verb--object, and \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}-\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}), we coded these as $-0.5$ for Japanese-like order, and $+0.5$ for Arabic-like order.
 This analysis models the relative frequency $p_{(v_1, \dots, v_9)}$ of a particular configuration of such a configuration $(v_1, \dots, v_9)$ by a log-linear model:
@@
 where $C$ is some set of (unordered) pairs of relations $\in \{1, \dots, 9\}$, modeling those pairs of relations that directly correlate with each other, and where $u_0, i_i, i_{i,j}$ are real-valued parameters.
-For instance, if all relations directly correlate with the verb-object order, and not with any other relation, $C$ would contain all the unordered pairs containing the verb-object relation.
+For instance, if all relations directly correlate with the verb--object order, and not with any other relation, $C$ would contain all the unordered pairs containing the verb--object relation.
 
@@
 This resulting model is shown in Figure~\ref{fig:loglinear}; following \cite{justeson1990explanation}, we show those links between nodes that are included in this selected model.
-In agreement with the results of \cite{justeson1990explanation}, a network is identified in which all relations are connected at least indirectly, but several relations are not directly connected to the verb-object relation:
-In particular, in accordance with the typological data analysed by \cite{justeson1990explanation}, the observed correlation between the verb-object and noun-genitive relation is entirely mediated through correlations with other relations (adposition-noun and verb-adpositional phrase) that directly correlate with the verb-object relation.
-A difference is that, in our analysis and unlike the analysis by \cite{justeson1990explanation}, the noun-relative clause dependency is not directly linked to the verb-object relation; this might be because our analysis takes a different set of relations into account compared to \cite{justeson1990explanation}.
+In agreement with the results of \cite{justeson1990explanation}, a network is identified in which all relations are connected at least indirectly, but several relations are not directly connected to the verb--object relation:
+In particular, in accordance with the typological data analysed by \cite{justeson1990explanation}, the observed correlation between the verb--object and noun--genitive relation is entirely mediated through correlations with other relations (adposition--noun and verb--adpositional phrase) that directly correlate with the verb--object relation.
+A difference is that, in our analysis and unlike the analysis by \cite{justeson1990explanation}, the noun--relative clause dependency is not directly linked to the verb--object relation; this might be because our analysis takes a different set of relations into account compared to \cite{justeson1990explanation}.
 
@@
 }
-	\caption{Detailed results for the two correlations found by \citet{justeson1990explanation} that do not involve the verb-object dependency, for grammars optimized for efficiency.
+	\caption{Detailed results for the two correlations found by \citet{justeson1990explanation} that do not involve the verb--object dependency, for grammars optimized for efficiency.
 Both correlations are strongly supported by optimized grammars, holding in about 90\% of optimized grammars.
@@
 	\caption{Network of pairwise correlations among the nine syntactic relations examined in this study, estimated from grammars optimized for efficiency, identified using a log-linear model following \citet{justeson1990explanation}.
-The verb-object relation is at the center of the network.
+The verb--object relation is at the center of the network.
 Relations between verbs and their dependents are colored in blue; relations between nouns and their dependents are colored in red; other relations are colored in green.}\label{fig:loglinear}
@@
 For instance, it does not represent cases where ordering varies between main and embedded clauses, as it does not condition ordering decisions on the larger context.
-It also does not nonprojective orderings, which -- while generally rare -- do occur in many languages (CITE).
-More complex and powerful ordering grammar models have been proposed \cite{futrell2015experiments, wang2016galactic}; however, they have similar limitations, and for our purposes, the model adopted here has the advantage of being simple and interpretable.
+It also does not nonprojective orderings, which---while generally rare---do occur in many languages.
+More complex and powerful ordering grammar models have been proposed \cite{futrell2015experiments,wang2016galactic}; however, they have similar limitations, and for our purposes, the model adopted here has the advantage of being simple and interpretable.
 
@@
 Among the eight Greenbergian correlations that we were able to test, five are annotated in WALS: adpositions, complementizers, relative clauses, genitives, and oblique PPs.
-In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb-object relation.
+In Table~\ref{tab:grammars-wals}, we compare our grammars with WALS on these five relations, and the verb--object relation.
 WALS has data for 74\% of the entries\footnote{Serbian and Croatian are listed as a single language Serbian-Croatian in WALS.
@@
 \begin{equation}\label{eq:efficiency}
-\argmax_\theta\	R_{\textit{Eff}}^{\theta} = \argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi} 
+\argmax_\theta\	R_{\textit{Eff}}^{\theta} = \argmax_\theta\	\max_{\phi, \psi} R_{\textit{Eff}}^{\theta, \phi, \psi}.
 %= \argmax_\theta\	\max_{\phi, \psi} \E_t \E_{u \sim p_\theta(u|t)} \left[R_{Pars}^{\phi}(u,t) + \lambda R_{Pred}^{\psi}(u)\right]
@@
 We enforced balanced distribution of object--verb and verb--object ordering among optimized languages by fixing $a_\tau$ for the \textit{obj} dependency to be 0.0 in four of these languages, and 1.0 in the other four.
-This maximizes statistical precision in detecting and quantifying correlations between the verb-object relation and other relations.
+This maximizes statistical precision in detecting and quantifying correlations between the verb--object relation and other relations.
 
@@
 
-In this section, we address the idea that the graph-based parser might have a built-in bias toward certain kinds of orderings, and the question whether this might be responsible for our findings.
+In this section, we provide evidence against the idea that the graph-based parser might have a built-in bias toward certain kinds of orderings.%, and the question whether this might be responsible for our findings.
 In particular, we address the idea that the graph-based parser might have a bias toward parses involving short dependencies, which we call a \key{locality bias}. 
-We address this by changing the order in which the parser sees words, in such a way that the distance between words in the input is not indicative of syntactic relations.
+We address this by changing the order in which the parser sees words, such that word adjacency in the input to the parser does not correspond to linear adjacency in the true utterance.
+If the parser has a locality bias, then this bias will be disrupted when it sees words in these distorted orders.
+We consider a number of possible distorted orders:
 
@@
 This is the idea that word order minimizes the average distance between syntactically related words.
-It is known that human languages reduce dependency length compared to random baselines~\cite{liu2010dependency,futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+It is known that human languages reduce dependency length compared to random baselines~\cite{liu2008dependency,futrell2015largescale,liu2017dependency,temperley2018minimizing}.
 Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,hawkins2004efficiency,futrell2017memory, futrell2017generalizing}.
@@
 
-Dependency length is typically quantified as the average distance between all pairs of syntactically related words, measured by the number of intervening words~\cite{futrell2015largescale}.
+Dependency length is typically quantified as the average distance between all pairs of syntactically related words, measured by the number of intervening words~\cite{liu2008dependency,futrell2015largescale}.
 Dependency length quantified in this manner is a heuristic measure of complexity:
@@
 
-\section{Simulation for Parseability and Correlating Orders}
+\section{Efficiency and correlating orders in toy grammars}
 
@@
 When we optimize grammars for efficiency, we find that the optimized grammars exhibit dependency length minimization and the Greenbergian word order correlations.
-To some extent, this result is surprising, because previous functional explanations for DLM (and the Greenbergian correlations, which are predicted by DLM) have been based on the idea of limitations in working memory, and yet our models do not instantiate any explicit working memory pressures.
-Nor is it the case that our parsers have implicit locality biases, as demonstrated by the experiments in Section~\ref{sec:distorted} above.
+To some extent, this result is surprising, because previous functional explanations for DLM (and the Greenbergian correlations, which have been argued to arise from DLM) have been based on the idea of limitations in working memory, and yet our models do not instantiate any explicit working memory pressures; see also Section~\ref{sec:distorted} above for evidence against the idea that a locality bias arises from our parsers.
 Our results therefore suggest that DLM and word order correlations might arise purely because they enable tree structures to be better recovered from trees, and/or they make sequences more predictable.
@@
 
-Our toy grammar consists of the following head-outward generative model.
+Our toy grammar consists of the following head-outward generative model \citep{eisner1996three}.
 Verbs generate verb dependents (\emph{xcomp}) and noun dependents (\emph{obj}), independently.
@@
 	\caption{Difference in parseability $I[\mathcal{T}, \mathcal{U}]$ (normalized by sentence length) between the correlating and anti-correlating grammars, for $p_\text{acl} = 0.0$ (left) and $p_\text{acl} = 0.3$ (right).
-Positive values indicate greater values of $I[\mathcal{T}, \mathcal{U}]$ for correlating grammars.}\label{fig:toy-parseability}
+Positive values indicate greater values of $I[\mathcal{T}, \mathcal{U}]$ for correlating grammars, i.e.
+cases where the grammars that exhibit natural-language-like correlations are more parseable than grammars that do not.}\label{fig:toy-parseability}
 \end{figure}
@@
 
-This simulation shows that correlations can in principle improve parseability in the controlled setting of such a model, without appeal to memory limitations; we leave a full graph-theoretical understanding to future work.
+This simulation shows that the Greenbergian word order correlations can in principle improve parseability in the controlled setting of such a model, without any appeal to memory limitations; we leave a full graph-theoretical understanding of this phenomenon to future work.
 

### 2019-10-25 results/plane/pureUD/analyze_pareto_optimality/plotOptimalityByLambda.R
cleaning

@@ -1,0 +1,41 @@
+data = read.csv("optimality-by-lambda.tsv", sep="\t")
+
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+
+plot = ggplot(data, aes(x=Lambda, y=Quantile)) + geom_line() + facet_wrap(~Language)
+
+cBoundUpper = c()
+cBoundLower = c()
+pValue = c()
+for(i in (1:nrow(data))) {
+   quantile = data$Quantile[[i]]
+   samples = data$Samples[[i]]
+   test = binom.test(round(quantile*samples), round(samples), alternative="two.sided", conf.level=1-0.05)
+   cBoundUpper = c(cBoundUpper, test$conf.int[[2]])
+   cBoundLower = c(cBoundLower, test$conf.int[[1]])
+   pValue = c(pValue, test$p.value*3)
+}
+data$UpperBound = cBoundUpper
+data$LowerBound = cBoundLower
+data$pValue = pValue
+
+
+plot = ggplot(data, aes(x=Lambda, y=Quantile)) + geom_line() + geom_line(aes(x=Lambda,y=LowerBound)) + facet_wrap(~Language)
+
+
+
+corpusSize = read.csv("../../../corpus-size/corpus-sizes.tsv", sep="\t")
+languagesOrdered = corpusSize$language[order(-corpusSize$sents_train)]
+
+data$Language = factor(data$Language, levels=languagesOrdered)
+
+
+
+plot = ggplot(data, aes(x=Lambda, y=Quantile)) + geom_line() + geom_line(aes(x=Lambda,y=UpperBound), linetype="dotted") + geom_line(aes(x=Lambda,y=LowerBound), linetype="dotted") + facet_wrap(~Language)
+
+
+ggsave(plot, file="figures/quantileByLambda.pdf")
+
+

### 2019-10-25 writeup/pnas-word-order-si.tex
cleaning

@@
 &  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
-\input{../results/plane/unlexicalized/per-language-statistics-unlexicalized-ground.tex}
+\input{../results/plane/unlexicalized/per-language-statistics-lexicalized.tex}
 \end{tabular}
@@
 	\centering
-	\includegraphics[width=\textwidth]{../results/plane/unlexicalized/analyze_pareto_optimality/figures/quantileByLambda-mle.pdf}
+	\includegraphics[width=\textwidth]{../results/plane/unlexicalized/analyze_pareto_optimality/figures/quantileByLambda.pdf}
 	\caption{Study 1, replication with lexicalized models: Optimality of real grammars for efficiency, compared to baselines, across values of $\lambda$: The $x$-axis shows $\lambda \in [0,1)$, the $y$-axis shows the fraction of baselines that have lower efficiency than the real grammar at this value of $\lambda$, with 95\% confidence bands obtained from a two-sided binomial test.
@@
 	\centering
-	\includegraphics[width=\textwidth]{../results/plane/pureUD/analyze_pareto_optimality/figures/quantileByLambda-mle.pdf}
+	\includegraphics[width=\textwidth]{../results/plane/pureUD/analyze_pareto_optimality/figures/quantileByLambda.pdf}
 	\caption{Study 1, replication with the original UD format: Optimality of real grammars for efficiency, compared to baselines, across values of $\lambda$: The $x$-axis shows $\lambda \in [0,1)$, the $y$-axis shows the fraction of baselines that have lower efficiency than the real grammar at this value of $\lambda$, with 95\% confidence bands obtained from a two-sided binomial test.

### 2019-10-25 results/plane/unlexicalized/analyze_pareto_optimality/plotOptimalityByLambda.R

@@ -6,6 +6,36 @@ library(ggplot2)
 
 plot = ggplot(data, aes(x=Lambda, y=Quantile)) + geom_line() + facet_wrap(~Language)
 
+cBoundUpper = c()
+cBoundLower = c()
+pValue = c()
+for(i in (1:nrow(data))) {
+   quantile = data$Quantile[[i]]
+   samples = data$Samples[[i]]
+   test = binom.test(round(quantile*samples), round(samples), alternative="two.sided", conf.level=1-0.05)
+   cBoundUpper = c(cBoundUpper, test$conf.int[[2]])
+   cBoundLower = c(cBoundLower, test$conf.int[[1]])
+   pValue = c(pValue, test$p.value*3)
+}
+data$UpperBound = cBoundUpper
+data$LowerBound = cBoundLower
+data$pValue = pValue
 
+
+plot = ggplot(data, aes(x=Lambda, y=Quantile)) + geom_line() + geom_line(aes(x=Lambda,y=LowerBound)) + facet_wrap(~Language)
+
+
+
+corpusSize = read.csv("../../../corpus-size/corpus-sizes.tsv", sep="\t")
+languagesOrdered = corpusSize$language[order(-corpusSize$sents_train)]
+
+data$Language = factor(data$Language, levels=languagesOrdered)
+
+
+
+plot = ggplot(data, aes(x=Lambda, y=Quantile)) + geom_line() + geom_line(aes(x=Lambda,y=UpperBound), linetype="dotted") + geom_line(aes(x=Lambda,y=LowerBound), linetype="dotted") + facet_wrap(~Language)
+
+
+ggsave(plot, file="figures/quantileByLambda.pdf")
 
 

### 2019-10-25 results/plane/unlexicalized/analyze_pareto_optimality/plotOptimalityByLambda_real.R

@@ -1,43 +1,0 @@
-data = read.csv("optimality-by-lambda-real.tsv", sep="\t")
-
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-
-plot = ggplot(data, aes(x=Lambda, y=Quantile)) + geom_line() + facet_wrap(~Language)
-
-cBoundUpper = c()
-cBoundLower = c()
-pValue = c()
-for(i in (1:nrow(data))) {
-   quantile = data$Quantile[[i]]
-   samples = data$Samples[[i]]
-   if(is.na(quantile)) {
-      cBoundUpper = c(cBoundUpper, 1)
-      cBoundLower = c(cBoundLower, 0)
-      pValue = c(pValue, 1)
-
-   } else{
-      test = binom.test(round(quantile*samples), round(samples), alternative="two.sided", conf.level=1-0.05)
-      cBoundUpper = c(cBoundUpper, test$conf.int[[2]])
-      cBoundLower = c(cBoundLower, test$conf.int[[1]])
-      pValue = c(pValue, test$p.value*3)
-   }
-}
-data$UpperBound = cBoundUpper
-data$LowerBound = cBoundLower
-data$pValue = pValue
-
-
-plot = ggplot(data, aes(x=Lambda, y=Quantile)) + geom_line() + geom_line(aes(x=Lambda,y=LowerBound)) + facet_wrap(~Language)
-
-
-#lambdas = c(0.0, 0.5, 0.98)
-#plot = ggplot(data %>% filter(Lambda %in% lambdas), aes(x=Lambda, y=Quantile)) + geom_line() + geom_line(aes(x=Lambda,y=LowerBound)) + facet_wrap(~Language)
-
-plot = ggplot(data, aes(x=Lambda, y=Quantile)) + geom_line() + geom_line(aes(x=Lambda,y=UpperBound), linetype="dotted") + geom_line(aes(x=Lambda,y=LowerBound), linetype="dotted") + facet_wrap(~Language)
-
-
-ggsave(plot, file="figures/quantileByLambda-real.pdf")
-
-

### 2019-10-25 results/plane/unlexicalized/plane-2D-best-visualization-large-3-unlexicalized-Real.R

@@ -1,220 +1,0 @@
-# per language
-
-
-library(lme4)
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-depl = read.csv("../../../grammars/dependency_length/total_summary_funchead_coarse.tsv", sep="\t")# %>% rename(Quality=AverageLength)
-library(tidyr)
-library(dplyr)
-dataS =  read.csv("../../../grammars/plane/controls/plane-fixed-withoutPOS.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataP =  read.csv("../../../grammars/plane/controls/plane-parse-lexicalized2.tsv", sep="\t") %>% mutate(Model = as.character(Model))
-dataS = dataS %>% group_by(Language, Type, Model) %>% summarise(Surp = mean(Surp, na.rm=TRUE))
-dataP = dataP %>% group_by(Language, Type, Model) %>% summarise(UAS = mean(UAS, na.rm=TRUE), Pars = mean(ParsU, na.rm=TRUE))
-dataS = as.data.frame(dataS)
-dataP = as.data.frame(dataP)
-library(lme4)
-dataS = dataS %>% mutate(Type = as.character(Type))
-dataP = dataP %>% mutate(Type = as.character(Type))
-dataS = dataS %>% mutate(Model = as.character(Model))
-dataP = dataP %>% mutate(Model = as.character(Model))
-dataS = dataS %>% mutate(Language = as.character(Language))
-dataP = dataP %>% mutate(Language = as.character(Language))
-data = merge(dataS, dataP, by=c("Language", "Model", "Type"), all.x=TRUE, all.y=TRUE)
-
-
-data = data %>% mutate(Type = ifelse(Type == "manual_output_funchead_RANDOM2", "manual_output_funchead_RANDOM", as.character(Type)))
-data = data %>% mutate(Type = ifelse(Type == "manual_output_funchead_RANDOM3", "manual_output_funchead_RANDOM", as.character(Type)))
-data = data %>% mutate(Type = ifelse(Type == "manual_output_funchead_RANDOM4", "manual_output_funchead_RANDOM", as.character(Type)))
-data = data %>% mutate(Type = ifelse(Type == "manual_output_funchead_RANDOM5", "manual_output_funchead_RANDOM", as.character(Type)))
-
-
-
-
-
[…]

### 2019-10-25 results/plane/unlexicalized/recordBaselinePositions_lexicalized.R

@@ -1,0 +1,38 @@
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+
+# Read predictability estimates
+dataS =  read.csv("../../../grammars/plane/controls/plane-fixed-withoutPOS.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataS = dataS %>% group_by(Language, Type, Model) %>% summarise(Surprisal = mean(Surp, na.rm=TRUE))
+dataS = as.data.frame(dataS)
+
+# Read parseability estimates
+dataP =  read.csv("../../../grammars/plane/controls/plane-parse-withoutPOS.tsv", sep="\t") %>% mutate(Model = as.character(Model))
+dataP = dataP %>% group_by(Language, Type, Model) %>% summarise(UAS = mean(UAS, na.rm=TRUE), Pars = mean(Pars, na.rm=TRUE))
+dataP = as.data.frame(dataP)
+
+# Ensure everything is treated as a string, not a factor, to ensure the frames can be merged
+dataS = dataS %>% mutate(Type = as.character(Type))
+dataP = dataP %>% mutate(Type = as.character(Type))
+dataS = dataS %>% mutate(Model = as.character(Model))
+dataP = dataP %>% mutate(Model = as.character(Model))
+dataS = dataS %>% mutate(Language = as.character(Language))
+dataP = dataP %>% mutate(Language = as.character(Language))
+data = merge(dataS, dataP, by=c("Language", "Model", "Type"), all.x=TRUE, all.y=TRUE)
+
+# Collapse all five groups of random grammars (10 random grammars per group per language)
+data = data %>% mutate(Type = ifelse(Type == "manual_output_funchead_RANDOM2", "manual_output_funchead_RANDOM", as.character(Type)))
+data = data %>% mutate(Type = ifelse(Type == "manual_output_funchead_RANDOM3", "manual_output_funchead_RANDOM", as.character(Type)))
+data = data %>% mutate(Type = ifelse(Type == "manual_output_funchead_RANDOM4", "manual_output_funchead_RANDOM", as.character(Type)))
+data = data %>% mutate(Type = ifelse(Type == "manual_output_funchead_RANDOM5", "manual_output_funchead_RANDOM", as.character(Type)))
+
+
+dataBaseline = data %>% filter(Type == "manual_output_funchead_RANDOM")
[…]

### 2019-10-25 writeup/pnas-word-order-si.tex
typo

@@
 
-Taken together, the order of the xcomp dependency tends to agree with that of most other \emph{xcomp} dependencies, with the sole exception of Gothic.
+Taken together, the order of `want' and its complement tends to agree with that of most other \emph{xcomp} dependencies, with the sole exception of Gothic.
 \end{enumerate}

### 2019-10-26 writeup/pnas-word-order-si.tex
code

@@
 An anonymous reviewer notes that the mapping between Dryer's relations and UD is not perfect, since some of the UD relations subsume other relations.
-Here we show that this is not a problem, since the ordering of the various relations subsumed under the UD label strongly agree typologically.
-
-TODO - say that, beyond just what Dryer reported, we see this distribution in the UD languages, and reproduce this
+Here we provide evidence that this is not impact our conclusions, since the ordering of the various relations subsumed under the UD label strongly agree typologically.
 
@@
 \paragraph{Relation to Formalizations of Pragmatics}
-In addition to these models, which  the concernefficiency and evolution of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context.
+In addition to these models, which concern the efficiency and evolution of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context.
 Our work is most closely related to the Rational Speech Acts model of pragmatic reasoning \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
@@
 In fact, comparing Figure~\ref{fig:lambda-quantile-noPOS} to Figure~\ref{fig:lambda-quantile} suggests that optimality of real grammars is \emph{more} pronounced when modeling predictability and parseability fully on the level of word forms.
+%This might be because such word-level models can account for morphology.
 
@@
 The Branching Direction Entropy measures the extent of freedom in choice between head-final and head-initial orderings, and it is a corpus-based quantitative measure of word order freedom \cite{futrell2015quantifying}.
-For a given syntactic relation, the \emph{branching direction entropy} measures the entropy of the Bernoulli variable that is $1$ whenever the head is ordered before the dependent, and $0$ if the dependent is ordered before the head.
+For a given syntactic relation, its branching direction entropy measures the entropy of the Bernoulli random variable that is $1$ whenever the head is ordered before the dependent, and $0$ if the dependent is ordered before the head.
 The branching direction entropy is $0$ if only one of the two orders can occur, and it is $\log 2$ if both orders are equally frequent.
@@
 For every one of the baseline grammars, we show both its deterministic and its nondeterministic version.
-Nondeterministic grammars are less efficient than deterministic grammars, in particular in language with greater degrees of word order freedom (Czech).
+Nondeterministic grammars are less efficient than deterministic grammars, in particular in languages with greater degrees of word order freedom (Czech).
 This shows that deterministic grammars constitute stronger baselines.
@@
 	\includegraphics[width=0.95\textwidth]{../results/plane/nondeterministic/pareto-plane-perLanguage-nondeterministic-mle.pdf}
-	\caption{Parseability and predictability for three languages, including both deterministic (green) and nondeterministic  (blue) versions of the 50 baseline grammars.}\label{fig:nondeterministic}
+	\caption{Parseability and predictability for three languages, including both deterministic (green, light) and nondeterministic  (blue, dark) versions of the 50 baseline grammars.}\label{fig:nondeterministic}
 \end{figure}
@@
 This example provides some intuitive idea of why the correlating grammar might lead to improved parseability:
-Note that the red token labeled `N' occupies the same structural position in both versions.
+Note that the red boldface token labeled `N' occupies the same structural position in both versions.
 In the anti-correlating version (right), when given only the token sequence, without the syntactic structure, this word could a priori be an \emph{obj} dependent of any of the three verbs occurring to its right.
@@
    \begin{deptext}[column sep=1em]
-	   V \& \textcolor{red}{N} \& V \& N \& V  \\
+	   V \& \textbf{\textcolor{red}{N}} \& V \& N \& V  \\
    \end{deptext}
@@
    \begin{deptext}[column sep=1em]
-	   \textcolor{red}{N} \& N \& V \&  V \&  V   \\
+	  \textbf{\textcolor{red}{N}} \& N \& V \&  V \&  V   \\
    \end{deptext}

### 2019-10-27 writeup/pnas-word-order-si.tex
code

@@
 \caption{Models estimating the log-odds of a random baseline grammar improving over a real grammar in efficiency ($\lambda=0.9$), surprisal, or parseability, with random effects for languages and language families.
-The strongly negative estimates of $\beta$ confirm that, across languages and language families, real grammars improve over most baselines.}\label{tab:study1-glmer}
+The strongly negative estimates of $\beta$ confirm that, across languages and language families, real grammars improve over most baselines in predictability, parseability, and overall efficiency.
+This model shows that the optimization observed in Study 1 cannot be attributed to family-specific effects.}\label{tab:study1-glmer}
 \end{table}
@@
 Additionally, we plot the efficiency values obtained when applying the Japanese grammar to English tree structures (purple square, left), and when applying the English grammar to Japanese tree structures (purple square, right).
-In both languages, the respective real grammars (crosses) are more efficient than grammars from the other language (squares).
+In both languages, the respective real grammars (crosses) are more efficient than grammars from the other language (squares), even though the grammar from the other language still is more efficient than the baselines.
 This suggests that the grammars of languages are specifically optimized for their individual tree structure distributions.}\label{fig:english-japanese-crossed}
@@
 The results are shown in Figure~\ref{fig:english-japanese-crossed}.
-In both languages, the respective real grammars (crosses) are more efficient than grammars from the other language (squares).
-This suggests that the grammars of languages are specifically optimized for their individual tree structure distributions.
+In both languages, the respective real grammars (crosses) are more efficient than grammars from the other language (squares), even though the grammar from other language still is more efficient than the baseline grammars.
+This suggests that the grammars of languages, beyond reflecting generic optimization for efficiency across tree structures, may also be specifically optimized for their individual tree structure distributions.
 Furthermore, it demonstrates that the tree structure distribution, and therefore the optimality of a given grammar, is language-specific.

### 2019-10-27 writeup/pnas-word-order-si.tex
code

@@
 \centering
-\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-WithN.pdf}
 	\caption[Predictability and Parseability]{Predictability and parseability of 51 languages, ordered by corpus size, measured by the number of sentences in the training partition, from largest (Czech) to smallest (Irish).
@@
 \centering
-\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-raw.pdf}
+\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage-raw-WithN.pdf}
 	\caption[Predictability and Parseability]{Raw numerical values estimated for Predictability (negative surprisal), and negative syntactic ambiguity $-H[T|U]$, before $z$-scoring.
@@
 \includegraphics[width=\textwidth]{../results/plane/analyze_pareto_optimality/pareto-plane-perLanguage-arrows-smoothed-halfspace-untransformed.pdf}
-	\caption[]{Per-language results as in Figure~\ref{fig:pareto-per-lang-raw}, representing the distribution of baseline grammars with a Kernel density estimate.
+	\caption[]{Per-language results as in Figure~\ref{fig:pareto-per-lang-raw}, representing the distribution of baseline grammars with a kernel density estimate.
 We add lines connecting those points that have the same efficiency as the real grammar at $\lambda = 0.0$ (dotted) and $\lambda = 0.9$ (dashed).

### 2019-10-27 results/plane/visualizePlane_perLanguage.R

@@ -191,5 +191,37 @@ ggsave(plot, file="pareto-plane-perLanguage.pdf", width=12, height=12)
 
 
 
+data = merge(data, corpusSize %>% rename(Language=language), by=c("Language"))
+subData = merge(subData, corpusSize %>% rename(Language=language), by=c("Language"))
+
+data$Language_ = paste(data$Language, "\nn = ", data$sents_train, sep="")
+subData$Language_ = paste(subData$Language, "\nn = ", subData$sents_train, sep="")
+corpusSize$Language_ = paste(corpusSize$language, "\nn = ", corpusSize$sents_train, sep="")
+
+
+languagesOrdered = corpusSize$Language[order(-corpusSize$sents_train)]
+
+data$Language_ = factor(data$Language_, levels=languagesOrdered)
+subData$Language_ = factor(subData$Language_, levels=languagesOrdered)
+
+
+
+plot = ggplot(data %>% filter(Type %in% c("manual_output_funchead_RANDOM")) %>% filter(Surprisal_z < 3), aes(x=-Pars_z, y=-Surprisal_z, color=Type, group=Type))
+plot = plot + geom_point()
+plot = plot + geom_path(data=subData, aes(x=-Pars_z, y=-Surprisal_z, group=1), size=1.5)
+plot = plot + geom_point(data=data %>% filter(Type %in% c("manual_output_funchead_ground_coarse_final")) %>% filter(Surprisal_z < 3), shape=4, size=1.5, stroke=2)
+plot = plot + facet_wrap(~Language_, scales="free")
+plot = plot + theme_bw()
+plot = plot + scale_x_continuous(name="Parseability") + scale_y_continuous(name="Predictability")
+plot = plot + theme(legend.title = element_blank())  
+plot = plot + guides(color=guide_legend(nrow=2,ncol=4,byrow=TRUE)) 
+plot = plot + theme(legend.title = element_blank(), legend.position="bottom")
+plot = plot + theme(axis.title.x = element_text(size=17))
+plot = plot + theme(axis.title.y = element_text(size=17))
+plot = plot + theme(legend.text = element_text(size=12))
+plot = plot + theme(legend.margin=margin(t = 0, unit='cm'))
+plot = plot + theme(legend.position = "none")
+ggsave(plot, file="pareto-plane-perLanguage-WithN.pdf", width=12, height=12)
+
 
 

### 2019-10-27 writeup/pnas-word-order-si.tex

@@
 
-\subsection{Quantifying Degree of Optimality}
+\subsection{Quantifying Degree of Optimality for Overall Efficiency}
 
@@
 In the main paper (Study 1), we showed that languages tend to be optimized for parseability and/or predictability.
-Efficiency is a combination of both components; in this section we address the question whether languages are generally optimized for efficiency, i.e., the multi-objective optimization problem of optimizing for parseability and predictability.
+Efficiency is a combination of both components; in this section we address the question whether languages are generally optimized for efficiency as a multi-objective optimization problem of optimizing for parseability and predictability.
 Recall the efficiency metric
@@
 For instance, the real grammar of Czech does not improve over all baselines in predictability (see Figure~\ref{fig:pareto-per-lang}), but it has higher overall efficiency than the vast majority of baselines in efficiency, for all values $\lambda \in [0,1)$.
+There are also languages for which the degree of optimality does strongly depend on $\lambda$; however, we note that estimated optimimality is stronger when estimating efficiency using lexicalized parsers that can take morphology into account (Figures~\ref{fig:pareto-plane-noPOS}-\ref{fig:lambda-quantile-noPOS}).
+
+This analysis is similar to that reported by \citet{zaslavsky2018efficient} in a study of color names; they found that observed color naming systems have higher efficiency than almost all baseline systems at \emph{a specific value of $\lambda$}.
+Here, we have shown that grammars tend to be more efficient than baselines across most values of $\lambda$.
 

### 2019-11-05 writeup/pnas-word-order-si.tex
code

@@
 Efficiency is a combination of both components; in this section we address the question whether languages are generally optimized for efficiency as a multi-objective optimization problem of optimizing for parseability and predictability.
+
 Recall the efficiency metric
@@
 with the tradeoff parameter $\lambda \in [0,1)$.
-For each possible value of $\lambda \in [0,1)$ trading off parseability and predictability, we quantify how many baseline grammars are less efficient than the real language.
+For each possible value of $\lambda \in [0,1)$ trading off parseability and predictability, we quantify what fraction of the baseline grammars are less efficient than the real language.
 
@@
 For instance, it does not represent cases where ordering varies between main and embedded clauses, as it does not condition ordering decisions on the larger context.
-It also does not nonprojective orderings, which---while generally rare---do occur in many languages.
+It also does not model nonprojective orderings, which---while generally rare---do occur in many languages.
 More complex and powerful ordering grammar models have been proposed \cite{futrell2015experiments,wang2016galactic}; however, they have similar limitations, and for our purposes, the model adopted here has the advantage of being simple and interpretable.
@@
 \begin{equation}\label{eq:rpars}
-	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p_\phi(t|u)}{p(t)} = \frac{1}{|\text{Heldout Data}|} \sum_{t,u \in \text{Heldout Data}} \log \frac{p_\phi(t|u)}{p(t)}
+	R_{Pars} := \operatorname{I}[\utterance,\tree] = \sum_{t,u} p(t,u) \log \frac{p_\phi(t|u)}{p(t)} \approx \frac{1}{|\text{Heldout Data}|} \sum_{t,u \in \text{Heldout Data}} \log \frac{p_\phi(t|u)}{p(t)}
 \end{equation}
@@
 
-In Study 1, we considered deterministic ordering grammars, even though real languages have some degree of word order freedom.
-Here, we show that deterministic baseline grammars provide conservative baselines: They have higher efficiency than baseline grammars with word order freedom comparable to the orders found in the original corpora, and thus provide conservative baselines for comparison with other grammars.
-
+In Study 1, we considered deterministic ordering grammars, and represented real languages using deterministic grammars extracted from observed orderings.
+This allowed us to ensure that we only compare baseline and real grammars that have exactly the same representational constraints, and utilize the same information encoded in the tree structures.
+
+In this section, we consider baselines that allow word order freedom to degrees comparable to that found in orders observed in the actual corpus data.
 In order to obtain baselines whose freedom is comparable to that of real languages, we constructed baselines that have the same Branching Direction Entropy \cite{futrell2015quantifying} as observed in the original corpora.
@@
 
-We compared these nondeterministic baseline grammars to their deterministic versions, for one language with relatively free order (Czech), and for two languages with relatively fixed order (English and Japanese).
+\paragraph{Comparing deterministic and nondeterministic grammars}
+Here, we compared nondeterministic baseline grammars to their deterministic versions, for one language with relatively free order (Czech), and for two languages with relatively fixed order (English and Japanese).
 Results are shown in Figure~\ref{fig:nondeterministic}.
@@
 Nondeterministic grammars are less efficient than deterministic grammars, in particular in languages with greater degrees of word order freedom (Czech).
-This shows that deterministic grammars constitute stronger baselines.
+%This shows that deterministic grammars constitute stronger baselines.
+This shows that deterministic baseline grammars provide conservative baselines:
+They have higher efficiency than baseline grammars with word order freedom comparable to the orders found in the original corpora, and thus provide conservative baselines for comparison with other deterministic grammars.
 
@@
 	\caption{Parseability and predictability for three languages, including both deterministic (green, light) and nondeterministic  (blue, dark) versions of the 50 baseline grammars.}\label{fig:nondeterministic}
+\end{figure}
+
+\paragraph{Comparing observed orders to baselines with matched degree of nondeterminism}
+Here, we compare the efficiency of the orders observed in the corpora with baselines whose degree of nondeterminism, quantified by branching direction entropy, is matched to that of the observed orders.
+We show results in Figures~\ref{fig:pareto-plane-nondet} and~\ref{fig:lambda-quantile-nondet}.
+Figure~\ref{fig:pareto-plane-nondet} shows that observed orders are mostly to the top and/or right of baselines with matched degree of nondetermminism.
+Figure~\ref{fig:lambda-quantile-nondet} shows that, with the exception of Telugu (a language with a small corpus), the observed orders have higher efficiency than most baselines at least for some values of $\lambda$.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../results/plane/nondeterministic/pareto-plane-perLanguage-nondeterministic-mle-allLanguages.pdf}
+	\caption[Predictability and Parseability]{Comparing observed orders (red crosses) with baselines (green) whose degree of nondeterminism is matched to the observed order.
+Compare Figure~\ref{fig:pareto-per-lang}.}\label{fig:pareto-plane-nondet}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{../results/plane/nondeterministic/analyze_pareto_optimality/figures/quantileByLambda.pdf}
+	\caption{Optimality of observed orders for efficiency, compared to nondeterministic baselines, across values of $\lambda$: The $x$-axis shows $\lambda \in [0,1)$, the $y$-axis shows the fraction of baselines that have lower efficiency than the observed orders at this value of $\lambda$, with 95\% confidence bands obtained from a two-sided binomial test.
+Compare Figure~\ref{fig:lambda-quantile}.}\label{fig:lambda-quantile-nondet}
 \end{figure}

### 2019-11-05 results/plane/nondeterministic/recordBaselinePositions.R
code

@@ -1,0 +1,48 @@
+# per language
+
+
+library(lme4)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+# Read predictability estimates
+dataS =  read.csv("../../../grammars/plane/controls/plane-fixed-nondeterministic.tsv", sep="\t") %>% mutate(Model = as.character(Model), Type=as.character(Type)) %>% mutate(Type = ifelse(grepl("RANDOM", Type), "manual_output_funchead_RANDOM_nondet", Type)) %>% mutate(Model = paste(Model, "nondet", sep="_"))
+dataS1 = read.csv("../../../grammars/plane/plane-fixed.tsv", sep="\t") %>% mutate(Model = as.character(Model), Type=as.character(Type)) %>% mutate(Type = ifelse(grepl("RANDOM", Type), "manual_output_funchead_RANDOM_det", Type)) %>% mutate(Model = paste(Model, "det", sep="_"))
+dataS2 = read.csv("../../../grammars/plane/plane-fixed-best.tsv", sep="\t") %>% mutate(Model = as.character(Model), Type=as.character(Type)) %>% mutate(Type = ifelse(grepl("RANDOM", Type), "manual_output_funchead_RANDOM_det", Type)) %>% mutate(Model = paste(Model, "det", sep="_"))
+dataS3 = read.csv("../../../grammars/plane/plane-fixed-best-large.tsv", sep="\t") %>% mutate(Model = as.character(Model), Type=as.character(Type)) %>% mutate(Type = ifelse(grepl("RANDOM", Type), "manual_output_funchead_RANDOM_det", Type)) %>% mutate(FullSurp=NULL) %>% mutate(Model = paste(Model, "det", sep="_"))
+dataS4 = read.csv("../../../grammars/plane/plane-fixed-random2.tsv", sep="\t") %>% mutate(Model = as.character(Model), Type=as.character(Type)) %>% mutate(Type = ifelse(grepl("RANDOM", Type), "manual_output_funchead_RANDOM_det", Type)) %>% mutate(FullSurp=NULL) %>% mutate(Model = paste(Model, "det", sep="_"))
[…]

### 2019-11-05 results/plane/visualizePlane_perLanguage.R

@@ -175,7 +175,7 @@ subData$Language = factor(subData$Language, levels=languagesOrdered)
 plot = ggplot(data %>% filter(Type %in% c("manual_output_funchead_RANDOM")) %>% filter(Surprisal_z < 3), aes(x=-Pars_z, y=-Surprisal_z, color=Type, group=Type))
 plot = plot + geom_point()
 plot = plot + geom_path(data=subData, aes(x=-Pars_z, y=-Surprisal_z, group=1), size=1.5)
-plot = plot + geom_point(data=data %>% filter(Type %in% c("manual_output_funchead_ground_coarse_final")) %>% filter(Surprisal_z < 3), shape=4, size=1.5, stroke=2)
+plot = plot + geom_point(data=data %>% filter(Type == "manual_output_funchead_ground_coarse_final") %>% filter(Surprisal_z < 3), shape=4, size=1.5, stroke=2)
 plot = plot + facet_wrap(~Language, scales="free")
 plot = plot + theme_bw()
 plot = plot + scale_x_continuous(name="Parseability") + scale_y_continuous(name="Predictability")

### 2019-11-11 writeup/3-paper-names.tex
code

--- 2019-04-18 second-draft/3-paper-names.tex
@@
 \newcommand{\xmark}{\ding{55}}%
+
+%https://tex.stackexchange.com/questions/219676/how-to-change-spacing-before-and-after-paragraph-command
+%\usepackage{xpatch}
 
@@
 
-\title{Universals of word order result from optimization of grammars for efficient communication}
+\renewcommand{\baselinestretch}{0.99}
+\addtolength{\textfloatsep}{-0.18in}
+\setlength{\abovecaptionskip}{5pt plus 1pt minus 0pt} % Chosen fairly arbitrarily
+%\setlength{\belowcaptionskip}{-5pt}
+%https://tex.stackexchange.com/questions/23313/how-can-i-reduce-padding-after-figure
+
+\title{Universals of word order reflect optimization of grammars for efficient communication}
 
 % Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-%\author[a,1,2]{Michael Hahn}
-%\author[a]{Daniel Jurafsky}
-%\author[b]{Richard Futrell}
+\author[a,1]{Michael Hahn}
+\author[a]{Dan Jurafsky}
+\author[b]{Richard Futrell}
 
-%\affil[a]{Stanford University}
-%\affil[b]{University of California, Irvine}
+\affil[a]{Stanford University}
+\affil[b]{University of California, Irvine}
 
@@
 \significancestatement{
-What explains the universal properties of human languages?
-We present evidence that a major subset of these properties can be explained by viewing languages as codes for efficient communication among agents with highly generic cognitive constraints.
-In doing so, we provide the first full formalization and computational implementation of ideas which have been stated informally in the functional linguistics literature for decades.
-The success of this approach suggests a new way to conceptualize human language in quantitative and computational work, as an information-theoretic code dynamically shaped by communicative and cognitive pressures.
-Our results argue against the idea that the distinctive properties of human language result from essentially arbitrary genetic constraints.% (``universal grammar'').
-}
+Human languages share many grammatical properties.
+We show that some of these properties 
+can be explained by the need for languages to offer
+efficient communication between humans given our cognitive constraints.
+Grammars of languages seem to find a balance between two communicative pressures: to be simple enough to allow the speaker to easily produce sentences, but complex enough to be unambiguous to the hearer, and this balance explains well-known word order generalizations across our sample of 51 varied languages.
+Our results offer new quantitative and computational evidence that language structure is dynamically shaped by communicative and cognitive pressures.}
 
 % Please include corresponding author, author contribution and author declaration information
-%\authorcontributions{MH and RF designed research. MH implemented models and experiments. MH and RF wrote the paper. MH, DJ, and RF provided comments on the paper.}
+\authorcontributions{MH, DJ, and RF designed research. MH implemented models and experiments. MH, DJ, and RF wrote the paper.}
 \authordeclaration{The authors declare no conflict of interest.}
-%\equalauthors{\textsuperscript{1}A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable).}
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: mhahn2@stanford.edu}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: mhahn2@stanford.edu}
 
@@
 \begin{abstract}
-%	TODO rewrite
-The universal properties of human languages have been the subject of intense study across disciplines.
-	We report novel computational and corpus-based evidence for the hypothesis that a prominent subset of these universal properties---those related to word order---result from a process of optimization for efficient communication among humans.
-	We computationally formalize functional theories of language efficiencies using information theory and state-of-the-art machine learning techniques.
-Applying these methods to data from 51 languages, we show that evolution towards efficiency results in word order patterns that predict a large subset of the major word order correlations reported in the linguistic typological literature.
-%	TODO explain
-	%We develop a probabilistic, differentiable model of word order grammars: the means by which different languages convert underlying hierarchical structures into strings of words.
-	%We show how the parameters of a word order grammar can be optimized for efficiency of information transfer, quantified as a tradeoff between incremental predictability and mutual information with latent tree structures.
-%	Applying these grammars to tree structures found in dependency corpora from 51 languages, we show that optimizing the grammar parameters for efficiency results in word order patterns that predict a large subset of the major word order correlations reported in the linguistic typological literature, and subsumes the predictions of previous heuristic theories such as dependency length minimization.
+The universal properties of human languages have been the subject
+of intense study across the language sciences.
+We report novel computational and corpus evidence for
+the hypothesis that a prominent subset of these universal
+properties---those related to word order---result from a
+process of optimization for efficient communication among humans,
+trading off the need to reduce complexity with the need to reduce ambiguity.
+We formalize these two pressures with information-theoretic and neural network models
+of complexity and ambiguity, and simulate grammars with optimized
+word order parameters on large-scale data from 51 languages. 
+Evolution of grammars towards efficiency results in word order patterns that
+predict a large subset of the major word order correlations 
+across languages.
 \end{abstract}
@@
 \begin{document}
-% TODO mention the preregistration somewhere
 \maketitle
@@
 
-%Ray Jackendoff. 2002. Foundations of Language:
-%Brain, Meaning, Grammar, Evolution. Oxford
-%University Press, Oxford, UK
+\dropcap{U}nderstanding what is universal and what varies across human languages is a central goal of linguistics.
+Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, kirby2015compression, pinker1990natural, nowak1999evolution}.
+%nowak2001towards, nowak2002computational
+But formalizing how these pressures explain specific grammatical universals has proved difficult.
+Here we pair new computational models that measure the communicative efficiency of grammars with a new simulation framework for finding optimal grammars, and show that the most efficient grammars also exhibit a large class of language universals.
+% information-theoretic
 
-%William Croft and Alan Cruse. 2004. Cognitive
-%Linguistics. Cambridge University Press, Cam-
-%bridge, UK.
-% --> emphasize language use
-
-%Adele Goldberg. 2005. Constructions at work:
-%The nature of generalization in language. Ox-
-%ford University Press, Oxford, UK.
-% --> emphasize usage-based grammar
-% we are all born with the same basic conceptual apparatus, with the
-%same basic communicative demands, and we all live in the physical world with
-%forces of gravity, bodies, and night and day (cf. also LakoV 1987; Webelhuth and
-%Ackerman 1998). Certain semantic/pragmatic functions are so relevant and
-%useful that they can be found in language after language (Croft 2001).
-%Cross-linguistic generalizations that relate form and function are explained
-%by appeal to general cognitive constraints together with the functions of the
-%constructions involved. Constructionists turn to grammar-external explan-
-%ations such as universal functional pressures, iconic principles, and processing
-%and learning constraints to explain such empirically observable cross-linguis-
-%tic generalizations.
-%For example, the fact
-%that that most languages seem to have noun and verb categories may be
-%explained by the existence of corresponding basic semantic categories (Baker
-%2004). Hauser, Chomsky, and Fitch go so far as to suggest that the only innate
-%ability speciWc to language that may be required are the recursive interfaces to
-%phonology and semantics, and they allow that even these may turn out not to
-%be speciWc to language (Hauser, Chomsky, and Fitch 2002).
-
-% Haspelmath 1999 talks about types of explanations
-
-%Joan Bresnan, Ash Asudeh, Ida Toivonen, and
-%Stephen Wechsler. 2016. Lexical-Functional
-%Syntax, 2nd ed. Blackwell, Malden, MA.
-
-%Andrew Radford. 2006. Minimalist syntax re-
-%visited. http://www.public.asu.edu/
-%~gelderen/Radford2009.pdf.
-
-%Givon
-
-% TODO optimality theory and syntax citation?
-%Blutner R. (2000) Some aspects of optimality in natural language interpretation. Journal of Semantics 17: 189–216
-
-% ,lin2017critical
-\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics. % \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
-% behaghel1909beziehungen,
-% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
-%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-% TODO mention universals here
-Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic}.
-%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
-In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which a large class of language universals can be explained through the computational optimization of grammars for efficiency.
-% information-theoretic
-Specifically, we show that efficiency optimization explains  linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}. 
-% TODO change wording so it's clear that this is a very major thing
-
-% Things TODO
-% - make clear that efficiency is something both functional and generative linguists have been appealing to
-% - at some point address head direction idea
-%TODO at some point address head direction idea
-%    - Vennemann 1972, 1974ab, 1976
-%    -Hawkins 1980 limitations of head direction, also Hawkins 1982
-
-%foreshadow functional explanations?
-%
-%- these linguistic patterns reflect general principles of communication
-%
-%- big idea has been around for a long time
-%
-%- here for the first time direct computational test
-%
-%- we can predict broad patterns, and make language-specific predictions based on these general principles 
-%
-
-% TODO add a cogsci person to citation chain at the beginning
-
-%\dropcap{F}or decades, researchers in fields ranging from philology to cognitive science to statistical physics have been involved in documenting and trying to explain the universal syntactic and statistical properties of human language \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical}.
-An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
-
-%Berwick and Chomsky ‘gene rative process  is optimal  ’, ‘efficient computa  tion ’(71), ‘this newly  emerged  computationalsystem forthought  ...ispe rfe ct, in sofar asSMT is corr ect ’(80).
-‘op -tim al’or ‘effici ent ’or ‘per fect ’.
-
-%Chomsky Three Factors, 2003:
-% --> emphasizes 'computational efficiency'
-
-%\textbf{TODO maybe should be more general in the beginning, as in the first version. Also make sure the reader doesn't think 'it's just headedness, why bother'.}
-
-%We aim to explain the linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}. 
+The language universals we study are the well-known Greenberg universals of word order \cite{greenberg1963universals}. 
 Human languages vary in the order in which they express information.
 Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
-Both sentences contain a verb meaning `wrote', a noun expressing  the object `letter', and an adpositional phrase translating to `to a friend'.
-%In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
-Yet, the order of these words are entirely different in the two languages:
-%However, the languages differ in the order in which these words and phrases are arranged: 
-The verb stands at the beginning in Arabic, and at the end in Japanese.
-Arabic expresses `to' by a \emph{preposition} (\emph{preceding} the noun `friend'); Japanese uses a \emph{postposition} (\emph{following} it). % and a postposition in Japanese, following `friend'.
+Both sentences contain a verb meaning `wrote', a noun expressing  the object `letter', and a phrase translating to `to a friend'.
+Yet the order of these words are entirely different in the two languages:
+the verb stands at the beginning in Arabic, and at the end in Japanese.
+Arabic expresses `to' by a \emph{preposition} (\emph{preceding} the noun `friend'); Japanese uses a \emph{postposition} (\emph{following} it).
 
-This variation reflects a deep and stable regularity:
-While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
+Yet this variation reflects a deep and stable regularity:
+while languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
 this is strongly correlated with the occurrence of pre- or postpositions (Figure~\ref{fig:corr-table}, top):
-Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
+Languages ordering their objects the way Japanese does have postpositions; languages ordering them as Arabic does have prepositions.
 
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{figures/corr-full.png}
-\caption{ The concept of a word order correlation: Languages can order the object after (Arabic) or before (Japanese) the verb, and have prepositions (Arabic) or postpositions (Japanese).
-For each combination, we indicate how many languages satisfy it, as documented in the World Atlas of Language Structure~\cite{wals}.
+    \includegraphics[width=0.4\textwidth]{figure1.pdf}
+\caption{ One word order correlation: Languages can order the object after (Arabic) or before (Japanese) the verb, and have prepositions (Arabic) or postpositions (Japanese).
+For each combination, we indicate how many languages satisfy it, as documented in the World Atlas of Language Structures~\cite{wals}.
 Combinations on the diagonal are vastly more common than off-diagonal ones.
-% A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.
 	}\label{fig:arabic-japanese-simple}	\label{fig:corr-table}
@@
 
-This generalization falls in a group of language universals that were originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}:
+This generalization lies in a group of language universals originally documented by Greenberg \cite{greenberg1963universals}, known as \key{word order correlations}.
 These describe correlations between the relative positions of different types of expressions across languages.
@@
 %adpositional phrases (`to a friend') relative to the verb.
-Greenberg also found that the order between verb and object is correlated with other aspects of a language's word order (Table~\ref{table:corr-dryer}), such the order of verb and adpositional phrase (`wrote -- to friend' in Arabic vs `friend to -- wrote' in Japanese), and that of noun and genitive (`book -- of friend' in Arabic, `friend of -- book' in Japanese).
+Greenberg also found that the order of verb and object is correlated with other aspects of a language's word order (Table~\ref{table:corr-dryer}), such as the order of verb and adpositional phrase (`wrote -- to friend' in Arabic vs `friend to -- wrote' in Japanese), and that of noun and genitive (`book -- of friend' in Arabic, `friend of -- book' in Japanese).
+
+\begin{table}[ht]
+	\begin{center}
+\begin{tabular}{|c|ll|ll|}
+	\hline
+	&	\multicolumn{2}{c|}{\textbf{Arabic} (English, ...)}   &        \multicolumn{2}{c|}{\textbf{Japanese} (Turkish, ...)} \\ \hline\hline
+	& \multicolumn{2}{c|}{Correlates with...} & \multicolumn{2}{c|}{Correlates with...}  \\
+	&	Verb & Object     & Object & Verb    \\ \hline\hline
+	& kataba & ris{\= a}la 	&tegami-o & kaita \\
+	&	\emph{wrote} & \emph{letter} & 	\emph{letter} & \emph{wrote} \\ \hline 
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    {\d s}ad{\= i}q       &	tomodachi & ni \\ % G3, G4
+	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline 
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&k{\= a}na    &    {\d s}ad{\= i}q         &	tomodachi & datta \\ % not in Greenberg
+	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak- & -udesho \\ % G16, G13 TODO this should only be inflected auxiliaries, replace Arabic
+	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& {\d s}ad{\= i}q  &    John    & 	John no & tomodachi \\ % G2, G23
+	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uh{\= a}       & 	anata-ga yonda & hon \\ % G24
+	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tu{\d s}il        & 	toochaku suru & koto \\ % 
+	&	\emph{that} & \emph{she arrives}  &	\emph{arrives} & \emph{that} \\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'il{\= a} lmadrasa     & 	gakkoo ni & itta \\ % G22
+	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'ur{\= i}d    &    'an 'ugh{\= a}dir         & 	ik- & -itai \\ % G15
+	& \emph{wants}   &  \emph{to leave}  &	\emph{to go} & \emph{want} \\ \hline 
+\end{tabular}
+	\end{center}
+	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples: Across the world, the orders of different constituents are strikingly correlated with that of verb and object.
+	Selection is based on a more recent typological study by Dryer~\cite{dryer1992greenbergian}, restricted to those correlations that are annotated in available corpus data.
+See SI Section S1 for more on Greenberg correlations.
+	}\label{table:corr-dryer}
+\end{table}
 
@@
 
-Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study~\cite{lehmann1973structural, vennemann1974theoretical,jackendoff1977x,frazier1985syntactic,chomsky1988language, dryer1992greenbergian, hawkins1994performance}.
-
-A prominent line of research has argued that language universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language}).
+Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study~\cite{lehmann1973structural, jackendoff1977x,frazier1985syntactic,chomsky1993theory, kayne2003antisymmetry, baker2008macroparameter, dryer1992greenbergian, hawkins1994performance}.
+% cinque2005deriving, vennemann1974theoretical,
+Prominent among this research is the argument that
+language universals arise for \key{functional} reasons: that is, because they make human communication and language processing maximally efficient, and regularities across languages hold because these efficiency constraints are rooted in general principles of communication and cognition (e.g., \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,pinker1990natural,givon1991markedness,hawkins1994performance,hawkins2004efficiency,croft2001functional,haspelmath2008parametric,jaeger2011language,kirby2015compression}).
 Under this view, the various human languages represent multiple solutions to the problem of efficient information transfer given human cognitive constraints.
 
-Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between the Forces of \emph{Unification} and \emph{Diversification}.
-The Force of Unification is a pressure to reduce the complexity of the language to make production and processing as easy as possible.
-Conversely, the Force of Diversification favors languages that provide different utterances for different meanings, so that the listener can unambiguously identify the meaning from the utterance.
+In an early and  influential functional framework,
+Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between two pressures: to reduce complexity and to reduce ambiguity.
+What Zipf called the `Force of Unification' is a pressure to reduce the complexity of the language by reducing the number of distinctions made in the language, in order to make production and processing as easy as possible.
+The countervailing `Force of Diversification' favors languages that provide different utterances for different meanings, so that the listener can unambiguously identify the meaning from the utterance.
 These two forces act in opposing directions:
-Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required in order to provide higher informativity.
-The idea that many properties of language arise from the tension between these two pressures has a long and fruitful history in theoretical studies of language~\cite{gabelentz1901sprachwissenschaft,horn1984toward, lindblom1990explaining, schwartz1997dispersion, hawkins2004efficiency}.
+Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to provide enough information.
+The idea that many properties of language arise from the tension between these two pressures has a long and fruitful history in linguistics~\cite{gabelentz1901sprachwissenschaft,horn1984toward,lindblom1990explaining,hawkins2004efficiency,haspelmath2008functional}.
+% , wedel2013high, schwartz1997dispersion
 
-Recent work has used an information-theoretic formalization to computationally test these ideas in various domains of language, showing that they predict
-both basic statistical properties of languages~\cite{ferreri2003least} and language development~\cite{smith2013linguistic},
+Recent work has drawn on information theory to computationally test this ``dual pressures'' idea in various domains of language, showing that it predicts
+both basic statistical properties of languages~\cite{ferreri2003least,torre2019physical} and language evolution~\cite{kirby2015compression},
 and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.
+While it has been suggested that the dual pressure should also apply to grammar~\cite{hawkins2004efficiency}, testing these accounts is more difficult, as this requires large amounts of data representative of language use across languages,  computational methods for estimating the efficiency of entire languages, and a simulation methodology for comparing different possible grammars.
 
-While it has been suggested that these ideas should also apply to grammar~\cite{hawkins2004efficiency}, testing these accounts is more difficult in the domain of grammar, as this requires both large amounts of data representative of language use across languages and cultures, and computational methods for estimating the efficiency of entire languages.
-In this work, we address these challenges by combining large-scale text data from 51 languages with state-of-the-art machine learning techiques to estimate the communicative efficiency of grammar.
+In this work, we address these challenges by combining large-scale text data from 51 languages with machine learning techniques to estimate both aspects of the communicative efficiency of grammar: complexity and ambiguity.
+We use machine learning models based on neural networks to model the evolution of grammars towards efficiency.
 We apply this approach to the problem of explaining Greenberg's word order correlation universals.
-We first show that the word order of languages has evolved to optimize efficiency.
-We then show that efficiency optimization accounts for the Greenbergian word order correlations, and makes fine-grained per-language predictions about individual syntactic patterns.
-Our results show how general properties of efficient communication gives rise to the universal word order properties of human language.
 
-\begin{table}
-	\begin{center}
-%		\setlength\tabcolsep{1.5pt}
-\begin{tabular}{|c|ll|ll|}
-	\hline
-	%&
-	&	\multicolumn{2}{c|}{\textbf{Arabic} (English, ...)}   &        \multicolumn{2}{c|}{\textbf{Japanese} (Hindi, ...)} \\ \hline\hline
-	& \multicolumn{2}{c}{Correlates with...} & \multicolumn{2}{|c|}{Correlates with...}  \\
-	&	Verb & Object     & Object & Verb    \\ \hline\hline
-	& kataba & rasaa'il 	&tegami-o & kaita \\
-	&	\emph{wrote} & \emph{letters} & 	\emph{letter} & \emph{wrote} \\ \hline 
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    sadiiq       &	tomodachi & ni \\
-	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&kaana    &    sadiiq         &	tomodachi & datta \\
-	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak & udesho \\
-	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& sadiiq  &    John    & 	John no & tomodachi \\
-	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uha       & 	anata-ga yonda & hon \\
-	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tusil        & 	toochaku suru & koto \\
-	&	\emph{that} & \emph{she arrives}  &	\emph{has.arrived} & \emph{that} \\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'ila lmadrasa     & 	gakkoo ni & ittekimashita \\
-	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'uriid    &    'an 'ughaadir         & 	yom & itai \\
-	& \emph{wants}   &  \emph{to leave}  &	\emph{to leave} & \emph{want} \\ \hline \hline
-\end{tabular}
-	\end{center}
-	\caption{Greenberg's word order correlations, exemplified by Arabic (left) and Japanese (right) examples: Across the world, the orders of different constituents are strikingly correlated with that of verb and object.
-See Materials and Methods.}\label{table:corr-dryer}
-\end{table}
+In Study 1 we compare the word order of actual grammars of 51 languages with alternative ``counterfactual'' grammars parameterized by different word orders.
+We use our model to measure the communicative efficiency of each possible grammar, showing  that the grammars of real languages are more efficient than alternative grammars.
+The fact that real grammars lie at the Pareto frontier of the efficiency space of possible grammars suggests that the word order of languages has evolved to optimize communicative efficiency.
 
-\section{Formalizing Efficiency in Word Order}
+In Study 2 we test whether efficiency optimization accounts for the Greenbergian word order correlations.
+For each of the 51 languages, we create hypothetical grammars optimized for efficiency.
+We then test statistically whether these optimized grammars exhibit the Greenberg correlations, using a Bayesian mixed-effects logistic regression to control for language and language family.
+Efficiency optimization indeed predicts all eight Greenberg correlations.
+Our results show that general properties of efficient communication can give rise to these universal word order properties of human language.
+
+\section*{Grammars and Grammar Data}
+
+Following a long tradition in theoretical and computational linguistics, we formalize  the grammatical structure of languages using dependency trees~\cite{hays1964dependency,hudson1984word,melcuk1988dependency,corbett1993heads,tesniere2015elements}.
+This linguistic formalism represents grammatical dependencies as directed arcs between syntactically related words, annotated with grammatical relations like subject or object (Figure~\ref{fig:sent-dep}).
+While syntactic formalisms vary, the dependency grammar community has an agreed representation format  which has been used to  annotate corpora of text from dozens of languages \cite{ud2.1}, and there are
+computational methods for deriving such representations from other standard linguistic formalisms \cite{boston2009dependency}.
+
+\begin{figure}[ht]
+    \centering
+{
+\centering{
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+          She \& met \& friends  \\
+   \end{deptext}
+   \depedge{2}{1}{subject}
+   \depedge{2}{3}{object}
+\end{dependency}
+}
+}
+	\caption{An English sentence with annotated syntactic relations.}
+	\label{fig:sent-dep}
+\end{figure}
+
+Our models require a sample of syntactic structures as actually used by speakers across different languages, for which we draw on the recent
+Universal Dependencies project~\cite{ud2.1}, which has collected and created syntactic annotations for several dozens of different languages.
+51 languages had sufficient data for our purposes. 
+These corpora represent a typologically and genetically diverse group of languages. %  (see Figure~\ref{fig:per-lang})
+We obtained a total of 11.7M words in 700K sentences annotated with syntactic structures; with a median of 117K words and 7K sentences for each individual language.
+
+\section*{Study 1: Efficiency of Languages}
+\label{sec:relative-efficiency}
+
+We first ask whether the grammars of human languages reflect optimization for efficiency of communication.  To do this we
+compare the efficiency of the actual grammars of the 51 languages from the Universal Dependencies datasets to randomly constructed baseline grammars.
+
+The grammars of natural languages specify how the different words in a syntactic structure are ordered into a sentence, i.e., a string of words  \cite{adger2015syntax}.
+This is illustrated in Figure~\ref{fig:grammars}:
+We show how four different grammars order objects, adpositional phrases, and adpositions.
+For instance, Grammar 1---corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple}---orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
+Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
+Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.48\textwidth]{figure3.pdf}
+	\caption{Grammars define consistent ordering rules for syntactic structures.
+Here, Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb.
+Grammars 1 and 3 conform to the Greenberg correlations and are common around the world; Grammars 2 and 4 are rare or impossible.} %  \textcolor{red}{TODO bubble misleading?}
+	\label{fig:grammars}
+\end{figure}
+
+Beyond the syntactic relations exemplified in Figure~\ref{fig:grammars}, human languages have further types of syntactic relations.
+The Universal Dependencies project, the source of our data, defines a total of 37 syntactic relations.
+We adopt a variant of the grammar model developed by Gildea and colleagues~\cite{gildea2007optimizing,gildea2010grammars,gildea2015human}: a grammar assigns a weight from $[-1,1]$ to each of these 37 syntactic relations, and orders words according to the weights assigned to their relations
+(See \textit{Materials and Methods} for details).
+
+Given a large database of sentences annotated with syntactic structures (such as those at the top of Figure~\ref{fig:grammars}), obtained from a corpus of some real language $L$, we can apply a grammar to reorder the structures in the database into a dataset of counterfactual sentences belonging to a hypothetical language defined by that grammar (Figure~\ref{fig:grammars}).
+This hypothetical language has identical syntactic structures and grammatical relations as the true language $L$, but different word order.
+
+We create baseline grammars by randomly sampling the weights for each syntactic relation.
+These baseline grammars have systematic word order rules similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations.
+All four grammars in Figure~\ref{fig:grammars} are equally likely under this baseline distribution.
+
+For every one of the 51 languages, we construct 50 counterfactual baseline versions by randomly creating 50 baseline grammars and applying them to obtain counterfactual orderings for all syntactic structures that were available for that language.
+
+Having defined our space of possible word order grammars, we now turn to how to define and measure efficiency. 
+Following the information-theoretical literature on language processing, 
+we formalize the communicative efficiency of a language as a weighted combination of two terms:  the 
+amount of information that utterances contain about the underlying messages, and the  cost or difficulty of communication~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
+We model the informativity term as the 
+ degree to which listeners can reconstruct syntactic structures from an utterance, i.e., the \key{parseability} of the language. We model the cost or complexity term as the 
+ \key{predictability}, or negative entropy, of the utterances, since entropy is a standard measure of the complexity of any system of messages \cite{shannon1948mathematical}.
+ We use standard neural network methods to estimate the numerical values of parseability and predictability from counterfactually ordered corpora.
+ Efficiency is a weighted sum of parseability and predictability.
+ See \textit{Materials and Methods} for details, and SI Section 7 for experiments demonstrating that our results are robust to different methods of estimating parseability and predictability.
+
+For each language, we computationally construct grammars that are \emph{optimized} for efficiency (see \textit{Materials and Methods}).
+This optimization problem is challenging because both the parseability and predictability of a sentence can only be evaluated globally, in the context of an entire language.
+We address this challenge by introducing a simple, differentiable computational formalism for describing grammatical regularities. 
+Our formalism makes it possible to find optimal grammars by standard methods such as stochastic gradient descent (see SI Section S5).
+For each grammar, we report predictability and parseability as estimated on the data resulting from ordering the syntactic structures from the corpus according to the grammar.
 
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{figures/communication-two-langs.png}
-	\caption{Grammars define communication channels: A speaker (left) expresses a dependency structure into a string of words forming a sentence.
-The order of words is determined by the grammar of the language.
-A listener recovers a dependency structure.
-Communication is successful if the listener recovers the correct depenency structure.}
-	\label{fig:comm}
+    \includegraphics[scale=.5]{figure4.pdf}
+    \caption{Predictability and parseability of the real word order grammars of 51 languages (red), indicated by ISO codes, compared to baseline word order grammars (blue distribution).
+Predictability and parseability scores are $z$-scored within language, to enable comparison across languages.
+The gray curve indicates the approximate Pareto frontier of computationally optimized grammars, averaged over the 51 languages, with dashed standard deviations.}
+    \label{fig:pareto-plane}
 \end{figure}
 
-% Given a set of dependency structures, grammars define a set of sentences. Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figures/grammars-langs.png}
-	\caption{Grammars define consistent ordering rules for dependency structures.
-Here, Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb.
-Grammars 1 and 3 conform to the Greenberg correlations and are common around the world; Grammars 2 and 4 are rare or impossible.}
-	\label{fig:grammars}
-\end{figure}
+In Figure~\ref{fig:pareto-plane}, we plot predictability and parseability of the grammars of 51 languages, together with the distribution of random baseline grammars, and the approximate Pareto frontier defined by computationally optimized grammars.
+This Pareto frontier is approximate because it is an average of the positions of the optimized grammars generated for the corpus of each language.
+To enable fair comparison with baselines and the estimated frontier, we represent real languages by grammars extracted from the actual orderings observed in the databases. 
+These extracted grammars have the same representational constraints as the baseline and optimized grammars,
+including the fact that the orders are purely a function of the tree structure, and do not take into account other factors, such as discourse structure, which are not annotated in the corpora.
+For a comparison of the raw word orders from corpora against appropriate baseline grammars, see SI Section 8.
 
-We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
-In this model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
-Applying this model to word order, we take the message to consist of people, objects, events, and a set of syntactic and semantic relations that hold between them~\cite{fodor1975language,jackendoff1983semantics,pinker1990natural}.
-Following a long tradition in formal and computational liguistics, we formalize these in the format of dependency trees~\cite{hays1964dependency,melcuk1988dependency,corbett1993heads,tesniere2015elements}:
-This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
-Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.
+In Figure~\ref{fig:pareto-plane}, we see that real grammars are attracted towards the approximate Pareto frontier and away from the region of the baseline grammars.
+The majority of real grammars are above and/or to the right of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
+100\% of real grammars improve over their baselines on either predictability or parseability ($p<0.05$, by one-sided $t$-test, with Bonferroni correction and Hochberg's step-up procedure).
+%\cite{hochberg1988sharper}).
+90\% of real grammars improve over the baselines in parseability ($p < 0.05$), and 80~\% improve in predictability ($p < 0.05$).
+See SI Section S3 for additional analyses.
 
-%TODO also need to say these are syntax \& principle of compositionality
+\section*{Study 2: Greenberg Word Order Correlations}
+\label{sec:greenberg}
 
-Consider Figure~\ref{fig:comm}.
-Imagine the speaker wants to communicate that Mary saw John.
-The message (left) consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
-When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
-In human language, this order is largely determined by the \key{grammar} of the language (CITE).
-The speaker will order it as \emph{Mary saw John} (Figure~\ref{fig:comm}, top) in English, and as \emph{Mary John saw} (Figure~\ref{fig:comm}, bottom) in Japanese.
-The listener receives a sentence -- a linear string of words -- and decodes the underlying message in accordance with the grammar of the language (Figure~\ref{fig:comm}, right).
+We have found that the grammars of human languages concentrate along the Pareto frontier of parseability and predictability.
+Which grammatical properties characterize Pareto-optimal languages in general, and which properties of human languages make them efficient?
+Here we show that all languages close to the Pareto frontier---both real and counterfactual ones---are highly likely to satisfy Greenberg's correlation universals.
+That is, optimizing for efficiency produces languages that satisfy these correlations.
+In contrast, the baseline grammars are constructed without any correlations between the ordering of different syntactic relations, and will therefore mostly not satisfy those universals.
 
-The grammars of human languages provide consistent ordering rules for all syntactic relations.
-This is illustrated in Figure~\ref{fig:grammars}:
-We show how four different grammars order objects, adpositional phrases, and adpositions.
-For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
-Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
-Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
+We first considered the 51 real languages.
+Among the grammars fit to the 51 languages, the number of satisfied correlations is strongly correlated with efficiency ($\rho = 0.61$, $p<0.0001$), suggesting that satisfying the correlations improves language efficiency.
 
-The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
-Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
-Common to these accounts is the idea of maximizing the amount of information that linguistic forms provide about meanings, while constraining the diversity of forms.
+We next examine those grammars from Study 1 that we had computationally optimized for efficiency.
+We controlled for variation across different optima by creating eight optimized grammars for each of the 51 datasets of syntactic structures from real languages.
+For each real language, we created four optimized grammars with verb-object order, and four object-verb grammars.
+We test whether the process of efficiency optimization produces the Greenberg correlations.
 
-As we take the underlying messages to be dependency structures, the first term is the degree to which listeners can reconstruct dependency structures from an utterance, i.e., the \key{parseability} if the language.
-This is formalized as the amount of information that utterances $u$ provide about their underlying dependency structures $t$:
-\begin{equation}
-	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)}
-\end{equation}
-where the sum runs over all possible pairs of sentences $l$ and dependency structures $t$ in the language.
-This quantity describes the degree to which dependency structures can be unambiguously recovered from sentences, and quantifies how successful communication is (cf. Figure~\ref{fig:comm}).
-
-We formalize the cost of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
-In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
-Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
-In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
-%The \key{predictability} of the language is the negative entropy of the sentences:
-\begin{equation}
-	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
-\end{equation}
-where the sum runs over all possible sentences $u$ that belong to the language.
-In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences.
-%how many distinct sentences there are.
-
-Then, the efficiency of a language is the weighted combination of these two scoring functions~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}:
-%The \key{efficiency} of a language is a weighted combination
-\begin{equation}\label{eq:efficiency}
-	R_{\textit{Eff}} := R_{\textit{Parseability}} + \lambda R_\textit{Pred}
-\end{equation}
-with an interpolation weight $\lambda \in [0,1)$ (See SI appendix section 5 for mathematical justification).
-%In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
-
-\section{Results}
-
-We examined 
-(1) whether the grammars of human languages evolve towards optimizing efficiency of communication, and
-(2) whether this process of optimization accounts  for Greenberg's order correlation universals.
-
-Answering these questions requires a sample of dependency structures as actually used by speakers across different languages.
-Such samples have recently become available with the Universal Dependencies project~\cite{nivre2017universal}, which has collected and created dependency annotations for several dozens of different languages.
-51 languages had sufficient data for our purposes. %There was sufficient data from 51 languages for our We use data from 51 languages.
-These corpora represent a typologically and genetically diverse group of languages.
-
-Second, addressing these questions requires a computational method for estimating predictability and parseability.
-We draw on recent machine learning methods that form the foundation of state-of-the-art natural language processing.
-We estimate predictability using LSTM recurrent neural networks, general sequence models that are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
-We estimate parseability using a generic neural network architecture that casts recovery of dependency structures as a minimum spanning tree problem \citep{dozat2017stanford, zhang2017dependency}.
-See SI for details and for robustness of our results to these choices.
-
-\subsection{Relative efficiency of languages}
-\label{sec:relative-efficiency}
-
-We first demonstrate that the grammars of real languages are relatively efficient compared to random baselines, providing evidence that grammars evolve towards efficiency.
-We compare the efficiency of the actual grammars of the 51 languages from the Universal Dependencies datasets to randomly constructed baseline grammars.
-For every one of the 51 languages, we construct 20 counterfactual versions by randomly creating 20 grammars.
-These baseline grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}, see \textit{Materials and Methods} for details).
-We estimate predictability and parsability using start-of-the-art machine learning methods (see \textit{Materials and Methods}).
-Furthermore, for every one of the 51 languages, we computationally construct grammars that are \emph{optimized} for parsability, predictability, and efficiency (see \textit{Materials and Methods}).
-For each grammar, we report predictability and parsability as estimated on the data resulting from ordering the dependency structures from the corpus according to the grammar.
-
-%For every one of the 51 corpora, we computationally construct eight optimal grammars.
-%Note that the original orders of the actual languages do not enter this objective.
-%See SI for our method for creating optimized grammars.
-% For each language, we generated ten baseline word order grammars for the language.
-In Figure~\ref{fig:pareto-plane}, we plot predictability and parsability of the grammars of 51 languages, together with the distribution of random baseline grammars, and the Pareto frontier defined by computationally optimized grammars.
-Languages are attracted towards the Pareto frontier and away from the region of the baseline languages.
-The majority of real languages are above and to the right of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
-100~\% of languages improve on either predictability or parsability ($p<0.05$ each, by one-sided $t$-test, with Bonferroni correction).
-%By a one-sided \textit{t}-test, 
-92~\% of languages improve over the baselines in parsability ($p < 0.05$ each), 84~\% improve in predictability ($p < 0.05$ each).
-%At $\lambda = 0.9$, 86~\% of languages improve in overall efficiency ($p<0.05$ each).
-%TODO fix bug with czech data
+For each grammar (baseline, optimized, and real), we computed how many of the eight relations in Table~\ref{table:corr-dryer} had the same order as  Japanese (in contrast to Arabic).
+Figure~\ref{fig:joint} shows the results, separately for grammars with verb-object and object-verb orders.
+In optimized grammars, the order of the eight relations is strongly correlated with the placement of the object, similar to the 51 real languages in our sample.
+In contrast, baseline languages show no correlation.
 
@@
     \centering
-    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend-viz-10.pdf}
-    \caption{Predictability and parseability of 51 languages (red), indicated by ISO codes, compared to baseline word order grammars (blue distribution).
-Predictability and parseability scores are $z$-scored within language.
-The gray curve indicates the Pareto frontier of computationally optimized grammars, averaged over the 51 languages.} % TODO colors, fontsize, etc could conceivably be made prettier
-    \label{fig:pareto-plane}
+    \includegraphics[width=0.48\textwidth]{figure5.pdf}
+    \caption{
+    Efficiency optimization produces grammars where the orders of the eight relations in Table~\ref{table:corr-dryer} are strongly correlated with the order of verb and object.
+We arrange grammars (baseline, optimized, real) by the number of relations where the language patterns with Japanese (as opposed to with Arabic), and plot a kernel density estimate.
+Object-Verb order leads to grammars where object patterners precede (like Japanese); verb-object order leads to verb patterners preceding (like Arabic).
+Baseline grammars show no such correlation.
+    }
+    \label{fig:joint}
 \end{figure}
 
-\subsection{Greenbergian Word Order Correlations}
-\label{sec:greenberg}
-
-\begin{figure} % need Efficiency and (maybe) MLE langs on here
-%	\includegraphics[width=0.3\textwidth]{../results/correlations/figures/coverage-ground_eff-circles.pdf}
-	\begin{center}
-	\includegraphics[width=0.32\textwidth]{../results/correlations/figures/pred-eff-families-2.pdf}
-	\end{center}
-	\caption{Order of the eight correlates across 51 languages, in the real grammars (left) and predicted by efficiency optimization (right).
-Dark blue: Verb patterner \emph{precedes} object patterner(English, Arabic, ...).
-Light blue: Verb patterner \emph{follows} object patterner (Japanese, Hindi , ...).}\label{fig:per-lang}
-\end{figure}
+%So far, our results show that languages are efficient in part because they mostly satisfy the Greenberg correlations.
+We asked whether efficiency optimization predicts the eight correlations to hold in most languages.
+To answer this question, we constructed a Bayesian multivariate mixed-effects logistic regression model predicting which of the eight correlations an optimized grammar satisfies.
+We controlled for variation between the syntactic structures used in different languages and language families by entering the language and language family as random effects.
+See SI Section S4.3 for robustness to modeling choices.
 
 \begin{table}
-	\begin{center}
-%		\setlength\tabcolsep{1.5pt}
-\begin{tabular}{|c|ll|c|cc|ccc}
-	\hline
-	%&
-	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
-	&	verb & object     & & &   \\ 
-	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
-				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
-	&		\emph{to}            & \emph{a friend} &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
-	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
-	&	\emph{has}          & \emph{written}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
-	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
-	&	\emph{books} & \emph{that you read}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
-	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
-	&	\emph{went} & \emph{to school}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
-	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-\end{tabular}
-	\end{center}
+    \centering
+    \includegraphics[width=0.48\textwidth]{table2.pdf}
+
 \caption{Efficiency optimization accurately predicts the Greenbergian correlations.
-For each correlation, we provide the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
+For each correlation, we provide its prevalence (between 0\% and 100\%) among the actual grammars of the 51 languages (left), and the posterior distribution of the prevalence among grammars optimized for efficiency (right) on datasets from the 51 languages.
+Efficiency optimization predicts all eight correlations to hold in the majority of grammars, matching the distribution observed in real languages.
+%\textcolor{red}{make visually clear that it's 51 languages}
 }\label{table:corr-resu}
@@
 
-We now show that this process of efficiency optimization accounts for Greenberg's correlation universals.
-% we examine the computatinally optimized grammars.
-%computationally construct grammars that optimize efficiency, and evaluate these for dependency length and for Greenberg's universals.
-For every one of the 51 languages, we create hypothetical grammars that are computationally optimized for efficiency (see \textit{Materials and Methods}).
-In all results reported here, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency}.
-(See SI appendix section 5 for mathematical justification and for robustness to other choices).
-We test whether the process of efficiency optimization can explain these typological patterns, and make correct predictions about the orderings used in individual languages.
-%, this shows that efficiency optimization predicts Greenberg's word order correlation universals.
-
-We compare grammars optimized for efficiency with the actual grammars of the 51 languages, as extracted from the orderings attested in the original corpora and validated against expert judgments (see \textit{Materials and Methods}).
-
-We first evaluated which of the eight correlations emerge from efficiency optimization across language families.
-We controlled for variation across different optima by creating eight optimized grammars for each language.
-We controlled for variation between languages and language families by constructing a single Bayesian mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family as random effects.
-In this analysis, efficiency optimization predicts all eight correlations to hold across languages with posterior probability $0.9911$.
-
 In Table~\ref{table:corr-resu}, we compare the prevalence of the eight correlations in real and optimized languages.
 For the real languages, we indicate how many of the 51 languages satisfy a correlation.
-For the optimized languages, we indicate the posterior distribution of the proportion of satisfying languages, obtained from the mixed-effects analysis (see \textit{Materials and Methods}).
-Grammars optimized for efficiency accurately predict all eight correlations to hold at prevalences significantly greater than 50\%, similar to actual human languages.
-Optimizing for only predicability or only parseability does not predict all of the correlations (see SI).
+For the optimized languages, we indicate the posterior distribution of the proportion of satisfying languages, obtained from the mixed-effects analysis. % (see \textit{Materials and Methods}).
+Grammars optimized for efficiency predict all eight correlations to hold at prevalences significantly greater than 50\%, similar to actual human languages.
+In the multivariate mixed-effects analysis, efficiency optimization predicts all eight correlations to hold across languages (posterior probability $0.9911$). 
+Optimizing for only predicability or only parseability does not predict all of the correlations (See SI Section S4).
 
-We confirmed these results by evaluating predictions of efficiency optimization on the level of individual laguages (Figure~\ref{fig:per-lang}).
-For every one of the 51 languages, we show orders predicted by efficiency, for the eight correlating pairs.
-For each language, we consider the optimized grammar that had the best efficiency value, among those with the direction of the verb-object relation fixed to correspond to the actual language.
-Predictions of efficiency optimization agree with the actual word order patterns in 82~\% of the cases.
-Note that perfect agreement is not to be expected, as grammars will be impacted by factors other than efficiency optimization, such as their historical development.
-For instance, Dutch (nld) and Afrikaans (afr) have a prevalence of object-verb order in corpus data, and efficiency optimization accordingly predicts them to have corresponding orders (light blue).
-However, Dutch and Afrikaans have orders similar to the other Germanic languages (dark blue), which they are historically closely related to.
+\section*{Discussion}
 
-We have directly implemented parseability and predictability using general-purpose statistical models of language and dependency structures.
-Prior work has considered Dependency Length Minimization (DLM) as an approximate formalization of efficiency optimization~\cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-This is the idea that languages order words so as to minimize the overall length of dependencies.
-This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in parsing and prediction~\cite{hawkins1994performance,gibson1998linguistic,gibson2000dependency, futrell2017memory}.
-Indeed, directly modeling efficiency optimization predicts DLM, and appears to predict word order universals with greater accuracy than optimizing for DLM (see SI, \emph{Relation to Dependency Length Minimization}).
+We found that the grammars of natural languages are more efficient than baseline grammars, and that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication.
 
-\section{Conclusion}
+Our work makes crucial use of neural network models for estimating the efficiency of languages.
+This method currently requires large computational resources; it still takes about three weeks to create optimized grammars for 51 languages, even with specialized hardware.
+We believe that further advances in machine learning will reduce the computational cost, making this approach more widely applicable.
 
-We found that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
-We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals.
-Beyond our present results, we provide a complete formalization and computational framework in which theories of the functional optimization of languages can be tested.
-Other objective functions could be hypothesized and tested within our framework; furthermore, future advances in machine learning will enable the optimization of richer and richer models of grammar.
+What makes the grammars of human languages efficient?
+Study 2 shows that Greenberg's correlations are one key property that real languages share with optimal grammars.
+Prior work has suggested \emph{Dependency Length Minimization} as another characteristic of efficient word order.
+This is the idea that word order minimizes the average distance between syntactically related words.
+It is known that human languages reduce this distance compared to random baselines~\cite{liu2008dependency,futrell2015largescale,liu2017dependency,temperley2018minimizing}.
+Our optimized grammars also share this property: we find that 100\% of grammars optimized for efficiency also reduce average distance between related words compared to baselines ($p < 0.05$, by one-sided $t$-test).
 
-A major question for functional explanations for linguistic universals is: \emph{how} do languages come to be optimized?
+To some extent, the Greenberg correlations and dependency length minimization are related, because the Greenberg correlations help reduce the distance between related words \citep{hawkins1994performance,temperley2008dependency}.
+Consider again the sentence `I wrote letters to friends' (cf.
+Figures~\ref{fig:corr-table} and \ref{fig:grammars}).
+Both real and optimized grammars of English linearize its syntactic structure as follows:
+\begin{center}
+\begin{dependency}[theme = simple]
+   \begin{deptext}[column sep=1em]
+          I \& wrote \& letters \& to \& friends  \\
+   \end{deptext}
+   \depedge{2}{1}{subject}
+   \depedge{2}{3}{object}
+   \depedge{2}{4}{oblique}
+   \depedge{4}{5}{case}
+\end{dependency}
+\end{center}
+This ordering exhibits correlations (1) and (7) from Table~\ref{table:corr-dryer}.
+Among all possible ways of ordering this syntactic structure, this one also minimizes the average distance between any two syntactically related words: E.g., inverting `to' and `friends' would increase the distance between `wrote' and `to'.
+
+It may come as a surprise that grammars which are efficient according to our metric also have low dependency length, even though dependency length is never considered explicitly during the calculation of efficiency nor the procedure for optimizing grammars.
+The result is especially surprising given that our efficiency metric does not incorporate any kind of memory limitations, whereas previous functional explanations for dependency length minimization have typically been based on the idea of limited working memory resources available during language production and comprehension \citep{gibson1998linguistic,futrell2017noisycontext} (although see \citep{hawkins1994performance} for a motivation of dependency length minimization which is not based in memory limitations).
+Our results suggest that both Greenberg's correlations and dependency length minimization might be explainable purely in terms of maximizing the general parseability and predictability of utterances, without a need for further constraints.
+See SI Section S12 for further discussion along with some simulations demonstrating how grammars that satisfy Greenberg's correlations can be more efficient in a generic sense.
+
+An idea related to functional optimization as we have explored it here is the idea that grammars are biased towards simplicity in terms of the number of parameters required to specify the grammar \cite{culbertson2016simplicity}.
+For example, it has been proposed that languages have a single head-directionality parameter and that this accounts for the Greenberg correlations \cite{chomsky1993theory,baker2001atoms}.
+As an explanation of correlations, this idea turns out to overpredict correlations \cite{dryer1992greenbergian,baker2008macroparameter}, and more recent research in syntactic theory has provided evidence against it \cite{kayne1994antisymmetry,kandybowicz2003directionality,kayne2011there}.
+Nevertheless, future research should examine whether there are more principled connections between communicative efficiency and grammar simplicity.
+
+A major question for functional explanations for linguistic universals is: \emph{how} do languages end up optimized?
 Do speakers actively seek out new communicative conventions that allow better efficiency?
@@
 To the extent that language universals arise from biases in learning or in the representational capacity of the human brain, our results suggest that those biases tilt toward communicative efficiency.
-Our work does provide evidence against the idea that word order universals are best explained in terms of learning biases that are irreducibly arbitrary and genetic in nature \cite{chomsky2010some}.
 
-While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into dependency trees.
-The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication.
-Relatedly, there are  remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
-Our models cannot capture this trade-off for technical reasons: the parseability objective does not operate over wordforms, only POS tags, and so it does not take advantage of morphological cues to syntactic structure.
-Future work can investigate whether these and other remaining universals can be explained using more sophisticated models of how meaning representations are transduced into strings of words, based on larger databases with richer annotations.
+Unlike crosslinguistic efficiency studies in the domain of lexical semantics~\cite{kemp2012kinship,regier2015word,zaslavsky2018efficient}, we did not derive a single universal bound for the efficiency across all 51 languages in Study 1; instead, we constructed optimized grammars individually for each language.
+Each language $L$ has its own distribution of tree structures that speakers communicate, and different grammars may be optimal for different tree structure distributions (see SI Section S3.5).
+Our results show that the word order of each language $L$ is approximately optimal for the tree structures used in $L$.
 
-\matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
+While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into trees of syntactic relations.
+We believe a promising avenue for future work is to determine whether these more basic properties themselves might also be explainable in terms of efficient communication.
+
+Our work provides evidence that the grammatical structure of languages is shaped by the need to support efficient communication.
+Beyond our present results, our contribution is to provide a computational framework in which theories of the efficiency optimization of languages can be tested rigorously.
+While our study has focused on syntax, our results suggest that this method can be fruitfully applied to testing efficiency explanations in other domains of language structure.
+
+\matmethods{
+\unskip
+\subsection*{Corpus Data}
+We use the Universal Dependencies (UD) 2.1 data~\cite{ud2.1}.
 We use all languages for which at least one treebank with a training partition was available, a total of 51 languages.
@@
 Punctuation was removed.
-%We removed punctuation. %, which we operationalized by removing all dependents of \emph{punct} dependencies, and recursively removing their dependents.
-
 Universal dependencies represents as dependents some words that are typically classified as heads in syntactic theory.
 This particularly applies to the \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark} dependencies.
-Following prior work studying dependency length minimization \cite{futrell2015largescale}, we modified each treebank by inverting these dependencies, promoting the dependent to the head position.
-We report results on this modified version of UD.
+Following prior work studying dependency length minimization \cite{futrell2015largescale}, we applied automated conversion to a more standard formalism, modifying each treebank by inverting these dependencies, and promoting the dependent to the head position.
+When a head had multiple such dependents, we iteratively applied the conversion until no such dependents were left.
+Language-specific relation types were truncated to their universal counterparts both in the design of word order grammars, and for modeling parseability.
 
-The efficiency optimization results from Table~\ref{tab:results-dryer} were preregistered: \url{http://aspredicted.org/blind.php?x=8gp2bt}.
+\subsection*{Word Order Grammars}
+We adapt the grammar model of~\cite{gildea2007optimizing} to UD.
+A grammar assigns a parameter $x_\tau \in [-1,1]$ to every relation  $\tau$ belonging to the 37 universal syntactic relations defined by UD 2.1.
+A syntactic structure, consisting of a set of words and syntactic relations between them, is then ordered into a string of words recursively starting from the root; the dependents of a word then are ordered around the head according to the values $x_\tau$ corresponding to their syntactic relations; those dependents where $x_\tau < 0$ are ordered before the head; the others after the head.
+See SI Section S5.2 for the methodology used to extract the languages' actual grammars from datasets, and for validation against expert judgments.
 
+\subsection*{Formalizing Efficiency}
+We adopt the formalization of language efficiency of \cite{ferreri2003least}, closely related to the Information Bottleneck \citep{strouse2017deterministic}, which has recently been successfully applied to model lexical semantics \cite{zaslavsky2018efficient}.
+Very similar formalizations of Zipf's ideas have been proposed across the information-theoretic literature on language~\cite{frank2012predicting,kemp2012kinship,regier2015word, xu2016historical}. 
+See SI Section S2.1 for discussion.
+
+In this framework, the overall efficiency of language is a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication  \cite{ferreri2003least,frank2012predicting,kemp2012kinship,regier2015word,zaslavsky2018efficient}.
+We model the first term as the
+ degree to which listeners can reconstruct syntactic structures from an utterance, i.e., the \key{parseability} of the language.
+ This is formalized as the amount of information that utterances $u$ provide about their underlying syntactic structures $t$:
+\begin{equation}
+	R_{Pars} := I[\mathcal{U},\mathcal{T}] = \sum_{t,u} p(t,u) \log \frac{p(t|u)}{p(t)},
+\end{equation}
+where the sum runs over all possible pairs of utterances $u$ and syntactic structures $t$ in the language.
+
+Again following \cite{ferreri2003least}, we formalize the complexity of a language as its entropy.
+This corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
+Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
+Entropy is also a general measure of the complexity of any system of messages \cite{shannon1948mathematical}.
+In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utterances:
+\begin{equation}
+	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u),
+\end{equation}
+where the sum runs over all possible sentences $u$ in the language.
+
+Maximizing one of the two scoring functions under a constraint on the other function (e.g., maximizing parseability under a constraint on the minimal predictability) amounts to maximizing a weighted combination of the two scoring functions~\cite{ferreri2003least}:
+\begin{equation}\label{eq:efficiency}
+	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred},
+\end{equation}
+with an interpolation weight $\lambda \in [0,1)$ that controls the relative strength of the two pressures.
+When optimizing grammars for efficiency, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency} in order to give approximately equal weight to both components.
+See SI Section S2.2 for mathematical discussion of $\lambda$, and robustness to other choices.
+
+We estimate predictability using LSTM recurrent neural networks~\citep{hochreiter1997long}, general sequence models that are the strongest known predictors of the surprisal effect on human processing effort~\citep{frank2011insensitivity,goodkind2018predictive}.
+We estimate parseability using a generic neural network architecture that casts recovery of syntactic structures as a minimum spanning tree problem \citep{dozat2017stanford, zhang2017dependency}.
+In order to reduce overfitting in the optimization process, we use an unlexicalized parsing setup, and add POS tags when estimating predictability.
+Grammars are optimized for efficiency by simultaneous gradient descent on the parameters of the grammar and these neural models.
+All parseability and predictability values are reported on the held-out (\emph{dev}) partitions from the predefined split for each UD corpus.
+See SI Sections S5--S8 for details and for robustness of our results to modeling choices, including evidence that our results are not specific to any particular language model or parser.
+
+\subsection*{Data Availability}
 Code and results are available at \url{https://github.com/m-hahn/grammar-optim}.
-
-See the SI for details on the neural language models and parsers used, and on the optimization procedures.}
-
-\showmatmethods{} % Display the Materials and Methods section
-
-%\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Noah Goodman, Chris Manning, and audiences at CAMP 2018 and CUNY 2019 for helpful discussion.}
-
-\showacknow{} % Display the acknowledgments section
-
-%\bibliographystyle{acl_natbib}
-%\bibliography{extracted}
-\bibliography{everything}
-
-\end{document}
-
-%%
-%%
-%
-%\begin{table}
-%	\begin{center}
-%%		\setlength\tabcolsep{1.5pt}
-%\begin{tabular}{c|ll|c|cc|ccc}
-%	\hline
-%	%&
-%	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
-%	&	verb & object     & & &   \\ 
-%	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
-%		& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_case.pdf} } %\textbf{86}  }
-%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_case.pdf}       }
-%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
-%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_case.pdf}   }   
-%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_case.pdf}     } & \\
-%	&		\emph{to}            & \emph{a friend} &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
-%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_cop.pdf} }% \textbf{94} }
-%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
-%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_cop.pdf} } % \textbf{53}     }
-%	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_cop.pdf}   }   
-%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_cop.pdf}   } & \\
-%	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
-%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-aux.pdf}}
-%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
-%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-aux.pdf}     }
-%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-aux.pdf}  }   
-%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-aux.pdf}  } & \\
-%	&	\emph{has}          & \emph{written}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
-%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-nmod.pdf}}
-%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
-%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-nmod.pdf}   }
-%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
-%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-nmod.pdf}  } & \\
-%	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
-%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-acl.pdf}}
-%	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
-%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-acl.pdf}   }
-%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-acl.pdf}  }   
-%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-acl.pdf}  } & \\
-%	&	\emph{books} & \emph{that you read}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
-%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_mark.pdf}}
-%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
-%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_mark.pdf}     }
-%	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_mark.pdf} }   
-%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_mark.pdf}  }  &\\
-%	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-%	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
-%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-obl.pdf}}
-%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
-%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-obl.pdf}     }
-%	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-obl.pdf} }   
-%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-obl.pdf}  } &  \\
-%	&	\emph{went} & \emph{to school}  &&&\\ \hline
-%	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
-%		&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-xcomp.pdf}}
-%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
-%	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-xcomp.pdf}     }
-%	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
-%		&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-xcomp.pdf}  } & \\
-%	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
-% \hline
-%%comp. &
-%%\multirow{3}{*}{NP}&
-%%	AP &
-%%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-%\end{tabular}
-%	\end{center}
-%\caption{OPTION 1: Efficiency optimization accurately predicts the Greenbergian correlations.
-Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-%Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
-%Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
-%%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%%Not all correlations are satisfied by every natu
-%}\label{table:corr-dryer}
-%\end{table}
-%
-
-%
-%\begin{table}
-%	\begin{center}
-%%		\setlength\tabcolsep{1.5pt}
-%\begin{tabular}{|ll|c|cc|}
-%	\hline
-%	%&
-%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  &  \multirow{2}{*}{Eff}  \\ 
-%	verb & object     & & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%%adp. &
-%	adposition    &    NP       
-%	& \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
-%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}       }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}   }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}     }  \\
-%	\emph{to}            & \emph{a friend} &&&\\ \hline
-%copula    &    NP         
-%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
-%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}       }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
-%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}   }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}   }  \\
-%	\emph{is}        & \emph{a friend}  &&&\\ \hline
-%auxiliary    &    VP       
-%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
-%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}       }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}     }
-%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  }  \\
-%	\emph{has}          & \emph{written}  &&&\\ \hline
-%noun    &    genitive      
-%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
-%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
-%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  }  \\
-%	\emph{friend} &  \emph{of John}  &&&\\ \hline
-%noun    &    relative clause      
-%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
-%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}     }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
-%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  }  \\
-%	\emph{books} & \emph{that you read}  &&&\\ \hline
-%complementizer    &    S        
-%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
-%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf}     }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}     }
-%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf} }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  }  \\
-%	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-%verb    &    PP         
-%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
-%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf}     }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}     }
-%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  }  \\
-%	\emph{went} & \emph{to school}  &&&\\ \hline
-%want    &    VP        
-%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
-%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}     }
-%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
-%	&   \parbox[c]{0.05\textwidth}{\multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}  }}  \\
-%	\emph{wants}   &  \emph{to leave}  &&&\\ \hline
-% \hline
-%%comp. &
-%%\multirow{3}{*}{NP}&
-%%	AP &
-%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-%\end{tabular}
-%	\end{center}
-%	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%For each correlation, we provide an English example, and show the distribution of real languages, of languages optimized for DLM, and of languages optimized for efficiency.
-The process of optimization fir efficiency and DLM explains all eight universals.
-%%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%%Not all correlations are satisfied by every natu
-%}\label{table:corr-dryer}
-%\end{table}
-%
-%
-
-%
-%
-%\begin{table}
-%	\begin{center}
-%%		\setlength\tabcolsep{1.5pt}
-%\begin{tabular}{|ll|c|cc|cc|}
-%	\hline
-%	%&
-%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  &  \multirow{2}{*}{Parser}  &  \multirow{2}{*}{Eff}  \\ 
-%	verb & object     & & &   \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%%adp. &
-%	adposition    &    NP       
-%	& \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_case.pdf} } %\textbf{86}  }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_case.pdf}       }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
-%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_case.pdf}   }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_case.pdf}     }  \\
-%	\emph{to}            & \emph{a friend} &&&\\ \hline
-%copula    &    NP         
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_cop.pdf} }% \textbf{94} }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_cop.pdf} } % \textbf{53}     }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_cop.pdf}   }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_cop.pdf}   }  \\
-%	\emph{is}        & \emph{a friend}  &&&\\ \hline
-%auxiliary    &    VP       
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-aux.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-aux.pdf}     }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-aux.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-aux.pdf}  }  \\
-%	\emph{has}          & \emph{written}  &&&\\ \hline
-%noun    &    genitive      
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-nmod.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-nmod.pdf}   }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-nmod.pdf}  }  \\
-%	\emph{friend} &  \emph{of John}  &&&\\ \hline
-%noun    &    relative clause      
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-acl.pdf}}
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-acl.pdf}   }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-acl.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-acl.pdf}  }  \\
-%	\emph{books} & \emph{that you read}  &&&\\ \hline
-%complementizer    &    S        
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-lifted_mark.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_mark.pdf}     }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_mark.pdf} }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-lifted_mark.pdf}  }  \\
-%	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-%verb    &    PP         
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-obl.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-obl.pdf}     }
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-obl.pdf} }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-obl.pdf}  }  \\
-%	\emph{went} & \emph{to school}  &&&\\ \hline
-%want    &    VP        
-%	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-ground-xcomp.pdf}}
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
-%%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-xcomp.pdf}     }
-%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
-%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-efficiency_large_best-xcomp.pdf}  }  \\
-%	\emph{wants}   &  \emph{to leave}  &&&\\ \hline
-% \hline
-%%comp. &
-%%\multirow{3}{*}{NP}&
-%%	AP &
-%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-%\end{tabular}
-%	\end{center}
-%	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%For each correlation, we provide an English example, and show the distribution of real languages, of languages optimized for DLM, and of languages optimized for efficiency.
-The process of optimization fir efficiency and DLM explains all eight universals.
-%%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%%Not all correlations are satisfied by every natu
-%}\label{table:corr-dryer}
-%\end{table}
-%
-%
-
-%\begin{table*}
-%	\begin{center}
-%\begin{tabular}{|ll|l|l|l|ll|l|}
-%	\hline
-%	%&
-%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
-%	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%%adp. &
-%adposition    &    NP       &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
-%copula    &    NP        &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
-%auxiliary    &    VP       &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
-%noun    &    genitive      &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
-%noun    &    relative clause       &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
-%complementizer    &    S        &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
-%verb    &    PP     &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
-%want    &    VP      &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
-%verb    &    subject    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
-%verb    &    manner adverb     &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
-% \hline
-%%comp. &
-%%\multirow{3}{*}{NP}&
-%%	AP &
-%    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
-%\end{tabular}
-%	\end{center}
-%	\caption{Greenbergian Correlations.
-Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%%	For clarity, we organize the correlations by the category of the `verb patterner'.
-%	For each correlation, we give our operationalization in terms of UD.
-%A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
-%	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
-%	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
-%\end{table*}
-%
-
-%We show that Efficiency optimization predicts DLM, providing a first-principles functional explanation.
-%%Functional studies of word order have proposed Dependency Length Minimization (DLM) as general principle of word order \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-%Figure~\ref{fig:deplength} shows average dependency length for real languages, random baselines, and languages optimized for dependency length and efficiency.
-%Optimizing for efficiency lowers dependency length relative to random baselines.
-
-%It has also been argued theoretically that this principle should account for Greenberg's correlations, though this link is nontrivial and depends on the precise distributions of tree topologies in syntactic structures.
-%As a principle of word order, DLM is stipulative: There are other processing models, such as surprisal, that are strongly predictive of human processing cost, and make predictions distinct from, and partly conflicting with, those of DLM~\cite{hale2001probabilistic,levy2008expectation}.
-%\cite{futrell2017memory} suggests that efficiency predicts DLM.
-
-%Our results suggest that dependency length minimization is a by-product of efficiency maximization.
-In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the dependency length of the real language; on average, the real language is best matched by efficiency optimization.
-
-%\begin{figure} % need Efficiency and (maybe) MLE langs on here
-%\includegraphics[width=0.4\textwidth]{../results/dependency-length/figures/depl-violin.pdf}
-%	\caption{Average dependency length for real languages (blue), random baselines (purple), and for languages optimized for efficiency (green) and explicitly for dependency length (red).
-Dependency length was computed from sentences of lengths 8-12 and averaged over words.}\label{fig:deplength}
-%\end{figure}
-
-\begin{table}
-	\begin{center}
-%		\setlength\tabcolsep{1.5pt}
-\begin{tabular}{|c|ll|c|cc|ccc}
-	\hline
-	%&
-	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
-	&	verb & object     & & &   \\ 
-	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
-				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } & \\
-	&		\emph{to}            & \emph{a friend} &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } & \\
-	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } & \\
-	&	\emph{has}          & \emph{written}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
-	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
-	&	\emph{books} & \emph{that you read}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
-	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
-	&	\emph{went} & \emph{to school}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
-	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-\end{tabular}
-	\end{center}
-\caption{OPTION 2: Efficiency optimization accurately predicts the Greenbergian correlations.
-Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
-Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
-%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%Not all correlations are satisfied by every natu
-}\label{table:corr-dryer}
-\end{table}
-
-\begin{table}
-	\begin{center}
-%		\setlength\tabcolsep{1.5pt}
-\begin{tabular}{c|ll|c|cc|ccc}
-	\hline
-	%&
-	&	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Predicted} & \\ 
-	&	verb & object     & & &   \\ 
-	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    NP       
-				&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_case.pdf}     } 
-		&    & \\
-	&		\emph{to}            & \emph{a friend} &&&\\ 
-	&&& {  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_case.pdf}     } \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    NP         
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_cop.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_cop.pdf} } % \textbf{53}     }
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_cop.pdf}   }   
-		& & \\
-	&	\emph{is}        & \emph{a friend}  &&&\\ 
-	&&&   {  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_cop.pdf}     } \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    VP       
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_aux.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-aux.pdf}     }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-aux.pdf}  }   
-		&  & \\
-	&	\emph{has}          & \emph{written}  &&&\\ 
-	&&&  {  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_aux.pdf}     } \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_nmod.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-nmod.pdf}   }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_nmod.pdf}     } & \\
-	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_acl.pdf}     } 
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-acl.pdf}   }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-acl.pdf}  }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_acl.pdf}     } & \\
-	&	\emph{books} & \emph{that you read}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    S        
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_lifted_mark.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-lifted_mark.pdf}     }
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-lifted_mark.pdf} }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_lifted_mark.pdf}     } & \\
-	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    PP         
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_obl.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-obl.pdf}     }
-	%	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-obl.pdf} }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_obl.pdf}     } & \\
-	&	\emph{went} & \emph{to school}  &&&\\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    VP        
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Real_xcomp.pdf}     } 
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }
-	%	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-langmod-xcomp.pdf}     }
-	%	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-parser-nmod.pdf}  }   
-		&   \multirow{2}{*}{  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_Efficiency_xcomp.pdf}     } & \\
-	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-%    \multicolumn{4}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, : $p < 0.01$, : $p < 0.001$}}
-\end{tabular}
-	\end{center}
-\caption{OPTION 2: Efficiency optimization accurately predicts the Greenbergian correlations.
-Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-For each correlation, we provide an English example, and show the distribution among the actual grammars of the 51 languages (left), and among grammars optimized for efficiency (right).
-Blue bars count grammars conforming to the correlation, red bars count grammars violating it.
-Efficiency optimization predicts all eight correlations to hold in the majority of grammars.
-%Given the statistical nature of the correlations, not every language satisfies every one of them.
-%Not all correlations are satisfied by every natu
-}\label{table:corr-dryer}
-\end{table}
-
-%the meaning
-
-\begin{figure}
-\centering
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
-	   katabt \& risāla \& li \& sadīq  \\
-	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
-	   wrote \& letter   \& to \& friend  \\
-   \end{deptext}
-\end{dependency}
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-      tomodachi \& ni \& tegami-o \& kaita \\
- \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
-	   friend \& to \& letter  \& wrote \\
-   \end{deptext}
-\end{dependency}
-
-%\begin{dependency}[theme=simple]
-%    \begin{deptext}
-%    I \& wrote \& a \& letter \& to \& a \& friend \\
-%    \end{deptext}
-%    \depedge{2}{1}{nsubj}
-%    \depedge{2}{4}{obj}
-%    \depedge{4}{3}{ }
-%    \depedge{4}{3}{ }
-%    \depedge{2}{5}{}
-%    \depedge{5}{7}{ }
-%    \depedge{7}{6}{ }
-%\end{dependency}
-	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese-simple}
-\end{figure}
-
-\section{Minimizing Dependency Length Predicts Universals}
-
-..
-
-\begin{figure}
-\centering
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
-	   katabt \& risāla \& li \& sadīq  \\
-	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
-	   wrote \& letter   \& to \& friend  \\
-   \end{deptext}
-	%   \deproot{3}{ROOT}
-   \depedge{1}{2}{obj}
-	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
-   \depedge{1}{3}{}
-   \depedge{3}{4}{}
-   %\depedge[arc angle=50]{7}{6}{ATT}
-\end{dependency}
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-      tomodachi \& ni \& tegami-o \& kaita \\
- \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
-	   friend \& to \& letter  \& wrote \\
-   \end{deptext}
-	%   \deproot{3}{ROOT}
-   \depedge{4}{2}{}
-	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
-   \depedge{2}{1}{}
-   \depedge{4}{3}{obj}
-   %\depedge[arc angle=50]{7}{6}{ATT}
-\end{dependency}
-\end{figure}
-
-Dependency Length:
-
-The sentences in Figure~\ref{fig:arabic-japanese} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words.
-For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} if the verb, rather than e.g.
-the subject or an indirect object.
-A dependency arc is drawn from a \key{head} (e.g.
-\emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g.
-\emph{tegami-o}).
-Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}.
-Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.
-
-In order to interpret a sentence correctly, a listener must be able to reconstruct the information expressed in the dependency tree and its labels.
-This is because, according to the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.
-
-The contribution of this work is (i) to provide a principled formalization of the notion of communicative efficiency in word order, based on related work in information theory \cite{cover2006elements,tishby1999information}, bounded rationality \cite{genewein2015bounded}, and computational neuroscience \cite{friston2009predictive}; (ii) to introduce a computational framework in which we can produce simulated grammars directly optimized for our notion of efficiency, using state-of-the-art machine learning techniques \cite{hochreiter1997long,goldberg2017neural,dozat2017stanford}; and (iii) to demonstrate that in doing so, we reproduce the major word order patterns reported in \cite{greenberg1963universals} and related work \cite{dryer1992greenbergian}, as well as reproducing the predictions of previous heuristic theories of word order optimization \cite{hawkins1994performance,ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
-
-We formalize the notion of communicative efficiency using two terms: one having to do with minimizing the difficulty of linguistic communication, and one having to do with maximizing the information transferred in such communication.
-We quantify the difficulty of linguistic communication in terms of the word-by-word \key{predictability} of strings, which has been found to be a highly accurate and general predictor of human online comprehension difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and which can be justified in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
-We quantify the information transferred in language as \key{parseability}: the quantity of information linguistic utterances contain about latent hierarchical tree structures.
-We also compare the results of optimizing for communicative efficiency to the results of optimizing for dependency length minimization, a previously proposed approximate metric of the amount of memory involved in language production and comprehension \cite{hudson1995measuring,gibson1998linguistic,gibson2000dependency,liu2017dependency,temperley2018minimizing}.
-
-To evaluate whether communicative efficiency can explain word order universals, we apply the following method \cite{gildea2007optimizing}.
-First, we develop word order grammars: simple, interpretable models of word order, implemented as functions that take in an unordered syntactic dependency tree and output a distribution over possible word orders for that tree.
-Second, we take dependency treebanks of natural languages and compute word order grammars which are optimized for an objective functions representing communicative efficiency.
-Third, we evaluate whether the optimized word order grammars reproduce the desired word order universals when applied to attested dependency trees from the treebanks across many languages.
-Our favorable results suggest that it may be possible to generally model human languages as efficient codes for communication among agents with generic human-like cognitive and communicative constraints.
-%The best way to describe a human language might be as a solution to an equation involving such constraints.
-
-%We draw on data from the Universal Dependencies project \cite{nivre-universal-2017}, which has been developing treebanks in a unified format for several dozens of diverse languages.
-%This universal annotation scheme is particularly useful for studying word order universals, as many of them can be operationalized  in terms of crosslinguistically valid syntactic notions.
-% such as \emph{verbs} and \emph{objects}.
-
-\section{Background}
-
-% TODO where do we say that we're basing our work on data from 51 languages?
-
-%\subsection{Explananda: Word Order Correlations}
-%
-%Working with a database of 30 languages, Greenberg \cite{greenberg1963universals} proposed 45 universals of language that he determined to be true of most or all languages.
-%31 of them were concerned with word order, the remainder with morphology.
-%Many of the word order universals state \emph{correlations} between patterns:
-%For instance, Universals 3 and 4 state that ``\emph{Languages with dominant VSO order are always prepositional.}''  (e.g.
-Arabic), and ``\emph{With overwhelmingly greater than chance frequency, languages with normal SOV order are postpositional.}'' (e.g.
-Japanese).
-%These are illustrated in Figure~\ref{fig:arabic-japanese}:
-%Arabic has basic order verb-object (VO) and prepositions (\emph{li} `to'), while Japanese has order object-verb (OV) and postpositions (\emph{ni} `to').
-Our goal is to explain Greenberg's reported word order correlations, as updated in \cite{dryer1992greenbergian}.
-
-\begin{figure}
-\centering
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
-	   katabt \& risāla \& li \& sadīq  \\
-	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
-	   wrote \& letter   \& to \& friend  \\
-   \end{deptext}
-	%   \deproot{3}{ROOT}
-   \depedge{1}{2}{obj}
-	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
-   \depedge{1}{3}{}
-   \depedge{3}{4}{}
-   %\depedge[arc angle=50]{7}{6}{ATT}
-\end{dependency}
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-      tomodachi \& ni \& tegami-o \& kaita \\
- \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
-	   friend \& to \& letter  \& wrote \\
-   \end{deptext}
-	%   \deproot{3}{ROOT}
-   \depedge{4}{2}{}
-	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
-   \depedge{2}{1}{}
-   \depedge{4}{3}{obj}
-   %\depedge[arc angle=50]{7}{6}{ATT}
-\end{dependency}
-
-%\begin{dependency}[theme=simple]
-%    \begin{deptext}
-%    I \& wrote \& a \& letter \& to \& a \& friend \\
-%    \end{deptext}
-%    \depedge{2}{1}{nsubj}
-%    \depedge{2}{4}{obj}
-%    \depedge{4}{3}{ }
-%    \depedge{4}{3}{ }
-%    \depedge{2}{5}{}
-%    \depedge{5}{7}{ }
-%    \depedge{7}{6}{ }
-%\end{dependency}
-	\caption{A sentence in Arabic (top) and Japanese (bottom), translating to `I wrote a letter to a friend.' Note the reversal of word order: Arabic has verb-object order and prepositions, while Japanese has object-verb order and postpositions.}\label{fig:arabic-japanese}
-\end{figure}
-
-% TODO what to do about these labels...? They're not UD...
-
-\subsection{Dependency trees}
-
-\section{Optimizing word order grammars}
-
-Here we describe our model of word order and how we derive optimal grammars within that model in terms of communicative efficiency.
-
-Let $t$ be an unordered dependency tree, i.e.
-a structure as in Figure~\ref{fig:arabic-japanese} but without any specified ordering of the words.
-We think of an unordered dependency tree $t$ as representing the compositional meaning to be conveyed by a sentence.
-We define a \key{word order grammar} as a function $L_\theta(t)$, whose behavior is specified by parameters $\theta$, which takes an unordered dependency tree $t$ as input and produces as output a probability distribution over ordered sequences of words $\mathbf{w}$ linearizing the tree, or equivalently, a probability distribution over orders on the nodes of the tree.
-% do we need to define "linearize"?
-This process is visualized in Figure~\ref{fig:process-visualization}.
-
-Next we define how a word order grammar linearizes trees in terms of its parameters $\theta$, and then we describe how $\theta$ can be set to optimize objective functions related to communicative efficiency.
-
-\begin{figure}
-\includegraphics[width=0.48\textwidth]{figures/process-visualization.png}
-\caption{Our method for constructing counterfactual versions of languages: We extract unordered dependency tree topologies from dependency corpora, and apply word order grammars to obtain counterfactually reordered corpora.}\label{fig:process-visualization}
-\end{figure}
-
-We think of it as the following problem:
-
-A speaker has a tree in mind, and they have to produce an utterance.
-
-%Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a word order grammar.
-Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
-%\begin{equation}
-%\label{eq:exp-over-utts}
-%J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim L_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
-%\end{equation}
-%where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
-In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
-%Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
-%
-%
-
-All natural languages have some degree of word order regularity.
-Thus, it is not sufficient to optimize the word orders of individual sentences in the corpora---instead, we will optimize the word order rules of entire languages.
-We develop a method to create word order rules that optimize given processing measures on the resulting counterfactual corpus.
-
-\subsection{DLM Predicts Universals}
-
-We first apply our method to show that DLM predicts the correlations.
-
-%Dependency length has been proposed as a metric of processing difficulty on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, and minimization of dependency length is known to recover a large subset of the Greenbergian universals \cite{hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic}.
-%We demonstrate that the minimization of dependency length can be formalized in our framework, reproducing the finding that optimizing grammars for dependency length recovers several word order universals \cite{gildea2010grammars}.
-%
-%To quantify dependency length, we sum the lengths of syntactic dependencies in the sentence \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
-%\begin{equation}
-%    \label{eq:depl}
-%	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
-%\end{equation}
-%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
-Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
-
-\subsection{Efficiency Predicts Universals}
-
-DLM is stipulative and at most a heuristic approximation to psycholinguistic findings
-
-We now introduce a general information-theoretic notion of language efficiency.
-
-use the general speaker-listener picture
-
-\paragraph{Predictability}
-Predictability represents the incremental difficulty of online language processing.
-We formalize predictability as the \key{surprisal} (negative log probability) of each word given its preceding context \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
-Optimization of predictability has been proposed as an explanation for certain word order patterns by \cite{ferrericancho2017optimality}.
-
-We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
-\begin{equation*}
-R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
-\end{equation*}
-%where the language model parameters $\phi$ are chosen to predict the language induced by tree distribution $T$ and word order grammar $\L_\theta$ as accurately as possible (maximizing predictability):
-%\begin{equation*}
-%	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-%\end{equation*}
-We implement the language model $P_\phi$ using an LSTM neural network architecture \cite{hochreiter1997long}, currently the state of the art for this task \cite{goldberg2017neural}.
-%Note that the language model parameters $\phi$ and the word order grammar parameters $\theta$ are recursively dependent on each other.
-See the SI for details on how $\theta$ and $\phi$ are jointly optimized, as well as on the neural network architecture.
-
-\paragraph{Parseability}
-Parseability is the extent to which the syntactic and semantic relationships among words can be recovered from the string of words. 
-%Some notion of parseability has been proposed informally as a desirable property of languages in the linguistic literature going back at least to the 19th century \cite{humboldt,sapir,jesperson}, and parseability has been offered as an underlying justification for syntactic locality as a cost function for language processing \cite{hawkins-performance-1994}.
-%At a basic level, a language can encode information about syntactic relations into fixed word order patterns or into morphological marking on individual words.
-Much linguistic work has been devoted to investigating the trade-offs of various ways of indicating syntactic relations \cite{}.
-In this work, we present what is to our knowledge the first fully implemented computational formalization of this idea.
-We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
-Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
-\begin{align}
-\nonumber
-R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
-\nonumber
-&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
-\end{align}
-%where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
-%\begin{equation*}
-%\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-%\end{equation*}
-We implement $P_\psi$ using a highly generic graph-based parsing architecture \cite{kiperwasser2016simple,dozat2017stanford} which recovers graph structures from word strings by solving a minimal spanning tree problem.
-%Due to data sparsity, we train parsers to predict dependency trees from POS tags alone, not from full wordforms, therefore we do not capture the extent to which morphological wordforms are informative about dependency tree structure.
-See the SI (appendix section 6-7) for further details on the joint optimization of $\theta$ and $\psi$ and on the neural parser architecture, and see SI (section 8) for robustness to different parsing models.
-
-\paragraph{Efficiency}
-The two scoring functions---predictability and parseability---can be combined into an overall equation of the form:
-\begin{equation}
-\label{eq:combined-score}
-	R_{\textit{Eff}} = R_{Pars} + \lambda R_{Pred},
-\end{equation}
-with an interpolation weight $\lambda \in [0,1)$.
-In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification).
-This combined equation represents efficiency: bits of information transferred about dependency trees minus the cost of the transmission.
-In expectation over utterances (substituting Eq.~\ref{eq:combined-score} into Eq.~\ref{eq:exp-over-utts}), the overall objective function to be maximized by a word order grammar can be written in information-theoretic terms as:
-\begin{equation}
-    \label{eq:efficiency}
-    J_T(\theta) = I[L_\theta(T);T] - \lambda H[L_\theta(T)],
-\end{equation}
-where $I[L_\theta(T);T]$ is the \key{mutual information} between strings and trees, and $H[L_\theta(T)]$ is the \key{entropy} of strings \cite{cover2006elements}.
-Maximizing Eq.~\ref{eq:efficiency} is equivalent to maximizing mutual information between strings and trees subject to a constraint on the maximum entropy of strings.
-
-The two terms of Eq.~\ref{eq:efficiency} implement two opposing pressures.
-Parseability (the mutual information term) pushes to make strings more heterogeneous, so that they can indicate different tree structures unambiguously.
-Predictability (the entropy term) pushes to make strings more homogeneous, to reduce the complexity of using the language.
-The idea that natural language arises from a tension between these two forces goes back over a century \cite{gabelentz1901sprachwissenschaft}.
-Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking unordered dependency trees $T$ as the underlying meanings to be conveyed.
-Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
-
-%In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-\section{Results}
-
-\subsection{Relative efficiency of languages}
-\label{sec:relative-efficiency}
-
-We first demonstrate that real languages are relatively efficient compared to random baselines.
-For each language, we generated ten baseline word order grammars for the language by choosing all word order grammar parameters randomly at uniform from $[0,1]$.
-Figure~\ref{fig:pareto-plane} shows the predictability and parseability for each real language relative to its baseline grammars.
-In order to control for limitations due to our word order grammar formalism, we represent real languages in the figure by maximum likelihood fits of word order grammars to the real language data.
-%; in general these maximum-likelihood word order grammars end up with lower predictability and parseability than the real languages. % TODO verify this is true? TODO awkward sentence
-For the calculation of predictability and parseability, we make all (baseline and real) word order grammars deterministic by always choosing the highest-probability linearization of each tree; by making the grammars deterministic in this way we eliminate an anticonservative bias toward low predictability in the baseline languages, which are highly nondeterministic.
-The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
-
-\begin{figure}
-    \centering
-    \includegraphics[scale=.45]{../results/plane/pareto-plane-iso-best-balanced-legend-viz.pdf}
-    \caption{Predictability and parseability of 51 UD languages (red), indicated by ISO codes, compared to ten baseline word order grammars per language (green).
-Predictability and parseability scores are $z$-scored within language.
-Each point for a real language has a line pointing in the direction of the center of mass of its baselines.
-The green contour shows the density of baseline languages.
-Unlabeled dots represent the centroid for real languages (red), baseline languages (green), and languages optimized for predictability (yellow), parseability (pink), efficiency (blue), and dependency length (red).
-When a language is to the bottom--left of its baselines, this indicates that it is relatively optimal for efficiency.} % TODO colors, fontsize, etc could conceivably be made prettier
-    \label{fig:pareto-plane}
-\end{figure}
-
-Figure~\ref{fig:pareto-plane} also shows the average position of optimized languages.
-Languages appear to be attracted toward these points and away from the region of the baseline languages.
-We also see that several languages actually end up \emph{more} efficient than the computationally optimized languages.
-% TODO add something about dependency length-optimized languages?
-
-These results are shown in more detail for two typologically distinct languages (English and Japanese) in Figure~\ref{fig:predict-pars}.
-Here we include also the predictability and parseability of the original language data and of the word order grammars optimized for dependency length.
-We see that real languages are even more predictable than any optimized or baseline language, while optimization for either parseability or dependency length suffices to raise parseability to the level of the real languages.
-This result is in line with the finding that languages with longer dependencies are harder to parse in general \cite{gulordava2016multilingual}.
-
-\begin{figure*}
-\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
-	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right).
-We provide the entropy $H[L_\theta(T)]$ for predictability (lower is better) and the conditional entropy $H[T|L_\theta(T)]$ for parseability (lower is better).
-Both measures are normalized by sentence length.}\label{fig:predict-pars}
-\end{figure*}
-
-Finally we demonstrate the relationship between dependency length minimization and the maximization of efficiency.
-Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency.
-We see that optimizing for efficiency lowers dependency length relative to random baselines, in keeping with the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}.
-In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the depenndency length of the real language; in 3/4 of the lanuages shown, the real language is best matched by efficiency optimization.
-
-\begin{figure*} % need Efficiency and (maybe) MLE langs on here
-\includegraphics[width=0.98\textwidth]{../results/dependency-length/figures/depLength-facet.pdf}
-\caption{Average dependency length as a function of sentence length in four languages.
-Across languages, real and optimized languages have shorter dependencies than random baseline orderings.}\label{fig:deplength}
-\end{figure*}
-
-\subsection{Greenbergian Word Order Correlations}
-\label{sec:greenberg}
-\begin{table*}
-	\begin{center}
-\begin{tabular}{|ll|l|l|l|ll|l|}
-	\hline
-	%&
-	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  & \multirow{2}{*}{Pred}   &  \multirow{2}{*}{Pars} &  \multirow{2}{*}{Efficiency}  \\ 
-	verb & object     & & & &  & \\ \hline \hline % \textsc{verb} $\xrightarrow{obj}$ \textsc{noun}
-%adp. &
-adposition    &    NP       &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
-copula    &    NP        &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
-auxiliary    &    VP       &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
-noun    &    genitive      &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
-noun    &    relative clause       &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
-complementizer    &    S        &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
-verb    &    PP     &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
-want    &    VP      &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
-verb    &    subject    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
-verb    &    manner adverb     &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
- \hline
-%comp. &
-%\multirow{3}{*}{NP}&
-%	AP &
-    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
-\end{tabular}
-	\end{center}
-	\caption{Greenbergian Correlations.
-Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
-%	For clarity, we organize the correlations by the category of the `verb patterner'.
-	For each correlation, we give our operationalization in terms of UD.
-%A rightwards arrow indicates that the dependency is predicted to have the same direction as the \textit{obj} dependency; a leftwards arrow indicates an inverse correlation.
-	For each correlation we report what percentage of the languages in our sample satisfied it (`Real').
-	We then report, for each correlation and each objective function, how many (in \%) of the optimized grammars satisfy the correlation, with the significance level in a logistic mixed-effects analysis across language families.}\label{table:results-dryer}
-\end{table*}
-
-%\begin{table*}
-%\begin{tabular}{lllllllllllllllllll}
-%	&	          & \multicolumn{2}{c}{Dep. Length}      & \multicolumn{2}{c}{Lang. Mod.}              &  \multicolumn{2}{c}{Parsing}  &\multicolumn{2}{c}{Operationalization} & & GU\\ \hline
-%	Adp &	Adp+Noun &                  & \cmark       &             &            &         &             & \textsc{noun} $\xrightarrow{case}$ \textsc{adp} & - &  3, 4\\ \hline
-%	Verb &	Verb+PP & \checkmark        & \cmark      & \checkmark  &   \cmark         &    &             & \textsc{verb} $\xrightarrow{obl}$ \textsc{noun}  & + & 7  \\
-%	Verb &	Copula & \checkmark         &  \cmark     &  \checkmark &   \cmark    &  \checkmark  &     \cmark        & \textsc{noun} $\xrightarrow{cop}$  \textsc{aux} & - \\
-%	Verb &	want & \checkmark           & \cmark      & \checkmark  & \cmark     &  \checkmark  &    \cmark         & \textsc{verb} $\xrightarrow{xcomp}$ \textsc{verb} & + & 15 \\
-%	Verb &	subject &                   &       & \checkmark  &      \cmark      & \xmark   &    \xmark         & \textsc{verb} $\xrightarrow{nsubj}$ \textsc{noun} & + \\
-%	Verb &	auxiliary & \cmark          &       & \cmark      & \cmark     &           &             & \textsc{verb}  $\xrightarrow{aux}$ \textsc{verb}/\textsc{aux}  & - & 16\\
-%	verb & manner adverb &              &  -     &  \xmark     &   \xmark     &          &             & \textsc{verb} $\xrightarrow{advmod}$ ADV  & + & 7 \\ \hline
-%	Comp &	comp; adv sub. &            &   \cmark    &             &   \cmark         &          &             & \textsc{verb} $\xrightarrow{mark}$ \textsc{sconj} & - & \\ \hline
-%	Noun &	article N' &                &   \xmark/\xmark    &             &    -/\cmark        &          &   -/\cmark          & \textsc{noun} $\xrightarrow{det}$ \textsc{det}/\textsc{pron}  & -\\
-%	Noun &	noun+genitive & \cmark      &   \cmark    &             &            &              &    \cmark  & \textsc{noun} $\xrightarrow{nmod}$ \textsc{noun}  & + & 2\\
-%	Noun &	noun+relative clse & \cmark &  \cmark   &               &            &  \checkmark &      \cmark       & \textsc{noun} $\xrightarrow{acl}$ \textsc{verb} & + & 24 \\ \hline
-%	Adj &	adj+std            &        &  \cmark   &               &            &             &             &  ADJ $\xrightarrow{nmod}$ \textsc{adp}  & + & 23\\
-%\end{tabular}
-%	\caption{Greenbergian Correlations, as provided by \cite{dryer-greenbergian-1992}.
-For each correlation, we give our operationalization in terms of UD and whether the direction of this dependency correlates positively ($+$) or negatively ($-$) with the verb-object-dependency.
-For each correlation and each objective function, we indicate whether the correlation is predicted correctly (\cmark), in the wrong direction (\xmark), or no prediction is made (blank).
-For each objective function, we show results for plain UD (left column) and UD with function word heads (right column).
-}\label{table:results-dryer}
-%\end{table*}
-%%, with corresponding numbers in the original Greenberg universals, where applicable.
-
-%In the last sections, we confirmed the effectiveness of our optimization method and elucidated the relation between the three optimization functions on the level of .
-
-We now examine to what extent we can recover Greenberg's word order correlations in optimized grammars. 
-Dryer \cite{dryer1992greenbergian} presents a comprehensive updated version of the word order correlations, drawing on 625 languages, which we take as the basis of our evaluation.
-In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
-Most of them can be straightforwardly implemented in UD, allowing us to check which correlations a word order grammar satisfies.
-%This dependency corresponds to dependencies of the form $\xrightarrow{obj}$ in UD.
-%Similarly, most of the other dependencies in Dryer's formulation can be formalized:
-%for instance, the dependency between nouns and relative clauses can be formalized as $\xrightarrow{acl}$.
-%Testing whether an ordering model satisfies the correlation between relative clauses and objects reduces to checking whether these two relation types have direction parameters $a_\tau$ that are either both $>0.5$ (both kinds of dependents precede their heads) or both $< 0.5$ (they both follow their heads).
-
-Dryer \cite{dryer1992greenbergian} presents three correlations which do not correspond to dependencies annotated in UD: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
-Two pairs of Dryer's correlations, namely those for the dependencies between complementizers and adverbial subordinators and their complement clauses,and those for the dependencies between verbs and adpositional phrases, and adjectives and their standard of comparison, had to be collapsed into two correlations in UD.
-From Dryer's 15 correlations, we obtained 10 formalized correlations, roughly covering nine of Greenberg's original universals.
-%TODO update to 10?
-What happened to the 11th?
-Was that amod?
-[The 11th was adjective+standard of comparison, which is merged into obl] Maybe we could include amod as a "true rejection" of efficiency, but a false alarm for DepL
-
-In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
-We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau$ show the same direction for the correlating dependency and for the verb-object dependency, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
-%We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau > 0.5$ holds for the correlating dependency depending on whether $a_\tau > 0.5$ for the verb-object dependency, with per-language-family random intercept and slope.\footnote{We coded language families according to \url{universaldependencies.org}.}
-We are interested in the direction and significance of this effect: 
-If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.
-
-We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
-To do evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
-The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS.
-%in all except four languages, for WALS did not have data.
-
-\paragraph{Results}
-Results are shown in Table~\ref{table:results-dryer}.
-All correlations but two are confirmed in the models estimated from the real orderings.
-The exceptions are the subject--verb dependency and the verb--adverb dependency, which typically go in the opposite direction from the standard description.
-We will discuss these exceptions further below.
-
-In keeping with previous work, we see that optimizing for dependency length correctly accounts for nine word order correlations, missing only the verb--adverb dependency. 
-Predictability and parseability predict five and seven correlations, respectively, making largely complementary predictions.
-Efficiency significantly predicts all the word order correlations, each in the same direction as attested in the dependency corpora.
-
-We now address the two word order correlations whose direction in the dependency corpora is opposite from what would be expected in the typological literature.
-The first is the correlation of the order of verb--subject and verb--object dependencies.
-Our sample of mainly European languages highly over-represents languages with the general order subject--verb--object (such as English), in which the order of the verb--subject and verb--object dependencies are anti-correlated.
-Surprisingly, given the sample of tree structures of these languages, it turns out that the optimal languages tend to have anti-correlated orders for subjects and objects order similar to the real languages.
-
-The second anomalous dependency is the verb--manner adverb dependency.
-We believe the anti-correlation in the UD corpora arises because the \emph{advmod} dependency does not distinguish between manner adverbs---the subject of the typological judgment---and various other types of modifiers such as sentence-level adverbs.
-Nevertheless, the languages optimized for efficiency reproduce the anti-correlation of the orders of verb--object and verb--adverb at around the same rate as the real languages.
-
-We further evaluate the word order predictions of efficiency, showing that efficiency is most successful in predicting correlations in the direction found in the UD corpora.
-We constructed a single logistic model predicting, for each of the ten dependencies, whether it is correlated or anti-correlated with the \emph{obj} dependency in languages optimized for efficiency, with random effects for language and language family, correlated across the ten dependencies.
-We conducted the same analysis for predictability, parseability, and dependency length.
-We used this model to estimate the posterior distribution of the number of correlations that an objective function predicts to be in the same direction as found in the UD treebanks.
-The resulting distributions are shown in Figure~\ref{fig:posterior-satisfied-count}.
-The estimated posterior probability that efficiency predicts less than all ten dependencies to correlate in the same direction as in the UD treebanks is $0.0242$.
-The probability that it predicts less than nine of the correlations is $3\cdot 10^{-4}$.
-For dependency length, the posterior puts much of the probability mass on predicting only nine of the correlations; predictability and parseability predict significantly less correlations.
-
-\begin{figure}
-    \centering
-%     \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-efficiency.pdf}
- %    \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-langmod.pdf}
-  %   \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-parseability.pdf}
-   %  \includegraphics[scale=.2]{../results/correlations/figures/posterior-satisfied-universals-depl.pdf}
-     \includegraphics[scale=.45]{../results/correlations/figures/posterior-satisfied-universals-together.pdf}
-
-	\caption{Posterior of the number of correlations predicted in the direction found in the UD treebanks, computed from a mixed-effects logistic regression jointly modeling all ten dependencies.
-Efficiency predicts all ten correlations with high posterior probability.}
-    \label{fig:posterior-satisfied-count}
-\end{figure}
-
-\subsection{Discussion}
-
-In Section~\ref{sec:relative-efficiency}, we found that word order grammars in the majority of UD languages have better efficiency than baseline word order grammars.
-Furthermore, in Section~\ref{sec:greenberg} we found that explicitly optimizing grammars for efficiency reproduces 10 major word order correlations reported in the typological literature, and with greater accuracy than the typological literature when it comes to the languages studied.
-
-%While the efficiency objective succeeds in predicting the presence of the most word order correlations, the dependency length objective seems to predict a greater prevalence for those correlations that it does predict.
-We believe this discrepancy is a result of noise in the optimization process.
-Optimizing for efficiency is computationally challenging, requiring stochastic gradient descent on word order grammar, language model, and parser parameters, while optimizing for dependency length is simpler and thus more optimal grammars can be reached easily (See SI section 2 for supporting data).
-Inasmuch as dependency length minimization is computationally easier to implement than efficiency maximization, we believe it will remain useful as an easy-to-use heuristic for predicting and explaining word order universals.
-% TODO we can also report (in the SI) prevalences among the best-performing of the optimized languages, i.e., those where optimization was most successful.
-I would expect that this should narrow the gap between DepL and Efficiency, which would confirm that the gap is really about the difficulty of optimization, not about the objective functions.
-
-\section{Conclusion}
-
-We found that a large subset of the Greenbergian word order correlations can be explained in terms of optimization of grammars for efficient communication, as defined by information theoretic criteria and implemented using state-of-the-art machine learning methods.
-We defined a space of word order grammars, as well as objective functions reflecting communicative efficiency, and found that the word order grammars that maximize communicative efficiency reproduce the word order universals.
-Beyond our present results, we provide a complete formalization and computational framework in which theories of the functional optimization of languages can be tested.
-Other objective functions could be hypothesized and tested within our framework; furthermore, future advances in machine learning will enable the optimization of richer and richer models of grammar.
-
-A major question for functional explanations for linguistic universals is: \emph{how} do languages come to be optimized?
-Do speakers actively seek out new communicative conventions that allow better efficiency?
-Or do languages change in response to biases that come into play during language acquisition \cite{fedzechkina2012language,culbertson2012learning}?
-Our work is neutral toward such questions.
-To the extent that language universals arise from biases in learning or in the representational capacity of the human brain, our results suggest that those biases tilt toward communicative efficiency.
-Our work does provide against the idea that word order universals are best explained in terms of learning biases that are irreducibly arbitrary and genetic in nature \cite{chomsky2010some}.
-
-While our work has shown that certain word order universals can be explained by efficiency in communication, we have made a number of basic assumptions about how language works in constructing our word order grammars: for example, that sentences can be syntactically analyzed into dependency trees.
-The question arises of whether these more basic properties themselves might be explainable in terms of efficient communication.
-Relatedly, there are many remaining word order universals not captured by our model, such as the trade-off of rich morphological marking and flexibility in word order \cite{jespersen1922,mcfadden2003morphological,futrell2015quantifying}.
-Our models cannot capture this trade-off for technical reasons: the parseability objective does not operate over wordforms, only POS tags, and so it does not take advantage of morphological cues to syntactic structure.
-Future work can investigate whether these and other remaining universals can be explained using more sophisticated models of how meaning representations are transduced into strings of words, based on larger databases with richer annotations.
-
-\matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
-We use all languages for which at least one treebank with a training partition was available, a total of 50 languages.
-For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.
-Punctuation was removed.
-%We removed punctuation. %, which we operationalized by removing all dependents of \emph{punct} dependencies, and recursively removing their dependents.
-
-Universal dependencies represents as dependents some words that are typically classified as heads in syntactic theory.
-This particularly applies to the \textit{cc}, \textit{case}, \textit{cop}, and \textit{mark} dependencies.
-Following prior work studying dependency length minimization \cite{futrell2015largescale}, we modified each treebank by inverting these dependencies, promoting the dependent to the head position.
-We report results on this modified version of UD.
-
-The efficiency optimization results from Table~\ref{tab:results-dryer} were preregistered: \url{http://aspredicted.org/blind.php?x=8gp2bt}.
-
-See the SI for details on the neural language models and parsers used, and on the optimization procedures.}
-
-\subsection{Specification of word order grammars}
-
-All natural languages have some degree of word order regularity.
-Thus, it is not sufficient to optimize the word orders of individual sentences in the corpora---instead, we will optimize the word order rules of entire languages.
-
-Our goal is to specify word order grammars that transduce unordered dependency trees into strings of words.
-We have three additional goals: first, we would like our grammars to be able to operate over dependency trees as specified in large human-annotated treebank corpora, such as those developed by the Universal Dependencies project \cite{nivre2017universal}.
-Second, we would like our grammars to be simple and easily interpretable, so we can determine whether they exhibit word order universals.
-Third, we would like our grammars to be differentiable, so that it is easy to optimize their parameters for efficiency.
-
-Our model linearizes dependency trees such as the ones in Figure~\ref{fig:arabic-japanese}.
-We assume nodes are labeled with part-of-speech (POS) tags, and arcs are labeled with syntactic relations.
-The parameters of a word order grammar are as follows.
-For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
-Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
-Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p.
-184) without replacement.
-
-This simple parameterization lets us directly test whether a given grammar follows word order universals.
-For example, we can test whether a Greenbergian word order correlation holds in a grammar by checking whether $a_\tau>\frac{1}{2}$ for different syntactic relations $\tau$.
-
-\subsection{Objective functions}
-\label{ref:objective-funcs}
-
-Let $T$ be a distribution over unordered dependency trees and let $L_\theta$ be a word order grammar.
-Our goal is to find parameters $\theta$ which maximize an objective function $J$ in expectation over utterances:
-\begin{equation}
-\label{eq:exp-over-utts}
-J_T(\theta) = \E_{t \sim T} \E_{\mathbf{w} \sim L_\theta(t)} [R(t; \mathbf{w})], % TODO make the expectations look better!!
-\end{equation}
-where the function $R$ specifies a score for an individual utterance $\mathbf{w}$ expressing a tree $t$.
-In practice, $T$ will be the empirical distribution over unordered trees observed in a dependency treebank.
-Below we discuss possible values of $R$, which may implement predictability, parseability, or a weighted combination of the two, representing overall efficiency.
-
-\paragraph{Predictability}
-Predictability represents the incremental difficulty of online language processing.
-We formalize predictability as the \key{surprisal} (negative log probability) of each word given its preceding context \cite{hale2001probabilistic,levy2008expectation,smith2013effect}.
-Optimization of predictability has been proposed as an explanation for certain word order patterns by \cite{ferrericancho2017optimality}.
-
-We formalize predictability as follows. Given a language model $P_\phi$ with parameters $\phi$, the log probability of a sentence ${\bf w}$ is:
-\begin{equation*}
-R_{Pred}^\phi(t; {\bf w}) = \sum_{i=1}^{\#{\bf w}} \log P_\phi(w_{i+1}|w_{1\dots i}),
-\end{equation*}
-where the language model parameters $\phi$ are chosen to predict the language induced by tree distribution $T$ and word order grammar $\L_\theta$ as accurately as possible (maximizing predictability):
-\begin{equation*}
-	\phi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R^\phi_{Pred}({\bf w}, \phi) ].
-\end{equation*}
-We implement the language model $P_\phi$ using an LSTM neural network architecture \cite{hochreiter1997long}, currently the state of the art for this task \cite{goldberg2017neural}.
-Note that the language model parameters $\phi$ and the word order grammar parameters $\theta$ are recursively dependent on each other.
-See the SI for details on how $\theta$ and $\phi$ are jointly optimized, as well as on the neural network architecture.
-
-\paragraph{Parseability}
-Parseability is the extent to which the syntactic and semantic relationships among words can be recovered from the string of words. 
-%Some notion of parseability has been proposed informally as a desirable property of languages in the linguistic literature going back at least to the 19th century \cite{humboldt,sapir,jesperson}, and parseability has been offered as an underlying justification for syntactic locality as a cost function for language processing \cite{hawkins-performance-1994}.
-%At a basic level, a language can encode information about syntactic relations into fixed word order patterns or into morphological marking on individual words.
-Much linguistic work has been devoted to investigating the trade-offs of various ways of indicating syntactic relations \cite{}.
-In this work, we present what is to our knowledge the first fully implemented computational formalization of this idea.
-We specify parseability as the mutual information between strings and trees, or equivalently the quantity of information provided by strings about trees.
-Formally, we say the parseability score of an utterance $\mathbf{w}$ is the pointwise mutual information of the $\mathbf{w}$ and the tree $t$ it was generated from, calculated using a parser with parameters $\psi$:\footnote{Note that $P(t)$ is a constant term which will be ignored in the optimization process.}
-\begin{align}
-\nonumber
-R_{Pars}^\psi(t; {\bf w}) &= \log \frac{P_\psi (t|\mathbf{w})}{P(t)} \\
-\nonumber
-&= \displaystyle \sum_{i=1}^{\#\mathbf{w}} \log P_\psi\left(\begin{array}{@{}l@{}}head(w_{i}) \\ label(w_i)\end{array}{\big |}\textbf{w}\right) - \log P(t),
-\end{align}
-where the parser parameters $\psi$ are chosen to maximize parsing accuracy:
-\begin{equation*}
-\psi(\theta) = \argmax_\phi \displaystyle \E_{t \sim T} \E_{{\bf w} \sim L_\theta(t)} [ R_{Pars}({\bf w}, \psi) ].
-\end{equation*}
-We implement $P_\psi$ using a highly generic graph-based parsing architecture \cite{kiperwasser2016simple,dozat2017stanford} which recovers graph structures from word strings by solving a minimal spanning tree problem.
-Due to data sparsity, we train parsers to predict dependency trees from POS tags alone, not from full wordforms, therefore we do not capture the extent to which morphological wordforms are informative about dependency tree structure.
-
-See the SI (appendix section 6-7) for further details on the joint optimization of $\theta$ and $\psi$ and on the neural parser architecture, and see SI (section 8) for robustness to different parsing models.
-
-\paragraph{Efficiency}
-The two scoring functions---predictability and parseability---can be combined into an overall equation of the form:
-\begin{equation}
-\label{eq:combined-score}
-R_{Efficiency} = R_{Pars} + \lambda R_{Pred},
-\end{equation}
-with an interpolation weight $\lambda \in [0,1)$.
-In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification).
-This combined equation represents efficiency: bits of information transferred about dependency trees minus the cost of the transmission.
-In expectation over utterances (substituting Eq.~\ref{eq:combined-score} into Eq.~\ref{eq:exp-over-utts}), the overall objective function to be maximized by a word order grammar can be written in information-theoretic terms as:
-\begin{equation}
-    \label{eq:efficiency}
-    J_T(\theta) = I[L_\theta(T);T] - \lambda H[L_\theta(T)],
-\end{equation}
-where $I[L_\theta(T);T]$ is the \key{mutual information} between strings and trees, and $H[L_\theta(T)]$ is the \key{entropy} of strings \cite{cover2006elements}.
-Maximizing Eq.~\ref{eq:efficiency} is equivalent to maximizing mutual information between strings and trees subject to a constraint on the maximum entropy of strings.
-
-The two terms of Eq.~\ref{eq:efficiency} implement two opposing pressures.
-Parseability (the mutual information term) pushes to make strings more heterogeneous, so that they can indicate different tree structures unambiguously.
-Predictability (the entropy term) pushes to make strings more homogeneous, to reduce the complexity of using the language.
-The idea that natural language arises from a tension between these two forces goes back over a century \cite{gabelentz1901sprachwissenschaft}.
-Eq.~\ref{eq:efficiency} is a special case of the objective function proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems, taking unordered dependency trees $T$ as the underlying meanings to be conveyed.
-Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship).
-
-\paragraph{Dependency length}
-
-In previous studies of word order optimization, it has been common to propose that word orders are chosen to minimize \key{dependency length}, the linear distance between words linked in head--dependent relationships \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
-Dependency length has been proposed as a metric of processing difficulty on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, and minimization of dependency length is known to recover a large subset of the Greenbergian universals \cite{hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic}.
-We demonstrate that the minimization of dependency length can be formalized in our framework, reproducing the finding that optimizing grammars for dependency length recovers several word order universals \cite{gildea2010grammars}.
-
-To quantify dependency length, we sum the lengths of syntactic dependencies in the sentence \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
-\begin{equation}
-    \label{eq:depl}
-	R_{DepL}(t; {\bf w}) = -\sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
-\end{equation}
-where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost. 
-Dependency length quantified in this manner is a heuristic measure of complexity: the actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}.
-
-\subsection{Discussion and Related Work}
-
-Our work builds on previous work on optimizing word order grammars for various objectives \cite{gildea2007optimizing,gildea2010grammars,gildea2015human}.
-Previous work has used deterministic word order grammars optimized by simple hill-climbing. 
-Our work departs from previous work in three ways. 
-First, we use word order grammars that are probabilistic and differentiable, enabling greater expressivity and optimization by stochastic gradient descent.
-Second, we use state-of-the-art neural language models for predictability, jointly optimizing parameters of word order grammars and language models.
-Third, we implement parseability and efficiency as a novel objective functions, which would not be feasible with non-differentiable word order grammars.
-
-In the space of word order grammars, more complex models than ours have been defined \cite{belz2011first,futrell2015experiments,wang2016galactic}.
-However, the simple parametric form of our model enables easy interpretation across languages.
-Like these previous models, our model makes simplifying assumptions and will not fit the rules of natural languages perfectly.
-For instance, none of the models accounts for word order phenomena that are conditioned on the larger context---e.g., differences in word order between embedded and main clauses. 
-Also like previous models, our model only generates word orders that are \key{projective}, meaning that dependency lines do not cross.
-There are two reasons for constraint.
-First, there is currently no easily interpretable probabilistic model of non-projective word ordering to build on from the natural language processing literature (cf.
-\cite{bohnet2012generating}).
-Second, natural language dependency trees are overwhelmingly projective \cite{kuhlmann2013mildly}.
-Note that the preponderance of projectivity in natural language is itself a linguistic universal, equivalent to context-freeness in formal language theory \cite{marcus1965notion,kuhlmann2013mildly}, and explanations have been offered for it in terms of dependency length minimization  \cite{ferrericancho2006syntactic}.
-
-We test efficiency metrics based both on information theory (Eq.~\ref{eq:efficiency}) and dependency length minimization (Eq.~\ref{eq:depl}).
-We do not see these as competing theories.
-Rather, we see the information-theoretic objective as a broad description of efficiency in general, and dependency length as a heuristic description of a particular component of complexity.
-There is reason to believe that dependency length minimization might emerge as a byproduct of optimization for parseability and predictability under resource constraints \cite{futrell2017memory}: in particular, short dependencies are optimal for predictability given lossily compressed representations of context \cite{futrell2017noisycontext,futrell2017generalizing}.
-In the results below, we demonstrate that optimizing for our Eq.~\ref{eq:efficiency} does indeed produce dependency length minimization as a byproduct.
-
+The efficiency optimization results from Table~\ref{table:corr-resu} were preregistered: \url{ http://aspredicted.org/blind.php?x=ya4qf8} (see also SI Section S4.6). 
 }
@@
 
-\acknow{We thank Ted Gibson, Michael C. Frank, Judith Degen, Chris Manning, and audiences at CAMP 2018 for helpful discussion.}
+\acknow{We thank Ted Gibson, Michael C.
+Frank, Judith Degen, Chris Manning, Paul Kiparsky, and audiences at CAMP 2018 and CUNY 2019 for helpful discussion.
+We also thank the reviewers and the editor for their helpful comments.}
 
@@
 
-%\bibliographystyle{acl_natbib}
 %\bibliography{extracted}
-\bibliography{everything}
+\bibliography{paper-revised}
 

### 2019-11-11 writeup/table2.tex
code

% https://tex.stackexchange.com/questions/171982/how-to-make-table-floats-work-with-preview-standalone-environment/171983

\documentclass[%
    ,float=false % this is the new default and can be left away.
    ,preview=true
    ,class=scrartcl
%    ,fontsize=20pt
    ]{standalone}

\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}

\usepackage{xcite}
\externalcitedocument{pnas-word-order-si}

\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage{siunitx}

\usepackage{longtable}

\usepackage[numbers]{natbib}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{arydshln}

\newcommand{\key}[1]{\textbf{#1}}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}

\usepackage{mathtools}
\DeclarePairedDelimiter{\ceiling}{\lceil}{\rceil}

\begin{document}
\begin{tabular}{|c|ll|c|cc|ccc}
	\hline
	%&
	&	\multicolumn{2}{c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Optimized} & \\ 
	&	verb & object     & & &   \\ 
	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    noun phrase       
				&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Real_lifted_case.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Efficiency_lifted_case.pdf}     } & \\
	&		\emph{to}            & \emph{a friend} &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    noun phrase         
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Real_lifted_cop.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Efficiency_lifted_cop.pdf}     } & \\
	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    verb phrase       
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Real_aux.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Efficiency_aux.pdf}     } & \\
	&	\emph{has}          & \emph{written}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Real_nmod.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Efficiency_nmod.pdf}     } & \\
	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Real_acl.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Efficiency_acl.pdf}     } & \\
	&	\emph{books} & \emph{that you read}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    sentence        
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Real_lifted_mark.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Efficiency_lifted_mark.pdf}     } & \\
	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    adp. phrase         
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Real_obl.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Efficiency_obl.pdf}     } & \\
	&	\emph{went} & \emph{to school}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    verb phrase        
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Real_xcomp.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{figures/posterior_Efficiency_xcomp.pdf}     } & \\
	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
\end{tabular}
\end{document}

### 2019-11-11 writeup/prepareRevisedSubmission.sh

@@ -1,11 +1,11 @@
-cp 3-paper-names.tex revised-submission/
-cp ../results/plane/pareto-plane-iso-best-balanced-legend-viz-10-fontsize_pareto_SD.pdf revised-submission/figures/
-cp ../results/correlations/figures/correlations-curve-whiteaxis-bw.pdf revised-submission/figures/
-cp ../results/correlations/figures/posteriors/posterior_Real_*.pdf revised-submission/figures/
-cp ../results/correlations/figures/posteriors/posterior_Efficiency_*.pdf revised-submission/figures/
-cp everything_mainPaper.bib revised-submission/
-cp figures/corr-full.png revised-submission/figures/
-cp figures/grammar-langs-fontsize-friends-rect.png revised-submission/figures/
+cp 3-paper-names.tex revised-submission/paper-revised.tex
+cp ../results/plane/pareto-plane-iso-best-balanced-legend-viz-10-fontsize_pareto_SD.pdf revised-submission/figure4.pdf
+cp ../results/correlations/figures/correlations-curve-whiteaxis-bw.pdf revised-submission/figure5.pdf
+cp paper-revised.bib revised-submission/
 cp pnas-new.cls revised-submission/
 cp pnas-new.bst revised-submission/
 cp pnasresearcharticle.sty revised-submission/
+cp figures/figure1.pdf revised-submission/
+cp figures/figure3.pdf revised-submission/
+cp table2.pdf revised-submission/
+

### 2019-11-12 writeup/3-paper-names.tex
code

@@
 	\begin{center}
-\begin{tabular}{|c|ll|ll|}
-	\hline
-	&	\multicolumn{2}{c|}{\textbf{Arabic} (English, ...)}   &        \multicolumn{2}{c|}{\textbf{Japanese} (Turkish, ...)} \\ \hline\hline
-	& \multicolumn{2}{c|}{Correlates with...} & \multicolumn{2}{c|}{Correlates with...}  \\
-	&	Verb & Object     & Object & Verb    \\ \hline\hline
-	& kataba & ris{\= a}la 	&tegami-o & kaita \\
-	&	\emph{wrote} & \emph{letter} & 	\emph{letter} & \emph{wrote} \\ \hline 
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    {\d s}ad{\= i}q       &	tomodachi & ni \\ % G3, G4
-	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline 
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&k{\= a}na    &    {\d s}ad{\= i}q         &	tomodachi & datta \\ % not in Greenberg
-	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak- & -udesho \\ % G16, G13 TODO this should only be inflected auxiliaries, replace Arabic
-	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& {\d s}ad{\= i}q  &    John    & 	John no & tomodachi \\ % G2, G23
-	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uh{\= a}       & 	anata-ga yonda & hon \\ % G24
-	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tu{\d s}il        & 	toochaku suru & koto \\ % 
-	&	\emph{that} & \emph{she arrives}  &	\emph{arrives} & \emph{that} \\ \hline
-	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'il{\= a} lmadrasa     & 	gakkoo ni & itta \\ % G22
-	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
-	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'ur{\= i}d    &    'an 'ugh{\= a}dir         & 	ik- & -itai \\ % G15
-	& \emph{wants}   &  \emph{to leave}  &	\emph{to go} & \emph{want} \\ \hline 
-\end{tabular}
+    \includegraphics[width=0.45\textwidth]{table1.pdf}
 	\end{center}
@@
 	Selection is based on a more recent typological study by Dryer~\cite{dryer1992greenbergian}, restricted to those correlations that are annotated in available corpus data.
-See SI Section S1 for more on Greenberg correlations.
+See SI Appendix, Section S1 for more on Greenberg correlations.
 	}\label{table:corr-dryer}
@@
     \centering
-{
-\centering{
-\begin{dependency}[theme = simple]
-   \begin{deptext}[column sep=1em]
-          She \& met \& friends  \\
-   \end{deptext}
-   \depedge{2}{1}{subject}
-   \depedge{2}{3}{object}
-\end{dependency}
-}
-}
+    \includegraphics[width=0.2\textwidth]{figure2.pdf}
 	\caption{An English sentence with annotated syntactic relations.}
@@
 Efficiency is a weighted sum of parseability and predictability.
- See \textit{Materials and Methods} for details, and SI Section 7 for experiments demonstrating that our results are robust to different methods of estimating parseability and predictability.
+See \textit{Materials and Methods} for details, and SI Appendix, Section S7 for experiments demonstrating that our results are robust to different methods of estimating parseability and predictability.
 
@@
 We address this challenge by introducing a simple, differentiable computational formalism for describing grammatical regularities. 
-Our formalism makes it possible to find optimal grammars by standard methods such as stochastic gradient descent (see SI Section S5).
+Our formalism makes it possible to find optimal grammars by standard methods such as stochastic gradient descent (see SI Appendix, Section S5).
 For each grammar, we report predictability and parseability as estimated on the data resulting from ordering the syntactic structures from the corpus according to the grammar.
@@
 including the fact that the orders are purely a function of the tree structure, and do not take into account other factors, such as discourse structure, which are not annotated in the corpora.
-For a comparison of the raw word orders from corpora against appropriate baseline grammars, see SI Section 8.
+For a comparison of the raw word orders from corpora against appropriate baseline grammars, see SI Appendix, Section S8.
 
@@
 90\% of real grammars improve over the baselines in parseability ($p < 0.05$), and 80~\% improve in predictability ($p < 0.05$).
-See SI Section S3 for additional analyses.
+See SI Appendix, Section S3 for additional analyses.
 
@@
 \begin{figure}
-    \centering
+	\begin{center}
 		\includegraphics[width=0.48\textwidth]{figure5.pdf}
+	\end{center}
     \caption{
@@
 We controlled for variation between the syntactic structures used in different languages and language families by entering the language and language family as random effects.
-See SI Section S4.3 for robustness to modeling choices.
+See SI Appendix, Section S4.3 for robustness to modeling choices.
 
-\begin{table}
+\begin{figure}
     \centering
-    \includegraphics[width=0.48\textwidth]{table2.pdf}
+    \includegraphics[width=0.48\textwidth]{figure6.pdf}
 
@@
 }\label{table:corr-resu}
-\end{table}
+\end{figure}
 
-In Table~\ref{table:corr-resu}, we compare the prevalence of the eight correlations in real and optimized languages.
+In Figure~\ref{table:corr-resu}, we compare the prevalence of the eight correlations in real and optimized languages.
 For the real languages, we indicate how many of the 51 languages satisfy a correlation.
@@
 In the multivariate mixed-effects analysis, efficiency optimization predicts all eight correlations to hold across languages (posterior probability $0.9911$). 
-Optimizing for only predicability or only parseability does not predict all of the correlations (See SI Section S4).
+Optimizing for only predicability or only parseability does not predict all of the correlations (See SI Appendix, Section S4).
 
@@
 Our results suggest that both Greenberg's correlations and dependency length minimization might be explainable purely in terms of maximizing the general parseability and predictability of utterances, without a need for further constraints.
-See SI Section S12 for further discussion along with some simulations demonstrating how grammars that satisfy Greenberg's correlations can be more efficient in a generic sense.
+See SI Appendix, Section S12 for further discussion along with some simulations demonstrating how grammars that satisfy Greenberg's correlations can be more efficient in a generic sense.
 
@@
 Unlike crosslinguistic efficiency studies in the domain of lexical semantics~\cite{kemp2012kinship,regier2015word,zaslavsky2018efficient}, we did not derive a single universal bound for the efficiency across all 51 languages in Study 1; instead, we constructed optimized grammars individually for each language.
-Each language $L$ has its own distribution of tree structures that speakers communicate, and different grammars may be optimal for different tree structure distributions (see SI Section S3.5).
+Each language $L$ has its own distribution of tree structures that speakers communicate, and different grammars may be optimal for different tree structure distributions (see SI Appendix, Section S3.5).
 Our results show that the word order of each language $L$ is approximately optimal for the tree structures used in $L$.
@@
 A syntactic structure, consisting of a set of words and syntactic relations between them, is then ordered into a string of words recursively starting from the root; the dependents of a word then are ordered around the head according to the values $x_\tau$ corresponding to their syntactic relations; those dependents where $x_\tau < 0$ are ordered before the head; the others after the head.
-See SI Section S5.2 for the methodology used to extract the languages' actual grammars from datasets, and for validation against expert judgments.
+See SI Appendix, Section S5.2 for the methodology used to extract the languages' actual grammars from datasets, and for validation against expert judgments.
 
@@
 Very similar formalizations of Zipf's ideas have been proposed across the information-theoretic literature on language~\cite{frank2012predicting,kemp2012kinship,regier2015word, xu2016historical}. 
-See SI Section S2.1 for discussion.
+See SI Appendix, Section S2.1 for discussion.
 
@@
 When optimizing grammars for efficiency, we set $\lambda := 0.9$ in Equation~\ref{eq:efficiency} in order to give approximately equal weight to both components.
-See SI Section S2.2 for mathematical discussion of $\lambda$, and robustness to other choices.
+See SI Appendix, Section S2.2 for mathematical discussion of $\lambda$, and robustness to other choices.
 
@@
 All parseability and predictability values are reported on the held-out (\emph{dev}) partitions from the predefined split for each UD corpus.
-See SI Sections S5--S8 for details and for robustness of our results to modeling choices, including evidence that our results are not specific to any particular language model or parser.
+See SI Appendix, Sections S5--S8 for details and for robustness of our results to modeling choices, including evidence that our results are not specific to any particular language model or parser.
 
@@
 Code and results are available at \url{https://github.com/m-hahn/grammar-optim}.
-The efficiency optimization results from Table~\ref{table:corr-resu} were preregistered: \url{ http://aspredicted.org/blind.php?x=ya4qf8} (see also SI Section S4.6). 
+The efficiency optimization results from Figure~\ref{table:corr-resu} were preregistered: \url{ http://aspredicted.org/blind.php?x=ya4qf8} (see also SI Appendix, Section S4.6). 
 }

### 2019-11-12 writeup/figure6.tex
code

% https://tex.stackexchange.com/questions/171982/how-to-make-table-floats-work-with-preview-standalone-environment/171983

\documentclass[%
    ,float=false % this is the new default and can be left away.
    ,preview=true
    ,class=scrartcl
%    ,fontsize=20pt
    ]{standalone}

\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}

\usepackage{xcite}
\externalcitedocument{pnas-word-order-si}

\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage{siunitx}

\usepackage{longtable}

\usepackage[numbers]{natbib}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{arydshln}

\newcommand{\key}[1]{\textbf{#1}}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}

\usepackage{mathtools}
\DeclarePairedDelimiter{\ceiling}{\lceil}{\rceil}

\begin{document}
\begin{tabular}{|c|ll|c|cc|ccc}
	\hline
	%&
	&	\multicolumn{2}{c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Optimized} & \\ 
	&	verb & object     & & &   \\ 
	&	\emph{wrote} & \emph{letters} & & & \\ \hline \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	adposition    &    noun phrase       
				&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Real_lifted_case.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Efficiency_lifted_case.pdf}     } & \\
	&		\emph{to}            & \emph{a friend} &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&copula    &    noun phrase         
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Real_lifted_cop.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Efficiency_lifted_cop.pdf}     } & \\
	&	\emph{is}        & \emph{a friend}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&auxiliary    &    verb phrase       
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Real_aux.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Efficiency_aux.pdf}     } & \\
	&	\emph{has}          & \emph{written}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	&noun    &    genitive      
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Real_nmod.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Efficiency_nmod.pdf}     } & \\
	&	\emph{friend} &  \emph{of John}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&noun    &    relative clause      
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Real_acl.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Efficiency_acl.pdf}     } & \\
	&	\emph{books} & \emph{that you read}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&complementizer    &    sentence        
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Real_lifted_mark.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Efficiency_lifted_mark.pdf}     } & \\
	&	\emph{that} & \emph{she has arrived}  &&&\\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&verb    &    adp. phrase         
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Real_obl.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Efficiency_obl.pdf}     } & \\
	&	\emph{went} & \emph{to school}  &&&\\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&want    &    verb phrase        
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Real_xcomp.pdf}     } 
		&   \multirow{2}{*}{  \includegraphics[width=0.12\textwidth]{../results/correlations/figures/posteriors//posterior_Efficiency_xcomp.pdf}     } & \\
	& \emph{wants}   &  \emph{to leave}  &&&\\ \hline
\end{tabular}
\end{document}

### 2019-11-12 writeup/prepareRevisedSubmission.sh

@@ -5,7 +5,10 @@ cp paper-revised.bib revised-submission/
 cp pnas-new.cls revised-submission/
 cp pnas-new.bst revised-submission/
 cp pnasresearcharticle.sty revised-submission/
-cp figures/figure1.pdf revised-submission/
-cp figures/figure3.pdf revised-submission/
-cp table2.pdf revised-submission/
+cp figure1.pdf revised-submission/
+cp figure3.pdf revised-submission/
+cp figure6.pdf revised-submission/
+cp figure2.pdf revised-submission/
+cp table1.pdf revised-submission/
+cp table1.tex revised-submission/
 

### 2019-11-12 writeup/table1.tex
code

% https://tex.stackexchange.com/questions/171982/how-to-make-table-floats-work-with-preview-standalone-environment/171983

\documentclass[%
    ,float=false % this is the new default and can be left away.
    ,preview=true
    ,class=scrartcl
%    ,fontsize=20pt
    ]{standalone}

\usepackage{amsmath}
\usepackage{tikz-dependency}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\E}{\mathop{\mathbb{E}}}

\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}

\usepackage{xcite}
\externalcitedocument{pnas-word-order-si}

\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage{siunitx}

\usepackage{longtable}

\usepackage[numbers]{natbib}

\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{arydshln}

\newcommand{\key}[1]{\textbf{#1}}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}

\usepackage{mathtools}
\DeclarePairedDelimiter{\ceiling}{\lceil}{\rceil}

\begin{document}
\begin{tabular}{|c|ll|ll|}
	\hline
	&	\multicolumn{2}{c|}{\textbf{Arabic} (English, ...)}   &        \multicolumn{2}{c|}{\textbf{Japanese} (Turkish, ...)} \\ \hline\hline
	& \multicolumn{2}{c|}{Correlates with...} & \multicolumn{2}{c|}{Correlates with...}  \\
	&	Verb & Object     & Object & Verb    \\ \hline\hline
	& kataba & ris{\= a}la 	&tegami-o & kaita \\
	&	\emph{wrote} & \emph{letter} & 	\emph{letter} & \emph{wrote} \\ \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    {\d s}ad{\= i}q       &	tomodachi & ni \\ % G3, G4
	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline 
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&k{\= a}na    &    {\d s}ad{\= i}q         &	tomodachi & datta \\ % not in Greenberg
	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak- & -udesho \\ % G16, G13 TODO this should only be inflected auxiliaries, replace Arabic
	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& {\d s}ad{\= i}q  &    John    & 	John no & tomodachi \\ % G2, G23
	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uh{\= a}       & 	anata-ga yonda & hon \\ % G24
	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tu{\d s}il        & 	toochaku suru & koto \\ % 
	&	\emph{that} & \emph{she arrives}  &	\emph{arrives} & \emph{that} \\ \hline
	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'il{\= a} lmadrasa     & 	gakkoo ni & itta \\ % G22
	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'ur{\= i}d    &    'an 'ugh{\= a}dir         & 	ik- & -itai \\ % G15
	& \emph{wants}   &  \emph{to leave}  &	\emph{to go} & \emph{want} \\ \hline 
\end{tabular}
\end{document}

### 2019-11-12 writeup/3-paper-names.tex
code

@@
 	\begin{center}
-    \includegraphics[width=0.45\textwidth]{table1.pdf}
+\begin{tabular}{|c|ll|ll|}
+	\hline
+	&	\multicolumn{2}{c|}{\textbf{Arabic} (English, ...)}   &        \multicolumn{2}{c|}{\textbf{Japanese} (Turkish, ...)} \\ \hline\hline
+	& \multicolumn{2}{c|}{Correlates with...} & \multicolumn{2}{c|}{Correlates with...}  \\
+	&	Verb & Object     & Object & Verb    \\ \hline\hline
+	& kataba & ris{\= a}la 	&tegami-o & kaita \\
+	&	\emph{wrote} & \emph{letter} & 	\emph{letter} & \emph{wrote} \\ \hline 
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}}	&	li    &    {\d s}ad{\= i}q       &	tomodachi & ni \\ % G3, G4
+	&		\emph{to}            & \emph{a friend} &		\emph{friend} & \emph{to} \\ \hline 
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}}}	&k{\= a}na    &    {\d s}ad{\= i}q         &	tomodachi & datta \\ % not in Greenberg
+	&	\emph{was}        & \emph{a friend} 	&\emph{friend} & \emph{was} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}}	&sawfa    &    yaktub       & 	kak- & -udesho \\ % G16, G13 TODO this should only be inflected auxiliaries, replace Arabic
+	&	\emph{will}          & \emph{write}  &    	\emph{write} & \emph{will} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}}	& {\d s}ad{\= i}q  &    John    & 	John no & tomodachi \\ % G2, G23
+	&	\emph{friend} &  \emph{of John}  &	\emph{John of} & \emph{friend} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}}	&kutub    &    taqra'uh{\= a}       & 	anata-ga yonda & hon \\ % G24
+	&	\emph{books} & \emph{that you read}  &	\emph{that you read} & \emph{book} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {6}}}}	&'an    &    tu{\d s}il        & 	toochaku suru & koto \\ % 
+	&	\emph{that} & \emph{she arrives}  &	\emph{arrives} & \emph{that} \\ \hline
+	\multirow{2}{*}{	\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}}	&dhahabt	    &    'il{\= a} lmadrasa     & 	gakkoo ni & itta \\ % G22
+	&	\emph{went} & \emph{to school}  &	\emph{school to} & \emph{went} \\ \hline
+	\multirow{2}{*}{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {8}}}}	&'ur{\= i}d    &    'an 'ugh{\= a}dir         & 	ik- & -itai \\ % G15
+	& \emph{wants}   &  \emph{to leave}  &	\emph{to go} & \emph{want} \\ \hline 
+\end{tabular}
 	\end{center}

### 2019-11-12 writeup/pnas-word-order-si.tex
code

@@
 \item Correlation {\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {7}}}} is formalized using \emph{obl}, which covers not only PPs and standards of comparison, but also adjunct NPs.
-In the WALS evaluation (Table~\ref{tab:grammars-wals}), the dominant order of \emph{nmod} agrees with that annotated for obliques in all 18 languages for which WALS has an entry.
+In the WALS evaluation (Table~\ref{tab:grammars-wals}), the dominant order of \emph{obl} agrees with that annotated for obliques in all 18 languages for which WALS has an entry.
 
@@
 
-We obtained the posterior density plots in Table 2 (Main Paper) and in Figure (\ref{tab:all-predictions-1}) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\alpha_j$ (\ref{eq:mixed-effects}).
+We obtained the posterior density plots in Figure 6 (Main Paper) and in Figure (\ref{tab:all-predictions-1}) by applying the logistic transformation ($x \mapsto \frac{1}{1+\exp(-x)}$) to the posterior samples of $\alpha_j$ (\ref{eq:mixed-effects}).
 As the logistic transformation is inverse to the logit transform (\ref{eq:mixed-effects}), this corresponds to the posterior distribution of the prevalence (between 0.0 and 1.0) of each correlation, controlling for languages and language families.
@@
 The `Real' column provides the prevalence among the 51 languages in the Universal Dependencies data.
-We provide posterior prevalences for grammars optimized for Efficiency, and for grammars optimized for Pars(eability) and Pred(ictability) alone, obtained from the Bayesian mixed-effects analysis controlling for languages and language families (as in Table 2 of the main paper).
+We provide posterior prevalences for grammars optimized for Efficiency, and for grammars optimized for Pars(eability) and Pred(ictability) alone, obtained from the Bayesian mixed-effects analysis controlling for languages and language families (as in Figure 6 of the main paper).
 In the last column, we indicate what prevalence is expected according to the typological literature. 

### 2019-11-12 writeup/prepareRevisedSubmission.sh

@@ -9,6 +9,4 @@ cp figure1.pdf revised-submission/
 cp figure3.pdf revised-submission/
 cp figure6.pdf revised-submission/
 cp figure2.pdf revised-submission/
-cp table1.pdf revised-submission/
-cp table1.tex revised-submission/
 
